### 2018-12-18 Reading Notes/reading_notes.tex
Initial commit

\documentclass[12pt]{article}

\usepackage{setspace}

\usepackage{xcolor, amsmath, amsfonts, amssymb, graphicx, color, fancyhdr, lipsum, scalerel, stackengine, mathrsfs, tikz-cd, mdframed, enumitem}
\usepackage[amsthm]{ntheorem}
\usepackage[mathscr]{euscript}
%set margins
\usepackage[
top    = 1.0in,
bottom = 1.0in,
left   = 1.0in,
right  = 1.0in]{geometry}

%header stuff
\setlength{\headsep}{24pt}  % space between header and text
\pagestyle{fancy}     % set pagestyle for document
\lhead{ Problems from our reading course } % put text in header (left side)
\rhead{Nico Courts} % put text in header (right side)
\cfoot{\it p. \thepage}
\setlength{\headheight}{15pt}
\allowdisplaybreaks

%Set of Integers
\newcommand*{\Z}{
\mathbb{Z}
}
%Set of Natural Numbers
\newcommand*{\N}{
\mathbb{N}
}
%Set of Real Numbers
\newcommand*{\R}{
\mathbb{R}
}
%Set of Complex Numbers
\newcommand*{\C}{
\mathbb{C}
}
%Field
\newcommand*{\F}{
\mathbb{F}
}
%Rationals
\newcommand*{\Q}{
\mathbb{Q}
}

%Section break
\newcommand*{\brk}{
\rule{2in}{.1pt}
}

\DeclareMathOperator{\Aut}{Aut}

%raise that Chi!
\DeclareRobustCommand{\Chi}{{\mathpalette\irchi\relax}}
\newcommand{\irchi}[2]{\raisebox{\depth}{$#1\chi$}}

%Image
\DeclareMathOperator{\im}{Im}

%Hom
\DeclareMathOperator{\Hom}{Hom}

%Ext
\DeclareMathOperator{\Ext}{Ext}

%Coker
\DeclareMathOperator{\coker}{coker}

%characteristic
\DeclareMathOperator{\ch}{char}

%restriction
\DeclareMathOperator{\Res}{Res_H}

%socle
%restriction
\DeclareMathOperator{\Soc}{Soc}

%induction
\DeclareMathOperator{\Ind}{Ind_H^G}

%fix tilde
\let\tilde\relax
\newcommand*{\tilde}[1]{\widetilde{#1}}

\newtheorem{lem}{Lemma}[subsection]
\newtheorem{thm}[lem]{Theorem}
\newtheorem{pro}[lem]{Problem}

\setenumerate[0]{label=(\alph*)}

%New environments for problem solving
\newenvironment{prob}{\par\smallskip
	\noindent\begin{mdframed}\small \begin{pro}}{\end{pro}\end{mdframed}\medskip}
\newenvironment{sol}{\noindent \textbf{Solution.} \,}{\\\hspace*{\fill}$\square$\medskip}

%\onehalfspacing
\begin{document}
%make the title page
\title{ Notes and Problems from My Research\vspace{-1ex}}
\author{Nico Courts}
\date{}
\maketitle

%begin problems

\section{Autumn 2018}
\subsection{Problems}
\begin{prob}
	Assume that $k$ is a field and let $K=k(t)$ (notice $K$ is a transcendental extension). Prove that $\Hom_k(K,k)\not\cong K$.
\end{prob}

\begin{sol}
	This is basically just a cardinality argument. I don't think it's particularly worth doing at this juncture.
\end{sol}

\begin{prob}
	Let $G$ be a finite group scheme (actually we need only assume that $G$ is a Frobenius algebra so that a module
	is injective if and only if it is projective). Prove that unless $M$ is projective, its projective dimension is 
	infinite. Conclude that $H^n(G,M)=0$ for $n>N$ implies that $M$ is projective.
\end{prob}
\begin{sol}
	Assume $M$ itself is not projective so that its minimal projective resolution is nontrivial and furthermore
	that it is finite. That is, let $P_i$ be projective modules such that
	\[0\to P_n\xrightarrow{f_n} P_{n-1}\to\cdots \xrightarrow{f_{1}}P_0\xrightarrow{f_0} M\to 0\]
	is a minimal length projective resolution of $M$ (notice here that $n\ge 1$).

	Next consider the short exact sequence
	\[0\to P_n\xrightarrow{f_n}P_{n-1}\to \coker f_n\to 0\]
	since $P_n$ is projective (and thus injective!) this sequence splits and therefore $P_{n-1}\cong P_n\oplus\coker f_n.$
	But then consider the sequence
	\[0\to P_n\xrightarrow{g} P_{n-2}\to\cdots\xrightarrow{f_0}M\to 0\]
	where above we are using $P_{n-1}\supseteq P_n\cong f_n(P_n)$  and that
	$g=f_{n-1}|_{f_n(P_n)}$. This map is injective since $\ker f_{n-1}=\coker f_n$, which is disjoint
	from $f_n(P_n)\cong P_n$. Exactness everywhere else is evident since the maps are not effectively changed.

	But then the existence of this sequence contradicts the minimality of the original sequence, so
	no finite sequence can exist.
\end{sol}

\begin{prob}
	Establish the five-term exact sequence for spectral sequences.
\end{prob}
\begin{sol}
	I plan to return to this problem in the future. I have other priorities at the moment,
	but I will eventually return to cohomology and spectral sequences and this will be a good
	exercise at that point.
\end{sol}

\section{Winter 2019}
\subsection{Preparation: Waterhouse and G\"ortz \& Weddhorn}
Problems I worked from Waterhouse can be found in the appropriate file.

\end{document}

### 2018-12-18 Waterhouse Problems/waterhouse_problems.tex
Initial commit

\documentclass[12pt]{article}

\usepackage{setspace}

\usepackage{xcolor, amsmath, amsfonts, amssymb, graphicx, color, fancyhdr, lipsum, scalerel, stackengine, mathrsfs, tikz-cd, mdframed, enumitem}
\usepackage[amsthm]{ntheorem}
\usepackage[mathscr]{euscript}
%set margins
\usepackage[
top    = 1.0in,
bottom = 1.0in,
left   = 1.0in,
right  = 1.0in]{geometry}

%header stuff
\setlength{\headsep}{24pt}  % space between header and text
\pagestyle{fancy}     % set pagestyle for document
\lhead{ {William Waterhouse's \it Introduction to Affine Group Schemes} } % put text in header (left side)
\rhead{Nico Courts} % put text in header (right side)
\cfoot{\it p. \thepage}
\setlength{\headheight}{15pt}
\allowdisplaybreaks

%Set of Integers
\newcommand*{\Z}{
\mathbb{Z}
}
%Set of Natural Numbers
\newcommand*{\N}{
\mathbb{N}
}
%Set of Real Numbers
\newcommand*{\R}{
\mathbb{R}
}
%Set of Complex Numbers
\newcommand*{\C}{
\mathbb{C}
}
%Field
\newcommand*{\F}{
\mathbb{F}
}
%Rationals
\newcommand*{\Q}{
\mathbb{Q}
}

%Section break
\newcommand*{\brk}{
\rule{2in}{.1pt}
}

\DeclareMathOperator{\Aut}{Aut}

%raise that Chi!
\DeclareRobustCommand{\Chi}{{\mathpalette\irchi\relax}}
\newcommand{\irchi}[2]{\raisebox{\depth}{$#1\chi$}}

%Image
\DeclareMathOperator{\im}{Im}

%Hom
\DeclareMathOperator{\Hom}{Hom}

%Coker
\DeclareMathOperator{\coker}{coker}

%characteristic
\DeclareMathOperator{\ch}{char}

%restriction
\DeclareMathOperator{\Res}{Res_H}

%socle
%restriction
\DeclareMathOperator{\Soc}{Soc}

%induction
\DeclareMathOperator{\Ind}{Ind_H^G}

%fix tilde
\let\tilde\relax
\newcommand*{\tilde}[1]{\widetilde{#1}}

\newtheorem{lem}{Lemma}
\newtheorem{thm}{Theorem}

\setenumerate[0]{label=(\alph*)}

%New environments for problem solving
\newenvironment{prob}[1]{\par\smallskip
	\noindent\begin{mdframed}\small \textbf{Problem~\thesection.#1} \rmfamily\quad}{\end{mdframed}\medskip}
\newenvironment{sol}{\noindent \textbf{Solution.} \,}{\\\hspace*{\fill}$\square$\medskip}

%\onehalfspacing
\begin{document}
%make the title page
\title{ Problems from William Waterhouse's\\ \textit{Introduction to Affine Group Schemes}\vspace{-1ex}}
\author{Nico Courts}
\date{}
\maketitle

%begin problems

\section{Affine Group Schemes}
\begin{prob}{1}
	\begin{enumerate}
		\item If $R$ and $S$ are two $k$ algebras and $F$ is a representable functor, show $F(R\times S)\cong F(R)\times F(S)$.
		\item Show there is no representable functor $R$ such that every $F(R)$ has exactly two elements.
		\item Let $F$ be the functor represented by $k\times k$. Show that $F(R)$ has two elements exactly when $R$ has no idempotents besides 0 and 1.
	\end{enumerate}
\end{prob}

\begin{sol}
	\subsubsection*{(a)}
	Let $A$ be the $k$-algebra representing $F$. Thus $F(R)$ is naturally isomorphic to
	$\Hom_k(A,R)$ and $F(S)\simeq\Hom(A,S)$. Then define the map $\Phi:\Hom(A,R\times S)\to\Hom(A,R)\times\Hom(A,S)$
	via
	\[\Phi(\varphi)=(\pi_R\circ\varphi,\pi_S\circ\varphi)\]
	where $\pi_X$ is the canonical projection onto $X$.

	This is surjective since (by the universal property of products) any pair of maps
	$\varphi_R:A\to R$ and $\varphi_S:A\to S$ factors through the product $R\times S$
	and furthermore it does so \textit{uniquely}, giving us injectivity. Thus this map
	(which is clearly a homomorphism since $\pi_X$ is) is a bijection.
	\subsubsection*{(b)}
	By the last problem this is impossible since there are more than 2 $k$-algebras for any $k$.
	\subsubsection*{(c)}
	Let $F$ be such a functor. Consider any $\varphi\in\Hom(k\times k, R)\simeq F(R)$.
	Assume first that $F(R)\cong \Z/2$ and let $r$ be an idempotent in $R$. 
\end{sol}

\begin{prob}{2}
	Let $E$ be a functor represented by $A$ and let $F$ be any functor. Show that the natural
	maps $\eta:E\to F$ correspond to elements in $F(A)$.
\end{prob}
\begin{sol}
	Consider the map $\Phi$ from natural maps $E\to F$ to elements in $F(A)$ defined by
	(again leveraging the representability of $E$)
	\[\eta\mapsto \eta(\operatorname{id}_A)\in F(A).\]
	Conversely, consider the map $\Psi$ from $F(A)$ to the natural maps $E\to F$ via
	\[x\mapsto \xi_x\]
	where $\xi_x$ where for any $Y$ and $y\in E(Y)\cong \Hom(A,Y)$ we define the $Y^{th}$ component of $\xi_x$
	as 
	\[\xi_x(y)=F(y)(x)\in F(Y)\]
	where (for clarity while I get a grasp here) $F(y):F(A)\to F(Y)$.

	Since we are only looking for a bijection, we only need that these maps are inverses.
	Consider that for all $Y$ and $y\in E(Y)$,
	\begin{align*}
		\Psi\circ\Phi(\eta)(y)&=\Psi\left(\eta(\operatorname{id}_A)\right)(y)\\
		&= \xi_{\eta(\operatorname{id}_A)}(y)\\
		&= F(y)\circ \eta (\operatorname{id}_A)\\
		&= \eta\circ E(y) (\operatorname{id}_A)\\
		&= \eta(y\circ\operatorname{id}_A)=\eta(y)
	\end{align*}
	where above we used the naturality of $\eta$ along with the fact that $E(y)$ is just
	precomposition with $y$. Thus $\Psi\circ\Phi(\eta)=\eta.$

	But then for any $x\in F(A)$,
	\begin{align*}
		\Phi\circ\Psi(x)&=\Phi\circ\xi_x\\
		&=\xi_x(\operatorname{id}_A)\\
		&=F(\operatorname{id}_A)(x)\\
		&=\operatorname{id}_{F(A)}(x)=x
	\end{align*}
	completing the proof.
\end{sol}

\begin{prob}{3}
	Let $E$ be a functor represented by $A$, and let $F$ be any functor. Let $\Psi:F\to E$
	be a natural map with surjective component maps. Show there is a natural map $\Phi:E\to F$
	with $\Psi\circ\Phi=\operatorname{id}_E$.
\end{prob}
\begin{sol}
	Since in particular $\Psi_A$ is surjective, there is an $x\in F(A)$ such that $\Psi(x)=\operatorname{id}_A$.
	Then using the map from the last problem, let $\Phi=\xi_x$. Then we can compute for any $R$ and $g\in E(R)$
	\begin{align*}
		\Psi\circ\Phi (g) &= \Psi\circ F(g) (x)\\
		&=E(g)\circ \Psi(x)\\
		&=E(g)(\operatorname{id}_A)\\
		&=g\circ \operatorname{id}_A=g
	\end{align*}
	since $g:A\to R$, so $E(g):E(A)\to E(R)$, which is just composition with $g$.
\end{sol}

\begin{prob}{5}
	Write out $\Delta,\varepsilon,$ and $S$ for the Hopf algebras representing $SL_2, \mu_n,$ and $\alpha_p$.
\end{prob}
\begin{sol}
	\subsubsection*{$SL_2$:}
	Notice $SL_2$ is represented by $A=k[X_1,X_2,X_3,X_4]/(X_1X_4-X_3X_2-1)$ so take two elements
	$f,g\in\Hom(A,R)$ where $f(X_i)=a_i\in R$ and $g(X_i)=b_i\in R$ and notice that we want
	\[(f,g)\Delta=h\]
	where since
	\[\begin{pmatrix}
		a_1 & a_2\\ a_3 & a_4
	\end{pmatrix}\begin{pmatrix}
		b_1 & b_2\\ b_3 & b_4
	\end{pmatrix}=\begin{pmatrix}
		a_1b_1+a_2b_3 & a_1b_2+a_2b_4\\ a_3b_1+a_4b_3 & a_3b_2+a_4b_4
	\end{pmatrix}=\begin{pmatrix}
		c_1 & c_2\\ c_3 & c_4
	\end{pmatrix}\]
	we want to have that $h(X_i)=c_i.$

	So then if $\Delta:A\to A\otimes A$ is defined as follows:
	\begin{align*}
		X_1&\mapsto X_1\otimes X_1+X_2\otimes X_3
	\end{align*}
\end{sol}

\begin{prob}{6}
	In $A=k[X_{11},\dots,X_{nn},1/\operatorname{det}]$ representing $GL_n$, show that $\Delta(X_{ij})=\sum X_{ik}\otimes X_{kj}$.
	What is $\varepsilon(X_{ij})$?
\end{prob}
\begin{sol}
	Due to the uniqueness of $\Delta,\varepsilon,$ and $S$, we need only find maps satisfying 
	the diagrams. I claim that $\varepsilon(X_{ij})=\delta_{ij}$. In this case, notice
	\[(\varepsilon\otimes\operatorname{id})\circ\Delta(X_{ij})=\varepsilon\otimes\operatorname{id}\left( \sum X_{ik}\otimes X_{kj} \right)=\sum \delta_{ik}\otimes X_{kj}=1\otimes X_{ij}\]
	exactly as we want.

	For associativity, notice
	\[(\Delta\otimes \operatorname{id})\circ \Delta (X_{ij})=\Delta\otimes\operatorname{id}\left( \sum_k X_{ik}\otimes X_{kj} \right)=\sum_k\left(\sum_l X_{il}\otimes X_{lk}\right)\otimes X_{kj}\]
	and then the associativity of $\Delta$ follows simply from the associativity of the tensor product.

	For the last axiom, we compute $S$ such that $(S,\operatorname{id})\circ\Delta=\iota\circ\varepsilon$ where $\iota:K\to A$ is the map sending $k\mapsto k\cdot 1_A$.
	That is, we define $S:A\to A$ so that
	\[\sum_k S(X_{ik})X_{kj}=\delta_{ij}.\]

	We want to leverage the fact that for a fixed $i$ and $j$, the determinant is
	\begin{align*}
		\operatorname{det} &= \sum_{\sigma\in S_n}\operatorname{sgn}(\sigma)\prod_l X_{\sigma(l)l}\\
		&=\sum_\sigma \operatorname{sgn}(\sigma)X_{\sigma{j}j}\prod_{l\ne j} X_{\sigma(l)l}\\
		&=\sum_i X_{ij}\left(\sum_{\sigma(j)=i}\operatorname{sgn}(\sigma)\prod_{l\ne j}X_{\sigma(l)l}\right)
	\end{align*}
	and so we want that 
	\[S(X_{ik})=\frac{1}{\operatorname{det}}\sum_{\sigma(i)=k}\operatorname{sgn}(\sigma)\prod_{l\ne k}X_{\sigma(l)l}\]
	so that when $i=j$,
	\[\sum_k S(X_{ik})X_{kj}=\frac{1}{\operatorname{det}}\sum_k X_{kj}\sum_{\sigma(j)=k}\operatorname{sgn}(\sigma)\prod_{l\ne k}X_{\sigma(l)l}=1=\delta_{ij}\]
	whenever $i\ne j$, however, this equation is the determinant of the matrix where we have replaced
	the $j^{th}$ column with a copy of the $i^{th}$ column. This is linearly dependent, so 
	\[\frac{1}{\operatorname{det}}\sum_k S(X_{ik})S_{kj}=0=\delta_{ij}.\]
	Thus these are precisely the maps we desire.
\end{sol}

\begin{prob}{}

\end{prob}
\begin{sol}
	a
\end{sol}

\begin{prob}{}

\end{prob}
\begin{sol}
	a
\end{sol}

\end{document}

### 2018-12-18 Webb Problems/webb_problems.tex
Initial commit

\documentclass[12pt]{article}

\usepackage{setspace}

\usepackage{xcolor, amsmath, amsfonts, amssymb, graphicx, color, fancyhdr, lipsum, scalerel, stackengine, mathrsfs, tikz-cd, mdframed}
\usepackage[amsthm]{ntheorem}
\usepackage[mathscr]{euscript}
%set margins
\usepackage[
top    = 1.0in,
bottom = 1.0in,
left   = 1.0in,
right  = 1.0in]{geometry}

%header stuff
\setlength{\headsep}{24pt}  % space between header and text
\pagestyle{fancy}     % set pagestyle for document
\lhead{ {Peter Webb's \it A Course in Finite Group Representation Theory} } % put text in header (left side)
\rhead{Nico Courts} % put text in header (right side)
\cfoot{\it p. \thepage}
\setlength{\headheight}{15pt}
\allowdisplaybreaks

%Set of Integers
\newcommand*{\Z}{
\mathbb{Z}
}
%Set of Natural Numbers
\newcommand*{\N}{
\mathbb{N}
}
%Set of Real Numbers
\newcommand*{\R}{
\mathbb{R}
}
%Set of Complex Numbers
\newcommand*{\C}{
\mathbb{C}
}
%Field
\newcommand*{\F}{
\mathbb{F}
}
%Rationals
\newcommand*{\Q}{
\mathbb{Q}
}

%Section break
\newcommand*{\brk}{
\rule{2in}{.1pt}
}

\DeclareMathOperator{\Aut}{Aut}

%raise that Chi!
\DeclareRobustCommand{\Chi}{{\mathpalette\irchi\relax}}
\newcommand{\irchi}[2]{\raisebox{\depth}{$#1\chi$}}

%Image
\DeclareMathOperator{\im}{Im}

%Hom
\DeclareMathOperator{\Hom}{Hom}

%Coker
\DeclareMathOperator{\coker}{coker}

%characteristic
\DeclareMathOperator{\ch}{char}

%restriction
\DeclareMathOperator{\Res}{Res_H}

%socle
%restriction
\DeclareMathOperator{\Soc}{Soc}

%induction
\DeclareMathOperator{\Ind}{Ind_H^G}

%fix tilde
\let\tilde\relax
\newcommand*{\tilde}[1]{\widetilde{#1}}

\newtheorem{lem}{Lemma}
\newtheorem{thm}{Theorem}

%New environments for problem solving
\newenvironment{prob}[1]{\par\smallskip
	\noindent\begin{mdframed}\small \textbf{Problem~\thesection.#1} \rmfamily\quad}{\end{mdframed}\medskip}
\newenvironment{sol}{\noindent \textbf{Solution.} \,}{\\\hspace*{\fill}$\square$\medskip}

%\onehalfspacing
\begin{document}
%make the title page
\title{ Problems from Peter Webb's\\ \textit{A Course in Finite Group Representation Theory}\vspace{-1ex}}
\author{Nico Courts}
\date{}
\maketitle

%begin problems

\section{Representations, Maschke's Theorem, and Semisimplicity}
\begin{prob}{1}
	In Example 1.1.6, prove that there are no invariant subspaces other than the ones listed.
\end{prob}

\begin{sol}
	Recall that in this example we are considering the representation of $C_2=\Z_2$ given by $\rho:C_2\to GL(\R^2)$ via
	\[\rho(1)=\begin{pmatrix}
	1 & 0\\ 0 & 1
	\end{pmatrix},\qquad \rho(-1)=\begin{pmatrix}
	1 & 0\\ 0 & -1
	\end{pmatrix}\tag{1}\label{eq:1.1.1}.\]
	In the example we noted that $\{0\}$, $\langle\binom{1}{0}\rangle$, $\langle\binom{0}{1}\rangle$ and $\R^2$ are all invariant subspaces.

	Assume that $V$ is any $\R$-subspace of $\R^2$ not listed above. Then necessarily it has $\R$-dimension one, so is spanned by an element $\binom{a}{b}\in\R^2.$ If $V$ is stable under the $C_2$ action, it must be that (in particular) the image of $\binom{a}{b}$ under nontrivial matrix in (\ref{eq:1.1.1}) above lands back in $V$. But then for some $\alpha\in\R$,
	\[\alpha\binom{a}{b}=\binom{a}{-b}\]
	whence either $a\ne 0$ and so $\alpha=1$ and $b=0$ or else $a=0$ and (since $V$ is nontrivial) $b\ne 0$ whence $\alpha=-1$ and $a=0$. But these two case describe precisely the two one dimensional subspaces given above so this must in fact be a complete list.
\end{sol}

\begin{prob}{2}
	(\textit{The Modular Law}) -- Let $A$ be a ring and $U=V\oplus W$ an $A$-module that is a direct sum of $A$-modules $V$ and $W$. Show by example that if $X$ is any submodule of $U$ then it need not be the case that $X=(V\cap X)\oplus(W\cap X)$. Show that if we make the assumption that $V\subseteq X$ then it is true that $X=(V\cap X)\oplus (W\cap X).$
\end{prob}
\begin{sol}
	Consider $A=\R$ and let $U=\langle e_1,e_2\rangle$ be $\R^2$ spanned by the standard basis vectors. Now define $V=\langle e_1+e_2\rangle$ and $W=\langle e_1\rangle$. $V\cap W=\{0\}$ and $V+W=U$, so the hypotheses are satisfied. Now let $X=\{e_2\}$ and notice
	\[(X\cap A)\oplus (X\cap B)=\{0\}\oplus\{0\}=\{0\}\ne X.\]

	Now assume that $V\subseteq X$, so that $V\cap X=V.$ Now we want to prove that $X=V\oplus(W\cap X).$ Clearly since $V$ and $W\cap X$ are in $X$, their sum is as well. Therefore $V\oplus (W\cap X)\subseteq X.$

	Assume now that $x\in X$. But then $x\in U$ so $x=v+w$ for $v\in V$ and $w\in W$. Since $V\subseteq X$, $v\in X$ and the same goes for $x-v=w$. But then $x=(v)+(x-v)$ where the former part is in $V$ and the latter is in $X\cap W$. This gives us the reverse inclusion and thus equality.
\end{sol}

\begin{prob}{3}
	Suppose that $\rho$ is a finite-dimensional representation of a finite group $G$ over $\C$. Show that for each $g\in G$, the matrix $\rho(g)$ is diagonalizable.
\end{prob}
\begin{sol}
	Let $|G|=n$ and $\rho:G\to GL(\C)$ be a degree $k<\infty$ representation. Since $G$ has finite order, $\rho(g)^n=I_k$ so $\rho(g)$ satisfies the polynomial $x^n-1$. Since $\C$ is algebraically closed, $\rho(g)$ has $k$ eigenvalues, each of which must be an $n^{th}$ root of unity. Furthermore $x^n-1$ has no repeated roots (separable) so the minimal polynomial of $\rho(g)$ has distinct roots.

	Therefore $\rho(g)$ has $k$ distinct eigenvalues in $\C$ which is sufficient to show that it is diagonalizable.
\end{sol}

\begin{prob}{4}
	Let $\phi:U\to V$ be an $A$-module homomorphism. Show that $\phi(\Soc U)\subseteq \Soc V$ and that if $\phi$ is an isomorphism then it restricts to an isomorphism $\Soc U$ to $\Soc V$.
\end{prob}
\begin{sol}
	Recall that the image of a simple module is itself simple. To see this, assume that $S$ is simple and $0\ne X\subseteq \phi(S)$ is a submodule. Then $\phi^{-1}(X)$ is a submodule of $U$ intersecting $S$ nontrivially and by simplicity $\phi^{-1}(X)\cap S=S$. But then $\phi(S)\subseteq X$, so they are equal. Thus the only nonzero submodule of $\phi(S)$ is itself.

	Thus since $\Soc U$ is a sum of simple modules of $U$, $\varphi(\Soc U)$ is a sum of simple modules of $V$. So $\phi(\Soc U)\subseteq \Soc V$, the sum of \textit{all} simple modules of $V$.

	Assume now that $\phi$ is an isomorphism. Then by the same argument as above $\phi^{-1}(\Soc V)\subseteq \Soc U$ and so
	\[\phi\circ\phi^{-1}(\Soc V)\subseteq \phi(\Soc U)\quad\Rightarrow\quad \Soc V\subseteq \phi(\Soc U)\]
	so they are equal.

	Finally that $\Soc V=\phi(\Soc U)$ and $\phi^{-1}(\Soc V)=\Soc U$ and the injectivity of $\phi$ and its inverse gives us that $\phi$ restricts to an isomorphism $\Soc U\to\Soc V$.
\end{sol}

\begin{prob}{5}
	Let $U=S_1\oplus\cdots \oplus S_r$ be an $A$-module that is the direct sum of finitely many simple $S_i$. Show that if $T$ is any simple submodule of $U$ then $T\cong S_i$ for some $i$.
\end{prob}
\begin{sol}
	Since $T\subseteq U$, $T\cap S_i$ is nontrivial for some $i$. But then $T\cap S_i=S_i$ by the simplicity of $S_i$ and so $S_i\subseteq T$. By problem 1.2, since we have
	\[T\subset S_i\oplus\left(S_1\oplus\cdots\oplus\hat S_i\oplus\cdots\oplus S_r\right)=S_i+S\]
	we have that
	\[T=(T\cap S_i)\oplus (T\cap S)=S_i\oplus (T\cap S)\]
	and by the simplicity of $T$, $T\cap S=0$ and $T\cong S_i$.
\end{sol}

\begin{prob}{6}
	Let $V$ be an $A$-module for some ring $A$ and suppose that $V$ is a sum $V=V_1+\cdots+V_n$ of simple submodules. Assume further that the $V_i$ are pairwise nonisomorphic. Show that the $V_i$ are the only simple submodules of $V$ and that $V=V_1\oplus\cdots\oplus V_n$ is their direct sum.
\end{prob}
\begin{sol}
	Let $W$ be any simple submodule of $V$. By Lemma 1.2.3 in the book $V=V_{i_1}\oplus\cdots\oplus V_{i_k}$ for some subset of the $V_i$. By problem 1.5 above, this gives us $W\cong V_j$ for some $j$. But the $V_i$ are pairwise nonisomorphic so $W=V_j$ so the $V_i$ are the only simple submodules of $V$.

	That all $V_i$ occur in the direct sum holds by again applying the last problem and noticing that each $V_i$ must be isomorphic to some $V_j$ in the direct sum, whence $V_i$ itself must appear.
\end{sol}

\begin{prob}{7}
	Let $G=\langle x,y| x^2=y^2=1=[x,y]\rangle$ be the Klein 4-group, $R=\F_2$, and consider the two representations $\rho_1$ and $\rho_2$ given by
	\[\rho_1(x)=\begin{pmatrix}
		1&1&0\\0&1&0\\0&0&1
	\end{pmatrix},\qquad \rho_1(y)=\begin{pmatrix}
	1&0&1\\0&1&1\\0&0&1
	\end{pmatrix}\]
	and
	\[\rho_2(x)=\begin{pmatrix}
	1&0&0\\0&1&1\\0&0&1
	\end{pmatrix},\qquad \rho_2(y)=\begin{pmatrix}
	1&0&1\\0&1&0\\0&0&1
	\end{pmatrix}.\]
	Compute the socles of these representations. Show that neither representation is semisimple.
\end{prob}
\begin{sol}
	Notice that every subrepresentation is automatically a vector subspace of $\F_2^3$, so begin by considering the one dimensional subspaces. If $\langle (a,b,c)^T\rangle$ is invariant under $\rho_1$, consider the images:
	\[\rho_1(x)(a,b,c)^T=(a+b,b,c)^T,\quad\rho_1(y)(a,b,c)^T=(a+c,b,c)^T\]
	each of which are either $(a,b,c)^T$ or $\mathbf{0}$. In either case we are forced to have $b=c=0$, so the only invariant one dimensional subspace is the one spanned by $(1,0,0)^T$.

	In the two dimensional case, consider any subspace spanned by $\mathbf{v}=(a,b,c)^T$ and $\mathbf{w}=(d,e,f)^T$. If $b\ne e$ or $c\ne f$, we must have that $\rho_1$ fixes each of these vectors, so the resulting module must not be simple. Therefore we can assume that $b=e$ and $c=f.$ But then since it must be two dimensional, $a=d+1\pmod{2}$. But then
	\[\mathbf{v}+\mathbf{w}=(1,0,0)^T\]
	which spans an invariant subspace under $\rho_1$ so this is also not simple.

	So there are no simple degree two subrepresentations of $\rho_1$ so the socle of $\rho_1$ is the one dimensional space spanned by $(1,0,0)^T$. Thus $\rho_1$ is clearly not semisimple.

	\brk

	Let $\mathbf{v}=(a,b,c)^T$ be any vector spanning a degree one subrepresentation of $\rho_2$. Then 
	\[\rho_2(x)\mathbf{v}=(a,b+c,c)^T,\qquad \rho_2(y)\mathbf{v}=(a+c,b,c)^T.\]
	Neither can be $\mathbf{0}$ since otherwise $\mathbf{v}=\mathbf{0}$. But this implies $c=0$. Therefore there are three one-dimensional invariant subspaces spanned by $(1,0,0)^T$, $(0,1,0)^T$ and $(1,1,0)^T$.

	Now given vectors $\mathbf{v}=(a,b,c)^T$ and $\mathbf{w}=(d,e,f)^T$ spanning a degree 2 subrepresentation, notice that if $c=f$ then $\mathbf{v}+\mathbf{w}$ is either $\mathbf{0}$ (whence $\mathbf{v}=\mathbf{w}$) or else this sum is one of the vectors from the previous paragraph. In either case $\mathbf{v}$ and $\mathbf{w}$ don't span a simple degree two representation space, so we have $c\ne f.$

	But then without loss of generality if $c=0$, $\mathbf{v}$ spans an invariant subspace so we can conclude that there are \textit{no} degree two simple subrepresentations of $\rho_2$. Thus the socle is computed as the sum of the spaces spanned by the three vectors from the first paragraph:
	\[\langle(1,0,0)^T,(0,1,0)^T,(1,1,0)^T\rangle=\{(a,b,0)^T\in\F_2^3\}\subsetneq\F_2^3\]
	so this representation is not semisimple either.
\end{sol}

\begin{prob}{8}
	Let $G=C_p=\langle x\rangle$ and $R=\F_p$ for some prime $p\ge 3.$ Consider the two representations $\rho_1$ and $\rho_2$ specified by
	\[\rho_1(x)=\begin{pmatrix}
		1&1&0\\0&1&1\\0&0&1
	\end{pmatrix}\quad\text{and}\quad \rho_1(x)=\begin{pmatrix}
	1&1&1\\0&1&0\\0&0&1
	\end{pmatrix}.\]
	Calculate the socles of these two representations and show that neither representation is semisimple. Show that the second representation is nevertheless the direct sum of two nonzero subrepresentations.
\end{prob}
\begin{sol}
	Consider first $\rho_1$. If $\mathbf{v}=(a,b,c)$ (lazily eschewing the transpose in this problem for notational simplicity) spans a one dimensional invariant subspace, then for some $\alpha\in\Z_p$ we have
	\[\rho_1(x)\mathbf{v}=\alpha\begin{pmatrix}
		a+b\\b+c\\c
	\end{pmatrix}\]
	Assume by contradiction that $c\ne 0$. Then $\alpha=1$ and $b=b+c\Rightarrow c=0$, a contradiction. Thus $c=0$. Similarly if $b\ne 0$, $a+b=a$, and we get another contradiction. Thus $b=c=0$. Any vector $(a,0,0)\in\F_p^3$ spans an invariant subspace with the trivial action by $\rho_1$ so these are all the degree one subrepresentations.

	Notice that if $V$ is the representation space of a degree two subrepresentation of $\rho_1$ and if $\mathbf{v}\in V$ where the third component of $\mathbf{v}$ is zero, we have that $\mathbf{v}=0$. This can be seen by considering 
	\[\left(\rho_1(x^2)-\rho_1(x)\right)\begin{pmatrix}v_1\\v_2\\0\end{pmatrix}=\begin{pmatrix}v_1+2v_2\\v_2\\0\end{pmatrix}-\begin{pmatrix}v_1+v_2\\v_2\\0\end{pmatrix}=\begin{pmatrix}v_2\\0\\0\end{pmatrix}\]
	and if $v_2\ne 0$, $V$ contains one of the one-dimensional subspaces from above and is therefore not simple. Otherwise $v_2=0$ and $\mathbf{v}$ spans a degree 1 invariant subspace, so again $V$ is not simple.

	\textbf{Not finished.}
\end{sol}

\begin{prob}{9}
	Let $k$ be an infinite field of characteristic 2, and $G=\langle x,y\rangle\cong C_2\times C_2$ be the noncyclic group of order 4. For each $\lambda\in k$, let $\rho_\lambda(x),\rho_\lambda(y)$ be
	\[\rho_\lambda(x)=\begin{pmatrix}1&0\\1&1\end{pmatrix},\quad\rho_\lambda(y)=\begin{pmatrix}1&0\\\lambda & 1\end{pmatrix},\]
	regarded as linear maps $U_\lambda\to U_\lambda$ where $U_\lambda$ is a $k$-vector space of dimension 2 with basis $\{e_1,e_2\}.$
	\begin{itemize}
		\item[(a)] Show that $\rho_\lambda$ defines a representation of $G$ with representation space $U_\lambda$.
		\item[(b)] Find a basis for $\Soc U_\lambda$.
		\item[(c)] By considering the effect on $\Soc U_\lambda$, show that any $kG$-module homomorphism $\alpha:U_\lambda\to U_\mu$ has a triangular matrix $\alpha=(\begin{smallmatrix}a&0\\b&c\end{smallmatrix})$ with respect to the given bases.
		\item[(d)] Show that if $U_\lambda\cong U_\mu$ as $kG$-modules then $\lambda=\mu$. Deduce that $kG$ has infinitely many nonisomorphic 2-dimensional representations.
	\end{itemize}
\end{prob}
\begin{sol}
	Let $k,G,$ and $\rho_\lambda$ be defined as above.
	\subsubsection*{(a)}
	Some routine computations gives us
	\[\rho_\lambda(x)\rho_\lambda(y)=\begin{pmatrix}1&0\\\lambda+1& 1\end{pmatrix}=\rho_\lambda(y)\rho_\lambda(x)\]
	and
	\[(\rho_\lambda(x))^2=(\rho_\lambda(y))^2=I_2\]
	proving $\rho_\lambda$ is a group homomorphism into $GL_2(k)$, whence is a representation (which acts on $U_\lambda$ by definition).
	\subsubsection*{(b)}
	Any one dimensional subspace spanned by $(a,b)$ must have $a=0$ in order to be fixed by $x$ (referring for simplicity to the element and its action on $U_\lambda$ rather than its image under the map). But then the only proper subspace is the one spanned by $e_2.$
	\subsubsection*{(c)}
	Leveraging the result from problem 1.4, if $\alpha:U_\lambda\to U_\mu$ is a $kG$-module homomorphism, then 
	\[\alpha(\Soc U_\lambda)=\alpha(\langle e_2\rangle)\subseteq \langle e_2\rangle=\Soc U_\mu.\]
	That is, $\alpha(e_1)=ce_1$ for some $c\in k$, giving us that $\alpha$ has the form
	\[\begin{pmatrix}
		a & 0\\
		b & c
	\end{pmatrix}\]
	for some $a,b\in k$ in the standard basis.
	\subsubsection*{(d)}
	Assume that $U_\lambda\cong U_\mu$ via $\alpha$. By the fact that the $G$ action pulls through $\alpha$, we get that
	\[\begin{pmatrix}1&0\\1&1\end{pmatrix}
	\begin{pmatrix}a&0\\b&c\end{pmatrix}
	=\begin{pmatrix}a&0\\a+b&c\end{pmatrix}
	=\begin{pmatrix}a&0\\b+c&c\end{pmatrix}
	=\begin{pmatrix}a&0\\b&c\end{pmatrix}
	\begin{pmatrix}1&0\\1&1\end{pmatrix}\]
	and so $a+b=b+c$ whence $a=c=k$. That this matrix is invertible means $k\ne 0.$

	Using a similar computation, we know
	\[\begin{pmatrix}1&0\\\lambda&1\end{pmatrix}
	\begin{pmatrix}a&0\\b&c\end{pmatrix}
	=\begin{pmatrix}a&0\\\lambda a+b&c\end{pmatrix}
	=\begin{pmatrix}a&0\\b+\mu c&c\end{pmatrix}
	=\begin{pmatrix}a&0\\b&c\end{pmatrix}
	\begin{pmatrix}1&0\\\mu&1\end{pmatrix}\]
	so $\lambda a=\mu c$ or $\lambda k=\mu k$ and since $k\ne 0$, $\lambda=\mu.$

	Thus since $\rho_\lambda$ can be defined for any $\lambda\in k$, this gives us the existence of infinitely many non-isomorphic two dimensional representations, as desired.
\end{sol}

\begin{prob}{10}
	Let
	\[\rho_1:G\to GL(V),\qquad\rho_2:G\to GL(V)\]
	be two representations of $G$ on the same $R$-module $V$ that are injective as homomorphisms. (We say that such a representation is \textit{faithful}.) Consider the three properties that
	\begin{itemize}
		\item[(1)] the $RG$-modules given by $\rho_1$ and $\rho_2$ are isomorphic,
		\item[(2)] the subgroups $\rho_1(G)$ and $\rho_2(G)$ are conjugate in $GL(V)$,
		\item[(3)] for some automorphism $\alpha\in\Aut(G)$, the  representations $\rho_1$ and $\rho_2\alpha$ are isomorphic.
	\end{itemize}
	Show that (1)$\Rightarrow$(2) and that (2)$\Rightarrow$(3). Show also that if $\alpha\in\Aut(G)$ is an inner automorphism then $\rho_1$ and $\rho_1\alpha$ are isomorphic.
\end{prob}
\begin{sol}
	Let $\rho_i:G\to GL(V)$ be faithful representations acting on $_RV$.
	\subsubsection*{(1)$\Rightarrow$(2)}
	Let $V_1$ and $V_2$ be $V$ with the $RG$-module structure defined by $\rho_1$ and $\rho_2$, respectively and assume that $V_1\cong V_2$, say via a map $\alpha$. Let $v\in V$ be arbitrary and notice that by virtue of being a $RG$-module homomorphism,
	\[\alpha(\rho_1(g)(v))=\rho_2(g)\alpha(v),\quad\forall g\in G.\]
	That is, $\alpha\circ \rho_1(g)=\rho_2(g)\circ\alpha$ as maps in $GL(V)$ (since $\alpha$ is invertible and $R$-linear) and furthermore
	\[\rho_1(g)=\alpha^{-1}\circ\rho_2(g)\circ\alpha\]
	for each $g\in G$ whence $\rho_1(G)$ is conjugate to $\rho_2(G)$ in $GL(V)$.
	\subsubsection*{(2)$\Rightarrow$(3)}
	Now assume that $T^{-1}\rho_1(G)T=\rho_2(G)$ for some $T\in GL(V)$. Now since $\rho_1$ and $\rho_2$ are faithful, $\exists! h_g\in G$ for each $g\in G$ such that
	\[T^{-1}\rho_1(g)T=\rho_2(h).\]
	Define $\alpha\in\Aut(G)$ by $\alpha(g)=h_g$. This is a homomorphism since $\rho_2$ is injective and
	\[\rho_2(\alpha(gg'))=T^{-1}\rho_1(gg')T=T^{-1}\rho_1(g)TT^{-1}\rho_1(g')T=\rho_2(\alpha(g))\rho_2(\alpha(g'))=\rho_2(\alpha(g)\alpha(g'))\]
	and an automorphism since it is both surjective and injective.

	Now define the map $\varphi=T^{-1}\in GL(V)$. This map is bijective and $R$ linear, so we need only show it is $G$-linear to prove it is a module isomorphism. But consider
	\[\rho_2(\alpha^{-1}g)(\varphi(v))=T^{-1}\rho_1(g)T(T^{-1}v)=T^{-1}(\rho_1(g)(v))=\varphi(\rho_1(g)(v)).\]
	which proves $G$-linearity and we can conclude that $\rho_1\cong \rho_2\alpha^{-1}$.
	\subsubsection*{Finally,}
	Let $\alpha\in \Aut(G)$ be an inner automorphism corresponding to conjugation by $h\in G$. But then 
	\[\rho_1(\alpha(g))=\rho_1(h^{-1}gh)=\rho_1(h^{-1})\rho_1(g)\rho_1(h)=H^{-1}\rho_1(g)H\]
	for each $g\in G$, so $\rho_1(\alpha(G))=H^{-1}\rho_1(G)H$, so (2) applies and the result follows from the argument above.
\end{sol}

\begin{prob}{11}
	One version of the Jordan-Zassenhaus Theorem asserts that for each $n$, $GL(n,\Z)=\Aut(\Z^n)$ has only finitely many conjugacy classes of subgroups of finite order. Assuming this, show that for each finite group $G$ and each integer $n$, there are only finitely many isomorphism classes of representations of $G$ on $\Z^n$.
\end{prob}
\begin{sol}
	Fix a finite $G$ and $n\in\N$. Assume by contradiction that there were infinitely many pairwise nonisomophic representations $\rho_i:G\to\Z^n$. Now consider the subgroups $H_i=\rho_i(G)\le Z^n$.

	Now if $H_i=k^{-1}H_jk$ then $H_i=H_j$ since $\Z^n$ is abelian. But since $G$ is finite, so are the $H_i$, so there are only finitely many maps $\rho_i$ that can map onto each finite $H_i\le \Z^n$. Furthermore by Jordan-Zassenhaus, there exist only finitely many conjugacy classes among the $H_i$, giving that there must be, in fact, only finitely many $H_i$. This contradicts that there be infinitely many $\rho_i$ and proves the statement. 
\end{sol}

\begin{prob}{12}
	Write out a proof of Maschke's theorem in the case of representations over $\C$ along the following lines.
	\begin{itemize}
		\item[(a)] Given a representation $\rho:G\to GL(V)$ where $V$ is a vector space over $\C$, let $(-,-)$ be a positive definite Hermitian form on $V$. Define a new form $(-,-)_1$ on $V$ by
		\[(v,w)_1=\frac{1}{|G|}\sum_{g\in G}(gv,gw).\]
		Show that $(-,-)_1$ is a positive definite Hermitian form, preserved by the action by $G$; that is $(v,w)_1=(gv,gw)_1$ always. If $W$ is a subrepresentation of $V$ then show that $V=W\oplus W^\perp$, as representations.
		\item[(b)] Show that any finite subgroup of $GL(n,\C)$ is conjugate to a subgroup of $U(n,\C)$ (the unitary group with matrices satisfying $A\bar A^T=I$). Show that any finite subgroup of $GL(n,\R)$ is conjugate to a subgroup of $O(n,\R)$, the orthogonal group of matrices satisfying $AA^T=I$.
	\end{itemize}

\end{prob}
\begin{sol}
	\textbf{(a)}\hspace{1ex} Define $(-,-)_1$ as above and we can see that this is a bilinear form since
	\[(a\alpha+b\beta,\gamma)_1=\frac{1}{|G|}\sum_g (ag\alpha+bg\beta,g\gamma)=\frac{1}{|G|}\sum_g\left(a(g\alpha,g\gamma)+b(g\beta,g\gamma)\right)
	= a(\alpha,\gamma)_1+b(\beta,\gamma)_1\]
	and linearity in the second coordinate is proved via an identical argument. It is positive definite since $(-,-)$ is positive definite and the image of $(-,-)_1$ is simply the average of images of $(-,-).$

	Finally this form is Hermitian since 
	\[(v,w)_1=\frac{1}{|G|}\sum (gv,gw)=\frac{1}{|G|}\sum\overline{(gw,gv)}=\overline{\frac{1}{|G|}\sum(gw,gv)}=\overline{(w,v)_1}.\]
	That this form is preserved by action by $G$ is evident since multiplying $u$ and $v$ by $g$ is the same as simply permuting the summands in the definition.

	Now let $_{\C G}W\le V$. $W^\perp$ is a vector space in $V$, but we must show it is stable under the $G$-action. Assume that for $g\in G$ and $0\ne w\in W^\perp$, $gw\in W.$ But then since $W$ is a subrepresentation, $g^{-1}(gw)=w\in W$, a contradiction. Thus $W$ and $W^\perp$ are disjoint (from the definition/linear algebra) subrepresentations of $V$ and it suffices to show $V=W+W^\perp$.

	To see this,
\end{sol}

\begin{prob}{13}
	\begin{itemize}
		\item[(a)] Using proposition 1.2.4, show that if $A$ is a ring such that the regular representation $_AA$ is semisimple, then every finitely generated $A$-module is semisimple.
		\item[(b)] Extend the result of part (a) using Zorn's lemma, to show that if $A$ is a ring for which $_AA$ is semisimple then every $A$-module is semisimple.
	\end{itemize}
\end{prob}
\begin{sol}
	\textbf{(a)}\hspace{1ex} Let $A$ be a ring with semisimple regular representation.
\end{sol}

\end{document}

### 2019-01-04 Reading Notes/reading_notes.tex
testing
Tweak styles and add some content

@@
 
-\usepackage{xcolor, amsmath, amsfonts, amssymb, graphicx, color, fancyhdr, lipsum, scalerel, stackengine, mathrsfs, tikz-cd, mdframed, enumitem}
-\usepackage[amsthm]{ntheorem}
+\usepackage{xcolor, amsmath, amsfonts, amssymb, graphicx, color, fancyhdr, lipsum, scalerel, stackengine, mathrsfs, tikz-cd, mdframed, enumitem, framed, adjustbox}
+\usepackage[framed,thmmarks]{ntheorem}
 \usepackage[mathscr]{euscript}
+
+%set up theorem/definition/etc envs
+%Problems will be created using their own counter and style
+\theoreminframepreskip{0pt}
+\theoreminframepostskip{0pt}
+\newframedtheorem{prob}{Problem}[section]
+
+%solution template
+\theoremstyle{nonumberbreak}
+\theoremindent0.5cm
+\theorembodyfont{\upshape}
+\theoremseparator{:}
+\theoremsymbol{\ensuremath\spadesuit}
+\newtheorem{sol}{Solution}
+
+%Theorems, Lemmas, and Corollaries
+\theoremstyle{changebreak}
+\theoremseparator{}
+\theoremsymbol{}
+\theoremindent0.5cm
+\theoremheaderfont{\bfseries} 
+\newtheorem{thm}{Theorem}[subsection]
+\newtheorem{lem}[thm]{Lemma}
+\newtheorem{cor}[thm]{Corollary}
+
+%Create a new env that references a theorem and creates a 'primed' version
+%Note this can be used recursively to get double, triple, etc primes
+\newenvironment{thm-prime}[1]
+  {\renewcommand{\thethm}{\ref{#1}$'$}%
+   \addtocounter{thm}{-1}%
+   \begin{thm}}
+  {\end{thm}}
+
+\setlength\fboxsep{15pt}
+
+%Shade definitions
+\theoremindent0cm
+\theoremheaderfont{\normalfont\bfseries} 
+\def\theoremframecommand{\colorbox[rgb]{.9,.9,1}}
+\newshadedtheorem{defn}[thm]{Definition}
+
+%Man, that's really good! Let's use the same thing for definitons.
+\newenvironment{def-prime}[1]
+  {\renewcommand{\thethm}{\ref{#1}$'$}%
+   \addtocounter{thm}{-1}%
+   \begin{def}}
+  {\end{def}}
+
+%proofs
+\theoremstyle{nonumberbreak}
+\theoremindent1cm
+\theoremheaderfont{\sc}
+\theoremseparator{}
+\theoremsymbol{\ensuremath\spadesuit}
+\newtheorem{prf}{Proof}
+
+\usepackage{fullpage}
 %set margins
-\usepackage[
-top    = 1.0in,
-bottom = 1.0in,
-left   = 1.0in,
-right  = 1.0in]{geometry}
+%\usepackage[
+%top    = 1.0in,
+%bottom = 1.0in,
+%left   = 1.0in,
+%right  = 1.0in]{geometry}
 
@@
 \pagestyle{fancy}     % set pagestyle for document
-\lhead{ Problems from our reading course } % put text in header (left side)
+\lhead{Notes on Affine Group Schemes and Their Representations} % put text in header (left side)
 \rhead{Nico Courts} % put text in header (right side)
-\cfoot{\it p. \thepage}
+\cfoot{\itshape p. \thepage}
 \setlength{\headheight}{15pt}
@@
 %socle
-%restriction
 \DeclareMathOperator{\Soc}{Soc}
+
+%rank
+\DeclareMathOperator{\rank}{rank}
 
@@
 
-\newtheorem{lem}{Lemma}[subsection]
-\newtheorem{thm}[lem]{Theorem}
-\newtheorem{pro}[lem]{Problem}
-
+% Enumerate will automatically use letters (e.g. part a,b,c,...)
 \setenumerate[0]{label=(\alph*)}
 
-%New environments for problem solving
-\newenvironment{prob}{\par\smallskip
-	\noindent\begin{mdframed}\small \begin{pro}}{\end{pro}\end{mdframed}\medskip}
-\newenvironment{sol}{\noindent \textbf{Solution.} \,}{\\\hspace*{\fill}$\square$\medskip}
-
-%\onehalfspacing
 \begin{document}
 %make the title page
-\title{ Notes and Problems from My Research\vspace{-1ex}}
+\title{Notes and Problems from My Research\vspace{-1ex}}
 \author{Nico Courts}
@@
 \section{Winter 2019}
-\subsection{Preparation: Waterhouse and G\"ortz \& Weddhorn}
-Problems I worked from Waterhouse can be found in the appropriate file.
+This is my first official quarter as Julia's student!! 
+\subsection{Preparation, Waterhouse, and G\"ortz \& Weddhorn}
+To give a sense of direction, Julia recommended that I take a look at the following regularity theorem:
+\begin{thm}[Smoothness Theorem]\label{smooth}
+	Let $G$ be an algebraic affine group scheme over a field $k$. Then $k[G]\otimes \bar k$ is reduced if and
+	only if $\dim G = \rank \Omega_{k[G]}$.
+\end{thm}
+I have quite a bit of information to process before this, so I will get started!
+\begin{defn}[Closed Embedding]
+	If $G$ and $H$ are affine group schemes represented by $A$ and $B$, respectively and
+	if $\psi:H\to G$ is a homomorphism of affine group schemes (locally a group homomorphism)
+	then if the corresponding algebra map $A\to B$ is surjective, then $\psi$ is called
+	a \textbf{closed embedding.}
+\end{defn}
+
+As its name suggests, this means that $\psi$ is an isomorphism onto a \textbf{closed subgroup}
+$H'$ of $G$. This is, in fact, a definition of this property. One can also think about 
+it in the following ways: a group scheme $H$ is closed in $G$ if
+\begin{itemize}
+	\item $H$ is defined by the relations imposed by $G$ plus some additional ones.
+	\item $H=V(I)$ for some ideal $I\subset k[G]$. 
+\end{itemize}
+Thinking back to our algebraic geometry, these are not too hard to see as equivalent. 
+For instance, there is a closed embedding of $\mu_n$ in $G_m$ (simply adding in $x^n-1$)
+and of $SL_n$ in $GL_n$ (adding $\det = 1$).
+
+\subsubsection{Hopf Ideals}
+One problem with the above characterization that one cannot choose $I$ arbitrarily and end
+up with a group scheme. This is equivalent to arbitrarily adding relations to a group, which is
+not always guaranteed to work out well (think adding $\det = 2$ to $GL_n$).
+
+Actually, we can exactly categorize the closed embeddings of subgroups in $G$ by considering certain
+ideals of the algebra $A$ which represents it. 
+\begin{defn}
+	Let $A$ be an algebra and $I\lhd A$. Then if 
+	\begin{itemize}
+		\item $\Delta(I)$ goes to zero under the map $A\otimes A\to A/I\otimes A/I$,
+		\item $S(I)\subseteq I$
+		\item $\epsilon(I)=0$
+	\end{itemize}
+	then $I$ is called a \textbf{Hopf Ideal of $A$.}
+\end{defn}
 

### 2019-01-04 Waterhouse Problems/waterhouse_problems.tex
Tweak styles and add some content

@@
 
-\newtheorem{lem}{Lemma}
-\newtheorem{thm}{Theorem}
+\newtheorem{lem}{Lemma}[section]
+\newtheorem{thm}[lem]{Theorem}
+\newtheorem{defn}[lem]{Definition}
 
@@
 
-\begin{prob}{}
-
-\end{prob}
-\begin{sol}
-	a
-\end{sol}
-
-\begin{prob}{}
-
-\end{prob}
-\begin{sol}
-	a
-\end{sol}
+\section{Examples}
+\begin{defn}[Closed Embedding]
+	If $G$ and $H$ are affine group schemes represented by $A$ and $B$, respectively and
+	if $\psi:H\to G$ is a homomorphism of affine group schemes (locally a group homomorphism)
+	then if the corresponding algebra map $A\to B$ is surjective, then $\psi$ is called
+	a \textbf{closed embedding.}
+\end{defn}
 

### 2019-01-05 Reading Notes/reading_notes.tex
Continue working

@@
 \theoreminframepostskip{0pt}
-\newframedtheorem{prob}{Problem}[section]
+\newframedtheorem{prob}{Problem}[part]
+\renewcommand\theprob{\arabic{part}.\arabic{prob}}
 
@@
 \theoremstyle{nonumberbreak}
-\theoremindent1cm
+\theoremindent0.5cm
 \theoremheaderfont{\sc}
@@
 
+%remarks
+\theoremstyle{change}
+\theoremindent0.5cm
+\theoremheaderfont{\sc}
+\theoremseparator{:}
+\theoremsymbol{}
+\newtheorem{rmk}[thm]{Remark}
+
+%waterhouse problem
+\newenvironment{wprob}[1]{\begin{prob}{\normalfont\bfseries (Waterhouse #1) }\itshape}{\end{prob}}
+
+%Replacement for the old geometry package
 \usepackage{fullpage}
-%set margins
-%\usepackage[
-%top    = 1.0in,
-%bottom = 1.0in,
-%left   = 1.0in,
-%right  = 1.0in]{geometry}
+
+%Put page breaks before each part
+\let\oldpart\part%
+\renewcommand{\part}{\clearpage\oldpart}%
+
+%Center each figure by default
+\makeatletter
+\g@addto@macro\@floatboxreset{\centering}
+\makeatother
 
@@
 
+%identity map
+\DeclareMathOperator{\id}{id}
+
+%schemes
+\DeclareMathOperator{\Ga}{\ensuremath\mathbf{G_a}}
+\DeclareMathOperator{\Gm}{\ensuremath\mathbf{G_m}}
+\newcommand*{\GL}[1]{\ensuremath\mathbf{GL_#1}}
+\newcommand*{\SL}[1]{\ensuremath\mathbf{SL_#1}}
+\newcommand*{\mun}[1]{\ensuremath\mathbf{\mu_#1}}
+\newcommand*{\alp}[1]{\ensuremath\mathbf{\alpha_#1}}
+
 %fix tilde
@@
 
-\section{Autumn 2018}
-\subsection{Problems}
+\part{Autumn 2018}
+\section{Problems}
 \begin{prob}
@@
 	no finite sequence can exist.
+
+	The last statement (as discussed with Julia) is actually false. 
 \end{sol}
@@
 
-\section{Winter 2019}
-This is my first official quarter as Julia's student!! 
-\subsection{Preparation, Waterhouse, and G\"ortz \& Weddhorn}
+\begin{wprob}{1.1}
+	\begin{enumerate}
+		\item If $R$ and $S$ are two $k$ algebras and $F$ is a representable functor, show $F(R\times S)\cong F(R)\times F(S)$.
+		\item Show there is no representable functor $R$ such that every $F(R)$ has exactly two elements.
+		\item Let $F$ be the functor represented by $k\times k$. Show that $F(R)$ has two elements exactly when $R$ has no idempotents besides 0 and 1.
+	\end{enumerate}
+\end{wprob}
+
+\begin{sol}
+	\subsubsection*{(a)}
+	Let $A$ be the $k$-algebra representing $F$. Thus $F(R)$ is naturally isomorphic to
+	$\Hom_k(A,R)$ and $F(S)\simeq\Hom(A,S)$. Then define the map $\Phi:\Hom(A,R\times S)\to\Hom(A,R)\times\Hom(A,S)$
+	via
+	\[\Phi(\varphi)=(\pi_R\circ\varphi,\pi_S\circ\varphi)\]
+	where $\pi_X$ is the canonical projection onto $X$.
+
+	This is surjective since (by the universal property of products) any pair of maps
+	$\varphi_R:A\to R$ and $\varphi_S:A\to S$ factors through the product $R\times S$
+	and furthermore it does so \textit{uniquely}, giving us injectivity. Thus this map
+	(which is clearly a homomorphism since $\pi_X$ is) is a bijection.
+	\subsubsection*{(b)}
+	By the last problem this is impossible since if $|F(k)|=2$ then 
+	\[|F(k\times k)|=|F(k)\times F(k)|=4.\]
+	\subsubsection*{(c)}
+	Let $F$ be such a functor. Consider any $\varphi\in\Hom(k\times k, R)\simeq F(R)$.
+	Assume first that $F(R)\cong \Z/2$ and let $r$ be an idempotent in $R$. 
+\end{sol}
+
+\begin{wprob}{1.2}
+	Let $E$ be a functor represented by $A$ and let $F$ be any functor. Show that the natural
+	maps $\eta:E\to F$ correspond to elements in $F(A)$.
+\end{wprob}
+\begin{sol}
+	Consider the map $\Phi$ from natural maps $E\to F$ to elements in $F(A)$ defined by
+	(again leveraging the representability of $E$)
+	\[\eta\mapsto \eta(\operatorname{id}_A)\in F(A).\]
+	Conversely, consider the map $\Psi$ from $F(A)$ to the natural maps $E\to F$ via
+	\[x\mapsto \xi_x\]
+	where $\xi_x$ where for any $Y$ and $y\in E(Y)\cong \Hom(A,Y)$ we define the $Y^{th}$ component of $\xi_x$
+	as 
+	\[\xi_x(y)=F(y)(x)\in F(Y)\]
+	where (for clarity while I get a grasp here) $F(y):F(A)\to F(Y)$.
+
+	Since we are only looking for a bijection, we only need that these maps are inverses.
+	Consider that for all $Y$ and $y\in E(Y)$,
+	\begin{align*}
+		\Psi\circ\Phi(\eta)(y)&=\Psi\left(\eta(\operatorname{id}_A)\right)(y)\\
+		&= \xi_{\eta(\operatorname{id}_A)}(y)\\
+		&= F(y)\circ \eta (\operatorname{id}_A)\\
+		&= \eta\circ E(y) (\operatorname{id}_A)\\
+		&= \eta(y\circ\operatorname{id}_A)=\eta(y)
+	\end{align*}
+	where above we used the naturality of $\eta$ along with the fact that $E(y)$ is just
+	precomposition with $y$. Thus $\Psi\circ\Phi(\eta)=\eta.$
+
+	But then for any $x\in F(A)$,
+	\begin{align*}
+		\Phi\circ\Psi(x)&=\Phi\circ\xi_x\\
+		&=\xi_x(\operatorname{id}_A)\\
+		&=F(\operatorname{id}_A)(x)\\
+		&=\operatorname{id}_{F(A)}(x)=x
+	\end{align*}
+	completing the proof.
+\end{sol}
+
+\begin{wprob}{1.3}
+	Let $E$ be a functor represented by $A$, and let $F$ be any functor. Let $\Psi:F\to E$
+	be a natural map with surjective component maps. Show there is a natural map $\Phi:E\to F$
+	with $\Psi\circ\Phi=\operatorname{id}_E$.
+\end{wprob}
+\begin{sol}
+	Since in particular $\Psi_A$ is surjective, there is an $x\in F(A)$ such that $\Psi(x)=\operatorname{id}_A$.
+	Then using the map from the last problem, let $\Phi=\xi_x$. Then we can compute for any $R$ and $g\in E(R)$
+	\begin{align*}
+		\Psi\circ\Phi (g) &= \Psi\circ F(g) (x)\\
+		&=E(g)\circ \Psi(x)\\
+		&=E(g)(\operatorname{id}_A)\\
+		&=g\circ \operatorname{id}_A=g
+	\end{align*}
+	since $g:A\to R$, so $E(g):E(A)\to E(R)$, which is just composition with $g$.
+\end{sol}
+
+\begin{wprob}{1.5}
+	Write out $\Delta,\varepsilon,$ and $S$ for the Hopf algebras representing $\SL{2}, \mun{n},$ and $\alp{p}$.
+\end{wprob}
+\begin{sol}
+	\subsubsection*{$\SL{2}$:}
+	Notice $SL_2$ is represented by $A=k[X_1,X_2,X_3,X_4]/(X_1X_4-X_3X_2-1)$ so take two elements
+	$f,g\in\Hom(A,R)$ where $f(X_i)=a_i\in R$ and $g(X_i)=b_i\in R$ and notice that we want
+	\[(f,g)\Delta=h\]
+	where since
+	\[\begin{pmatrix}
+		a_1 & a_2\\ a_3 & a_4
+	\end{pmatrix}\begin{pmatrix}
+		b_1 & b_2\\ b_3 & b_4
+	\end{pmatrix}=\begin{pmatrix}
+		a_1b_1+a_2b_3 & a_1b_2+a_2b_4\\ a_3b_1+a_4b_3 & a_3b_2+a_4b_4
+	\end{pmatrix}=\begin{pmatrix}
+		c_1 & c_2\\ c_3 & c_4
+	\end{pmatrix}\]
+	we want to have that $h(X_i)=c_i.$
+
+	So then if $\Delta:A\to A\otimes A$ is defined as follows:
+	\begin{align*}
+		X_1&\mapsto X_1\otimes X_1+X_2\otimes X_3\\
+		X_2&\mapsto X_1\otimes X_2 +X_2\otimes X_4\\
+		X_3&\mapsto X_3\otimes X_1+X_4\otimes X_3\\
+		X_4&\mapsto X_3\otimes X_2+X_4\otimes X_4
+	\end{align*}
+	Where one can compute
+	\begin{align*}
+		\id\otimes\Delta\circ \Delta (X_1)&= (\id\otimes\Delta)(X_1\otimes X_1+X_2\otimes X_3)\\
+		&= X_1\otimes X_1\otimes X_1 + X_1\otimes X_2\otimes X_3 + X_2\otimes X_3\otimes X_1 + X_2\otimes X_4\otimes X_3\\
+		&= (\Delta\otimes\id)(X_1\otimes X_1+X_2\otimes X_3)=\Delta\otimes\id\circ \Delta (X_1)
+	\end{align*}
+	and similar equality holds for the other $X_i$, so this is $\Delta$.
+
+	Using that we want $\varepsilon\otimes\id\circ \Delta(X_i)=1\otimes X_i$, we see that the map
+	$\varepsilon$ sending $X_1$ and $X_4$ to 1 and $X_2$ and $X_3$ to zero is the map we want.
+
+	Notice that as a sanity check we get that 
+	\[\begin{pmatrix}
+		\varepsilon(X_1) & \varepsilon(X_2)\\
+		\varepsilon(X_3) & \varepsilon(X_4)
+	\end{pmatrix}=\begin{pmatrix}
+		1 & 0\\
+		0 & 1
+	\end{pmatrix}=I_2\]
+
+	Finally using that $(S,\id)\circ\Delta(X_i)=\varepsilon(X_i)$ and the fact that in $A$,
+	$\det=X_1X_4-X_3X_2=1$, we can define $S$ such that
+	\[\begin{pmatrix}
+		S(X_1) & S(X_2)\\
+		S(X_3) & S(X_4)
+	\end{pmatrix}=\begin{pmatrix}
+		X_4 & -X_2\\
+		-X_3 & X_1
+	\end{pmatrix}=\begin{pmatrix}
+		X_1 & X_2\\
+		X_3 & X_4
+	\end{pmatrix}^{-1}\]
+	and one can verify that this satisfies the relation above.
+	\subsubsection*{$\mun{n}:$}
+	For this scheme, $A=k[X]/(X^n-1)$ is the representing algebra. If $f,g\in\Hom(A,k)$ with
+	$f(X)=r$ and $g(X)=s$, then we want $(f,g)\Delta(X)=\sum f(X_{(1)})g(X_{(2)})=m(r,s)=rs$ where 
+	$\Delta(X)=\sum X_{(1)}\otimes X_{(2)}$. An obvious choice is the diagonal map.
+
+	Then choosing $\varepsilon(X)=1$ satisfies the diagrams (as before in $G_m$) and using our
+	intuition, $S(X)=X^5$ which also works.
+	\subsubsection*{$\alp{p}$:}
+	This time we are working with $A=\Z/p[X]/(X^p)$. This time (since the group is additive) we want
+	$\Delta(X)=X\otimes 1+1\otimes X$, which we can see works with associativity immediately.
+
+	Following suit with the other additive group scheme $G$, setting $\varepsilon(X)=0$ and $S(X)=-X$
+	we can quickly check these still satisfy the given axioms.
+\end{sol}
+
+\begin{wprob}{1.6}
+	In $A=k[X_{11},\dots,X_{nn},1/\operatorname{det}]$ representing $GL_n$, show that $\Delta(X_{ij})=\sum X_{ik}\otimes X_{kj}$.
+	What is $\varepsilon(X_{ij})$?
+\end{wprob}
+\begin{sol}
+	Due to the uniqueness of $\Delta,\varepsilon,$ and $S$, we need only find maps satisfying 
+	the diagrams. I claim that $\varepsilon(X_{ij})=\delta_{ij}$. In this case, notice
+	\[(\varepsilon\otimes\operatorname{id})\circ\Delta(X_{ij})=\varepsilon\otimes\operatorname{id}\left( \sum X_{ik}\otimes X_{kj} \right)=\sum \delta_{ik}\otimes X_{kj}=1\otimes X_{ij}\]
+	exactly as we want.
+
+	For associativity, notice
+	\[(\Delta\otimes \operatorname{id})\circ \Delta (X_{ij})=\Delta\otimes\operatorname{id}\left( \sum_k X_{ik}\otimes X_{kj} \right)=\sum_k\left(\sum_l X_{il}\otimes X_{lk}\right)\otimes X_{kj}\]
+	and then the associativity of $\Delta$ follows simply from the associativity of the tensor product.
+
+	For the last axiom, we compute $S$ such that $(S,\operatorname{id})\circ\Delta=\iota\circ\varepsilon$ where $\iota:K\to A$ is the map sending $k\mapsto k\cdot 1_A$.
+	That is, we define $S:A\to A$ so that
+	\[\sum_k S(X_{ik})X_{kj}=\delta_{ij}.\]
+
+	We want to leverage the fact that for a fixed $i$ and $j$, the determinant is
+	\begin{align*}
+		\operatorname{det} &= \sum_{\sigma\in S_n}\operatorname{sgn}(\sigma)\prod_l X_{\sigma(l)l}\\
+		&=\sum_\sigma \operatorname{sgn}(\sigma)X_{\sigma{j}j}\prod_{l\ne j} X_{\sigma(l)l}\\
+		&=\sum_i X_{ij}\left(\sum_{\sigma(j)=i}\operatorname{sgn}(\sigma)\prod_{l\ne j}X_{\sigma(l)l}\right)
+	\end{align*}
+	and so we want that 
+	\[S(X_{ik})=\frac{1}{\operatorname{det}}\sum_{\sigma(i)=k}\operatorname{sgn}(\sigma)\prod_{l\ne k}X_{\sigma(l)l}\]
+	so that when $i=j$,
+	\[\sum_k S(X_{ik})X_{kj}=\frac{1}{\operatorname{det}}\sum_k X_{kj}\sum_{\sigma(j)=k}\operatorname{sgn}(\sigma)\prod_{l\ne k}X_{\sigma(l)l}=1=\delta_{ij}\]
+	whenever $i\ne j$, however, this equation is the determinant of the matrix where we have replaced
+	the $j^{th}$ column with a copy of the $i^{th}$ column. This is linearly dependent, so 
+	\[\frac{1}{\operatorname{det}}\sum_k S(X_{ik})S_{kj}=0=\delta_{ij}.\]
+	Thus these are precisely the maps we desire.
+\end{sol}
+
+\begin{wprob}{1.10}
+	Prove the following Hopf algebra facts by interpreting them as statements about group functors:
+	\begin{enumerate}
+		\item $S\circ S=\id$
+		\item $\Delta\circ S=(\operatorname{twist})\circ(S\otimes S)\Delta$
+		\item $\varepsilon\circ S=\varepsilon$
+		\item The map $A\otimes A\to A\otimes A$ sending $a\otimes b$ to $(a\otimes 1)\Delta(b)$
+		is an algebra isomorphism.
+	\end{enumerate}
+\end{wprob}
+\begin{sol}
+	\subsubsection*{(a)}
+	Dualizing, we get
+	\begin{figure}[h]
+		\begin{tikzcd}[column sep=small]
+			A\ar[rr,equal]\ar[rd,"S",swap] & & A\\
+			& A\ar[ru,"S",swap] &
+		\end{tikzcd}
+		\begin{tikzcd}[column sep=small]
+			A\ar[rr,equal] & & A\ar[ld,"inv"]\\
+			& A\ar[lu,"inv"] &
+		\end{tikzcd}
+	\end{figure}
+
+	\noindent so this statement is equivalent to the group (scheme) fact that $(g^{-1})^{-1}=g$.
+	\subsubsection*{(b)}
+	Using a similar duality argument, this is equivalent to saying 
+	\[\operatorname{inv}\circ m = m\circ(\operatorname{inv}\times\operatorname{inv})\circ(\operatorname{twist})\]
+	but if we consider arbitrary elements $g,h\in G(R)$, this means
+	\[(gh)^{-1}=m\circ(\operatorname{inv}\times\operatorname{inv})(h,g)=m(h^{-1},g^{-1})=h^{-1}g^{-1}\]
+	which is clearly true.
+	\subsubsection*{(c)}
+	This one is equivalent to $(\operatorname{inv})\circ i=i$ where if $g\in G(R)$, 
+	\[(\operatorname{inv})\circ i(g)=(\operatorname{inv})(e)=e=i(g)\]
+	or in other words $e^{-1}=e.$
+	\subsubsection*{(d)}
+\end{sol}
+
+\part{Winter 2019}
+This is my first official quarter as Julia's student! My plan for now is to continue working on Waterhouse as well
+as learn about algebraic geometry (alongside my usual classes, of course).
+\section{Preparation, Waterhouse, and G\"ortz \& Weddhorn}
 To give a sense of direction, Julia recommended that I take a look at the following regularity theorem:
@@
 Thinking back to our algebraic geometry, these are not too hard to see as equivalent. 
-For instance, there is a closed embedding of $\mu_n$ in $G_m$ (simply adding in $x^n-1$)
-and of $SL_n$ in $GL_n$ (adding $\det = 1$).
+For instance, there is a closed embedding of $\mun n$ in $\Gm$ (simply adding in $x^n-1$)
+and of $\SL{n}$ in $\GL{n}$ (adding $\det = 1$).
 
-\subsubsection{Hopf Ideals}
+\subsection{Hopf Ideals}
 One problem with the above characterization that one cannot choose $I$ arbitrarily and end
 up with a group scheme. This is equivalent to arbitrarily adding relations to a group, which is
-not always guaranteed to work out well (think adding $\det = 2$ to $GL_n$).
+not always guaranteed to work out well (think adding $\det = 2$ to $\GL n$).
 
@@
 		\item $S(I)\subseteq I$
-		\item $\epsilon(I)=0$
+		\item $\varepsilon(I)=0$
 	\end{itemize}
@@
 
+\begin{rmk}
+	That these ideals exactly characterize closed subgroups follows since any group (scheme) represented
+	by $A'$ has the property that $\Delta'(A')\subseteq A'\otimes A'$. Since the new comultiplication map
+	$\Delta'$ is derived from the old one $\Delta$, we have the diagram
+	\begin{figure}[h]
+		\begin{tikzcd}
+			A \ar[r,"\Delta"]\ar[d] & A\otimes A\ar[d]\\
+			A/I\ar[r,"\Delta'"] & A/I\otimes A/I
+		\end{tikzcd}
+	\end{figure}
+
+	\noindent from which we see that $\Delta(I)$ must go to zero in the map on the right.
+	Similarly, we want that $S$ and $\varepsilon$ satisfy similar diagrams.
+
+	Note that an example of such an ideal is $I=\ker(\varepsilon)$.
+\end{rmk}
+
+\begin{defn}
+	A \textbf{character} is a scheme morphism $G\to \Gm$.
+\end{defn}
+
+Let $\Phi$ be a character of $G$ and notice that the corresponding hopf algebra map 
+$\varphi:k[X,1/X]\to A$ is defined by $\varphi(X)=b$, which
+is automatically invertible in $A$. Furthermore since for any $a,b\in G(R)$ we have
+\[\Phi\circ m(a,b)=\Phi(ab)=\Phi(a)\Phi(b)=m\circ(\Phi,\Phi)(a,b)\]
+we get the diagram
+\begin{figure}[h]
+	\begin{tikzcd}
+		G\times G\ar[r,"\Phi\times\Phi"]\ar[d,"m"] & \Gm\times \Gm\ar[d,"m"]\\
+		G\ar[r,"\Phi"] & \Gm
+	\end{tikzcd}
+\end{figure}
+
+\noindent and then by dualizing we get that
+\newpage
+\begin{figure}[h]
+	\begin{tikzcd}
+		A\otimes A & k[X,X^{-1}]\otimes k[X,X^{-1}]\ar[l,"\varphi\otimes\varphi",swap]\\
+		A\ar[u,"\Delta"] & k[X,X^{-1}]\ar[u,swap,"\Delta"]\ar[l,"\varphi"]
+	\end{tikzcd}
+\end{figure}
+
+\noindent where in both diagrams we are abusing notation by using the same symbols for 
+(co)multiplication in the different schemes. So we conclude that 
+\[\Delta(b)=\Delta\circ\varphi(X)=(\varphi\otimes\varphi)\circ \Delta(X)=(\varphi\otimes\varphi)(X\otimes X)=b\otimes b.\]
+
+Then we can easily compute that $\varepsilon(b)=1$ and $S(b)=b^{-1}$. This is precisely:
+\begin{defn}
+	Let $A$ be a Hopf algebra and $a\in A$ such that
+	\begin{itemize}
+		\item $a$ is invertible
+		\item $\Delta(a)=a\otimes a$
+		\item $\varepsilon(a)=1$
+		\item $S(a)=a^{-1}$
+	\end{itemize}
+	then $a$ is called a \textbf{group-like} element of $A$.
+\end{defn}
+\begin{rmk}
+	As we saw above, every character of an affine group scheme $G$ corresponds to a 
+	group-like element of its representing algebra.
+\end{rmk}
+\begin{rmk}
+	Using a parallel construction with any morphism $\Phi:G\to\Ga$, we get an element $b\in A$
+	such that $\Delta(b)=b\otimes 1+1\otimes b$, $\varepsilon(b)=0$ and $S(b)=-b$. These elements
+	are called \textbf{primitive}
+\end{rmk}
+
+\section{Problems}
+\subsection{Week 1}
+
+\subsection{Waterhouse}
+
 \end{document}

### 2019-01-08 Hopf Algebra Notes/Hopf Algebras.tex
Add Course Notes and make updates

\documentclass[12pt]{article}

\usepackage{setspace}

\usepackage{amsmath, amsfonts, amssymb, graphicx, color, fancyhdr, lipsum, scalerel, stackengine, mathrsfs, tikz-cd, mdframed, enumitem, framed, adjustbox, bm, upgreek, x	color}
\usepackage[framed,thmmarks]{ntheorem}
\usepackage[mathscr]{euscript}

%set up theorem/definition/etc envs
%Problems will be created using their own counter and style
\theoreminframepreskip{0pt}
\theoreminframepostskip{0pt}
\newframedtheorem{prob}{Problem}[section]

%solution template
\theoremstyle{nonumberbreak}
\theoremindent0.5cm
\theorembodyfont{\upshape}
\theoremseparator{:}
\theoremsymbol{\ensuremath\spadesuit}
\newtheorem{sol}{Solution}

%Theorems
\definecolor{thmcol}{RGB}{0,102,34}
\theoremstyle{changebreak}
\theoremseparator{}
\theoremsymbol{}
\theoremindent0.5cm
\theoremheaderfont{\color{thmcol}\bfseries} 
\newtheorem{thm}{Theorem}[subsection]

%Lemmas and Corollaries
\theoremheaderfont{\bfseries}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{cor}[thm]{Corollary}

%Create a new env that references a theorem and creates a 'primed' version
%Note this can be used recursively to get double, triple, etc primes
\newenvironment{thm-prime}[1]
  {\renewcommand{\thethm}{\ref{#1}$'$}%
   \addtocounter{thm}{-1}%
   \begin{thm}}
  {\end{thm}}

\setlength\fboxsep{15pt}

%Shade definitions
\theoremindent0cm
\theoremheaderfont{\normalfont\bfseries} 
\def\theoremframecommand{\colorbox[rgb]{.85,1,.85}}
\newshadedtheorem{defn}[thm]{Definition}

%Example
\theoremstyle{break}
\def\theoremframecommand{\colorbox[rgb]{0.9,0.9,0.9}}
\newshadedtheorem{ex}{Example}[section]

%Man, that's really good! Let's use the same thing for definitons.
\newenvironment{def-prime}[1]
  {\renewcommand{\thethm}{\ref{#1}$'$}%
   \addtocounter{thm}{-1}%
   \begin{def}}
  {\end{def}}

%proofs
\theoremstyle{nonumberbreak}
\theoremindent0.5cm
\theoremheaderfont{\sc}
\theoremseparator{}
\theoremsymbol{\ensuremath\spadesuit}
\newtheorem{prf}{Proof}

\theoremstyle{nonumberplain}
\theoremseparator{:}
\theoremsymbol{}
\newtheorem{conj}{Conjecture}

%remarks
\theoremstyle{change}
\theoremindent0.5cm
\theoremheaderfont{\sc}
\theoremseparator{:}
\theoremsymbol{}
\newtheorem{rmk}[thm]{Remark}

%Replacement for the old geometry package
\usepackage{fullpage}

%Put page breaks before each part
\let\oldpart\part%
\renewcommand{\part}{\clearpage\oldpart}%

%Center each figure by default
\makeatletter
\g@addto@macro\@floatboxreset{\centering}
\makeatother

%header stuff
\setlength{\headsep}{24pt}  % space between header and text
\pagestyle{fancy}     % set pagestyle for document
\lhead{Notes on Hopf Algebras} % put text in header (left side)
\rhead{Nico Courts} % put text in header (right side)
\cfoot{\itshape p. \thepage}
\setlength{\headheight}{15pt}
\allowdisplaybreaks

%Set of Integers
\newcommand*{\Z}{
\mathbb{Z}
}
%Set of Natural Numbers
\newcommand*{\N}{
\mathbb{N}
}
%Set of Real Numbers
\newcommand*{\R}{
\mathbb{R}
}
%Set of Complex Numbers
\newcommand*{\C}{
\mathbb{C}
}
%Rationals
\newcommand*{\Q}{
\mathbb{Q}
}

%Section break
\newcommand*{\brk}{
\rule{2in}{.1pt}
}

\DeclareMathOperator{\Aut}{Aut}

%raise that Chi!
\DeclareRobustCommand{\Chi}{{\mathpalette\irchi\relax}}
\newcommand{\irchi}[2]{\raisebox{\depth}{$#1\chi$}}

%Image
\DeclareMathOperator{\im}{Im}

\DeclareMathOperator{\Ext}{Ext}

%Coker
\DeclareMathOperator{\coker}{coker}

%characteristic
\DeclareMathOperator{\ch}{char}

%rank
\DeclareMathOperator{\rank}{rank}

%identity map
\DeclareMathOperator{\id}{id}

%Hopf algebra stuff
\newcommand*{\Vectk}{\operatorname{Vect}_k}
\newcommand*{\Algk}{\operatorname{Alg}_k}
\newcommand*{\Coalgk}{\operatorname{Coalg}_k}

%fix tilde
\let\tilde\relax
\newcommand*{\tilde}[1]{\widetilde{#1}}

% Enumerate will automatically use letters (e.g. part a,b,c,...)
\setenumerate[0]{label=(\alph*)}

\begin{document}
%make the title page
\title{Hopf Algebras\vspace{-1ex}}
\author{A course by: Prof. James Zhang\\
Notes by: Nico Courts}
\date{Winter 2019}
\maketitle

\renewcommand{\abstractname}{Introduction}
\begin{abstract}
	These are the notes I took in class during the Winter 2019 topics course
	\textit{Math 582H - Hopf Algebras} at University of Washington, Seattle.

	The course description follows:

	\brk

	This course is an introduction to Hopf algebras. In addition to basic material in 
	Hopf algebra, we will present some latest developments in quantum groups and tensor 
	and fusion categories. One of the newer topics is homological properties of 
	Noetherian Hopf algebras of low Gelfand-Kirillov dimension. A good reference for 
	the first two topics in the book \textit{Hopf Algebras and Their Action Rings} by 
	Susan Montgomery. Here is a list of possible topics:

\begin{itemize}
	\item Classical theorems concerning finite dimensional Hopf algebras.
	\item Infinite dimensional Hopf algebras and quantum groups.
	\item Duality and Calabi-Yau property.
	\item Actions of Hopf algebras and invariant theory.
	\item Representations of Hopf algebras, tensor and fusion categories.
\end{itemize}
\end{abstract}

\section{January 7, 2019}
If you don't know what a symmetric tensor category is, today is going to be a three 
star day. Max is 5.

\subsection{Overview}
We are shooting to understand two conjectures:

\begin{conj}[Etingof-Ostrik `04]
	If $A$ is a finite dimentional Hopf algebra, then 
	\[\bigoplus_{i\ge 0}\Ext_A^i(k,k)\]
	is Noetherian.
\end{conj}

\begin{conj}[Brown-Goodearl `98]
	If $A$ is a Noetherian Hopf algebra, then the injective dimension of $A_A$ is finite.
\end{conj}

These are both still open! In fact there is a meeting at Oberwolfach this March concerning
exactly these conjectures.

\subsection{Symmetric Tensor Categories}
We are going to be using the following notation throughout:
\begin{itemize}
	\item $k$ is a field
	\item $\Vectk$ is the category of $k$-vector spaces
	\begin{itemize}
		\item $\Vectk$ is closed under tensor products
		\item There is an element $k\in\Vectk$ such that
		\[k\otimes_k V\cong V\cong V\otimes_k k\]
		where the above isomorphisms are natural.
		\item $V\otimes_k W\cong W\otimes_k V$
	\end{itemize}
	\item An algebra is an object in $\Vectk$.
\end{itemize}

\begin{defn}\label{def-alg}
	$V\in\Vectk$ is called an \textbf{algebra object} if there are two morphisms
	\begin{enumerate}
		\item $m:V\otimes V\to V$
		\item $u:k\to V$
	\end{enumerate}
	such that the diagrams in figure \ref{fig-alg} commute.

\end{defn}
\begin{figure}\label{fig-alg}
	\begin{tikzcd}
		V\otimes V\otimes V\ar[r,"\id_V\otimes m"]\ar[d,"m\otimes \id_V"] & V\otimes V\ar[d,"m"]\\
		V\otimes V \ar[r,"m"] & V
	\end{tikzcd}
	\begin{tikzcd}
		k\otimes V\ar[r,"\sim"]\ar[rd,swap,"u\otimes\id_V"] & V & V\otimes k\ar[l,"\sim",swap]\ar[dl, "\id_V\otimes u"]\\
		& V\otimes V \ar[u,"m"] &
	\end{tikzcd}
	\caption{Diagrams for definition \ref{def-alg}.}
\end{figure}

\begin{lem}
	$V\in\Vectk$ is an algebra object iff $V$ is an algebra over $k$.
\end{lem}

\begin{lem}
	If $C$ is a symmetric tensor category, so is $C^{op}$.
\end{lem}

Then the natural thing to ask is: what is an algebra object in this opposite category?
\begin{defn}
	A \textbf{coalgebra object} in $C$ is an algebra object in $C^{op}$. Here we have 
	comultiplication $\Delta$ and counit $\varepsilon$.
\end{defn}
\begin{rmk}
	Naturally you could go about defining this from first principles and drawing the diagrams
	in figure \ref{fig-alg} with the arrows reversed, but we are probably mature enough
	to do without that (saving my fingers from repetitive strain injury in the process.)
\end{rmk}

\begin{lem}
	$\Algk$, defined as the category of algebra objects in $\Vectk$, is a symmetric tensor category. Furthermore
	$\Coalgk$, the category of coalgebra objects in $\Vectk$, is a symmetric tensor category.
\end{lem}

\begin{lem}
	The following are equivalent:
	\begin{enumerate}
		\item $V$ is an algebra object in $\Coalgk$
		\item $V$ is a coalgebra object in $\Algk$
		\item There are morphisms $m,u,\Delta,\varepsilon$ such that
		\begin{itemize}
			\item $(V,m,u)$ is an algebra
			\item $(V,\Delta,\varepsilon)$ is a coalgebra
			\item Equivalently:
			\begin{itemize}
				\item $m$ and $u$ are coalgebra morphisms
				\item $\Delta$ and $\varepsilon$ are algebra morphisms.
			\end{itemize}
		\end{itemize}
	\end{enumerate}
\end{lem}
\begin{prf}
	The nice thing here is that the $(a)\Leftrightarrow (c)$ without the last condition. A similar fact holds
	for $(b)$ except the second-to-last. The last thing to do is to prove the last two conditions are equivalent.
\end{prf}
\begin{prob}
	Fill in the details for the proof above.
\end{prob}

\begin{defn}
	$V$ is called a \textbf{bialgebra object} if $V$ is an algebra object in $\Coalgk$.
\end{defn}

\begin{prob}
	\begin{enumerate}
		\item Suppose that $\ch k\ne 2$. Classify all bialgebras of $\dim 2.$
		\item Do the same for $\ch k = 2.$
	\end{enumerate}
\end{prob}
\begin{sol}
	\subsubsection*{Part (a)}
	Consider $\varepsilon:V\to k$ and consider $\ker\varepsilon\lhd V$. Since $\varepsilon$ cannot be the zero map,
	$\ker\varepsilon=kx$ for some $x\in V\setminus 0$. 
\end{sol}

\end{document}

### 2019-01-08 Lie Algebra Notes/Lie Algebras.tex
Add Course Notes and make updates

\documentclass[12pt]{article}

\usepackage{setspace}

\usepackage{amsmath, amsfonts, amssymb, graphicx, color, fancyhdr, lipsum, scalerel, stackengine, mathrsfs, tikz-cd, mdframed, enumitem, framed, adjustbox, bm, upgreek, x	color}
\usepackage[framed,thmmarks]{ntheorem}
\usepackage[mathscr]{euscript}

%set up theorem/definition/etc envs
%Problems will be created using their own counter and style
\theoreminframepreskip{0pt}
\theoreminframepostskip{0pt}
\newframedtheorem{prob}{Problem}[part]
\renewcommand\theprob{\arabic{part}.\arabic{prob}}

%solution template
\theoremstyle{nonumberbreak}
\theoremindent0.5cm
\theorembodyfont{\upshape}
\theoremseparator{:}
\theoremsymbol{\ensuremath\spadesuit}
\newtheorem{sol}{Solution}

%Theorems, Lemmas, and Corollaries
\theoremstyle{changebreak}
\theoremseparator{}
\theoremsymbol{}
\theoremindent0.5cm
\theoremheaderfont{\color{violet}\bfseries}

\newtheorem{thm}{Theorem}[subsection]
\theoremheaderfont{\bfseries}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{cor}[thm]{Corollary}

%Create a new env that references a theorem and creates a 'primed' version
%Note this can be used recursively to get double, triple, etc primes
\newenvironment{thm-prime}[1]
  {\renewcommand{\thethm}{\ref{#1}$'$}%
   \addtocounter{thm}{-1}%
   \begin{thm}}
  {\end{thm}}

\setlength\fboxsep{15pt}

%Shade definitions
\theoremindent0cm
\theoremheaderfont{\normalfont\bfseries} 
\def\theoremframecommand{\colorbox[rgb]{.9,.8,1}}
\newshadedtheorem{defn}[thm]{Definition}

%Man, that's really good! Let's use the same thing for definitons.
\newenvironment{def-prime}[1]
  {\renewcommand{\thethm}{\ref{#1}$'$}%
   \addtocounter{thm}{-1}%
   \begin{def}}
  {\end{def}}

%proofs
\theoremstyle{nonumberbreak}
\theoremindent0.5cm
\theoremheaderfont{\sc}
\theoremseparator{}
\theoremsymbol{\ensuremath\spadesuit}
\newtheorem{prf}{Proof}

%remarks
\theoremstyle{change}
\theoremindent0.5cm
\theoremheaderfont{\sc}
\theoremseparator{:}
\theoremsymbol{}
\newtheorem{rmk}[thm]{Remark}

%Replacement for the old geometry package
\usepackage{fullpage}

%Put page breaks before each part
\let\oldpart\part%
\renewcommand{\part}{\clearpage\oldpart}%

%Center each figure by default
\makeatletter
\g@addto@macro\@floatboxreset{\centering}
\makeatother

%header stuff
\setlength{\headsep}{24pt}  % space between header and text
\pagestyle{fancy}     % set pagestyle for document
\lhead{Notes on Lie Algebras} % put text in header (left side)
\rhead{Nico Courts} % put text in header (right side)
\cfoot{\itshape p. \thepage}
\setlength{\headheight}{15pt}
\allowdisplaybreaks

%Set of Integers
\newcommand*{\Z}{
\mathbb{Z}
}
%Set of Natural Numbers
\newcommand*{\N}{
\mathbb{N}
}
%Set of Real Numbers
\newcommand*{\R}{
\mathbb{R}
}
%Set of Complex Numbers
\newcommand*{\C}{
\mathbb{C}
}
%Rationals
\newcommand*{\Q}{
\mathbb{Q}
}

%Section break
\newcommand*{\brk}{
\rule{2in}{.1pt}
}

\DeclareMathOperator{\Aut}{Aut}

%raise that Chi!
\DeclareRobustCommand{\Chi}{{\mathpalette\irchi\relax}}
\newcommand{\irchi}[2]{\raisebox{\depth}{$#1\chi$}}

%Image
\DeclareMathOperator{\im}{Im}

%Coker
\DeclareMathOperator{\coker}{coker}

%characteristic
\DeclareMathOperator{\ch}{char}

%rank
\DeclareMathOperator{\rank}{rank}

%identity map
\DeclareMathOperator{\id}{id}

%Lie algebra stuff
\DeclareMathOperator{\gl}{\mathfrak{gl}}
\let\sl\relax
\DeclareMathOperator{\sl}{\mathfrak{sl}}
\DeclareMathOperator{\tr}{tr}

%fix tilde
\let\tilde\relax
\newcommand*{\tilde}[1]{\widetilde{#1}}

% Enumerate will automatically use letters (e.g. part a,b,c,...)
\setenumerate[0]{label=(\alph*)}

\begin{document}
%make the title page
\title{Lie Algebras and Groups\vspace{-1ex}}
\author{A course by: Monty McGovern\\
Notes by: Nico Courts}
\date{Winter 2019}
\maketitle

\renewcommand{\abstractname}{Introduction}
\begin{abstract}
	These notes are my best attempt at following along with our \textit{Math 508 --
	Lie Algebras} course at UW. This is my first time trying to type my 
	notes on-the-fly in class so we'll see how well this goes. The course reference
	is Humphreys' \textit{Introduction to Lie Algebras and Representation Theory.}

	The course description follows:

	\brk

	This is the second course in the Algebraic Structures sequence. I will classify 
	finite-dimensional complex semisimple Lie algebras, also proving some structural 
	results on general Lie algebras along the way. Although one usually first 
	encounters Lie algebras in a manifolds course, the treatment (following the text) 
	will be entirely algebraic.
\end{abstract}

\section{January 7, 2019}
The homework is posted on Monty's website. :) 
\subsection{Lie algebras}

This course will be studying Lie algebras, but as opposed to their treatment in manifolds, 
we will be studying them from a purely algebraic point of view. The book (Humphreys)
actually never defines a Lie group.

\begin{defn}
	A \textbf{Lie Algebra} $L$ or $\mathfrak{g}$ over a field $k$ is a $k$-vector space (usually f.d.)
	along with a \textit{bracket operation} $[vw]:L\times L\to L$ such that $[\cdot\cdot]$ is 
	\begin{itemize}
		\item anticommutative,
		\item bilinear,
		\item $[x[yz]]=[[xy]z]+[y[xz]]$
	\end{itemize}
\end{defn}

\begin{rmk}
	The last principle above is actually equivalent to the \textit{Jacobi identity:}
	\[[x[yz]]+[y[xz]]+[z[xy]].\]
	This follows from bilinearity and anticommutativity of the bracket.
\end{rmk}
The most natural place for these to arise is as \textit{derivations} on an algebra!
\begin{defn}
	A \textbf{$k$-derivation} $d:A\to A$ on an algebra $A$ over $k$ is a $k$-linear map
	satisfying the Leibniz rule.
\end{defn}
\begin{rmk}
	Some key facts about derivations (for us):
	\begin{itemize}
		\item Given a fixed $a\in A$, the map $d_a$ sending $b\mapsto ab-ba$, the \textbf{commutator}
		$[ab]$ is a derivation.
		\item If $d,e$ are derivations, then so is $[de]=de-ed$, where $de$ is the \textit{composite}
		of $d$ and $e$ as opposed to the product.
	\end{itemize}
\end{rmk}

\subsection{Examples}

A main source of Lie algebras is (associative) algebras! \textit{Any associative $k$-algebra $A$}
becomes a Lie algebra over $k$, taking $[ab]=ab-ba.$ In particular, one obvious choice for $k$-algebra
is $M_n(k)=\gl_n(k)$, the (Lie) algebra of $n\times n$ matrices over $k$.

\textbf{Lie subalgebras} are what you'd expect (including closure under brackets). Notice
that if $L'\le L$, then they \textbf{must both be over the same field.}

If $L$ is a $k$-Lie algebra and $I\lhd L$ is an ideal of $L$, then the quotient space $L/I$ becomes a 
Lie algebra with $[x+I,y+I]=[xy]+I$ as the bracket.

A \textbf{Lie algebra homomorphism} is a map $\varphi:L\to L'$ such that $\varphi$ is $k$-linear
and $\varphi([xy])=[\varphi(x)\varphi(y)].$

We get the usual first isomorphism theorem $L/\ker\varphi\cong \varphi(L).$

\brk

Associative algebras are not the only source of Lie algebras, however! One example is 
$\sl(n,k)=\{n\times n\text{ matrices over } k\text{ with trace zero}\}$

Note that this is \textbf{not closed under product} since $\tr(AB)\ne\tr A\tr B$ but $\tr(AB)=\tr(BA)$
so $\tr(AB-BA)=\tr(AB)-\tr(BA)=0$.
\begin{defn}
	We call this algebra (or, in fact any subalgebra of $\gl(n,k)$) \textbf{linear}. Think 
	``Linear'' means ``of matrices.''
\end{defn}

We say that $\sl(n,k)$ has \textbf{type} $A_{n-1}$. Eventually we will see seven types
$A-G$ of semisimple Lie algebras. The shift in index will emerge later.

$\sl(n,k)$ is, in fact, a simple Lie algebra: for $k=\C$, $\sl(n,\C)$ has no ideals
apart from the trivial ones.

\brk

Other non-associative examples include $k^n$ with a bilinear form $(\cdot,\cdot)$ which
is either symmetric or skew-symmetric and (in either case) is nondegenerate.
\begin{defn}
	$(\cdot,\cdot)$ is \textbf{nondegenerate} if the map $v\mapsto (v,\cdot)$ is injective. Equivalently
	there is no $v\in V$ such that $(v,w)=0$ for all $w\in V$.
\end{defn}

Given $V=k^n$ and a bilinear form on $V$, we can look at all $X\in \gl(n,k)=\gl(V)$ such that 
$(Xv,w)=(v,Xw)$. Then $X$ is \textbf{skew-adjoint} with respect to the form. One can check that 
$[XY]$ is skew-adjoint whenever both $X$ and $Y$ are.

\subsection{Generating (skew) symmetric forms}
It ends up that the dot product (which a symmetric form) is misleadingly simple -- thus
we will look elsewhere.

If $M\in \gl(n,k)$ is symmetric, so that $M^t=M$, then $(v,w)=v^tMw$ is a symmetric. If 
instead $M$ is skew-symmetric, then the same definition yields a skew-symmetric form. 
This actually induces a one-to-one correspondence between matrices and forms.

In both cases, if $M$ is invertible, then the form will be nondegenerate. As a consequence, 
since skew-symmetric matrices are always singular in odd dimensions, we see that 
nondegenerate skew-symmetric forms (over $\ch k\ne 2$ where the two families of forms
coincide) exist only in even dimensions.

\subsection{A peek at classifications}
If we have a nondegenerate symmetric form where $n=2m$ is going to give us an algebra
of type $D_m$. If $n=2m+1$, then it is of type $B_m$. Both of these cases are called
\textbf{orthogonal.}

If instead we have a skew-symmetric form and $n=2m$, then this is of type $C_m$, and we 
call this algebra \textbf{symplectic.}

\brk

We will make a particular choice for our matrix $M$ and then study the resulting
Lie algebras in much more detail next time. The choices will be: 
\begin{itemize}
	\item For type $D_m$:
	\[\begin{pmatrix}
		0 & I_m\\
		I_m & 0
	\end{pmatrix}\]
	\item For type $C_m$:
	\[\begin{pmatrix}
		0 & -I_m\\
		I_m & 0
	\end{pmatrix}\]
	\item For type $B_m$:
	\[\begin{pmatrix}
		1 & 0 & 0\\
		0 & 0 & I_m\\
		0 & I_m & 0
	\end{pmatrix}\]
\end{itemize}
\end{document}

### 2019-01-08 Reading Notes/reading_notes.tex
Add Course Notes and make updates

@@
 
-\usepackage{xcolor, amsmath, amsfonts, amssymb, graphicx, color, fancyhdr, lipsum, scalerel, stackengine, mathrsfs, tikz-cd, mdframed, enumitem, framed, adjustbox}
+\usepackage{amsmath, amsfonts, amssymb, graphicx, color, fancyhdr, lipsum, scalerel, stackengine, mathrsfs, tikz-cd, mdframed, enumitem, framed, adjustbox, bm, upgreek, x	color}
 \usepackage[framed,thmmarks]{ntheorem}
@@
 \theoremindent0.5cm
+\theoremheaderfont{\color{blue}\bfseries}
+
+\newtheorem{thm}{Theorem}[subsection]
 \theoremheaderfont{\bfseries}
-\newtheorem{thm}{Theorem}[subsection]
 \newtheorem{lem}[thm]{Lemma}
@@
 \theoremheaderfont{\normalfont\bfseries} 
-\def\theoremframecommand{\colorbox[rgb]{.9,.9,1}}
+\def\theoremframecommand{\colorbox[rgb]{.7,.9,1}}
 \newshadedtheorem{defn}[thm]{Definition}
@@
 %proofs
-\theoremstyle{nonumberbreak}
+\theoremstyle{nonumberplain}
 \theoremindent0.5cm
 \theoremheaderfont{\sc}
-\theoremseparator{}
+\theoremseparator{:}
 \theoremsymbol{\ensuremath\spadesuit}
@@
 \newcommand*{\SL}[1]{\ensuremath\mathbf{SL_#1}}
-\newcommand*{\mun}[1]{\ensuremath\mathbf{\mu_#1}}
-\newcommand*{\alp}[1]{\ensuremath\mathbf{\alpha_#1}}
+\newcommand*{\mun}[1]{\ensuremath\mathbf{\bm\upmu_#1}}
+\newcommand*{\alp}[1]{\ensuremath\mathbf{\bm\upalpha_#1}}
 
@@
 as learn about algebraic geometry (alongside my usual classes, of course).
-\section{Preparation, Waterhouse, and G\"ortz \& Weddhorn}
+\section{Affine Group Schemes (Waterhouse)	}
 To give a sense of direction, Julia recommended that I take a look at the following regularity theorem:
@@
 
+\begin{defn}
+	A group scheme $G$ represented by $A$ that consists entirely of group-like elements is
+	called \textbf{diagonalizable.}
+\end{defn}
+\begin{rmk}\label{rmk-diag}
+	Notice that an alternative definition is to begin with an abelian group $M$ and to 
+	define $S,\Delta,$ and $\varepsilon$ such that each element is group-like. The resulting
+	Hopf algebra represents a diagonalizable group scheme.
+\end{rmk}
+
+Consider the group algebra $k[M]$ on the group in Remark \ref{rmk-diag}. If this algebra
+is finitely generated, we get the following:
+\begin{thm}
+	Let $G$ be diagonalizable and represented by $A$ and assume $A$ is finitely generated
+	as a $k$-algebra. Then $G$ is a finite product of copies of $\Gm$ and $\mun n$.
+\end{thm}
+\begin{prf}
+	Let $x_1,\dots, x_n$ be generators for $k[M]=A$. Each $x_i$ can be written as a (finite!)
+	$k$-linear combination of elements in $M$, so we can instead use the finitely many $m_i$
+	which generate these generators, and this gives us a $k$-algebra basis for $k[M]$. Call this
+	new generating set $U$.
+
+	Let $M'$ be the abelian group generated by $U$ and notice that $k[M']$ is a subalgebra of
+	$k[M]$ containing $U$ so $k[M']=k[M]$ and therefore $M'=M$. This establishes that $M$ is 
+	a finitely generated abelian group, so we can split up the algebra into a tensor product
+	of $k[\Z]$ and $k[\Z/n\Z]$.
+
+	When $M=\Z$, 
+\end{prf}
+
+\section{Algebraic Geometry}
+
 \section{Problems}

### 2019-01-10 Hopf Algebra Notes/Hopf Algebras.tex
update notes

@@
 %Problems will be created using their own counter and style
+\theoremstyle{break}
 \theoreminframepreskip{0pt}
@@
 	If $A$ is a finite dimentional Hopf algebra, then 
-	\[\bigoplus_{i\ge 0}\Ext_A^i(k,k)\]
+	\[\bigoplus_{i\ge 0}\Ext_A^i(_Ak, _Ak)\]
 	is Noetherian.
@@
 \end{prob}
+\begin{sol}
+	Assume that $(V,m,u,\Delta,\varepsilon)$ is an algebra and coalgebra and further that
+	$m$ and $u$ are coalgebra morphisms. That means in particular that the diagrams in figure \ref{fig-coalg-mor} commute.
+	\begin{figure}[h]\label{fig-coalg-mor}
+		\centering
+		\begin{tikzcd}
+			V\otimes V \ar[d,swap,"\Delta\otimes\Delta"]\ar[r,"m"] & V\ar[r,"\Delta"] & V\otimes V\\
+			V^{\otimes 4}\ar[rr,"T_{2,3}"] & & V^{\otimes 4}\ar[u,swap,"m\otimes m"]
+		\end{tikzcd}
+		\begin{tikzcd}
+			k\ar[r,"\Delta"]\ar[d,"u"] & k\otimes k\ar[d,"u\otimes u"]\\
+			V\ar[r,"\Delta"] & V\otimes V
+		\end{tikzcd}
+		\label{fig-coalg-mor}
+		\caption{$m$ and $u$ are coalgebra morphisms.}
+	\end{figure}
+
+	We are looking to prove that $\Delta$ and $\varepsilon$ are algebra morphisms, or that 
+	the diagrams in figure \ref{fig-coalg-mor}  commute.
+	\begin{figure}[h]
+		\centering
+		\begin{tikzcd}
+			V\otimes V\ar[r,"m"]\ar[d,"\Delta\otimes \Delta"] & V\ar[d,"\Delta"]\\
+			V^{\otimes 4}\ar[r,"m"] & V\otimes V
+		\end{tikzcd}
+		\begin{tikzcd}
+			V\otimes V\ar[r,"\varepsilon\otimes\varepsilon"]\ar[d,"m"] & k\otimes k\ar[d,"m"]\\
+			V\ar[r,"\varepsilon"] & k
+		\end{tikzcd}
+		\label{fig-alg-mor}
+		\caption{$\Delta$ and $\varepsilon$ are algebra morphisms.}
+	\end{figure}
+
+	From here it's actually a bit boring because it's kinda just a definition/notation game.
+	It boils down to the fact that the (co)multiplication on $V\otimes V$ has a twist that 
+	exactly lines up so that each square is saying the same thing.
+\end{sol}
 
@@
 	\subsubsection*{Part (a)}
-	Consider $\varepsilon:V\to k$ and consider $\ker\varepsilon\lhd V$. Since $\varepsilon$ cannot be the zero map,
-	$\ker\varepsilon=kx$ for some $x\in V\setminus 0$. 
+	Consider $\varepsilon:V\to k$ and consider $\ker\varepsilon\lhd V$. By rank-nullity, $\dim\ker\varepsilon= 1$, so
+	$\ker\varepsilon=kx$ for some $x\in V$. Therefore $x^2=cx$ for some $c$, and if $c=0$, then (as an algebra) $V\cong k[x]/(x^2)$.
+	Otherwise consider $y=\frac{x}{c}$. In this case $y^2=\frac{x^2}{c^2}=\frac{x}{c}=y$, and $V\cong k[x]/(x^2-x)$.
+
+	Notice that in either case $\varepsilon(x)=0$, so let
+	\[\Delta(x)=a(1\otimes 1)+b(1\otimes x)+c(x\otimes 1) + d(x\otimes x)\]
+	and using that $\varepsilon\otimes \id\circ\Delta=\id\otimes\varepsilon\circ\Delta$
+	and that each should be (essentially) the identity (this is just the diagram we saw before),
+	we get $a=0$ and $b=c=1$. Thus the coalgebra structure of any Hopf algebra is given by 
+	\[\varepsilon(x)=0,\quad \Delta(x)=1\otimes x+x\otimes 1+d(x\otimes x).\]
+
+	Consider first the case when $x^2=0$. Then since comultiplication will be an algebra morphism,
+	\[0=\Delta(x^2)=\Delta(x)^2=1\otimes x^2+x^2\otimes 1+d^2(x^2\otimes x^2)+2(x\otimes x)+2d(x\otimes x^2)+2d(x^2\otimes x)\]
+	and since $x^2=0$, 
+	\[0=2(x\otimes x).\]
+	But $x\otimes x$ is a basis element of $V\otimes V$, so $V$ can only have this algebra structure
+	when $\ch k=2$. We will return to this in the next part.
+
+	So then $x^2=x$ and using the computation above,
+	\[1\otimes x+x\otimes 1+d(x\otimes x)=\Delta(x)=\Delta(x^2)=1\otimes x+x\otimes 1+(d^2+4d+2)(x\otimes x)\]
+	so
+	\[(d^2+3d+2)(x\otimes x)=0\quad\Rightarrow\quad d^2+3d+2=(d+2)(d+1)=0\]
+	and so either $d=-1$ or $d=-2$.
+
+	One can verify that $\Delta$ is coassociative, so we can conclude that when $\ch k\ne 2$, 
+	there are precisely two Hopf algebra structures with algebra structure $k[x]/(x^2-x)$ and
+	comultiplication either
+	\[\Delta(x)=1\otimes x+x\otimes 1-x\otimes x\quad\text{or}\quad \Delta(x)=1\otimes x+x\otimes 1-2(x\otimes x)\]
+
+	\subsection*{(b)}
+	Now assume that $\ch k=2$ and that $V\cong k[x]/(x^2-x)$ as an algebra. Then using the analysis above, 
+	we see that we can choose comultiplication either
+	\[\Delta(x)=x\otimes 1+1\otimes x\quad\text{or}\quad \Delta(x)=1\otimes x+x\otimes 1+x\otimes x.\]
+
+	If instead $V\cong k[x]/(x^2)$, then \textit{any} value of $d$ will suffice, so there are a full $k$'s worth
+	of Hopf algebra structures that can appear.
 \end{sol}
 
+\section{January 9, 2019}
+Today we are going to rely heavily on Sweedler notation. :) Notice that if we are looking at actual objects in the 
+diagram for coassociativity, we get
+\begin{figure}[h]
+	\centering
+	\begin{tikzcd}
+		v\ar[r]\ar[d] & \sum v_{(1)}\otimes v_{(2)}\ar[d]\\
+		\sum v_{(1)}\otimes v_{(2)}\ar[r] & \sum (v_{(1)})_{(1)}\otimes (v_{(1)})_{(2)}\otimes v_{(2)}=\sum v_{(1)}\otimes (v_{(2)})_{(1)}\otimes (v_{(2)})_{(2)}
+	\end{tikzcd}
+	\caption{Coassociativity on elements in Sweedler notation}
+\end{figure}
+
+\begin{ex}
+	Let $G$ be a group and $kG$ be the group algebra. The algebra structure arises as normal
+	where $g\cdot h$ comes from the structure on $G$. Then $\Delta(g)=g\otimes g$ and this extends linearly.
+
+	But then if you consider $\Delta(\sum c_gg)$, notice that by the nature of tensors this is not unique!
+	So we will just write
+	\[\Delta\left(\sum c_gg\right)=\sum_Gc_g(g\otimes g)=\sum v_{(1)}\otimes v_{(2)}\]
+\end{ex}
+
+\subsection{Algebra structure on $V\otimes V$}
+We said earlier on that $\Algk$ is a symmetric \textit{tensor} category. But how do we
+define the multiplication on the tensor product?
+
+Well it all comes from the twist! We define
+\[m_{V\otimes W}=(m_V\otimes m_W)\circ(\id_V\otimes \tau_{2,3}\otimes \id_W)\]
+where $\tau_{2,3}$ is the twist morphism.
+
+More simply, $u_{V\otimes W}:k=k\otimes k\to V\otimes V$ simply defined by $u_V\otimes u_W$.
+
+So then when we say that $\Delta$ is an algebra morphism, we are saying that for all $v,w\in V$
+\[\Delta(vw)=\sum(vw)_{(1)}\otimes(vw)_{(2)}=(\sum v_{(1)}\otimes v_{(2)})(\sum w_{(1)}\otimes w_{(2)})=\sum v_{(1)}w_{(1)}\otimes v_{(2)}w_{(2)}\]
+
+\subsection{Hopf Algebras}
+Already to the good stuff!
+\begin{defn}
+	$V\in\Vectk$ is a \textbf{Hopf algebra} if $V$ is a bialgebra together with an \textbf{antipode}
+	$S:V\to V$ satisfying
+	\[(S,\id_V)\circ \Delta=\varepsilon=(\id_V,S)\circ\Delta\]
+\end{defn}
+\begin{conj}
+	If $V\in\Vectk$ is a Noetherian Hopf algebra, then $S$ is bijective.
+\end{conj}
+
+\subsection{History and Motivation}
+Hopf himself was a topologist, so this is the first context in which it arose. In the 1940's,
+he began studying Hopf algebras over $\Z_2$ graded $k$ vector spaces. For instance, the 
+cohomology ring of topological space $X$ with coefficients in $k$.
+
+Later, in combinatorics, they ended popping up. Looking at rings of symmetric functions and other places
+gave some interesting examples.
+
+Then in group theory you can define a functor from groups to Hopf algebras by $F(G)=kG$
+with the diagonal map. The antipode is just the inverse.
+
+Then with Lie algebras, you can look at $\mathcal{U}(L)$, the universal enveloping algebra is a Hopf algebra.
+
+Finally with algebraic groups (yay!) we take an algebraic group $G$ and consider the ring of functions on it, 
+which is again a Hopf algebra.
+
+\brk
+
+Some ``cousins'' of Hopf algebras: quasi, weak, multiplier, ribbon, quasi-triangular, etc Hopf algebras. 
+Each has slightly different base category or restrictions.
+
 \end{document}

### 2019-01-10 Lie Algebra Notes/Lie Algebras.tex
update notes

@@
 \DeclareMathOperator{\sl}{\mathfrak{sl}}
+\DeclareMathOperator{\so}{\mathfrak{so}}
+\let\sp\relax
+\DeclareMathOperator{\sp}{\mathfrak{sp}}
 \DeclareMathOperator{\tr}{tr}
+\DeclareMathOperator{\ad}{ad}
 
@@
 Given $V=k^n$ and a bilinear form on $V$, we can look at all $X\in \gl(n,k)=\gl(V)$ such that 
-$(Xv,w)=(v,Xw)$. Then $X$ is \textbf{skew-adjoint} with respect to the form. One can check that 
+$(Xv,w)=(v,Xw)$. Then $X$ is \textbf{adjoint} with respect to the form. There is a similar definition for when
+$X$ is \textbf{skew-adjoint.} One can check that 
 $[XY]$ is skew-adjoint whenever both $X$ and $Y$ are.
@@
 \subsection{Generating (skew) symmetric forms}
-It ends up that the dot product (which a symmetric form) is misleadingly simple -- thus
+It ends up that the dot product (which is a symmetric form) is misleadingly simple -- thus
 we will look elsewhere.
@@
 \end{itemize}
+
+\section{January 9, 2019}
+Today we will be looking deeply into the stucture of linear Lie algebras of types A-D.
+
+\subsection{Linear Lie Algebras Revisited}
+Recall that the \textbf{matrix unit} $e_{ij}$ is the matrix with 1 in the $(i,j)$ entry and 
+zero elsewhere. And then $e_{ij}e_{kl}=\delta_{jk}e_{il}$ and furthermore
+\[[e_{ij}e_{kl}]=\delta_{jk}e_{il}-\delta_{li}e_{kj}.\]
+This is especially nice when $j=i$, called the \textbf{diagonal matrix unit}.
+
+Then we look at type $A_{n-1}$ ($\sl(n,k)$). Let $D$ be the set of diagonal matrices in this algebra.
+Notice the dimension is $n-1$ since then $n^{th}$ term on the diagonal is determined as the
+negative of the sum of the other $n-1$ terms. Let $A=\operatorname{diag}(d_1,\dots,d_n)$. Then consider the eigenvalues
+associated with $e_{ij}$: 
+\[[Ae_{ij}]=(d_i-d_j)e_{ij}=(E_i-E_j)Ae_{ij}\]
+where $E_i$ is the linear functional selecting the $i^{th}$ entry in $A$. Moreover, $D$ is abelian as a Lie algebra, so $D$
+acts diagonally on $L=\sl(n,k)$ by commutation with eigenvalues $E_i-E_j$ and zero
+for $1\le i,j\le n$ and $i\ne j$.
+
+In the other classical cases B-D, there is always a matrix $M$ which defines the form $(v,w)=v^tMw$ 
+as we saw yesterday. In all three cases, the Lie algebra exists consists of all skew-adjoint matrices
+$X$ relative to the form. $B_m=\mathfrak{so}(2m+1,k)$ as well as $D_m=\mathfrak{so}(2m,k)$ and $C_m=\mathfrak{sp}(2m,k)$.
+
+This condition translates to the form of matrices in the above Lie groups and the condition is always $Mx=-x^tM$ in all cases.
+
+Type B:
+\[\begin{pmatrix}
+	0 & b_1 & b_2\\
+	c_1 & m & n\\
+	c_2 & p & q
+\end{pmatrix}\]
+where $c_1=-b_2^t$, $c_2=-b_1^t$, $q=-m^t$, $n^t=-n$ and $p^t=-p$.
+
+Type C:
+\[\begin{pmatrix}
+	m& n\\ p & q
+\end{pmatrix}\]
+where $n^t=n, p^t=p,$ and $m^t=-q$.
+
+Type D:
+\[\begin{pmatrix}
+	m& n\\ p & q
+\end{pmatrix}\]
+where $n^t=-n, p^t=-p,$ and $m^t=-q$.
+
+Looking at the eigenvalues of elements of $D$ associated to vectors $e_{ij}-e_{m+i,m+j}$. Look at photos
+
+Using a similar analysis, we can look at types $B$ and $D$. We define the functions $E_i$ similarly on the space of diagonalmatrices and gives a rise to the following collection of linear functions:
+in $B_m:$ $\pm E_i$ and $\pm(E_i\pm E_j)$ and $D_m$ gives us $\pm(E_i\pm E_j)$.
+
+This collection of functions in each case is called the \textbf{root system of the Lie algebra},
+Then any complex simple finite-dimensional Lie algebra is classified by its root system.
+The (perhaps surprising) fact is that this already encompasses all but finitely many of these
+things up to ismorphism: the classical Lie algebras. Eventually we will learn more about the
+\textbf{exceptional Lie groups.}
+
+This section was a little hard to follow and the handling in Humphreys is easier to follow, 
+but delays speaking about root systems and actually deriving the eigenfunctions (is that the right word?)
+until significantly later. Monty seemed to think it was acceptable to delay the understanding of this a bit.
+
+\subsection{Derivations and $\exp$}
+Look an an arbitrary Lie algebra over a field $k$ where $\ch k=0$ (which we will mostly be assuming from here on)
+. Let $\delta$ be a derivation of $L$, so that $[\delta x,y]+[x\delta y]=\delta[x,y]$. Assume that $\delta$ is nilpotent.
+
+Then the ``power series'' (polynomial) is
+\[\exp\delta=\sum_{i>0}\frac{\delta^i}{i!}\]
+\begin{prob}
+	This is a good exercise to go through: Check that
+	\[[(\exp\delta)x,(\exp \delta)y)=[xy]\]
+	for each $x,y\in L$
+\end{prob}
+
+\begin{rmk}
+	This actually shows that $\exp\delta$ is an automorphism of $L$. Furthermore you'd find that
+	\[(\exp\delta)(\exp(-\delta))=1.\]
+\end{rmk}
+
+What if $k=\R$ or $\C$? Then the power series (even when $\delta$ is not nilpotent!) always converges
+and defines an automorphism as before.
+
+\begin{lem}
+	For all \textbf{complex} semisimple Lie algebras $L$ it turns out that the group
+	generated by $\exp\ad x$ ($\ad x(y)=[xy]$) coincides with the group generated by all 
+	nilpotent $\ad x$.
+\end{lem}
+\subsection{Adjoint group}
+The last thing for today is to define the adjoint group:
+\begin{defn}
+	Let $L$ be a Lie algebra, then 
+	\[\operatorname{Int}(L)=\exp\ad L\]
+	is the \textbf{Adjoint group} of $L$. It is a subgroup (so we believe) of the Lie
+	Group associated to $L$.
+\end{defn}
+
+Some examples of adjoint groups:
+\begin{itemize}
+	\item If $L=\sl(n,\C)$, then $\operatorname{Int}(L)=PSL(n,\C)=SL(n,\C)/\text{center}$
+	\item If $L=\so(n,\C)$, then $\operatorname{Int}(L)=PSO(n,\C)$
+	\item If $L=\sp(2n,\C)$ then $\operatorname{Int}(L)=PSp(2n,\C)$.
+\end{itemize}
+
 \end{document}

### 2019-01-10 Reading Notes/reading_notes.tex
update notes

@@
 \part{Winter 2019}
+\section{Preparation for the Quarter}
 This is my first official quarter as Julia's student! My plan for now is to continue working on Waterhouse as well
 as learn about algebraic geometry (alongside my usual classes, of course).
-\section{Affine Group Schemes (Waterhouse)	}
 To give a sense of direction, Julia recommended that I take a look at the following regularity theorem:
@@
 \end{thm}
+\section{Affine Group Schemes (Waterhouse and Jantzen)	}
 I have quite a bit of information to process before this, so I will get started!
@@
 
-	When $M=\Z$, 
+	When $M=\Z$, $k[M]\cong k[X,X^{-1}]$, so $G= \Gm$\footnote{Notice that here Waterhouse also
+	uses that the basis elements $e_i$ are group-like and so $\Delta(e_i)=e_i\otimes e_i$. I am not
+	quite sure why this is necessary.} Similarly if $M=\Z/n$ then $G\cong \mun{n}$
 \end{prf}
+
+Looking at my progress over the last several days, I am thinking that I am getting
+too much in the weeds writing out all the lemmas and proofs. Perhaps it is better to read more quickly
+and only take note of theorems as I need/use them.
+
+\subsection{Cartier Duals}
+The gist here is that if $G$ is represented by $A$, then we can define the dual 
+$A^D=\Hom(A,k)$ and then $A^D$ represents a new scheme $G^D$ called the \textbf{Cartier Dual.}
+
+We do some work to show that, in fact, elements of $G^D$ are in correspondence with the 
+group-like elements of $A$, or equivalently the character group of $G$, $X_G$.
+
+It is easy enough to evaluate $G^D(k)\cong\Hom(A^D,k)$, but luckily one shows that
+``dualizing'' commutes with tensor products (and thus with extension of scalars) and $\Hom$. Thus
+if $G_R$ is the scheme represented by $A\otimes_k R$, then using that $\Hom_k(A^D,R)\cong\Hom_R(A^D\otimes R,R)$,
+\[G^D(R)=(G^D)_R(R)\]
+which is represented by $(A\otimes R)^D=A^D\otimes R$, so finally 
+\[G^D(R)=(G_R)^D(R)=\{\text{group-like elements in }A\otimes_k R\}\cong \Hom(G_R,(\Gm)_R)\]
 
@@
 \subsection{Waterhouse}
+\begin{wprob}{2.1}
+	\begin{enumerate}
+		\item Show that there are no nontrivial homomorphisms from $\Gm$ to $\Ga$.
+		\item If $k$ is reduced, show that there are no nontrivial homomorphisms from $\Ga$ to $\Gm$.
+		\item For each nonzero $b\in k$ with $b^2=0$, find a nontrivial homomorphism $\Ga\to\Gm$.
+	\end{enumerate}
+\end{wprob}
+Hmm, there must be a small bug in my macro but I can't find it. I get an error about \texttt{missing `\textbackslash item'}
+in the definition below. It seems to only occur when I use \texttt{\textbackslash subsubsection*} and the like.
+
+\begin{sol}
+	\subsubsection*{(a)}
+	Let $\Phi:\Gm\to\Ga$ and let $\varphi:k[X]\to k[X,X^{-1}]$ be the corresponding Hopf
+	algebra map. But then we have that 
+	\[\varphi\otimes\varphi \circ \Delta_a(X)=\Delta_m\circ\varphi(X)\]
+	and so using that $\Delta_m(a)=a\otimes a$ and $\Delta_a(b)=b\otimes1+1\otimes b$, we get
+	\[\varphi(X)\otimes\varphi(1)+\varphi(1)\otimes\varphi(X)=\varphi(X)\otimes\varphi(X)\]
+
+\end{sol}
 

### 2019-01-11 Hopf Algebra Notes/Hopf Algebras.tex
Update Notes

@@
 
+\DeclareMathOperator{\Hom}{Hom}
+
 %Hopf algebra stuff
@@
 \newcommand*{\Coalgk}{\operatorname{Coalg}_k}
+\DeclareMathOperator{\Coalg}{Coalg}
+\DeclareMathOperator{\Alg}{Alg}
+\newcommand*{\scrT}{\ensuremath\mathscr{T}}
 
@@
 
+\section{January 11, 2019}
+The plan for today is to talk about:
+\begin{itemize}
+	\item Convolution Algebras
+	\item Antipodes
+	\item Duality
+	\item (Co-)Modules
+\end{itemize}
+
+\subsection{Convolution Algebras}
+ Let $\scrT$ be a symmetric tensor category. We can usually think of $\scrT=\Vectk$, but 
+ there is a problem since $\Vectk$ is equivalent to the category of Hopf algebras over $k$,
+ while this is not generally true.
+
+ We also need that $\scrT$ is $k$-linear (that is, enriched as a category over $k$). This
+ means that $\Hom_\scrT (A,B)\in\Vectk$.
+
+ \begin{thm}
+	Let $\scrT$ be as above. Then $\Hom_\scrT(C,A)$ is an algebra and $\Hom_\scrT(-,-):(\Coalg_\scrT)^{op}\times\Alg_\scrT\to\Alg_k$
+	is a functor.
+ \end{thm}
+ \begin{prf}
+	Let $A$ be an algebra object in $\scrT$ and $C$ be a coalgebra object in $\scrT$. Then $1_{\Hom} :=u_A\circ\varepsilon_C:C\to 1_\scrT$ and define
+	\[f\ast g := m_A(f\otimes g)\Delta_C:C\to A.\]
+
+	Then using Lemma \ref{lem-func} and the fact that $A$ and $C$ are (co)algebra objects,
+	we can see that the product $\ast$ satisfies the axioms required.
+ \end{prf}
+ \begin{lem}
+	\begin{enumerate}
+		\item $1_{\Hom}\ast f = m_A(u\otimes 1)(1\otimes f)(\varepsilon\otimes 1)\Delta.$
+		\item $f\ast 1_{\Hom}=m_A(1\otimes u)(f\otimes 1)(1\otimes \varepsilon)$
+		\item $(f\ast g)\ast h=m_A(m_A\otimes 1)((f\otimes g)\otimes h)(\Delta_C\otimes 1)\Delta_C$
+		\item $f\ast(g\ast h)=m_A(1\otimes m_A)(f\otimes(g\otimes h))(1\otimes \Delta)\Delta$
+	\end{enumerate}
+	\label{lem-func}
+ \end{lem}
+
+ \begin{prf}
+	\subsubsection*{(a)}
+	\begin{align*}
+		1_{\Hom}\ast f&= m_A(1_{\Hom}\otimes f)\Delta_C\\
+		&=m_A(u_A\circ\varepsilon\otimes f)\Delta_C\\
+		&=m_A(u\otimes 1)(\varepsilon\otimes 1)(1\otimes f)\Delta
+	\end{align*}
+	\subsubsection*{(b)}
+	Same as (a), essentially.
+	\subsubsection*{(c) and (d)}
+	\begin{align*}
+		(f\ast g)\ast h&= m_A((f\ast g)\otimes h)\Delta\\
+		&=m_A((m_A(f\otimes g)\Delta)\otimes h)\Delta\\
+		&=m_A(m_A\otimes 1)((f\otimes g)\otimes h)(\Delta\otimes 1)\Delta
+	\end{align*}
+	and the other is analogous.
+ \end{prf}
+
+ \begin{defn}
+	$V\in\scrT$ is a \textbf{Hopf algebra object} if:
+	\begin{itemize}
+		\item $V$ is a bialgebra object in $\scrT$ and
+		\item There is a map $S:V\to V$ that is $(\id_V)^{-1}$ with respect to $\ast$.
+	\end{itemize}
+ \end{defn}
+ \begin{rmk}
+	Notice here that $\id_V\in\Hom_\scrT(V,V)$, the identity map in $\scrT$. We are \textit{not}
+	taking about $1_{\Hom}=u\circ\varepsilon$.
+
+	Also, we call $S$ an \textbf{antipode.}
+ \end{rmk}
+
+ \subsection{Duality}
+ Notice that when $C=1_\scrT$ (that is the tensor identity), $\Hom_\scrT(1_\scrT,-):\Alg_\scrT\to\Algk$
+ is a functor. Same for the dual from $\Coalg_\scrT$. This second one gives us a chance to talk about duality.
+
+ \begin{lem}
+	Let $\scrT$ be the category of finite dimensional vector spaces over $k$.
+	Then $(-)^*:\scrT\to\scrT^{op}$ is an equivalence.
+ \end{lem}
+ This uses $(V\otimes W)^*=W^*\otimes V^*$.
+ \begin{cor}
+	$V$ is an algebra over $k$ $\Leftrightarrow$ $V^*$ is a coalgebra over $k$.
+	And vice versa.
+ \end{cor}
+
+ Recall $S=(\id_V)^{-1}$. Thus
+ \[S\ast \id_V=1_{\Hom}=\id_V\ast S\]
+ The diagram we have here is
+ \begin{figure}[h]
+	 \centering
+	 \begin{tikzcd}
+		V\otimes V\ar[rr,"S\otimes \id_V"] & & V\otimes V\ar[d,"m_V"]\\
+		V\ar[u,"\Delta"]\ar[d,"\Delta"]\ar[r,"\varepsilon_V"] & k \ar[r,"u_V"] & V\\
+		V\otimes V\ar[rr,"\id_V\otimes S"] & & V\otimes V\ar[u,"m_V"]
+	 \end{tikzcd}
+ \end{figure}
+
+ \subsection*{Modules/Comodules}
+ \begin{defn}
+	Let $A$ be an algebra object in $\scrT$. A \textbf{left $A$ module} is $M\in\scrT$
+	with a morphism
+	\[m_M:A\otimes M\to M\]
+	such that the diagrams in Figure \ref{fig-mod} commute.
+\end{defn}
+\begin{figure}[h]\label{fig-mod}
+	\centering
+	\begin{tikzcd}
+		A\otimes A\otimes M\ar[r,"m_A\otimes 1"]\ar[d,"1\otimes m_M"] & A\otimes M\ar[d,"m_M"]\\
+		A\otimes M\ar[r,"m_M"] & M
+	\end{tikzcd}
+	\begin{tikzcd}
+		A\otimes M\ar[r,"m_M"] & M\\
+		1_\scrT\otimes M\ar[ur,"\sim"]\ar[u,"u_A\otimes 1"] &
+	\end{tikzcd}
+	\caption{Module diagrams}
+\end{figure}
+\begin{rmk}
+	Note that we don't necessarily need that $M$ lie in $\scrT$. We could instead just 
+	rely on an algebra homomorphism $\varphi:A\to \Hom_{\scrT}(M,M)$ and proceed as usual.
+\end{rmk}
+
 \end{document}

### 2019-01-11 Lie Algebra Notes/Lie Algebras.tex
Update Notes

@@
 \theoreminframepostskip{0pt}
-\newframedtheorem{prob}{Problem}[part]
-\renewcommand\theprob{\arabic{part}.\arabic{prob}}
+\newframedtheorem{prob}{Problem}[section]
+\newenvironment{hwprob}[1]
+{\renewcommand{\theprob}{#1}%
+ \addtocounter{thm}{-1}%
+ \begin{prob}}
+{\end{prob}}
 
@@
 %Note this can be used recursively to get double, triple, etc primes
-\newenvironment{thm-prime}[1]
+\newenvironment{thmprime}[1]
   {\renewcommand{\thethm}{\ref{#1}$'$}%
@@
 %Man, that's really good! Let's use the same thing for definitons.
-\newenvironment{def-prime}[1]
+\newenvironment{defprime}[1]
   {\renewcommand{\thethm}{\ref{#1}$'$}%
    \addtocounter{thm}{-1}%
-   \begin{def}}
-  {\end{def}}
+   \begin{defn}}
+  {\end{defn}}
 
@@
 
+%some nice shortcuts
+\DeclareMathOperator{\calB}{\mathcal{B}}
+
 %Lie algebra stuff
@@
 \DeclareMathOperator{\ad}{ad}
+\DeclareMathOperator{\Der}{Der}
+\DeclareMathOperator{\Int}{Int}
+\DeclareMathOperator{\Rad}{Rad}
+\let\t\relax
+\DeclareMathOperator{\t}{\mathfrak{t}}
+\DeclareMathOperator{\n}{\mathfrak{n}}
+\DeclareMathOperator{\diag}{diag}
 
@@
 \end{defn}
+\begin{rmk}
+	Actually after talking to Monty, $\exp\ad L$ \textit{is} (essentially) the Lie group
+	corresponding to $L$. Such a group is not unique, however.
+\end{rmk}
 
@@
 
+%%%% HW 1 - Due Jan 18 %%%%
+\newpage
+
+\section*{HW 1 -- Due January 18}
+Do problems 1.9, 2.1, 3.8. 3.9, and 4.3.
+
+\begin{hwprob}{1.9}
+	When $\ch k=0$, show that each classical Lie algebra $A_l,B_l, C_l$ and $D_l$ satisfies $[LL]=L$. (This shows again that each algebra consists of trace zero matrices.)
+\end{hwprob}
+\begin{sol}
+	\subsubsection*{Type $A_k$:}
+	Recall that a basis for $A_l$ is the collection $\mathcal{B}=\{e_{ij}|i\ne j\}\cup\{e_{ii}-e_{i+1,i+1}|1\le i\le l\}$. Using linearity of the bracket,
+	it then suffices to check that $\calB$ is contained in the algebra generated by $[\calB \calB]$. To see this, we use that
+	\[[e_{ij}e_{kl}]=\delta_{jk}e_{il}-\delta_{li}e_{kj}.\]
+
+	First, notice we have
+	\[[e_{i,i+1}e_{i+1,i}]=e_{ii}-e_{i+1,i+1},\]
+	so it suffices to show we also have the $e_{ij}$ for $i\ne j$.
+
+\end{sol}
+
+\begin{hwprob}{2.1}
+	Prove that the set of all inner derivations $\ad x$ for $x\in L$ is an ideal of $\Der L$.
+\end{hwprob}
+\begin{sol}
+
+\end{sol}
+
+\begin{hwprob}{3.8}
+	Let $L$ be nilpotent. Prove that $L$ has an ideal of codimension 1.
+\end{hwprob}
+\begin{sol}
+
+\end{sol}
+
+\begin{hwprob}{3.9}
+	Prove that every nilpotent Lie algebra L has an outer derivation (see 1.3) as follows: Write $L=K+Fx$
+	for some ideal $K$ of codimension 1. Then $C_L(K)\ne 0$ (why?). Choose $n$ so that $C_L(K)\subseteq L^n$,
+	$C_L(K)\not\subseteq L^{n+1}$ and let $z\in C_L(K)-L^{n+1}$. Then the linear map $\delta$ sending $K$ to zero
+	and $x$ to $z$ is an outer derivation.
+\end{hwprob}
+\begin{sol}
+
+\end{sol}
+
+\begin{hwprob}{4.3}
+	This exercise illustrates the failure of Lie's theorem when $F$ is allowed to have positive characteristic.
+	Consider the $p\times p$ matrices
+	\[x=\begin{pmatrix}
+		0 & 1 & 0 & \cdots & 0\\
+		0 & 0 & 1 & \cdots & 0\\
+		\vdots & \vdots & \vdots & \ddots & \vdots\\
+		0 & 0 & 0 & \cdots & 1\\
+		1 & 0 & 0 & \cdots & 0
+	\end{pmatrix}\quad y=\diag(0,1,\dots, p-1)\]
+	Check that $[xy]=x$, hence that $x$ and $y$ span a two dimensional solvable subalgebra $L$ of $\gl(p,F)$. Verify that $x$ and $y$ have no common eigenvector.
+\end{hwprob}
+\begin{sol}
+
+\end{sol}
+\newpage
+%%%% End HW 1 %%%%
+
+\section{January 11, 2019}
+A few more things about $\Int L$ in classical cases:
+
+Begin with $n\times n$ matrices $M$ over $\R$ or $\C$. Given any such $M$, we have
+\[e^M=\sum_{i>0}\frac{M^i}{i!}\]
+always converges (that is, the series for each entry converges). Futhermore $e^Me^{-M}=I$.
+\begin{rmk}
+	Notice that over \textit{any} field over characteristic zero whenever $M$ is nilpotent 
+	this still holds.
+\end{rmk}
+
+Now if $M$ is skew-adjoint with respect to a bilinear form $(\cdot,\cdot)$, then 
+\[(M^iv,w)=(-1)^i(v,M^iw)\]
+whence
+\[(e^Mv, w)=(v,e^{-M}w)\]
+and so 
+\[(e^Mv,e^Mw)=(v,w)\]
+for all $v,w\in\R^n$ or $\C^n$.
+
+But notice that then $e^M$ is an \textbf{isometry of} or \textbf{preserves} the form. Moreover, 
+if we set $A=\ad M$, then $A$ is given as the difference of two \textit{commuting} linear maps, 
+left and right multiplication by $M$.
+
+\brk
+
+Using the binomial theorem: $e^AN=e^MNe^{-M}$. This, in a nutshell, is why the formulas for
+$\Int L$ all involved modding out by the center, since conjugation by a scalar matrix is trivial! :)
+
+\subsection{Combining Ideals}
+Now we return to the case of general Lie algebras over arbitrary fields (no longer assuming $\ch k=0$). 
+We have seen that a subspace $I\le L$ is an ideal if $[IL]\subseteq I$.
+
+If $I,J\lhd L$, then $I+J$ is an ideal, as well as
+\[[IJ]=\{\sum_1^n [x_iy_i]|x_i\in I, y_i\in J\}.\]
+
+We also are sometimes interested in the direct sum of ideals $\oplus L_i$ where $[L_iL_j]=0$ when $i\ne j$.
+
+\subsection{Examples}
+$0$ and $L$ are ideals as well as the \textbf{center} $Z(L)$.
+
+\subsection{Characterizing Lie Algebras as Linear}
+We now want to realize many Lie algebras as (isomorphic to) linear ones. For now, note that the adjoint map $\ad$
+is a Lie algebra homomorphism by the Jacobi identity. Then $\ker\ad=Z(L)$, so when $Z=0$, then $L$ is isomorphic to 
+a linear Lie algebra.
+
+\begin{defn}
+	If $Z(L)$, and the map $\ad:L\to L$ is an isomorphism, we say $L$ \textbf{acts faithfully}
+	on itself (via $\ad$).
+\end{defn}
+
+\begin{defn}
+	If we have a homomorphism $L\to\gl(n,k)$, then we call $k^n$ an $L$-\textbf{module}.
+	We define the action of $L$ on $K^n$ through the homomorphism.
+\end{defn}
+
+Notice that $L$ is always an $L$ module over itself.
+
+One can check, using the realization of $\sl(n,k)$ we saw last time that (when $\ch k\nmid n$)
+that $\sl(n,k)$ is simple. This is actually the Lie algebra analogue of the fact that $M_n(K)$ is simple
+as a ring over any field $K$.
+
+\begin{rmk}
+	Notice that if $\ch k |n$$,$ then \textit{any} scalar matrix $aI\in\sl(n,k)$
+	so $Z(\sl(n,k))$ is nontrivial.
+\end{rmk}
+
+\brk
+
+In any event you can look up the proof for $\sl(2,k)$ in the book. The thing to glean here is that
+\[[xy]=h,\quad [hx]=2x,\quad [hy]=-2y\]
+for $h=e_{11}-e_{22}$, $x=e_{12}$ and $y=e_{21}$. This will keep popping up throughout the course.
+
+\subsection{Ideals of Linear Lie Algebras}
+Many subalgebras of $\sl(n,k)$ or $\gl(n,k)$ have many ideals (which happen to be ring-theoretic ideals).
+
+\begin{defn}
+	Some notation:
+	\[\t(n,k)=\{\text{upper-triangular matrices}\}\] 
+	also 
+	\[\n(n,k)=\{\text{strictly-upper-triangular matrices}\}\]
+\end{defn}
+
+\subsection{Derived Series}
+For any Lie algebra, we can define
+\begin{defn}
+	The \textbf{derived series} of a Lie algebra $L$ is defined to be
+	\[L^{(0)}=L,\quad L^{(1)}=[LL],\quad L^{(n+1)}=[L^{(n)}L^{(n)}]\]
+	which form an ascending chain of ideals.
+\end{defn}
+
+\begin{defn}
+	If $L^{(n)}$ for some $n$, and thus for all sufficiently large $n$, we call $L$ \textbf{solvable.}
+\end{defn}
+\begin{rmk}
+	Note that actually there is a cooresponding definition for groups using commutators $[G,G]$.
+	This was defined first and ported to Lie algebras. It actually happened in the reverse direction for:
+\end{rmk}
+
+\begin{defn}
+	The \textbf{central series} of $L$ is 
+	\[L^0=L,\quad L^1=[LL],\quad L^2=[L,L^1],\dots\]
+	and if $L^n=0$ for some $n$, then we call $L$ \textbf{nilpotent.}
+\end{defn}
+\begin{rmk}
+	Equivalently, we can say that $L$ is nilpotent if \textit{any $n$-fold bracket} in $L$ is zero.
+\end{rmk}
+\begin{rmk}
+	Notice that every nilpotent Lie group is solvable, but the converse fails: $t(n)$ is solvable by not nilpotent.
+\end{rmk}
+
+\subsection{Examples of solvable algebras}
+One can easily check that every ideal and homomorphic image of a solvable algebra is solvable. Therefore if $I$ and $J$ are solvable ideals of $L$, then 
+so is $(I+J)/J\cong I/(I\cap J)$ is solvable, as is $I+J$. Furthermore if $I$ and $L/I$ are solvable, then $L$ is.
+
+\begin{defn}
+	$\Rad L$, the \textbf{radical of $L$}, is the unique largest solvable ideal of $L$.
+\end{defn}
+\begin{rmk}
+	Any finite dimensional algebra has a radical (taken to be the sum of all solvable ideals in $L$).
+\end{rmk}
+
+Using this definition, we are able to give two different characterizations:
+\begin{defn}[Semisimple Lie Algebra]\label{def-semisimple}
+	$L$ is semisimple if $\Rad L=0$.
+\end{defn}
+\begin{defprime}{def-semisimple}[Semisimple Lie Algebra]
+	$L$ is semisimple if it is a direct sum of simple Lie algebras.
+\end{defprime}
+This equivalence will not become clear until we learn about the Killing form, but at least it jives
+with our experience with representation theory and semisimple modules.
 \end{document}

### 2019-01-18 Hopf Algebra Notes/Hopf Algebras.tex
Update Notes

@@
 
-\DeclareMathOperator{\Hom}{Hom}
+\newcommand*{\Hom}{\ensuremath\text{Hom}}
 
@@
 	we can see that the product $\ast$ satisfies the axioms required.
+
+	Note that actually
  \end{prf}
@@
 		\item $1_{\Hom}\ast f = m_A(u\otimes 1)(1\otimes f)(\varepsilon\otimes 1)\Delta.$
-		\item $f\ast 1_{\Hom}=m_A(1\otimes u)(f\otimes 1)(1\otimes \varepsilon)$
+		\item $f\ast 1_{\Hom}=m_A(1\otimes u)(f\otimes 1)(1\otimes \varepsilon)\Delta$
 		\item $(f\ast g)\ast h=m_A(m_A\otimes 1)((f\otimes g)\otimes h)(\Delta_C\otimes 1)\Delta_C$
@@
 
+\section{January 14, 2019}
+Today we are talking about Hopf modules and later in the week we will see the fundamental
+theorem of Hopf modules as well as a neat result.
+
+Here is the ``example of the day.''
+\begin{ex}
+	Consider $k[x]$ with maps $\Delta(x^n)=\sum_0^nx^i\otimes x^{n-i}$ and $\varepsilon(x^n)=\delta_{1,n}$. 
+	$S(x^n)=(-x)^n$. Then notice that $\Delta(x)=1\otimes x+x\otimes 1$, so $\delta(x^n)=(\delta x)^n$.
+
+	James doesn't want to do the rest of the computations, but they can be done. :)
+\end{ex}
+
+\begin{prob}\label{prob-4.1}
+	(**) Working with a Hopf algebra $V$, consider the convolution algebra $\Hom(V\otimes V,V)$
+	\begin{enumerate}
+		\item $(S\circ m)\ast m=1_{\Hom(V\otimes V,V)}=m\ast(m\circ s\otimes s\circ\tau)$ where $s$ is the antipode in $V$.
+		\item $(\Delta\circ S)\ast \Delta=1_{\Hom(V,V\otimes V)}=\Delta\ast((s\otimes s)\circ\tau\circ \Delta)$
+	\end{enumerate}
+\end{prob}
+\begin{sol}
+	\begin{align*}
+		(s\circ m)\ast m
+	\end{align*}
+\end{sol}
+\begin{prob}
+	Classify all Frobenius (to be defined) Hopf algebras of dimension 3.
+\end{prob}
+
+\subsection{Returning to (co)modules}\label{subsec-comod}
+\begin{lem}
+	Let $\scrT=\Vectk$, and $V$ and algebra over $k$. The following are equivalent:
+	\begin{enumerate}
+		\item $M$ is a left $V$-module (see last lecture).
+		\item There is an action of $A$ on $M$ such that $(ab)\cdot m=a\cdot (b\cdot m)$ and $1\cdot m=m$.
+		\item There is an algebra morphism $\varphi:V\to\Hom_k(M,M)$
+	\end{enumerate}
+\end{lem}
+
+\begin{defn}
+	Let $V$ be a coalgebra over $k$. We say $M$ is a left \textbf{comodule} if there is a map
+	\[\rho_M:M\to V\otimes M\]
+	such that the diagrams in figure \ref{fig-com} commute.
+\end{defn}
+\begin{figure}[h]\label{fig-com}
+	\begin{tikzcd}
+		M\ar[r,"\rho"]\ar[d,"\rho"] & V\otimes M\ar[d,"\id_V\otimes\rho"]\\
+		V\otimes M\ar[r,"\Delta_V\otimes\id_M"] &V\otimes V\otimes M
+	\end{tikzcd}
+	\begin{tikzcd}
+		M\ar[r,"\rho"]\ar[d,"\cong"] & V\otimes M\ar[dl, "\varepsilon\otimes\id_M"]\\
+		k\otimes M
+	\end{tikzcd}	
+	\caption{Diagrams for the definition of a comodule.}
+\end{figure}
+
+\begin{rmk}
+	Notationally speaking, we write $_V\mathcal{M}$ and $^V\mathcal{M}$ for the categories of 
+	left modules and comodules. Similar for right boyes.
+\end{rmk}
+
+\subsection{Tensors of $V$-modules}
+Say $V$ is a bialgebra (or Hopf if you prefer, but it's not necessary). Then we get the following:
+\begin{lem}
+	Let $M$ and $N$ be two right $V$-modules. Then
+	\begin{enumerate}
+		\item $M\otimes N$ is a right $V$-module.
+		\item If $V$ is cocommutative, then $M\otimes N\cong N\otimes M$ as right $V$-modules.
+		\item $\mathcal{M}_V$ is a tensor (monoidal) category.
+	\end{enumerate}
+\end{lem}
+\begin{prf}
+	For (a), use the fact that $V\otimes V$ acts on $M\otimes N$ in a natural way. Then you get
+	a map $V\otimes V\to \Hom_k(M\otimes N,M\otimes N)^{op}$ (why opposite?) and by precomposing
+	with $\Delta$ to show they are $V$-modules. This establishes (b).
+
+	Finally (c) follows because $\Vectk$ is a tensor category.
+\end{prf}
+
+\subsection{Hopf Modules}\label{subsec-hopfmod}
+\begin{defn}
+	Let $V$ be a bialgebra over $k$. We say that $M$ is a $\binom{r}{r}$ Hopf $V$-module if
+	\begin{enumerate}
+		\item $(M,m_M)$ is a right $V$-module
+		\item $(M,\rho_M)$ is a right $V$ comodule.
+		\item $\rho_M$ is a right $V$-module map.
+		\begin{itemize}
+			\item Equivalently, $m_M$ is a right $V$-comodule map.
+		\end{itemize}
+	\end{enumerate}
+\end{defn}
+\begin{rmk}
+	The fact that these two last conditions are equivalent are perhaps not immediately obvious but we
+	have been assured it is true. :)
+\end{rmk}
+
+\section{January 16, 2019}
+\begin{thm}[Larson-Sweedler `69]\label{thm-LS69}
+	Let $V$ be a Hopf algebra over $k$. Then the following categories are equivalent:
+	\[_V^V\mathcal{M}\cong \Vectk\cong \mathcal{M}_V^V\]
+\end{thm}
+
+\begin{rmk}
+	A natural question that one may ask is how to extend this theorem to the enriched setting.
+\end{rmk}
+
+\begin{lem}
+	$S$ is an anti-homomorphism of an algebra $V$ and an anti-homomorphism of the coalgebra $V$.
+\end{lem}
+\begin{prf}
+	Use the facts proved in problem~\ref{prob-4.1}.
+\end{prf}
+
+\begin{prob}
+	If $\ch k=0$ any Hopf quotient of $k[x]$ is either $k$ or $k[x]$ itself.
+\end{prob}
+
+\begin{prob}
+	Suppose $k$ has positive characteristic. Construct two different Hopf algebra quotients of $k[x]$ of dimension $p=\ch k$.
+\end{prob}
+
+\begin{ex}
+	$\ch k=p>0$. Let $V=k[x]/(x^p)$ as an algebra and define $\Delta(x)=x\otimes 1+1\otimes x$,
+	$\varepsilon(x)=0$ and $S(x)=x$. Then notice
+	\[\Delta(x^p)=\sum\binom{p}{i}x^i\otimes x^{p-i}=x^p\otimes 1+1\otimes x^p.\]
+\end{ex}
+
+\subsection{Hopf Modules Once More}
+Recall the definitions from section~\ref{subsec-comod} and section~\ref{subsec-hopfmod}.
+
+\begin{rmk}
+	An equivalent (more category-theoretical) definition of a Hopf module is a $V$-comodule object in 
+	the category of $V$-modules. Also can dualize everything.
+\end{rmk}
+
+\begin{ex}
+	Consider $V\in _V^V\mathcal{V}$. THen you can define $\rho=\Delta$ and check it satisfies all the requirements.
+\end{ex}
+\begin{defn}
+	An $\binom{l}{l}$ Hopf $V$-module $M$ is called \textbf{trivial} if $M\cong V\otimes M_0$ for some $M_0\in\Vectk.$
+\end{defn}
+\begin{rmk}
+	Basically you can think of this being trivial since we can \textit{always} define a module in this way
+	where the entire module structure is inherited from the structure of $V$ (that is, irrespective of $M_0$).
+\end{rmk}
+\begin{thm-prime}{thm-LS69}%[Fundamental Thm of Hopf Modules (L-S `69)]
+	Suppose that $V$ is a Hopf algebra. Then
+	\begin{enumerate}
+		\item Every $\binom{l}{l}$ Hopf $V$-module is trivial.
+		\item Let $M\in _V^V\mathcal{M}$. Then 
+		\[M\xrightarrow{\sim} V\otimes M^{Cov}\]
+		where
+		\[M^{Cov}:=\{m\in M|\rho(m)=1\otimes m\}.\]
+		\item $_V^V\mathcal{M}\cong \Vectk$.
+	\end{enumerate}
+\end{thm-prime}
+\begin{rmk}
+	This is actually just a reformulation of theorem~\ref{thm-LS69} in less compact (but ultimately more usable) notation.
+\end{rmk}
+\begin{prf}
+	Some identites that will be useful:
+	\begin{align*}
+		\Delta(v)=\sum v_1\otimes v_2\\
+		\sum S(v_1)\otimes v_2=\varepsilon(v)=\sum v_1S(v_2)\\
+		\rho(m)=\sum m_{-1}\otimes m_1\\
+		\sum v_1\otimes (v_2)_1\otimes (v_2)_2=\sum (v_1)_1\otimes (v_1)_2\otimes v_2=\sum v_1\otimes v_2\otimes v_3\\
+		\sum (m_{-1})_1\otimes (m_{-1})_2\otimes m_0=\sum m_{-1}\otimes (m_0)_{-1}\otimes (m_0)_0=\sum m_{-2}\otimes m_{-1}\otimes m_0
+	\end{align*}
+
+	Then we defin $\phi:M\to M^{Cov}$ by 
+	\[\phi(x)=\sum S(x_{-1})x_0\in M\]
+	We claim first that $\phi(x)\in M^{Cov}$ -- namely $\rho(\phi(x))=1\otimes\phi(x)$. To see this, compute
+	\begin{align*}
+		\rho(\phi(x)) &= \sum (\phi(x))_{-1}\otimes (\phi(x))_0\\
+		&= (S(x_{-1})x_0)_{-1}\otimes (S(x_{-1})x_0)\\
+		&= \sum\Delta(S(x_{-1}))\rho(x_0)\\
+		&= \sum (S\otimes S)\circ \tau(\Delta(x_{-2}))\cdot [x_{-1}\otimes x_0]\\
+		&=\sum[S(x_{-2})\otimes S(x_{-3})][x_{-1}\otimes x_0]\\
+		&= \text{my fingers are aching...}\\
+		&= 1\otimes\phi(x)
+	\end{align*}
+
+	Now define $F:M\to V\otimes M^{Cov}$ by $F=(\id\otimes\phi)\circ\rho$ and $G:V\otimes M^{Cov}$ 
+	to be the map taking $v\otimes m$ to $vm$. The next claim is that $GF$ and $FG$ are the identity.
+	You can see this by similarly pushing around notation.
+
+	Finally the last claim will be seen on Friday.
+\end{prf}
 \end{document}

### 2019-01-18 Lie Algebra Notes/Lie Algebras.tex
Update Notes

@@
 \begin{sol}
+	Let $I=\ad L$. $I\subseteq \Der L$ since $\ad x$ is a derivation. This is a subspace since $\ad 0=0$ and since the bracket is bilinear
+	\[\ad x(ay+bz)=[x,ay+bz]=[x,ay]+[x,bz]=a[xy]+b[xz]=a\ad x(y)+b\ad x(z).\]
 
+	Take any $x\in I$ and let $f\in\Der L$ be arbitrary. But then for $y,z\in L$
+	\begin{align*}
+		[\ad x,f](yz)&=(\ad x f-f\ad x)(yz)\\
+		&=[x,f(yz)]-f[x,yz]\\
+		&=[x,f(y)z+yf(z)]-f([x,y]z+y[x,z])\\
+		&=[x,f(y)z]+[x,yf(z)]-f([x,y]z)-f(y[x,z])\\
+		&=[x,f(y)]z+f(y)[x,z]+[xy]f(z)+y[x,f(z)]-f([x,y])z\\
+		&\hspace{2in} -[x,y]f(z)-f(y)[x,z]-yf[x,z]\\
+		&=([x,f(y)]-f[x,y])z+y([x,f(z)]-f[x,z])\\
+		&=([\ad x, f](y))z+y([\ad x,f](z))
+	\end{align*}
+	so $[IL]\subseteq I$ whence $I$ is an ideal of $L$.
 \end{sol}
@@
 \begin{sol}
-
+	Assume that $L$ is nilpotent. Then defining $L^0=L$, and by induction $L^i=[LL^{i-1}]$, this 
+	means that $L^n=0$.
 \end{sol}
@@
 \begin{defn}
-	If $L^{(n)}$ for some $n$, and thus for all sufficiently large $n$, we call $L$ \textbf{solvable.}
+	If $L^{(n)}=0$ for some $n$, and thus for all sufficiently large $n$, we call $L$ \textbf{solvable.}
 \end{defn}
@@
 with our experience with representation theory and semisimple modules.
+
+\section{January 14, 2019}
+I missed this day for an event. I will try to catch up if I have time.
+
+\section{January 16, 2019}
+We are continuing from last time when we were showing that as solvable Lie subalgebra of $\gl(n,k)$
+(when $\ch k=0$ and $\bar k=k$) necessarily admits a vector $v\in k^n\setminus 0$ that is
+a common eigenvector:
+\[xv=\lambda(x)v\]
+for some \textbf{linear} function $\lambda:L\to k$.
+
+We argues (as in the parllel case of a Lie subalgebra of $\gl(n,k)$ consisting of nilpotent matrices
+by induction on dimension and found the same result (except it is zero). In this case when $\dim L=1$,
+since $k$ is algebraically closed, $L$ is spanned by asingle matrix $x$ and any eigenvector of $x$ does the trick.
+
+Now assume that the result holds for solvable Lie algebras of dimension $<d$. Then since $L$ is sovable, we know
+$[LL]\subsetneq L$. Let $I$ be any subspace containing $[LL]$ and properly contained in $L$ of codimension one.
+Then $I$ is an ideal since $[LI]\subseteq[LL]\subseteq I$, so by the inductive hypothesis $I$ has a common eigenvector, 
+so that there is a linear function $\lambda$ such that the \textbf{weight space}
+\[V_\lambda=\{v\in k^n:xv=\lambda(x)v,\,\forall x\in I\}\]
+is nonzero.
+
+We know that $L=I+kx$ for some $x\in L$, so we must show that $x$ preserves $V_\lambda$: then any eigenvector
+of $x$ on $V_\lambda$ will do the job. Given $v\in V_\lambda$, we must show that 
+\[yxv=xyv+[yx]v=\lambda(y)xv+\lambda([yx])v\]
+so we must show that $\lambda([xy])=0$.
+
+Given $v\in V_\lambda$, look at hte powers $v,xv, x^2v,\dots$. Let $m$ be the least power such that the
+$v,xv,\dots,x^mv$ are linearly independent and let $W_i$ be the span of the first $i$ of these.
+
+Then one checks by induction for an $y\in I$ that $yx^iv=\lambda(y)x^iv\pmod{v,xv,\dots,x^iv}$
+It follows that $y$ was in the subspace $W_m$ (oh no missed some stuff).
+
+So the trave of $y$ on $W_m$ is $m\lambda(y)$. Now for any $y\in I$, the matrix $[xy]$ acts on $W_m$ ascendingthe commutator of two matrices having trace zero, whence $m\lambda([xy])=0$,
+so since $\ch k=0$, $\lambda([xy])=0$, as desired.
+
+\brk
+
+As a consequence, given a solvable Lie subalgebra of $\gl(n,k)$ here is a chain of subspaces
+\[0=V_0\le V_1\le\cdots\le V_n=k^n\]
+or a \textbf{flag} such that $LV_i\subseteq V_i$ for all $i$: $L$ acts by upper triangular 
+matrices with respect to a suitable basis.
+
+Another way of putting this, using module language, is that given any \textbf{finite dimensional} module over
+a solvable Lie algebra (over an algebraically closed characteristic zero field) has a one-dimensional submodules $N$.
+
+Equivalently, the only finite-dimensional irreducible modules over a solvable Lie algebra are one-dimensional. We 
+will use this later in the context of semisimple Lie algebras with large solvable subalgebra.
+
+\begin{rmk}
+	For the record, Lie's theorem \textit{definitely fails} when $k$ has positive characteristic.
+	We see one in our homework, and another example comes from $\sl(2,k)$ where $\ch k=2.$
+
+	We know that $\sl(2,k)$ has a basis: $(\begin{smallmatrix}
+		1&0\\0&-1
+	\end{smallmatrix})=I_2$, $x=e_{12}$, $y=e_{21}$. Here $[hx]=0=[hy]$ and $[xy]=h$. It is shown in the book
+	and is easy to see that when $\ch k\ne 2$, $\sl(2,k)$ is simple, but here $\sl(2,k)$ is nilpotent and its
+	bracket relations become $[xy]=z$, $[xz]=[yz]=0$.
+
+	It ends up we can define (actually over any field) the \textbf{Heisenberg} Lie algebra, spanned
+	by $x=e_{12},y=e_{23},z=e_{13}$ or more simply the $3\times 3$ strictly-upper-triangular matrices.
+
+	The matrices in $\sl(2,k)$ for any $k$ have no common eigenvector, so our proof of Lie's theorem must
+	fail somewhere? We indeed have an ideal $I$ of coimension one spanned by $h$ and $x$ and there is acommon eigenvector for $I$, 
+	namely $e_1$. Then $I$ acts diagonally on $ke_1$ via the linear function sending $h$ to 1, $x$ to $1$. 
+	Here $y$ plays the role of $x$ in the proof, but is outside the ideal. That is,
+	\[\lambda[yx]=\lambda(h)\ne 0.\]
+\end{rmk}
+
+Returning to the good case of $k$, we can now apply our result on a linear solvable Lie algebras to arbitrary ones:
+uGIven any solvable Lie algebra $L$, there is a flag $0=L_0\le L_1\le\cdots\le L_n=L$ of ideas of $L$
+so that $[LL_i]\subseteq L_i$ Also $[LL]$ is nilpotent, since it acts on $L$ by strictly upper triangular matrices.
+
+Both results fail for general solvable Lie algebras when the basefield is not nice. If a solvable Lie algebra $L$ 
+\textbf{does} have a chain of ideals $0=L_0\le L_1\le\cdots\le L_n=L$ where $\dim L_i=i$, then we call $L$ \textbf{completely solvable.}
+
+\subsection{Returning to Linear Algebra}
+At this point we will take some time to return to an important idea from Linear algebra, namely the \textbf{Jordan canonical form}
+of a square matrix. Recall Jordan blocks and similarity when eigenvalues lie in the field.
+
+If $M$ is already in this form, then we can write $M=M_s+M_n$ where $M_s$ consists of the diagonal entries of each
+block (zeroes elsewhere) and $M_n$ the off-diagonal entries. Then $M_s$ is diagonal (or for general $M$
+diagonalizable), or \textbf{semisimple} and $M_n$ is nilpotent and $[M_sM_n]=0$ (they commute).
+
+\begin{defn}
+	$M=M_s+M_n$ is called the \textbf{Jordan decomposition of $M$.} We call $M_s$ and $M_n$ the \textbf{semisimple} 
+	and \textbf{nilpotent} parts of $M$.
+\end{defn}
+\begin{rmk}
+	Each are uniquely determined for $M$ and (and this is not generally proved in the first year 
+	algebra sequence) each are equal to polynomials in $M$ with no constant term.x`
+\end{rmk}
+
 \end{document}

### 2019-01-18 Reading Notes/reading_notes.tex
Update Notes

@@
 
+Consider the representations of A over a field $k$ (or more generally $A$ modules. W can make this a tensor catecory)
+Recall that in this category we have enough projectives and (with finitely generated hopf algebras) we get that all projectives are injective.
+
+Then we end up defining $\Ext_A^*(k,k)=H^*_A(A,k)$, the cohomology ring of $A$ over $k$.
+
+The conjecture here is that when $A$ is finite dimensional that $H^*(A,k)$ is finitely generated as a $k$-algebra.
+
+This is known in some special cases:
+\begin{itemize}
+	\item Finite groups
+	\item Finite group schemes (in positive characteristic) due to Friedlander and Suslin.
+	\item In characteristic zero we have
+	\begin{itemize}
+		\item Quantum groups
+		\item Hopf algebras that come from Nichols algebras
+	\end{itemize}
+\end{itemize}
+
+Papers to look at: Ginzburg and Kumar '94 Something like ``Cohomology of small quantum groups''. There is also a '10 paper on Julia's website with Sarah Witherspoon and two others on Nichols algebras.
+
+Sarah Witherspoon has a book she is writing on Hochschild cohomology that could be very good. It's on her website (Texas A\& M). Check out chapter 9 and appendix A.
+
 \end{document}

### 2019-02-13 Hopf Algebra Notes/Hopf Algebras.tex
I should push more frequently

@@
 \DeclareMathOperator{\Ext}{Ext}
+\DeclareMathOperator{\gldim}{gldim}
+\DeclareMathOperator{\projdim}{projdim}
+\DeclareMathOperator{\findim}{findim}
 
@@
 \newcommand*{\Coalgk}{\operatorname{Coalg}_k}
+\newcommand*{\calM}{\mathcal{M}}
 \DeclareMathOperator{\Coalg}{Coalg}
@@
 \newcommand*{\scrT}{\ensuremath\mathscr{T}}
+\newcommand*{\Ch}{\mathbf{Ch}}
 
@@
 \begin{ex}
-	Consider $V\in _V^V\mathcal{V}$. THen you can define $\rho=\Delta$ and check it satisfies all the requirements.
+	Consider $V\in _V^V\mathcal{M}$. THen you can define $\rho=\Delta$ and check it satisfies all the requirements.
 \end{ex}
@@
 \end{prf}
+\section{January 18, 2019}
+Recall the Fundamental Theorem of Hopf Modules: $_V^V\mathcal{M}\simeq \Vectk$ or equivalently that
+every (not necessarily finite dimensional) Hopf module is trivial: $M\cong V\otimes M^{Cov}$.
+
+\section{Frobenius Algebras}
+Today we will discuss the result that
+\begin{thm}[Larson-Sweedler, `69]\label{LS-69-2}
+	Every finite dimensional Hopf algebra is Frobenius.
+\end{thm}
+
+There are some nice categorical equivalences: finite dimensional Frobenius algebras over $k$ is
+equivalent to 2-D quantum field theories, or equivalently the symmetric functor category from the 2-D cobordism category to $\Vectk$.
+
+Whoa.
+
+\begin{ex}
+	$V=(kG)^*=\oplus_{g\in G}k\delta g$ where $G$ is a finite group. Define the algebra structure with
+	\[1_V=\sum_{g\in G}\delta_g\]
+	and $\delta_g\delta_h=\delta_g$ when $g=h$ and 0 otherwise.
+
+	The coalgebra structure is $\varepsilon(\delta_g)=\delta_{g,1_G}$ where on the right it is the Kroneker delta.
+	Furthermore $\Delta(\delta_g)=\sum_{h\in G}\delta_h\otimes\delta_{h^{-1}g}$ and $S(\delta_g)=\delta_{g^{-1}}$.
+\end{ex}
+
+\begin{prob}
+	Prove that $kS_3\cong(kS_3)^*$.
+\end{prob}
+\begin{prob}
+	Prove that $k\Z_3\cong(k\Z_3)^*$ when $\ch k\ne 3$.
+\end{prob}
+
+\begin{lem}
+	Let $A$ and $B$ be algebras over $k$.
+	\begin{enumerate}
+		\item If $M\in_A\mathcal{M}_B$, $N\in \mathcal{M}_B$, then $\Hom_{\mathcal{M}_B}(_AM,N)\in\mathcal{M}_A$.
+		\item Take $B=k$, $M\in _A\mathcal{M}$. Then
+		\[M^*:=\Hom_k(M,k)\in\mathcal{M}_A.\]
+		\item $(-)^*:_A\mathcal{M}_{f.d.}\to (\mathcal{M}_A)_{f.d.}^{op}$ is an equivalence of categories. This is called the \textbf{reflection principle.}
+	\end{enumerate}
+\end{lem}
+\begin{prf}
+	This is a known fact. (b) follows quickly from (a) and (c) involves extending the dual map to maps of left $A$-modules.
+\end{prf}
+By the principle of duality, we have:
+\begin{lem}
+	If $C$ and $d$ are finite dimensional coalgebras over $k$, then
+	\begin{enumerate}
+		\item if $M\in ^C\mathcal{M}^D_{f.d.}$ and $N\in \mathcal{M}^D_{f.d.}$ then $\Hom_{\mathcal{M}^D}(M,N)\in\mathcal{M}^C_{f.d.}$.
+		\item If $M\in ^C\mathcal{M}_{f.d.}$ then $\Hom_k(M,k)\in\mathcal{M}^C$
+		\item $(-)^*:^C\mathcal{M}_{f.d.}\to(\mathcal{M}_{f.d.}^C)^{op}$ is an equivalence of categories.
+	\end{enumerate}
+\end{lem}
+
+\begin{defn}
+	A finite dimensional algebra $A$ is called \textbf{Frobenius} if one of the following (equivalent)
+	conditions hold:
+	\begin{enumerate}
+		\item $(A)^*_A\cong A_A$
+		\item $_A(A)^*\cong _AA$
+	\end{enumerate}
+\end{defn}
+\begin{rmk}
+	Note that the equivalence of these two things follow due to the equivalence of categories we get
+	from the two lemmas above.
+\end{rmk}
+We actually don't need the following lemma, but James is a fan so we're going to write it down.
+\begin{lem}\label{lem-not-needed}
+	Let $A$ be a finite dimensional algebra over $k$. Then $M$ is a finite dimensional left $A$-module if
+	and only if $M$ is a finite dimensional right $A^*$-comodule.
+\end{lem}
+\begin{rmk}
+	The reason we get the module to comodule switch is (partly?) due to the fact that in our
+	monoidal category $(V\otimes W)^*=W^*\otimes V^*.$
+\end{rmk}
+
+\begin{lem}
+	Let $V$ be a finite dimensional Hopf algebra and $M$ be a finite dimensional $\binom{l}{l}$ Hopf $V$-module.
+	Then $M^*$ is a finite dimensional $\binom{r}{r}$ Hopf $V$-module.
+\end{lem}
+\begin{prf}
+	$M\in {_V^V\mathcal{M}}$ if and only if $M\in_V\mathcal{M}\cap {^V\mathcal{M}}$ and the two structures are compatible.
+
+	The fact that $M^*\in \mathcal{M}_V\cap\mathcal{M}^V$ is straightforward and checking the compatibility is easy to check.
+\end{prf}
+
+Now we finally prove that theorem~\ref{LS-69-2}:
+\begin{prf}
+	First of all $V\in{^V_V\mathcal{M}}$ so by the lemma $V^*\in\mathcal{M}_V^V$. By the fundamental theorem of Hopf modules,
+	\[V^*\cong V\otimes(V^*)^{Cov}\cong V\]
+	as Hopf modules. Thus $V_V^*\cong V_V$ in particular.
+\end{prf}
+\begin{rmk}
+	The second isomorphism above comes from the fact that $(V^*)^{Cov}$ is one-dimensional, so is the trivial module. :)
+\end{rmk}
+\begin{cor}
+	If $A$ is \textit{not} a Forbenius algebra, then there is \textit{no} Hopf algebra structure on $A$.
+\end{cor}
+\begin{ex}
+	Some examples of algebras with(out) this property.
+	\begin{enumerate}
+		\item $\C[x]/(x^n)$ is finite dimensional and Frobenius, but if $n\ge 2$, there is \textbf{no Hopf structure.}
+		\item $\C[x,y]/(x^2,y^2,xy)=\C1+\C x+\C y$. This has the property that it is local (unique maximal submodule over itself)
+		but $A^*$ has two maximal submodules, so $A\not\cong A^*$, so not Frobenius.
+	\end{enumerate}
+\end{ex}
+Finally we conclude with a proof of lemma~\ref{lem-not-needed}:
+\begin{prf}
+	Say that $\{a_i\}$ is a basis for $A$ and $\{a_i^*\}$ is a basis for $A^*$ ($a_i^*(a_j)=\delta_{ij}$).
+
+	Define $\rho(m)=\sum(a_i\cdot m)\otimes a_i^*$. A lemma (use change of basis matrices) shows that
+	$\sum a_i\otimes a_i^*$ is independent of choice of basis. Another lemma says (scalar) multiplication is associative
+	if and only if $\rho$ is coassociative. This gets us the forward direction.
+
+	For the reverse direction, say $\rho(m)=\sum m_0\otimes m_{-1}.$ then define $a\cdot m=\sum m_1(a)m_0$.
+\end{prf}
+
+\section{January 23, 2019}
+Today we are interested in studying the representation category $_V\mathcal{M}$ for any Hopf algebra $V$. It will end up
+(and this is a bit cryptic for now) that $k$ controls everything here.
+\section{Represetnations and Modules}
+Notice that for any Hopf algebra $V$ we have the exact sequence:
+\[0\to \ker\varepsilon\to V\xrightarrow{\varepsilon} k\to 0\]
+\begin{defn}
+	The trivial $V$ modules is $V/\ker\varepsilon=k\in _V\mathcal{M}_V$.
+\end{defn}
+\begin{ex}
+	$V=kG$ where $G$ is a group, $\varepsilon(g)=1$ for all $g\in G$. Then $\ker\varepsilon=\oplus k(g-1)$ and so
+	\[k=V/\oplus k(g-1).\]
+\end{ex}
+\begin{ex}
+	$V=k[x]$ where $\Delta(x)=1\otimes x+x\otimes 1$ and $\varepsilon(x)=0$. Then $\ker\varepsilon=\oplus_{n\ge 1}kx^n$, whence
+	\[V/\ker\varepsilon=k[x]/(x).\]
+\end{ex}
+\begin{ex}
+	Let $V=(kG)^*=\oplus_G k\delta_g$ and $\varepsilon(\delta_g)=\delta_{1,1_G}$. THen $\ker\varepsilon=\oplus_{g\ne 1_G}k\delta_g$ and
+	\[V/\ker\varepsilon=k\delta_1\]
+\end{ex}
+The following uses $T_4$, the example of the day I missed due to my meeting.
+\begin{ex}
+	$T_4$: $\varepsilon(g)=1$, $\varepsilon(p)=0$. Then $\ker\varepsilon=k(g-1)\oplus kp\oplus kpg$ and
+	\[V/\ker\varepsilon=k.\]
+\end{ex}
+\begin{lem}
+	\begin{itemize}
+		\item For all $x \in V$ and $a\in k$ (the trivial module), $xa=\varepsilon(x)a$.
+		\item $k$ is the identity object in $_V\mathcal{M}$.
+		\item If $V$ is cocommutative, then $_V\mathcal{M}$ is a symmetric tensor category.
+	\end{itemize}
+\end{lem}
+\begin{prf}
+	\[x\cdot a=(x-\varepsilon(x))a+\varepsilon(x)a=\varepsilon(x)a\]
+	since $x-\varepsilon(x)\in\ker\varepsilon$.
+
+	For the next part, consider that for every $M\in _V\mathcal{M}$, we have 
+	\[M\otimes k\cong M\cong k\otimes M.\]
+	But then for all $x\in V$, we have
+	\begin{align*}
+		\varphi[x\dot(m\otimes 1)]&=\varphi(x_1\cdot m\otimes x_2\cdot 1)\\
+		&=\varphi(x_1\cdot m\otimes \varepsilon(x_2)\cdot 1)\\
+		&=\varphi((x_1\varepsilon(x_2))m\otimes 1)\\
+		&=\varphi(xm\otimes 1)=xm=x\varphi(m\otimes 1).
+	\end{align*}
+\end{prf}
+\subsection{Integrals}
+\begin{defn}
+	An element $x\in V$ is called a \textbf{left integral} if $vx=\varepsilon(v)x$ for all $v\in V$.
+	Similar for \textbf{right integral.}
+\end{defn}
+\begin{lem}
+	The following are equivalent:
+	\begin{itemize}
+		\item $x\in V$ is a left integral.
+		\item $kx\cong$ the trivial module.
+		\item $1\mapsto x$ defines a left $V$-module morphism $k\to kx$.
+	\end{itemize}
+\end{lem}
+\begin{lem}
+	Let $\int_V^l$ denote the set of left integrals in $V$.
+	\begin{itemize}
+		\item $\int_V^l$ forms a vector subspace of $V$.
+		\item $\int_V^l$ forms a left $V$ submodule of $V$. Thus it forms a left ideal of $V$.
+		\item $\int_V^l$ forms a right ideal of $V$.
+		\item $\int_V^l\cong\Hom_V(k,V)$ (left $V$-mods)
+	\end{itemize}
+\end{lem}
+\begin{prf}
+	The first three are relatively obvious. The last one comes from the map $\Phi:x\in\int_V^l\mapsto f_x:1\to x.$
+\end{prf}
+\begin{rmk}
+	Unfortunately it ends up that the collection above ends up usually being zero. 
+\end{rmk}
+\begin{thm}
+	Let $V$ be a finite dimensional Hopf algebra. Then $\dim\int_V^l=1$.
+\end{thm}
+\begin{prf}
+	\begin{align*}
+		\dim\int_V^l&= \dim\Hom_V(_Vk,_VV)\\
+		&=\dim\Hom_V((V^*)_V,k_V^*)\\
+		&=\dim\Hom_V(V_V,k)=\dim k=1
+	\end{align*}
+	Where we used above that $V$ was finite dimensional whence Frobenius to get $V^*\cong V$.
+\end{prf}
+\begin{ex}
+	$V=kG$ for $G$ a finite group. Then $\int_V^l=\sum_G g$ and
+	\[v\cdot(\sum_G g)=\varepsilon(v)(\sum_G g)=\int_V^r\]
+	and when $v=h\in G$, we can compute it.
+\end{ex}
+\begin{ex}
+	When $V=k[x]$, which is infinite dimensional, we get no integrals. This is because it is an 
+	integral domain, so if
+	\[x\cdot\int_V^l=\varepsilon(x)\int_V^l=0\cdot\int_V^l=0\]
+	then this forces $\int_V^l=0$.
+\end{ex}
+\begin{ex}
+	When $V=(kG)^*$ for $G$ a finite group, then 
+	\[\int_V^l=\int_V^r=\delta_{1_G}\]
+\end{ex}
+\begin{ex}
+	When $V=T_4=k1\oplus kg\oplus kp\oplus kgp$,
+	\[\int_V^l=(g+1)p\ne\int_V^r=p(1+g)=(1-g)p\]
+\end{ex}
+\begin{ex}
+	When $\ch k=p$ and $V=k[x]/(x^p)$, then 
+	\[\int_V^l=\int_V^r=x^{p-1}.\]
+\end{ex}
+
+\section{January 28, 2019}
+Today we will be mostly doing things in homological algebra. :)
+\begin{ex}
+	An example of an infinite dimension Hopf algebra is of the following: fix $q\in k^*$. Then 
+	\[T_\infty=k<g, g^{-1}, p>/\langle gg^{-1}=g^{-1}g=1, gp=qpg\rangle\]
+	and we can find $\{g^ip^j:i\in\Z, j\in\N\}$ is a $k$-linear basis.
+
+	For the coalgebra structure, define $\Delta(g)=g\otimes g$ and $\varepsilon(g)=1$ and 
+	$\Delta(p)=p\otimes 1+g\otimes p$ and $\varepsilon(p)=0$. Finally $S(g)=g^{-1}$ and $S(p)=q^{-1}pg^{-1}$.
+\end{ex}
+
+\subsection{Some nice results in Homological algebra}
+\begin{prob}
+	Prove that $\Q$ is an injective $\Z$-module.
+\end{prob}
+\begin{thm}[Auslander-Buchsbaum `59]
+	Every local commutative algebra with finite global dimension is a UFD.
+\end{thm}
+Another result is about algebraic groups:
+\begin{thm}
+	If $\ch k=0$, then every Noetherian commutative Hopf algebra has finite global dimension
+	and is a direct sum of integral domains, each of which is isomorphic.
+\end{thm}
+\begin{defn}[Semiprime]
+	An ideal $I\lhd R$ is called \textbf{semiprime} if it is the intersection of (possibly infinitely many) primes.
+\end{defn}
+\begin{conj}
+	Every Noetherian Hopf algebra of finite global dimension is semiprime (that is, 0 is semiprime in $V$).
+\end{conj}
+
+\begin{defn}[Connected Graded]
+	An algebra $A$ is called \textbf{connected graded} if 
+	\[A=k\oplus A_1\oplus A_2\oplus \cdots\]
+	and $1_A=1_k\in k$, $A_iA_j\subseteq A_{i+j}$.
+
+	The new part (connected) refers to the fact that $k$ is a summand.
+\end{defn}
+\begin{thm}
+	Suppose that $\ch k=0$. Let $V$ be a Noetherian Hopf algebra that is connected graded as an algebra (says nothing about the coalgebra structure).
+	Then the following hold:
+	\begin{itemize}
+		\item $V$ has finite global dimension.
+		\item $V$ is a domain.
+		\item $V$ is Artin-Schelter regular.
+		\item $V$ is Auslander regular and Cohen-Macaulay.
+		\item $V$ is Calabi-Yau
+		\item $V$ is an iterated Ore extension.
+	\end{itemize}
+\end{thm}
+That's a lot of word salad.
+
+\subsection{Projective Modules}
+\begin{defn}[Module-Theoretic Definitions]
+	Let $M$ be a (left) $A$ module. 
+	\begin{enumerate}
+		\item $M$ is \textbf{free} if $M\cong \oplus_\mathcal{I} A$
+		\item $M$ is \textbf{projective} if it is a direct summand of a free module.
+	\end{enumerate}
+\end{defn}
+\begin{lem}
+	$p\in{_A\mathscr{M}}$ is projective if and only if the diagram in figure~\ref{fig-proj2} commutes.
+\end{lem}
+\begin{figure}[h]
+	\centering
+	\begin{tikzcd}
+		& P\ar[d,"f"]\ar[dl,dashed,"F",swap] &\\
+		M\ar[r,"\pi"] & N\ar[r] & 0
+	\end{tikzcd}
+	\caption{The definition of a projective object.}
+	\label{fig-proj2}
+\end{figure}
+\begin{rmk}
+	This actually is just the definition for any Abelian category. This may generalize even further, but this is enough for us. :)
+\end{rmk}
+
+\begin{defn}[Injective Module]
+	An \textbf{injective} $V$-module is a projective $V^{op}$ module.
+\end{defn}
+
+\section{January 30, 2019}
+Today we're talking more about homological algebra. In particular, we'll learn about (or see again)
+\begin{itemize}
+	\item Complexes
+	\item Projective resolutions and dimension
+\end{itemize}
+First the example of the day:
+\begin{ex}
+	Consider $GL_2(k)$, a group -- in fact, an algebraic group! Then 
+	\[\mathcal{O}(GL_2)=k[x_{11},x_{12},x_{21},x_{22},det^{-1}]/\langle det^{-1}(x_{11}x_{22}-x_{21}x_{12}-1\rangle\]
+	Let $X=(x_{ij})\in GL_2(k)$.
+
+	Define the coalgebra structure via $\Delta(X)=X\otimes X$ and $\Delta(x_{ij})=\sum_1^2 x_{is}\otimes x_{sj}$,
+	$\varepsilon(X)=I_2$, $\varepsilon(x_{ij})=\delta_{ij}$. Some computation (this is not trivial!) give us
+	\[\Delta(det)=det\otimes det\quad\Rightarrow\quad \Delta(det^{-1})=det^{-1}\otimes det^{-1}\]
+	Then we can find that $S(X)=X^{-1}$ and $S(det^{\pm})=det^{\mp}$.
+\end{ex}
+\begin{prob}
+	Construct $\mathcal{O}(GL_n)$.
+\end{prob}
+\begin{prob}
+	Construct $\mathcal{O}(G)$.
+\end{prob}
+\subsection{Homological stuff}
+\begin{defn}[Complex]
+	A \textbf{complex} of $A$-modules is a sequence of $A$-modules connected by homomorphisms
+	\[\cdots\to M_2\xrightarrow{d_2}M_1\xrightarrow{d_1} M_0\to \cdots\]
+	where $d_i\circ d_{i+1}=0$ for all $i$.
+\end{defn}
+Then we went over the standard definitions for:
+\begin{itemize}
+	\item Homology
+	\item (Short) Exact sequences
+	\item Projective resolutions
+	\item Projective dimension
+\end{itemize}
+\begin{defn}[Global (Homological) Dimension]
+	The (left) \textbf{global dimension} is
+	\[\gldim A=\max\{\projdim M| M\in {_A\mathcal{M}}\}\]
+\end{defn}
+\begin{thm}[Hilbert Syzygy Theorem]
+	\[\gldim k[x_1,\dots,x_n]=n\]
+\end{thm}
+\begin{cor}
+	\[\gldim k= 0\]
+\end{cor}
+\begin{defn}
+	The \textbf{finitistic global dimension} of $A$ is defined to be
+	\[\findim A=\max\{\projdim M| M\in{_A\mathcal{M}}, M\text{ is f.g.},\projdim M <\infty\}\]
+\end{defn}
+\begin{conj}[Finitistic Dimension Conjecture]
+	Let $A$ be a finitely generated algebra. Then $\findim A<\infty$.
+\end{conj}
+\section{February 1, 2019}
+A really cool theorem:
+\begin{thm}[Lorenz-Lorenz `95]
+	Let $V$ be a Hopf algebra. Then $\gldim V=\projdim {_Vk}$.
+\end{thm}
+\begin{prf}
+	It is enough to show that $\gldim V\le\projdim{_Vk}$. Or equivalently that
+	\[\projdim M\le \projdim k,\forall M\in{_V\calM}\]
+	but then
+	\[\projdim(M)=\projdim(k\otimes M)\le \projdim(k)\]
+	by lemma~\ref{lem-tensorproj}
+\end{prf}
+\begin{ex}
+	Today's example: Quantum group $V=\mathcal{O}_q(GL_2)$. Fix some $q\in k^*$. Then as an algebra, 
+	\[V=k\langle x_{11},x_{12},x_{21},x_{22},det_q^{-1}\rangle/(\text{relations})\]
+	where the relations are given by 
+	\begin{align*}
+		x_{ij}x_{kl}=(\delta_{ik}+\delta_{jl})x_{kl}x_{ij}\quad\text{except when } i=k \text{ and } j=l\\
+		x_{12}x_{21}=x_{21}x_{12}\\
+		x_{22}x_{11}-x_{11}x_{22}=(q-q^{-1})X_{12}x_{21}\\
+		det_q^{-1}det_q=det_q det_q^{-1}=1
+	\end{align*}
+	where $det_q=x_{22}x_{11}-q^{-1}x_{12}x_{21}x_{11}x_{22}-qx_{12}x_{21}$. One can show $det_q$ is central in $V$.
+
+	Now for the coalgebra: $\delta(X)=X\otimes X$, so $\Delta(x_{ij})=\sum_{s=1}^2 x_{is}\otimes x_{sj}$, $\varepsilon(X)=I_2$ and
+	\[S(X)=X^{-1}=\begin{pmatrix}
+		x_{22}det_q^{-1} & -qx_{12}det_q^{-1}\\
+		-q^{-1}x_{21}det_q^{-1} & x_{11}det_q^{-1}
+	\end{pmatrix}\]
+\end{ex}
+\begin{prob}
+	Show that $\gldim(\begin{smallmatrix}\Z & \Q\\0 & \Q\end{smallmatrix})=2$ while $\gldim(\begin{smallmatrix}\Z & \Q\\0 & \Q\end{smallmatrix})^{op}=1$
+\end{prob}
+
+\subsection{Returning to Homological Algebra}
+Let $A$ be an algebra, $M\in {_A\mathcal{M}}$. Recall the definitions of projective and global dimension.
+\begin{rmk}
+	The following are equivalent:
+	\begin{enumerate}
+		\item $\gldim A=0$
+		\item $\gldim A^{op}=0$
+		\item $A$ is semisimple Artininan
+		\item $A=\oplus_1^b M_{n_i}(D_i)$ by Artin-Wedderburn where $D_i$ is a division ring.
+	\end{enumerate}
+\end{rmk}
+\begin{rmk}
+	\begin{enumerate}
+		\item A PID has global dimension 1. 
+		\item The free algebra $k\langle x_1,\dots,x_n\rangle$ has global dimension 1.
+		\item The path algebra of a finite quiver has global dimension 1.
+		\item $\mathcal{O}_q(GL_2)$ has global dimension 4.
+	\end{enumerate}
+\end{rmk}
+
+From now on, let $V$ always stand for some Hopf algebra.
+\begin{lem}
+	${_V\mathcal{M}}$ is a tensor category. The action on $M\otimes N$ is the following:
+	\[v\cdot (m\otimes n)=\sum (v_{(1)}\cdot m)\otimes (v_{(2)}\cdot n)\]
+\end{lem}
+\begin{lem}
+	If $M\in {_V\mathcal{M}}$, then $V\otimes M\in {_V^V\mathcal{M}}$. As a corollary,
+	\[V\otimes M\cong V\otimes (V\otimes M)^{cov}\]
+	and as a consequence $V\otimes M$ is a free $V$-module.
+\end{lem}
+
+\begin{thm}
+	Let $P\in {_V\calM}$ be projective. Then $P\otimes N\in {_V\calM}$ is projective for all $M\in{_V\calM}$
+\end{thm}
+\begin{prf}
+	To see this, notice that $P$ is a summand of $V^{\oplus b}$. But then if $P\oplus Q\cong V^{b}$, then 
+	\[[P\otimes N]\oplus[Q\otimes N]=[P\oplus Q]\otimes N=(V^b)\otimes N=(V\otimes N)^b\]
+	and by the last lemma, this is free.
+\end{prf}
+
+\begin{lem}\label{lem-tensorproj}
+	\[\projdim(M\otimes N)\le \min\{\projdim M,\projdim N\}\]
+\end{lem}
+\begin{prf}
+	We only show that $\projdim M\otimes N\le \projdim M$. If $\projdim M=\infty$, we are done.
+	If it is finite, take any minimal projective resolution of $M$ and tensor with $N$. This is a projective 
+	resolution of $M\otimes N$ of length at most $\projdim M$. 
+\end{prf}
+\begin{rmk}
+	The fact it is still exact holds because we can appeal to the vector space structure.
+\end{rmk}
+
+\brk
+
+Some more results (mostly for finite dimensional algebras):
+\begin{thm}[Radford `75]
+	If $V$ is finite dimensional, then the antipode $S$ has finite order:
+	\[S^d=\id_V\]
+\end{thm}
+\begin{thm}[Larson-Radford `88]
+	Suppose $\ch k\ne 0$ (can be relaxed but gets more ugly). Then the following are equivalent:
+	\begin{enumerate}
+		\item $\gldim V=0$ ($V$ is semisimple Artinian)
+		\item $\gldim V^*=0$
+		\item $S^2=\id_V$
+	\end{enumerate}
+\end{thm}
+\begin{thm}[Nichols-Zoeller `89]
+	Let $V$ be finite dimensional. If $Q$ is a Hopf subalgebra of $V$, then $_WV$ and $V_W$ are free.
+\end{thm}
+
+\section{February 6, 2019}
+First some history: in 1899, Maschke proves the following:
+\begin{thm}
+	Let $G$ be a finite group. Then $kG$ is semisimple (Artinian) if and only if $\ch k\nmid |G|$.
+\end{thm}
+Today, we will see an analog of Maschke for Hopf algebras:
+\begin{thm}[``Maschke's Theorem'' (Larson-Sweedler `69)]
+	Let $V$ be a finite-dimensional Hopf algebra. Then the following are equivalent:
+	\begin{enumerate}
+		\item $\gldim V = 0$ ($V$ is semisimple Artinian)
+		\item $\varepsilon(\int^l)\ne 0$
+		\item $\varepsilon(\int^r)\ne 0$
+	\end{enumerate}
+\end{thm}
+
+Now say $V=kG$. Then $\int=\sum_{g\in G} g$ is both a left and right integral of $V$. This uses that $\varepsilon(g)=1$.
+
+But then if $\varepsilon(\int)=\sum_G \varepsilon(g)=\sum_G 1=|G|\ne 0\Leftrightarrow \ch k\nmid |G|$.
+
+\subsection{Example of the Day}
+\begin{ex}[Kac-Paljutkin Algebra]
+	$k=\C$ or any field with $\ch k\ne 2$. Then $V=k\langle x,y,z\rangle$ modulo the relations
+	\begin{align*}
+		x^2=y^2=1\\
+		xy=yx\\
+		xz=yz\\
+		xy=xz\\
+		x^2=\frac{1}{2}(1+x+y-xy)
+	\end{align*}
+	Then $V$ has a $k$-basis $\{1,x,y,z,xy,xz,yz,xyz\}$, so as an algebra $V\cong k^4\oplus M_2(k)$.
+
+	Then $\Delta(x)=x\otimes x$, $\Delta(y)=y\otimes y$, $\Delta(z)=(1\otimes 1+1\otimes x+y\otimes 1-y\otimes x)(z\otimes z)$.
+
+	And $\varepsilon(x)=\varepsilon(y)=\varepsilon(z)=1$ and $S(x)=x$, $S(y)=y$ and $S(z)=z$. Recall that although $S$ looks like the identity map, 
+	$S$ is an \textbf{antihomomophism}. So actually $S^2=\id_V$.
+
+	Claim:
+	\[\int=1+x+y+z+xy+xz+yz+xyz\]
+	then $\varepsilon(\int)=8\ne 0$, so $V$ is semisimple.
+
+	Some nice facts:
+	\begin{enumerate}
+		\item $V^*\cong V$
+		\item $V$ is the unique eight dimensional noncommutative, noncocommutative semisimple Hopf algebra.
+		\item Let $Q_8$ be the quaternion group. Then $kQ_8$ is eight dimensional. Therefore $V\cong kQ_8$ as algebras, but \textbf{not as coalgebras.}
+		The reason behind this is that $kG$ is \textit{always} cocommutative.
+	\end{enumerate}
+\end{ex}
+
+\subsection{HW of the day}
+\begin{prob}
+	If $V$ is eight dimensional and noncommutative, then $V\cong k^4\oplus M_2(k)$ as an algebra. Thus if $G$ is nonabelian group of order 8, then $\C G\cong \C^4\oplus M_2(\C)$.
+\end{prob}
+
+\subsection{Proof of ``Maschke's theorem''}
+Now we actually do the proof:
+\begin{prf}
+	(a) $\Rightarrow$ (b): Say the global dimension of $V$ is zero. Then $V$ is semisimple by Artin-Wedderburn and we have a decompositon
+	\[V\cong \oplus_1^s M_{n_i}(D_i)\]
+	and every minimal nonzero ideal has the form $M_{n_i}(D_i)$.
+
+	Consider the ideal $k\int^l=I$ which is an ideal since $v\int^l=\varepsilon(v)\int^l\in I$. But then consider that if $v'\int^l\in I$, 
+	\[v'\int^l v=\varepsilon(v')\int^l v\in k\int^l\]
+	so in fact it is a two-sided ideal. Thus $k\int^l\cong M_{n_i}(D_i)$, so 
+	\begin{align*}
+		(k\int^l)^2&=k\int^l
+		\varepsilon(\int^l)\int^l=(\int^l)^2&=a\int^l
+	\end{align*}
+	for some nonzero $a$. But then $\varepsilon(\int^l)\ne 0$.
+
+	(b) $\Rightarrow$ (a): Now let $e=\frac{1}{\varepsilon(\int^l)}\int^l$. Then $e^2=e$ and $ke=k\int^l$ is a left $V$-module.
+	But then $V=Ve\oplus V(1-e)$ so $ke=k\int^l$ is projective, so $\projdim_v k=0$.
+
+	The last part follows by Lorenz-Lorenz -- $\gldim V=\projdim_V k$.
+\end{prf}
+
+\subsection{$\Ext$ groups}
+We define these in the usual way through projective resolutions:
+
+Let $A$ be an algebra. Then $_A\calM$ is an abelian cateogry, so $\Hom_A(M,N)\in\Vectk$.
+\begin{defn}
+	A (covariant) functor $F:{_A\calM}\to{_B\calM}$ is called \textbf{left exact} if, for any short exact sequence
+	\[0\to L\to M\to N\to 0\]
+	we have the exact sequence
+	\[0\to F(L)\to F(M)\to F(N).\]
+
+	Similar for right exact. If both, it's just exact.
+\end{defn}
+\begin{lem}
+	For any $W\in{_A\calM}$, the functor $\Hom(W,-)$ is left exact.
+\end{lem}
+We also have that $\Hom_A(P,-)$ is exact iff $P$ is projective, so this gives us a nice characterization of projective objects in Abelian categories.
+Then the category $\Ch({_A\calM})$ is the category of chain complexes in $A$ modules with chain maps.
+\begin{lem}
+	$\Ch({_A\calM})$ is an Abelian category.
+\end{lem}
+\begin{lem}
+	A functor $F:{_A\calM}\to{_B\calM}$ can be extended to a functor between the appropriate chain complex categories.
+\end{lem}
+
+\subsection{Looking Forward}
+On Friday, we may see the following result:
+\begin{thm}
+	Under some mild hypothesis (including finite dimensionality),
+	\[\projdim_V k=\gldim V\]
+\end{thm}
 \end{document}

### 2019-02-13 Lie Algebra Notes/Lie Algebras.tex
I should push more frequently

@@
 
+%Shade definitions
+\theoremindent0cm
+\theoremheaderfont{\normalfont\bfseries} 
+\def\theoremframecommand{\colorbox[rgb]{.9,.9,.9}}
+\newshadedtheorem{ex}[thm]{Example}
+
 %Man, that's really good! Let's use the same thing for definitons.
@@
 \newtheorem{prf}{Proof}
+\newtheorem{conj}{Conjecture}
 
@@
 %Image
-\DeclareMathOperator{\im}{Im}
+\DeclareMathOperator{\im}{im}
+\DeclareMathOperator{\Hom}{Hom}
+\DeclareMathOperator{\End}{End}
 
@@
 \DeclareMathOperator{\n}{\mathfrak{n}}
+\DeclareMathOperator{\h}{\mathfrak{h}}
+\DeclareMathOperator{\g}{\mathfrak{g}}
 \DeclareMathOperator{\diag}{diag}
@@
 	so it suffices to show we also have the $e_{ij}$ for $i\ne j$.
-
+	This is clear whenever $k\ge 2$ (so that we have three columns in our matrices). Let $i\ne j$ be arbitrary
+	and $l$ be a (the) third index. Then 
+	\[e_{il}e_{lj}]=e_{ij}.\]
+
+	If $k=1,$ then 
+	\[e_{21},e_{11}-e_{22}]=e_{21}+e_{21}=2e_{21}\]
+	and so $e_{21}$ is in the space generated by $[\calB \calB]$ and an analagous proof
+	gives us the rest of the basis.
+
+	\subsubsection*{Type $B_k$:}
+	This one is getting a bit long and is low on my priority list so might not be finished.
 \end{sol}
@@
 \begin{sol}
-	Assume that $L$ is nilpotent. Then defining $L^0=L$, and by induction $L^i=[LL^{i-1}]$, this 
+	Assume that $L$ is nilpotent. That is, defining $L^0=L$, and by induction $L^i=[LL^{i-1}]$, this 
 	means that $L^n=0$.
+
+	We proceed by induction on $\dim L$: for the base case, if $L=0$ the result doesn't really mean anything,
+	and when $\dim L=1$, the trivial subspace has codimension 1.
+
+	Assume now that $\dim L=i$ and that every Lie algebra of dimension less than $i$ has a codimension 1 ideal. 
+	Then we have two subcases to consider. If $L$ is commutative then
+	every vector subspace is an ideal, so take a codimension one subspace and we are finished.
+
+	Finally assume $[LL]\ne 0$ and consider the quotient $Q=L/[LL]$. By 
+	our assumptions (and since $L$ is nilpotent $L\ne[LL]$) we get that
+	\[0<\dim Q<\dim L\]
+	and that, in fact, by rank-nullity
+	\[\dim Q = \dim L - \dim[LL].\]
+	But then by the induction hypothesis there is an ideal $J\lhd Q$ with dimension $\dim Q - 1$
+	and pulling this back through the quotient yields the ideal $\bar J\lhd L$ that has dimension $\dim Q-1+\dim[LL]=\dim L-1$. 
+	This is our codimension one ideal and we complete the proof.
 \end{sol}
@@
 \begin{sol}
-
+	Using the result above, we can write every nilpotent $L$ as $K+Fx$ for some $K\lhd L$.
+	Now since $K\le L$, and since $L$ is simple, for some $n$: $L^n=[L^{n-1}L^{n-1}]=0$,
+	or equivalently every $n$-fold bracket in $L$ is zero.
+
+	Assuming that $n$ is minimal, this means that one can construct a $(n-1)$-fold bracket
+	\[x=[x_1[x_2[x_3\cdots[x_{n-2}x_{n-1}]\cdots]]]\]
+	which is nonzero. But then $[xL]=0$ so in particular $[xK]=0$ whence $x\in C_L(K).$
+
+	Therefore we can choose an $m$, as suggested, such that $C_L(K)\in L^m\setminus L^{m-1}$
+	and let $\delta$ be the linear map sending $K$ to zero and $x$ to some $z\in C_L(K)\setminus L^{m+1}$.
+	We will now establish this is an outer derivation.
+
+	Let $a,b\in L$ be arbitrary with $a=k_a+l_ax$ and $b=k_b+l_bx$ and compute
+	\begin{align*}
+		\delta(ab) &= \delta(k_a+l_ax)(k_b+l_bx)\\
+		&= \delta(k_ak_b+k_al_bx+l_axk_b+l_axl_bx)\\
+		&=0+k_a\delta(l_bx)+\delta(l_ax)k_b+\delta(l_ax)l_bx+l_ax\delta(l_bx)\\
+		&=(k_a+l_ax)\delta(l_bx)+\delta(l_ax)(k_b+l_bx)\\
+		&=a\delta(k_b+l_bx)+\delta(k_a+l_ax)b\\
+		&=a\delta(b)+ \delta(a)b
+	\end{align*}
+	so $\delta$ is a derivation. It remains to show it is not inner.
+
+	Assume by contradiction that it were. Then $\delta =\ad w$ for some $w\in L$. But then notice
+	that $[wK]=0$ so $w\in C_L(K)\subseteq L^m$. But then bracketing with one more thing puts us in $L^{m+1}$, so
+	$[w,x]\in L^{m+1}$. But then
+	\[[w,x]=\ad w(x)=\delta(x)=z\in L^{m+1}\]
+	a contradiction. Thus $\delta$ is an outer derivation.
 \end{sol}
@@
 \begin{sol}
-
+	Compute
+	\[[xy]=\begin{pmatrix} 
+		0 & 1 & 0 & \cdots & 0\\
+		0 & 0 & 2 & \cdots & 0\\
+		\vdots & \vdots & \vdots & \ddots & \vdots\\
+		0 & 0 & 0 & \cdots & p-1\\
+		0 & 0 & 0 & \cdots & 0
+	\end{pmatrix}-\begin{pmatrix} 
+		0 & 0 & 0 & \cdots & 0\\
+		0 & 0 & 1 & \cdots & 0\\
+		\vdots & \vdots & \vdots & \ddots & \vdots\\
+		0 & 0 & 0 & \cdots & p-2\\
+		p-1 & 0 & 0 & \cdots & 0
+	\end{pmatrix}=\begin{pmatrix} 
+		0 & 1 & 0 & \cdots & 0\\
+		0 & 0 & 1 & \cdots & 0\\
+		\vdots & \vdots & \vdots & \ddots & \vdots\\
+		0 & 0 & 0 & \cdots & 1\\
+		1 & 0 & 0 & \cdots & 0
+	\end{pmatrix}=x\]
 \end{sol}
+Thus if $L$ is the vector space spanned by $x$ an $y$, since it is closed under bracket, 
+it is a Lie subalgebra of $\gl(p,F)$. Now because of this computation it is easy to see that
+$[LL]=Fx$, so $L^2=0$, whence $L$ is solvable.
+
+Notice that multiplication by $x$ on the right acts by permuting the columns of a matrix so $x^p=\id$,
+whence the eigenvalues of $x$ are roots of $x^p-1=(x-1)^p$ since $\ch F=p$. Let $v$ me any element of $L$ 
+fixed by the action of $x$ via $\ad x$ on $L$. Then
+\[bx=b[x,y]=[x,ax+by]=[x,v]=xv-vx=v\]
+but then since the only eigenvalue can be 1, $v=x$. But $[y,x]=-x$, which cannot be a scalar multiple of $y$,
+so $x$ and $y$ can have no common eigenvector.
 \newpage
@@
 
+\section{January 18, 2019}
+\begin{prob}
+	The ``board problem'' for this week: Show that a semisimple Lie algebra over $\C$ is generated as a Lie
+	algebra by two elements.
+\end{prob}
+
+\subsection{Returning to Jordan Decomposition}
+Given an $n\times n$ over an algebraically closed field $k$ of any characteristic, let $T$ be the corresponding linear
+transformation on $k^n$. We know from the proof of the Jordan canonical form that 
+\[V=k^n=\oplus_{a\in k}\ker(T-aI)^n\]
+the sum of the \textbf{generalized eigenspaces} of $T$ on $V$. We then use the CRT:
+there is a polynomial $p(T)$ such that $p(T)=a\pmod (T-a)^n$ for all eigenvalues $a$ and
+$p(T)=0\pmod{T}$.
+
+This is clear since the eigenvalues are distinct, so that all moduli are pairwise relatively prime.
+SO then $p(T)$ acts on $\ker(T-A)^n$ so it is \textbf{semisimple} (diagonal) and $T-p(T)$, acting by $T-a$ on
+$\ker(T-a)^n$, acts nilpotently.
+
+Hence $p(x)$ and $x-p(x)$ are polnomials in $x$ without constant term such that as matrices they commute
+and $x$ is semisimple (diagonal) and $x-p(x)$ is nilpotent. Moreover, tere is only one way to write $x$ as $s+n$ where $s$ is semisimple, $n$ is nilpotent, and $[sn]=0$.
+This is since given $x=s+n$ where $s,n$ both commute with $x$, so with $p(x)$ and $x-p(x)$, we get
+\[p(x)-s+(p(x)-x)+n\]
+(check this) with the left side the sum of \textbf{commuting} semisimple matrices is semisimple and
+the right side is the sum of \textbf{commuting} nilpotent matrices whence nilpotent, but the only matrix
+which is both semisimple and nilpotent is the zero matrix.
+\begin{rmk}
+	Notice that commutativity above is vital to the statement.
+\end{rmk}
+
+Why is this important? We want to study the representations of Lie algebras, even the linear ones, and we would like to know that the semisimple and nilpotent parts $x_s, x_n$ of
+any $x$ lying in such an algebra \textbf{continue} to act semisimply and nilotently in the representation 
+(and that the parts $x_s$ and $x_n$ lie in L if $x$ does, which is true when $L$ is semisimple).
+
+We have already seen that if $d$ is diagonal, then 
+<FILL THIS IN>
+
+Now we want to head towards a criterion in terms of traces alone for a matrix to be nilpotent.
+Eventually we want to determine when a Lie algebra is solvable.
+
+\begin{lem}
+	Suppose that $k$ is algebraically closed of characteristic zero and that $B\subseteq A$ are subspaces of $\gl(n,k)$ and that $x(\in M?)$ is a matrix such that 
+	$\tr(xy)=0$ for all $y\in M=\{z\in\gl(n,k)|[zA]\subseteq B\}$. Then $x$ is nilpotent.
+\end{lem}
+\begin{prf}
+	Write $x=s+n$, the Jordan decomposition. Since $x\in M, s,n\in M$ since $s$ and $n$ are polynomials in 
+	$x$ without constant term. We must show that $s=0$ -- that all eigenvalues of $s$ are zero.
+
+	Let $a_1,\dots,a_n$ be the eigenvalues of $s$. These all lie in $k$, so they span a finite-dimensonal $\Q$-subspace $E$
+	of $k$. To see this, we will show that any $\Q$-linear function $f:E\to\Q$ must be zero. Given $f$, 
+	choose a polynomial $r(s)$ such that $r(a_i-a_j)=f(a_i)-f(a_j)$ for all $i$ and $j$ (using the Lagrange interpolation theorem).
+
+	This is consistent since $a_i-a)j=a_k-a_l$ (using the linearity of $f$) implies the image of these under $r$ are equal.
+	Then looking at $r(\ad s)$, we find that its eigenvalues are $r(a_i-a_j)$ as well as $f(a_i)-f(a_j)$,
+	whence $\sum f(a_i)a_i=0=\tr(sy)$ for asuitable $y$.
+
+	Applying $f$, we get $\sum f(a_i)^2=0$ whence all $f(a_i)=0$ (since $f(a_i)\in\Q$) and $f=0$, as desired.
+\end{prf}
+
+\subsection{Cartan's Criterion}
+How can we turn this into something useful for us in the Lie algebra world?
+\begin{thm}[Cartan's Criterion]
+	Let $L$ be a linear Lie algebra over (alg. closed and char zero) $k$. Suppose that $\tr(xy)=0$
+	for all $x\in L$ and $y\in[LL]$. Then $L$ is solvable.
+
+	As a consequence, suppose that $L$ is a Lie algebra such that 
+	$\tr((\ad x)(\ad y))=-$ for all $x$ and $y$ above. Then $L$ is solvable.
+\end{thm}
+\begin{rmk}
+	Notice that the consequence really does follow from the theorem: if $\tr((\ad x)(\ad y))=0$
+	then $\ad L\cong L/Z$ is solvable where $Z$ is the center of $L$ whence $L$ is also solvable
+	since $Z$ is abelian.
+
+	To prove that $L$ is solvable, under our hypothesis, it is enough to show that $[LL]$ is nilpotent, 
+	but (due to Engel's theorem) it suffices to show that $[LL]$ consists of nilpotent matrices.
+\end{rmk}
+To prove this criterion, we need a fact about traces (the \textbf{associativity of traces}):
+\[\tr([xy]z)=\tr(x[yz]).\]
+\begin{prf}
+	We know that $\tr(xy)=0$ for all $x\in L$, $y\in [LL]$ and any such $x$ maps $L$ to $[LL]$ under $\ad$.
+	We would like to know that for any $x\in M$, that $\tr(xz)=0$ for all  $x\in[LL]$. We know this is true if $z\in L$
+	so by the associativity of trace, this follows.
+
+	Returning to the Jordan decomposition, let $A$ be any finite-dimensional algebra over $k$ (not necessarily associative)
+	and let $d$ be a derivation of $A$. THen the semisimple parts $d_s$ and $d_n$ of $d$ are also derivations.
+
+	To see this, we write $A$ as the direct sum of the generalized eigenspaces $A_a$ of itself under the action of $d$. Since $d$ is a derivation,
+	it ends up becoming clear that $A_aA_b\subseteq A_{a+b}$, so that the transformation from $A$ to itself action by $a$ on $A_a$ is a derivation.
+	But this is $d_s$. Then $d_n=d-d_s$ is also a derivation.
+\end{prf}
+
+\section{January 23, 2019}
+Some clarifications from last time:
+\begin{lem}
+	Let $A\subseteq B$ be subspaces of $\gl(n,l)$ for some algebraically closed field of char zero.
+	Let $M=\{x\in\gl(n,k)|[xB]\subseteq A\}$. Suppose that $x\in M$ such that $\tr(xy)=0$ for all $y\in M$.
+	Then $x$ is nilpotent.
+\end{lem}
+\begin{rmk}
+	This used the fact that if $x=x_s+x_n$ then $\ad x=\ad x_s+\ad x_n$ is the Jordan decomposition for $\ad x$.
+
+	Furthermore, when applying this lemma to Cartan's criterion for sovability of Lie subalgebras of $\gl(n,k)$, 
+	we claimed that, if $\tr(xy)=0$ for all $x\in L$ and $y\in [LL]$, then $L$ is solvable. This follows by our lemma
+	since if $x\in M=\{z\in\gl(n,k)|[zL]\subseteq[LL]\}$ and if $y,z\in L$ then $\tr(x[yz])=\tr([xy]z)=0$ by hypothesis.
+
+	So by the lemma $[LL]$ consists of nilpotents, so is nilpotent and thus solvable.
+\end{rmk}
+
+\begin{defn}
+	If $L$ is any Lie algebra over $k$, then $\kappa(x,y)=\tr((\ad x)(\ad y))$, is a bilinear 
+	form on $L$ and is called the \textbf{Killing form.}
+
+	It is symmetric and associative ($\kappa([xy],z)=\kappa(x,[yz])$).
+\end{defn}
+\begin{rmk}
+	Now in the context of \textit{general (as opposed to linear) Lie algebras} we can restate
+	Cartan to say:
+\end{rmk}
+\begin{thm}[Cartan Redux]
+	If $\kappa(x,y)$ for all $x\in [LL]$ and $y\in L$, then $L$ is solvable.
+\end{thm}
+
+\subsection{Semisimple Lie Algebras}
+What we are interested in, actually, is the polar opposite of solvable algebras: namely,
+\textbf{semisimple} Lie algebras, having no nonzero solvable ideals.
+\begin{thm}
+	$L$ is semisimple if and only if the Killing form $\kappa$ is nondegenerate.
+\end{thm}
+\begin{rmk}
+	This sort of naturally follows from our discussion above -- we want the opposite property for our algebra
+	and we require that the form be as far from zero as possible.
+\end{rmk}
+\begin{rmk}
+	If $I\lhd L$, we note that, if $\kappa_I$ is the Killing form for $I$ and $\kappa$ the Killing form for
+	$L$, then $\kappa_I=\kappa|_{I\times I}$. This is because if $x,y\in I$ then $\ad x$ and $\ad y$ both map $L$ into $I$.
+	Thus if $M$ is any $n\times n$ matrix over $k$ mappting $V=k^n$ to a proper subspace $W$, then $\tr M=\tr M|_W$, since 
+	extending a basis from $W$ to $M$ is such that the added basis vectors make no contribution to the trace of $M$.
+\end{rmk}
+
+\begin{prf}
+	If $\kappa$ is nondegenerate and $S$ is a solvable ideal of $L$, then $\kappa|X=\kappa S$ has
+	\textbf{radical} $\{z\in S:\kappa(z,y)=0,\forall y\in S\}$ which is at least $[SS]$, and then $K$ would have a 
+	radical as well.
+
+	More precisely, if $S$ is a solvable ideal of $L$ then there is a nonzero abelian ideal $A$ of $L$
+	(the last nonzero term in the derived series for $L$) and then if $x\in A$ and $y\in L$, then $\ad x\ad y$ maps $L$
+	into $A$ and $(\ad x\ad y)^2=0$, so $\kappa$ has radical at least $A$.
+
+	Conversely, if $L$ is semisimple, then the radical of $\kappa$ as defined aobve would be a solvable ideal of $L$,
+	by associativity and Cartan's criterion, so $\operatorname{Rad}(\kappa)=0$.
+\end{prf}
+
+The nondegeneracy of the Killing form is \textbf{the key} result that makes the classification
+of semisimple Lie algebras tractable. For instance, in the case of finite simple groups the classification required 
+10,000 pages of journal papers and it boils down to not having any kind of simple criterion like this.
+
+\subsection{Applications}
+As a first step, we show 
+\begin{thm}
+	A Lie algebra $L$ is semisimple if and only if it is a finite direct sum of simple ideals $I$.
+\end{thm}
+\begin{prf}
+	This is clear since if $I$ is an ideal then $I^\perp=\{y\in K:\kappa(y,x)=0\forall x\in I\}$
+	is again an ideal of $L$ and $I\cap I^\perp=0$ by Cartan. Then we can keep doing this, pulling off minimal ideals.
+\end{prf}
+
+As a second application, we will show that ever derivation $d$ of a semisimple Lie algebra $L$ is inner.
+\begin{prf}
+	$L$ is semisimple, so $L=[LL]$ and $Z(L)=0$, so $\ad L\cong L$. Now if $D=\operatorname{Der} L$, then
+	$L\lhd D$ (as in homework). If $L\ne D$, then take the orthogonal of $L$ under the killing form in $D$.
+
+	Thus $L\cap L^\perp=0$ and $\dim L^\perp\ge\dim D-\dim L$, but $[LL^\perp]=0$, which says exactly that if $m\in L^\perp$, then $m(x)=0$
+	for all $x\in L$, so $L^\perp=0$ and $L=D$.
+\end{prf}
+
+Next we can extend the Jordan decomposition to semisimple Lie algebras. We have already seen that if $L$ is semisimple then $\Der L$ is closed
+under Jordan decomposition but $\Der L=L$, so any $x\in L$ can be written as $x_s+x_n$ \textbf{FOR $x_s,x_n\in L$ (important)} where $\ad x_s$ and $\ad x_n$
+act semisimply and nilpotently on $L$ and that $[x_sx_n]=0.$
+
+\begin{rmk}
+But what if $L$ is a semisimple linear Lie algebra over $k$? Then if $x\in L$, we have  as above $x_s,x_n\in L$
+but also $x=s+n$ for some $s,n\in\gl(n,k)$ where $s$ is semisimple and $n$ is nilpotent and $[sn]=0$. Then we can ask:
+do $s$ and $n$ lie in $L$? The answer is yes, but requires more machinery Once we know this, there are immediate applications to 
+representation theory of $L$ (maps $\varphi:L\to\gl(n,k)$):
+
+Given any such representation $\varphi$ and $x\in L$ with $x=x_s+x_n$, we get that $\varphi(x_s)$ and $\varphi(x_n)$
+\textbf{continue to act semisimply and nilpotently} on $k^n$. This is what allows us to get a real handle on the representation theory.
+\end{rmk}
+
+\begin{thm}
+	Given any \textbf{finite dimensional} module $M$ over a finite dimensional $M$ over a semisimple Lie
+	algebra $L$ is completely reducible.
+\end{thm}
+\begin{rmk}
+	We will prove this next time, using also certain constructions on modules as well as the matrix
+	representations of elements.
+
+	This result ends up being extremely useful not only in the (more obvious) representation theory of $L$, 
+	but also in the structure theory of $L$ by considering the adjoint action of $L$ on itself.
+\end{rmk}
+
+\section{January 25, 2019}
+\subsection{Weyl's Theorem}
+TOday we'll prove Weyl's theorem:
+\begin{thm}[Weyl]
+	Any finite dimensional module over a semisimple Lie algebra is complete reducible ( a finite direct sum of simple modules).
+\end{thm}
+To prove this, we need some preparation: namely a couple of methods of producing new moduels over any Lie algebra $L$ from old ones.
+
+First of all, if $M,N$ are $L$-modules, then $M\otimes_k N$ is an $L$ module of dimension the product
+of the dimensions of $M$ and $N$. Here we define the action
+\[x\cdot(m\otimes n)=x\cdot m\otimes n+m\otimes x\cdot n\]
+where we can quickly check that 
+\[[xy]\cdot(m\otimes n)=x(y(m\otimes n))-y(x(m\otimes n)).\]
+
+Similarly, $\Hom_k(M,N)$ is an $L$ module via
+\[(xf)(v)=xf(v)-f(xv)\]
+and again we can check this defines a module structure, noting that $xf=0$ for all $x\in L$ if and only if $f$ is an $L$ module homomorphism.
+
+\begin{lem}[Shur's Lemma]
+	If $M$ is an irreducible $L$ module, then the only $L$ module maps in $\operatorname{End}(M)$ are scalars.
+\end{lem}
+
+Last we need
+\begin{defn}
+	We have the \textbf{trace form} on $L$:
+	\[(x,y)=(\tr\varphi(x))(\tr\varphi(y))\]
+	and by associativity and Cartan's criterion, we know (assuming that $L$ acts faithfully) that $(-,-)$ is nondegenerate.
+\end{defn}
+Recall that given any basis $x_i$ of $L$, we can define the dual basis $y_i$ such that $(x_i,y_j)=\delta_{ij}$.
+\begin{defn}
+	The \textbf{Casimir element} attached to a homomorphism $\varphi:L\to\gl(n,k)$, for some $n$, to be the matrix
+	\[c_\varphi=\sum_1^m\varphi(x_i)\varphi(y_i).\]
+\end{defn}
+\begin{rmk}
+	This matrix commutes with $\varphi(x)$ for all $x\in L$. This can be done by showing $a_{ij}=-b_{ji}$
+
+	Then in computing $[xc_\varphi]$, we find that the coefficients of $\varphi(x_i)\varphi(y_j)$
+	is $a_{ij}+b_{ji}=0$. Thus $c_\varphi$ commutes with the matrices of $\varphi(L)$.
+
+	Hence if $V=k^n$ ws te module being acted upon and if $V$ is irreducible, then $c_\varphi$ must be a scalar (Schur).
+	What is this scalar? Well since $\tr c_\varphi=m=\dim L$, we get
+	\[\frac{\dim L}{\dim V}I=\frac{m}{n}I.\]
+\end{rmk}
+\begin{prf}
+	The proof here will basically follow Humphreys, but will include a comment that \textit{should have appeared} in the text
+	(much to Monty's chagrin).
+
+	Notice that we must show that every submodule $N$ of $M$ has a complement $N'$ so that $M$ is the direct sum of $N$ and $N'$ and $N'$ is a submodule.
+
+	For the rest of it, peek at the book. :)
+\end{prf}
+\begin{rmk}
+	This result fails for semisimple Lie algebras in positive characteristic, although we won't really be dealing with this much.
+	This parallels the fact that Maschke's theorem fails when $\ch k$ divides the order of $G$.
+\end{rmk}
+\subsection{Structure}
+We now consider the simples case of a semisimple Lie algebra, namely $\sl(2,k)$. Here we know that we have the basis
+$h=e_{11}-e_{22}$, $x=e_{12}$ and $y=e_{21}$ where
+\[[hx]=2x\quad [hy]=-2y\quad [xy]=h\]
+
+Next time we will see that the diagonal matrix $hin L$ continues to act diagonally on \textit{any} finite dimensional representation of $L$.
+Thus the first thing to look at here is how the representation $V$ decomposes as a sum of its eigenspaces
+\[V=\oplus_{\lambda\in k}V_\lambda\]
+but since $V$ is assumed to be finite dimensional, this is actually a finite sum. In particular, there is $\lambda$ such that 
+$V_{\lambda+2}=0$.
+\begin{defn}
+	We call the $\lambda$ where $V_\lambda\ne 0$ \textbf{weights} of $V$.
+\end{defn}
+So there is a weight $\lambda$ such that $\lambda+2$ is not a weight. In general, due to the relations above on $L$,
+\[xV_\mu\subseteq V_{\mu+2}\quad yV_\mu\subseteq V_{\mu-2}\]
+and we call $x$ and $y$ (or their matrices) the \textbf{raising and lowering operators}.
+
+Then we can pick $\lambda$ a weight such that $\lambda+2$ is not and pick $v_0\in V\setminus 0$. Then set
+\[v_i=\frac{y^iv_0}{i!}\]
+so 
+\[hv_i=(\lambda-2i)v_i,\quad yv_i=(i+1)v_{i+1}\quad xv_i=(x-i+1)v_{i-1}\]
+where this can be proved by induction.
+
+But the nonzero vectors $v_i$ are independent, so there is a least $m$ with $v_m\ne 0$ and $v_{m+1}=0$. 
+The last equation with $i=m+1$ then forces $\lambda-m=0$, so $\lambda=m$ is a nonnegative integer!
+
+It follows that if $V$ is irreducible that $\dim V=m+1$ and $v_0,\dots,v_m$ are a basis for $V$.
+
+This ends up capturing the situation with \textit{any} irreducible representation of $\sl(2,k)$. We get integer eigenvalues, all of the same parity, from $-m$ to $m$.
+
+\section{January 28, 2019}
+We'll start off with tying up some loose ends from last time: 
+\begin{lem}
+	Let $L$ be a semisimple subalgebra of $\gl(n,k)$ (once more, $k=\bar k$ and $\ch k=0$). Let $x\in L$. Then 
+	the semisimple and nilpotent parts $x_s$ and $x_n$ of $x$ lie in $L$. Thus the Jordan decompositions of $x$ 
+	(as a matrix and as an element in $L$) coincide.
+\end{lem}
+\begin{prf}
+	To see this, recall that we know that $\ad x_x$ and $]ad x_n$ are polynomials with zero constant term in $\ad x$ and
+	so bracket $L$ into $\gl(V)$. So if $x_s,x_n\in N=\{y\in\gl(n,k):[yL]\subseteq L\}$ and $L$ is an ideal of $N$.
+
+	We want to show that $L-N$ but this is false! we need to get the scalar matrices out. So replace $N$ by those elements 
+	such that if $y\in N$ then $y$ restricted to any irreducible submodule $k^n$ of $L$ has trace zero.
+
+	We have $L\subseteq N$... (snip)
+
+	So by Weyl's thoerem we can write $N=L\oplus M$ as an $L$ module. But $L$ is an ideal of $N$ and $[LM]=0$, so the matrices in $M$
+	all commute with $L$. By Schur's lemma, they act by scalars on any irreducible submodule $W$ of $k^n$, which must be zero, by the trace condition.
+
+	Since $k^n$ is the direct sum of irreducible modules, we get $M=$, so $L=N$.
+\end{prf}
+
+Hence we have fully justified our characterization of irreducible $(m+1)$-dimensional $\sl(2,k)$ modules $V$:
+\begin{itemize}
+	\item We always have a basis $v_0,\dots,v_m$ with $v_{-1}=v_{m+1}=0$ such that:
+	\item $hv_i=(m-2i)v_i$
+	\item $yv_i=(i+1)v_{i+1}$
+	\item $xv_i=(m-i+1)v_{i-1}$
+\end{itemize}
+we can check that the bcracket relations are satisfied starting from this as a definition.
+\begin{rmk}
+	In fact, if you replace $m$ by any $\lambda\in k$ and let all $v_i$'s form a basis for all non-negative $i$ (that is easing
+	the restriction $v_{m+1}=0$), you get an infinite dimensional $L$-module. Whenever $\lambda\in\N$, this module is \textbf{not}
+	completely reducible.
+\end{rmk}
+\subsection{One more $\sl(2,k)$ Computation}
+\[(\exp\ad ax)(\exp\ad hy)(\exp\ad ax)=\begin{pmatrix}
+	1+ab & a(2+ab)\\
+	b & 1+ab
+\end{pmatrix}=\begin{pmatrix}
+	0 & -b^{-1}\\
+	b & 0
+\end{pmatrix}\]
+if $ab=-1$. Multiplying by its transpose, we get
+\[\begin{pmatrix}
+	b^{-2} & 0 \\
+	0 & b^2
+\end{pmatrix}\]
+So whenever $K$ is algebraically closed, we get a set of generators of $\sl(2,k)$ as products of matrices in $\Int L$ which is therefore 
+isomorphic to $\operatorname{PSL}(2,k)$.
+
+The action of $L$ on any irreducible module $V$ always lifts to $\operatorname{SL}(2,k)$ and lifts to $\operatorname{PSL}(2,k)$ if and only if $\dim V$ is odd.
+
+\brk
+
+The moral to draw from our classification of $\sl(2,k)$ moduoes of finite dimension is that the weights ($h$-eigenvalues) always occur in strings
+of integers. In particular the weight $i$ occurs overall with the same multiplicity (eigenspace dimension) as $-i$.
+
+Also for an irreducible module either the weight zero or the weight 1 occurs with multiplicity one (but not both).
+
+\subsection{Return to semisimple Lie algebras}
+Consider decomposing such an $L$ as a module over itself via the adjoint representation. If $L$ had only nilpotent elements, it itself would be nilpotent (it isn't)
+so $L$ has a \textbf{toral subalgebra} $T$ consisting of only semisimple elements.
+
+Then, in fact, $T$ is abelian, since otherwise we would have $[xy]=ky$ for $x,y\in T$ and $k\ne 0$.
+But on the other hand, since $\ad y$ is semisimple, but on the other hand $[yx]$ is a sum of eigenvectors of $\ad y$
+each with nonzero eigenvalues, a contradiction.
+
+\subsection{Decomposing $L$}
+Now let's fix a \textbf{maximal toral subalgebra} $\h$ (that is, maximal among toral subalgebras). Then we have
+\[L=\oplus_{\alpha\in\h^*}L_\alpha\]
+where $L_\alpha=\{x\in L:[xh]=\alpha(h)x\}$, since any commuting set of semisimple (diagonalizable) matrices are simultaneously diagonalizable.
+
+Now by Jacobi, $[L_\alpha L_\beta]\subseteq L_{\alpha+\beta}$ and bringing back in the Killing form,
+\[\kappa(L_\alpha,L_\beta)=0,\quad\text{if}\quad \alpha+\beta\ne 0\]
+so that $\kappa$ is nondegenerate on $L_0$:
+\[\kappa(L_0,L_\alpha)=0\quad\forall \alpha\ne 0.\]
+
+\begin{lem}
+	$H=L_0$, or equivalently, $H$ is its own centralizer.
+\end{lem}
+
+The proof can be found in the book, but using this, we can rewrite our decomposition as
+\[L=\h\oplus\bigoplus_{\alpha\in\h^*}L_\alpha\]
+and this is called the \textbf{root space decomposition} of $L$ in terms of $\alpha\in\h^*$.
+The terms $\alpha\in\h^*$  occurring above are called the \textbf{roots of $L$} (relative to $\h$). By definition \textbf{no root is zero.}
+
+Notation: Write $\Phi$ to be the set of all (nonzero) roots in $\h^*$.
+
+Since the restriction of $\kappa$ to $\h$ is nondegenerate, we get an identification of $\h^*$ with $\h$:
+\[\alpha\in\h^*\Leftrightarrow t_\alpha\in \h\]
+where $\alpha(h)=\kappa(t_\alpha,h)$. Moreover we can transfer $\kappa$ to $\h^*$:
+\[(\alpha,\beta)=\kappa(t_\alpha, t_\beta)\]
+which gives us a symmetric bilinear form on $\h^*$.
+
+\brk
+
+\textbf{Next Time: } we will place the form on a real vector space of dimension $\dim_k\h$. Best of all, 
+$(-,-)$ will be positive definite, giving us a nice orthonormal basis.
+
+\section{January 30, 2019}
+Let $L$ be a semisimple Lie algebra, $\h$ a maximal toral subalgebra. Last time we saw the root space decomposition above. Note
+we actually have $L=\h\oplus \bigoplus_{\alpha\in\Phi}L_{\alpha}$ where $\Phi\in\h^*$ are the values that give you nontrivial weight spaces,
+called the \textbf{root system} of $L$ (relative to $\h$) with elements called \textbf{roots.} It ends up we can embed these root systems in
+a Euclidean space, giving us a context in which to write some relatively simple axioms.
+
+\subsection{Root Systems}
+First,
+\begin{lem}
+	$\Phi$ spans $\h^*$.
+\end{lem}
+\begin{prf}
+	Othersiwe, some nonzero $h\in \h$ would have $\alpha(h)=0$ for all $alpha\in\Phi$, but then $h$ would be in the center of $L$, a contradiction.
+\end{prf}
+\begin{lem}
+	If $\alpha\in \Phi$, then $-\alpha\in\Phi$.
+\end{lem}
+\begin{prf}
+	This is due to the fact that otherwise $L_\alpha$ would lie in the radical of the Killing form. Thus
+	$[L_\alpha L_{-\alpha}]=\kappa t_\alpha$ where $\alpha(g)=\kappa(t_\alpha,h)$. Then you can follow the text.
+\end{prf}
+The takeaway from the discussion, which I admittedly slacked on typing, was that in considering $L_\alpha$ and $L_{-\alpha}$ we find that the structure
+has to be precisely the structure of $\sl(2,k).$
+
+\brk
+
+Then $L$ becomes a module for $S_\alpha\cong\sl(2,k)$ under the adjoint action. Look at the submodule
+\[M_\alpha=\h\oplus\bigoplus_{c\in k^*}L_{c\alpha}\]
+Clearly $M_\alpha$ is stable under the $S_\alpha$ action. Weights of $h_\alpha$ are integers, so we know that 
+$L_{c\alpha}=0$ if $c\notin\Z\cup\frac{1}{2}\Z$.
+
+But also all occurrences of the weight zero come from $\h$, a codimension one space, namely $\ker\alpha$, dies under the action of $x_\alpha$ and $y_\alpha$.
+The only other occurrence is $h_\alpha$ itself, which leads to $x_\alpha$ on the action $x_\alpha$ then dies (what)? Thus the weights occurring are $0,\cdots,0,2,-2$.
+
+So the takeaway is that the weight 4 doesn't occur and $L_{2\alpha}=L_{\alpha/2}=0$, so the weight 1 never occurs.
+Finally we get that $L_{c\alpha}=0$ if $c\ne \pm 1$ and $\dim L_{\alpha}=1$.
+
+\brk
+
+By the way, we also know that $[L_\alpha L_\beta]=L_{\alpha+\beta}$ by the way $x_\alpha$ and $y_\alpha$ act on any irreducible $S_\alpha$ module.
+
+Finally we also know that if $\alpha$ and $\beta$ are roots, then 
+\[\beta-\beta(h_\alpha)\alpha=\beta-\frac{2(\alpha,\beta)}{(\alpha,\alpha)}\alpha\]
+is a root as well and $\frac{2(\alpha,\beta)}{(\alpha,\alpha)}\in\Z$. The equation above is just a reflection in Euclidean space.
+
+\subsection{Rationality Properties}
+We know that $\Phi$ spans $\h^*$: choose a basis $\alpha_1,\dots,\alpha_n$ of $\h^*$ consisting of roots. Then any root $\beta$ takes the form 
+$\sum_i c_i\alpha_i$ for $\alpha_i\in k$. Apply $(-,\alpha_i)$ and divide by $(\alpha_i,\alpha_i)$, then we get
+\[\frac{2(\beta,\alpha_i)}{(\alpha_i,\alpha_i)}=\sum_j c_j\frac{2(\alpha_j,\alpha_i)}{(\alpha_i,\alpha_i)}.\]
+
+Finally we can do some row operations to find the unique solutions for the $c_j$ must lie entirely in $\Q$, so \textbf{all roots} lie in the $\Q$ subspace $E_\Q$
+of $\h^*$ whose dimension over $\Q$ is $n=\dim_k\h^*=\dim_k\h$.
+
+Moreover, for any $\alpha,\beta\in\h$,
+\[(\beta,\alpha)=\kappa(t_\beta,t_\alpha)=\sum_{\gamma\in\Phi}\gamma(t_\alpha)\gamma(t_\beta)=\sum (\gamma,\alpha)(\gamma,\beta)\]
+so in particular $(\beta,\beta)=\sum (\beta,\gamma)^2$ and dividing both sides by $(\beta,\beta)^2$ we get that $(\beta,\alpha)\in \Q$ for all roots and so more generally
+for all $\alpha,\beta\in E_\Q$. Apply our formula one last time to see
+that $(\lambda,\lambda)>0$ when $\lambda\ne 0$, so our form $(-,-)$ is symmetric, bilinear, and positive definite on $E_\Q$.
+
+We know set up a formal vector space $E_\R$ as the $\R$ span of a basis of $\h^*$ fixed earlier consisting of roots. Then we know that
+$\Phi\in\h^*$ satisfies the following properties:
+\begin{enumerate}
+	\item $\Phi$ spans $\R^n$
+	\item If $\alpha\in\Phi$, the only multiples of $\alpha$ also in $\Phi$ are $\pm\alpha$.
+	\item If $\alpha,\beta\in\Phi$, then $\beta-\frac{2(\beta,\alpha)}{)\alpha,\alpha)}\alpha\in\Phi$.
+	\item $\frac{2(\beta,\alpha)}{(\alpha,\alpha)}\in\Z$ for all roots.
+\end{enumerate}
+\begin{defn}
+	Any system $\Phi$ satisfying properties (a)-(c) above is an \textbf{(abstract) root system in $\R^n$}. If, furthermore,
+	$\Phi$ satisfies $(d)$, we call it a \textbf{crystallographic root system.} In this case, the $\Z$-span of $\Phi$ is called the
+	\textbf{root lattice} $\Lambda$, which is stable under all reflections $s_\alpha$ for $\alpha\in\Phi$.
+\end{defn}
+\begin{rmk}
+	We note that for any root system $\Phi$ that the subgroup of $O_n(\R)$ generated by all $s_\alpha$ is a finite group $W$
+	(finite since it acts on a finite set $\Phi$) called the \textbf{Weyl group.} $W$ consists of automorphisms of $\Phi$ -- that is,
+	linear maps on $\R^n$ sending $\Phi$ to $\Phi$ and preserving the dot product.
+\end{rmk}
+We will classify (mostly) all finite groups $W$ arising in this way, both in the crystallographic and non-crystallographic cases, although we are primarily interested in the former.
+
+\brk
+
+We end with an example.
+
+Look at $S_n$, the symmetry group on $n$ letters acting on $\R^n$ by coordinate permutations. Any coordinate flip, 
+say the $i^{th}$ and $j^{th}$ coordinates, is the reflection $S_{e_i-e_j}$. These coordinate flips generate all of $S_n$, 
+so $S_n$ is a Weyl group.
+
+\newpage
+
+\section*{Homework 2 -- Due Feb 1}
+%4.5; 7.4,6,7; board problem: show that a semisimple Lie algebra is generated as a Lie algebra by two elements
+\begin{hwprob}{4.5}
+	If $x,y\in\End V$ commute, prove that $(x+y)_s=x_s+y_s$ and $(x+y)_n=x_n+y_n$.
+	Show by example that this can fail if $x,y$ fail to commute. [Show first that $x,y$ semisimple (nilpotent) implies their sum is.]
+\end{hwprob}
+\begin{sol}
+	Following the hint, we will first establish:
+	\begin{lem}
+		If $x$ and $y$ in $\End V$ commute and are semisimple (resp. nilpotent) then their sum $x+y$ is semisimple (resp. nilpotent).
+	\end{lem}
+	\begin{prf}
+		Assume first that $x$ and $y$ are semisimple. Then $\ad x$ and $ad y$ can both be diagonalized. In particular, since they commute
+		they can be diagonalized \textit{simultaneously} via some matrix $A$ to become diagonal matrices $D_x$ and $D_y$. But then their sum
+		conjugated by $A$ is
+		\[A[(\ad x)+(\ad y)]A^{-1}=A(\ad x)A^{-1}+A(\ad y)A^{-1}=D_x+D_y\]
+		which is clearly diagonal, thus $x+y$ itself is semisimple.
+
+		\brk
+
+		Assume now that $x$ and $y$ are nilpotent, with $(\ad x)^\alpha=(\ad y)^\beta=0$. Then since
+		\[[\ad x,\ad y]=\ad [xy]=\ad 0=0,\]
+		we have that $\ad x=X$ and $\ad y=Y$ commute. So consider 
+		\[(\ad(x+y))^{\alpha+\beta-1}=(X+Y)^{\alpha+\beta-1}=\sum_{i=0}^{\alpha+\beta-1}\binom{\alpha+\beta-1}{i}X^iY^{\alpha+\beta-1-i}=0\]
+		where the last equality follows since the degree of each monomial is $\alpha+\beta-1$, which either forces the power of $X$ to be greater than $\alpha$
+		or the power of $Y$ greater than $\beta$ (since if $i<\alpha,$ we get $\alpha+\beta-1-i>\beta-1$).
+
+		Thus $x+y$ is nilpotent as well, completing the proof of the lemma.
+	\end{prf}
+
+	But then the statement itself follows simply by the uniqueness of the decomposition of $x+y$ into semisimple and nilpotent components: since 
+	\[x+y=x_s+x_n+y_s+y_n=(x_s+y_s)+(x_n+y_n)\]
+	is such a decomposition of $x+y$, these are \textit{the} semisimple and nilpotent components.
+\end{sol}
+\newpage
+In the following three problems, let $L=\sl(2,k)$.
+\begin{hwprob}{7.4}
+	The irreducible representation of $L$ of highest weight $m$ can also be realized ``naturally''
+	as follows. Let $X,Y$ be a basis for the two dimensional vector space $k^2$, on which $L$ acts as usual. Let $\mathscr{R}=k[X,Y]$
+	be the polynomial algebra in two variables, and extend the action of $L$ to $\mathscr{R}$ by the derivation rule
+	$z\cdot fg=(z\cdot f)g+f(z\cdot g)$. Show that this extension is well defined and that $\mathscr{R}$ becomes an $L$ module. Then show that
+	the subspace of homogeneous polynomials of degree $m$, with basis $X^m,X^{m-1}Y,\dots,Y^m$ is invariant under $L$ and irreducible of highest weight $m$.
+\end{hwprob}
+\begin{sol}
+	We simply extend the action linearly both for $\mathscr{R}$ and $L$ -- that is
+	\[(\alpha x+\beta y) f=\alpha(x\cdot f)+\beta(y\cdot f)\]
+	and
+	\[x\cdot (\alpha f+ \beta g)=\alpha(x\cdot f)+\beta(x\cdot g)\]
+	so by the linearity of derivations (and the bracket in $L$), all we have to check is that the bracket acts the right way under the derivation rule:
+	\begin{align*}
+		[xy]\cdot fg &= (xy-yx)\cdot fg\\
+		&= xy \cdot fg - yx\cdot fg\\
+		&= (xy\cdot f)g+f(xy\cdot g)-(yx\cdot f)g-f(yx\cdot g)\\
+		&= (xy\cdot f - yx\cdot f)g + f(xy\cdot g - yx\cdot g)\\
+		&= ([xy]\cdot f)g +f([xy]\cdot g)
+	\end{align*}
+	so this does in fact define an action on $L$.
+
+	Since the action is defined to be $k$-linear, we need only check that the $L$ action preserves the degree of a monomial in $X$ and $Y$.
+	\[z\cdot X^\alpha Y^\beta = (z\cdot X^\alpha)Y^\beta+ X^\alpha(z\cdot Y^\beta)=\alpha(z\cdot X)X^{\alpha-1}Y^\beta+\beta(z\cdot Y)X^\alpha Y^{\beta-1}\]
+	but since $z\in L$, $z$ takes $X$ and $Y$ to some linear combination of themselves (perhaps zero!), so the total degree of each of the
+	above monomials is $\alpha+\beta$. Thus these elements span an $L$-module.
+
+	Using $A=(\begin{smallmatrix}0&0\\1 & 0\end{smallmatrix})$ and $B=(\begin{smallmatrix}0&1\\0 & 0\end{smallmatrix})$ in $L$,
+	we can get between any of these basis vectors since 
+	\[A\cdot X^aY^b=aX^{a-1}Y^{b+1}\quad\text{and}\quad B\cdot X^aY^b=bX^{a+1}Y^{b-1}\]
+	so this module is irreducible,
+
+	Finally by considering the action of $C=(\begin{smallmatrix}1&0\\0 & -1\end{smallmatrix})$ on the basis, we see 
+	\[C\cdot X^aY^b=aX^aY^b-bX^aY^b=(a-b)X^aY^b\]
+	so the weights are $a+b=m,m-2, \dots, 2-m, -m$, so this is of highest weight $m$.
+\end{sol}
+\begin{hwprob}{7.6}
+	Decompose the tensor product of the two $L$ modules $V(3)$ and $V(7)$ into the sum of irreducible submodules:
+	$V(4)\oplus V(6)\oplus V(8)\oplus V(10)$. Try to develop a general formula for the decomposition of $V(m)\otimes V(n)$.
+\end{hwprob}
+\begin{sol}
+	Using the fact that, if $v_i$ and $w_i$ are bases for $V(3)$ and $V(7)$, respectively, then $v_i\otimes w_j$ form a basis for $V(3)\otimes V(7)$ and that
+	\begin{align*}
+		h\cdot v_i\otimes v_j&=(h\cdot v_i)\otimes v_j+v_i\otimes(h\cdot v_j)\\
+		&=(3-2i)v_i\otimes v_j+(7-2j)v_i\otimes v_j\\
+		&=(10-2(i+j))v_i\otimes v_j
+	\end{align*}
+	Some quick analysis of the possible values $a_{ij}=10-2(i+j)$ can attain for $0\le i\le 3$ and $0\le j\le 7$
+	tells us $-10\le a_{ij}\le 10$. Counting the number of $(i,j)$ pairs giving us each weight yields the list
+	\begin{center}
+		\begin{tabular}{c|c|c|c|c|c}
+			0 & $\pm 2$ & $\pm 4$ & $\pm 6$ & $\pm 8$ & $\pm 10$\\\hline
+			4 & 4 & 4 & 3 & 2 & 1
+		\end{tabular}
+	\end{center}
+
+	Take $v_0\otimes w_0$, the basis element of highest weight in the tensor product. We have characterized irreducible $\sl_2$ modules,
+	so we know that the $\sl_2$-span of this element must be a copy of $V(10)$, consisting of the $k$-span of a single vector in each weight space.
+
+	I am certain one can find a vector $v$ in $L_8$ such that $y^9\cdot v=0$, but I had a difficult time finding one.
+	So rather than demonstrating that each of these simple modules whose sum (whence direct sum) is all of $V(3)\otimes V(7)$, I appeal
+	to theorem 5.2 in Humphreys (the Lie analogue of Maschke from group reps): since $V(10)$ is simple and we're over a characteristic zero field there is a submodules $W$ of $V(3)\otimes V(7)$ such that
+	\[V(3)\otimes V(7)=V(10)\oplus W.\]
+
+	Since $V(10)$ spans a one-dimensional subspace of each weight space, the dimension of each must decrease by one in $W$, giving us 
+	dimensions (for weight spaces of $W$):
+	\begin{center}
+		\begin{tabular}{c|c|c|c|c|c}
+			0 & $\pm 2$ & $\pm 4$ & $\pm 6$ & $\pm 8$ & $\pm 10$\\\hline
+			3 & 3 & 3 & 2 & 1 & 0
+		\end{tabular}
+	\end{center}
+	Then we can pick any nonzero vector in the highest-weight space $L_8$ and under $\sl_2$ we will generate a copy of $V(8)$.
+	Iterating this process we get a copy of $V(6)$ and $V(4)$ until (by dimension considerations, if you like), the process terminates and
+	\[V(3)\otimes V(7)=V(10)\oplus V(8)\oplus V(6)\oplus V(4).\]
+
+	\brk
+
+	After gaining some intuition for how these decompositions work in practice, I believe this general theory holds:
+	\begin{conj}
+		Let $m$ and $n$ be integers and let $V=V(m)\otimes V(n)$. Without loss of generality, assume that $m\le n$.
+		Then 
+		\[V\cong V(m+n)\oplus\cdots\oplus V(m+n-2\cdot k)\oplus\cdots\oplus V(n-m)\]
+	\end{conj}
+\end{sol}
+\begin{hwprob}{7.7}
+	In this exercise, we construct certain infinite dimensional $L$-modules. Let $\lambda\in k$ be
+	an arbitrary scalar. Let $Z(\lambda)$ be the vector space over $k$ with countably-infinite basis $(v_0,v_1,\dots)$.
+	\begin{enumerate}
+		\item Prove that the formulas (a)-(c) of Lemma 7.2 define an $L$-module structure on $Z(\lambda)$, and that every nonzero $L$-submodule of $Z(\lambda)$
+		contains at least one maximal vector.
+		\item Suppose $\lambda+1=i$ is a nonnegative integer. Prove that $v_i$ is a maximal weight vector (e.g. $\lambda=-1$, $i=0$).
+		This induces an $L$-module homomorphism $Z(\mu)\xrightarrow{\varphi}Z(\lambda)$ for $\mu=\lambda-2i$ sending $v_0\mapsto v_i$.
+		Show that $\varphi$ is a monomorphism, and that $\im\varphi$ and $Z(\lambda)/\im\varphi$ are both irreducible $L$-modules
+		(but $Z(\lambda)$ fails to be completely reducible when $i>0$).
+		\item Suppose $\lambda+1$ is not a nonnegative integer. Prove that $Z(\lambda)$ is irreducible.
+	\end{enumerate}
+\end{hwprob}
+\begin{sol}
+	\textbf{Part (a):}
+
+	To see that these relations define an $L$-module structure, we can rely on the fact that all vectors are
+	\textit{finite} linear combinations of the $v_i$. This enables us to look ``locally'' at the $L$-module structure to confirm it 
+	satisfies all required properties on a case-by-case basis.
+
+	More specifically, for any $g\in L$ and any element $\sum_i c_i v_i\in Z(\lambda)$, take 
+	\[S=\{i,g(i,k)|c_i,a^i_k\ne 0\}\] 
+	where we define $g(i, k)$ and $a^i_k\in k$ to be such that 
+	\[g\cdot v_i=\sum_k a^i_k v_{g(i,k)}\]
+
+	Then $|S|<\infty$, so we can take $N=\max S$ and all $v_i$ that appear in the computation $g\cdot \sum c_i v_i$ are contained in 
+	$V(N)$, so the $L$ structure follows from the $L$ structure there.
+
+	\brk
+
+	To see the last claim, notice that if $L'$ is a nonzero submodule of $Z(\lambda)$, then if $0\ne v_i\in L'$ then 
+	since $x^{i+1}\cdot v_i=0$, there is some minimal $j$ such that $x^{j}\cdot v_i=0$. Then $v_{i-j+1}$ is a maximal
+	vector.
+
+	Furthermore examining the structure of $L'$, we get that we get $v_k\in L'$ for all $k\ge i$, which follows
+	trivially from $y\cdot v_i=(i+1)v_{i+1}$.
+
+	\textbf{Part (b):}
+
+	Let $i=\lambda+1$ for some scalar $\lambda\in k$. Then by applying $x$, we get
+	\[x\cdot v_i=(\lambda-i+1)v_{i-1}=(i-1-i+1)v_{i-1}=0\]
+	so $v_i$ is a maximal weight vector in $Z(\lambda)$.
+
+	Define $\varphi:Z(\lambda-2i)\to Z(\lambda)$ to be the map sending $v_0\to v_i$. Then we get
+	\[\varphi(v_{1})=\varphi\left(y\cdot v_0\right)=y\cdot\varphi(v_i)=y\cdot v_i=(i+1)v_{i+1}\]
+	and more generally 
+	\begin{align*}
+		\varphi(v_k)&=\varphi\left(\frac{1}{k!}y^k\cdot v_0\right)\\
+		&=\frac{1}{k!}y^k\cdot v_i\\
+		&=\frac{(i+1)(i+2)\cdots (i+k)}{k!} v_{i+k}\\
+		&=\frac{(i+k)!}{i!k!}v_{i+k}\\
+		&=\binom{i+k}{k}v_{i+k}
+	\end{align*}
+
+	But then the $\varphi(v_j)$ form a basis for the image (clearly) so if 
+	\[\varphi\left(\sum c_k v_k\right)=\sum c_k\varphi(v_k)=0\]
+	this implies that all $c_k=0$, so $\varphi$ in a monomorphism.
+
+	Notice that $Z(\lambda)/\im\varphi\cong V(i-1)$, so is irreducible. Now since $\im\varphi\cong Z(\lambda-2i)$
+	the problem reduces to showing the latter is irreducible. From (a) we know that any nonzero submodule of $Z(\mu)$ has a maximal vector $v_j$.
+	Consider that in $Z(\mu)$
+	\[x\cdot v_j=(\lambda-2i-j+1)v_{j-1}=0\]
+	thus either $v_{j-1}=0$ (whence $j=0$ and this submodule is all over $Z(\mu)$) or else
+	\[\lambda=2i+j-1=i-1\quad\Rightarrow\quad i+j=0\]
+	but $v_{-i}=0$ for positive $i$, so this forces $i=j=0$, so the submodule is the entire module.
+
+	\textbf{Part (c):}
+
+	Assume now that $\lambda+1$ is not a nonnegative integer. Similar to above, any nonzero submodule has a maximum vector $v_j$.
+	Then the equation we saw before
+	\[2i+j-1=i-1\]
+	has no solution, which forces $v_{j-1}=0$ and so $j=0$ and so the submodule is all of $Z(\lambda)$.
+\end{sol}
+\begin{hwprob}{(Board)}
+	Show that a semisimple Lie algebra is generated (as a Lie algebra) by two elements.
+\end{hwprob}
+\begin{sol}
+	Consider the decomposition (as a vector space) of any semisimple Lie algebra $\g$ with Cartan subalgebra $\h$ 
+	into the sum of its root spaces
+	\[\g = \h\oplus \bigoplus_{\alpha\in\Phi}\g_\alpha\]
+	and chose $h\in\h$ and from each $\g_\alpha$ choose a nonzero $v_\alpha$. Define $v =h+\sum_{\alpha\in\Phi}v_\alpha$. I claim that 
+	this vector, along with another vector generates everything. I'll keep thinking about this. :)
+\end{sol}
+\newpage
+
+\section{February 1, 2019}
+\begin{prob}
+	This assignment's board problem: Show, using large tensor powers of the adjoint representation and $\sl_2$ theory,
+	that any semisimple Lie algebra $L$ has irreducible modules of arbitrarily high degree.
+\end{prob}
+
+\subsection{Revisiting classical Lie Algebras}
+In all cases the diagonal matrices form a maximal toral subalgebra with roots given by (relative the standard basis $e_i$ for $\R^n$):
+\begin{itemize}
+	\item $A_{n-1}$ is spanned by $e_i-e_j$ for $i\le i,j\le n$
+	\item $B_{n}$ is spanned by $e_i-e_j$ and $e_i+e_j$ as well as $e_i$
+	\item $C_{n}$ is spanned by $e_i-e_j$ and $e_i+e_j$ as well as $2ei$
+	\item $D_{n}$ is spanned by (only) $e_i-e_j$ and $e_i+e_j$.
+\end{itemize}
+
+We can check in all cases that the axioms of a root system are satisfied: the differences, sums, and monomials act
+on the roots by flipping, flipping and changing the sign, and just changing the sign, respectively.
+In all cases, the root system is obtained from a lattice ($\Z$-span of an $\R$-basis) as follows:
+\begin{itemize}
+	\item $A_{n-1}$ the vector space is spanned by $e_1-e_2,\dots$ and the lattice is generated by the same vectors.
+	Then the root system is realized as the set of such vectors of square length 2.
+	\item $B_n$ we take the $\Z^n\subseteq\R^n$ vectors of square length 1 or 2.
+	\item $D^n$ is the same, but just square length two.
+	\item $C_n$ is a bit sketchy in terms of lengths -- we can't quite say all vectors of length 2 or 4. So instead
+	we derive it from $B_n$ by replacing all $e_i$ by $2e_i$.
+\end{itemize}
+What are the Weyl groups of these sweet babies?
+\begin{itemize}
+	\item $A_{n-1}$: $S_n$ action on $\R^n$ by coordinate permutations. This is also the symmetry group of the unit simplex in $\R^n$.
+	\item In $B_n$ and $C_n$ (reflections act the same way irrespective of magnitude) is the group of all permutations and sign changes 
+	of the coordinates in $\R^n$. This can also be recognized as the hypercube or hyperoctahedron in $\R^n$ with vertices $\pm e_i$. This is 
+	sometimes called the hyperoctahedral group.
+	\item For $D_n$, this consists of all permutations \textit{even} sign changes of coordinates in $\R^n$. This has no 
+	standard name. This can also be viewed as the group of symmetries of a polytope whose vertices are a subset of 
+	those of a hypercube, but with \textit{evenly-many} negative. Again, there is no consensus about what to call these.
+\end{itemize}
+Later we will see the exceptional root systems (non-classical) and again they are taken from a lattice by taking all vectors of one or two specified square lengths.
+\subsection{Some Non-Crystallographic Examples}
+Here there are some $\alpha$ and $\beta$ where $\frac{2(\alpha,\beta)}{(\alpha,\alpha)}\notin\Z$.
+\begin{ex}[$I_2(m)$]
+	Start off with a regular $m$-gon in $\R^2$. The dihedral group of symmetries of this $m$-gon is not a Weyl group, 
+	but it is a generalization: a \textbf{Coxeter group}: a finite group generated by reflections.
+
+	To find the roots: if $m$ is odd, place the $m$-gon so its center is at $(0,0)$ and take $\pm v$ as $v$ runs over the vertices. If, instead,
+	$m$ is even, center the $m$-gon at $(0,0)$ and take $\pm v$ for vertices $v$ along with $\pm 2m$, where the $m$ are 
+	midpoints of edges.
+
+	Why $2m$? Well in some special cases ($m=3,4,6$), this actually becomes a crystallographic system -- but not otherwise.
+
+	There is also a degenerate case when $m=2$. Then the group is $S_2\times S_2$, generated by two orthogonal reflections.
+
+	In all cases, the dihedral group is generated by just \textbf{two} reflections, corresponding to two roots which make as large an angle as possible.
+\end{ex}
+
+In general, the root system described in the above example is denoted $I_2(m)$ and the corresponding Coxeter group is
+$D_m$, the dihedral group on an $m$-gon.
+
+\subsection{Exceptional Root Systems}
+\begin{ex}[$E_8$]
+	The biggest such system is $E_8$. We begin with vector space $\R^8$ and a lattice spanned by all $e_i\pm e_j$
+	as well as $(\frac{1}{2},\cdots,\frac{1}{2})$ and take all vectors in the lattice of square length 2.
+
+	The lattice consists of all $(a_1,\dots,a_8$ where all $a_i\in\Z$ OR all $a_i\in \Z+\frac{1}{2}$ AND $\sum a_i\in 2\Z$.
+
+	The roots are $e_i\pm e_j$ and vectors of the form $(\pm\frac{1}{2},\cdots,\pm\frac{1}{2})$ with \textbf{evenly many} $-\frac{1}{2}$. 
+	There are 240 of them.
+\end{ex}
+Then we can define:
+\begin{itemize}
+	\item $E_7$ to be the set of vectors in $E_8$ orthogonal to $e_7+e_8$ (120 roots)
+	\item $E_6$ to be the set of vectors in $E_8$ orthogonal to $e_7+e_8$ and $e_6+e_8$ (72 roots).
+\end{itemize}
+
+\begin{ex}[$F_4$]
+	Vector space $\R^4$, with lattice spanned by $e_1,e_2,e_3,$ and $\frac{1}{2}(1,1,1,1)$.
+	Then we get all $(a_i)\in\R^4$ with all $a_i\in\Z$ or all $a_i\in \Z+\frac{1}{2}$. Then we take all vectors of square length one or two.
+
+	We get $\pm e_i$, $\pm(e_i\pm e_j)$ and $(\pm\frac{1}{2},\cdots,\pm\frac{1}{2})$ to get a total of 48 roots.
+\end{ex}
+\begin{ex}[$G_2=I_2(6)$]
+	Vectors space is the one spanned by $e_1-e_2$ and $e_2-e_3$ in $\R^3$ and lattice is spanned by these and the roots 
+	consist of these vectors as well as all vectors of square length 2 or six:
+	\[(\pm (-2,1,1),\pm (1,-2,1),\pm (1,1,-2).\]
+\end{ex}
+\subsection{Positive and Negative Roots}
+Given a root system $\Phi$, we can cut it in half by realizing $\R^n$ as an ordered space by saying $(v_1,\dots,v_n)$ is positive
+if and only if the smallest index $i$ with $v_i\ne 0$ has $v_i>0$. Then the sum of two positive vectors is positive, as is a positive multiple.
+
+Then define
+\begin{defn}[Positive Roots]
+	Set $\Phi^+=\Phi\cap\{\text{positive vectors}\}$. Call the roots in $\Phi^+$ and call $\Phi^+$ a \textbf{positive subsystem.}
+\end{defn}
+
+\section{February 6, 2019}
+Recall from last itme that we were letting $\Phi$ be a root system with positive and negative roots $\Phi^+$ and $\Phi^-$ such that they partition $\Phi$,
+are disjoint, and are closed under positive linear combinations.
+
+Next we want ot exhibit a basis for $\R^n$, the ambient vector space, consising of positive roots such that
+every root in $\Phi^+$ is a \textbf{positive} combination of basis vectors.
+
+To do this, we already know that $\Phi^+$ has subsets $S$ (e.g. the whole thing) such that every root in $\Phi^+$ is a positive combination of things in $S$.
+Let $\Delta$ be a minimal such subset under inclusion. I am having a slow morning -- take a look at the book. :)
+
+\brk
+
+The next claim is that $\Delta$ is independent. Otherwise, we have the relation $\sum_{\alpha\in\Delta}r_\alpha \alpha=0$ with
+at least one nonzero $r_\alpha$. Moving all the negative coefficients over to the other side, we get the equality of two
+(possibly empty) sums. If either side is trivial, this is a contradiction since $0\notin\Phi^+$.
+
+If both are nonempty, then you can use the positivity of the inner product of one side with some $\beta$,
+which then implies that the inner product of the other side with $\beta$ is negative -- again a contradiction.
+
+\begin{defn}
+	We call such a $\Delta$ a \textbf{simple subsystem} of $\Phi$.
+\end{defn}
+
+\begin{rmk}
+	Notice that $\Delta$ is uniquely determined -- it consists exactly of those $\alpha\in\Phi^+$ that are \textit{not} positice combinations of two or more positive roots.
+
+	Pick some $\alpha\in\Delta$, a simle root. The corresponding reflection $s_\alpha$ is also called simple. Then for any $\beta\in\Phi^+$, 
+	$s_\alpha\beta$ is positive unless $\beta=\alpha$, as it must involve some simple root with a positive coefficent, which stays the same upon applying $s_\alpha$.
+
+	Now let $\Phi^+$ and $(\Phi^+)'$ be two choies of positice roots, corresponding to two choices of positive vecors in $\R^n$. Let $\Delta,\Delta'$ be the simple subsystems
+	corresponding two these two systems.
+
+	If $\Phi^+\ne(\Phi^+)'$, there is some \textbf{simple} $\alpha\in\Delta$ which is \textbf{negative} in $(\Phi^+)'$. Applying $S_\alpha$ to $\Phi^+$, we get another positive
+	system which contains the same elements of $(\Phi^+)'$ that $\Phi^+$ did, plus now $-\alpha$.
+
+	Doing this repeatedly, we can bump up the size of the intersection $\Phi^+\cap(\Phi^+)'$ at each step, eventually yielding
+	two equal positive sytems.
+
+	So any two positive systems are conjugate under a product of simple reflections with respect to any system. That product is 
+	an orthogonal transformation, so the angles between all pairs of simple roots don't depend on the choice of simple roots.
+
+	Finally, if $\beta$ is a positive root, but not simple, then since it is a positive linear combination of simple roots, 
+	$(\beta,\alpha)>0$ for some simple $\alpha$. Then $s_\alpha\beta=\beta-c\alpha$ for some $c>0$ is another positive root.
+	Continuing this process must end with a simple root, so any positive root is a product of simple reflections applies to a simple root (in fact, any root)
+
+	So the \textbf{entire root system} is completely determined by a knowledge of the square length of the simple roots and the angles between them.
+\end{rmk}
+
+\begin{rmk}
+	In the crystallographic case, we can say more! Here
+	\[s_\alpha\beta=\beta=\frac{2(\beta,\alpha)}{(\alpha,\alpha)}\alpha\]
+	where the coefficient is in $\Z$! Here any positive root is a \textbf{positive integer combination} of simple roots.
+\end{rmk}
+
+\begin{rmk}
+	Finally, what possible angles can occur between pairs of simple roots? If $\alpha,\beta$ are simple roots with simple reflections $s_\alpha,s_\beta$, then it must be that $s_\alpha s_\beta$
+	must be an element of finite order lying in a finite group. Thus the angle must be a rational multiple of $2\pi$.
+
+	In orderr to capture all positive roots that are combinations of $\alpha$ and $\beta$ that are \textbf{positive} combinations, the angle between $\alpha$ and $\beta$ msut be as wide as possible:
+	$\pi-\frac{\pi}{m}$ for some $m$ (the order) if $\alpha\ne c\beta$.
+
+	In the Crystallographic case, we get that $m=2,3,4,$ or $6$.
+\end{rmk}
+\subsection{Dynkin Diagrams}
+Now we construct the \textbf{Dyinkin diagram} of our root system, if it is crystallographic, or its
+\textbf{Coxeter graph} in general.
+
+This is a graph in which the vertices correspond to simple roots $\alpha$. The vertices corresponding to $\alpha$ and $\beta$
+are joined by an edge if $\alpha\not\perp\beta$. This splits into cases:
+\begin{itemize}
+	\item $m=3$, we take a single edge.
+	\item $m=4$, take a double edge with arrow pointing toward the shorter root.
+	\item $m=6$, we take a triple edge with the same condition as above.
+\end{itemize}
+
+In a Coxeter graph, we do the same except there is just one edge, labeled with $m$.
+
+\begin{ex}
+	Recall in the classical cases that a root is positive if its leftmost nonzero coordinate is positive. Then for $A_{n-1}$:
+	\[\]
+	I want to fill these all in later. For now, just including the notes:
+
+	Notice that the diagrams for $B_2$ and $C_2$ are graph isomorphic (consisting of a single double edge!). This 
+	tells us that, in fact, the Lie algebras are isomorphic. This relates one of the spin groups to one of the special linear groups although i forget which.
+\end{ex}
+
+\section{February 8, 2019}
+Continuing from last time, we want to write down the simple roots that give rise to the Dynkin diagrams 
+for the exceptional Lie algebras we saw (and I just realized I haven't copied in yet.)
+
+\subsection{Exceptional Lie Algebra Dynkin Diagrams}
+For $E_8$he roots are in the images I took. :)
+
+For $E_7$ and $E_6$ you just remove the roots that we talked about last time.
+
+For $H_3$ we take as the simple roots the edge midpoints of a regular dodecahedron or icosahedron.
+
+For $H_4$, take the face centers of a 120-sided regular polytope in $\R^4$, called the hecatonicosahedroid.
+
+\subsection{Have we found them all?}
+Now we want ot see that htis list of connected Dynking diagrams is complete! 
+\begin{rmk}
+	We say ``connected'' because if $\Phi$ and $\Phi'$ are two root systems in $\R^n$ and $\R^m$,
+	then $\Phi\sqcup\Phi'$ is a root system in $\R^{n+m}$, the orthogonal direct sum. This gives two disjoint
+	Dynkin diagrams (since they are orthogonal) so we restrict to the connected components.
+\end{rmk}
+
+We are actually going to classify the \textit{Coxeter graphs} corresponding to (crystallographic) Dynkin diagrams
+starting with a simple subsystem of a crystallographic root system. We obtain a basis of $n$ unit vectors in $\R^n$ such that hte angle between any 
+two of them is $\pi-\frac{\pi}{m}$ for $m\in\{2,3,4,6\}$, encoded by a coxeter graph where there is no edge if $m=2$, unlabelled if $m=3$ and unlabelled otherwise.
+
+\subsection{Classifying Coxeter Graphs}
+To classify these, we actually classify the graphs that are \textbf{not} of this type. That is, for each connected Coxeter graph
+that we've seen so far, we are going to attach a slightly larger graph to it that \textit{doesn't occur} as a Coxeter graph of any root system
+or, in fact, even a \textit{subgraph} of a Coxeter graph.
+
+By section 10 of Humphreys, every connected Dynkin diagram and root system $\Phi$ corresponds to a unique lowest (highest) root.
+Then every other root, if the lowest is subtracted, becomes a positive combination of simple roots and has nonpositive dot
+product with the simple roots.
+
+The resulting ``Dynkin diagram'' (it won't be) corresponding to the simple roots plus the lowest root 
+comes from a \textbf{dependent} set of vectors -- in fact, one admitting a dependence relation with all
+positive coefficients. It is not a Dynkin diagram since if it were, that combination of roots would have square length zero,
+but this is impossible for any Dynkin diagram or subdiagram thereof. This is true \textit{even if some of the edge labels are increased.}
+
+\subsubsection{What do we get when we do this?}
+We call these the \textbf{extended} or \textbf{affine} Coxeter graphs.
+
+For $\tilde A_{n-1}$ our lowest root is $e_n-e_1$, which attaches to both $e_1-e_2$ and $e_{n-1}-e_n$ and creates a cycle.
+
+For $\tilde B_{n}$, we add $-e_1-e_2$ which links with the secondmost left root in $B_n$.
+
+For $\tilde C_n$, we append $-2e_1$ to the leftmost root.
+
+For $\tilde D_n$, we append $-e_1-e_2$ to the secondmost left root as in $\tilde B_n$.
+
+\brk
+
+For the exceptional algebras: $\tilde E_6$ gets $\frac{1}{2}(-1,\cdots,-1,1,1,-1)$. I got a picture of the rest.
+
+The argument here can be found in Humphreys' \textit{Reflection Groups and Coxeter Groups} on page 37 that, once the forbidding configurations
+above have been ruled out, the only connected diagrams left are the ones we've seen.
+
+As a sketch, $B_n$ and $C_n$ are easy to see that the crystallographic property allows only two 
+possibilities: that the rightmost simple root has square length \textbf{twice or half} the length of the other simple roots, but no other multiples.
+
+In the type $E$ cases, we have one central vertex and three chains of lengths $p,q,$ and $r$ coming from it.
+Let's consider the posibilities we get:
+\begin{itemize}
+	\item $\tilde E_6$: $(3,3,3)$
+	\item $\tilde E_7$: $(2,4,4)$
+	\item $\tilde E_8$: $(2,3,6)$
+\end{itemize}
+Notice that in all three cases the sum of the reciprocals of these values is 1! In fact, these are the only 
+solutions to this this in the integers!
+
+But then in the regular Dynkin diagrams:
+\begin{itemize}
+	\item $E_6$: $(3,3,3)$
+	\item $E_7$: $(2,4,4)$
+	\item $E_8$: $(2,3,6)$
+\end{itemize}
+Here we have the only solutions where the sum of the reciprocals are greater than 1 (with $p,q,r>1$). 
+Well we actually have the case $(2,2,n)$, but these correspond to $D_{n+2}$.
+
+In non-Euclidean geometry, this condition on $(p,q,r)$ arises all the time. The sum of reciprocals plays a role in the geometry of triangles.
+
+\subsection{Finite subgroups of $SO(3,\R)$}
+We actually played the $\frac{1}{p}+\frac{1}{q}+\frac{1}{r}>1$ game before in algebra!
+\begin{itemize}
+	\item $(2,2,n)$ is the group $D_n$
+	\item $(2,3,3)$ is $T\cong A_4$
+	\item $(2,3,4)$ is $O\cong S_4$
+	\item $(2,3,5)$ is $I\cong A_5$, the icosahedral group.
+\end{itemize}
+
+The icosahedrom (or dodecahedron) pops up in the non-crystallographic Coxeter groups:
+\begin{itemize}
+	\item $H_3$ is all symmatries of an icosahedron or of a dodecahedron in $O(3,\R)$. This group is $A_5\times\Z_2$
+	\item $H_4$ is a crazy picture, but the group is
+	\[\frac{(A_5\times \Z_2)^2\ltimes \Z_2}{\langle \pm 1\rangle}\]
+	which we call the Hecatonicosahedroidal group. :)
+\end{itemize}
+
+\section{February 13, 2019}
+One final remark about the inequalities that came up: $\frac{1}{p}+\frac{1}{q}+\frac{1}{r}>1$ and $<1$ for $p,q,r\ge 2$.
+
+\begin{rmk}
+	if $\frac{1}{p}+\frac{1}{q}+\frac{1}{r}<1$, then the closest to 1 it can come is $\frac{1}{42}=\frac{1}{2}+\frac{1}{3}+\frac{1}{7}=1-\frac{1}{42}$.
+	This actually comes up in enumerative algebraic geometry for exactly this reason. Of course, it is also the
+	answer to the question of life, the universe, and everything as well. :)
+\end{rmk}
+
+We can also work out the coefficents of all simple roots in the highest root quite easily by induction:
+
+Start with the extended Dynkin diagram: place the label 1 on the new vertex (highest root) and label
+the others so that the corresponding combination of vectors has square length zero. This amounts to ensuring
+that the sum of the labels of all vertices adjacent to agive one is twice the label of that vertex, where labels are counted doubly/triply
+if they lie at the left end of a double/triple arrow.
+
+In $\tilde A_{n-1}$, this forces labels all to be ones. In $\tilde B_n$, we get 1's at the ``forked'' end 
+and twos elsewhere. $\tilde C_n$ has ones at the endpoints and twos elsewhere. $\tilde D_n$ has ones at the ends of both forks.
+You can keep doing this for the exceptional ones but I will have to take a picture because they are hard to transcribe.
+
+\brk
+
+Now we go off ``in a completely different direction'': Let $\Phi$ e a crystallographic root system. Then there are two lattices attached to $\Phi\in\R^n$:
+one is the \textbf{root lattice} spanned over $\Z$ by the roots -- denoted $R$. Another lattice is the \textbf{weight lattice} $P$
+consisting of all $\lambda\in\R^n$ such that $\langle\lambda,\alpha\rangle=2(\lambda,\alpha)/(\alpha,\alpha)\in\Z$ for all $\alpha\in\Phi$.
+
+Then $P$ and $R$ are free abelian groups of rank $n$. $P$ is generated by the \textbf{fundamental dominant weights} $\lambda_\alpha$
+defined by $\langle\lambda_\alpha,\alpha\rangle = 1$ and $\langle\lambda_\alpha,\beta\rangle=0$
+for $\alpha\ne\beta\in\Delta$.
+
+Hence $P/R$ is a finite abeliang group, called the \textbf{fundamental group of $\Phi$}. Its importance is the following:
+Complex semisimple Lie algebras are classified by their root systems. But what about complex semisimple Lie groups?
+These are classified by their root system \textit{plus} a lattice lying between the root and weight lattices. Taking the root lattice 
+leads to the adjoind group $\Int L$ (the smallest possible choice) if we take the weight lattice, then we get the simply-connected 
+universal covering space of $\Int L$.
+
+\begin{ex}
+	For type $A_{n-1}$, $\Int \sl_n(\C)=PSL(n,\C)$. The simply-connected cover is $SL(n,\C)$ itself.
+\end{ex}
+
+\begin{rmk}
+	When $k$ is not algebraically closed, we end up getting more forms to work with, but you can follow
+	a similar technique to classifying the Lie groups. In positive characteristic things blow up pretty badly.
+\end{rmk}
+
+\brk
+
+What did that have to do with the coefficients of the simple roots in the highest root? Well, in Humphreys exercise 13.13,
+we see that every coset has a unique representative that is minimal in the sense that $\langle\lambda,\alpha\rangle=0$ or $1$ for all positive roots
+$\alpha.$
+
+It turns out that $\Phi^v$ (see Humphreys) is another root system called the \textbf{dual} to $\Phi$. Then if 
+$\Delta$ is a simple subsystem for $\Phi$, then $\Delta v$ is a simple subsustem for $\Phi^v$. Then the condition
+that $\langle\lambda,\alpha\rangle\in\Z$ is linear in $\alpha^v$.
+
+\subsection{Takeaway}
+We got a bit in the weeds for me, but the takeaway is that the fundamental groups are as follows:
+\begin{itemize}
+	\item $A_{n-1}$: $\Z/(n)$
+	\item $B_n$: $\Z/(2)$
+	\item $c_n$: $\Z/(2)$
+	\item $D_n$: $\Z/(4)$ if $n$ is odd; $\Z_2\times\Z_2$ otherwise
+	\item $E_6$: $\Z/(3)$
+	\item $E_7$: $\Z/(2)$
+	\item $E_8,F_4,G_2$: $\Z$
+\end{itemize}
+
+\begin{rmk}
+	One more cool calculation (explained better in Humphreys' 1990 book): We can compute the order $|W|$:
+	look at the orbit and stabilizer of the highest roots under the action of $W$:
+
+	The orbit consists of all roots with the same length as the heighest weight: the long roots.
+
+	The stabilizer (discussed in Humphreys \S 10) is generated by all reflections fixing the highest root, corresponding to the roots \textbf{not}
+	adjacent to the longest root in the extended diagram.
+\end{rmk}
+
+We have a final remark on root systems before returning to the land of Lie algebras:
+\begin{rmk}
+	Given a positive root $\beta$ in a root system $\Phi$ , we have seen that (if $\Delta$ is the corresponding simple subsystem)
+	there is a product of simple reflections (attached to simple roots) sending $\beta$ to $\gamma$, a simple root.
+
+	Thus there is a concatenated string
+	\[\beta,\beta-\alpha,\cdots,\beta-\langle\beta,\alpha\rangle\alpha,\beta-\langle\beta,\alpha\rangle\alpha-\alpha',\cdots\]
+	in which all terms of roots and the last root is simple. But then we can reverse this string to deduce that
+	$\beta=\alpha_1+\alpha_2+\cdots+\alpha_m$ where all $\alpha_i$ are simple and every partial sum is a positive root.
+
+	What we get from this, in terms of Lie algebras, is that one generator from each of the positive root spaces generates all the positive root spaces.
+\end{rmk}
 \end{document}

### 2019-02-13 Reading Notes/reading_notes.tex
I should push more frequently

--- 2019-01-10 Reading Notes/reading_notes.tex
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% Package Includes
 \usepackage{setspace}
-
-\usepackage{amsmath, amsfonts, amssymb, graphicx, color, fancyhdr, lipsum, scalerel, stackengine, mathrsfs, tikz-cd, mdframed, enumitem, framed, adjustbox, bm, upgreek, x	color}
+\usepackage{amsmath, amsfonts, amssymb, graphicx, color, fancyhdr, scalerel, mathrsfs, tikz-cd, mdframed, enumitem, framed, adjustbox, bm, upgreek, xcolor}
 \usepackage[framed,thmmarks]{ntheorem}
@@
 
-%set up theorem/definition/etc envs
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%% Theorems, definitions, problems, solutions, etc
 %Problems will be created using their own counter and style
@@
 
-%solution template
+%solution
 \theoremstyle{nonumberbreak}
@@
 
-%Theorems, Lemmas, and Corollaries
+%Theorems, Lemmas, Corollaries, and Conjectures
 \theoremstyle{changebreak}
@@
 \theoremheaderfont{\color{blue}\bfseries} 
-
 \newtheorem{thm}{Theorem}[subsection]
@@
 \newtheorem{cor}[thm]{Corollary}
+\newtheorem{conj}[thm]{Conjecture}
 
@@
 
-\setlength\fboxsep{15pt}
+%Man, that's really good! Let's use the same thing for definitions.
+\newenvironment{def-prime}[1]
+  {\renewcommand{\thethm}{\ref{#1}$'$}%
+   \addtocounter{thm}{-1}%
+   \begin{def}}
+  {\end{def}}
 
@@
 \newshadedtheorem{defn}[thm]{Definition}
-
-%Man, that's really good! Let's use the same thing for definitons.
-\newenvironment{def-prime}[1]
-  {\renewcommand{\thethm}{\ref{#1}$'$}%
-   \addtocounter{thm}{-1}%
-   \begin{def}}
-  {\end{def}}
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% MATH DEFINITIONS
+%Blackboard Sets
+\newcommand*{\bbZ}{\mathbb{Z}}
+\newcommand*{\bbN}{\mathbb{N}}
+\newcommand*{\bbR}{\mathbb{R}}
+\newcommand*{\bbC}{\mathbb{C}}
+\newcommand*{\bbF}{\mathbb{F}}
+\newcommand*{\bbQ}{\mathbb{Q}}
+
+%Homological Algebra homological 
+\DeclareMathOperator{\im}{Im}
+\DeclareMathOperator{\Hom}{Hom}
+\DeclareMathOperator{\Ext}{Ext}
+\DeclareMathOperator{\HH}{HH}
+
+%Representations
+\DeclareMathOperator{\Res}{Res_H}
+\DeclareMathOperator{\Ind}{Ind_H^G}
+\DeclareMathOperator{\Soc}{Soc}
+
+%schemes
+\newcommand{\Ga}{\ensuremath\mathbf{G_a}}
+\newcommand{\Gm}{\ensuremath\mathbf{G_m}}
+\newcommand*{\GL}[1]{\ensuremath\mathbf{GL_#1}}
+\newcommand*{\SL}[1]{\ensuremath\mathbf{SL_#1}}
+\newcommand*{\mun}[1]{\ensuremath\mathbf{\bm\upmu_#1}}
+\newcommand*{\alp}[1]{\ensuremath\mathbf{\bm\upalpha_#1}}
+
+%Lie Algebras
+\newcommand*{\g}{\mathfrak{g}}
+\newcommand*{\n}{\mathfrak{n}}
+\newcommand*{\h}{\mathfrak{h}}
+\let\sl\relax
+\newcommand*{\sl}{\mathfrak{sl}}
+\newcommand*{\gl}{\mathfrak{gl}}
+\DeclareMathOperator{\ad}{ad}
+\DeclareMathOperator{\Int}{Int}
+\DeclareMathOperator{\Der}{Der}
+
+%categories
+\newcommand*{\Mod}{\text{\textbf{Mod}}}
+\newcommand*{\C}{\mathcal{C}}
+\newcommand*{\T}{\mathcal{T}}
+\let\O\relax
+\newcommand*{\O}{\mathcal{O}}
+\newcommand*{\Ass}{\text{\textbf{Ass}}}
+\newcommand*{\Vect}{\text{\textbf{Vect}}}
+\newcommand*{\Grp}{\text{\textbf{Grp}}}
+\newcommand*{\Ab}{\text{\textbf{Ab}}}
+\newcommand*{\Ring}{\text{\textbf{Ring}}}
+\newcommand*{\Lie}{\text{\textbf{Lie}}}
+
+%general commands
+\DeclareMathOperator{\ch}{char}
+\DeclareMathOperator{\Aut}{Aut}
+\DeclareMathOperator{\rank}{rank}
+\DeclareMathOperator{\id}{id}
+\DeclareMathOperator{\coker}{coker}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% Hacks, tweaks, and general helpful stuff
+%fix tilde
+\let\tilde\relax
+\newcommand*{\tilde}[1]{\widetilde{#1}}
+
+% Enumerate will automatically use letters (e.g. part a,b,c,...)
+\setenumerate[0]{label=(\alph*)}
+
+%Why is chi set so low?
+\DeclareRobustCommand{\Chi}{{\mathpalette\irchi\relax}}
+\newcommand{\irchi}[2]{\raisebox{\depth}{$#1\chi$}}
+
+%Center each figure by default
+\makeatletter
+\g@addto@macro\@floatboxreset{\centering}
+\makeatother
+
+\setlength\fboxsep{15pt}
+
+%Section break
+\newcommand*{\brk}{
+\rule{2in}{.1pt}
+}
+
 %Replacement for the old geometry package
@@
 
-%Center each figure by default
-\makeatletter
-\g@addto@macro\@floatboxreset{\centering}
-\makeatother
+%Get q-binomial coefficients
+\newcommand{\qbinom}{\genfrac{[}{]}{0pt}{}}
 
-%header stuff
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% Header Stuff
 \setlength{\headsep}{24pt}  % space between header and text
@@
 
-%Set of Integers
-\newcommand*{\Z}{
-\mathbb{Z}
-}
-%Set of Natural Numbers
-\newcommand*{\N}{
-\mathbb{N}
-}
-%Set of Real Numbers
-\newcommand*{\R}{
-\mathbb{R}
-}
-%Set of Complex Numbers
-\newcommand*{\C}{
-\mathbb{C}
-}
-%Field
-\newcommand*{\F}{
-\mathbb{F}
-}
-%Rationals
-\newcommand*{\Q}{
-\mathbb{Q}
-}
-
-%Section break
-\newcommand*{\brk}{
-\rule{2in}{.1pt}
-}
-
-\DeclareMathOperator{\Aut}{Aut}
-
-%raise that Chi!
-\DeclareRobustCommand{\Chi}{{\mathpalette\irchi\relax}}
-\newcommand{\irchi}[2]{\raisebox{\depth}{$#1\chi$}}
-
-%Image
-\DeclareMathOperator{\im}{Im}
-
-%Hom
-\DeclareMathOperator{\Hom}{Hom}
-
-%Ext
-\DeclareMathOperator{\Ext}{Ext}
-
-%Coker
-\DeclareMathOperator{\coker}{coker}
-
-%characteristic
-\DeclareMathOperator{\ch}{char}
-
-%restriction
-\DeclareMathOperator{\Res}{Res_H}
-
-%socle
-\DeclareMathOperator{\Soc}{Soc}
-
-%rank
-\DeclareMathOperator{\rank}{rank}
-
-%induction
-\DeclareMathOperator{\Ind}{Ind_H^G}
-
-%identity map
-\DeclareMathOperator{\id}{id}
-
-%schemes
-\DeclareMathOperator{\Ga}{\ensuremath\mathbf{G_a}}
-\DeclareMathOperator{\Gm}{\ensuremath\mathbf{G_m}}
-\newcommand*{\GL}[1]{\ensuremath\mathbf{GL_#1}}
-\newcommand*{\SL}[1]{\ensuremath\mathbf{SL_#1}}
-\newcommand*{\mun}[1]{\ensuremath\mathbf{\bm\upmu_#1}}
-\newcommand*{\alp}[1]{\ensuremath\mathbf{\bm\upalpha_#1}}
-
-%fix tilde
-\let\tilde\relax
-\newcommand*{\tilde}[1]{\widetilde{#1}}
-
-% Enumerate will automatically use letters (e.g. part a,b,c,...)
-\setenumerate[0]{label=(\alph*)}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 \maketitle
-
-%begin problems
 
@@
 	Let $F$ be such a functor. Consider any $\varphi\in\Hom(k\times k, R)\simeq F(R)$.
-	Assume first that $F(R)\cong \Z/2$ and let $r$ be an idempotent in $R$. 
+	Assume first that $F(R)\cong \bbZ/2$ and let $r$ be an idempotent in $R$. 
 \end{sol}
@@
 	\subsubsection*{$\alp{p}$:}
-	This time we are working with $A=\Z/p[X]/(X^p)$. This time (since the group is additive) we want
+	This time we are working with $A=\bbZ/p[X]/(X^p)$. This time (since the group is additive) we want
 	$\Delta(X)=X\otimes 1+1\otimes X$, which we can see works with associativity immediately.
@@
 \end{thm}
-\section{Affine Group Schemes (Waterhouse and Jantzen)	}
+
+\section{Overarching Ideas and Notes}
+\subsection{Week 2}
+Consider the representations of A over a field $k$ (or more generally $A$ modules, which we can make into a tensor category).
+Recall that in this category we have enough projectives and (with finitely generated Hopf algebras) we get that all projectives are injective.
+
+The object of study here:
+\begin{defn}
+	The \textbf{Hochschild cohomology} of a Hopf algebra $A$ over the field $k$ is 
+	\[\Ext_A^*(k,k)=H^*_A(A,k)\].
+\end{defn}
+
+The conjecture here is
+\begin{conj}
+	when $A$ is a finite dimensional Hopf algebra, $H^*(A,k)$ (the Hochschild cohomology) 
+	is finitely generated as a $k$-algebra.
+\end{conj}
+This is known in some special cases:
+\begin{itemize}
+	\item Finite groups
+	\item Finite group schemes (in positive characteristic) due to Friedlander and Suslin.
+	\item In characteristic zero we have
+	\begin{itemize}
+		\item Quantum groups
+		\item Hopf algebras that come from Nichols algebras
+	\end{itemize}
+\end{itemize}
+
+Papers to look at: 
+\begin{itemize}
+	\item Ginzburg and Kumar '94 ``Cohomology of Quantum Groups at Roots of Unity''
+	\item Mastnak, Pevtsova, Shauenburg, and Witherspoon '09 ``Cohomology of Finite Dimensional Pointed Hopf Algebras''
+	\item Witherspoon '17 ``Varieties for Modules of Finite Dimensional Hopf Algebras''
+\end{itemize}.
+
+Sarah Witherspoon has a book she is writing on Hochschild cohomology that could be very good. It's on her website (Texas A\& M). Check out chapter 9 and appendix A.
+
+\subsection{Week 3}
+We primarily discussed support varieties in the context of Sarah Witherspoon's \textit{Varieties for Modules of Finite Dimensional Hopf Algebras.}
+
+\subsubsection{Questions}
+\begin{enumerate}
+	\item On page 5, in the definition of the support variety of a Hopf module $M$ (see Definition~\ref{def-support-var}),
+	the construction seems to depend rather heavily on the choice (and existence!) of a graded subalgebra $H\le \HH^*(A)$
+	satisfying (fg1) and (fg2). But the notation seems to imply there is no dependence there. What's going on?
+	\item On page 9, why is $\mathcal{Z}(kG)=\HH^0(kG)$ a local ring (for $G$ a finite group)?
+	\item Directly after the last question, why would this ring being local imply that, on the variety level,
+	we can use $H^{ev}(G,k)$ instead of $H^{ev}(G,k)\cdot \HH^0(kG)$?
+\end{enumerate}
+
+\subsubsection{Answers/Hints}
+\begin{enumerate}
+	\item Actually the definition of support variety, when properly generalized to triangulated tensor categories,
+	doesn't rely on the choice of this algebra. But to see this, we need to dive deeper into that subject. Julia gave 
+	me a book by Dave Benson, Srikanth Iyengar, and Henning Krause on local cohomology and support. It was written
+	as a compendium from a course at Oberwolfach that they gave. It is available on the Arxiv.
+	\item Julia suggests that this is a ``standard'' result that one might find in Webb's book (or another book where the author has taken care to be thorough).
+	I couldn't locate it there immediately, but I found a page online that gave me the maximal ideal, so I am going to try to prove it as lemma~\ref{lem-local}.
+	\item The take-away here was that when computing the varieties the nilpotent elements are irrelevant. Since 
+	the only maximal ideal in $\HH^0(kG)$ is composed of nilpotents, it doesn't contribute to the support variety.
+\end{enumerate}
+
+\begin{defn}[Support Variety]\label{def-support-var}
+	Let $A$ be a finite dimensional algebra and assume $_AA$ is injective (notice that finite dimensional Hopf algebras are
+	Frobenius whence self-injective). Assume further that there is a graded subalgebra $H\le \HH^*(A)$ that satisfies the following 
+	finiteness conditions:
+	\begin{itemize}
+		\item \textbf{(fg1):} $H$ is finitely generated, commutative, and $H^0:=H\cap \HH^0(A)=\HH^0(A)$;
+		\item \textbf{(fg2):} for all finite dimensional $A$-modules $M$, $\Ext^*_A(M,M)$ is finitely generated as an $H$-module.
+	\end{itemize}
+
+	Then if $I_A(M)$ is the annihilator of $\Ext_A^*(M,M)$ in $H$, the \textbf{support variety} is
+	\[V_A(M)=\operatorname{MaxSpec(H/I_A(M))}\]
+\end{defn}
+\begin{rmk}
+	In taking with Julia about this definition she mentioned that the use of the max spectrum instead of the entire
+	spectrum is a bit watered-down and, while simpler, perhaps hearkens back to the beginnings of AG where that is what one considered.
+\end{rmk}
+
+\begin{lem}\label{lem-local}
+	If $G$ is a finite $p$-group and $k$ a field of characteristic $p$, $\mathcal{Z}(kG)$ is local.
+\end{lem}
+\begin{prf}
+	I claim that the maximal ideal is
+	\[I=\left\{\sum_{g\in G}a_gg:\sum_ga_g=0\right\}\]
+	To see this, let $A=\mathcal{Z}(kG)$, and notice that $A=\ker\varphi$ where $\varphi(\sum a_g g)=\sum a_i$.
+	Since $\varphi(kG)=k$, this gives us that $I$ is a maximal ideal in $A$.
+
+	Notice that if we prove that $I$ consists entirely of nilpotents, we will be done. This is because
+	the nilradical of $A$ is the intersection of its prime ideals. But if $I$ is contained in the nilradical, 
+	it must, in particular, be contained in every maximal ideal of $A$. Thus it must be the only maximal ideal.
+
+	To see this, we proceed by induction on $n$ where $|G|=p^n$. When $n=0$ this 
+	is obvious and when $n=1$, this means $G=\bbZ/p\bbZ=\langle\alpha\rangle$. In this case each element is nilpotent of 
+	degree $p$ due to the fact that (since in this case our ring is commutative):
+	\[\left(\sum_1^p a_i\alpha^i\right)^p=\sum_1^p a_i^p\alpha^{pi}=\sum_1^p a_i1_G=0.\]
+
+	Now assume that $I$ consists of nilpotents for all $n\le k$ and let $|G|=p^{k+1}$. Let $Z\le G$ be the
+	(necessarily nontrivial) center of $G$. Extend the canonical quotient map $q:G\to G/Z$ to a map 
+	between group rings $\tilde q:kG\to k(G/Z)$. Let $I_Z$ be the augmentation ideal of $k(Z)$. Then
+	$\tilde q$ is surjective with kernel $I_Z\cdot kG$. That this ideal is contained within the kernel is clear
+	since if $y$ is a $kG$-linear sum of elements of the form $\sum_Z a_zz$ where $\sum_Z a_z=0$, then
+	the image is just the sum of zeros.
+
+	To see that every element of the kernel is of this form, let $y=\sum k_i g_i$ and notice
+	\[\tilde q\left(\sum k_ig_i\right)=\sum k_i\tilde q(g_i)=\sum_j\left(\sum_{g_i-g_j\in Z}k_i\right)g_j\]
+	and this is zero precisely when all the sums are. But then the sum of the coefficients of $y$ restricted to any coset of $Z$
+	are zero.
+
+	Thus in particular $kG/(I_Z)\cong k(G/Z)$.
+
+	Then look at the augmentation ideal $I_{G/Z}\lhd k(G/Z)$ which, by the induction hypothesis, consists of nilpotents.
+	Pull this back through $\tilde q$ to the ideal $I=(I_{G/Z})+ (I_Z)$. Since some power of $p$ kills all elements of the two 
+	summands, this implies that some power of $p$ kills each element in $I$, completing the proof.
+\end{prf}
+
+\subsection{Week 4}
+Recall that $\mathcal{U}:\Lie\to \Ass$ is a functor that assigns to each 
+Lie algebra the universal enveloping algebra: the tensor algebra over $\g$ modulo the relations
+$x\otimes y-y\otimes x-[x,y]$.
+
+Then we have the PBW theorem:
+\begin{thm}[Poincar\'e-Birkhoff-Witt]
+	If $\g$ is a finite dimensional Lie algebra, with basis $x_1,\dots,x_n$, then $\mathcal{U}(\g)$ has a basis
+	in the form $\prod x_i^{a_i}$ for $a_i\in\bbZ_{\ge 0}$.
+\end{thm}
+\begin{rmk}
+	A fancier way of saying this is to say that $\mathcal{U}(\g)$ admits a filtration such 
+	that $\operatorname{gr}\mathcal{U}(\g)\cong S^*(\g)$.
+\end{rmk}
+
+Furthermore, there is a more general class of algebras called \textbf{PBW algebras} where you get commutation
+modulo lower-order terms. This idea is also what forms the basis of the proof for the above remark.
+
+The idea here, then, is to take the center and see how it acts on the universal enveloping algebra.
+
+\subsubsection{The BGG category $\mathcal{O}$}
+The representations of finite dimensional Lie algebras are in correspondence with the dominant integral weights. This
+will be worth looking into, perhaps.
+
+The general question we want to answer is what he structure of finite dimensional representations of $\g$. In the case
+when $\g=\sl_2$, we have the representation $V=ke_1\oplus ke_2$.
+
+Here I need to work out some of the notation, but $k[x,y]$ is a representation and $S^d(V)=k_d[x,y]$ is as well. (F\&H lecture 11 for more details there).
+
+So we end up realizing that we can generate any such Lie algebra by taking a single highest-weight vector and 
+applying $\mathcal{U}(\mathfrak{n}^-)$ to get the whole thing. This help motivate:
+\begin{defn}[BGG Category $\mathcal{O}$]
+	$\mathcal{O}$ is the full subcategory of $\mathcal{U}(\g)$ modules such that
+	\begin{itemize}
+		\item $M$ is finitely generated over $\mathcal{U}(\g)$ -- this is necessary since we want free modules
+		and $\mathcal{U}(\g)$ itself to be representations under study and these are excluded from finite ($k$-) dimensional algebras.
+		\item $M$ is $\mathfrak{h}$ semisimple.
+		\item $M$ is $\n$-locally finite -- for all $v\in M$, $\mathcal{U}(\n)v$ is finite dimensional.
+	\end{itemize}
+\end{defn}
+
+Ideas for things to understand/cover for next week:
+\begin{itemize}
+	\item First things first, understand the representations of $\sl(2,k)$.
+	\item Then in Humphreys(I), read/do section \S 6.21 to understand the universal enveloping algebra.
+	\item Read Humphreys(II) chapter 1 as much as possible.
+\end{itemize}
+
+\subsection{Week 5}
+We spent most of the time deriving the fact that any simple $\sl_2$ module is isomorphic to one of the ``standard''
+highest weight modules.
+
+Our assignment for this upcoming week:
+Write down the Serre relations. Then do some computations and read section 21 in Humphreys. Read those lemmas and theorems and start to describe decomposition of general modules of Lie algebras.
+
+\newpage
+\section{Lie Algebras}
+We're planning on digging into the category $\O$, the category of (reasonably finite) $\mathcal{U}(\g)$ modules.
+To begin, however, we are going to make sure we have our fundamentals vis a vis (semi) simple representations of Lie algebras.
+
+\subsection{$\sl_2$}
+In the case when $L=\sl_2$, everything is relatively nice. We read (and discussed) the fact that any simple, finite-dimensional
+$\sl_2$-module is isomorphic to $V(\lambda)$ for some integer $\lambda$ (highest weight).
+
+The action of $L$ on these modules are precisely as one would hope: $x$ and $y$ (the off-diagonal generators) act by raising 
+and lowering the weight of a vector by 2 and the central generator fixes each eigenspace.
+
+\subsection{The Serre Relations}
+Let $L$ be a semisimple Lie algebra and $\h$ a Cartan subalgebra with corresponding roots $\Phi$ with simple root system $\Delta$.
+Recall that 
+\[\langle\alpha_i,\alpha_j\rangle=\frac{2(\alpha_i,\alpha_j)}{(\alpha_j,\alpha_j)}=\alpha_i(h_{\alpha_j})\]
+Fix generators $x_i$ and $y_i$ of $L_{\alpha_i}$ and $L_{-\alpha_i}$, respectively, so that $[x_i,y_i]=h_i=h_{\alpha_i}$
+Recall that the $x_i$ and $y_i$ generate $L$ (as a Lie algebra).
+
+Then in particular we have the relations:
+\begin{align*}
+	[h_i,h_j]&=0\tag{S1}\\
+	[x_i,y_j]&=\delta_{ij}h_i\tag{S2}\\
+	[h_i,x_j]=\langle\alpha_j,\alpha_i\rangle x_j\text{ and }& [h_i,y_j]=-\langle\alpha_j,\alpha_i\rangle y_j\tag{S3}
+\end{align*}
+\begin{defn}[Serre Relations]
+	With the notation above, the relations S1-3 above are called the \textbf{Serre relations}.
+\end{defn}
+
+These relations end up being crucial in making the connection between root systems and semisimple Lie algebras concrete:
+taking the abstract free Lie algebra on generators satisfying these properties, we get a finite dimensional semisimple Lie algebra
+with roots corresponding exactly to the $\alpha_i$. This also gets us uniqueness of such a root system.
+
+\subsection{The more general case.}
+This primarily borrows from \S 21 in Humphreys (I) but also relies on some results from earlier that I will cite as 
+necessary.
+
+Throughout this discussion, we rely on the theory developed in sections 15 and 16 regarding Cartan and Borel subalgebras:
+\begin{defn}[Cartan Subalgebra (CSA)]
+	If $L$ is a Lie algebra, then $C\subseteq L$ is called a \textbf{Cartan subalgebra} if $C$ is
+	\begin{enumerate}
+		\item Nilpotent; and
+		\item $N_L(C)=C$.
+	\end{enumerate}
+\end{defn}
+\begin{defn}[Borel Subalgebra]
+	If $L$ is a Lie algebra, $B\subseteq L$ is called a \textbf{Borel subalgebra} if $B$ is a
+	maximal solvable subalgebra.
+\end{defn}
+\begin{rmk}
+	Since every nilpotent Lie algebra is solvable, each CSA lies in some Borel subalgebra $B$ of $L$.
+	then by showing that any two Borel subalgebras are conjugate under $\mathscr{E}(L)$ (the subalgebra of $\Int L$ generated
+	by the strongly $\ad$-nilpotent elements of $L$), it suffices to 
+	show that two CSAs are $\mathscr{E}(L)$-conjugate in any \textit{semisimple} Lie algebra.
+\end{rmk}
+
+In all that follows, we've fixed a CSA $\h$ of $L$ and root system $\Phi$ with simple root system $\Delta=\alpha_1,\dots,\alpha_l$
+of positive roots with $\mathcal{W}$ the Weyl group.
+
+\subsection{Representations}
+Since $\h$ is semisimple (since $L$ is), it acts diagonally on any finite dimensional $L$-module. Thus we can always
+do the same trick of decomposing $V$ into eigenspaces $V=\sqcup_{\lambda\in \h^*}V_\lambda$ that we do for $\sl_2$.
+Slightly more generally (when $V$ is infinite-dimensional), we can consider the sum of weight spaces $V_\lambda$ of $V$, 
+which is necessarily  direct\footnote{Think: vectors can't have multiple eigenvalues.}, so there always
+exists a direct sum of simple modules contained in $V$. When $\dim V<\infty$, this is equal to $V$ itself.
+
+\subsubsection{Standard Cyclic Modules}
+Here
+
+\newpage
+\section{Quantum Groups}
+I am mostly working out of Jantzen's \textit{Lectures on Quantum Groups} with the goal of reading through Ginzburg \& Kumar's
+(apparently influential!) paper \textit{Cohomology of Quantum Groups at Roots of Unity.}
+
+\subsection{What they are}
+The core idea to keep in mind is that quantum groups (or quantum enveloping algebras) are ``deformations'' (or perhaps $q$-analogues) of
+regular enveloping algebras. Well, at least that is the kind we're mostly working with here. A general definition eludes me (and Jantzen).
+
+Mostly we will be interested in $U=\mathcal{U}_q(\sl_2)$. In this setting, we get the definition
+\begin{defn}[Quantum Enveloping Algebra of $\sl_2$]
+	$\mathscr{U}_q(\sl_2)$ is the associative algebra $k\langle E,F,K,K^{-1}\rangle$ under the following relations:
+	\begin{align*}
+		KK^{-1}&=K^{-1}K=1\tag{R1}\\
+		KEK^{-1} &= q^2 E\tag{R2}\\
+		KFK^{-1} &= q^{-2}F\tag{R3}\\
+		EF-FE &= \frac{K-K^{-1}}{q-q^{-1}}\tag{R4}
+	\end{align*}
+	where $q$ is some nonzero element of $k$ such that $q^2\ne 1$.
+\end{defn}
+
+Some notation that may be helpful in the future:
+\begin{itemize}
+	\item For $a\in\bbZ$, define $[a]=\frac{v^a-v^{-a}}{v-v^{-1}}\in\bbQ(v)$ -- the $v$ analogue of $a$. Equivalently, (and perhaps suggestively):
+	\[[a]=v^{a-1}+v^{a-3}+\cdots+v^{-a+3}+v^{-a+1}.\]
+	\item the $v$ analogue of the factorial is
+	\[[a]^!=[a][a-1]\cdots[1]\]
+	\item Then you get the binomial coefficients:
+	\[\qbinom{a}{n}=\frac{[a]^!}{[n]^![a-n]^!}\]
+	\item \[[K;a]=\frac{Kq^a-K^{-1}q^{-a}}{q-q^{-1}}\]
+\end{itemize}
+
+\subsection{Getting a basis}
+The first part of the first chapter focuses on getting a PBW-style basis for $U$, which comes in the form $F^s,K^n,E^r$ for $r,s\in\bbN$, $n\in\bbZ$.
+This is mostly standard but I thought this proof had a neat idea in it:
+
+\begin{prf}
+	Let $A=k[X,Y,Z]$ and define $f,e,h\in \Hom_k(A,A)$ such that 
+	\begin{align*}
+		f(Y^sZ^nX^r)&=Y^{s+1}Z^nX^r,\\
+		e(Y^sZ^nX^r)&=q^{-2n}Y^sZ^nX^{r+1}+[s]Y^{s-1}[Z;1-s]Z^n X^r,\\
+		h(Y^sZ^nX^r)&=q^{-2s}Y^sZ^{n+1}X^r
+	\end{align*}
+	One can check that these vector space homomorphisms exactly mirror multiplication of these monomials
+	by $F, E,$ and $H$, respectively, so we get a map $\varphi:U\to\Hom_k(A,A)$. Then take any linear combination of 
+	$F^sK^nE^r$, mapped though to $A$ via $\varphi$ and then evaluation at 1, gives us a linear combination of the $Y^sZ^nX^r$,
+	so it must be trivial. Sick.
+\end{prf}
+
+\brk
+
+After that, the book focuses on $U_0=k\langle Z,Z^{-1}\rangle$ as well as $U^+=k\langle E\rangle$ and $U^-=k\langle F\rangle$,
+the latter of two which are isomorphic to $k[x]$ by the linear independence above.
+
+\brk
+
+Here's another cool bit: By studying products of monomials (or just defining degree outright) we can impose a graded structure on $U$:
+let $\deg(E)=1$, $\deg(F)=-1$ and $\deg(K)=\deg(K^{-1})=0$ and notice (R1-4) are homogeneous according to this degree.
+Thus the quotient by this homogeneous ideal gives a graded structure to $U$.
+
+Furthermore if $u\in U$ is homogeneous of degree $i$, 
+\[KuK^{-1}=q^{2i}u\]
+so when the powers of $q$ are distinct ($|q|\ne 1$), the graded pieces are precisely the 
+ of $K$ under the adjoint action.
+
+\subsection{Representations of $U$}
+
+\newpage
+\section{Affine Group Schemes (Waterhouse and Jantzen)}
 I have quite a bit of information to process before this, so I will get started!
@@
 \begin{rmk}\label{rmk-diag}
-	Notice that an alternative definition is to begin with an abelian group $M$ and to 
+	Notice that an alternative definition is to begin with an Abelian group $M$ and to 
 	define $S,\Delta,$ and $\varepsilon$ such that each element is group-like. The resulting
@@
 	a finitely generated abelian group, so we can split up the algebra into a tensor product
-	of $k[\Z]$ and $k[\Z/n\Z]$.
+	of $k[\bbZ]$ and $k[\bbZ/n\bbZ]$.
 
-	When $M=\Z$, $k[M]\cong k[X,X^{-1}]$, so $G= \Gm$\footnote{Notice that here Waterhouse also
+	When $M=\bbZ$, $k[M]\cong k[X,X^{-1}]$, so $G= \Gm$\footnote{Notice that here Waterhouse also
 	uses that the basis elements $e_i$ are group-like and so $\Delta(e_i)=e_i\otimes e_i$. I am not
-	quite sure why this is necessary.} Similarly if $M=\Z/n$ then $G\cong \mun{n}$
+	quite sure why this is necessary.} Similarly if $M=\bbZ/n$ then $G\cong \mun{n}$
 \end{prf}
@@
 
+\newpage
 \section{Algebraic Geometry}
 
+\newpage
 \section{Problems}
@@
 
+\section{To be filed:}
+
 \end{document}

### 2019-03-02 Hopf Algebra Notes/Hopf Algebras.tex
Updates

@@
 \DeclareMathOperator{\projdim}{projdim}
+\DeclareMathOperator{\injdim}{injdim}
 \DeclareMathOperator{\findim}{findim}
@@
 \DeclareMathOperator{\id}{id}
+
+\DeclareMathOperator{\End}{End}
 
@@
 \newcommand*{\Ch}{\mathbf{Ch}}
+\newcommand*{\g}{\mathfrak{g}}
+\DeclareMathOperator{\Der}{Der}
 
@@
 \end{thm}
-\begin{thm}[Larson-Radford `88]
+\begin{thm}[Larson-Radford `88]\label{thm-LR88}
 	Suppose $\ch k\ne 0$ (can be relaxed but gets more ugly). Then the following are equivalent:
@@
 \end{thm}
+
+\section{February 13, 2019}
+Here's the theorem for today (with the hypotheses in place):
+\begin{thm}\label{thm-213}
+	Assume $\ch k=0$ and $S^2=\id_V$. Then 
+	\[\gldim V=\projdim{_VM}\]
+	for all finite dimensional $V$ modules $M$.
+\end{thm}
+Another result is the theorem~\ref{thm-LR88}:
+\begin{prf}
+	Assuming theorem~\ref{thm-213}, (c)$\Rightarrow$(a) is obvious. That's only part of it, 
+	obviously. :)
+\end{prf}
+\subsection{Example of the Day}
+\begin{ex}
+	$\mathcal{U}(\g)$, the universal enveloping algebra which (as an algebra) is isomorphic to
+	$k\langle \g\rangle/(xy-yx-[xy])$. As a coalgebra, we define $\Delta(x)=x\otimes 1+1\otimes x$,
+	$\varepsilon=0$ and $S(x)=-x$.
+
+	This is a nice example for today, because notice that $S^2(x)=x$ and in fact $S^2=\id_{\mathcal{U}(\g)}$,
+	so we get that $\gldim\mathcal{U}(\g)=\projdim_{\mathcal{U}(\g)}M$ for any finite dimensional $M$ (by 
+	the theorem).
+\end{ex}
+\subsection{Today's Content}
+\begin{lem}[Reflection Principle]\label{lem-ref]}
+	Let $A$ be an algebra. Then
+	\begin{enumerate}
+		\item $(-)^*=\Hom)k(-,k):({_A\calM})_{fd}\to(\calM_A)_{fd}$ is a contravariant equivalence.
+		\item $(M^*)^*\cong M$ for each $M\in{_AM}_{fd}$. In linear algebra, this says that we have a perfect pairing $M\otimes M^*\to k$.
+	\end{enumerate}
+\end{lem}
+\begin{lem}[Restriction of Scalars]\label{lem-res1]}
+	Let $f:A\to B$ be a map of algebras. Then
+	\begin{enumerate}
+		\item There is a functor $f^*:{_B\calM}\to{_A\calM}$; also 
+		\item $f^*:\calM_B\to\calM_A$.
+	\end{enumerate}
+\end{lem}
+\begin{lem}[Restriction of Scalars]\label{lem-res2}
+	Let $f:A\to B$ be an antihomomophism of algebras. Then
+	\begin{enumerate}
+		\item There is a functor $f^*:{_B\calM}\to{\calM_A}$; also 
+		\item $f^*:\calM_B\to{_A\calM}$.
+	\end{enumerate}
+\end{lem}
+\begin{lem}
+	Let $V$ be a Hopf algebra with antipode $S$. Then 
+	\[\tilde S\circ (-)^*:{_V\calM}\to \calM_V\to {_V\calM}\]
+	is a contravariant functor.
+\end{lem}
+\begin{cor}
+	If $V$ is finite dimensional, then $\tilde S\circ(-)^*$ is a duality of $(_V\calM)_{fd}$.
+\end{cor}
+For the future we will use the notation $M^{*S}=\tilde S(M^*)$. Notice that for all $M\in{_V\calM},$ $M^{*S}\in{_V\calM}$.
+
+This leads us to our proposition:
+\begin{thm}
+	$V$ is a Hopf algebra.
+	\begin{enumerate}
+		\item If $M\in{_V\calM}$ then there is a natural morphism
+		\[ev_M:M^{*S}\otimes M\to {_Vk}\]
+		\item if $M\in{_V\calM_{fd}}$, then there is a natural map
+		\[coev_M:{_Vk}\to M\otimes M^{*S}\]
+	\end{enumerate}
+\end{thm}
+\begin{prf}
+	$ev_M(f\otimes m)=f(m)\in k$ is the map we're talking about. For all $v\in V$, 
+	\begin{align*}
+		ev_M(v\cdot(f\otimes m))&=ev_M(v_1f\otimes v_2 m)\\
+		&=ev_M(fS(v_1)\otimes v_2m)\\
+		fS(v_1)(v_2 m)\\
+		&= f(S(v_1)v_2m)\\
+		&= f(\varepsilon(v) m)=\varepsilon(v)f(m)=v\cdot ev_M(f\otimes m).
+	\end{align*}
+
+	If instead $M$ is finite dimensional, pick a $k$ basis $m_1,\dots,m_d$ for $M$ and let 
+	$m_i^*$ be the dual basis. Set $coev_M(1)=\sum m_i\otimes m_i^*=\Phi$. Then use the fact that (under the identification of 
+	$M\otimes M^*$ with $\Hom_k(M,M)$) we get $\Phi=\id_M$. Then prove 
+	\[coev_M(v\cdot 1)=v\cdot coev_M(1).\]
+\end{prf}
+\begin{defn}
+	The map $ev_M$ above is called the \textbf{evaluation map} and $coev_M$ is called the 
+	\textbf{coevaluation map.}
+\end{defn}
+
+\begin{lem}
+	\begin{enumerate}
+		\item $(M^{*S})^{*S}\cong \tilde S^2(M)$ for all finite dimensional $M\in{_V\calM}$.
+		\item If $S^2=\id$ then $(M^{*S})^{*S}\cong M$.
+	\end{enumerate}
+\end{lem}
+\begin{lem}
+	Let $M\in {_VM}_{fd}$ and suppose $S^2=1$. Then 
+	\[ev_{M^{*S}}\circ coev_M=(\dim M)\id_k\]
+\end{lem}
+
+\section{February 15, 2019}
+
+Here we use the (slightly extended) lemma:
+\begin{lem}
+	\begin{enumerate}
+		\item $ev_M\circ coev_{M^{*S}}=(\dim M)\id_k$
+		\item $\frac{1}{\dim M}ev_M\circ coev_{M^{*S}}=\id$
+		\item (since the identity factors through $M^{*S}\otimes M$) $M^{*S}\otimes M\cong {_Vk}\oplus N$ as $V$-modules.
+		\item $\projdim k\le \projdim M^{*S}\otimes M$.
+	\end{enumerate}
+\end{lem}
+Then using this lemma, we can prove the theorem~\ref{thm-213}:
+\begin{prf}
+	Basically use this, Lorenz-Lorenz, and the result bounding $\projdim$ of a tensor module by its factors.
+\end{prf}
+
+Notice that if we change things up slightly we get a new theorem:
+\begin{thm-prime}{thm-213}
+	Let $V$ be aHopf algebra. Suppose that $S^2=\id$. Then 
+	\[\projdim_V M=\gldim V\]
+	for all $M\in {_VM}_{fd}$ such that $\dim M\ne 0$ in $k$.
+\end{thm-prime}
+
+\begin{conj}
+	Let $V$ be a Noetherian Hopf algebra over a field $k$ of characteristic zero and $S^2=\id$.
+	Then
+	\[\gldim V <\infty\]
+\end{conj}
+
+\begin{cor}
+	If $V$ is a finite module over $\mathcal Z(V)$, its center, then the above conjecture holds.
+\end{cor}
+
+\subsection{Today's example}
+He did the quantized enveloping algebra of $\sl_2$! Might be worth looking up.
+
+An interesting part of this example is that $S^2$ is \textbf{not} the identity! In fact, 
+this example is important in constructing Joon (sp?) polynomials in knot theory.
+
+\subsection{Today's Homework}
+\begin{prob}
+	Prove the following ``dimensional shift'' lemma: Let
+	\[0\to X\to P\to Y\to 0\]
+	be a SES of $A$ modules, where $P$ is projective. If $M$ is an $A$ module, then for all $n\ge 1$:
+	\[\Ext_A^{n+1}(Y,M)\cong\Ext^{n}_A(X,M).\]
+\end{prob}
+\begin{sol}
+	Take a projective resolution for $X$ and try to attach it to the SES.
+\end{sol}
+\subsection{Classifying Algebras}
+We have the (descending) chain of classes of algebras:
+\begin{itemize}
+	\item Algebras
+	\item Noetherian Algebras
+	\item Cohen-Macaulay Algebras
+	\item Gorenstein Algebras
+	\item Complete Intersections
+	\item Hypersurface Rings
+	\item Regular Algebras
+\end{itemize}
+The question we have is: where to (Noetherian) Hopf algebras lie in this hierarchy?
+
+If we restrict to finite-dimensional Hopf algebras, Brown-Goodearl proved that Noetherian Hopf algebras
+are all Gorenstein rings.
+
+\begin{defn}
+	A Noetherian algebra $A$ is called \textbf{regular} if $\gldim A<\infty$.
+\end{defn}
+\begin{rmk}
+	Notice that since $A$ is Noetherian, we didn't have to talk about left vs. right global dimension.
+\end{rmk}
+\begin{ex}
+	If $L$ is a finite dimensional Lie algebra, then $U(L)$ is regular:
+	\[\gldim U(L)=\dim L.\]
+\end{ex}
+\begin{defn}
+	An element $x\in A$ is called \textbf{normal} if $xA=Ax$ (NOT $xa=ax$)
+\end{defn}
+\begin{defn}
+	A Noetherian algebra $A$ is called a \textbf{hypersurface ring} if there exists a Noetherian 
+	regular algebra $B$ and a normal non-zero-divisor $x\in B$ such that $A\cong B/(x)$
+\end{defn}
+\begin{ex}
+	$\ch k=p>0$, $V=k[x]/(x^p)$, a Hopf algebra with comultiplication $\Delta(x)=x\otimes 1+1\otimes x$.
+	Then $V$ is a hypersurface ring.
+\end{ex}
+\begin{defn}
+	A Noetherian algebra $A$ is called a \textbf{complete intersection} if there is a Noetherian regular algebra
+	$B$ and a sequence of elements $x_1,\dots,x_d$ such that:
+	\begin{itemize}
+		\item $x_1$ is a non-zero-divisor in $B$.
+		\item $x_i$ is a NZD in $B/(x_1,\dots,x_{i-1})$ for all $i$
+		\item $A\cong B/(x_i)$.
+	\end{itemize}
+\end{defn}
+
+\begin{defn}
+	A Noetherian algebra $A$ is called \textbf{Gorenstein} if 
+	\[\injdim{_AA}=\injdim A_A<\infty.\]
+\end{defn}
+\begin{defn}
+	A Noetherian algebra $A$ is called \textbf{Cohen-Macaulay} if there is an $A$-bimodule
+	$M$ such that
+	\begin{itemize}
+		\item $\injdim {_AM}<\infty$, $\injdim M_A<\infty$
+		\item $\End(M_A)\cong A,$ $\End({_AM})\cong A^{op}$
+		\item ${_AM}$ and $M_A$ are Noetherian.
+	\end{itemize}
+\end{defn}
+
+\section{February 20, 2019}
+Last time we wrote out a big chain of inclusions for algebras. We didn't really discuss why all complete
+intersections are Gorenstein algebras. Actually, this is one of the harder inclusions to prove. We will at least 
+try today, but will definitely get to it by Friday.
+
+\subsection{Example of the day}
+\begin{defn}
+	$V=k[g^{\pm 1}][y;\delta]$, an Orr extension of the Laurent polynomial ring by the derivation $\delta=(g^n-g)\frac{d}{dg}$.
+	For those (me) wo haven't seen Orr extensions before, we also have
+	\[V\cong k\langle g, g^{-1}, y\rangle/(yg-gy-(g^n-g), gg^{-1}-1, g^{-1}g-1)\]
+
+	This is a domain where $\gldim V=2$ (also the GK dimension, which we haven't defined). As a coalgebra, we get
+	$g$ is grouplike and $\Delta y= y\otimes g^{n-1}+1\otimes y$, $\varepsilon(y)=0$, and $S(y)=-yg^{-(n+1)}$.
+
+	Notice that $V$ is regular.
+\end{defn}
+\subsection{Problem of the day}
+\begin{prob}
+	Let $x$ be a normal non-zero-divisor in $A$ where $xA=Ax$. Then for all $a\in A$ let $\sigma\in\Aut A$ be defined such that
+	$\sigma(a) x= xa$.
+
+	Let $M\in {_A\calM}$ and define a left $A$ action on ${^\sigma M}$ whose underlying set is $M$ and the action is defined by $a\cdot m=\sigma(a)m$. In terms of the action (morphism), this is 
+	equivalent to precomposing with $\sigma:$ 
+	\[A\xrightarrow{\sigma} A\xrightarrow{\rho}\End(M)\]
+
+	Prove the following lemma:
+	\begin{lem}[Rees Lemma]
+		Let $x$ be a normal non-zero-divisor in $A$. If $L$ is a left $A/(x)$-module and $M$ is an $x$-torsion-free $A$-module, then
+		\[\Ext_{A/(x)}^n(L,M/xM)\cong \Ext_A^{n+1}(L,{^\sigma M}).\]
+	\end{lem}
+\end{prob}
+
+\subsection{Today's Results}
+We begin with some corollaries of Rees Lemma:
+\begin{cor}
+	\begin{enumerate}
+		\item $\Ext_{A/(x)}^n(L, A/(x))\cong \Ext_A^{n+1}(L,{^\sigma A})\cong\Ext_A^{n+1}(L,A)$
+		\item $\injdim_{A/(x)}(A/(x))\le \injdim_AA-1$
+		\item If $A$ is Gorenstein, so is $A/(x)$.
+	\end{enumerate}
+\end{cor}
+\begin{rmk}
+	Then the idea here is that we just have to prove that regular rings are Gorenstein, since then complete intersections
+	are just (finite) sequences of quotients by regular elements.
+\end{rmk}
+\subsection{Doing $\Ext$ again}
+Let $P^\bullet$ and $Q^\bullet$ be complexes of $A$-modules. Then $\Hom_A(P^\bullet, Q^\bullet)$ is a complex. Define
+\[\Hom_A(P^\bullet, Q^\bullet):=\prod_{i\in\Z}\Hom_A(P^i,Q^{i+n}=\prod_{-i+j=n}\Hom(P^i,Q^j)\]
+where we define the boundary map to be 
+\[d_{\Hom_A(P^\bullet,Q^\bullet)}:(f_i)_{i\in\Z}\mapsto (d_Q^{n+1}\circ f^i - (-1)^nf^{i+1}\circ d_P^i)_{i\in\Z}\]
+\begin{prob}
+	The \textbf{real} homework for this class (the other one is what we're showing, apparently) is to show that this is a
+	chain complex -- i.e. that $d^2=0$.
+\end{prob}
+
+Define the projective resolution to be the regular projective resolution, but do not include $M$ itself. But then we have
+\begin{align*}
+	R\underline{\Hom}=\Ext_A^n(M,N)&:=H^n(\Hom_A(P^\bullet_M,N))\\
+	&:=H^n(\Hom_A(P_M^\bullet,P_N^\bullet))\\
+	&:=H^n(\Hom_A(I_M^\bullet, I_N^\bullet))\\
+	&:=H^n(\Hom_A(M,I_N^\bullet))
+\end{align*}
+
+\subsection{Proof plz}
+Today we will try to get $n=0$ in the inductive argument.
+\begin{lem}
+	\begin{enumerate}
+		\item  There is a short exact sequence of left $A$ modules
+		\[0\to {^\sigma A}\xrightarrow{l_x}A\to A/(x)\to 0\]
+		where $l_x$ sis left multiplication by $x$.
+		\item If $M$ is $x$-torsion-free, then there is a SES of left $A$ modules
+		\[0\to {^\sigma M}\xrightarrow{l_x}M\to M/xM\to 0\]
+	\end{enumerate}
+\end{lem}
+The second item implies the first and neither is hard to show directly.
+\begin{lem}
+	Let $M$ be an $x$-torsion-free left $A$ modules and let $I_M$ be an injective resolution of $M$. THen there is a morphism of complexes
+	$l_x:{^\sigma I_M}\to I_m$.
+\end{lem}
+Again, not too hard to check.
+
+\brk
+
+Notice now that $xL=0$ by the hypotheses. Combining this with the last result, we have
+\begin{lem}
+	\begin{enumerate}
+		\item For any $L\in {_A\calM}$,
+		\[\Ext_A^n(L,l_x):\Ext_A^n(L,{^\sigma M})\to \Ext_A^n(L,M)\]
+		\item If $xL=0$ then the above map is zero for all $n\ge 0$.
+	\end{enumerate}
+\end{lem}
+Again, follow your nose a bit. $f:L\to {^\sigma I^i}$ is an $A$ homomorphism and pull that $x$ in.
+
+Applying $\Hom_A(L,-)$ to $0\to {^\sigma M}\xrightarrow{l_x}M\to M/xM\to 0$, we obtain the long exact sequence
+for $\Ext$. But then notice that $\Hom_A(L,M)=\Ext_A^1(L,M)=0$ since $M$ is $x$-torsion-free (and $L$ is killed by $x$).
+and since $\Ext_A^1(L,{^\sigma M})\cong \Hom_A(L,M/xM)\cong\Hom_{A/(x)}(L,M/xM)$, we get the base case proven.
+
+\section{February 22, 2019}
+Today we are going to finish the proof of Rees, then give some remarks, then give the example of the day. :)
+
+\subsubsection{Continuation of Proof of Rees}
+Last time we proved that $\Ext_{A}^1(L,{^\sigma M})\cong \Ext_{A/(x)}^0(L,M/xM)$. Now we proceed by induction.
+
+Consider the SES $0\to K\to F\to L\to 0$ of left $A/(x)$-modules where $F$ is free. Then we get a long exact sequence
+\begin{align*}
+	0\to \Hom_{A/(x)}(L,M/xM)\to \Hom_{A/(x)}&(F,M/xM)\to \Hom_{A/(x)}(K,M/xM)\\
+	&\to \Ext^1_{A/(x)}(L,M/xM)\to\Ext^1_{A/(x)}(F,M/xM)\to\cdots
+\end{align*}
+as well as the sequence
+\begin{align*}
+	\cdots \to \Ext^1_{A}(L,{^\sigma M})\to \Ext^1_{A}&(F,{^\sigma M})\to \Ext^1_{A}(K,{^\sigma M})\\
+	&\to \Ext^2_{A}(L,{^\sigma M})\to \Ext^2_{A}(F,{^\sigma M}\to\cdots)
+\end{align*}
+and we have isomorphisms between all the corresponding parts (except the fourth term). THe first three 
+follow from $n=0$ and the last one follows since $F$ is a projective $A/(x)$ module and since $F$ has projective dimension (over $A$) 
+less than or equal to one. Then five lemma proves that this holds for $n=1$.
+
+Then you can just extend both sequences and iteratively prove this holds for each term. Woot. :)
+
+\subsection{Return to Classifications}
+Recall we have the (descending) category chain
+\begin{itemize}
+	\item Algebras
+	\item Noetherian Algebras
+	\item Cohen-Macaulay Algebras
+	\item Gorenstein Algebras
+	\item Complete Intersections
+	\item Hypersurface Rings
+	\item Regular Algebras
+\end{itemize}
+Now we can show by applying Rees directly:
+\begin{cor}
+	$\injdim A/(x)\le \injdim A$
+\end{cor}
+Which then allows us to prove:
+\begin{cor}
+	Every complete intersection ring is Gorenstein.
+\end{cor}
+\begin{prf}
+	Let $A$ be a CI ring. Then there exists a Noetherian regular ring $B$ and a sequence of elements $b_1,\dots, b_n\in B$ 
+	such that each $b_i$ is a normal nonzerodivisor on $B/(b_1,\dots,b_{i-1})$.
+
+	Since $B$ is regular, it is Gorenstein. By the corollary $B/(b_1)$ is Gorenstein. Continuing this way, $A$
+	itself is Gorenstein and we are done. 
+\end{prf}
+
+\subsection{Where do Hopf algebras fit in?}
+Recall that we proved that (finite dimensional) Hopf algebras are Frobenius algebras. Well it ends up
+that Frobenius algebras are also Gorenstein! Thus all (f.d) Hopf algebra is Gorenstein.
+
+Recall the Brown-Goodearl conjecture: all Noetherian Hopf algebras are Gorenstein. This is still an open conjecture!
+But it gets worse!
+\begin{conj}
+	Every Noetherian Hopf algebra is a complete intersection.
+\end{conj}
+\begin{conj}
+	Every finite dimensional Hopf algebra is a complete intersection.
+\end{conj}
+There are no known counterexamples to either! This has been open for decades with no satisfactory result.
+\subsubsection{Even more classes of algebras!}
+Inside the category of Gorenstein rings, there are a couple more classes (not related to complete intersections):
+We get the decreasing chain
+\begin{itemize}
+	\item Gorenstein rings
+	\item Artin-Schelter Gorenstein
+	\item A-S Gorenstein with skew Calabi-Yau property
+\end{itemize}
+and over at regular algebras we get the decreasing chain
+\begin{itemize}
+	\item Regular algebras
+	\item A-S regular algebras
+	\item Skew C-Y algebras
+	\item C-Y algebras
+\end{itemize}
+\begin{defn}
+	A Noetherian Gorenstein ring $A$ of $\injdim$ d is called $A-S$ Gorenstein if, for all $S\in{_A\calM}_{fd},$
+	\[\Ext_A^i(S,A)=\left\{\begin{array}{lr}0 & i\ne d,\\ \in {_A\cal M}_{fd}, & i=d\end{array}\right.\]
+	where the above property is called the Artin-Schelter (AS) property.
+\end{defn}
+\begin{defn}
+	A Noetherian A-S Gorenstein algebra $A$ of injective dimension $d$ is called \textbf{skew Calabi-Yau} if
+	\[\Ext_{A^e}^i(A,A^e)=\left\{\begin{array}{lr}0 & i\ne d,\\ {^\sigma A}, & i=d\end{array}\right.\]
+	(remember that $A^e\cong A\otimes A^{op}$).
+\end{defn}
+\begin{defn}
+	$A$ is Calabi-Yau if it is skew C-Y and also $\Ext_{A^e}^d(A,A^e)=A$.
+\end{defn}
+\begin{rmk}
+	The idea here (and honestly I missed some parts of the discussion) is that C-Y algebras are ones where we can get 
+	Poincar\'e duality.
+\end{rmk}
+\subsection{Example of the Day}
+\begin{ex}
+	$V$ is generated by $g,g^{-1},x_2,x_3$ subject to
+	\[gg^{-1}=g^{-1}g=1,\quad gx_i=\eta^ix_i g,\quad x_2^3=x_3^2\quad x_2x_3=x_3x_2\]
+	where $\eta$ is a primitive $6^{th}$ root of unity.
+
+	Then $\gldim V=\infty$ and it is a kypersurface ring, since
+	\[V\cong k[x_1,x_2][g^{\pm 1};\phi]/(x_2^3-x_3^2)\]
+	where $\phi:x_i\to \eta^ix_i$.
+
+	As a coalgebra, $g$ is grouplike and $x_i$ are such that $\Delta(x_i)=x_i\otimes 1+g^i\otimes x_i$
+	and $\varepsilon(x_i)=0$. Then we get $S(x_i)=-g^{-i}x_i.$
+\end{ex}
+\subsection{HW for the day}
+\begin{prob}
+	Let $A$ and $B$ be rings. Let $_AM$, $_AN_B$, and $_BC$. Then 
+	\[\Hom_A(N\otimes_B C,M)\cong \Hom)B(C, \Hom_A(N,M))\]
+	and if instead $C_A$, $_AN_B$ and $M_B$, then you can do things over $A^{op}$ and $B^{op}$.
+\end{prob}
+
+\section{February 25, 2019}
+Topic for today: \textbf{Hopf actions on algebras.} One can naturally extend everything we talk about
+today to (co)actions on (co)algebras. :)
+\subsection{Example of the Day}
+\begin{ex}\label{ex-22519}
+	$V=k\langle g,g^{-1},x\rangle/(gg^{-1}=1=g^{-1}g)$. Then $g$ is grouplike and $\Delta(x)=x\otimes1+g\otimes x$, 
+	$\varepsilon(x)=0$, and $S(x)=-g^{-1}x$.
+
+	We can see that $\gldim V=1$ (use resolution $0\to V\oplus V\to V\to {_V k}\to 0$). Note that
+	it is NOT AS-Gorenstein. $\Ext_V^1(k,V)$ is infinite dimensional over $k$ (not easy!).
+\end{ex}
+\begin{rmk}
+	We've seen elements like $x$ above several times -- these elements are called $(1,g)$-primitive 
+	(generalizing the concept of primitive, obviously).
+
+	Also, these are examples of \textbf{pointed Hopf algebras}(!!) These are $V$ such that (as a
+	coalgebra) the coradical is a sum of one-dimensional modules. This is actually related to the fact 
+	that every simple submodule of $V$ is of the form $kg^n$ (in this case $n=1$).
+\end{rmk}
+\subsection{HW of the Day}
+\begin{prob}
+	This problem considers (one of the) internal $\Hom$(s) in $_V\calM$. Suppose $V$ has bijective 
+	antipode. Then for any $M$ and $N$,
+	\[\Hom^{S^{-1}}(M,N):=\Hom_k(M,N)\]
+	where a $V$ action is defined by (for $v\in V$ and $f\in\Hom^{S^{-1}}(M,N)$) 
+	\[(v\cdot f)(m)=\sum v_2f(S^{-1}(v_1)m)\]
+	where, as usual, $\Delta v=\sum v_1\otimes v_2$.
+	\begin{enumerate}
+		\item Show that this makes $\Hom^{S^{-1}}(M,N)$ into a left $V$-module.
+		\item Show that $\Hom^{S^{-1}}(-,-)$ is right adjoint to $-\otimes_k -$.
+	\end{enumerate}
+\end{prob}
+\subsection{Automorphisms and Group Actions}
+Let $A$ be an algebra. $\Aut_{alg}(A)$ denotes the \textbf{group of autormorphisms of $A$.}
+\begin{defn}\label{def-action}
+	Let $G$ be a group. We say $G$ acts on $A$ ($G\curvearrowright A$) fi there is a group homomorphism
+	$G\to \Aut_{alg}(A)$.
+\end{defn}
+\begin{defn}
+	Alternatively, we say $G\curvearrowright A$ if
+	\begin{itemize}
+		\item $A$ is a $G$-module.
+		\item $\forall g\in G$, $g(1_A)=1_A$
+		\item $\forall a,b\in A$, $g(ab)=g(a)g(b)$
+		\item $g$ is bijective.
+	\end{itemize}
+\end{defn}
+\begin{defn}
+	Suppse $G\curvearrowright A$. Then the \textbf{invariant subring} of this action is 
+	\[A^G:=\{a\in A|g(a)=a,\forall g\in G\}\subseteq A\]
+\end{defn}
+This was fairly popular around 1880-1960 or so.
+\begin{ex}
+	$A=k[x_1,x_2]$, $G=\langle\sigma\rangle\cong C_2$ where
+	$\sigma\cdot f(x_i)=f(-x_i)$. One can compute that this just switches signs on monomial 
+	of odd degree (and fixes the even ones) so
+	\[A^G=\left\{\sum c_{ij}x_1^ix_2^j|i+j=2k\right\}\cong k[a,b,c]/(ac-b^2)\]
+\end{ex}
+Generalizing this, many of the questions of this time revolved around asking what category of 
+algebra $k[x_1,\dots,x_n]^G$ is for arbitrary $G$. We may return to this later.
+
+\subsection{Derivations}
+\begin{defn}
+	Let $A$ be an algebra. A map $d\in\Hom_k(A,A)$ is called a \textbf{derivation} if
+	it satisfies the Leibniz rule.
+\end{defn}
+\begin{lem}
+	The collection of all derivations of an algebra from a Lie algebra -- denoted $\Der(A)$.
+\end{lem}
+\begin{defn}
+	Let $L$ be a Lie algebra. We say $L$ acts on $A$ if there is a Lie algebra morphism $L\to \Der(A)$.
+	Equivalently, if
+	\begin{enumerate}
+		\item $A$ is a left $L$-module
+		\item $\forall l\in L$, $l(1_A)=0$
+		\item $\forall a,b\in A$, $l(ab)=l(a)b+al(b)$.
+	\end{enumerate}
+\end{defn}
+\begin{rmk}
+	The second item above is implied by the first (in fact, James included a similar statement in 
+	the definition of a derivation) but apparently there is a reason to include it!
+\end{rmk}
+\begin{ex}
+	$A=k[x_1,x_2]$, $d=\frac{\partial}{\partial x_1}$. Then $L=kd$ is a Lie algebra that acts on $A$.
+
+	Then $A^L$ (see below) is $k[x_2]$
+\end{ex}
+\begin{defn}
+	If $L$ is a Lie algebra acting on $A$, then the invariant subring under the action is
+	\[A^L=\{a\in A|l(a)=0,\forall l\in L\}\]
+\end{defn}
+
+\subsection{Putting These Things Together}
+Let $A$ be an algebra and $V$ be a Hopf algebra.
+\begin{defn}
+	We say $V$ actions on $A$ if
+	\begin{itemize}
+		\item $A$ is a left $V$-module
+		\item $A$ is an algebra object in $({_V\cal M},\otimes)$
+	\end{itemize}
+	Note that the second item above implies the first!
+
+	Equivalently, $V$ acts on $A$ if
+	\begin{itemize}
+		\item $A\in {_V\cal M}$
+		\item For all $v\in V$, $v\cdot(1_A)=\varepsilon(v)1_A$
+		\item For all $a,b\in A$, $v(ab)=\sum v_1(a)v_2(b)$.
+	\end{itemize}
+\end{defn}
+
+Returning to the example from the beginning of class (\ref{ex-22519}), if $V$ acts on $A$,
+then 
+\begin{itemize}
+	\item $g\in \Aut_{alg}(A)$
+	\item $x\in\Der_g(A)$
+\end{itemize}
+In particular, notice that $x(1)=0$ and $x(ab)=x(a)b+g(a)x(b)$.
 \end{document}

### 2019-03-02 Lie Algebra Notes/Lie Algebras.tex
Updates

@@
 \end{rmk}
+
+\newpage
+\section*{Homework 3 -- Due Feb 15}
+%10.10,13; 11.2; 13.4, board problem: using large tensor powers of the adjoint representation and sl(2) theory, show that any semisimple Lie algebra L admits irreducible representaitons of arbitrarily large dimension
+\begin{hwprob}{10.10}
+	Given $\Delta=\{\alpha_1,\dots,\alpha_l\}$ in $\Phi$, let $\lambda=\sum_1^lk_i\alpha_i$ with all $k_i\in\Z$ and all nonpositive or all nonnegative. Prove that
+	either $\lambda$ is a multiple (possibly 0) of a root or else there exists $\sigma\in W$ such that $\sigma\lambda=\sum k_i'\alpha_i$ with 
+	some $k_i'>0$ and some $k_i'<0$.
+
+	[Sketch: if $\lambda$ is not a multiple of any root, then the hyperplane
+	$P_\lambda$ orthogonal to $\lambda$ is not included in $\cup_{\alpha\in\Phi}P_\alpha$. Take $\mu\in P_\lambda-\cup_\Phi P_\alpha$. Then find
+	$\sigma\in W$ for which all $(\alpha_i,\sigma\mu)>0$. It follows that $0=(\lambda,\mu)=(\sigma\lambda,\sigma\mu)=\sum k_i(\alpha_i,\sigma\mu).$]
+\end{hwprob}
+\begin{sol}
+	Assume that $\lambda$ is not a scalar multiple of any root in $\Phi$. As suggested above, this means that $P_\lambda$ 
+	is not coincident with any $P_\alpha$ for $\alpha\in\Phi$. Let $\mu\in P_\lambda-\cup_\Phi P_\alpha$.
+
+	Now by the theorem in section 10.3 (part a) gives us (since $\mu$ is regular) a $\sigma\in W$ such that $(\alpha_i,\sigma\mu)>0$ for all $i$.
+	Then since $\lambda\perp\mu,$ (if $\sigma\lambda=\sum_i c_i\alpha$ where $c_i\in\Z$ since the integer lattice is fixed by $W$):
+	\[0=(\lambda,\mu)=(\sigma\lambda,\sigma\mu)=\sum_i c_i(\alpha_i,\sigma\mu)\]
+	which (since all the inner products are strictly positive) forces some $c_i>0$ and $c_j<0$, as desired.
+\end{sol}
+
+\begin{hwprob}{10.13}
+	The only reflections in $W$ are those of the form $\sigma_\alpha$ $(\alpha\in\Phi)$.
+
+	[A vector in the reflecting hyperplane would, if othogonal to no root, be fixed only by the identity in $W$.]
+\end{hwprob}
+\begin{sol}
+	Let $\sigma\in W$ be an arbitrary reflection. Say that $\sigma$ is not a reflection through the hyperplane
+	with normal vector $\lambda$ and let $\mu$ be a vector fixed by $\sigma$. If $\mu$ is not orthogonal to any root,
+	it lies in the interior of one of the regions carved out by the hyperplanes $P_\alpha$ for $\alpha\in\Phi$. 
+	Then by Lemma B in \S10.3, since $\mathfrak{C}(\Delta)$ is a fundamental domain for the $W$ action, let $\mu'=\sigma'\mu$
+	be the (unique) representative of $\mu$ in $\mathfrak{C}(\Delta)$.
+
+	But then $\sigma'\sigma(\sigma')^{-1}(\sigma'\mu)=\sigma'\mu$, so by the above lemma $\sigma'\sigma(\sigma')^{-1}$
+	is the identity map. But then 
+	\[\sigma'\sigma(\sigma')^{-1}=\id\quad\Rightarrow\quad \sigma'\sigma=\sigma'\quad\Rightarrow\quad \sigma=\id.\]
+\end{sol}
+
+\begin{hwprob}{11.2}
+	Calculate the determinants of the Cartan matrices (using induction on $l$ for types $A_l-D_l$).
+\end{hwprob}
+\begin{sol}
+	In the following, let the caligraphic letters $\mathcal{A,B,C,D,E,F,G}$ correspond to the Cartan matrix 
+	for type $A,B,$ etc.
+\subsection*{($A_l$)}
+	The claim is that $|\mathcal{A}_l|$ (determinant of the Cartan matrix) is $l+1$. For $A_1$ this is trivial.
+	Assume now that this holds up to $A_{k}$. But notice that $\mathcal A_k$ is a minor of $\mathcal A_{k+1}.$ In particular,
+	\[\begin{pmatrix}
+		2 & -1 & \mathbf{0}^T\\
+		-1 & \mathcal A_k &\\
+		\mathbf{0} & &
+	\end{pmatrix}\]
+	where $\mathbf{0}\in\R^{k-1}$ and so
+	\[|\mathcal A_{k+1}|=2\cdot |\mathcal A_k|+|\tilde{\mathcal A}_k|\]
+	where $\tilde{\mathcal A}_k$ is the same as $\mathcal A_k$ with thr first row replaced with $(-1,0,\dots,0)$.
+	But then subtracting the first row from the second (which leave the determinant unchanged):
+	\[\tilde{\mathcal A}_k=\begin{pmatrix}
+		-1 & \mathbf{0}^T\\
+		-1 & \mathcal A_{k-1}\\
+		\hat{\mathbf{0}} &
+	\end{pmatrix}\rightsquigarrow\begin{pmatrix}
+		-1 & \mathbf{0}^T\\
+		\mathbf{0} & \mathcal A_{k-1}
+	\end{pmatrix} \]
+	($\hat{\mathbf{0}}\in\R^{k-2}$) and so by the inductive hypothesis $|\tilde{\mathcal A}_k|=-|\mathcal A_{k-1}|=-k$.
+
+	But then we can conclude that
+	\[|\mathcal A_{k+1}|=2\cdot |\mathcal A_k|+|\tilde{\mathcal A}_k|=2(k+1)-k=k+2\]
+	as desired.
+
+	\brk
+	\subsection*{($B_l$)}
+	Here the claim is $|\mathcal B_l|=2$. This can be seen directly (without induction) by using that 
+	$\mathcal B_l$ differs in only one spot from $\mathcal A_l$: expanding the determinant from the 
+	bottom right along the last column, we get 
+	\[|\mathcal B_l|=2|\mathcal A_{l-1}|+2|\hat{\mathcal A}_{l-2}\]
+	where $\hat{\mathcal A_{l-2}}$ is $\mathcal A_{l-2}$ with the last column replaced by $(0,\dots, 0,-1)$.
+	Luckily, it can be seen by expanding upon the last column that
+	\[|\hat{\mathcal A}_{l-1}|=|\tilde{\mathcal A}_{l-1}|=-(l-1)\]
+
+	So finally:
+	\[|\mathcal B_l|=2|\mathcal A_{l-1}|+2|\hat{\mathcal A}_{l-2}|=2(l)-2(l-1)=2.\]
+
+	\brk
+	\subsection*{($C_l$)}
+	This isn't bad at all since $|\mathcal C_l|=|\mathcal B_l^T|=|\mathcal B_l|=2$.
+
+	\brk
+	\subsection*{($D_l$)}
+	Here the claim is that $|\mathcal D_l|=4$ (for $l\ge 4$). Here we can again compute the determinant from the 
+	bottom right, to see that 
+	\[|\mathcal D_l|=2|\mathcal A_{l-1}|+|H|\]
+	where $H$ is $\mathcal A_{l-1}$ with the second-to-last row replaced by $(0,\dots,0,-1,0)$.
+	Doing row reductions, we can compute $|H|=-2|\mathcal A_{l-3}|$, and sol
+	\[|\mathcal D_l|=2l-2(l-2)=4.\]
+
+	\brk
+
+	The matrices for the exceptional Lie algebras are just regular matrices and their determinants can 
+	be computed (either by hand or your favorite CAS) :).
+\end{sol}
+
+\begin{hwprob}{13.4}
+	Prove that the Weyl group of a root system $\Phi$ is isomorphic to the direct product of the
+	respective Weyl groups of its irreducible components.
+\end{hwprob}
+\begin{sol}
+
+\end{sol}
+
+\begin{hwprob}{(Board)}
+	Using large tensor powers of the adjoint representation and $\sl_2$ theory, show that any semisimple Lie algebra $L$ admits irreducible representaitons of arbitrarily large dimension.
+\end{hwprob}
+\begin{sol}
+
+\end{sol}
+\newpage
+
+\section{February 15, 2019}
+One last fact about root systems for now:  if we let $\Phi$ be a root system and $g$ an automorphism of $\Phi$
+(orthogonal transformation of $\R^n$ permuting $\Phi$). If $\Phi^+$ is a positive system for $\Phi$, so is $g\Phi^+$ 
+and we can find $w\in W$ sending $g\Phi^+$ to $\Phi^+$. Then $wg$ is an automorphism of $\Phi$ preserving $\Phi^+$,
+thus also the corresponding simple subsystem $\Delta$, thus acting on the dynking diagram by an automorphism.
+
+If $\Phi$ is such that this diagram is connected, then running through all the possibilities, there are just 3 of them 
+where nondrivial diagram automorphisms occur: types $A_n, D_n,$ and $E_6.$
+
+$A_n, E_6,$ and $D_n$ for $n>4$ all admit two automorphisms while $D_4$, most interestingly, admits a group $S_3$ of automorphisms.
+The study of these kinds of things is called triality. It turns up that $S_3$ also acts by automorphisms on $\mathfrak{so}(8)$, the Lie algebra
+of type $D_4$. If you compute the fixed point subalgebra of $D_4$, you end up finding out that it is, in fact, $G_2$. This can be seen in 
+the diagram by ``folding''.
+
+\subsection{(Officially) Returning to Lie Algebras}
+FOr now, let $L$ be any Lie algebra over a field $k$ of characteristic zero admitting a vector space decomposition
+$L=H\oplus_{\alpha\in\Phi\subseteq H^*}L_\alpha$
+such that
+\begin{enumerate}
+	\item $H$ is abelian
+	\item $L_\alpha=\{x\in L:[hx]=\alpha(h)x\}$
+	\item $\Phi$ is a root system
+	\item $\dim L_\alpha=1$; and
+	\item $[L_\alpha,L_\beta]=L_{\alpha+\beta}$ when this makes sense
+\end{enumerate}
+Then $L$ is simple if the Dynkin diagram is connected and semisimple in general.
+
+\begin{rmk}
+	Monty mentioned today that he will not be following the approach in Humphreys -- showing that 
+	any two Cartan subalgebras are conjugate. Instead, he will show this using an algebro-geometric 
+	argument. This means we can skip reading chapters 15 and 16 of Humphreys.
+\end{rmk}
+
+\subsection{Uniqueness of Root Systems}
+Now we want to see that any two semisimple Lie algebras with the same roots systems are isomorphic,
+
+We have the following: let $L,L'$ be two semisimple Lie algebras with roots systems $\Phi, \Phi'$ and suppose that $\pi$ is an isomorphism from 
+$\Phi$ to $\Phi'$ (thus an orthogonal isomorphism from a maximal toral subalgebra $H$ of $L$ to its
+counterpart $H'$ in $L'$). Fix simple subsystems $\Delta$ and $\Delta'$ that correspond under $\pi$
+and choose nonzero $x_\alpha\in L_alpha$ and $x'_{\alpha'}\in L_{\alpha'}'$ where $\alpha'=\pi(\alpha).$
+
+There is a unique isomorphism from $L$ to $L'$ extending $\pi$ and sending $x_\alpha$ to $x_{\alpha'}'$. To
+see this, assume first that $L,L'$ are simple. We know that there are $y_\alpha\in L_{-\alpha}$ and 
+$y_{\alpha'}'\in L_{-\alpha'}'$ such that the $x_\alpha,y_\alpha,h_\alpha$ and the primed 
+versions each span subalgebras $S_\alpha$ and $S_{\alpha'}$ isomoprhic to $\sl_2$.
+
+Then you have this nice trick where you look at $L\oplus L'$ and end up proving they are, in fact, isomorphic. :)
+
+\brk
+
+\begin{thm}
+	Let $g$ be any automorphism of the root system $\Phi$ of a semisiple Lie algebra $L$.
+	Then there is a unique automorphism of $L$ extending $g$.
+\end{thm}
+\begin{rmk}
+	To see this, we notice that $g$ is an automorphism of a maximal toral subalgebra 
+	$H$ of $L$ sending roots to roots. So then we can pick a simple subsystem $\Delta$
+	nonzero $x_\alpha\in L_\alpha$ and $x_{g\alpha}$ in $L_{g\alpha}$ and the above automorphism
+	sends $x_\alpha\mapsto x_{g\alpha}$.
+\end{rmk}
+
+\section{February 20, 2019}
+I missed the first section here but the definition of a Chevalley automorphism was discussed.
+
+In the special case when our automorphism $\pi$ lies in $W$ (where $\pi$ is our simple system automorphism), 
+there is a direct comstruction of $g_\pi$, which we have seen: since $W$ is generated by simple reflections,
+it is enough to do this for $w=s_\alpha$, and then
+\[g_{s_\alpha}=(\exp\ad x_\alpha)(\exp\ad -y_\alpha)(\exp\ad x_\alpha)\]
+
+\subsection{The next logical quesetion}
+Given a semisimple Lie algebra $L$ over an \textbf{algebraically closed} field $k$ of characteristic zero,
+does its root system $\Phi$ depend on the choice of maximal toral subalgebra $H$?
+\begin{thm}
+	No.
+\end{thm}
+\begin{prf}
+	We will prove this using some (very basic) algebraic geometry. There is an elementary (but much more ridiculous)
+	argument in the text. Notice that in either case, the result \textbf{is not true} if $k$ is not algebraically closed.
+
+	We will prove this for a general Lie algebra (not necessarily semisimple). First we will recall some facts from
+	algebraic geometry: in particular the Zariski topology -- recall that here any two nonempty open sets intersect nontrivially.
+	FUrthermore, we have a theory of morphisms from $k^n\to k^m$: in particular given a map $f:k^n\to k^m$ and a point $x\in k^n$,
+	we get the differential $df_x:T_xk^n\to T_{f(x)}k^m$. Finally recall that if such an $f$ is surjective, then it maps
+	an open subset containing $f(x)$.
+
+	Let $L$ be any Lie algebra over $k$, and let $H$ be a maximal toral subalgebra. We have a root space decomposition
+	\[L=L_0\bigoplus_{\alpha\in \Phi} L_\alpha\]
+	and $H\subseteq L_0$. Fix a vector space complement $H'$ to $H$ in $L_0$ and look at the Zariski open subset $H''=\{h\in H':\alpha(h)\ne 0, \alpha\in\Phi\}$.
+	Set $\mathcal H=H''+H_i$, which is Zariski open in $L_0$.
+
+	Now let $b_1,\dots,b_m$ be a basis for $\oplus_\alpha L\alpha$ and we obtained by combining bases of each $L_\alpha$ 
+	separately, and consider the map from $k^m\times L_0$  to $L$ sending
+	\[(k_1,\dots,k_m,h)\mapsto (\exp\ad k b_1)\cdots(\exp\ad k_m b_m)(h)\]
+	this map has a surjective differential at any $h\in\mathcal H$, by dimension counding, so it follows that given two maximal toral subalgebras
+	$H_1,H_2$ of $L$, some automorphism of this type sends a regular element of $H_1$ to a regular element of $H_2$
+	(for which no nonzero --- takes the value zer). But eh centralizer of the centralizer of such an element of $H_1$, if semisimple,
+	lies in $H_2$, and similarly for $H_2$, so $H_1$ is sent to $H_2$.
+\end{prf}
+
+\subsection{Cartan Subalgebras}
+Monty (and Humphreys) have actually be decieving us this whole time: in the literature what we really care about, 
+instead of maximal toral subalgebras, are \textbf{Cartan subalgebras}, which are defined as a nilpotent subalgebra
+which is normal in $L$: $N_L(H)=H$. This definition is actually equivalent to what we've been using
+whenever $L$ is semisimple and in general over any algebraically closed field. FUrthermore any two Cartan subalgebras are
+conjugate y $\Int L$.
+
+\subsection{Classification continued}
+We still must show that given any root system there is a semisimple Lie algebra (over $k=\bar k$)
+which has $\Phi$ as its root system. We will only sketch the proof here -- see section 10 of Humphreys for more
+information. We will instead give a defining set of relations for such a Lie algebra using only $\Phi$.
+
+Notice that if we had $L$ with a root system $\Phi$ then we would have the basis $x_\alpha,y_\alpha,h_\alpha$ for $\alpha\in\Delta$.
+Wow rough morning, huh?
+
+\section{February 22, 2019}
+Last time we learned that any two maximal toral (Cartan) subalgebras  of a semisimple Lie algebra over an algebraically closed field $k$ 
+of characteristic 0 are conjugate under $\Int L$. In particular, $\dim H_1=\dim H_2=l$, which is called the \textbf{rank of $L$.}
+
+We have anothe conjugacy result:
+\begin{lem}
+	Given a Cartan subalgebra $H$ with root space decomposition
+	\[L=H\oplus\bigoplus_\alpha L_\alpha\]
+	and a choice $\Phi^+$ of positive roots in $\Phi$, the subalgebra $B=H\oplus\bigoplus_{\alpha\in\Phi^+}L_\alpha$
+	has a derived subalgebra $\oplus_{\alpha\in \Phi^+}L_\alpha$, which is nilpotent, so $B$ is solvable. In fact
+	$B$ is maximal among solvable subalgebras of $L$.
+\end{lem}
+
+\begin{defn}
+	We call any maximal solvable subalgebra of $L$ a \textbf{Borel subalgebra}.
+\end{defn}
+\begin{rmk}
+	Any two Borel subalgebras are conjugate under $\Int L$. The quotient of $\Int L$ by the normalizer of such a subalgebra $B$ is called the \textbf{flag variety of $L$.}
+\end{rmk}
+
+\subsection{Today}
+We are focusing today on how to construct a semisimple Lie algebra $L$ from aroot system $\Phi$, together with a basefield $k$ of characteristic zero via generators and relations.
+We need some definitions:
+\begin{defn}
+	The \textbf{free Lie algebra} on a set $x_1,\dots,x_n$ over $k$ is a subalgebra of the associative free tensor $k$-algebra
+	generated by the $x_i$. In particular, it is the Lie subalgebra of this algebra generated by the $x_i$ -- spanned by all the $m$-fold brackets
+	with the only dependence relations among the brackets following from the axioms of a Lie algebra (anticommutativity and the Jacobi 
+	identity).
+\end{defn}
+
+We will follow this approach starting from $\Phi$: fix a simple subsystem $\Delta=(\alpha_i)_1^n$, a basis of $\Phi\subseteq\R^n$.
+Then we take the 3n generators $x_\alpha,y_\alpha,h_\alpha$ of $L$ given the relations
+\begin{align*}
+	[h_i,h_j]&=0\\
+	[h_i,x_j]&=\langle \alpha_i,\alpha_j\rangle x_j\\
+	[h_i,y_j]&=-\langle\alpha_i,\alpha,j\rangle y_j\\
+	[x_i,y_j]&=\delta_{ij}h_i
+\end{align*}
+with more relations to follow later.
+
+\brk
+
+First look at hte Lie algebra $L_0$ defined by these relations alone. THis is infinite-dimensional, but still has nice enough structure to lead to what we want.
+The structure of $L_0$ is (as a cector space) $L_0=Y+H+X$, where $Y$ is genreated by the $y_i$, etc. Moreover, the $h_i,x_i,y_i$ are independent in $L_0$,
+which follows from an explicit representation of $L_0$ on page 97 of Humphreys.
+
+$H$ is easy: generators $h)i$ commute, so $H$ is jsut the polynomial ring over $k$ of the $h_i$, regarded as independent variables.
+
+$Y$ and $X$ are harder, but it is easy enough to observe that these are both graded: an $m$-fold bracket of various
+$x_i$ has weight the sum of the corresponding $\alpha_i$, which is an element of $H^*$. Similar for the $y_i$'s.
+
+So we can write $X=\oplus X_\lambda$ where $X_\lambda=\{x\in X:[hx]=\lambda(h)x\}$ and similar for $Y$. Now each nonzero $X_\lambda$ corresponds to a $\lambda$ 
+that is a nonnegative linear combination of $\alpha_i$ and the $Y_\eta$ correspond to $\eta$ which are 
+nonpositive linear combinations. This gets us that
+\begin{lem}
+	\[\dim X_{\alpha_i}=1=\dim Y_{\alpha_i}\]
+	and furthermore if $c\notin\{0,\pm 1\}$,
+	\[\dim X_{c\alpha_i}=0=\dim Y_{c\alpha_i}.\]
+\end{lem}
+
+\brk
+
+So how can we cut down $L_0$ to the size we want? We know that for all $i\ne j$ that $\alpha_i-\alpha_j$ is not a root,
+whence $S_{\alpha_j}(\alpha_i-\alpha_j)=\alpha_i-(\langle\alpha_i,\alpha_j\rangle+1)\alpha j$
+is also not a root. Call it $\alpha_i+k_{ij}\alpha_j$. We know the bracket of $x_i$ with $x_j$ $k_{ij}$ times has to be zero in our quotient, so
+we must impose the further relations
+\[(\ad x_j)^{k_{ij}}x_i=0.\]
+
+If these relations are imposed (along with the correspinding ones for the $y_i$), then the ideal $I$ of $X$ generated by the $x_i$ relations are, in
+fact, an ideal of $L_0$ (Humphreys p. 99) (same for $Y$). These ideals are generated by graded elements, so
+\[(X+H+Y)/(I+J)\]
+retains a graded structure for which the root spaces are still dimension one.
+
+\brk
+
+\begin{defn}
+	An element $x$ where for each element $l\in L$ there exists a positive integer $a_l$ such that
+	\[(\ad x)^{a_l}l=0\]
+	is said to act \textbf{locally $\ad$-nilpotently} on $L$.
+\end{defn}
+So we have seen that $x_i$ and $y_i$ act locally $\ad$-nilpotently, so we get an automorephism of the quotient
+\[(\exp\ad x_i)(\exp \ad -y_i)(\exp\ad x_i)\]
+acts as the reflection $s_{\alpha_i}$ on the root spaces, preserving their dimensions.
+
+It follows that the Weyl group $W$ acts on the root spaces of our quotient, preserving their dimensions.
+Let $L$ be this quotient. We know that $\dim L_\beta=1$ for any $\beta\in\Phi$, since any such $\beta$ is 
+conjugate to a simple root (where the root space has dimension 1).
+
+Finall, if $\lambda\in H^*$, then some $W$ conjugate of $\beta$ is a linear combination of simple roots
+with some positive and some negative coefficients. This doesn't occur in $L_0$.
+
+The upshot here is that $L$ consists of $H$ plus a one-dimensional $L_\beta$ for all $\beta\in\Phi$ and 
+nothing else. By $\sl_2$ theory, we get the bracket relations among root spaces. By a previous result, any Lie algebra with a root 
+system is semisimple, so $L$ is semisimple with root system $\Phi$. This gives us our conclusion:
+\begin{thm}
+	A semisimple Lie algebra $L$ exists with \textbf{any} specified root system $\Phi$ and base
+	field $k$.
+\end{thm}
+
+\subsection{Extending this construction}
+This begs to be embedded in a larger construction: Give $k$ and our generators $x_i$, and an ideal $I$ 
+of $F$ (the free $k$ Lie algebra on the $x_i$), we can form $T/\langle I\rangle$ ($T$ the free tensor algebra), the quotient of $T$ by the 
+\textbf{ordinary two-sided} generated by $I$ in $T$. This will be a larger associative algebra containing $F/I$ such that
+$[x,y]=xy-yx$.
+
+More generally, let $L$ be a finite dimensional Lie algebra with basi $x_i$. Then we have 
+\[[x_ix_j]=\sum m_{ijk}x_k\]
+and imposing the relations on the tensor algebra, and this is called the universal enveloping algebra of $L$.
+
+Sick.
+
+\section{February 25, 2019}
+Notice that (having classified all complex semisimple Lie algebras) we have thus far used lots of linear algebra
+but relatively little abstract algebra. But now the theory of associative algebras is going to come into the fold
+as we proceed to classify the finite dimensional $L$ modules.
+
+By Weyl's theorem, it is enough to classify the finite-dimensional \textit{irreducible} $L$-modules $M$.
+For this we need an asoociaitive algebra containing $L$ as a Lie algera , and then we use the method of associative algebras:
+\begin{ex}
+	If $A$ is such an algebra, then a typical irreducible left $A$-module takes the form $A/M$
+	for some maximal left ideal $M$ of $A$.
+\end{ex}
+
+\subsection{Universal Enveloping Algebras}
+But we need to choose $A$ in such a way that all irreducible finite-dimensional $L$-modules are captured
+in this way (as $A$-modules). For this purpose,
+\begin{defn}
+	Let $\mathcal U:=\mathcal U(L)$ be the associative algebra generated by $L$ (or any basis) with relations exactly tose of $L$,
+	that is $xy-yx=[xy]$.
+
+	There are only finite many relations since it is enough to take $x$ and $y$ above to be elements of a basis of $L$.
+\end{defn}
+
+How big is $\mathcal U$? What is a basis? To this we answer
+\begin{thm}[Poincar\'e-Birkhoff-Witt (PBW)]\label{thm-pbw}
+	If $x_1,\dots,x_n$ is an ordered basis for $L$, then a basis of $U$ (over $k$) is given by 
+	monomials in the $x_i$: 
+	\[x_1^{m_1}\cdots x_n^{m_n},\quad m_i\ge 0\].
+\end{thm}
+\begin{rmk}
+	The proof of this can be found in the book (p.93-94) (or other places). The idea is that, to show span,
+	using that whenever $j>i$, $x_jx_i=x_ix_j+[x_jx_i]$, converting an element in ``wrong'' order into the
+	``right'' order.
+\end{rmk}
+A more sophistochated statement of PBW: Let $\mathcal U_m=\mathcal U_m(L)$ be the span of all products of at most $m$ geneartors $x_i$ (allowing
+repetitions) so that 
+\[0=\mathcal U_{-1}\subseteq\mathcal U_{0}\subseteq\mathcal U_{1}\subseteq\cdots\]
+where $\cup \mathcal U_i=\mathcal U$.
+
+Then $\mathcal U_m\mathcal U_n\subseteq \mathcal U_{m+n}$, so we can form the associated graded algebra
+$G=\oplus_0^\infty \mathcal U_m/\mathcal U_{m-1}$.
+
+In particular, we have $x_1,\dots,x_n$ as a basis for $L$ representa by elements labelled in the same way in $\mathcal U_1/\mathcal U_0$.
+Then $x_ix_j=x_jx_i$ for all $i,j$, regarding the $x_i$ as lying in $G$. Moreover, the images in $G$ generate $G$, so 
+we have a map $\varphi$ from $S$, the polynomial ring in the $x_i$ to $G$, sending each $x_i$ to its image in $G$.
+\begin{thmprime}{thm-pbw}[PBW']
+	In the above construction, the map $\varphi$ is an isomorphism.
+\end{thmprime}
+\begin{rmk}
+	An interesting consequence of this is that $\mathcal U$ has no zero divisors. For if $x,y$ are nonzero elements just choose $m,n$ minimal so that 
+	$x\in\mathcal U_m$ and $y\in\mathcal U_n$. Then the product is nonzero in $\mathcal U_{m+n}/\mathcal U_{m+n-1}$!
+\end{rmk}
+
+It is clear from the construction of $|mathcal U$ that any left $L$ module may be regarded as a left $\mathcal U$-module. Thus we can construct an irreducible such module
+as $\mathcal U/M$ for a maximal (left) module $M$. Now let $N$ be an irreducible $L$-module,
+so that $N$ is an irreducible $\mathcal U$-module.
+
+If $L$ is semisimple, it has root space decomposition $L=\h\oplus_{\alpha\in\Phi}L_\alpha$. Choose a positive root system
+and we ahve seen that the Borel subalgebra $B$ is a (maximal) solvable subalgebra for $L$. Thus by Lie's theorem, there is a 
+one-dimensional submodule $K_v$ of $L$, acted on semisimple (thus diagonally) by $\h$,
+while $x_\alpha v=0$ for all $\alpha\in\Phi^+$, since $L_\alpha\in [BB]$.
+
+Accordingly, $v$ is a weight vector in $N$ (simultaneous eigenvector for $\h$) and so now we bring in PBW:
+we know by irreducibility that $\mathcal U v=N$. Now fix a basis of $L$ consisting of root vectors, together 
+with a basis for $\h$, in which all negative root spaces come first, then the basis for $\h$, then the basis for $[BB]=\oplus_{\alpha\in\Phi^+}L_\alpha.$
+
+We have observed that $N$ (since it is irreducible) is a quotient of $\mathcal U$ by a maximal left ideal. We 
+know this ideal contains all $x_\alpha$ for $\alpha\in\Phi^+$. It follows that if $N'=\oplus_{\alpha\in\Phi^-}$,
+then $\mathcal Uv=\mathcal U(N')v$, so all vectors in $N$ are obtained from $v$ by applying suitable products of 
+negative roots vectors and taking the span.
+
+It follows that our module $N$, the sum of its weight spaces, is such that any of its weights $\mu\in\h^*$
+is such that $\mu$ is obtained from $\lambda$ (where $hv=\lambda(h) v$ for all $h\in\h$) by subtracting anonnegative integer combination
+of simple roots. This says that the weight $\lambda$ of $v$ occurs with multiplicity one (the only product $P$ of negative root vectors such that
+$Pv$ has weight $\lambda$ is the empty product) and all other weights occurring (whoops, it got erased.)
+
+\section{February 27, 2019}
+Last time we saw that, given any semisimple Lie algebra, we can identify a \textbf{highest weight vector} of any $L$-module $V$
+which is killed by the action of any positive root. We also saw (via $\sl_2$ theory) that this highest weight $\lambda$ satisfies
+$\langle\lambda, \alpha_i\rangle\in\N$ for all $i$.
+
+We also saw the enveloping algebra $\mathcal U$ of $L$ and that $V=\mathcal Uv$ and $\h v=kv$. Let $I$ be the ideal of $\mathcal U$ gnerated by $N^+$,
+the sum of the positive weight spaces, and all $h-\lambda(h)$. Notce $I\subseteq M=\operatorname{Ann}(V).$
+Notice that any proper ideal of $\mathcal U$ cannot contain a vector of weight zero (since the only one is 1)
+and thus the sum of every ideal containing $I$ is proper, whence there is a unique maximal $M$ over $A$.
+
+Hence immediately if $V$ and $W$ are irreducible modules generated by vectors of the same highest weight $\lambda\in\h^*$,
+then $V\cong W\cong \mathcal U/M$.
+
+\subsection{Integral Weights}
+It remains to show the following:
+\begin{lem}
+	If (conversely) $\lambda$ is such that $\langle\lambda,\alpha_i\rangle\in\N$ for all $i$, 
+	then there is a unique irreducible highest weight module $V$ with highest weight $\lambda$ 
+	that, furthermore, is finite dimensional!
+\end{lem}
+\begin{rmk}
+	Such a $\lambda$ (regardless of the context in which we see it) is called a \textbf{dominant (integral) weight.}
+\end{rmk}
+\begin{prf}
+	We will construct $M$ explicitly and show $\mathcal U/M$ is finite dimensional. Set
+	\[m_i=\langle\lambda,\alpha_i\rangle\]
+	and let $J$ be the ideal generated by the $y_i^{m+i+1}$ where $y_i\in L_{-\alpha_i}$. We claim that $M=I+J$. To see this, note first that
+	$I+J$ is generated as a left $\mathcal U$=module by $I$  and $\mathcal U(N^+)J$ since $x_jy_j^{m+j+1}=y_j^{m_j+1}$ if $i\ne j$ where $x_j\in L_{\alpha_j}$
+	and $x_jy_j^{m_j+1}\in I+J$ by a $\sl_2$ calculation.
+
+	Thus $I+J$ is proper and $U/(I+J)$ has a finite-dimensional $S_{\alpha_i}$-module spanned by $v, y_iv,\dots, y_i^{m_i}v$
+	where each $S_{\alpha_i}$ is spanned by the $x_i,y_i,h_i$. Eventually we can prove that $U/(I+J)$ is stable under 
+	the Weyl group, having the same dimenion of its weight space.
+
+	Then given any weight $\eta$ smaller than $\lambda$ in the partial order, recall that, if the dot product with any simple root 
+	is negative, one can apply a reflection to "bump up" $\eta$ in the partial order. But of course you can only do this so many times!
+
+	Thus any dominant weight $\mu$ of $V'=\mathcal U/(I+J)$ has length at most that of $\lambda$ and it lies on the integer lattice,
+	so as a discrete subset of a compact space there are only finitely many.
+
+	Thus any weight occuring in $V'$ is $W$-conjugate to some dominant weight, so only finite many weights occur in $V'$ 
+	and therefore $V$ is finite-dimensional. Furthermore, we see that $V$ has only one highest weight vector (if it is irreducible and finite-dimensional). 
+\end{prf}
+
+Next time we will try to explain the structure of $V$ as it decomposes into $V_\lambda$ which is analgous to, but not
+exactly the same as what we see in the adjoint representation: in this case, we will see that one can get higher multiplicities
+occuring for certain weights, but the pattern will be that the highest weights will occur with multiplicity one and the rest 
+will lie in the convex hull of these vectors, with multiplicity increasing as you go further inside the hull.
+
+\section{February 28, 2019}
+
 \end{document}

### 2019-03-02 Reading Notes/reading_notes.tex
Updates

@@
 \usepackage{setspace}
-\usepackage{amsmath, amsfonts, amssymb, graphicx, color, fancyhdr, scalerel, mathrsfs, tikz-cd, mdframed, enumitem, framed, adjustbox, bm, upgreek, xcolor}
+\usepackage{amsmath, amsfonts, amssymb, graphicx, color, fancyhdr, scalerel, mathrsfs, tikz-cd, mdframed, enumitem, framed, adjustbox, bm, upgreek, xcolor, hyperref}
+\usepackage[style=alphabetic]{biblatex}
 \usepackage[framed,thmmarks]{ntheorem}
@@
 
+%questions
+\theoremheaderfont{\sc}
+\newframedtheorem{qst}{Question}[subsection]
+%\newtheorem{qst}[thm]{Question}
+
 %waterhouse problem
@@
 \DeclareMathOperator{\HH}{HH}
+\DeclareMathOperator{\Gr}{Gr}
 
@@
 \DeclareMathOperator{\coker}{coker}
+\DeclareRobustCommand{\sqbinom}{\genfrac[]{0pt}{}}
 
@@
 \newcommand{\qbinom}{\genfrac{[}{]}{0pt}{}}
+
+%Set the bibliography file
+\bibliography{sources}
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \part{Winter 2019}
@@
 \begin{itemize}
-	\item Ginzburg and Kumar '94 ``Cohomology of Quantum Groups at Roots of Unity''
-	\item Mastnak, Pevtsova, Shauenburg, and Witherspoon '09 ``Cohomology of Finite Dimensional Pointed Hopf Algebras''
-	\item Witherspoon '17 ``Varieties for Modules of Finite Dimensional Hopf Algebras''
+	\item Ginzburg and Kumar '94 ``Cohomology of Quantum Groups at Roots of Unity'' \cite{ginzburg-kumar}
+	\item Mastnak, Pevtsova, Shauenburg, and Witherspoon '09 ``Cohomology of Finite Dimensional Pointed Hopf Algebras''\cite{julia10-pointed}
+	\item Witherspoon '17 ``Varieties for Modules of Finite Dimensional Hopf Algebras''\cite{witherspoon-expository}
 \end{itemize}.
@@
 \subsection{Week 3}
-We primarily discussed support varieties in the context of Sarah Witherspoon's \textit{Varieties for Modules of Finite Dimensional Hopf Algebras.}
+We primarily discussed support varieties in the context of Sarah Witherspoon's \textit{Varieties for Modules of Finite Dimensional Hopf Algebras.}\cite{witherspoon-expository}
 
@@
 
+\subsection{Week 6}
+We skipped Lie algebras this week because of snow (and then Ale couldn't make our meeting). So instead
+Julia just met up to talk about my reading.
+
+My primary question (regarding the Ginzburg and Kumar paper): why are we factoring out the toral elements in quantum groups? 
+The answer here is that these are called the ``small quantum groups''. So certainly having a finite-dimensional algebra makes certain things
+more tractable, although when computing cohomology that isn't always the case. Sometimes when you get smaller 
+objects you end up getting more nontrivial elements in the cohomology. There's a clear analog to topology here.
+
+So then maybe a refinement of the question is this: why is it necessary at all to have this larger object serving as the ambient 
+space for the object we truly want to study? Here the answer more-or-less boils down to the fact that it end up playing a central
+role in computations involving these objects.
+
+Having the big algebra $U$ over $u$ (modulo something central) gives you that thePBW basis for $U$ descends to 
+$u$. In general (e.g. Nichols algebras) you fail to have a PBW basis for the cover, but somehow (for a technical reason that 
+will become apparent as we read more) you still can use similar techniques to do the computations (the more general theory arises in the 2010 paper).
+
+Julia is going to sent me a paper that is a survey of finite dimensional algebras that will give some sense for how the small quantum groups fit in.
+
+\subsubsection{This week's assignment}
+Continue reading through GK until we get to the spectral sequence nonsense. Then this points to Jantzen (AG). Try to recreate the
+argument for constructing the first page of the spectral sequence, but for the GZ case (note, you'll use projectives instead of injectives)
+After that, go to Weibel and try to read the section of spectral sequences over filtered complexes \S 5.4 (what we are usually interested in).
+Don't worry too much about issues of convergence -- we are mostly interested in the ``first quadrant'' case.
+
+David Anick is the name of an algebraic topologist who started off created (free) resolutions that we will end up looking at. 
+His paper \cite{anick-resolution} will probably be useful at some point.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \newpage
@@
 \subsection{The more general case.}
-This primarily borrows from \S 21 in Humphreys (I) but also relies on some results from earlier that I will cite as 
+This primarily borrows from \S 21 in \cite{humphreys1} but also relies on some results from earlier that I will cite as 
 necessary.
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \newpage
-\section{Quantum Groups}
-I am mostly working out of Jantzen's \textit{Lectures on Quantum Groups} with the goal of reading through Ginzburg \& Kumar's
-(apparently influential!) paper \textit{Cohomology of Quantum Groups at Roots of Unity.}
+\section{Quantum Groups}\label{sec-quantum}
+I am mostly working out of Jantzen's \textit{Lectures on Quantum Groups} (\cite{janzten-quantum}) with the goal of reading through Ginzburg \& Kumar's
+(apparently influential!) paper \textit{Cohomology of Quantum Groups at Roots of Unity} (\cite{ginzburg-kumar}).
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\newpage
+\section{Reading Ginzburg \& Kumar}
+This section contains my notes and thoughts about \cite{ginzburg-kumar}. As a high-level summary, this paper establishes
+methods for computing (Hochschild) cohomology of the small quantum groups. It begins by establishing how they are constructed
+as quotients or subalgebras of quantized universal enveloping algebras (see section~\ref{sec-quantum}), and then proceeds to compute
+the cohomology using a filtration (this is vital!) and spectral sequences.
+
+This paper is important because it ends up that this general formula is the same as is applied later in \cite{julia10-pointed} to prove
+a much more general result. This also extends a method originally created by \cite{FP} for computing the cohomology of Lie algebras and
+algebraic groups.
+
+\subsection{The Main Result}
+In this paper they prove the following (see \ref{sec-defs} for notation):
+\begin{thm}[GK '93]
+	If $k=\bbQ(\xi)$ for a primitive root of unity of odd order (not 1) and $u_\xi$ the restricted enveloping algebra of some $\g$. Then
+	\[H^{odd}(u_\xi,k)=0\]
+	and there is a natural graded algebra isomorphism
+	\[H^{ev}(u_\xi,k)\cong k^\bullet[\mathcal N].\]
+\end{thm}
+
+\subsection{The Objects in Question}
+We begin by considering the objects in question and how they are constructed. One thing we haven't seen (yet!) in Lie algebras
+but comes up in the structure theory is the 
+\begin{defn}[Kostant $\bbZ$-form]\label{def-kostant}
+	Let $U(\g)$ be the universal enveloping algebra of $\g$, spanned (over $k$) by the $x_\alpha,y\alpha,$ and $h_\alpha$ for all $\alpha\in\Phi$. 
+	Then the \textbf{Kostant $\bbZ$-form} is the $\bbZ$-span of the elements
+	\[X^{(r)}_\alpha=\frac{x_\alpha^r}{r!},\quad Y^{(r)}_\alpha=\frac{y_\alpha^r}{r!},\quad H^{(r)}_\alpha=\frac{h_\alpha^r}{r!},\quad \]
+	the so-called \textbf{divided powers} of $x_\alpha, y_\alpha,$ and $h_\alpha$. This ($\bbZ$-)algebra is denoted $U_\bbZ$.
+\end{defn}
+
+Before we use it, we had better write down the definiton (via \cite{janzten-quantum}) of a
+\begin{defn}[Quantized Enveloping Algebra]
+	Fix some odd root of unity $q$ in the field\footnote{Here Jantzen also requires that $\ch k\ne2$ and that $\ch k\ne3$ if $\Phi$ has an irreducible component of type $G_2$.
+These can be treated, but not in quite the same manner.} $k$.
+	Let $\g$ be a Lie algebra over $k$ and let $U_q(\g)$ denote the $k$ algebra with generators $E_\alpha,F_\alpha, K_\alpha,$ and $K_\alpha^{-1}$
+	for all $\alpha$ in a basis $\Pi$ of $\Phi$, subject to the relations for all $\alpha,\beta\in\Pi$
+	\begin{align*}
+		K_\alpha K_\alpha^{-1}=&1=K^{-1}_\alpha K_\alpha\\
+		K_\alpha K_\beta &= K_\beta K_\alpha\\
+		K_\alpha E_\beta K_\alpha^{-1} &= q^{(\alpha,\beta)}E_\beta\\
+		K_\alpha F_\beta K_\alpha^{-1} &= q^{-(\alpha,\beta)}F_\beta\\
+		E_\alpha F_\beta - F_\beta E_\alpha &= \delta_{\alpha\beta}\frac{K_\alpha-K_\alpha^{-1}}{q_\alpha-q_\alpha^{-1}}\\
+		\sum_{s=0}^{1-a_{\alpha\beta}}(-1)^s\sqbinom{1-a_{\alpha\beta}}{s}_{q_\alpha} &E_\alpha^{1-a_{\alpha\beta}-s}E_\beta E_\alpha^s = 0\\
+		\sum_{s=0}^{1-a_{\alpha\beta}}(-1)^s\sqbinom{1-a_{\alpha\beta}}{s}_{q_\alpha} &F_\alpha^{1-a_{\alpha\beta}-s}F_\beta F_\alpha^s = 0
+	\end{align*}
+	where $q_\alpha=q^{d_\alpha}=q^{(\alpha,\alpha)/2}$ and $a_{\alpha,\beta}=\langle\alpha,\beta\rangle=2(\alpha,\beta)/(\alpha,\alpha)$.
+\end{defn}
+
+\begin{rmk}
+	In \cite{janzten-quantum}, he mentions that he will often work in the algebra that includes all the above relations
+	but \textbf{omitting the last two}. Apparently this will be done for computational simplicity and then studying $U_q(\g)$ 
+	itself will amount to taking a quotient. This may not be relevant to me, but it's good to keep in mind for this reference.
+\end{rmk}
+
+\brk
+
+For the case of quantum groups, we are extending all of our notions to quantized enveloping algebras in the case when our field $k$ is the rational function field
+$\bbQ(v)$.
+\begin{defn}[$q$-analog of the Kostant $\bbZ$-form]
+	Let $U=U_q(\g)$ be the (Hopf!) algebra described above. Then we define the $q$-analog of the Kostant $\bbZ$-form
+	to be the $\bbZ[v,v^{-1}]$-span of the $q$-divided powers the $E_\alpha,F_\alpha$ and $K_\alpha^{\pm 1}$, which are the same as
+	the divided powers in definiton~\ref{def-kostant} except the factorial is replaced by the $q$-analog thereof.
+\end{defn}
+
+\subsubsection{Notation and Related Algebras}\label{sec-defs}
+But this isn't what we're studying yet! (This are the so-called ``big'' quantum groups: what Drinfeld studies in \cite{drinfeld}). 
+We are interested in taking particular quotients and subalgebras. Here we define the notation used throughout and how they are related to $U=U_q(\g)$.
+
+We will be interested primarily in the case when $v=\xi$, that is, when $k$ is the cyclotomic field over $\bbQ$ found by adjoining
+an \textbf{odd} primitive $l^{th}$ root of unity ($l$ not 1). For the rest of this paper, fix such a $k$. Then define
+\[U_\varepsilon :=(k\otimes_{\bbZ[v,v^{-1}]}U_\bbZ)/(K_\alpha^l-1)=:U_k/(K_\alpha^l-1)\]
+where we defined $U_k$ by extension of scalars from $U_\bbZ$ and we are using here the fact that any $\bbZ[\xi]$-module 
+becomes a $\bbZ[v,v^{-1}]$-module by letting $v$ act by $\xi$.
+
+\brk
+
+Another construction is the so called \textbf{restricted enveloping algebra} $u_\xi$ of $U_\xi$ generated by the 
+degree one generators $F_\alpha, E_\alpha$, and $K^{\pm 1}_\alpha$. This is a finite-dimensional algebra of dimension $l^{\dim \g}$.
+This is actually the object that we are studying here. It amounts to restricting coefficients in $U_\xi$ to polynomials in $\xi$ with integer coefficients.
+
+\brk
+
+Define $\mathcal N$ to be the \textbf{nilpotent cone} of $\g$, consisting of all $\ad$-nilpotent elements in $\g$. This admits a 
+natural graded structure over the nonnegative integers by recognizing it as the ring of regular functions on a cone.
+
+\subsubsection{Another formulation}
+
+There is a parallel construction of $u_\xi$ that highlights a ``duality'' property of these objects. Set $\mathscr U_\bbQ$ to be the 
+$\bbQ[v,v^{-1}]$-span of $E_\alpha,F_\alpha,K^{\pm 1}_\alpha$ (notice we are not using divided powers here, but allowing coefficients in $\bbQ$) 
+and extend scalars again to $\mathscr U_k$. Again let $\mathscr U_\xi$ be the quotient
+where we identify $K_\alpha^l=1$. Define $\mathscr L$ to be the subalgebra of $\mathscr U_\xi$ generated by $E_\alpha^l$ and $F_\alpha^l$.
+
+Then it is a result of Lusztig that $\mathscr L$ and $u_\xi$ are normal (algebras) in $\mathscr U_\xi$ and $U_\xi$, respectively,
+and that 
+\[\mathscr U_\xi//\mathscr L\cong u_\xi\qquad\text{and}\qquad U_\xi//u_\xi\cong U(\g)\]
+where
+\begin{defn}
+	Given an augmented algebra $B$ with augmentation ideal $B_+$, a subalgebra $A\subseteq B$ is called \textbf{normal} in $B$ if 
+	\[B\cdot (A\cap B_+)=:B\cdot A_+=A_+\cdot B.\]
+\end{defn}
+
+(My) intuition for augmented algebras come from group algebras $kG$, where you have the augmentation map $\epsilon(\sum_G c_gg)=\sum_G c_g$.
+Here the augmentation ideal is $\ker\varphi$. In fact, any map $\epsilon:A\to k$ you have induces such a structure. Usually in the case of Hopf algebras
+we pick the counit (as we do in group algebras).
+
+\begin{defn}
+	If $A$ is normal in $B$, then we write 
+	\[B//A:=B/(B\cdot A_+)=B/(A_+\cdot B).\]
+\end{defn}
+
+The duality we mentioned earlier comes from the fact that $\mathscr L$ is a commutative Hopf algebra that is ``dual'' to (the cocommutative) $U(\g)$.
+
+\subsubsection{Triangular Decompositions}
+We use superscripts $+,-,$ and $0$ to denote the subalgebras generated, respectively, by the $E$'s, $F$'s, and $K$'s\footnote{with the exception of $U_0$ which has another set of generators due to the divided powers}.
+
+Then for $U_\xi$, $\mathscr U_\xi$, and $u_\xi$ we get the standard triangular decompositions; e.g.
+\[u_\xi = u_\xi^+\oplus u_\xi^0\oplus u_\xi^-.\]
+
+For each, define the \textbf{Borel part} to be (e.g.) $\mathscr B_\xi=\langle K_\alpha, E_\alpha\rangle$ (i.e. the nonnegative part).
+
+\subsection{The Argument}
+\subsubsection{Filtrations}
+The core idea (according to Julia) is the existence of certain filtrations such that the associated graded algebras have nice structure.
+
+In this paper, we are using an earlier result in \cite{de-concini} which tells us that $\mathscr B_\xi$ admits such a filtration:
+(Recall that we are using roots $\alpha$ in some basis $\Pi$ of $\Phi$)
+\begin{thm}[DK '90]
+	The algebra $\mathscr B_\xi$ has a multiplicative filtration such that $\Gr \mathscr B_\xi$ is generated by the homogeneous elements $\{E_\alpha,K_\beta\}_{\alpha\in \Delta_+,\beta\in \Pi}$
+	subject to the relations
+	\[K_iK_j=K_jK_i\quad K_i^l=1\quad K_iE_\alpha=\xi^{\langle \alpha,\alpha_i\rangle} E_\alpha K_i\]
+	and
+	\[E_\alpha E_\beta=\xi^{\langle\alpha,\beta\rangle}E_\beta E_\alpha\]
+	if $\alpha\succ\beta$.
+\end{thm}
+Then we can use this filtration to get filtrations on all quotients and subalgebras, most importantly
+$\mathscr B_\xi,\mathscr U_\xi^+, b_\xi,$ and $u_\xi^+$.
+\begin{qst}
+	What is the filtration here? 
+\end{qst}
+I ended up finding the book in which \cite{de-concini} is published (in the library) and the answer is that you look at 
+monomials 
+\[M_{k,r;u}=F^kuE^r\]
+where $F^k=\prod_\alpha F_\alpha^{k_{\alpha}}$ and similar for $E^r$ and $u\in\mathcal U^0$.
+
+Then you can define the \textbf{height:}
+\[\operatorname{ht}(M_{k,r;u})=\prod_\alpha (k_\alpha+r_\alpha)\operatorname{ht}\alpha\in\bbZ_+\]
+(What is $\operatorname{ht}(\alpha)?$) and then finally define \textbf{degree:}
+\[d(M_{k,r;u})=(k_N,k_{N-1},\dots,k_1,r_1,\dots,r_N,\operatorname{ht}(M_{k,r;u}))\in\bbZ_+^{2N+1}\]
+which is totally ordered via the lexicographic ordering and therefore gives us a $\bbZ^{2N+1}$ grading.
+
+\brk
+
+Then we proceed to prove that $H^\bullet(\Gr\mathcal U_\xi^+)\cong\Lambda_\xi$, which follows from some arguments that can be found 
+in \cite{priddy}.
+\begin{qst}
+	What role do augmented algebras play in this theory? I have noted on several occasions that augmentations
+	have been used to define normal subalgebras. It also popped up in the discussion of Koszul algebras in \cite{priddy}.
+\end{qst}
+
+The idea here is to give $k$ the structure of an $A$-module. This is non-trivial for the folloiwng reason:
+consider $A=k[x_i]$ and localize at a non-maximal prime $p$. Then the residue field will have positive transcendence degree
+over $k$, and since $p$ will now be the (only!) maximal ideal, this will be the smallest field to which there is such a map.
+
+\begin{qst}
+	In the computation of cohomology for $\mathcal L^+$, why do the terms in the Koszul complex
+	look like $\mathcal L^+\otimes_k \Lambda^i(\mathfrak N)$? Is this exploiting somehow that 
+	all Hopf modules are trivial (e.g. $\mathcal L^+$) determines the entire action by 
+\end{qst}
+
+Let's take a regular sequence $(x_1,\dots,x_n)$. Then $R\xrightarrow{x_1} R$ is $K(x_1)$, the 
+Koszul complex for this regular subsequence. If we have two elements, then $K(x_1)\otimes K(x_2)$
+is the complex $R\to R\oplus R\to R$. Corollary 4.5.5 in Weibel says that whenever the sequence is regular, if you 
+construct the entire Koszul complex to get a resolution of the quotient $R/(x_i)$.
+
+Note that this essentially boils down to the computation that $\Lambda^iS^n\cong \Lambda^i(k^n\otimes_k S)=(\lambda^i k^n)\otimes S$.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \newpage
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \newpage
@@
 \section{Problems}
-\subsection{Week 1}
 
@@
 	\[\varphi(X)\otimes\varphi(1)+\varphi(1)\otimes\varphi(X)=\varphi(X)\otimes\varphi(X)\]
-
 \end{sol}
@@
 \section{To be filed:}
+The last paper on Julia's page is a preprint with her and Dave Benson. There is a spectral sequence computation that might be illustrative.
+There is also an older paper...
+
+One of the things Julia is interested in at the moment is a super group scheme. These arise by having a $\bbZ_2$ grading on your Hopf algebra.
+For instance we can look at exterior algebra $\Lambda(x)$, which represents the super group scheme $G_a^-$.
+
+The first non-elementary example is $G_a^-\times G_a^-$ with a $\bbZ_p$ action on it that ``folds'' the second factor
+onto the first. To write this out explicitly, we consider the group algebra $\Lambda(u,v)$ where $uv=-vu$ and the coordinate algebra of $\bbZ_p$
+is $k[t]/t^p$ where $t\cdot u=0$ and $t\cdot v=\alpha u$ for some $\alpha$. This is a $4p$ dimensional algebra.
+
+Then we want to compute coholomology. We know $H^{**}(k\bbZ/p,k)\cong k[x]\oplus\Lambda(y)$ for $|x|=2,0$.
+Note there is a notion of dimension that is an ordered pair $(a,b)$ of the graded degree and cohomology degree.
+THis give us new elements of degree $(1,1)$
+
+Look up Koszul duality! It is in the representations and local cohomology book!
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%  Bibliography %%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\medskip
+
+\printbibliography
 

### 2019-03-12 Hopf Algebra Notes/Hopf Algebras.tex
Fortnightly Update

@@
 In particular, notice that $x(1)=0$ and $x(ab)=x(a)b+g(a)x(b)$.
+
+\section{March 11, 2019}
+We took a week off since James was in Germany for a conference. We're coming back together for this
+last week of the quarter.
+
+Here is one result that was brought up at the conference. It was proved by a grad student and might 
+have some missing hypotheses:
+\begin{thm}
+	If $\ch k=0$ (and some other mild hypotheses) then the Etingof-Ostrik conjecture holds for $H$ 
+	if and only if it holds for $H^*$. That is,
+	\[\operatorname{kdim}\oplus \Ext_H^i(k,k)=\operatorname{kdim}\oplus \Ext^i_{H^*}(k,k)\]
+	where $\operatorname{kdim}$ means Krull dimension.
+\end{thm}
+
+\subsection{Example of the Day}
+\begin{ex}[Small Quantum Group of $\sl_2$]
+	Let $q\in k$ be a primitive $l^{th}$ root of unity ($l\ge 2$). Then
+	\[u_q(\sl_2)=k\langle K,K^{-1},E,F\rangle/R\]
+	where $R$ is the ideal for the relations
+	\begin{align*}
+		KK^{-1}=K^{-1}K=1\\
+		KEK^{-1}=q^2E\\
+		KFK^{-1}=q^{-2}F\\
+		EF-FE=\frac{K-K^{-1}}{q-q^{-1}}\\
+		E^l=F^l=K^0\\
+		K^l=1
+	\end{align*}
+	Now $\dim_k u_q(\sl2)=l^3$ and this is (symmetric) Frobenius and noncommutative and noncocommutative.
+	As a coalgebra, $K$ is grouplike and $E,F$ are almost primitive:
+	\[\Delta E=E\otimes 1+K\otimes E,\quad \Delta F = F\times K^{-1}+1\otimes F\]
+	and $S(E)=-K^{-1}E$ and $S(F)=-FK$.
+
+	Then $\sl_2=ke\oplus kf\oplus kh$ and there is a map $e\mapsto E$, $f\mapsto F$ and $K=\exp(h)$. Hmm
+\end{ex}
+
+\subsection{Last Lecture}
+I wasn't here, so I missed this part. Let $A$ be a Noetherian connected graded AS-regular algebra and let
+$V$ be a Hopf algebra action on $A$ such that each $A_i$ is a left $V$-module.
+\begin{enumerate}
+	\item $k:=A/A_{\ge 1}$ and $\oplus_0^\infty \Ext_A^i(k,k)$ is an algebra.
+	\item (S. Paul Smith's Result) $\oplus \Ext_A^i(k,k)$ is Frobenius and $\Ext_A^d(k,k)$ is one
+	dimensional over $k$ when $d=\gldim A$.
+	\item  Each $\Ext^i_A(k,k)$ is a left $V$ module.
+	\item $(\oplus \Ext_A^i(k,k))^{op}$ is a left $V$ algebra.
+\end{enumerate}
+
+Now pick any nonzero element $e\in \Ext_A^d(k,k)$ where $d$ is as above. For all $v\in V$, 
+$v\cdot e\in \Ext_A^d(k,k)$ and we write $v\cdot e=\operatorname{hdet}(v)v$ for $\operatorname{hdet}(v)\in k$.
+\begin{defn}
+	A map $\operatorname{hdet}:V\to k$ that fits in as above is called the \textbf{homological determinant}
+	of the $V$-action on $A$.
+\end{defn}
+\begin{lem}
+	$\operatorname{hdet}$ is an algebra map.
+\end{lem}
+\begin{lem}
+	If $A=k[x_1,\dots,x_n]$, $V=kG$, $G\subseteq GL(n)$ then
+	\[\operatorname{hdet} g = \det g\]
+	for all $g\in G$.
+\end{lem}
+\begin{defn}
+	We say that $\operatorname{hdet}$ is trivial if $\operatorname{hdet}(v)=\varepsilon(v)$ for all $v\in V$.
+\end{defn}
+
+Recall the theorem
+\begin{thm}[Watanabe]
+	Suppose $G$ is a finite subgroup in $GL_n$. If $\deg g=1$ for all $g\in G$, then $k[x_1,\dots,x_n]^G$ is Gorenstein.
+\end{thm}
+\begin{thm}[Noncommutative Watanabe]
+	Let $A$ be a Noetherian connected graded AS-regular algebra. Let $V$ be a semisimple Hopf algebra acting on 
+	$A$ such that each $A)i$ is a left $V$-module. If $\operatorname{hdet}$ is trivial, then $A^V$ is AS Gorenstein.
+\end{thm}
+\begin{rmk}
+	This last result is important because it distinguishes between the commutative and noncommutative cases: in the commutative
+	case the only AS-regular ring is the polynomial ring. This shows us that there is in fact a vast class of them 
+	in the noncommutative case. In fact it's even bigger!
+\end{rmk}
+Recall that $g\in GL_n$ is a reflection if $g$ is similar to $\operatorname{diag}(1,\dots,1,\sqrt{-1})$. Let $R$ be the subgroup of $G$ generated
+by the reflections in $G$. 
+\begin{thm}[Watanabe II]
+	Suppose $G$ is a finite subgroup of $GL_n$. Then $k[x_1,\dots,x_n]^G$ is Gorenstein if and only if
+	$\det \bar g=1$ for all $g\in G/R$.
+\end{thm}
+
+\subsection{Noncommutative Reflections}
+\begin{defn}
+	Let $g\in\Aut_{gralg}(A)$. Then
+	\[\operatorname{Tr}(g)=\sum_0^\infty \operatorname{tr}(g|_{A_i})t^i\in k[[t]].\]
+\end{defn}
+Then a result (which?) gets us that $\operatorname{Tr}(g)=\frac{1}{p(t)}$ for a polynomial $p$.
+Consider $\tilde A=k_{p_{ij}}[x_1,\dots,x_n]$ where $x_jx_i=p_{ij}x_ix_j$ for all $i<j$. Then 
+\begin{defn}
+	$g\in\Aut_{gralg}(\tilde A)$ is a reflection if $\operatorname{Tr}(g)=\frac{1}{(1-t)^{n-1}(1-\lambda t)}$.
+\end{defn}
 \end{document}

### 2019-03-12 Lie Algebra Notes/Lie Algebras.tex
Fortnightly Update

@@
 %characteristic
-\DeclareMathOperator{\ch}{char}
+%\DeclareMathOperator{\char}{char}
+\DeclareMathOperator{\ch}{ch}
 
@@
 \section{February 28, 2019}
-
+Had an appointment. :(
+
+\section{March 4, 2019}
+Notice that we should read sections 21 and 24 of the text, but not 22 or 23. There are notoriously hard sections and it will end 
+up that we don't actually need 23, using the appendix to section 24.
+
+Last itme we identitified the highest weight of the definitng representation $V=k^n$ of any classical Lie algebra
+$L\subseteq \gl_n(k)$. The highest weight is $e_1$ in essentially all cases of type $A_{n-1}$ -- modify to
+($\frac{n-1}{n},\frac{-1}{n},\dots, \frac{-1}{n}$).
+
+\subsection{Creating lager representations}
+We have see the tensor product $\otimes V_i$ of representations $V_i$ over a Lie algebra $L$ that has a Lie sructure defined by 
+\[x(v_1\otimes v_n)=\sum_{i=1}^nv_1\otimes\cdots\otimes xv_i\otimes\cdots\otimes v_n.\]
+
+Certain quotients of $T^nV$, the $n$-fold tensor product, also inherit an $L$ module structure. Specifically, $S^nV=T^nV$ modulo
+the ideal generated by $v\otimes w-w\otimes v$ intersected with $T^nV$. These are the homogeneous polynomials of degree $n$ on a basis of $V$
+as an $L$-module. Similarly one can corrects the exterior power $\Lambda^nV$.
+
+Recall that $S^nV$ grows in dimension while the exterior power grows like binomial coefficients.
+
+In particular, any classical Lie algebra $L\subseteq\gl_n(k)$ acts on $S^mk^n$ and $\Lambda^mk^n$. Luckily
+there is one phrase that sums up how $L$ acts:
+
+\textit{All summetric and exterior powers of $k^n$ in this situation are irreducible, modulo the obvious relations.}
+
+\begin{itemize}
+	\item For $A_{n-1}$ ($\sl_n(k)$), all powers $S^mk^n$ and $\Lambda^mk^n$ are irreducible.
+	\item For $B_n$ and $D_n$, recall that the corresponding group $\so_n(k)$preserves a symmetric bilinear nondegenerate
+	form, which may be thought of as an element of $S^2k^n$. Thus $S^mk^n$ contains a copy of $S^{m-2}k^n$ and then
+	$S^mk^n/S^{m-2}k^n$ is irreducible under the $L$-action, identifying $S^{m-2}k^n$ with its copy in $S^mk^n$. the highest weight is $me.$
+	\item For type $C$, the exterior power $\Lambda^mk^n$, as long as $2\le m\le\frac{n}{2}$. Then the underling group $\sp_n(k)$ preserves
+	a skew-symmetric nondegenerate bilinear form, whence $\Lambda^mk^n$ contains acopy of $\Lambda^{m-2}k^n$ and the corresponding 
+	quotient is irreducible. The highest weight is $m$ 1's followed by $n-m$ zeros.
+\end{itemize}
+
+More generally, there is a purely multilinear algebraic construction starting from $V=k^n$ which produces all irreducible finite dimensional representations in 
+types $A$ and $C$. For $B$ and $D$, we need an additional ingredient of the half-splin representations seen in hw.
+
+\subsection{A Road Map}
+We now head towards a road map of all finite dimensional irreducible represetnations of a semisimple Lie algebra $L$. That is, recall 
+that every such representation $V$ is the sum $\oplus V_\mu$ of its weight spaces, each occuring with some multiplicity $m_\mu=\dim V_\mu$.
+
+We want an implicite formulat for all the multiplicities. We introduce some formalisms:
+\begin{defn}
+	$\ch V$ is the formal character of $V$ and is defined to be
+	\[\ch V=\oplus_\mu m_\mu e^\mu\]
+	where the sum is over all weights of $V$ (the weight lattice).
+\end{defn}
+
+We add any two finite formal sums in the obvious way and multiplication is also obvious in that $e^\mu s^\sigma=e^{\mu+\sigma}$.
+
+Then we can actually take the formal characters $\ch M_\lambda$ of \textbf{Verma modules} $M_\lambda :=\mathcal U/(L_\alpha,h-\lambda(h),\alpha\in\Phi^+)$
+where $\ch M_\lambda=\sum a_\mu e^\mu$ where (as long as $\lambda$ is a weight) all $\mu$ occuring will lie in the same lattice.
+
+We will work out an expression for $\ch L_\lambda$, the formal character of the irreducible highest weight module of heighest weight $\lambda$, a dominant integral weight $\lambda$
+in terms of various $M_\mu$. We will see that $\ch L_\lambda$ will be a quotient of two simple alternating sums of ${e^\mu}'s$. To do this, we will have to dive into the structure of $M_\lambda$
+for a general $\lambda$.
+
+Notice that any subquotient (quotient of a submodule) of $M_\lambda$ has a weight that is maximal in the partial order, so also a highest weight vector of that highest weight. 
+Then that vector generates a highest weight module with a unique irreducible quotient $L_\mu$ for some $\mu$ so every irreducible subquotient 
+of $M_\lambda$ takes the form $L_\mu$ for some $\mu$.
+
+Now only finitely many $L_\mu$ can occur as subquotients of $M_\lambda$ for a fixed $\lambda$. 
+To prove theis, we need to revisit and generalize the idea of a \textbf{Kasimir element.} Recall the Killing form,
+which is nondegenerate on $L$. Set up dual basis of this form so that $\kappa(x_\alpha,z_\alpha)=1$ for $x_\alpha\in L_\alpha$ and $z\in L_{-\alpha}$.
+
+Set
+\[c=\sum_{\alpha\in\Phi^+}(x_\alpha z_\alpha+z_\alpha x_\alpha)+\sum_{i=1}^nh_i^2\]
+and notice that $c$ lies in the center of $\mathcal U$ -- it commutes with $L$. Then consider 
+(for $\lambda$ a highest weight vector)
+\[c v_\lambda=\sum_1^r\lambda(h_i)^2+\sum_{\alpha\in\Phi^+}(z_\alpha x_\alpha+[x_\alpha,z_\alpha]+z_\alpha x_\alpha)v_\lambda=(\lambda\cdot \lambda+\lambda(2\rho))v_\lambda\]
+where $2\rho$ is the sum of the positive roots (called the ``rho shift'')
+
+\section{March 6, 2019}
+Last time we stated that in the classical case, if $V=k^n$ is the defining (or natural) representation of
+$L\in\gl_n(k)$, then the symmetric and exterior owers are irreducible under $L$, except for the ``obvious relations.''
+
+The obvious relations were that, if $L$ is orthogonal, $S^2V$ has an $L$-invariant vector (sent to zero)
+so $S^mV$ is not irreducible for $m\ge 2$, but $\Lambda^mV$ is irreducible of highest weight $(1,\cdots,1,0\cdots,0)$  (with $m$ ones)
+for $m\le\lfloor\frac{n}{2}\rfloor$ where $n=\dim V$.
+
+Likewise if $L$ is symplectic, the $S^mV$ are all irreducible with highest weight $(m,0,\dots,0)$.
+
+\subsection{Return to (Formal) Characters}
+Recall that last time we studyies the Verma modules $M_\lambda=\mathcal U/(L_\alpha, h-\lambda(h): h\in H, \alpha>0)$.
+These all have irreducible quotients $L_\lambda$. SO we introduced the formal character $\ch M_\lambda$ of $M_\lambda$
+to be $\ch M_\lambda=\sum n_\mu e^\mu$ where $|mu$ runds over the $H$ weights of $M_\lambda$ 
+and $n_\mu=(M_\lambda)_\mu$, the dimension of the corresponding weight space.
+
+\begin{rmk}
+	We'll do a quick aside to explain the connection between the character of a group representation vs 
+	one of a Lie algebra. Recall that in the group case, we are just taking the determinant of a matrix 
+	(the image of $g$ under the representation).
+
+	In general, there is an exponential map from any Lie algebra to a Lie group $G$ with Lie algebra $L\subseteq \gl_n(k)$.
+	Here $G=GL(n,k)$ and $\exp \mu=\sum_{i=0}^\infty \frac{\mu^i}{i!}$ (we can work here over $\C$ so the series 
+	always converges.)
+
+	Then if $h\in H$, the Cartan subalgebra and if $N$ is a finidte simensional subquotient, then
+	\[\tr e^h)=\sum n_\mu e^{\mu(h)}.\]
+	we are extending to the case of infinite dimensional $M_\lambda$ without worrying about convergence.
+
+	We saw last time that any $M_\lambda$ admits a finite composition series $0=M_0\subseteq\cdots\subseteq M_m=M_\lambda$
+	where successive quotients are irreducible and each is isomorphic to some $L_\mu$. Then we are saying
+	\[\ch M_\lambda=\sum_{1}^m\ch(M_i/M_{i-1}).\]
+\end{rmk}
+
+Let's specialize this discussion to the case when $\lambda$ is \textbf{dominant integral}, which is the case of 
+greatest interest, so that $L_\lambda$ is finite-dimensional. Now
+\[\ch L_\lambda=\sum_\mu m_\mu\ch M_\mu\]
+and we will work out exactly which $\ch M_\mu$ occur and the $m_\mu$ in each case. More precisely,
+\[\ch M_\mu=e^\mu\prod_{\alpha\in\Phi^+}(1+e^{-\alpha}+e^{-2\alpha}+\cdots)=e^\mu\prod_{\alpha\in\Phi^+}\frac{1}{1-e^{-\alpha}}=\frac{e^{\mu+\rho}}{\prod(e^{\alpha/2}-e^{-\alpha/2})}\]
+
+Now if $|lambda$ is dominant integral we know that $\dim(L_\lambda)_\mu=\dim(L_\lambda)_{w\mu}$
+for any $w\in W$. So $w(\ch L_\lambda)=\ch L_\lambda$. Now applying any simple reflection $s_\alpha$ to the product above
+just premutes the terms except the one including $\alpha$, which is sent to its negative. Thus if $\Delta$ is 
+the denominator, $s_\alpha\Delta=-\Delta.$
+
+Hence $w\Delta=(\det w)\Delta$ where $\det w$ is computed with respect to the $W$ action on $H^*$. If 
+$\ch L_\lambda=\sum m_\mu \ch M_\mu=(\sum m_\mu e^{\mu+\rho})/\Delta$, it must be the case that 
+$w(\sum m_\mu e^{\mu+\rho})=(\det w)\sum m_\mu e^{\mu+\rho}$. THus we get that the coefficent of 
+$e^{w\lambda+\rho}$ in the sum must be $\det w$. But no other coefficients can appear -- $m_\mu=0$
+if $\mu\ne w(\lambda+\rho)-\rho$ for if any $m_\mu\ne -$ then $m_{w\mu}\ne 0$ for some
+\textbf{dominant} $w\mu$ lying below $|lambda$< but $w\mu+\rho$ would have the same length as $\lambda+\rho$.
+
+This leads us to
+\begin{thm}
+	The Weyl Charcter formula:
+	\[\ch L_\lambda=\frac{\sum_{w\in W}(\det w)e^{w(\lambda+\rho)}}{\prod_{\alpha\in \Phi^+(e^{\alpha/2}-e^{-\alpha/2}}}.\]
+\end{thm}
+\begin{rmk}
+	If $\lambda=0$ then $L_\lambda$ is the trivial representation having weight zero with multiplicity one and no other weight. So 
+	\[\prod(e^{\alpha/2}-e^{-\alpha/2}=\sum_W(\det w)e^{w\rho}\]
+	so 
+	\[\ch L_\lambda=\frac{\sum_{w\in W}(\det w)e^{w(\lambda+\rho)}}{\sum_{w\in W}(\det w)e^{w\rho}}\]
+
+	There is one small annoyance: that there is no product expression for the nmerator, except for the case when $\lambda=n\rho$.
+	Then we see that the weights of $L_n\rho$, with heighest weight $n\rho$ are of the form $n\rho-\sum n_\alpha\alpha$ for positive roots $\alpha$
+
+	The multiplicitiy of any weight is just the number of ways it can be written in this form.
+\end{rmk}
+
+\section{March 8, 2019}
+We will begin by mopping up some minor technical issues in the Weyl character formula from last time.
+
+We have Kostant's multiplicity formula: in computing $\ch L_\lambda$, we used that it is the sum over $w\in W$
+of $(\det w)\ch M_{w(\lambda+\rho)-\rho}$, then $m_\mu$, the multiplicity of $\mu$ in $L_\lambda$ is 
+\[m_\mu=\sum_{w\in W}(\det w)P(w(\lambda+\rho)-(\mu+\rho))\]
+where $P(\nu)$ is the number of ways to write $\nu$ as nonnegative ineger combination of positive roots.
+
+\subsection{Computing Dimension of $L_\lambda$}
+We use the Weyl character formula to compute a finite dimensional representation of highest weight $\lambda$, assumed dominant integral.
+Start with $R$, the rings of $\Z$ combinations of $e^\mu$ for $\mu$ in the root lattice. We have a homomorphism $v:R\to \Z$ taking $e^\mu\to 1.$
+The basic idea here is to apply this to the top and bottom of the Weyl character formula. We get zero in both
+cases. So we use something like L'H\^opital's rule.
+
+More precisely, given any positive root $\alpha$, there is a derivation $\delta_\alpha$ of $R$ sending $e^\mu$ to $(\mu \alpha)e^\mu$ (or something similar).
+Let $\delta=\prod_{\alpha>0}\delta_\alpha$ , which is not a derivation. But we can compute $\delta$
+on any product, term by term:
+\[\Delta(\ch L_\lambda)=\sum_{w\in W}(\det w)e^{w(\lambda+\rho)}\]
+where $\Delta$ is the Weyl denominator from before. Apply $v\delta$ to each side.
+
+On the left side, the nly remaining terms are from $v(\delta\Delta)\ch L_\lambda$ since $v(\Delta)=0$.
+On the other hand when we apply $\delta$ to a single $(\det w) e^{w(\lambda+\rho)}$, you get the 
+same result, independent of $w$ since any $w$ is the product of simple reflections and 
+\[\det(ws_\alpha)=\det w\prod_{\beta>0}(w(\lambda+\rho)\cdot \beta)=-\det w\prod_{\beta>0}(ws_\alpha(\lambda+\rho)\cdot \beta)\]
+
+With a little more computation we get 
+\begin{thm}
+	\textbf{Weyl's Dimension Formula}
+	\[\dim L_\lambda =\frac{\prod_{\beta>0}(\lambda+\rho)\cdot\beta)}{\prod_{\beta>0}\rho\cdot \beta}\]
+\end{thm}
+
+In general when applying this formula, we usually multiply each term $w(\lambda+\rho)\cdot\beta$ or $\rho\cdot \beta$ by 2, 
+divide by $\beta\cdot \beta$ so all terms on top and bottom are integers.
+
+\subsection{What this means}
+Recall that for any root system $\Phi$ there is a dual root system $\Phi^v=\{2\alpha/(\alpha\cdot\alpha)|\alpha\in\Phi\}$
+which is again a root system with a positive rootsystem going to a positive roots system and a simple subsystem going to a simple subsystem,
+
+Applying this formula to $G_2$, notice that the dual system is isomorphic to $G-2$, but the duality
+interchanges long and short roots. The simple roots are $\alpha=(0,1,-1)$ and $\beta=(1,-2,1)$
+and the positive roots are $\alpha,\beta,\alpha+\beta,\beta+2\alpha,\beta+3\alpha,2\beta+3\alpha.$
+
+In the dual system, we get simple roots the same thing but with $\alpha,\beta$ swapped for $\beta^v$ and $\alpha^v$, respectively. THen applying the formula 
+with $\lambda a^v=1$ and $\lambda\beta^v=0$, we get $\dim L_\lambda=7$. Thus $G_2$ has a seven dimensional representation. This tells us that a Lie algebra $L$ 
+of type $G_2$ embeds in $\gl_7$, and in fact in $\so_7$. So $L$ may be realized as a subalgebra of $\so_7$ (which is type $B_3$)
+and you can see these computations explicitly in $\S 19$ of the book.
+
+If instead we have $\lambda\cdot a^v=0$ and $\lambda\cdot \beta^v=1$, then $L_\lambda$ is 
+14 dimensional, and this is the adjoint representation of $G_2$. The largest exception simple Lie algebra s of type $E_8$
+and it is the unique simple Lie algebra $L$ such that $L$ itself is the unique nontrivial $L$-module of smallest dimension (which is 248).
+
+\subsection{Next Week}
+We are going to talk about simple Lie algebras over $\R$ -- we will find that it will somehow 
+mirror the classification we mentioned for Lie groups (where they were parameterized by subgroups between the weight
+and integer lattices). This time we will find that we get two types: complex simple Lie algebras which are just 
+considered to be real and real forms of complex simple lLie algebras such that $L\otimes_R\C$ is simple.
+
+\section{March 11, 2019}
+Today we are going to say something about semisimple Lie algebras over non algebraically closed fields (in particular $\R$).
+In the first week we defined the classical Lie algebras and gave bases for them. We say the commutator of any two basis vectors was an 
+integer combination of the the others.
+
+Now let $k$ be a field of characteristic zero (this isn't needed everywhere so we will point out where) and let $\Phi$
+be a root system. We have already seen the construction of an abstract semisimple Lie algebra over $k$ with root system $\Phi$
+by generators and relations. We didn't need anything about the field except that we needed formulas (power series)
+to implement the action of the Weyl group, which involves denominators. In fact, we only get twos and threes.
+
+This doesn't work over $\Z$, but let $L$ be the complex semisimple Lie algebra with root system $\Phi$. Let $\sigma$ be a
+Chevalley automorphism of $L$ (recall it acts on a Cartan subalgebra by $-1$ and interchanges the root spaces $L_{\alpha}$ and $L_{-\alpha}$).
+Now for all $\alpha\in\Phi^+,$ pick $0\ne x_\alpha\in L_\alpha$. We know that $[x_\alpha,-\sigma(x_\alpha)]$ is a multiple
+of $h_\alpha(w(2\alpha)/(\alpha,\alpha))$. If we replace $x_\alpha$ by $cx_\alpha$ for any $c\in\C$,
+then $-\sigma(x_\alpha)$ is also replaced by $-c\sigma(x_\alpha)$, so $[x_\alpha,-\sigma(x_\alpha)]$ is replaced by $c^2$
+times itself. Thus we can choose $c$ to scale $x_\alpha$ such that htis bracket is exactly $h_\alpha$.
+
+Choose a basis of $L$ as follows: $x_\alpha\in L_\alpha$ for $\alpha>0$ and $-\sigma(x_\alpha)\in L_{-\alpha}$ for $\alpha>0$. 
+Then pick $h_i=h_{\alpha_i}$ for all the simple roots $\alpha_i$. THen 
+\[[x_\alpha,x_{-\alpha}]=h_\alpha\quad [h_i,x_{\alpha}]=\alpha(\alpha_i)h_\alpha\in\Z h_\alpha\quad [x_\alpha x_\beta]=c_{\alpha\beta}x_{\alpha+\beta}\]
+and by applying $\sigma$, we get that $[x_{-\alpha}x_{-\beta}]=c_{-\alpha-\beta}$, so $c_{-\alpha-\beta}=-c_{\alpha\beta}$
+
+Then by thm 25.2(a) in the text, this implies $c_{\alpha\beta}=\pm(r+1)$ where $r$ is the largest integer 
+such that $\beta-r\alpha\in\Phi$. The upshot here is that any bracket of basis vectors is an integer combination of basis vectors.
+
+\brk
+
+So we see that htere is a Lie algebra (usually semisimple) over any basefield $k$ with root system $\Phi$, called \textbf{split}
+(having a root space decomposition and a lage abelian subalgebra consisting of semisimple elements). We now want to look at semisimple Lie algebras 
+\textit{not} admitting root space decompositions -- the main example is $\R$.
+
+Let $\Phi$ be a root system, $L$ the corresponding complex semisimple Lie algebra, and $\{h_i,x_\alpha\}$
+a Chevalley basis. Look at the $\R$-subalgebra $L'$ of $L$ spanned by $x_\alpha-x_{-\alpha}$ and $i(x_\alpha+x_{-\alpha})$ as well as $ih_1,\dots,ih_l,$ for all $\alpha\in\Phi^+$.
+Now $L'$ is closed under the bracket and the Killing form $\kappa$ on $L'$ is negative definite.
+
+But $L'$ doesn't admit a root space decomposition because $L'$ is toral -- it consists \textit{entirely} of semisimple elements:
+possible since in fact the semisimple elements are \textbf{not} diagonalizable (in a non-algebraically-closed field semisimple and diagonalizable don't have to coincide)
+and in fact, they will all have purely imaginariy eigenvalues.
+
+We call $L'$ \textbf{compact} -- note that it is not compact as a topological space, but its adjoint group $\Int L$, generated 
+by $\exp\ad x$, is a compact Lie group. It is, in fact, a closed subgroup of $O(n,\R)$ (a compact group)
+and $n=\dim_\R L'=\dim_\C L$, Since it preserves $\kappa$, a negative definite symmetric bilinear form, the \textbf{compact real form of $L$.}
+
+\subsection{Unitary matrices}
+An astute reader will notice that we haven't mentioned the Lie groups formed by unitary matrices, for instance (and 
+serving as a guide to understand the rest) $\mathfrak{su}(n,\C)$, the set of all $n\times n$ complex matrices that are skew-adjoint
+with respect to the Hermetian form $((v_i),(w_i))=\sum_1^n v_i\bar w_i$ and of trace zero.
+Notice that although these matrices are complex, the algebra is not stable under complex multiplication, so it is not a \textit{complex} Lie algebra.
+The correspoinding group $SU(n,\C)$ consists of all $g\in GL_n(k)$ such that $(gv,gw)=(v,w)$ for all $v,w\in\C^n$.
+
+\subsection{Orthogonal matrices}
+$\mathfrak{so}_n(\R)=\mathfrak{o}_n(\R)$, the skew-adjoint matrices with respect to the dot product on $\R^n$ has correspinding group $SO(n,\R)$.
+The lie algebra is all the $n\times n$ skey-symmetric matrices in the ordinary sense.
+
+Over $\R$, the group that we would get from our original construction is not $SL(n,\R)$, but something else.
+
+\subsection{Symplectic Case}
+Starting with $\sp_{2n}(\C)$, take the intersection with $\mathfrak{su}_{2n}(\C)$ -- this is still a real Lie algebra.
+We can think of this as right-linear transformations of $\mathbb{H}^n$ (quaternions) preserving a skew-Hermitian form:
+\[((v_1,\dots,v_n)(w_1,\dots,w_n))=\sum_1^nv_ij\bar w_i.\]
+
+Notice that we say ``right-linear'' because $\mathbb{H}^n$ admits (different) left- and right-vector space structures.
+
+Then the group in question is the symmetry group of this form, the Lie algebra consists of matrices skew-adjoint with respect to it.
+
+\subsection{Other Real Forms}
+The above concludes our discussion of the \textit{compact} Lie algebra over $\R$, but there are even 
+more real forms that are neither compact nor split. There are only finitely many of these, however, and each is partly compact and partly not.
+These are of the form $L=K+P$ where $K$ is compact and $[PP]\subseteq K$. All the elements in $K$ will be semisimple
+with purely imaginary eigenvalues and $P$ will consist of both semisimples and nilpotents, but the semisimples will have 
+real eigenvalues.
+
+One such example is defined by the form 
+\[(v_1,\dots,v_n)(w_1,\dots,w_n)=\sum_1^m v_i\bar w_i-\sum_{m+1}^n v_j\bar w_j.\]
+This is called the \textbf{symmetric group} $U(n,n-m)$.
+
+%TODO: FIX CHAR/CH
 \end{document}

### 2019-03-12 Reading Notes/reading_notes.tex
Fortnightly Update

@@
 %schemes
-\newcommand{\Ga}{\ensuremath\mathbf{G_a}}
-\newcommand{\Gm}{\ensuremath\mathbf{G_m}}
+\newcommand*{\bbG}{\ensuremath\mathbb{G}}
+\newcommand*{\Ga}{\ensuremath\mathbf{G_a}}
+\newcommand*{\Gm}{\ensuremath\mathbf{G_m}}
 \newcommand*{\GL}[1]{\ensuremath\mathbf{GL_#1}}
@@
 \newpage
+\section{Julia \& Dave Benson's preprint}
+As an aside, in response to me saying I needed to practice doing computations with spectral sequences,
+Julia recommended I read her current work \cite{julia-dave} with Dave Benson wherein there is a supposedly
+straightforward spectral computation in support of their larger work \cite{detecting-nilpotence}.
+
+I have already spent some time reading this paper, but the core objects in question are \textbf{supergroup schemes}
+which have coordinate Hopf algebras algebras with a $\bbZ/2$ grading that are furthermore (graded) cocommutative (just 
+as group schemes correspond to finite dimensional cocommutative Hopf algebras). The fundamental example is $\bbG_a^-$,
+which has as its coordinate ring $k[x]/(x^2)$, or equivalently the exterior algebra on a single element $x$,
+where $x$ is taken to be a primitive element.
+
+Then if we let $A$ be the coordinate algebra of $\bbG_a^-\times\bbG_a^-$ and $H$ that of $\bbG_{a(r)}\times(\bbZ/p)^s$ where $\bbG_{a(r)}$ is the
+$r^{th}$ Frobenius kernel of $\bbG_a$, we can define the \textbf{smash product} or in this case, since 
+$A$ is graded cocommutative, the \textbf{semidirect product} $A\#H=A\rtimes H$ where we define the action
+to be (if $\tau:H\otimes A\to A$ is the action of $H$ on $A$)
+\[(a\otimes h)\cdot(a'\otimes h')=\sum (-1)^{|h_1||b|}a\tau(h_1\otimes b)\otimes h_2g.\]
+
+So we use that $A\cong \Lambda(k[u,v])$, the exterior algebra on two (primitive) generators $u$ and $v$
+and that, if $t_i$ is the generator for the coordinate algebra of the $i^{th}$ copy of $\bbZ/p$, and if 
+we let $g_i=t_i-1$, these along with the $s_1,\dots s_r$ generating the Frobenius kernel, generate $H$.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\newpage
 \section{Affine Group Schemes (Waterhouse and Jantzen)}

### 2019-03-15 Hopf Algebra Notes/Hopf Algebras.tex
Update work

@@
 \end{defn}
+
+\section{March 13, 2019}
+Today we are going to discuss why considering algebras over $A$ is not quite sufficient in many cases --
+why we may be instead interested in studying $_A\calM$.
+
+Let $A$ be an algebra and $V$ a Hopf algebra. Let $V$ act on $A$ and define
+\[A^V=\{a\in A|va=\varepsilon(v)a,\forall v\in V\}\]
+We can see that we get $A^v=\Hom_V({_Vk},A)$ and then use this to compute the cohomology
+\[\Ext_V^i(k,A)\]
+
+We forgot the
+\subsection{(Non)-Examples of the Day}
+\begin{ex}
+	Let $A$ be an algebra, $n\ge 2$ an integer. Then $M_n(A)$ can't be a Hopf algebra.
+
+	The idea here is that there is no counit $\varepsilon:M_n(A)\to k$.
+\end{ex}
+\begin{ex}
+	If $V_1$ and $V_2$ are Hopf algebras, then $V_1\oplus V_2$ is not a Hopf algebra.
+	Notice here that it is both an algebra and a coalgebra!
+
+	Here, use that $1_V=1_{V_1}+1_{V_2}$ and compute $\Delta(1_V)\ne 1_V\otimes 1_V$.
+\end{ex}
+\begin{ex}
+	If $A$ is finite dimensional and not Frobenius, then $A$ can't be a Hopf algebra.
+\end{ex}
+
+\subsection{Some Category Theory}
+Let $k$ be af field.
+\begin{defn}
+	Let $\mathscr C$ be a category. $\mathscr C$ is called \textbf{$k$-linear} if
+	\begin{itemize}
+		\item $\Hom_{\mathscr C}(M,N)\in\Vectk$ for all $M$ and $N$ in $\mathscr C$
+		\item Composition is $k$-linear.
+	\end{itemize}
+\end{defn}
+\begin{defn}
+	A category $\mathscr C$ is called \textbf{finite} if $\mathscr C\cong {_A\calM_{fd}}$ 
+	for some finite dimensional algebra $A$.
+\end{defn}
+\begin{defn}
+	A category $\mathscr C$ is called \textbf{monoidal} (notice that sometimes this books 
+	call this a tensor category) if it is equipped with
+	\begin{itemize}
+		\item A bifunctor $-\otimes -:\mathscr C\times\mathscr C\to \mathscr C$
+		\item An associativity isomorphism: for all $X,Y,X\in\mathscr C$:
+		\[a_{X,Y,Z}:(X\otimes Y)\otimes Z\xrightarrow{\sim} X\otimes(Y\otimes Z)\]
+		\item An identity object $\mathbf{1}\in\mathscr{C}$
+		\item An isomorphism $i:\mathbf{1}\otimes\mathbf{1}\to \mathbf{1}.$
+	\end{itemize}
+	That satisfies the pentagon axiom (Figure~\ref{fig:pentagon}) and the unit axioms
+	\[L_\mathbf{1}X\xrightarrow{\sim} \mathbf{1}\otimes X,\qquad R_\mathbf{1}:X\xrightarrow{\sim}X\otimes \mathbf{1}\]
+\end{defn}
+\begin{figure}\label{fig:pentagon}
+	\centering
+	\begin{tikzcd}
+		((W\otimes X)\otimes Y)\otimes Z\ar[d,"a_{W,X,Y}\otimes \id_Z"]\ar[r,"a_{W\otimes X,Y,Z}"] & (W\otimes X)\otimes (Y\otimes Z)\ar[dd,"a_{W,X,Y\otimes Z}"]\\
+		(W\otimes(X\otimes Y))\otimes Z \ar[d,"a_{W,X\otimes Y,Z}"]& \\
+		W\otimes((X\otimes Y)\otimes Z)\ar[r,"\id_W\otimes a_{X,Y,Z}"] & W\otimes(X\otimes (Y\otimes Z))
+	\end{tikzcd}
+	\caption{The Pentagon Axioms}
+\end{figure}
+\begin{ex}
+	Let $A$ be an algebra and let $\mathscr C= ({_A\calM_A},\otimes_A,\mathbf{1}=A)$.
+	Clearly this works. :)
+\end{ex}
+
+Recall that if $V$ is a Hopf algebra and $M\in {_V\calM_{fd}}$, then $M^*:=\Hom_k(M,k)\in\calM_V$
+then we defined $M^{*S}=\tilde S\circ M^*\in {_V\calM}$. This is the right dual.
+
+We can also define the left dual if $S$ is bijective. In this case, we define $M^{*S^{-1}}=\tilde S^{-1}\circ M^*\in{_V\calM}$.
+
+Now let $(\mathscr C,\otimes, \mathbf{1})$ be a monoidal category.
+\begin{defn}
+	Let $X\in\mathscr C$. An object $X^*$ in $\mathscr C$ is said to be a \textbf{left dual of $X$}
+	if there exist an \textbf{evaluation map}
+	\[\operatorname{ev}_X:X^*\otimes X\to \mathbf{1}\]
+	and a \textbf{coevaluation map}
+	\[\operatorname{coev}_X:\mathbf{1}\to X\otimes X^*\]
+	such that
+	\[X\xrightarrow{\operatorname{coev}_X\otimes\id_X}(X\otimes X^*)\otimes X\to X\otimes (X^*\otimes X)\xrightarrow{\id_X\otimes \operatorname{ev}_X} X\]
+	is an isomorphism as well as the map ``on the other side''.
+\end{defn}
+\begin{defn}
+	In the context of the above, a \textbf{right dual of $X\in\mathscr C$} is a left dual of $X$ in $\mathscr C^{op}$.
+	The right dual is denoted (confusingly) by ${^*X}$.
+\end{defn}
+So in the Hopf algebra case, ${^*M}=M^{*S}$ and $M^*=M^{*S^{-1}}.$
+\begin{defn}
+	A category $\mathscr C$ is called \textbf{rigid} if every $X\in\mathscr C$ has both left and right duals.
+\end{defn}
+\begin{defn}
+	$\mathscr C$ is called a \textbf{tensor category} if 
+	\begin{itemize}
+		\item $\mathscr C$ is finite (this implies $k$-linear)
+		\item It is monoidal
+		\item it is rigid
+		\item $\End_\mathscr{C}(\mathbf{1})=k$
+	\end{itemize}
+\end{defn}
+\begin{rmk}
+	If we remove the last condition and replace it with $\End(\mathbf{1})=\oplus k$, then
+	we call $\mathscr C$ \textbf{multi-tensor.}
+\end{rmk}
+\begin{defn}
+	$\mathscr C$ is called \textbf{fusion} if 
+	\begin{itemize}
+		\item It is tensor
+		\item It is semisimple (every object is a direct sum of simple objects).
+	\end{itemize}
+\end{defn}
+\begin{rmk}
+	If instead $\mathscr C$ is multitensor and semisimple, then it is called \textbf{multifusion.}
+\end{rmk}
+\begin{ex}
+	\begin{enumerate}
+		\item $\Vectk$ is an example.
+		\item If $V$ is a finite dimensional Hopf algebra, ${_V\calM}$ is tensor.
+		\item If $V$ is also semisimple, ${_V\calM}$ is fusion.
+	\end{enumerate}
+\end{ex}
+Recall this conjecture for the first day (Etinghof-Ostrik restated):
+\begin{conj}
+	If $\mathscr C$ is a tensor category then $\Ext^*(\mathbf{1},\mathbf{1})$ is Noetherian.
+\end{conj}
+
+\section{March 15, 2019}
+On Wednesday, we said that if $M\in{_V\calM_{fd}}$ that $M^*=M^{*S^{-1}}$ and ${^*M}=M^{*S}$,
+but in fact these are reversed.
+\subsection{Example of the Day}
+\begin{ex}
+	Usually we have been defining Hopf algebras over vector spaces, but this will be an object in $\text{\sc grVect}_k$,
+	that is $V=\oplus_{i\in\Z}V_i$ and if $M=\oplus M_i$ and $N=\oplus N_i$, then $M\otimes N=\oplus_i(\oplus_{s+t=i}M_s\otimes N_t)$.
+
+	Now putting the ``symmetric'' back into the monoidal category, we define $\tau:M\otimes N\to N\otimes M$ via
+	\[m\otimes n\mapsto (-1)^{|m||n|}n\otimes m.\]
+	Then we have our symmetric monoidal category and we can define a Hopf algebra in it.
+
+	$V=\Lambda(x_1,\dots,x_n)$, the exterior algebra on $n$ elements. Define the coalgebra structure 
+	so that each $x_i$ is primitive. Note that here $V$ is a Hopf algebra over graded vector spaces, but 
+	NOT over vector spaces (unless $\ch k=2$).
+
+	To see this, we can compute 
+	\[0=\Delta(x_i^2)=\Delta(x_i)^2=(x_i\otimes 1+1\otimes x)^2=x_i^2\otimes 1+x_i\otimes x_i+\tau(x_i\otimes x_i)+1\otimes x_i^2\]
+	and if $\tau$ is the usual vector space twist (and the characteristic isn't 2) we get a contradiction.
+\end{ex}
+
+\subsection{Some History}
+Before 1941 people studied rings like $kG, \mathcal U(\g)$ and $\mathcal O(G)$. An open question at that
+point was ``If $\mathcal U(\g)$ is Noetherian, then is $\dim \g<\infty?$''
+
+In 1941, Hopf algebras were born, although they were originally defined (by Hopf) to be over graded
+vector spaces. Eventually people started considering them over vector spaces.
+
+In 1967, Sweedler wrote the first book about Hopf algebra over vector spaces. Then later, near 1970,
+Rota began working more on combinatorial Hopf algebras. It ends up in this context that often the 
+M\"obius function ends up being the co(unit? multiplication?).
+
+In 1975, Kaplansky posited 10 conjectures. Of those only one more remains:
+\begin{conj}{Kaplansky, `75 (\#6)}
+	If $V$ is a semisimple Hopf algebra and $M$ is a simple $V$-module, then 
+	\[\dim_kM|\dim_kV.\]
+\end{conj}
+
+From 1960-90, the studies of $kG, \mathcal U(G)$ and $\mathcal O(G)$ and semisimple Hopf algebra continued,
+led in large part by Radford. In 1987, Drinfeld and Jimbo came out with quantum groups.
+At the same time Drinfeld introduced quasi- and braided Hopf algebras and weak Hopf algebras.
+
+Of course we all know about Nichols algebras. Also we switched to different tensor categories, including
+fusion, pivotal, and spherical ones.
+
+In 1993 Lusztig invented the small quantum groups, which are of particular interest to representation theorists.
+`97 was Brown-Goodearl. Then around 1990 there were lots of classification projects including
+the one by Anruskiewitsch-Schneider project. Near this time Zhu proved that every Hopf algebra of dimension $p$
+is a group algebra $k\Z/p$.
+
+In 2004 was Etingof-Ostrik. The next year, Etingof-Nikshych-Ostrik wrote a big paper ``On Fusion Categories''
+which really shifted the perspective from Hopf algebras to things in fusion categories, where many of the big
+results were transferred.
+
+In 2015 Etingof, Nikshych, Ostrik (and someone else) wrote ``Tensor Categories''.
+
+\subsection{Extensions}
+We can look at extensins of Brown-Goodearl. The question ``Is every Noetherian X Gorenstein?'' can be extended to be asked about any noun 
+in \{Hopf Algebras, weak Hopf algebras, braided hopf algebras, Nichols algebgras, quasi-Hopf algebras\}.
+
+We can also consider extensions of Etingof-Ostrik: ``Let $V$ be a finite dimensional X and $A$ be a 
+Notherian left $V$-module such that $A^V$ is Noetherian. Is then $\Ext_V^*(k,A)$ Noetherian?'' Again
+we can take $X$ to be anything in the above set.
+
+In 2019, who knows!
+
 \end{document}

### 2019-03-15 Lie Algebra Notes/Lie Algebras.tex
Update work

@@
 %characteristic
+%\let\char\relax
 %\DeclareMathOperator{\char}{char}
@@
 \DeclareMathOperator{\so}{\mathfrak{so}}
+\DeclareMathOperator{\su}{\mathfrak{su}}
 \let\sp\relax
@@
 
-%TODO: FIX CHAR/CH
+\section{March 13, 2019}
+One correction from last week: the compact from of $\sp_{2n}(\C)$ is given by the skew-adjoing matrices of trace zero relative to the \textit{Hermitian}
+form $(v,w)=\sum_iv_i\bar w_i$ on $\mathbb{H}^n$ (not skew-Hermitian).
+
+\subsection{Continuing with Real Forms}
+Now we will address the non-compact forms. We've seen the compact and split real forms, which form 
+the opposite ends of a spectrum and now we consider the ones that lie in the middle.
+
+\subsubsection{Real forms of type $A$}
+In type $A$, we have already
+seen $\su_n(\C)$ consisting of skew-adjoint matrices of trace zero relative to the positive definite Hermitian form on $\C$.
+We've also seen $\sl_n(\R)$.
+
+In between, we have $\su(p,q)$, consisting of the skey adjoint matrixes with respect to the Hermitian form $(-,-)_{pq}$ where
+\[(v,w)=\sum_1^pv_i\bar w_i-\sum_{p+1}^{p_q}v_i\bar w_i\]
+which is Hermitian of \textbf{signature} $(p,q)$ on $\C^{p+q}=\C^n$. WHen $p=n$ and $q=0$, we get the compact form $\su_n(\C)$.
+
+We get one more real form -- only in even dimensions -- $\su^*(2n)$: all $n\times n$ matrices over the quaternions $\mathbb{H}$ whose trace is a real
+combination of $i,j,k$.
+
+\subsubsection{Types $B$ and $D$}
+Here the story is mostly the same for these types with a small difference: we have $\so(p,q)$
+in either, which is all skew-adjoint matrices with respect of the symmetric form $(-,-)_{pq}$
+on $\R^{p+q}$ where
+\[(v,w)=\sum_1^pv_iw_i-\sum_{p+1}^{p+q}v_iw_i.\]
+
+But we get one extra in type $D$: $\so^*(2n)$. This consits of all skew-adjoint matrices with respect to the skey Hermitian form
+\[(v,w)=\sum_1^n v_ij\bar w_i\]
+where $j\in\mathbb{H}$.
+
+\subsubsection{Type $C$}
+Here we get $\sp(p,q)$ -- the skew-adjoint matrices with respect to the form as usual ($v_i\bar w_i$).
+We also have $\sp(n,\R)$, the compact form.
+
+Finally we get $\sp(2n,\R)$, the real matrices skew adjoint with respect to a skew-sypmmetric form on $\R^{2n}$.
+
+\subsection{Cartan Involution}
+In general, recall from last time that any real semisimple Lie algebra admits a Cartan decomposition
+$L=K+P$ where $K$ is compact and $[P,P]\subseteq K$. Furthermore, the Killing form $\kappa$ restricts
+to $K$ as a negative definite form and to $P$ as a positive definite form. Finally, $P$ is a $K$-sumbodule:
+$[KP]\subseteq P$.
+
+It follows that we get an involution on $L$ by decreeing that $\theta=1$ on $K$ and $\theta=-1$ on $P$ (???) the \textbf{Cartan
+involution.} Moreover $L$ admits a complexification $L\otimes_\R\C=L'$, which is complex semisimple, and
+$\theta$ extends to an involution of $L'$. We classify all real forms up to isomorphism by classifying the involutions
+$\theta$ up to conjugacy. It ends up there are only finitely many real forms.
+
+\subsection{Exceptional Cases}
+It turns out htat in the exceptional cases we can label real forms explicitly as $X_N(a)$ where the real form of type $X_N$ has $a=\dim P-\dim K$.
+Here the compact form is always $X_N(-\dim_CX_N)$
+\subsubsection{$E_6$}
+This ends up being the most interesting type (although it sits inside of its bigger brother).
+Here we get the real forms:
+\begin{figure}[h!]
+	\centering
+	\begin{tabular}{c|c|c|c}
+		$E_6(6)$ & $E_6(2)$ & $E_6(-14)$ & $E_6(-26)$\\\hline
+		$C_4$ & $A_5\times A_1$ & $D_5\times \C$ & $F_4$
+	\end{tabular}
+\end{figure}
+where here the $\C$ refers to the fact that there will be a (one-dimensional) nontrivial center.
+\subsubsection{$E_7$}
+\begin{figure}[h]
+	\centering
+	\begin{tabular}{c|c|c}
+		$E_7(7)$ & $E_7(-5)$ & $E_7(-25)$ \\\hline
+		$A_7$ & $D_6\times A_1$ & $E_6\times \C$
+	\end{tabular}
+\end{figure}
+\subsubsection{$E_8$}
+\begin{figure}[h]
+	\centering
+	\begin{tabular}{c|c}
+		$E_8(8)$ & $E_8(-24)$ \\\hline
+		$D_8$ & $E_7\times A_1$
+	\end{tabular}
+\end{figure}
+\subsubsection{$F_4$}
+\begin{figure}[h!]
+	\centering
+	\begin{tabular}{c|c}
+		$F_4(4)$ & $F_4(-20)$ \\\hline
+		$C_3\times A_1$ & $B_4$
+	\end{tabular}
+\end{figure}
+So again we get two.
+\subsubsection{$G_2$}
+Here we only have one form: $G_2(2)=A_1\times A_1$.
+
+\subsection{Diagram Automorphisms}
+There are three types of Lie algebras where diagram automorphisms of order two occur: namely 
+$A_n$, $D_n$ and $E_6$. In these cases, we create a different kind of extended Dynkin diagram:
+we add a vertex representing the lowest weight of the quotient representation that is formed
+by beginning by looking at $L$ and then quotienting out by the fixed points of the diagram 
+automorphism.
+
+\subsection{Next time...}
+The last lecture of the quarter will be dedicated to what happens when $k$ has positive characteristic.
+There is a classification, but it is much longer and mroe complicated than the one we have given.
+
+In particular, one might at first think that at least the split form of a complex semisimple Lie algebra
+$L$ would have a characteristic $p$ analog. While this is true, it is not always (even) semisimple. 
+The problem is that we completely lose the notion of positivity in positive characteristic, so 
+the Killing form, even when restricted to a Cartan subalgebra, needn't be nondegenerate.
+
+In the simplest case, consider $\ch k=p$ and the Lie algebra $\sl_n(k)$. The identity matrix
+then has trace zero! Problems abound. Finally we will talk about the finite simple groups of Lie type,
+which form the building blocks for the Lie groups we are interested in.
+
+\section{March 15, 2019}
+Let $L$ be a complex simple Lie algebra. We have seen that $L$ has a Chevalley basis
+consisting of a basis of $H$, a Cartan subalgebra, with root system $\Phi$, and for each
+(nonzero) root $\alpha\in\Phi$ a vector $x_\alpha\in L_\alpha$ such that whenever $\alpha,\beta,\alpha+\beta$
+are all roots, $[x_\alpha,x_\beta]=c_{\alpha\beta}x_{\alpha+\beta}$. Then $c_{\alpha\beta}=-c_{-\alpha.-\beta}$
+and $c_{\alpha\beta}=\pm(r+1)$ where $r$ is the largest integer (possibly zero) such that $\beta-r\alpha\in\Phi$.
+
+Then the span $L_\Z$ over the integers of a Chevalley basis is a Lie algebra ``over $\Z$'', which is 
+simple in the sense that any nonzero ideal comtains $nL_\Z$ for some nonzero $n\in\Z$. We have seen however that
+the reduction $L_p$ of $L_\Z$ modulo $p$ need not be simple.
+
+So we focus on the \textbf{group} generated by $\exp\ad cx_\alpha$ for $c\in\Z$ and $\alpha\in\Phi$.
+Suppose that $\alpha,\beta,\beta+\alpha,\beta+2\alpha,\dots,$ are all roots. Then
+\[\exp\ad x_\alpha(x_\beta)=\pm(r+1)(r+2)\cdots x_{\beta+n\alpha}\]
+(since all root strings are actually finite, the series actually can be considered to be
+finite in any one case). Thus $\pm(r+1)\cdots(r+n)/w!$ is an integer, so the terms in the series
+expainsion of $\exp\ad cx_\alpha$ all preserve the $\Z$-span $L_\Z$ whence there reductions modulo
+a prime $p$ preserve the corresponding span $L_p$ so that if $F$ is any field containing $\Z_p$
+and if we set $L_F=L_p\otimes_{\Z/p}F$ then the group $G_F$ generated by $\exp\ad cx_\alpha$ acts on $L_F$.
+
+Specializing down to the case when $F=\mathbb{F}_q$, the finite field of order $q=p^n$,
+we therefore get a finite group acting on $L_q:=L_{\mathbb{F}_q}$. Then Chevalley proved that $L_q$ is always simple as a group, 
+with just a few exceptions occurring for very small fields $F$, thereby exhibiting a number 
+of families of finite simple groups, now said to be \textbf{of Lie Type.} Many of the resulting simple groups were known
+before but Chevalley was the first to give a unified treatment.
+
+\subsection{Examples}
+An explicit example comes from $GL_n(\mathbb{F}_q)$. It can be shown (see many old prelims) to 
+have order $q^{\binom{n}{2}}(q^n-1)\cdots(q-1)$. This is actually never simple! But luckily we want some smaller
+groups! We have $SL_n(\mathbb{F}_q)$ which has order $q^{\binom{n}{2}}(q^n-1)\cdots(q^2-1)$.
+
+Still not simple, unfortunately, but if we mod out scalars to $PSL_n(\mathbb{F}_q)$, we get a simple
+group (whenever $q\ge 4$) of order $q^{\binom{n}{2}}(q^n-1)\cdots(q^2-1)/g$ where $g=\operatorname{gcd}(n,q-1)$.
+
+We notice that the formula for the order of the simple group $PSL_n(\mathbb{F}_q)$ is given 
+by a power of $q$ times a product of $q^i-1$ for varius $i$, divided the by the order of a central
+subgroups.
+
+Chevalley showed this pattern always holds: the exponent of $q$ is the dimension of the Borel
+subalgebra of the Lie algebra giving rise to the group. The $q^i-1$ powers are called the \textbf{exponents}
+of the Weyl group of the Lie algebra.
+
+\subsection{Classificiations}
+Let us look at all finite simple roups of nonprime order: we have
+\begin{itemize}
+	\item $A_n$, the alternating groups, for $n\ge 5$
+	\item Finite groups of Lie type 
+	\item Chevalley groups
+	\item Further anaglogues of these arising from a nontrivial diagram automorphisms fo the Lie algebra
+	\item Steinberg analogues on characteristic 2 and 3 on types $B_2$ and $G_2$
+	\item Ree, Suzuki, and Tits (are these the discoverers of above?)
+	\item Only finitely many others! 26 sporadic groups with the monster group being the largest.
+\end{itemize}
+\begin{rmk}
+	There is also a monster Lie algebra.
+\end{rmk}
+
+Moreover, one can form an analgof of $L_\Z$ for the enveloping algebra: it has a $\Z$-form
+$U_\Z$ consisting of all sums of products $\frac{x^{n_\alpha}_{-\alpha}}{n_\alpha!}\cdots\frac{x_\alpha^{n_\alpha}}{n_\alpha!}$
+and this is stable under the action of the Chevalley group.
+
+I missed the last bit because Monty was talking, but that wraps it up for the quarter!
 \end{document}

### 2019-03-15 Reading Notes/reading_notes.tex
Update work

@@
 
+\subsection{Week 8}
+The last paper on Julia's page is a preprint with her and Dave Benson. There is a spectral sequence computation that might be illustrative.
+There is also an older paper...
+
+One of the things Julia is interested in at the moment is a super group scheme. These arise by having a $\bbZ_2$ grading on your Hopf algebra.
+For instance we can look at exterior algebra $\Lambda(x)$, which represents the super group scheme $G_a^-$.
+
+The first non-elementary example is $G_a^-\times G_a^-$ with a $\bbZ_p$ action on it that ``folds'' the second factor
+onto the first. To write this out explicitly, we consider the group algebra $\Lambda(u,v)$ where $uv=-vu$ and the coordinate algebra of $\bbZ_p$
+is $k[t]/t^p$ where $t\cdot u=0$ and $t\cdot v=\alpha u$ for some $\alpha$. This is a $4p$ dimensional algebra.
+
+Then we want to compute coholomology. We know $H^{**}(k\bbZ/p,k)\cong k[x]\oplus\Lambda(y)$ for $|x|=2,0$.
+Note there is a notion of dimension that is an ordered pair $(a,b)$ of the graded degree and cohomology degree.
+THis give us new elements of degree $(1,1)$
+
+Look up Koszul duality! It is in the representations and local cohomology book!
+
+\subsection{Week 10}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+One of the first propositions (3.1) of the paper computes the cohomology of $\bbG_a^-$ with 
+coefficients in $k$ (actually, refers to another paper) and computes how the ($\bbZ/2$-graded)
+Steenrod operations act on it.
+
+\begin{defn}[Steenrod Operations]
+	If $p$ is odd and $A$ is a $\bbZ$-graded cocommutative Hopf algebra, then there are natural maps/operations
+	\[\mathcal P^i:H^{s,t}(A,k)\to H^{s+(2i-t)(p-1),pt}(A,k)\]
+	\[\beta \mathcal P^i:H^{s,t}(A,k)\to H^{s+1+(2i-t)(p-1),pt}(A,k)\]
+	satisfying the properties
+	\begin{enumerate}
+		\item $\mathcal P^i=0$ whenever $2i<t$ or $2i>s+t$,
+		\item $\beta\mathcal P^i=0$ whenever $2i<t$ or $2i\ge s+t$.
+		\item $\mathcal P^i(x)=x^p$ when $2i=s+t$
+		\item $\mathcal P^j(xy)=\sum_i\mathcal P^i(x)\mathcal P^{j-i}(y)$
+		\item $\beta\mathcal P^j(xy)=\sum_i(\beta\mathcal P^i(x)\mathcal P^{j-i}(y)+\mathcal P^i(x)\beta\mathcal P^{j-i}(y))$
+		\item $\mathcal P$ and $\beta\mathcal P$ satisfy the Adem relations.
+		\item $\mathcal P^0(\lambda u)=\lambda^p\mathcal P^0(u)$ for all $\lambda \in k$ ($\ch k= p$)
+	\end{enumerate}
+\end{defn}
+
+In the course of proving things here, the authors adapt these operations to the $\bbZ/2$ grading
+by reindexing, calling these new operations $\mathscr P_i$ and $\beta\mathscr P^i$ (they use \texttt{mathscr} but I've
+reassigned mine to a font I like better. :P)
+
+Okay, time is running out, so I am pulling together some questions:
+\begin{qst}
+	Is the cohomology you're using here group cohomology? It seems to me the Lyndon-Hochschild-Serre
+	applies to group cohomology.
+
+	If so, I will need to focus some time on getting familiar with it.
+\end{qst}
+
+Yes it is group cohomology, or more specifically group \textit{scheme} cohomology. Here we 
+are computing $\Ext_G^{i,\varepsilon}(k,k)$.
+
+We begin with our group algebra (here $\Lambda(v_1,v_2)\#k\bbG_{a(1)}$) and take 
+a projective resolutions and computing $H^*(\Hom_A(p^\bullet,k))$.
+
+Recall the syzygys, $\Omega^i k$ that occur in a projective resolution of $k$. Then we can compute the cohomology
+via the stable module category:
+\[\underline{\Hom}_G(\Omega^ik,k)=\Ext_G^{i,\varepsilon}(k,k)\]
+
+There is a result by Jeremy Rickard (rep theory) and by Buchweitz (more generally) that proves
+
+\begin{qst}
+	Where is $H^{*,*}(\bbG_a^-,k)$ computed? The papers claim it is $S^{*,*}(V^\sharp)\cong k[\zeta,\eta]$
+	where $\zeta$ and $\eta$ are the dual generators corresponding to the generators $u$ and $v$ of $V$,
+	the augmentation ideal of $k(\bbG_a^-\times \bbG_a^-)$.
+\end{qst}
+
+Again, this is an example of Koszul duality -- use that $k\bbG_a^-=\Lambda(v)$ where $v$ is primitive in degree 1.
+But then $k(\bbG_a^-\times\bbG_a^-)\cong\Lambda(v_1,v_2)$. Then applying Koszul duality you get the 
+symmetric algebra. Also you need the K\"unneth formula, where as algebras
+\[H^*(A\otimes B,k)\cong H^*(A,k)\otimes H^*(A,k).\]
+
+Evans has a book on The groups of cohomology. Here there are computations and resolutions of (cyclic)
+groups. There is also a discussion of spectral sequences which is relatively terse (although it defers
+somewhat to other sources).
+
+\begin{qst}
+	What is the importance of suspensions and what role do they play in this theory? They 
+	have come up at several points and I am realizing they are fairly pervasive.
+
+	In particular, I understand the topological definition, but I don't see why they correspond
+	to what we want. I am looking at the formula from \cite{detecting-nilpotence}:
+	\[H^{1,1}(G,k)=\Hom_{sGr/k}(G^0,\mathbb{G}_a^-)^\pi\]
+	where $\pi$ is the group of connected components of $G$.
+\end{qst}
+
+\begin{qst}
+	In $G=(\bbG_a^-\times\bbG_a^-)\rtimes G_{a(1)}$, where the action is said to be ``nontrivial'',
+	why does this seem to imply that the action on one of the copies of $\bbG_a^-$ is trivial?
+
+	The fact that there is a LHS spectral sequence 
+	\[H^{*,*}(\bbG_a^-\times \bbG_{a(1)},H^{*,*}(\bbG_a^-,k))\Rightarrow H^{*,*}(G,k)\]
+	seems to imply that there is a short exact sequence 
+	\[1\to \bbG_a^-\to G\to \bbG_a^-\times \bbG_{a(1)}\to 1\]
+	but since the product is direct on the right, doesn't this imply that the action of $\bbG_{a(1)}$ on 
+	$\bbG_a^-$ is trivial?
+\end{qst}
+
+Path 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \section{To be filed:}
-The last paper on Julia's page is a preprint with her and Dave Benson. There is a spectral sequence computation that might be illustrative.
-There is also an older paper...
-
-One of the things Julia is interested in at the moment is a super group scheme. These arise by having a $\bbZ_2$ grading on your Hopf algebra.
-For instance we can look at exterior algebra $\Lambda(x)$, which represents the super group scheme $G_a^-$.
-
-The first non-elementary example is $G_a^-\times G_a^-$ with a $\bbZ_p$ action on it that ``folds'' the second factor
-onto the first. To write this out explicitly, we consider the group algebra $\Lambda(u,v)$ where $uv=-vu$ and the coordinate algebra of $\bbZ_p$
-is $k[t]/t^p$ where $t\cdot u=0$ and $t\cdot v=\alpha u$ for some $\alpha$. This is a $4p$ dimensional algebra.
-
-Then we want to compute coholomology. We know $H^{**}(k\bbZ/p,k)\cong k[x]\oplus\Lambda(y)$ for $|x|=2,0$.
-Note there is a notion of dimension that is an ordered pair $(a,b)$ of the graded degree and cohomology degree.
-THis give us new elements of degree $(1,1)$
-
-Look up Koszul duality! It is in the representations and local cohomology book!
 

### 2019-05-16 Homological Algebra Notes/Hom Alg.tex
Haven't pushed in a while

\documentclass[12pt]{article}

\usepackage{setspace}

\usepackage{amsmath, amsfonts, amssymb, graphicx, color, fancyhdr, lipsum, scalerel, stackengine, mathrsfs, tikz-cd, mdframed, enumitem, framed, adjustbox, bm, upgreek, x	color}
\usepackage[framed,thmmarks]{ntheorem}
\usepackage[mathscr]{euscript}

%Replacement for the old geometry package
\usepackage{fullpage}

\input{/home/nico/latex-includes/mydefs.tex}

%class-specific macros
\DeclareMathOperator{\sExt}{\mathscr{E}xt}

%header stuff
\setlength{\headsep}{24pt}  % space between header and text
\pagestyle{fancy}     % set pagestyle for document
\lhead{Notes on Homological Algebra} % put text in header (left side)
\rhead{Nico Courts} % put text in header (right side)
\cfoot{\itshape p. \thepage}
\setlength{\headheight}{15pt}
\allowdisplaybreaks

\begin{document}
%make the title page
\title{Homological Algebra\vspace{-1ex}}
\author{A course by: Prof. Julia Pevtsova\\
Notes by: Nico Courts}
\date{Spring 2019}
\maketitle

\renewcommand{\abstractname}{Introduction}
\begin{abstract}
	These are the notes I took in class during the Winter 2019 topics course
	\textit{Math 509 - Homological Algebra} at University of Washington, Seattle.

	The course description follows:

	\brk

	This course is an introductory course on homological algebra. We will be following the 
	book \textit{An Introduction to Homological Algebra} by Charles Weibel. We will be
	covering the following topics:

\begin{itemize}
	\item Chain complexes, homotopies, homology and long exact sequence in homology
	\item Resolutions, derived functors, Ext and Tor. Koszul complexes
	\item Group (co)homology
	\item Triangulated and derived categories
	\item Spectral sequences or open topic depending on the class interests
\end{itemize}
\end{abstract}

\section{April 1, 2019}

The idea here is that we are going to spend a significant amount of time doing traditional lectures.
After a while we are going to be doing some presentations. We have two weeks to pick groups
(and topics, if desired) before we are assigned to them randomly.

Again, we are using Weibel's book. At the beginning we are going to be following the text quite closely,
so that is a perfect reference.

\subsection{Chain Complexes, Maps, Homotopies, and Homology}
Let $R$ be a ring. In general we will say it is unital and associative, but not much more.
We will be generally working in the category $\mathbf{Mod}_R$ of (left) $R$-modules.

\begin{defn}
	A chain complex $C_\bullet$ is a sequence of $R$-modules and $R$-module homomorphisms
	\[d_n:C_n\to C_{n-1}\]
	where the $C_n$ are $R$-modules and $f_n$ are $R$-module homomorphisms, subject to the relation
	\[d_n\circ d_{n-1}=0.\]
\end{defn}
The first HW problem:
\begin{prob}
	Show that $\Ch$, the category of chain complexes forms an abelian category.
\end{prob}

\begin{defn}
	Given two chain complexes $C_\bullet$ and $D_\bullet$, then $f_\bullet:C_\bullet\to D_\bullet$
	is a \textbf{chain map} if
	\[f_n:C_n\to D_n\]
	is a map of $R$-modules and the obvious diagram commutes. (figure~\ref{fig-commute})
\end{defn}
\begin{figure}\label{fig-commute}
	\centering
	\begin{tikzcd}
		C_n \ar[r,"d_n"]\ar[d,"f_n"] & C_{n-1}\ar[d,"f_{n-1}"]\\
		D_n\ar[r,"d'_n"] & D_{n-1}
	\end{tikzcd}
	\caption{Commuting Figure for Chain Maps}
\end{figure}

\begin{defn}
	Let $C_\bullet$ be a chain complex. Then $Z_n=Z_n(C_\bullet)$ is defined to be
	\[Z_n=\ker d_n\]
	called the $n$-\textbf{cycles} of $C_\bullet$.

	Similarly, define
	\[B_n=\Im d_{n+1}\]
	the $n^{th}$ \textbf{boundary} of $C_\bullet$.

	Then $H_n=Z_n/B_n$, the $n^{th}$ \textbf{homology group}. 
\end{defn}

\subsection{Some Notation}
The category $\mathbf{Ch}_\bullet$ (double underline!!!) is the category of (unbounded) complexes,
but we are also sometimes interested in $\mathbf{Ch}_\bullet^b$, the category of \textbf{bounded}
complexes. Furthermore, we may consider $\mathbf{Ch}_\bullet^{\ge 0}$ or $\mathbf{Ch}_\bullet^{\le 0}$

Furthermore, there are cochain complexes, where we switch the direction and the notation is changed as
we do for any dual notion.

\subsection{Types of Complexes}
\begin{defn}
	A complex $C_\bullet$ is called \textbf{exact} if it is an exact sequence. Equivalently the 
	homology groups are trivial.
\end{defn}
\begin{rmk}
	A complex with trivial homology is also called \textbf{acyclic}. This is obviously equivalent to being 
	exact.
\end{rmk}
\begin{lem}
	Let $f:C_\bullet\to D_\bullet$ be a map of chain complexes. Then it induces a well-defined map on 
	cycles, boundaries, and homology.
\end{lem}
\begin{prf}
	Homework.
\end{prf}

\begin{defn}
	Let $f:C_\bullet\to D_\bullet$ be a chain map. It is called a \textbf{quasi-isomorphism}
	if it induces isomorphisms on homology groups.
\end{defn}
\begin{rmk}
	An acyclic complex $C_\bullet$, then $C_\bullet\to 0$ is a quasi-isomorphism.
\end{rmk}
\subsection{Snake Lemma}
\begin{lem}[Snake]
	Given a commutative diagram with exact rows as in figure~\ref{fig-snake},
	we have a long exact sequence
	\[\ker f\to \ker g\to \ker h\xrightarrow{\delta} \coker f\to \coker g\to \coker h\to 0\]
\end{lem}
\begin{figure}[h!]\label{fig-snake}
	\centering
	\begin{tikzcd}
		& A'\ar[r]\ar[d,"f"] &B'\ar[r]\ar[d,"g"] &C\ar[d,"h"]\ar[r] & 0\\
		0\ar[r]& A\ar[r] & B\ar[r] & C &
	\end{tikzcd}
	\caption{Snake Lemma}
\end{figure}

\subsection{An Example}
We talked about singular homology. Cool!

\section{April 3, 2019}
\subsection{Long Exact Sequence of Homology}
\begin{defn}
	A short exact sequence $0\to A_\bullet\xrightarrow{f} B_\bullet\xrightarrow{g} C\bullet\to 0$
	of chain complexes is a width four double complex with exact rows.
\end{defn}
\begin{thm}
	Let $0\to A_\bullet\to B_\bullet\to C_\bullet\to 0$ be a short exact sequence of complexes. Then 
	there exists a (natural) long exact sequence in homology:
	\[\cdots\to H_n(A)\to H_n(B)\to H_n(C)\to H_{n-1}(A)\to\cdots\]
\end{thm}
\begin{rmk}
	The statement of naturality just means that the maps above are induced from those in the SES
	under application of $H_n$ functor.
\end{rmk}
\begin{prf}
	Snaaaaaaaake.

	(Sketch) Use first the fact that $d:A_n\to A_{n-1}$ factors through both $A_n/B_n(A_\bullet)$ and $Z(A_\bullet)$,
	so using the induced maps between these quotients to get a nice little snake-like diagram:

	\begin{center}
	\begin{tikzcd}
		& A_n/B(A_n)\ar[r]\ar[d] & B_n/B(B_n)\ar[r]\ar[d] & C_n/B(C_n)\ar[r]\ar[d] & 0\\
		0\ar[r] & Z_{n-1}(A)\ar[r] & Z_{n-1}(B)\ar[r] & Z_{n-1}(C)\ar[r] & 0
	\end{tikzcd}
	\end{center}

	where the exactness of the rows above follows from applications of snake lemma to the original
	SES of chains.

	A final application of snake gets us the SES we desire.
\end{prf}
\subsection{Some Topology}
\begin{defn}
	Let $f:C_\bullet\to D_\bullet$ be a chain map. We say $f$ is \textbf{nullhomotopic} if there are $s_n:C_n\to D_{n+1}$
	for all $n$ such that
	\[f_n=d_{n+1}\circ s_n+s_{n-1}\circ d_n.\]

	Sometimes we just write $f=ds+sd$ and let the indicies figure themselves out.
\end{defn}
\begin{defn}
	If $f,g:C_\bullet\to D_\bullet$ are chain maps, we say that they are \textbf{chain homotopic}
	if $f-g$ is nullhomotopic.

	Here we write $f\sim g$, so $f\sim 0$ means $f$ is nullhomopotic.
\end{defn}
\begin{defn}
	We say $C_\bullet$ and $D_\bullet $ are \textbf{chain homotopy equivalent} if there
	exist maps $f:C\to D$ and $g:D\to C$ such that $f\circ g\sim\id_D$ and $g\circ f\sim \id_C$.
\end{defn}
\begin{defn}
	We say a complex $C_\bullet$ is \textbf{contractible} if $\id_C\sim 0$.
\end{defn}
\begin{rmk}
	Some motivation: say we have $X,Y$ topological spaces and $f,g:X\to Y$ continuous maps. Then $f\sim g$
	if there is a continuous map $H:X\times [0,1]\to Y$ such that $H(x,0)=f(x)$ and $H(x,1)=g(x)$.

	Now if $f\sim g$, then the natural (induced) maps $S_\bullet(f)$ and $ S_\bullet(g)$ on singular 
	chain complexes are chain homotopic.
\end{rmk}
\begin{thm}
	If $f,g:C_\bullet\to D_\bullet$ are chain homotopic, then they induce the same map on $H_\bullet$.
\end{thm}
\begin{prf}
	We will be using that homology is an additive functor: $H_n(f)-H_n(g)=H_n(f-g)$. Then it suffices to show 
	that if $f\sim 0$ then $H_n(f)=0$ for all $n$.

	Assume $f\sim 0$ with homotopy $s_\bullet:C_\bullet\to D_\bullet[1]$. Then we have
	\begin{center}
		\begin{tikzcd}
			C_{n+1}\ar[r,"d"] & C_n\ar[r,"d"]\ar[ld,"s"]\ar[d,"f"] & C_{n-1}\ar[ld,"s"]\\
			D_{n+1}\ar[r,"d"] & D_n\ar[r,"d"] & D_{n-1}
		\end{tikzcd}
	\end{center}

	\noindent and so if $a\in Z_n(C)$, then say $H_n(f)(a)=f(a)$. But 
	\[f(a)=(ds+sd)(a)=d_{n+1}(s(a))+s(d_n(a))=d_{n+1}(s(a))\in B_n(D).\]
\end{prf}
\begin{rmk}
	We can define the \textbf{homotopy category} $K_\bullet(\mathbf{Mod}_R)$ or $K_\bullet(\mathbf{Ch}_\bullet)$
	where the objects are chain complexes and the morphisms are chain homotopy equivalence classes of maps.

	This category is additive and in our homework we will investigate whether it is abelian. So then 
	our homology functors $H_n:\Ch(R)\to \mathbf{Ab}$ factor through $K_\bullet(R)$.
\end{rmk}

\subsection{Split exact complexes}
\begin{defn}
	$C_\bullet$ is a \textbf{split exact sequence} if its an exact sequence that splits. That is,
	\[C_n=B_{n+1}(C_\bullet)\oplus B_n(C_\bullet)\]
	or since the sequence is exact, the sum of cycles.

	The boundary maps simply map isomorphic summands to one another.
\end{defn}
\begin{ex}
	$R=k$. Then $\mathbf{Mod}_R$ is actually $\mathbf{Vect}_k$. All exact sequences here split.
\end{ex}

\section{April 5, 2019}
\begin{ex}
	Continuing from our example on Wednesday, let $R=k$ and consider $\Ch(k)$. Everything in $\Vectk$ splits, 
	so we can write $C_n=Z_n\oplus C_n/Z_n\cong Z_n\oplus B_{n-1}$.

	But then $C_n\cong B_n\oplus H_n\oplus B_{n-1}$, so $C$ is exact iff $H_n=0$ (it is acyclic)
	iff $C$ is split exact.
\end{ex}
\begin{rmk}
	In general, split exact implies acyclic, but the converse is not true.
\end{rmk}
\begin{ex}
	Consider $\Ch(\bbZ)$ and consider the chain of $\cdots \to\bbZ/4\to\bbZ/4\to\bbZ/4\to\cdots$
	with differential $\ast 2$.

	Notice that $H_n(C)=0$, but it is NOT nullhomotopic/contractible! This is because $\id=sd+ds$ maps
	only to even numbers!
\end{ex}

\subsection{Operations with complexes}
We can do some operations on complexes $C\in\Ch$:
\begin{itemize}
	\item $(C[d])_i=C_{i+d}$
	\item Total complexes
\end{itemize}
\begin{defn}
	A \textbf{double (chain) complex} is a collection of $R$-modules and maps $C_{pq}$ 
	\begin{center}
		\begin{tikzcd}
			C_{pq}\ar[r,"d^h"]\ar[d,"d^v"] & C_{p-1\,q}\ar[d,"d^v"]\\
			C_{p\,q-1}\ar[r,"d^h"] & C_{p-1\,q-1}
		\end{tikzcd}
	\end{center}

	such that $(d^h)^2=(d^v)^2=d^h\circ d^v+d^v\circ d^h=0.$
\end{defn}
\begin{ex}
	Let $f:C\to C'$ be a chain map. Then

	\begin{center}
		\begin{tikzcd}
			C_{n+1}\ar[r,"-f"] \ar[d,"-d"] & C_{n+1}'\ar[d,"d'"]\\
			C_n\ar[r,"-f"]\ar[d,"-d"] & C_n'\ar[d,"d'"]\\
			C_{n-1}\ar[r,"-f"] & C_{n-1}'
		\end{tikzcd}
	\end{center}

	Is a double complex. Notice we also could use $f$ instead of $-f$ but for some reason this 
	is usually what we want.
\end{ex}
\begin{defn}
	Let $(C_{pq},d)$ be a double complex. Then
	\[\operatorname{Tot}_n(C_{\bullet\bullet})=\oplus_{p+q=n}C_{pq}\]
	with differential $d=d^v+d^h$. This is a chain complex.
\end{defn}
\begin{rmk}
	Notice that $(d^h+d^v)^2=d^h\circ d^h+d^h\circ d^v+d^v\circ d^h+d^v\circ d^v=0.$ So this
	is indeed a chain complex.
\end{rmk}
\begin{rmk}
	Notice that we have technically used the fact that our category contains coproducts to
	define $\operatorname{Tot}=\operatorname{Tot}^\oplus$ as we just did. These should exist 
	in an abelian category.
\end{rmk}
\begin{ex}
	Let $f:B\to C$ be a chain map. Then $\operatorname{Cone}(f)$ is the total complex of the associated
	double complex of $f$.

	Here $(\operatorname{Cone}(f))_n=B_{n-1}\oplus C_n$ (careful, the indices are a bit wonky here.)
	Then the differential is given by $(\begin{smallmatrix}
		-d_B & 0\\ -f & d_C
	\end{smallmatrix})$.
\end{ex}
\begin{rmk}
	Notice that
	\[B_\bullet\xrightarrow{f} C_\bullet\to (\operatorname{Cone}(f))_\bullet\to B_\bullet[-1]\]
	is a short(?!) exact sequence of complexes.
\end{rmk}
\begin{rmk}
	This idea comes from an idea in topology: given a continuous map $f:X\to Y$
\end{rmk}

\section*{April 8, 2019}
Some quick comments about the topological analogy we discussed last time (I trailed off because
there was a question about what the map between topological spaces induces this map on homology):

The idea is that although the image of $f$ is a deformed copy of $X$ in $\operatorname{Cone}(f)$, 
the suspension $\Sigma X$ is somehow recovered in the cone over $\Im f$.

This is a bit hand-wavy, but we then recover via the long exact sequence in homology the sequence
we were cosidering last time.

\subsection*{Derived Functors, $\Ext$, and $\Tor$}
Recall that we are working over $\Rmod$.
\begin{defn}
	An $\Rmod$ $P$ is projective if one of the following equivalent conditions hold:
	\begin{itemize}
		\item This diagram commutes:
		\begin{center}\begin{tikzcd}
		& P\ar[ld,swap,dashed,"\tilde f"]\ar[d,"f"]\\
		M\ar[r,two heads,"\varphi"]& N
		\end{tikzcd}\end{center}
		\item $\Hom_R(P,-):\Rmod\to\Ab$ is an exact functor.
		\item $P$ is a direct summand of a free module. \textit{Notice that this only works in $\Rmod$}
	\end{itemize}
\end{defn}
\begin{rmk}
	Notice that we are generally assuming (as in the definition of an exact functor) that our
	functors are additive -- that is, it preserves sums. 
\end{rmk}

A question one may ask: what are the projective objects in $\Ch(R)$? This will be the topic of a homework 
question.
\begin{defn}
	A \textbf{projective resolution} $P_\bullet$ of an $R$ module $M$ is a complex
	\[\cdots\to P_2\to P_1\to P_0\]
	of projective $R$ modules such that if we tack on $\varepsilon:P_0\to M$, it becomes exact.

	Equivalently, the chain map induced by $(\cdots,0,\varepsilon):P_\bullet\to (\cdots\to 0\to M\to 0)$ 
	is a quasi-isomorphism.
\end{defn}
\begin{defn}
	A category $\mathcal A$ has \textbf{enough projectives} if, for all $A\in\mathcal A$, there exists
	a projective $P\in\mathcal A$ and an epimorphism $P\to A$.
\end{defn}
\begin{lem}
	$\Rmod$ has enough projectives.
\end{lem}
\begin{prf}
	Let $P_0$ be any free module over $M$ (say the free module on all elements of $M$) and say $\varepsilon:P_0\to M$ is the covering map.
	and let $\Omega M=\ker \varepsilon$, the zeroth syzygy of $M$. Then let $P_1$ be a projective cover of $\Omega_0M$
	and continue in the fashion.
\end{prf}
\begin{cor}
	Any $R$-module has a projective resolution.
\end{cor}

\subsection*{Fundamental Theorem of Homological Algebra}
\begin{thm}[Fundamental Theorem]
	If $P_\bullet$ is a projective complex with map $\varepsilon_M:P_0\to M$ (not necessarily projective) and $Q_\bullet\to N$ is a projective 
	resolution of $N$ and $f:M\to N$ is an $R$-linear map, then there exists an $F:P_\bullet \to Q_\bullet$ 
	extending $f$. That is it makes following diagram commute:
	\begin{center}
		\begin{tikzcd}
			P_\bullet\ar[r,"\varepsilon_M"]\ar[d,"F",dashed] & M\ar[d,"f"]\\
			Q_\bullet \ar[r,"\varepsilon_N"] & N
		\end{tikzcd}
	\end{center}

	Furthermore $F$ is unique up to chain homotopy.
\end{thm}
\begin{rmk}
	Notice that while any two extensions are chain homotopic, this doesn't mean there exists
	a \textit{unique} homotopy between them and this is the cause of many problems in homological algebra.
\end{rmk}
\begin{prf}
	Begin by lifting $f\circ \varepsilon_M$ to a map $f_0:P_0\to Q_0$ using the definition of projectivity.
	Then construct $f_n$ by induction. Consider
	\begin{center}
		\begin{tikzcd}
			P_n\ar[r,"d^P"] \ar[rd] & P_{n-1}\ar[r]\ar[d,"f_{n-1}"] & P_{n-2}\ar[d,"f_{n-2}"]\\
			Q_n\ar[r] & Q_{n-1}\ar[r,"d^Q"] & Q_{n-2}
		\end{tikzcd}
	\end{center}
	and by the induction hypothesis, $d^Q\circ f_{n-1}\circ d^P=0$ whence $\varphi:P_n\to Q_{n-1}$ 
	maps to $\ker d^Q_{n-1}=\Im d^Q_n$. But this gives us a map $P_n\to \Im d^Q$ onto which $Q_n$ surjects,
	so again we leverage the projectivity of $P_n$ to lift to $f_n$.

	To see the uniqueness, assume $F$ and $F'$ are two liftings of $f$. We want to show $H=F-F'\sim 0$.
	It will be sufficient once we see that any lifting of $0:M\to N$ is nullhomotopic. We will see this next time.
\end{prf}

\section{April 10, 2019}
\begin{thm}
	... and it is unique up to chain homotopy.
\end{thm}
\begin{prf}
	Say that $F$ and $F'$ are two chain maps extending $f:M\to N$ in the above scenario.
	Notice that since we want to show $F\sim F'$, it suffices to show that $0:M\to N$ induces 
	a nullhomotopic map $H$.

	Let $H$ be a lifting of $0$ and define $s_i$ such that
	\begin{center}
		\begin{tikzcd}
			P_1\ar[r]\ar[d,"h_1"] & P_0\ar[dl,"s_0"]\ar[d,"h_0"],\ar[r] & M\ar[d,"0"]\ar[dl,"s_{-1}=0"]\\
			Q_1\ar[r] & Q_0 \ar[r,"\varepsilon_N"] & N
		\end{tikzcd}
	\end{center}
	where we need to have that $h_0=s_0\circ d_Q$ for some $s_0$. we know that $h_0(P_0)\subseteq\ker\varepsilon_N=d_Q(Q_1)$
	so lifting $h_0$ through $Q_1\twoheadrightarrow d(Q_1)$, we get a map $s_0:P_0\to Q_1$ (since $P_0$ is projective).

	Continuing my induction, consider
	\begin{center}
		\begin{tikzcd}
			P_{n+1}\ar[r]\ar[d,"h_{n+1}"] & P_n\ar[dl,"s_n"]\ar[d,"h_n"],\ar[r] & P_{n-1}\ar[d,"h_{n-1}"]\ar[dl,"s_{n-1}"]\\
			Q_1\ar[r] & Q_0 \ar[r,"\varepsilon_N"] & Q_{n-1}
		\end{tikzcd}
	\end{center}
	and consider the map $s_n=h_n-(s_{n-1}\circ d_P)$, which maps $P_n$ into $d_Q(Q_{n+1})$. Now $d_Q(Q_{n+1})=\ker d_Q$ (easy lemma)
	This establishes that $s_n$ exists and satisfies all the requisite properties.
\end{prf}
\begin{cor}
	Any two projective resolutions of $M$ induce the same homology.
\end{cor}
\begin{rmk}
	Lift the identity map $M\to M$ in both directions to get chain maps $(P\to M)\to(Q\to M)$
	and backwards. These compose to the identity map! Then apply the theorem!
\end{rmk}

\begin{lem}[Horseshoe]
	(\textbf{Weibel 2.2.8}) Say that we have a short exact sequence $0\to A\to B\to C\to 0$ 
	and projective resolutions $P^A_\bullet$ and $P_\bullet^C$ of $A$ and $C$, respectively.
	Then $P_\bullet^B=(P_i^A\oplus P_i^C)_{i\in\bbZ}$ is a resolution of $B$ and furthermore we have 
	maps for the following diagram (which commutes everywhere):

	\begin{center}
		\begin{tikzcd}
			P_2^A\ar[r] & P_1^A\ar[r]\ar[d] & P_0^A\ar[r]\ar[d] & A\ar[d]\\
			 & P_1^A\oplus P_1^C\ar[r]\ar[d] & P_0^A\oplus P_0^B\ar[r]\ar[d] & B\ar[d]\\
			P_2^C\ar[r] & P_1^C\ar[r] & P_0^C\ar[r]s & C
		\end{tikzcd}
	\end{center}
\end{lem}

\subsection{Left Derived Functors}
\begin{defn}
	Let $\mathcal A, \mathcal B$ be Abelian categories and let $F:\mathcal A\to \mathcal B$
	be an additive functor. We say that $F$ is \textbf{right exact} if for every exact sequence
	$A\to B\to C\to 0$ in $\mathcal A$, 
	\[F(A)\to F(B)\to F(C)\to 0\]
	is exact in $\mathcal B$.
\end{defn}
\begin{defn}
	Assume $\mathcal A$ is an abelian category with enough projective and that $\mathcal B$ is abelian.
	Let $F:\mathcal A\to \mathcal B$ be a right exact functor. Then $L_iF$ for $i\ge 0$ are called
	\textbf{the left derived functors of $F$} which are defined as follows:

	For all $A\in\mathcal A$, let $P_\bullet\to A$ be a projective resolution. Then we have an 
	exact sequence
	\[\cdots\to F(P_2)\to F(P_1)\to F(P_0)\]
	and we define $(L_iF)(A):=H_i(F(P_\bullet))$.
\end{defn}
\begin{rmk}
	\begin{enumerate}
		\item $L_i F$ are functors because of the fundamental theorem.
		\item The functors are well-defined because of the lemma.
	\end{enumerate}
\end{rmk}
The derived functors have some nice properties:
\begin{itemize}
	\item $L_i F= 0$ for $i<0$.
	\item Say $0\to A\to B\to C\to 0$ is short exact and $F$ is right exact. Then we get a long exact sequence
	\[\cdots \to L_1F(A)\to L_1F(B)\to L_1F(C)\to F(A)\to F(B)\to F(C)\to 0\]
\end{itemize}

\section{April 17, 2019}
Last time we wrote down the definition for (left) derived functors, but never wrote out why $\Tor$ is an example.
\subsection{Examples of $L_i$}
Let $R$ be a ring and $M$ and $N$ be right- and left-$R$-modules, respectively. Then $M\otimes_RN$ 
is the set of $m\otimes n$ for $m\in M$ and $n\in N$ where $m\cdot a\otimes n=m\otimes a\cdot n$
for all $a\in R$.

This can also be defined formally as a bifunctor $-\otimes_R-:\Rmod\times\modR\to\Ab$, or as a regular
functor by currying. So then

\begin{defn}
	\[\Tor_i^R(M,N):=L_i(M\otimes_R-)(N)\]
	and
	\[\Tor_i^R(M,N)=L_i(-\otimes_R M)(N)\]
\end{defn}
\begin{rmk}
	Notice that the two definitions are completely equivalent and compatible. That is, $\Tor_i^R(M,N)$ 
	is what we call \textbf{balanced.}

	To see this you can do a few computations including computing the homology of the total complex to show that 
	you always get the same thing.
\end{rmk}

\subsection{Injective Modules}
\begin{defn}
	Let $\mathcal A$ be an abelian category. Then $E\in \mathcal A$ is injective if we have the diagram
	\begin{center}
		\begin{tikzcd}
			M\ar[d,"f"]\ar[r,hook] & N\ar[dl,dashed,"\tilde f"]\\
			E &
		\end{tikzcd}
	\end{center}
\end{defn}
\begin{lem}
	If $\mathcal A=\Rmod$, then in the above definition it suffices to check that maps $f:I\to E$ from
	every \textit{ideal} lift to maps $\tilde R\to E$ from $R$.
\end{lem}
\begin{ex}
	Injective $\bbZ$-modules are precisely the divisible abelian groups. For instance, $\bbQ$, $\bbQ/\bbZ$, $\bbZ_{p^\infty}=\bbZ_{(p)}/\bbZ$.
\end{ex}
In fact:
\begin{lem}
	The indecomposable injective $\bbZ$ moduels are $\bbQ$ and $\bbZ_{p^\infty}$.
\end{lem}
\begin{thm}
	For any ring $R$, $\modR$ has enough injectives.
\end{thm}
\begin{cor}
	Every $M\in\modR$ has an injective resolution.
\end{cor}
\begin{prf}
	It suffices to show that for all $M\in\modR$, there is an injective $R$-module $E$ 
	such that $M\hookrightarrow E$.

	To begin, we will see that $\bbZ$-$\mathbf{mod}=\Ab$ has enough injectives. Let $A\in\Ab$
	and define the notation
	\[A^{\vee}=\Hom_{\Ab}(A,\bbQ/\bbZ).\]
	Then $\bbZ^\vee=\bbQ/\bbZ$, since any map is determined uniquely by its image. So then if $F=\bbZ^n$,
	we use linearity of $\Hom$ to prove $F^\vee=(\bbQ/\bbZ)^n$, so all free groups are divisible (whence injective).

	Finally, for all $A$, there is a natural map $A\to A^{\vee\vee}$. Now for any $A\in\Ab$, there exists a 
	surjective map $F\to A^{\vee}$ from a free module onto the dual of $A$. Thus we get an embedding
	$A^{\vee\vee}\hookrightarrow F^\vee$ (note that $\Hom(-,\bbQ/\bbZ)$ is exact). Since $A\hookrightarrow A^{\vee\vee}$,
	this gets us an embedding of $A$ into $F^\vee$.

	\brk

	Now let $R$ be any ring. Let $A\in\Ab$, then the claim is that $\Hom_\bbZ(R,A)$ is a (right) $R$-module via
	the action $(f\cdot a)(b)=f(ba)$ (\textbf{check this}).

	Next, we claim that there exists a natural isomorphism for all $M\in\modR$ and $A\in\Ab$
	such that
	\[\Hom_\bbZ(M,A)\cong\Hom_R(M,\Hom_\bbZ(R,A)).\]
	We claim that one can ``follow their nose'' here.

	Using these facts, we will finish the proof of the main theorem: Let $A$ be a divisible abelian group.
	Then $\Hom_\bbZ(R,A)$ is injective. This is because 
	\[\Hom_R(-,\Hom_\bbZ(R,A))\]
	is exact since $A$ is injective (by the last claim).

	Let $M$ be any right $R$-module. Then in particular it is an abelian group, so 
	$M\hookrightarrow A\in\Ab$, where $A$ is divisible. Thus (again by the last fact) we have a map
	$\varphi:M\to \Hom_\bbZ(R,A)$ is a map of $R$-modules. Then it suffices to show that $\varphi$ is injective.

	This follows from the explicit construction of the natural isomorphism between the functors.
\end{prf}
\begin{rmk}
	The map referred to above takes $f\in \Hom_Z(M,A)$ to a map $\varphi_f\in\Hom(M,\Hom_\bbZ(R,A))$
	where
	\[\varphi_f(a)(r)=(f\cdot a)(r).\]

	To show it is injective we just construct the inverse map: let $\psi_g$ for any $g:M\to\Hom_\bbZ(R,A)$
	to be
	\[\psi_g(a)=(g(a))(1).\]
	One can easily check that this composes to the identity in each direction.
\end{rmk}

\section{April 19, 2019}
I came late to the discussion, but it looks like we defined left exact contravariant functors --
those are functors that are left exact as functors $F:C^{op}\to D$.

\begin{rmk}
	$\Hom_R(-,W)$ is left exact.
\end{rmk}

From here, we can continue the sequence to a long exact sequence of homology. The point of this 
section was to show that we can compute $\Ext_R^i(V,W)$ \textit{either} from a projective resolution of $V$ or 
an injective resolution of $W$.

\begin{lem}
	$P$ is a projective $R$ module if and only if $\Ext_R^i(P,-)=0$ for all $i>0$.
\end{lem}
\begin{lem}
	$I$ is an injective $R$ module if and only if $\Ext_R^i(-,I)=0$ for all $i>0$
\end{lem}
\subsection{Adjoint Functors}
\begin{defn}
	We say the functors $\cal F:\cal A\to \cal B$ and $\cal G:\cal B\to \cal A$ are \textbf{adjoint functors}
	if there exists a natural isomorphisms
	\[\eta_{AB}:\Hom_{\cal B}({\cal F}(A),B)\to \Hom_{\cal A}(A,{\cal G}(B))\]
	and in this case we say $\cal F$ is left adjoint to $\cal G$.
\end{defn}
\begin{ex}
	Recall we hat the functors $\cal F$ and $\cal G$ from $\bbZ$-modules to $R$-modules and back, respectively,
	where $\cal F=\Hom_\bbZ(R,-)$, assigning to each $\bbZ$ module an $R$-module with the action we discussed last time.

	Then $\cal G=\operatorname{Res}_\bbZ^R$, the forgetful functor, is its left adjoint.
\end{ex}
\begin{ex}
	In rep theory we have Frobenius Reciprocity! Woo.
\end{ex}

\section{April 22, 2019}
Recall the tensor-hom adjunction: $-\otimes_S B$ and $\Hom_R(B,-)$ are adjunctions between $\modR$ and $\mathbf{mod}$-$S$.
There is some shadiness here (mirroring what we saw before) but the fact that 
\[(f.s)(b)=f(sb)\]
defines a right $S$ action on $\Hom_R(B,C)$ is not immediately evident to me.

Again, many of the common adjunctions we see are actually just special cases of tensor-hom.

\subsection{Why Adjointness Matters}
\begin{lem}
	If $\cal F$ and $\cal G$ are adjoints ($\cal F$ is left adjoint to $\cal G$), then $\cal F$ is
	right exact and $\cal G$ is left exact.
\end{lem}
\begin{prf}
	Let $0\to A' \to A\to A''\to 0$ be a short exact sequence. Then consider the diagram
	\begin{center}
		\begin{tikzcd}
			\Hom_{\cal A}(A',\cal G(B))\ar[d,"\sim"]& \ar[l] \Hom_{\cal A}(A,\cal G(B))\ar[d,"\sim"] &\ar[l]\Hom_{\cal A}(A'',\cal G(B)\ar[d,"\sim"])& \ar[l]0\ar[d,"\sim"]\\
			\Hom_{\cal B}(\cal F(A'),B) & \ar[l] \Hom_{\cal B}(\cal F(A),B) & \ar[l] \Hom_{\cal B}(\cal F(A''),B) & \ar[l] 0
		\end{tikzcd}
	\end{center}
	and since the top sequence is exact (by left exactness of $\Hom$), and by the naturality of the isomorphism
	between the functors, we get exactness below.
\end{prf}
\begin{rmk}
	There is a fact that $C'\to C\to C''\to 0$ is exact \textbf{if and only if} $\Hom_{\cal B}(C_\bullet,B)$ is 
	exact \textit{for all $B$.} One direction is just left exactness, the other basically uses Yoneda lemma.
\end{rmk}
\begin{thm}
	If $\cal F$ and $\cal G$ are adjoint (left and right, resp.), then
	\begin{enumerate}
		\item $\cal F$ commutes with colimits (e.g. coproducts, $\oplus$, cokernels).
		\item $\cal G$ commutes with limits (e.g. products and kernels).
	\end{enumerate}
\end{thm}
\begin{cor}
	Let $\cal A$ be an abelian category with enough projectives and let $\cal F:\cal A\to \cal B$ be right exact.
	Then
	\[(L_i\cal F)\left(\bigoplus_j A_j\right)\cong \bigoplus_j L_i\cal F(A_j).\]
\end{cor}

\begin{ex}
	Let $R$ be a PID. Consider torsion modules (free ones are uninteresting when computing $\Tor$).
	Let's compute $\Tor_\ast^R(R/(a),M)$ for any $M$. Take the resolution
	\[0\to R\xrightarrow{\cdot a} R\to R/(a)\]
	which is free whence projective. Then compute the homology of
	\[0\to R\otimes_R M\xrightarrow{\cdot a} R\otimes_RM\to 0\]
	and since $R\otimes_RM\cong M$, we get $\Tor_{\ge 2}^R(R/(a),M)=0$. Then $\Tor_0^R(R/(a),M)=R/(a)\otimes_R M\cong M/aM$.

	Finally $\Tor_1^R(R/(a),M)={_a}M$, the $a$-torison in $M$.
\end{ex}
\begin{rmk}
	Because of the fact that derived functors of right exact functors preserve colimits, we have actually effectively computed
	$\Tor_i^R(N,M)$ for all $N$.
\end{rmk}
\begin{ex}
	In the case when $R=Z$ and $N=\bbQ/\bbZ$, we can use the fact that $\bbQ/\bbZ$ is the colimit (directed limit)
	of its finite subgroups. Then you can pull this out and say that $\Tor_*^R(\bbQ/\bbZ,M)$ is the ``total torsion module'' of $M$,
	the directed limit of all finite torsion subgroups (is this last sentence correct?).
\end{ex}

\section{April 24, 2019}
We're going to start by doing some more computations.
\subsection{Computing $\Ext$}
Begin by letting $A=\bbZ$. Then $\Ext_\bbZ^{i>0}(\bbZ,B)=0$ since $\bbZ$ is free whence projective.
When $A=\bbZ/n$, we can compute $\Ext$ via the free resolution
\[0\to \bbZ\xrightarrow{\cdot n}\bbZ\to \bbZ/n\]
and then applying $\Hom_\bbZ(-,B)$, we get the complex 
\[0\to \Hom_\bbZ(\bbZ,B)\xrightarrow{\cdot n}\Hom_\bbZ(\bbZ,B)\to 0\]
so $\Ext_\bbZ^{i\ge 2}(\bbZ/n,B)=0$ and $\Ext_\bbZ^1(\bbZ/n,B)=B/nB$ and $\Ext_\bbZ^0(\bbZ/n,B)={_n}B/$.

\begin{thm}
	$\Ext_{\bbZ}^{i\ge 2}(A,B)=0.$
\end{thm}
\begin{prf}
	Take an injective resolution of $B$:
	\[B\to I^0\to I^1\cong\Omega^{-1}B\to 0\]
	and every resolution can end there! This is because the cokernel of $B\to I^0$ is the 
	quotient of a divisible group, whence divisible.
\end{prf}

\begin{rmk}
	Notice that $\operatorname{gldim}(\bbZ)=1$.
\end{rmk}
Notice that when $B=\bbZ$, we can consider the injective resolution
\[\bbZ\to \bbQ\to \bbQ/\bbZ\to 0\]
amd then compute from the complex
\[\Hom_\bbZ(A,\bbQ)\to\Hom_Z(A,\bbQ/\bbZ)\]
where the first is zero and the latter is $A^\vee$, the \textbf{Pontragin dual} of $A$. Here
we get that $\Ext_\bbZ^1(A,\bbZ)\cong \Hom_\bbZ(A,\bbQ/\bbZ)=A^\vee$.

Now let $A=Z/p^\infty=\lim_\to \bbZ/p^n$. Then 
\[\Ext_\bbZ^1(\lim_\to\bbZ/p^ny,\bbZ)=\lim_\leftarrow(\bbZ/p^n,\bbZ)=\lim_\leftarrow \bbZ/p^n=\widehat\bbZ_p\]
the $p$-adic numbers.

Recall:
\begin{lem}
	$P$ is projective iff $\Ext^1(P,-)$ is exact if and only if $\Ext^i(P,-)$ is exact for all $i\ge 1$.
\end{lem}
\begin{defn}
	$M$ is a flat module if $-\otimes_R M$ is exact.
\end{defn}
\begin{thm}
	THe following are equivalent
	\begin{itemize}
		\item $M$ is flat
		\item $-\otimes_RM$ is exact
		\item if $N'\hookrightarrow N$ then $N'\otimes_RM\hookrightarrow N\otimes_RM$.
		\item $\Tor_i^R(-,M)=0$ for $i\ge 1$
		\item $\Tor_1^R(-,M)=0$
	\end{itemize}
\end{thm}
\begin{rmk}
	Any projective module is flat.
\end{rmk}

\subsection{Local Properties of Modules}
Here let $R$ be a commutative ring, $\mathfrak p\lhd R$ is a prime, and $R_{\mathfrak p}$ is localization at $\mathfrak p$
and $(-)_{\mathfrak p}$ is a functor from $R$-mod to $R_{\mathfrak p}$-mod.
\begin{prop}
	$(-)_{\mathfrak p}$ is exact.
\end{prop}
\begin{defn}[Informal]
	We say that a property $P$ is local if $M$ satisfies $P$ if and only if $M_{\mathfrak p}$ satisfies $P$ for all $\mathfrak p\lhd R$.
\end{defn}
\begin{prop}
	$M\cong 0$ if and only if
	\begin{itemize}
		\item $M_{\mathfrak p}=0$ for all $\mathfrak p$
		\item $M_{\mathfrak m}=0$ for all maximals $\mathfrak m$
	\end{itemize}
\end{prop}
\begin{rmk}
	Other examples of local properties are flatness of a module and exactness of a short exact sequence.
\end{rmk}
\begin{prop}
	If $R$ is commutative, and if $S$ is flat as an $R$-algebra (flatness is just referring to the module structure), then
	\[S\otimes_R\Tor^R_i(A,B)\cong \Tor_i^S(S\otimes_R A,S\otimes_R B).\]
\end{prop}
\begin{rmk}
	The idea above is ``flat base changes commute with $\Tor$.''
\end{rmk}
\begin{cor}
	The following are equivalent:
	\begin{itemize}
		\item $\Tor_i^R(M,N)=0$
		\item For all $\mathfrak p$, $\Tor_i^{R_{\mathfrak p}}(M_{\mathfrak p},N_{\mathfrak p})=0$
		\item For all maximal $\mathfrak m$, $\Tor_i^{R_{\mathfrak m}}(M_{\mathfrak m},N_{\mathfrak m})=0$
	\end{itemize}
\end{cor}

\section{April 26, 2019}
\subsection{A bit more about flat modules}
\begin{prop}
	Let $R$ be a local ring and let $M$ be a finitely generated $R$-module. Then the following are equivalent:
	\begin{itemize}
		\item $M$ is flat.
		\item $M$ is free.
		\item $M$ is projective.
	\end{itemize}
\end{prop}
\begin{rmk}
	In other words, every flat module is locally free.
\end{rmk}
\begin{rmk}
	In general we have the inclusions: free, projective, flat, torsion free.
\end{rmk}

\subsection{$\Ext$ as Extensions}
Recall the five lemma 
\begin{lem}[Weak 5-lemma]
	If we have the commutative diagram with exact rows:
	\begin{center}
		\begin{tikzcd}
			0\ar[r] & A\ar[r]\ar[d,"\sim"] & B\ar[r]\ar[d,"f"] & C\ar[r]\ar[d,"\sim"] & 0\\
			0 \ar[r] & A'\ar[r] & B'\ar[r] & C'\ar[r] & 0
		\end{tikzcd}
		then $f$ is an isomorphism.
	\end{center}
\end{lem}

\begin{defn}
	Define $\sExt_R^1(M,N)$ to be the set of short exact sequences $0\to N\ to E\to M\to 0$
	modulo the relation that if there is a map (necessarily an ismomorphism) $\varphi:E\to E'$ extending 
	the identity maps on $M$ and $N$, then the two sequences are equivalent.

	We call $0\to N\to E\to M\to 0$ a \textbf{extension of $M$ by $N$ of length 1.}
\end{defn}
\begin{prop}
	$\sExt_R^1(N,M)$ has the structure of an abelian group.
\end{prop}
\begin{rmk}
	Let $\alpha:M'\to M$ and let $E\times^MM'$ be the pullback over
	\[E'\xrightarrow{g} M\xleftarrow{\alpha}M'\]
	where where $0\to N\to E\to M\to 0$ is an extension.

	Then setting $E'=E\times^MM'$, we can pull back along $\alpha$ to a short exact sequence 
	\[0\to N\to E'\to M'\to 0.\]
	Here we usually refer to this map (of short exact sequences) as $\alpha^*:\sExt^1(M,N)\to \sExt^1(M',N)$.

	Dually, we can use a $\beta:N\to N'$ and the pushforward $E''=N'\times_NE$ under
	\[N'\xleftarrow{\beta} N\xrightarrow{f}E\]
	and this gives us a map to the sequence $0\to N'\to E''\to M\to 0$. This gives us 
	the map $\beta_*:\sExt^1(M,N)\to \sExt^(M,N')$.
\end{rmk}

\begin{rmk}
	A natural thing to want to do is sum SES's by using direct sums of the elements in the sequence.
	Call this $\xi\oplus \xi'$.
	That is great, but it doesn't quite work. To do so, you use the (co)diagonal maps $\Delta:M:M\oplus M$
	sending $m\mapsto(m,m)$ and $\nabla:N\oplus N\to N$ via $(n,n')\mapsto n+n'$.

	Then we can define a sum of extensions:
	\[\xi+\xi':= \nabla_*\Delta^*(\xi\oplus\xi')=\Delta^*\nabla_*(\xi\oplus\xi')\]
	The last equality isn't completely trivial.
\end{rmk}
From here we talked about how one proves that this is indeed a group. The zero element is the split exact sequence.
\begin{prop}
	There exists a natural group isomorphism
	\[\sExt^1_R(M,N)\cong\Ext_R^1(M,N).\]
\end{prop}
\begin{rmk}
	The proof isn't too hard. Check Weibel.
\end{rmk}

\brk

In general, for $i\ge 1$, we use an analogous definition but now we define two extensions
to be equivalent if there exist a collection of ($i$) maps $f_k:E_k\to E_k'$ making the diagram commute.

The general group law via pushbacks and -forwards like before. Then we define the addition the same way.
\begin{prop}
	There exists a natural isomorphism $\sExt^i_R(M,N)\to \Ext^i_R(M,N)$.
\end{prop}
\begin{prf}
	(Sketch) Take a projective resolution $P_\bullet\to M\to 0$ and for any extension 
	\[0\to N\to E_1\to \cdots\to E_i\to M\to 0\]
	lift the identity map on $M$ to a map of resolutions (not both projective) of $M$.

	Say that $f:P_i\to N$ is the map you get in this chain map. We claim that $f$ is a cycle
	in $\Hom(P_\bullet,N)$. If $d_i^*$ is the differential on this complex, then since $d_i^*(f)=f\circ d_i$,
	and using the extension since the map $f\circ d_i:P_{i+1}\to P_i\to N$ is the same as $0:P_{i+1}\to 0\to N$,
	we have it.

	But then $\bar f$ is in $\Ext_R^i(M,N)$.
\end{prf}

\subsection{Yoneda Product}
There is a relatively natural pairing
\[\Ext_R^i(L,M)\times\Ext_R^j(M,N)\to \Ext_R^{i+j}(L,N)\]
where we just concatenate the two exact sequences (eliminating $M$ along the way). It is easy to 
check that it is still exact.

\section{April 29, 2019}
Today we will finish up the last bit on $\sExt$. Let $M=N$ and consider $\Ext_R^\ast(M,M)$.
Then the Yoneda product gives the the structure of a graded ring.

Usually when you are working in another context (e.g. group cohomology) where you have another 
product (e.g. the cup product) this ends up being the same as the Yoneda product. There is something 
to prove here.

\begin{ex}
	Recall $\Ext_\bbZ^1(\bbZ/p,\bbZ/p)\cong\bbZ/p$. Then consider the SES
	\[0\to \bbZ/p\hookrightarrow\bbZ/p^2\to \bbZ/p\]
	where the embedding on the first map takes (written multiplicatively) $a\mapsto ap$.
	Then there are $p-1$ different choices we have for the surjection where $a\mapsto ia$ for $1\le a\le p-1$.

	The one remaining expension comes from the split exact sequence.
\end{ex}
\begin{rmk}
	These are the same embedings we say in $\bbZ/p^\infty$!
\end{rmk}

\subsection{The K\"unneth formula and the Universal Coefficient Theorem}
The motivation here is that, given a space $X$, it is most natural to consider $H_\ast(X,\bbZ)$, 
but if there is some kind of an action of $M$ on these cycles, one may want to compute $H_\ast(X,M)$.

\begin{defn}
	Let $P_\bullet$ and $Q_\bullet$ be complexes. Then $P_\bullet\otimes Q_\bullet$ be defined by 
	\[(P_\bullet\otimes_RQ_\bullet)_n=\bigoplus_{i+j=n}P_i\otimes_R Q_j\]
	and the differential 
	\[d_n=\sum_{i+j=n}d_{ij}:(P\otimes Q)_n\to (P\otimes Q)_{n-1}\]
	is such that if $p$ is in degree $i$ and $q$ is in degree $j$, then 
	\[d_n(p\otimes q)=d_i(p)\otimes q+(-1)^ip\otimes d_j(q).\]
\end{defn}

A natural question to ask: what is the homology of this complex?
\begin{rmk}
	We will be using the fact that, to compute $\Tor$, it suffices to take flat (rather than projective)
	resolutions.

	There is a more general principle here that one can use modules that are ``acyclic'' (higher derived functors are zero)
	for a functor and these resolutions give us homology.
\end{rmk}

\begin{thm}
	Let $P$ and $Q$ be complexes of (right- and left-) $R$ modules such that the $P_n$ and 
	\[B_n^P=d_{n+1}(P_{n+1})\]
	are flat.

	Then there exists a short exact sequence
	\[0\to \bigoplus_{i+j=n}H_i(P)\otimes H_j(Q)\to H_n(P\otimes Q)\to \bigoplus_{i+j=n}\Tor_1^R(H_i(P),H_j(Q))\to 0\]
\end{thm}
\begin{prf}
	We will not use boundaries and cycles for $Q$, so throughout $Z$ and $B$ refer to those for $P$. 
	Now since we have the short exact sequence
	\[-\to Z_i\to P-i\to B_{i-1}to 0\]
	and the long exact sequence of $\Tor$ gives us that $Z_i$ are flat as well. Furthermore (again by the long exact sequence)
	we get 
	\[0\to Z_i\otimes Q_j\to P_i\otimes Q_j\to B_{i-1}\otimes Q_j\to 0\]
	is short exact.

	Now consider
	\[0\to\bigoplus_{i+j=n}Z_i\otimes Q_j\to \bigoplus_{i+j=n} P_i\otimes Q_j\to\bigoplus_{i+j=n-1}B_i\otimes Q_j\to 0\]
	which gives us a SES of complexes
	\[-\to Z_\bullet\otimes Q_\bullet \to P_\bullet\otimes Q_\bullet \to B_\bullet\otimes Q_\bullet[-1]\to 0\]
	and by taking the long exact sequence in homology (note the index shift because of the shift on $Q$):
	\[\cdots\to H_n(B\otimes Q)\to H_n(Z\otimes Q)\to H_n(P\otimes Q)\to H_{n-1}(B\otimes Q)\to H_{n-1}(Z\otimes Q)\to\cdots\]
	Let $\partial:H_n(B\otimes Q)\to H_n(Z\otimes Q)$ be the map above. Then we get the SES
	\[0\to \coker\partial\to H_n(P\otimes Q)\to \ker \delta\to 0\]
	where $\coker d_{ij}$ is $H_i(P)\otimes H_j(Q)$ (this needs some work -- see the book!)

	Finally $ker\partial$ can be computed via using the (flat!) resolution $0\to B_i\to Z_i\to H_i(P)$ and computing $\Tor$.
\end{prf}
\begin{rmk}
	Notice that it is \textit{not true} that if $A$ and $B$ are flat that $A/B$ are flat! 
\end{rmk}

\section{May 1, 2019}
\begin{cor}[Universal Coefficient Theorem -- Corollary of K\"unneth Formula]
	Let $P_\bullet$ be a chain complex of flat modules and that $d_n(P)$ are flat, and that $M$ is an $R$ module.
	Then we have a short exact sequence
	\[0\to H_n(P_\bullet)\otimes M\to H_n(P_\bullet\otimes M)\to \Tor_1^R(H_{n-1}(P_\bullet),M)\to 0\]
	and in particular if $M$ is flat, then $H_n(P_\bullet)\otimes M\cong H_n(P_\bullet\otimes M)$ 
\end{cor}
\begin{prop}[UCT for $\bbZ$]
	Let $P_\bullet$ be a complex of free abelian groups. Then there exists a (non-canonical) isomorphism
	\[H_n(P\otimes M)\cong H_n(P_\bullet)\otimes M\bigoplus \Tor_1^R(H_{n-1}(P_\bullet),M)\]
	where the statement that this isomorphism is non-canonical amounts to saying that this splitting
	in the sequence from the UCT doesn't behave well with respect to maps.
\end{prop}
\begin{rmk}
	What does this come from? Well in topology, if $X$ is a topological space, then ``homology with coefficients''
	is just $H_n(X,M)=H_n(S(X)\otimes M)$. Then the universal coefficient theorem gives us a way to compute this.
\end{rmk}

\begin{prop}[UCT for $H^n$]
	Let $P_\bullet$ be a chain complex of projective $R$-modules where the $d_n(P_\bullet)$ are projective and
	$M$ is any $R$-module. Then we can apply $\Hom(-,M)$ to $P_\bullet$ to compute cohomology, yielding the short exact sequence
	\[0\to \Ext_R^1(H_{n-1}(P_\bullet),M)\to H^n(\Hom(P_\bullet,M))\to \Hom(H_n(P_\bullet,M))\to 0\]
\end{prop}
\begin{prop}[K\"unneth for $H^n$]
	Let $P_\bullet$ be a chain complex of projectives with $d_n(P_\bullet)$ projective. Let $Q_\bullet$ be a cochain complex. 
	Then we have the short exact sequence
	\[0\to\prod_{i+j=n-1}\Ext_R^1(H_i(P),H^j(Q))\to H^n(\Hom(P,Q))\to \prod_{i+j=n}\Hom(H_i(P),H^j(Q))\]
\end{prop}

\subsection{A reminder on dimension theory}
\begin{defn}
	Let $M$ be a (left) $R$-module. Then
	\[\projdim_RM:=\inf_n\{0\to P_n\to\cdots\to P_0\to M\}\]
	Notice that $\projdim$ can be infinite.
\end{defn}
\begin{rmk}
	A module has zero projective dimension if and only if it is projective.
\end{rmk}
\begin{rmk}
	One can define (analogously) injective and flat dimension.
\end{rmk}
\begin{lem}
	Let $M$ be an $R$-module. The following are equivalent:
	\begin{enumerate}
		\item $\projdim_RM\le d$
		\item $\Ext_R^i(M,N)=0$ for all $i\ge d+1$, for all $N$.
		\item $\Ext_R^{d+1}(M,N)=0$ for all $N$.
		\item If $0\to\Omega^dM\to P_{d-1}\to P_{d-2}\to\cdots\to P_0\to M$ is a resolution, then $\Omega^d(M)$ is projective.
	\end{enumerate}
\end{lem}
\begin{rmk}
	Everything can be dualized (e.g. use injective resolutions and injective dimension and $\Omega^{-d}$) 
	and get a similar set of equivalences.
\end{rmk}
\begin{rmk}
	Notice here that the syzygies are not uniquely defined, but they are uniquely defined \textit{up to projective summands.}
	In fancier parlance, they are uniquely defined in the stable module category.
\end{rmk}
\begin{rmk}
	Notice that $\Omega(\Omega^{d-1}(M))=\Omega^d(M)$. Using this, we can compute
	\[\Ext_R^{d+1}(M,N)\cong \Ext_R^1(\Omega^dM,N)\cong\underline{\Hom}_R(\Omega^{d+1},N)\]
	where $\underline{\Hom}$ is $\Hom$ is the stable module category.
\end{rmk}
\begin{rmk}
	Sometimes the \textbf{syzygy} is called \textbf{Heller shift} in representation theory or just \textbf{shift}
	when using a categorical viewpoint.
\end{rmk}

\begin{thm}
	The following numbers are the same:
	\begin{enumerate}
		\item $\sup_{M\in\Rmod}\projdim_R M$
		\item $\sup_{M\in\Rmod}\injdim_R M$
		\item $\sup_{I\text{ a left ideal}}\projdim_RR/I$
		\item $\sup_{M,N\in\Rmod}\{d|\Ext_R^d(M,N)\ne 0\}$
	\end{enumerate}
\end{thm}
\begin{defn}
	The above number is called the \textbf{left global dimension of $R$.}
\end{defn}
\begin{rmk}
	This is where commutative algebra diverges: we have notions of \textit{right}
 global dimension (just change to $\mathbf{mod}$-$R$) as well as global dimension proper.
\end{rmk}

We also have a notion of \textit{weak global dimension} or $\Tor$ global dimension:
\begin{defn}
	The $\Tor$ dimension of $R$ is any of the following numbers:
	\begin{enumerate}
		\item $\sup_{M\in\Rmod}\operatorname{flatdim}_RM$
		\item $\sup_{M\in\modR}\operatorname{flatdim}_RM$
		\item $\sup_{I\text{ left ideal}}\operatorname{flatdim}_R R/I$
		\item $\sup_{J\text{ right ideal}}\operatorname{flatdim}_R R/J$
		\item $\sup_{M\in\modR,N\in\Rmod}\{d|\Tor_d^R(M,N)\ne 0\}$
	\end{enumerate}
\end{defn}
\begin{rmk}
	If $R$ is (left) Noetherian, then $\operatorname{flatdim}_R(M)=\projdim_R(M)$ for all finitely
	generated left $R$-modules. Furthermore the \textbf{left} $\gldim R=\operatorname{Tordim}R$.
\end{rmk}

\begin{rmk}
	If $R$ is (right) Noetherian, you get all the same stuff but flipped.
\end{rmk}
\begin{rmk}
	Rings of global dimension zero are exactly the semisimple ones!
\end{rmk}

\section{May 3, 2019}
Recall that semisimple rings are precisely those of global dimension zero. 
\subsection{Frobenius Algebras}
\begin{defn}
	Let $R$ be a finite dimensional algebra over $k$. Then $R$ is \textbf{Frobenius} if there exists a non-degenerate,
	associative, bilinear form $\sigma:R\times R\to k$.
\end{defn}
\begin{defn}
	If $R$ is Frobenius and furthermore $\sigma$ is a symmetric form, then $R$ is \textbf{symmetric}.
\end{defn}
\begin{prop}
	If $R$ is Frobenius, then
	\[R^\ast=\Hom_k(R,k)\cong R\]
	as (left) $R$-modules.
\end{prop}
\begin{prf}
	Consider the map $R\to R^\ast$  where we send $a\mapsto \varphi_a=\sigma(-,a)$. One can check that this
	actually gives you an isomorphism.
\end{prf}
\begin{cor}
	$R$ is injective as an $R$-module.
\end{cor}
\begin{rmk}
	The idea here is that $R^*$ is projective, and dualizing takes projectives to injectives.
\end{rmk}
\begin{cor}
	When $R$ is Frobenius, all projectives are injective (and vice versa).
\end{cor}
\begin{rmk}
	$R$ is Frobenius implies that $R$ is ``self-injective'', or that $\injdim_RR=0$. Note that this 
	then implies that $R$ is \textit{Gorenstein!}

	It often happens in representation theory that one proves something for finite dimensional algebras, 
	then Frobenius algebras, then sees that it generalizes to Gorenstein ones.
\end{rmk}
\begin{defn}
	$R$ is called \textbf{quasi-Frobenius} (according to Wikipedia) if the collection of injectives and 
	projectives are the same.
\end{defn}
\begin{ex}
	(This is homework) If $G$ is a finite group then $kG$ is symmetric. Note that it is uninteresting when 
	we have Artin-Wedderburn, so the fun stuff happens in the modular case.
\end{ex}
\begin{ex}
	Consider the Lie algebra $\frakg$ over $k$, a field of characteristic $p$. Then $[p]:\frakg\to \frakg$
	is a $p^{th}$ power map or \textbf{restriction operation} if
	\begin{enumerate}
		\item $\ad x^{[p]}=(\ad(x))^p$
		\item $(\lambda x)^{[p]}=\lambda^{[p]}x^{[p]}$
		\item $(x+y)^{[p]}=x^{[p]}+y^{[p]}+\sum_1^{p-1}s_i(x,y)$
	\end{enumerate}
	where $is_i(x,y)$ is the coefficient of $\lambda^{i-1}$ in 
	\[(\ad(\lambda x+y))^{p-1}(x)=[\lambda x+y,[\lambda x+y,[\cdots,[\lambda x + y,x]\cdots,x],x],x]\]
\end{ex}
\begin{rmk}
	If $\frakg$ is linear (sits in $\gl_n$), then we have the $p^{th}$ power map that takes every matrix 
	to its $p^{th}$ power. This is what we are trying to capture without having to rely on what is going to be 
	an inherently coordinate-dependent definition.
\end{rmk}

\begin{defn}
	The universal enveloping algebra $\mathfrak{U}(\frakg)$ is the universal object representing the smallest
	associative algebra capturing the relations of $\frakg$. Look to my Lie algebra notes for more details.
\end{defn}
\begin{defn}
	The \textbf{restricted enveloping algebra} $\mathfrak{u}(\frakg)$ is 
	\[\mathfrak{u}(\frakg)=\mathfrak{U}(\frakg)/\langle x^p-x^{[p]}\rangle\]
\end{defn}
\begin{prop}
	$\mathfrak{u}(\frakg)$ has a ``PBW'' style basis. The monomials take a basis for $\frakg$ and 
	then the monomials look like 
	\[x_1^{i_1}\cdots x_n^{i_n}\]
	for $0\le i_1,\dots\le p-1$.
\end{prop}
In general, the category of representations of $\frakg$ is equivalent to $\mathfrak{U}(\frakg)$-$\mathbf{mod}$
and the restricted representations of $\frakg$ are equivalent to $\mathfrak{u}(\frakg)$ modules.
\begin{prop}
	$\mathfrak{u}(\frakg)$ is Frobenius.
\end{prop}
\subsection{Hereditary Rings}
\begin{defn}
	A ring $R$ is \textbf{(left) hereditary} of every submodule of a free module is projective.
\end{defn}
\begin{rmk}
	Equivalently, the (left) $\gldim R\le 1$.
\end{rmk}
\begin{ex}
	PIDs are hereditary, of course. Furthermore, a commutative integral domain $R$ is hereditary if and only if $R$
	is a Dedekind domain.
\end{ex}

\begin{prop}
	If $f:R\to S$ is a ring homomorphism and $M$ is an $S$-module, then 
	\[\projdim_R(M)\le\projdim_R(S)+\projdim_S(M).\]
\end{prop}
\begin{prf}
	The proof of this statement (which we may have seen with S\'andor?) uses Cartan-Eilenberg resolutions.
	Let $\cal A$ be an abelian category with enough projectives and let $C$ be a chain complex in $\cal A$.
	Then $P=P_{\bullet,\bullet}$ (a double complex) is a \textbf{Cartan-Eilenberg Resolution of $C$} if we have 
	a commutative diagram
	\begin{center}
		\begin{tikzcd}
			& \vdots\ar[d] & \vdots\ar[d] & \\
			& P_{n+1,1}\ar[d] & P_{n,1}\ar[d] & \\
			& P_{n+1,0}\ar[d] & P_{n,0}\ar[d] & \\
			\cdots\ar[r] & C_{n+1}\ar[r] & C_n\ar[r] & \cdots
		\end{tikzcd}
	\end{center}
	where we have the properties: 
	\begin{itemize}
		\item $P_{n\bullet}\to C_n$ is a projective resolution.
		\item $P_{nj}=0$ for $j<0$
		\item $B(P_{n\bullet})\to B(C_\bullet)$ and $H(P_n)\to H_n(C_\bullet)$ are projective resolutions.
	\end{itemize}
\end{prf}

\section{May 8, 2019}
\begin{lem}
	Cartan-Eilenberg resolutions exist for any complex $C$.
\end{lem}
\begin{prop}
	Let $f:R\to S$ be ar ring homomorphism and let $M$ be an $S$-module and let $f^\ast(M)$ be the $R$-module
	taken by acting on $M$ via $f$. Then 
	\[\projdim_S(f^\ast(M))=\projdim_S(M)+\projdim_R(S)\]
\end{prop}
\begin{prf}
	\textit{Sketch.} Let $n=\projdim_S(M)$ and let $P_\bullet\to M$ be a projective resolution of $M$ (as an $S$-module)
	of length $n$. Now we use the fact (exercise) that $\projdim_R(P_i)\le \projdim_R(S)$.

	Let $Q_{\bullet\bullet}\to P_\bullet$ be a Cartan-Eilenberg resolution of $P_\bullet$ as $R$-modules. Now we can ``chop off''
	higher terms of the CE complex, replacing the $m^{th}$ row with $\Omega^m(Q_{n\bullet})$ and the higher rows with zeros. One can 
	show without too much work that this is still a CE resolution of $P_\bullet$.

	Now we have already that $P_\bullet\sim M$ is a quasi isomorphism and the total complex $\operatorname{Tot}(Q_{\bullet\bullet})$ 
	is quasi isomorphic to $P_\bullet$, so we have, in effect, created a projective resolution of $M$ using $R$-modules of length $n+m$.
\end{prf}
\subsection{Koszul Complex/Resolution}
As motivation, let $R$ be a regular local ring of dimension $n$. Then the Koszul resolution is a resolution of the 
residue field $k=R/\mathfrak{m}$ of length $n$.

\begin{ex}
	Let $x$ be a central element in $R$. Consider the map $R\xrightarrow{x}R$. This is a complex! We can compute 
	$H_1={_x}R$ and $H_0=R/xR$. In particular, if $x$ is not a zero divisor, then $K_\bullet(x)=R\to R$ is a 
	projective resolution of $R/xR$.

	This gives us the Koszul complex on a single element.
\end{ex}

\begin{defn}[Koszul Complex]
	Let $\underline x=(x_1,\dots,x_n)$ be a sequence (not yet regular) of central elements. Then 
	\[K_\bullet(\underline{x})=K_\bullet(x_1)\otimes\cdots\otimes K_\bullet(x_n)\]
	is the Koszul complex of $\underline{x}$.
\end{defn}
\begin{ex}
	Let's consider what happens when $n=2$. Then the length of this complex is going to be 2. This comes 
	from the diagonals of the double complex:
	\begin{center}
		\begin{tikzcd}
			R\ar[r,"x_1"]\ar[d,"-x_2",swap] & R\ar[d,"x_2"]\\
			R\ar[r,"x_1"] & R
		\end{tikzcd}
	\end{center}
	giving us the complex 
\[R\xrightarrow{\begin{pmatrix}-x_2 & x_1\end{pmatrix}} R\oplus R\xrightarrow{\begin{pmatrix}x_1\\x_2\end{pmatrix}} R\]
\end{ex}

Now in general if $K_p(\underline{x})$ is the $p^{th}$ component of the Koszul complex, we get $\dim_R(K_p(\underline{x})))=\binom{n}{p}$.
We denote the generators for $K_p(\underline{x})$ as $e_{i_1}\wedge\cdots\wedge e_{i_p}$ for $1\le i_1\cdots i_p\le n$.

Then we claim that the differential on this complex is 
\[d_p(e_{i_1}\wedge \cdots\wedge e_{i_p})=\sum (-1)^{k+1}x_k e_{i_1}\cdots\wedge \widehat{e_{i_k}}\wedge\cdots\wedge e_{i_p}\]

\begin{prop}
	$K_\ast(\underline{x})$ is a differential graded algebra with product formed by concatenation modulo
	the relation $e_i\wedge e_j=-e_j\wedge e_i$ and $e_i\wedge e_j=0$. If $V=Rx_1\oplus \cdots\oplus Rx_n$, then
	$K_\ast(\underline{x})\cong \Lambda^\ast(V)$ and $K_p(\underline{x})\cong\Lambda^p(V)\cong\Lambda^p(R^n)$.
\end{prop}
\begin{rmk}
	Note that in the proposition above, $K_\ast(\underline{x})\cong \Lambda^\ast(V)$ is meant to imply isomorphism \textit{on the level of graded algebras.}
\end{rmk}
\begin{defn}
	If $M$ is any (left) $M$ module then $K_\bullet(\underline{x})\otimes M$ is a Koszul complex of $M$.
\end{defn}
\begin{rmk}
	Then, as a matter of notation, $H_i(\underline{x},M)=H_i(K_\bullet(\underline{x}\otimes M))$.
\end{rmk}

\section{May 10. 2019}
\subsection{Koszul Complexes Continued}
Let $R$ be a commutative local ring $(R,\frakm,k)$. 
\begin{thm}
	Let $(x_1,\dots,x_n)\in \frakm$ is an $M$-regular sequence for $M$, then
	\begin{itemize}
		\item $H_0(\underline{x},M)=M/((x_1,\dots,x_n)M)$.
		\item $H_i(\underline{x},M)=0$ when $i>0$.
	\end{itemize}
\end{thm}
\begin{rmk}
	Equivalently, $K_\bullet(\underline{x})\otimes M$ is a projective resolution of $M/(x_1,\dots,x_n)M$.
\end{rmk}
In particular, when $M=R$, and when $\dim R=n$, then the (length $n$) sequence is just a regular sequence. Then $K_\bullet(\underline{x})\to k$ is a free resolution of $k=R/\frakm$.
\begin{cor}
	If $R$ is a regular ring, $\projdim_Rk=n=\dim R$
\end{cor}
\begin{lem}
	Let $x$ be a non-zero-divisor in the center of $R$ and let $C\in\Ch(\Rmod)$. Then there 
	exists a natural short exact sequence
	\[0\to H_i(C)/xH_i(C)\to H_i(K_\bullet(x)\otimes C)\to {_x}H_i(C)\to 0\]
\end{lem}
\begin{prf}
	K\"unneth formula!
\end{prf}
\begin{thm}
	Let $R$ be a commutative local ring, $M$ an $R$-module, and a sequence $(x_1,\dots,x_n)\in\frakm$. Then 
	\begin{itemize}
		\item $(x_1,\dots,x_n)$ is $M$-regular.
		\item $H_i(\underline{x},M)=0$ for $i>0$.
		\item $H_1(\underline{x},M)=0$
	\end{itemize}
\end{thm}
\begin{cor}
	If $(x_1,\dots,x_n)$ is a regular sequence in a local ring, then any permutation of it is also regular!
\end{cor}
\begin{rmk}
	Definitely not true when $R$ is non-local. We did an example with S\'andor.
\end{rmk}

\subsection{Group (co)homology}
Let $G$ be a group. Then we can consider $\bbZ G$ modules and compute 
\[\Ext_G^i(M,N)=R^i\Hom_G(M,N).\]

Recall that we can consider representations of $G$ over $R$ (a commutative ring). Then if $M$ and $N$ 
are two representations of $G$ ($RG$-modules), then $M\otimes_R N$ can be defined to have the diagonal action of $G$.
Similarly $\Hom_R(M,N)$ becomes a representation via $(g\cdot f)(m)=g\cdot(f(g^{-1}m))$.
\begin{rmk}
	Notice here that $\Hom_R(-,-)$ is \textit{internal hom} since it becomes again a $G$-module.
\end{rmk}
\begin{prop}
	$\Hom_R(M,N)^G\cong \Hom_{RG}(M,N)$ where the later is the external (categorical) $\Hom$.
\end{prop}

The upshot of all of this is that we can compute the group cohomology:
\[H^i(G,M):=\Ext_{RG}(R,M)=R^i\Hom_{RG}(R,M)=R^i\Hom_R(Z,M)^G=R^iM^G\]
so we can define the group cohomology functor 
\[H^i(G,-)=R^i(-)^G.\]

Now $H_i(G,M)=\Tor_i^{RG}(R,M)$. But $R\otimes_{RG}M$ is going to be isomorphic to $M_G$, the \textbf{coinvariance} of $M$,
which is $M/\langle m-gm|m\in M,g\in G\rangle$. So then $H_i(G,-)=L_i(-)_G$.

\begin{rmk}
	How do we compute this in general? This is not actually known. For instance, consider $GL_n(\bbF_p)$ is a relatively simple group (finite!)
	and we don't know the minimal $i>0$ such that $H^i(GL_n(\bbF_p),\bbF_p)\ne 0$.
\end{rmk}

Na\"ively, we can use that $H^i(G,M)=\Ext_G^i(\bbZ,M)$ (here setting $R=\bbZ$) and then compute this via a resolution 
either of $P_\bullet\to\bbZ$ and compute $H^i(\Hom(P_\bullet,M))$ or $H^i((I^\bullet)^G)$.

\section{May 13, 2019}
\subsection{Sample Computations}
\begin{ex}
	Let $G=\bbZ/p=C_p$ (not necessarily a prime yet). For instance, let $p=0$ first and $G=\bbZ$.
	Then to compute $H^i(C_p,A)=\Ext_{C_p}^i(\bbZ,A)$ is computed either with a resolution of $\bbZ$ or $A$. Since the 
	latter is allowed to be general, we need to do the former. Consider 
	\[\cdots\to\bbZ C_p\xrightarrow{N}\bbZ C_p\xrightarrow{\sigma-1}\bbZ C_p\xrightarrow{\varepsilon}\bbZ\to 0\]
	where $N=(\sigma^p-1)/(\sigma-1)$.

	Let's begin by noticing that $\Hom_{\bbZ C_p}(\bbZ C_p,A)\cong A$, so we get a resolution 
	\[\cdots\leftarrow A\xleftarrow{\sigma-1}A\xleftarrow{N}A\xleftarrow{\sigma-1}A\]
	Doing our computations, we get $H^0(C_p,A)=\ker\{\sigma-1:A\to A\}=A^{C_p}$. As we expected!

	Then $H^1(C_p,A)=\ker(N)/(\sigma-1)A$ and $H^2(C_p,A)=A^{C_p}/NA$, and this just cycles.
\end{ex}
\begin{rmk}
	The above is an example of a \textbf{periodic resolution of $\bbZ$}. We will not talk much about this, but there is a measure of 
	\textit{complexity of a module} which refers to the length of the period in such a resolution.
\end{rmk}

\begin{ex}
	In a specific case, consider $A=\bbZ$ (the trivial $\bbZ C_p$ module). Then $H^0$ is $\bbZ$, $H^1=0$ and $H^2=\bbZ/p$.
\end{ex}
\begin{ex}
	Now let's let $k$ be a field of characteristic $p$ (now $p$ is a prime, obviously) and let $A$ be a $k$-module. 
	Then the same thing goes through in computing $H^i(C_p,A)=\Ext_{kC_p}(k,A)$, but we just replace $\bbZ$ with $k$.

	If instead we compute $H^i(C_p,k)$, we get $k$ every time (since $\sigma-1$ is the zero map as well as $1+\sigma+\cdots+\sigma^{p-1}$).
\end{ex}
\begin{ex}
	Let $\langle t\rangle=C$ be a cyclic group and consider the group algebra $\bbZ C\cong \bbZ[t,t^{-1}]$.
	But then a resolution of $\bbZ$ is 
	\[0\to \bbZ\langle t\rangle\xrightarrow{t-1}\bbZ\langle t\rangle\xrightarrow{\varepsilon}\bbZ\]
	And so you get a result about Poincar\'e duality that I missed but don't have time to figure out.
\end{ex}
\begin{ex}
	Now if $A=\bbZ\langle t\rangle$ and $C=\langle t\rangle$. THen $H^0(C,\bbZ C)=0$ and $H^1(C,\bbZ C)=\bbZ$.
	Replacing $\bbZ$ with $k$, we get $H^0(C,\bbZ C)=0$ and $H^1(C,\bbZ C)=k$.
\end{ex}
\begin{rmk}
	This gives us one of the examples of the differences between finite and infinite groups: if $C$ were finite,
	$kC$ would be Frobenius, whence self-injective, whence $H^1(C,kC)=0$.
\end{rmk}

\subsection{Products in Cohomology}
We will begin by discussing the idea of a ``coss-product.'' Let $G$ and $H$ be groups. Then we want to define 
a map 
\[H^p(G,\bbZ)\otimes H^q(H,\bbZ)\xrightarrow{\times} H^{p+q}(G\times H,\bbZ).\]
Eventually we can let $G=H$ and using the coproduct (diagonal) map $\Delta(g)=(g,g)$, whose image under $H^{p+q}$ is 
$\Delta^\ast:H^{p+q}(G\times G,\bbZ)=H^{p+q}(G,\bbZ)$ to get our cup product.

\begin{prop}
	Let $P_\bullet\to\bbZ$ be a projective resolution as $\bbZ G$ modules and $Q_\bullet\to\bbZ$ a resolution as $\bbZ H$ modules.
	Then $P_\bullet\otimes Q_\bullet\to\bbZ$ is a projective resolution as $\bbZ G\otimes \bbZ H=\bbZ(G\times H)$ modules.
\end{prop}

Let's begin by defining a map
\[f:\Hom_G(P_\bullet,\bbZ)\otimes \Hom_H(Q,\bbZ)\to \Hom_{G\times H}(P\otimes Q,\bbZ)\]
via
\[f(\mu\otimes \nu)(x\otimes y)=\mu(x)\mu(y)\]
for all $x\in P$ and $y\in Q$. Notice here that if $|\mu|\ne |x|$, then we set $\mu(x)=0$.

Now using that $H^i(G,\bbZ)=H^i(\Hom_G(P_\bullet,\bbZ)),$ and using the first map from the K\"unneth formula,
\[H^i(\Hom_G(P_\bullet))\otimes H^j(\Hom_H(Q_\bullet,\bbZ))\to H^{i+j}(\Hom_G(P_\bullet,\bbZ)\otimes \Hom_H(Q_\bullet,\bbZ))\]
and then it can be checked that this map behaves well with respect to cycles and boundaries, so we 
get a map 
\[H^{i+j}(\Hom_G(P_\bullet,\bbZ)\otimes \Hom_H(Q_\bullet,\bbZ))\to H^{i+j}(\Hom_{G\times H}(P\otimes Q,\bbZ))\cong H^{i+j}(G\times H,\bbZ).\]

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%              Homework                %%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newpage
\section{HW1 -- Due April 17}
Throughout the assignment, $R$ is an associative ring with 1.
\begin{prob}
	Show that $\Ch(R)$ is an abelian category.
\end{prob}
\begin{sol}
	We begin by proving a lemma:
	\begin{lem}\label{lem-chain-additive} 
		$\Ch=\Ch(R)$ is additive. 
	\end{lem}
	\begin{prf}
		Let $C,D\in\Ch$ and $f\in\Hom_{\Ch}(C,D)$. Then 
		$f=(f_i)_{i\in\bbZ}$ is a collection of maps of $R$-modules. Then, leveraging that $\Rmod$ is Abelian
		(\textit{the} Abelian category in some ways) and the fact that addition of chain maps is perfomed 
		component-wise, this gives $\Hom_{\Ch}(C,D)$ the structure of an Abelian group as the direct product
		of Abelian groups. That the component-wise sum is actually a chain map is evident since 
		\[d((f+g)(x))=d(f(x)+g(x))=d(f(x))+d(g(x))=f(d(x))+g(d(x))=(f+g)(d(x)).\]

		The chain $0_\bullet=\cdots 0 \to 0\to 0\to \cdots$ is the zero object in $\Ch$ since there is a 
		unique map $\mathbf{0}=(0)_{i\in\bbZ}$ both from and to $0_\bullet$ from any chain (again, since maps
		are component-wise $R$-module maps, so each component must be $0$). This map is indeed a chain map
		since $0=0\circ d=d'\circ 0$. Therefore $\Ch$ is additive.
	\end{prf}

	\brk

	Now for the final three properties (that (co)kernels exist, that every monic is the kernel of its cokernel,
	and that every epi is the cokernel of its kernel) follow from the fact that $\Rmod$ is abelian (and thus satisfy
	these properties) and that (co)kernels in $\Ch$ are ``what we would hope they would be'' component-wise. 
	Therefore the properties pass from $\Rmod$ up to $\Ch$.

	\begin{lem}\label{lem-chain-ker}
		Let $C,D\in\Ch$ and $f\in\Hom_{\Ch}(C,D)$. Then there is a complex $K$ and a map $i:K\to C$ 
		that is the kernel of $f$ and furthermore corresponds to the ($\Rmod$) kernel 
		\[i_n:K_n=\ker f_n\to C_n\]
		on the components.
	\end{lem}
	\begin{prf}
		\begin{rmk}
		I have just noticed that I am basically re-proving part of snake lemma below for the existence of the maps between
		kernels. Oh well, good practice.
		\end{rmk}
		Consider the following diagram:

	\begin{center}
		\begin{tikzcd}
			\ker f_n \ar[r,"\partial_n"]\ar[d,hook,"i_n"] & \ker f_{n+1}\ar[d,hook,"i_{n+1}"]\\
			C_n\ar[r,"d_n"]\ar[d,"f_n"] & C_{n+1}\ar[d,"f_{n+1}"]\\
			D_n\ar[r,"d_n'"] & D_{n+1}
		\end{tikzcd}
	\end{center}

	Let $f\in\Hom_{\Ch}(C,D)$ as above. Define $K=(\ker f_i,\partial)\in\Ch$ where 
	$\partial^i:\ker f_i\to \ker f_{i+1}$ is defined as follows: let $x\in\ker f_n$. Then since $f$
	is a chain map, 
	\[0=d_n'(0)=d_n'(f_n(x))=f_{n+1}(d_n(x))\]
	so in particular $d_n(x)\in\ker f_{n+1}$. So $d_n(\ker f_n)\subseteq \ker f_{n+1}$, so we can lift
	$d_n\circ i_n$ through $i_{n+1}$ to a map $\partial_n:\ker f_n\to \ker f_{n+1}$. Then $K$ is a chain 
	and $i$ is a chain map by construction.

	That this is the kernel of $f$ in $\Ch$ follows since if $g\in\Hom(A,C)$ and $f\circ g=0$, each component
	map $g_n$ must factor through $\ker f_n$. It remains to see that this map is a chain map. Consider the diagram

	\begin{center}
		\begin{tikzcd}
			& A_n\ar[rr,"d_A"]\ar[ld,"\tilde g_n",swap]\ar[ldd,"g_n"] & & A_{n+1}\ar[ld,"\tilde g_{n+1}",swap]\ar[ldd,"g_{n+1}"]\\
			\ker f_n\ar[rr,"d_K"]\ar[d,hook,"i_n",swap] & & \ker f_{n+1}\ar[d,hook,"i_{n+1}",swap] &\\
			C_n\ar[rr,"d_C"] & & C_{n+1} &
		\end{tikzcd}
	\end{center}

	\noindent where $\tilde g_n$ is the canonical map through the kernel of $f_n$. Pick any $a\in A_n$ and notice
	\begin{align*}
		(i_{n+1}\circ d_K)\circ \tilde g_n(a)&=(d_C\circ i_n)\circ \tilde g_n(a) \\
		&= d_C\circ g_n(a)\\
		&=g_{n+1}\circ d_A(a)\\
		&=i_{n+1}\circ \tilde g_{n+1}\circ d_A(a)
	\end{align*}
	and so since $i_{n+1}$ is mono, this establishes
	\[d_K\circ\tilde g_n = \tilde g_{n+1}\circ d_A\]
	so $\tilde g$ is a chain map and therefore $i:K\to C$ is the kernel of $f$, which lies in $\Ch$.
	\end{prf}

	\brk
	\begin{lem}\label{lem-chain-coker}
		Let $C,D\in\Ch$ and $f\in\Hom_{\Ch}(C,D)$. Then there is a complex $E$ and a map $q:D\to E$ 
		that is the cokernel of $f$ and furthermore corresponds to the ($\Rmod$) cokernel 
		\[q_n:D_n\to E_n=\coker f_n\]
		on the components.
	\end{lem}
	\begin{prf}
	Consider

	\begin{center}
		\begin{tikzcd}
			C_n\ar[r,"d_n"]\ar[d,"f_n"] & C_{n+1}\ar[d,"f_{n+1}"]\\
			D_n\ar[r,"d_n'"]\ar[d,two heads,"q_n"] & D_{n+1}\ar[d, two heads,"q_{n+1}"]\\
			\coker f_n\ar[r,"\partial_n"] &\coker f_{n+1}
		\end{tikzcd}
	\end{center}

	\begin{rmk}
		This also follows from snake lemma and is less ``elementy''.
	\end{rmk}
	Given $\hat x\in\coker f_n$, let $x$ be any preimage of $\hat x$ under $q_n$ and define
	\[\partial_n(\hat x)=q_{n+1}\circ d_n'(x).\]

	To show that this map is well-defined, let $y$ be any other preimage of $x$ under $q_n$.
	Then since $q_n(x-y)=0$, we can pull $x-y$ back through $f_n$ to some $c\in C_n$. Then
	\begin{align*}
		(q_{n+1}\circ d_n')(x-y)&=q_{n+1}\circ (d_n'\circ f_n)(c)\\
		&=q_{n+1}\circ(f_{n+1}\circ d_n)(c)\\
		&=(q_{n+1}\circ f_{n+1})\circ d_n(c)\\
		&=0\circ d_n(c)=0
	\end{align*}
	so $q_{n+1}\circ d_n'(x)=q_{n_1}\circ d_n'(y)$, so $\partial_n(\hat x)$ is well-defined. Then 
	$E=(\coker f_n)_{n\in\bbZ}$ is a chain complex with differential $\partial$. To show that $q:D\to E$ 
	is the cokernel of $f$, let $A\in\Ch$ and $g\in\Hom_{\Ch}(D,A)$  with $g\circ f = 0$ then from the universal property in 
	$\Rmod$, the component maps must factor through the $E_n$:

	\begin{center}
		\begin{tikzcd}
			& A_n\ar[rr,"d_A"] & & A_{n+1}\\
			\coker f_n\ar[rr,"d_E"]\ar[ur,"\tilde g_n"]& & \coker f_{n+1}\ar[ur,"\tilde g_{n+1}"] &\\
			D_n\ar[rr,"d_D"]\ar[uur,"g_n",swap]\ar[u,two heads,"q_n"] & & D_{n+1}\ar[uur,"g_{n+1}",swap]\ar[u,two heads,"q_{n+1}"] &
		\end{tikzcd}
	\end{center}

	Then for any $d\in D_n$, 
	\begin{align*}
		\tilde g_{n+1}\circ d_E\circ q_n(d) &= \tilde g_{n+1}\circ q_{n+1}\circ d_D(d)\\
		&= g_{n+1}\circ d_D(d)\\
		&= d_A\circ g_n(d)\\
		&= d_A\circ \tilde g_n\circ q_n(d)
	\end{align*}
	and since $q_n$ is epi, we conclude that 
	\[\tilde g\circ d_E=d_A\circ \tilde g\]
	so since $E$ is the (unique) object satisfying this property, it is the cokernel of $f$ in $\Ch$.
	\end{prf}
\end{sol}

\begin{prob}
	Show that a chain map $f:C_\bullet\to D_\bullet$ induces a well-defined map $H_n(f)$ on homology.
\end{prob}
\begin{sol}
	Since $f$ is a chain map, it is easy to see it induces a map on cycles: let $c\in Z_n(C)$.
	Then 
	\[d_D(f(c))=f(d_C(c))=f(0)=0\]
	so $f(Z_n(C))\subseteq Z_n(D).$ Thus it suffices to show that if $x$ and $y$ differ by a boundary,
	$f(x)$ and $f(y)$ do as well.

	To see this, assume that $x-y\in B_n(C)$, so there is a $z\in C_{n+1}$ such that $d_C(z)=x-y$.
	But then 
	\[f(x)-f(y)=f(x-y)=f(d_C(z))=d_D(f(z))\in B_n(D)\]
	so the map $H_n(f)$ is well-defined.
\end{sol}

\begin{prob}
	(\textbf{Weibel 1.4.5}) 
	\begin{enumerate}
		\item Show that the chain homotopy equivalence is an equivalence relation on the set of all chain maps from $C$
		to $D$. Let $\Hom_\mathbf{K}(C,D)$ denote the equivalence classes of such maps. Show that $\Hom_\mathbf{K}(C,D)$
		is an Abelian group.
		\item Let $f\sim g$ where $f,g:C\to D$. If $u:B\to C$ and $v:D\to E$ are chain maps, show $vfu$ and $vgu$ are 
		chain homotopic. Deduce that there is a category $\mathbf K$ whose objects are chain complexes and whose morphisms
		are given in (a).
		\item Let $f_0,f_1,g_0,$ and $g_1$ be chain maps $C\to D$ such that $f_i$ is chain homotopic to $g_i$. Show that 
		$f_0+f_1\sim g_0+g_1$. Deduce that $\mathbf K$ is an additive category and that $\Ch\to\mathbf K$ is an 
		additive functor.
		\item (Optional) Is $\mathbf K$ an abelian category? Explain.
	\end{enumerate}
\end{prob}
\begin{sol}
	\textbf{(a):} Let $f,g,h\in\Hom_{\Ch}(C,D).$ It is clear that $f\sim f$ since $f-f=0$ is nullhomotopic via $s=0$. Thus $\sim$ is reflexive.
	If $f\sim g$, then there is a chain homotopy $s$ such that $f-g=ds+sd$. Setting $s'=-s$, we get
	\[ds'+s'd=-(ds+sd)=-(f-g)=g-f\]
	so $g\sim f$, whence $\sim$ is symmetric.

	Finally, assume $f\sim g$ via $s$ and $g\sim h$ via $t$. But then notice that
	\begin{align*}
		f-h &= f-g + g-h\\
		&= ds+sd+dt+td\\
		&= ds+dt+sd+td\\
		&= d(s+t)+(s+t)d
	\end{align*}
	so $f-h\sim 0$ via the chain homotopy $s+t$. Therefore $\sim$ is an equivalence relation.

	The last statement follows since 
	\[\Hom_{\mathbf K}(C,D)=\Hom_{\Ch}(C,D)/\sim\]
	and every quotient of an Abelian group is Abelian.

	\vspace{2ex}\noindent \textbf{(b):} Say $s$ is the homotopy from $f-g$ to $0$. Then notice since everything is linear:
	\begin{align*}
		vfu-vgu &= v(f-g)u\\
		&= v(sd_C+d_Ds)u\\
		&= vs(d_Cu)+(vd_D)su\\
		&= vsud_B+d_Evsu
	\end{align*}
	so $vfu\sim vgu$ via the chain homotopy $vsu$. This means that if $[f]$ is the chain homotopy
	equivalence class of $f$, that the composition of chain maps is well-defined. To see this, we can let
	either $B=C$ and $u=\id_C$ or else $D=E$ and $v=\id_D$.

	The identity map in $\Hom_{\mathbf K}(C,C)$ is $[\id_C]$, containing all nullhomotopic self-maps of $C$.
	Since $\mathbf K$ is a collection of objects and morphisms with an identity map for each object and that is closed 
	under composition, we get that $\mathbf K$ is a category.

	\vspace{2ex}\noindent \textbf{(c):} Let $f_0,f_1,g_0$ and $g_1$ be defined as provided. Then
	\begin{align*}
		f_0+f_1-g_0-g_1 &= (f_0-g_0)+(f_1-g_1)\\
		&= s_0d_C+d_Ds_0 +s_1d_C+d_D s_1\\
		&= (s_0+s_1)d_C+d_D(s_0+s_1)
	\end{align*}
	so $f_0+f_1\sim g_0+ g_1$. Then $[f]+[g]=[f+g]$ for all $f,g\in\Hom_{\Ch}(C,D)$. Now $0_\bullet$
	is the zero object in $\mathbf K$ since there is a unique map in $\Hom_{\Ch}(0,C)$ and $\Hom_{\Ch}(C,0)$,
	so the same is true in $\mathbf K$. Therefore $\mathbf K$ is additive.

	The functor $Q:\Ch\to\mathbf K$ sending complexes to themselves and chain maps $Q(f)=[f]$ (which is
	a functor since composition is respected by (b) and identity maps are respected by definition)
	is additive since for any $A,B\in \Ch$,
	\[\Hom_{\Ch}(A,B)\to \Hom_{\mathbf K}(A,B)\]
	is a group homomorphism as shown above.

	\vspace{2ex}\noindent\textbf{(d):} I suspect the answer to this is no.
\end{sol}

\begin{prob}
	Show that a chain complex $C$ is split exact if and only if it is nullhomotopic.
\end{prob}
\begin{sol}
	Begin by assuming that $C$ is split exact. Then consider the diagram

	\begin{center}
		\begin{tikzcd}
			C_{n-1}^0\oplus C_{n-1}^1\ar[r,"d"]\ar[d,"\id"] & C_n^0\oplus C_n^1\ar[r,"d"]\ar[d,"\id"]\ar[ld,"s"] & C_{n+1}^0\oplus C_{n+1}^1\ar[d,"\id"]\ar[ld,"s"]\\
			C_{n-1}^0\oplus C_{n-1}^1\ar[r,"d"] & C_n^0\oplus C_n^1\ar[r,"d"] & C_{n+1}^0\oplus C_{n+1}^1
		\end{tikzcd}
	\end{center}

	\noindent where $C_n=C_n^0\oplus C_n^1$ and $d_n|_{C_n^0}$ is the zero map and $d_n|_{C_n^1}$ is 
	an isomorphism onto $C_{n+1}^0.$

	Define $s$ to be 
	\[(0,\pi_0)_i:C_i^0\oplus C_i^1\to C_{i-1}^0\oplus C_{i-1}^1\]
	where 
	\[\pi_0:C_i^0\oplus C_i^1\to C_i^0\cong C_{i-1}^1\]
	is projection onto the first coordinate. Then clearly 
	\[f((a,b))=(a,b)=(a,0)+(0,b)=d((0,a))+s((b,0))=d(s((a,b)))+s(d((a,b)))\]
	whence $\id\sim 0$ so $C$ is nullhomotopic.

	\brk

	Now assume there exists a map $\id_C\sim 0$ via chain homotopy $s:C\to C[-1]$. The exactness of 
	$C$ is immediate since homotopic maps induce isomorphisms on homology. Consider the SES

	\begin{center}
		\begin{tikzcd}
			0\ar[r] & \ker d \ar[r,"i"] & C_n\ar[r,swap,"d"] & d(C_n)\ar[r]\ar[l,bend right,"s'",swap] & 0
		\end{tikzcd}
	\end{center}

	\noindent where 
	\[s'=s_{n+1}|_{d(C_n)}\]

	But then consider the restriction of $\id =\id_{C_{n+1}}$, to $d(C_n)\subseteq C_{n+1}$:
	\[\id|_{d(C_n)}=(s_{n+2}\circ d+d\circ s_{n+1})|_{d(C_n)}=(s_{n+2}\circ d)|_{d(C_n)}+(d\circ s_{n+1})|_{d(C_n)}=(d\circ s_{n+1})|_{d(C_n)}\]
	so $d\circ s'$ is the identity on $d(C_n)$ whence $s$ is a splitting of the above short exact sequence.

	Since this holds for an arbitrary $n$, this establishes that $C_n\cong \ker d\oplus d(C_n)$, 
	and thus by exactness $C$ is split exact.
\end{sol}

\begin{prob}
	Show that a chain complex $P_\bullet$ is a projective object in the category of
chain complexes over a ring $R$ if and only if $P_\bullet$ is a split exact complex of projective
modules.
\end{prob}
\begin{sol}
	Assume first that $P$ is a projective object in $\Ch$. Following a hint in Weibel, we consider the surjection
	$\operatorname{cone}(P)\xrightarrow{\delta} P[-1]\to 0$, which admits a splitting map $\sigma:P[-1]\to\operatorname{cone}(P)$
	via the standard argument -- lift $\id_P:P\to P$ to a map $\sigma:P\to \operatorname{cone}(P)$ such that
	$\delta\circ\sigma=\id_P$.

	From Weibel, we know $\delta(a,b)=-a$, so if $\sigma_i=(s,t)_i:P_{i-1}\to P_{i-1}\oplus P_i,$ we get that
	\[\delta\circ(s,t)(p)=-s(p)=p=\id_P(p)\]
	by the property of the lift $\sigma$ and therefore $s=-\id_P$.

	Then using the fact that $\sigma$ is a chain map, we can compute (if $C=\operatorname{cone}(P)$):
	\[
		(s\circ d_P,t\circ d_{P[-1]})=(s,t)\circ d_P=d_C\circ(s,t)=(d_P\circ s, d_P\circ t-\id_P\circ s)
	\]
	and therefore by focusing on the second component:
	\[t\circ d_{P[-1]}-d_P\circ s=-t\circ d_P-d_P\circ s=\id_P\]
	whence $-\id_P\sim\id_P\sim 0$, so by problem three above $P$ is split exact.

	To see every element is injective, we can ``localize'' by considering a lift of $P\to M_\bullet=(\cdots\to0\to M\to 0\to \cdots)$
	through the surjection $(\cdots\to 0\to N\to 0\to\cdots)\twoheadrightarrow M_\bullet$. Since a lift always exists for $M$ and $N$ 
	in any component, this shows all the terms in the complex $P$ are projective.

	\brk

	Now assume that $P$ is a split exact. Then since the $P_i$ are projective, for any chain map $f:P\to M$ and 
	surjection $g:N\to M$, we get maps $\tilde f_i:P_i\to N_i$. Recall that since $P$ is split exact, there 
	exists a splitting map $\sigma_n:P_n\to P_{n+1}$ for every $n$ such that $d_P\circ \sigma_n=\id_{P_n}.$

	Now define $F_n:P_n\to N_n$ by 
	\[F_n=d_N\circ \tilde f_{n+1}\circ \sigma_n+\tilde f_n\circ\sigma_{n-1}\circ d_P\]
	and so we can compute
	\begin{align*}
		d_N\circ F_n&=d_N^2\circ\tilde f_{n+1}\circ \sigma_n + d_N\circ\tilde f_n\circ \sigma_{n-1}\circ d_P\\
		&= d_N\circ\tilde f_n\circ \sigma_{n-1}\circ d_P\\
		&= d_N\circ\tilde f_n\circ \sigma_{n-1}\circ d_P + \tilde f_{n-1}\circ\sigma_{n-2}\circ d_P^2\\
		&= F_{n-1}\circ d_P
	\end{align*}
	so $F$ is a chain map, and therefore $F$ is a lift of $f$, so since $P$ satisfies the usual 
	diagram, it is a projective object.
\end{sol}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newpage
\section{HW2 -- Due XXXXX}
\begin{prob}
	Let $R$ be a commutative ring, and $N$ be an $R$-module. Show that the following are equivalent:
	\begin{enumerate}
		\item $N$ is flat 
		\item $\Tor_i^R(M,N) = 0$ for any $i>0$ and any $R$-module $M$
		\item $\Tor_1^R(M,N) = 0$ for any $R$-module $M$
		\item $\Tor_1^R(R/I, N) = 0$ for any ideal $I \subset R$.
	\end{enumerate}
\end{prob}
\begin{sol}
	(a)$\Rightarrow$(b)$\Rightarrow$(c) is clear since $\Tor_i^R(-,N)$ is a left derived functor of $-\otimes N$. But when $N$ is flat, $-\otimes N$ is exact.

	That (c)$\Rightarrow$(a) can be seen by using the long exact sequence of $\Tor$ arising from any short exact sequence $0\to A\to B\to C\to 0$:
	\[\cdots\to\Tor_2^R(C,N)\to 0\to 0\to 0\to A\otimes N\to B\otimes N\to C\otimes N\to 0\]
	but then in particular $0\to A\otimes N\to B\otimes N\to C\otimes N\to 0$ is exact, so $-\otimes N$ is exact whence $N$ is flat.

	Clearly (c)$\Rightarrow$(d), so it remains to see that (d) implies any of (a),(b), or (c).

\end{sol}

\begin{prob}
	Let $(R, \frakm, k)$ be a commutative local ring and $M$ be a finitely generated $R$-module. Prove that the following are equivalent: 
	\begin{enumerate}
		\item $M$ is free
		\item $M$ is flat
		\item The map $\frakm \otimes_R M \to R \otimes_R M =M$ induced by the embedding $\frakm \subset R$ is injective
		\item $\Tor_1(k,M)=0$.
	\end{enumerate}
\end{prob}
\begin{sol}
	(a)$\Leftrightarrow$(b): First let $M$ be free (and finitely generated). Then $M\cong R^k$ for some finite $k$. 
	Then take an injection
\end{sol}

\begin{prob}
	Let $R$ be a Frobenius algebra over a field $k$. Show that the global dimension of $R$ is either zero or infinity.
\end{prob}
\begin{sol}
	Assume that $\gldim(R)=n<\infty$. Then for any $R$-module $M$ both $\projdim_R(M)$ and $\injdim_R(M)$
	are less than or equal to $n$. Take minimal projective and injective resolutions $P_\bullet\xrightarrow{\varepsilon} M$ and $M\xrightarrow{\eta} I_\bullet$
	where in particular we have $I_k=P_k=0$ for all $k>n.$ But then consider the diagram 
	\begin{center}
		\begin{tikzcd}
			P_n\ar[r] & \cdots\ar[r] & P_1\ar[r] & P_0\ar[dr,"\varepsilon"]\ar[rr,"\eta\circ\varepsilon"] & & I_0\ar[r] & \cdots\ar[r] & I_n\ar[r] & 0\\
			& & & & M\ar[ur,"\eta"] & & & &
		\end{tikzcd}
	\end{center}
	where, since $\varepsilon$ is epi and $\eta$ is mono, the top row is still exact. Therefore since $R$ is Frobenius, 
	the $I_k$ are all projective as well, so this is a projective resolution of $I_n$. Then $\Ext^i_R(I_n,N)$ (for any $N$)
	is zero everywhere except (possibly) the zeroth position.

	Here we consider two cases. Notice that $n=0$ if and only if $M$ is projective and injective. In this case we are done since $\projdim_R(M)=\injdim_R(M)=0$. If this is not the case, 
	$P_1$ and $I_1$ are both not trivial. Then we can continue by noticing that the homology at $\Hom_R(P_1,N)$ is unaffected 
	by this splicing of sequences. Observe 
	\[\Ext_R^{n+1}(I_n,N)=\Ext_R^1(M,N)=0\]
	which implies that $M$ is projective (since $N$ was arbitrary) and injective. But then all $R$-modules are projective and injective,
	so $\gldim(R)=0$.
\end{sol}

\begin{prob}
	Let $k$ be a field of positive characteristic $p$. 
	\begin{enumerate} 
		\item Show that the group algebra $kG$ for a finite group $G$ is Frobenius (in fact, symmetric). 
		\item Show that the restricted enveloping algebra $\fraku(\gl_n)$ is Frobenius. 
	\end{enumerate} 
\end{prob}
\begin{sol}

\end{sol}

\begin{prob}
	Let $\underline{x} = (x_1, x_2,\ldots, x_n)$ be a sequence of 
	elements in a commutative ring $R$, and let $K(\underline{x})=K(x_1) \otimes \ldots \otimes K(x_n)$ 
	be the corresponding Koszul  complex.

	\begin{enumerate}  
		\item Prove the explicit formula for the differential in $K(\underline{x})$. 
		\item Show that $K(\underline{x})$ 
		is a graded commutative DGA and, moreover, that $K(\underline{x}) \simeq \Lambda^*(V)$ where $V = \bigoplus Rx_i$.
	\end{enumerate}
\end{prob}
\begin{sol}

\end{sol}

\end{document}

### 2019-05-16 Hopf Algebra Notes/Hopf Algebras.tex
Haven't pushed in a while

@@
 
-%identity map
+%Homological algebra stuff
 \DeclareMathOperator{\id}{id}
-
 \DeclareMathOperator{\End}{End}
-
 \newcommand*{\Hom}{\ensuremath\text{Hom}}
-
-%Hopf algebra stuff
-\newcommand*{\Vectk}{\operatorname{Vect}_k}
-\newcommand*{\Algk}{\operatorname{Alg}_k}
-\newcommand*{\Coalgk}{\operatorname{Coalg}_k}
-\newcommand*{\calM}{\mathcal{M}}
-\DeclareMathOperator{\Coalg}{Coalg}
-\DeclareMathOperator{\Alg}{Alg}
-\newcommand*{\scrT}{\ensuremath\mathscr{T}}
-\newcommand*{\Ch}{\mathbf{Ch}}
-\newcommand*{\g}{\mathfrak{g}}
-\DeclareMathOperator{\Der}{Der}
 
@@
 %make the title page
-\title{Hopf Algebras\vspace{-1ex}}
-\author{A course by: Prof. James Zhang\\
+\title{Homological Algebra\vspace{-1ex}}
+\author{A course by: Prof. Julia Pevtsova\\
 Notes by: Nico Courts}
-\date{Winter 2019}
+\date{Spring 2019}
 \maketitle
@@
 
-	This course is an introduction to Hopf algebras. In addition to basic material in 
-	Hopf algebra, we will present some latest developments in quantum groups and tensor 
-	and fusion categories. One of the newer topics is homological properties of 
-	Noetherian Hopf algebras of low Gelfand-Kirillov dimension. A good reference for 
-	the first two topics in the book \textit{Hopf Algebras and Their Action Rings} by 
-	Susan Montgomery. Here is a list of possible topics:
-
-\begin{itemize}
-	\item Classical theorems concerning finite dimensional Hopf algebras.
-	\item Infinite dimensional Hopf algebras and quantum groups.
-	\item Duality and Calabi-Yau property.
-	\item Actions of Hopf algebras and invariant theory.
-	\item Representations of Hopf algebras, tensor and fusion categories.
+	This course is an introductory course on homological algebra. We will be following the 
+	book \textit{An Introduction to Homological Algebra} by Charles Weibel. We will be
+	covering the following topics:
+
+\begin{itemize}
+	\item Chain complexes, homotopies, homology and long exact sequence in homology
+	\item Resolutions, derived functors, Ext and Tor. Koszul complexes
+	\item Group (co)homology
+	\item Triangulated and derived categories
+	\item Spectral sequences or open topic depending on the class interests
 \end{itemize}
@@
 
-\section{January 7, 2019}
+\section{April 1, 2019}
 If you don't know what a symmetric tensor category is, today is going to be a three 

### 2019-05-16 Reading Notes/reading_notes.tex
Haven't pushed in a while

@@
 \newcommand*{\Lie}{\text{\textbf{Lie}}}
+\DeclareMathOperator{\stHom}{\underline{\Hom}}
 
@@
 \maketitle
+
+\part{Spring 2019}
+\section{Meeting Notes}
+
+\subsection{Week 3}
+Due to travelling and some growing pains teaching my own 308 class research has gotten off to a slow
+start. Luckily Julia's class requires us to do a presentation on homological algebra and last week 
+we discussed some possible topics (mainly drawing on Dave Benson's course notes from Oberwolfach which I need 
+to add as a reference here).
+
+\begin{qst} Is the right way to think about triangles in a homotopy category where the suspension
+measures failure to be exact?
+
+Is a Cartesian square just a pushout/pullback ``modulo homotopy''? These seem to be related by 
+the fact that a Cartesian square
+\begin{center}
+	\begin{tikzcd}
+		A
+	\end{tikzcd}
+\end{center}
+corresponds to a sequence
+\[X\to Y'\sqcup Y''\to Z\to \Sigma X\]
+which appears to be ``almost'' the fiber product except the $\gamma$ map.
+\end{qst}
+
+Maybe. Maybe not. Let's talk about it in the context of modular representation theory. The triangles 
+here comes from short exact sequences $0\to M'\to M\to M''\to 0$. But a triangle in the 
+triangluated category corresponds to $M'\to M\to M''\to \Sigma M$.
+
+What this means is that 
+\[\Ext_G^1(M'',M')\cong \stHom_G(M'',\Omega^{-1}M')\]
+
+For next week, read Neeman's paper on triangulated categories and try reading through the computation
+in \cite{julia-dave}.
+
+\subsection{Week 4}
+Most of my personal time has been spent on reading \cite{julia-dave}. I will summarize where I am at the moment.
+
+I am reading theorem 4.1, which states that if $G=(\Ga^-\times \Ga^-)\rtimes \Ga_{(1)}$, then $H^{\ast,\ast}(G,k)$
+is generated by $\zeta$ in degree $(1,1)$, $x\in(2,0)$, $\kappa\in(p,1)$ and $\lambda_i\in (i,i+1)$ for $1\le i\le p-2$.
+All of these generators are found by computations earlier giving a presentation of $kG$ as a quotient of a polynomial ring:
+\[\frac{k[u,v,t]}{(u^2,v^2,uv+vu,t^p, tv-vt, tu-ut-v)}\]
+
+\begin{qst}
+	We let $\mu_p$ be the (finite) group scheme of $p^{th}$ roots of unity and say that $\mu_p\times\mu_p$
+	acts on $kG$ where the first element acts on $u$ and the second on $t$.
+\end{qst}
+Wait... is this literally just multiplication by a (principle? does that even necessarily exist in general?) $p^{th}$ root of unity?
+
+\begin{qst}
+	What does the ``fibre'' correspond to the the spectral sequence?
+\end{qst}
+
+\subsection{Week 5}
+The main thing I have been considering this week has been the computation
+\[d(\lambda^\varepsilon\eta^a\zeta^bx^c)\]
+where all terms in the sum (using the Leibniz rule) die (since $\lambda,\zeta,x$ lie in the bottom row) except 
+\[a\lambda^{\varepsilon+1}\eta^{a-1}\zeta^{b+1}x^c\]
+where we've used that $d(\eta)=\lambda\zeta$. It is clear that for this to be nonzero, we need $a\le p-1$
+and when $\varepsilon=1$. \textbf{It ends up there was an error}
+
+\brk
+
+\subsubsection{Elementary abelian group schemes}
+$\Ga_{(r)}$ has group aglebra isomorphic to the the truncated polynomial algebra $k[x_i]/(x^p)$. Then an elementary abelian
+group scheme is $\Ga_{(r)}\times(\bbZ/p)^k$. Now the homology (of the two factors?) are isomorphic, but
+the Steenrod operations, which can further distinguish between them.
+
+The Steenrod operations can bee seen as $\cal P$ and $\beta$, which is the connection homomorphism that arises 
+from the long exact sequence in cohomology arising from $\bbZ/p\to\bbZ/p^2\to\bbZ/p$: $\beta:H^1(G,\mathbb{F}_p)\to H^2(G,\mathbb{F}_p)$. We can generalize to other
+fields, but the $\beta$ map is no lnger linear, but $p$-semilinear!
+
+There is a theorem by Serre on Bocksteins (?). Frattini quoitient -- maximal elementary abelian quotient -- look this up!
+
+So the idea thien is we use this Frattini thing (think an analog to Nakayama) and we are trying to see when the Frattini quotient
+is the group itself. 
+\begin{lem}
+	Let $\varphi:H\to G$ be a (surjective?) map of $p$-groups and $\varphi^*$ ist induced map on the holomology rings with trivial coefficients.
+	then if $\varphi^*$ restricted to $H^1$ is an isomorphism and on $H^2$ is injectice, then $\varphi$ is an isomorphism.
+\end{lem}
+\begin{rmk}
+	This requires a result that $H^1(G,k)\ne 0$ when $G$ is a $p$ group.
+\end{rmk}
+
+\subsection{Week 6}
+
+\section{Group (Scheme) Cohomology}
+
+To get some more background for reading \cite{julia-dave}, I began reading Evens' book on group cohomology \cite{evens}.
+So I will begin here.
+
+I need to get better about typing notes! It is so useful!
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\part{Winter 2019}
+\section{Preparation for the Quarter}
+This is my first official quarter as Julia's student! My plan for now is to continue working on Waterhouse as well
+as learn about algebraic geometry (alongside my usual classes, of course).
+To give a sense of direction, Julia recommended that I take a look at the following regularity theorem:
+\begin{thm}[Smoothness Theorem]\label{smooth}
+	Let $G$ be an algebraic affine group scheme over a field $k$. Then $k[G]\otimes \bar k$ is reduced if and
+	only if $\dim G = \rank \Omega_{k[G]}$.
+\end{thm}
+
+\section{Overarching Ideas and Notes}
+\subsection{Week 2}
+Consider the representations of A over a field $k$ (or more generally $A$ modules, which we can make into a tensor category).
+Recall that in this category we have enough projectives and (with finitely generated Hopf algebras) we get that all projectives are injective.
+
+The object of study here:
+\begin{defn}
+	The \textbf{Hochschild cohomology} of a Hopf algebra $A$ over the field $k$ is 
+	\[\Ext_A^*(k,k)=H^*_A(A,k)\].
+\end{defn}
+
+The conjecture here is
+\begin{conj}
+	when $A$ is a finite dimensional Hopf algebra, $H^*(A,k)$ (the Hochschild cohomology) 
+	is finitely generated as a $k$-algebra.
+\end{conj}
+This is known in some special cases:
+\begin{itemize}
+	\item Finite groups
+	\item Finite group schemes (in positive characteristic) due to Friedlander and Suslin.
+	\item In characteristic zero we have
+	\begin{itemize}
+		\item Quantum groups
+		\item Hopf algebras that come from Nichols algebras
+	\end{itemize}
+\end{itemize}
+
+Papers to look at:
+\begin{itemize}
+	\item Ginzburg and Kumar '94 ``Cohomology of Quantum Groups at Roots of Unity'' \cite{ginzburg-kumar}
+	\item Mastnak, Pevtsova, Shauenburg, and Witherspoon '09 ``Cohomology of Finite Dimensional Pointed Hopf Algebras''\cite{julia10-pointed}
+	\item Witherspoon '17 ``Varieties for Modules of Finite Dimensional Hopf Algebras''\cite{witherspoon-expository}
+\end{itemize}.
+
+Sarah Witherspoon has a book she is writing on Hochschild cohomology that could be very good. It's on her website (Texas A\& M). Check out chapter 9 and appendix A.
+
+\subsection{Week 3}
+We primarily discussed support varieties in the context of Sarah Witherspoon's \textit{Varieties for Modules of Finite Dimensional Hopf Algebras.}\cite{witherspoon-expository}
+
+\subsubsection{Questions}
+\begin{enumerate}
+	\item On page 5, in the definition of the support variety of a Hopf module $M$ (see Definition~\ref{def-support-var}),
+	the construction seems to depend rather heavily on the choice (and existence!) of a graded subalgebra $H\le \HH^*(A)$
+	satisfying (fg1) and (fg2). But the notation seems to imply there is no dependence there. What's going on?
+	\item On page 9, why is $\mathcal{Z}(kG)=\HH^0(kG)$ a local ring (for $G$ a finite group)?
+	\item Directly after the last question, why would this ring being local imply that, on the variety level,
+	we can use $H^{ev}(G,k)$ instead of $H^{ev}(G,k)\cdot \HH^0(kG)$?
+\end{enumerate}
+
+\subsubsection{Answers/Hints}
+\begin{enumerate}
+	\item Actually the definition of support variety, when properly generalized to triangulated tensor categories,
+	doesn't rely on the choice of this algebra. But to see this, we need to dive deeper into that subject. Julia gave 
+	me a book by Dave Benson, Srikanth Iyengar, and Henning Krause on local cohomology and support. It was written
+	as a compendium from a course at Oberwolfach that they gave. It is available on the Arxiv.
+	\item Julia suggests that this is a ``standard'' result that one might find in Webb's book (or another book where the author has taken care to be thorough).
+	I couldn't locate it there immediately, but I found a page online that gave me the maximal ideal, so I am going to try to prove it as lemma~\ref{lem-local}.
+	\item The take-away here was that when computing the varieties the nilpotent elements are irrelevant. Since 
+	the only maximal ideal in $\HH^0(kG)$ is composed of nilpotents, it doesn't contribute to the support variety.
+\end{enumerate}
+
+\begin{defn}[Support Variety]\label{def-support-var}
+	Let $A$ be a finite dimensional algebra and assume $_AA$ is injective (notice that finite dimensional Hopf algebras are
+	Frobenius whence self-injective). Assume further that there is a graded subalgebra $H\le \HH^*(A)$ that satisfies the following 
+	finiteness conditions:
+	\begin{itemize}
+		\item \textbf{(fg1):} $H$ is finitely generated, commutative, and $H^0:=H\cap \HH^0(A)=\HH^0(A)$;
+		\item \textbf{(fg2):} for all finite dimensional $A$-modules $M$, $\Ext^*_A(M,M)$ is finitely generated as an $H$-module.
+	\end{itemize}
+
+	Then if $I_A(M)$ is the annihilator of $\Ext_A^*(M,M)$ in $H$, the \textbf{support variety} is
+	\[V_A(M)=\operatorname{MaxSpec(H/I_A(M))}\]
+\end{defn}
+\begin{rmk}
+	In taking with Julia about this definition she mentioned that the use of the max spectrum instead of the entire
+	spectrum is a bit watered-down and, while simpler, perhaps hearkens back to the beginnings of AG where that is what one considered.
+\end{rmk}
+
+\begin{lem}\label{lem-local}
+	If $G$ is a finite $p$-group and $k$ a field of characteristic $p$, $\mathcal{Z}(kG)$ is local.
+\end{lem}
+\begin{prf}
+	I claim that the maximal ideal is
+	\[I=\left\{\sum_{g\in G}a_gg:\sum_ga_g=0\right\}\]
+	To see this, let $A=\mathcal{Z}(kG)$, and notice that $A=\ker\varphi$ where $\varphi(\sum a_g g)=\sum a_i$.
+	Since $\varphi(kG)=k$, this gives us that $I$ is a maximal ideal in $A$.
+
+	Notice that if we prove that $I$ consists entirely of nilpotents, we will be done. This is because
+	the nilradical of $A$ is the intersection of its prime ideals. But if $I$ is contained in the nilradical, 
+	it must, in particular, be contained in every maximal ideal of $A$. Thus it must be the only maximal ideal.
+
+	To see this, we proceed by induction on $n$ where $|G|=p^n$. When $n=0$ this 
+	is obvious and when $n=1$, this means $G=\bbZ/p\bbZ=\langle\alpha\rangle$. In this case each element is nilpotent of 
+	degree $p$ due to the fact that (since in this case our ring is commutative):
+	\[\left(\sum_1^p a_i\alpha^i\right)^p=\sum_1^p a_i^p\alpha^{pi}=\sum_1^p a_i1_G=0.\]
+
+	Now assume that $I$ consists of nilpotents for all $n\le k$ and let $|G|=p^{k+1}$. Let $Z\le G$ be the
+	(necessarily nontrivial) center of $G$. Extend the canonical quotient map $q:G\to G/Z$ to a map 
+	between group rings $\tilde q:kG\to k(G/Z)$. Let $I_Z$ be the augmentation ideal of $k(Z)$. Then
+	$\tilde q$ is surjective with kernel $I_Z\cdot kG$. That this ideal is contained within the kernel is clear
+	since if $y$ is a $kG$-linear sum of elements of the form $\sum_Z a_zz$ where $\sum_Z a_z=0$, then
+	the image is just the sum of zeros.
+
+	To see that every element of the kernel is of this form, let $y=\sum k_i g_i$ and notice
+	\[\tilde q\left(\sum k_ig_i\right)=\sum k_i\tilde q(g_i)=\sum_j\left(\sum_{g_i-g_j\in Z}k_i\right)g_j\]
+	and this is zero precisely when all the sums are. But then the sum of the coefficients of $y$ restricted to any coset of $Z$
+	are zero.
+
+	Thus in particular $kG/(I_Z)\cong k(G/Z)$.
+
+	Then look at the augmentation ideal $I_{G/Z}\lhd k(G/Z)$ which, by the induction hypothesis, consists of nilpotents.
+	Pull this back through $\tilde q$ to the ideal $I=(I_{G/Z})+ (I_Z)$. Since some power of $p$ kills all elements of the two 
+	summands, this implies that some power of $p$ kills each element in $I$, completing the proof.
+\end{prf}
+
+\subsection{Week 4}
+Recall that $\mathcal{U}:\Lie\to \Ass$ is a functor that assigns to each 
+Lie algebra the universal enveloping algebra: the tensor algebra over $\g$ modulo the relations
+$x\otimes y-y\otimes x-[x,y]$.
+
+Then we have the PBW theorem:
+\begin{thm}[Poincar\'e-Birkhoff-Witt]
+	If $\g$ is a finite dimensional Lie algebra, with basis $x_1,\dots,x_n$, then $\mathcal{U}(\g)$ has a basis
+	in the form $\prod x_i^{a_i}$ for $a_i\in\bbZ_{\ge 0}$.
+\end{thm}
+\begin{rmk}
+	A fancier way of saying this is to say that $\mathcal{U}(\g)$ admits a filtration such 
+	that $\operatorname{gr}\mathcal{U}(\g)\cong S^*(\g)$.
+\end{rmk}
+
+Furthermore, there is a more general class of algebras called \textbf{PBW algebras} where you get commutation
+modulo lower-order terms. This idea is also what forms the basis of the proof for the above remark.
+
+The idea here, then, is to take the center and see how it acts on the universal enveloping algebra.
+
+\subsubsection{The BGG category $\mathcal{O}$}
+The representations of finite dimensional Lie algebras are in correspondence with the dominant integral weights. This
+will be worth looking into, perhaps.
+
+The general question we want to answer is what he structure of finite dimensional representations of $\g$. In the case
+when $\g=\sl_2$, we have the representation $V=ke_1\oplus ke_2$.
+
+Here I need to work out some of the notation, but $k[x,y]$ is a representation and $S^d(V)=k_d[x,y]$ is as well. (F\&H lecture 11 for more details there).
+
+So we end up realizing that we can generate any such Lie algebra by taking a single highest-weight vector and 
+applying $\mathcal{U}(\mathfrak{n}^-)$ to get the whole thing. This help motivate:
+\begin{defn}[BGG Category $\mathcal{O}$]
+	$\mathcal{O}$ is the full subcategory of $\mathcal{U}(\g)$ modules such that
+	\begin{itemize}
+		\item $M$ is finitely generated over $\mathcal{U}(\g)$ -- this is necessary since we want free modules
+		and $\mathcal{U}(\g)$ itself to be representations under study and these are excluded from finite ($k$-) dimensional algebras.
+		\item $M$ is $\mathfrak{h}$ semisimple.
+		\item $M$ is $\n$-locally finite -- for all $v\in M$, $\mathcal{U}(\n)v$ is finite dimensional.
+	\end{itemize}
+\end{defn}
+
+Ideas for things to understand/cover for next week:
+\begin{itemize}
+	\item First things first, understand the representations of $\sl(2,k)$.
+	\item Then in Humphreys(I), read/do section \S 6.21 to understand the universal enveloping algebra.
+	\item Read Humphreys(II) chapter 1 as much as possible.
+\end{itemize}
+
+\subsection{Week 5}
+We spent most of the time deriving the fact that any simple $\sl_2$ module is isomorphic to one of the ``standard''
+highest weight modules.
+
+Our assignment for this upcoming week:
+Write down the Serre relations. Then do some computations and read section 21 in Humphreys. Read those lemmas and theorems and start to describe decomposition of general modules of Lie algebras.
+
+\subsection{Week 6}
+We skipped Lie algebras this week because of snow (and then Ale couldn't make our meeting). So instead
+Julia just met up to talk about my reading.
+
+My primary question (regarding the Ginzburg and Kumar paper): why are we factoring out the toral elements in quantum groups? 
+The answer here is that these are called the ``small quantum groups''. So certainly having a finite-dimensional algebra makes certain things
+more tractable, although when computing cohomology that isn't always the case. Sometimes when you get smaller 
+objects you end up getting more nontrivial elements in the cohomology. There's a clear analog to topology here.
+
+So then maybe a refinement of the question is this: why is it necessary at all to have this larger object serving as the ambient 
+space for the object we truly want to study? Here the answer more-or-less boils down to the fact that it end up playing a central
+role in computations involving these objects.
+
+Having the big algebra $U$ over $u$ (modulo something central) gives you that thePBW basis for $U$ descends to 
+$u$. In general (e.g. Nichols algebras) you fail to have a PBW basis for the cover, but somehow (for a technical reason that 
+will become apparent as we read more) you still can use similar techniques to do the computations (the more general theory arises in the 2010 paper).
+
+Julia is going to sent me a paper that is a survey of finite dimensional algebras that will give some sense for how the small quantum groups fit in.
+
+\subsubsection{This week's assignment}
+Continue reading through GK until we get to the spectral sequence nonsense. Then this points to Jantzen (AG). Try to recreate the
+argument for constructing the first page of the spectral sequence, but for the GZ case (note, you'll use projectives instead of injectives)
+After that, go to Weibel and try to read the section of spectral sequences over filtered complexes \S 5.4 (what we are usually interested in).
+Don't worry too much about issues of convergence -- we are mostly interested in the ``first quadrant'' case.
+
+David Anick is the name of an algebraic topologist who started off created (free) resolutions that we will end up looking at. 
+His paper \cite{anick-resolution} will probably be useful at some point.
+
+\subsection{Week 8}
+The last paper on Julia's page is a preprint with her and Dave Benson. There is a spectral sequence computation that might be illustrative.
+There is also an older paper...
+
+One of the things Julia is interested in at the moment is a super group scheme. These arise by having a $\bbZ_2$ grading on your Hopf algebra.
+For instance we can look at exterior algebra $\Lambda(x)$, which represents the super group scheme $G_a^-$.
+
+The first non-elementary example is $G_a^-\times G_a^-$ with a $\bbZ_p$ action on it that ``folds'' the second factor
+onto the first. To write this out explicitly, we consider the group algebra $\Lambda(u,v)$ where $uv=-vu$ and the coordinate algebra of $\bbZ_p$
+is $k[t]/t^p$ where $t\cdot u=0$ and $t\cdot v=\alpha u$ for some $\alpha$. This is a $4p$ dimensional algebra.
+
+Then we want to compute coholomology. We know $H^{**}(k\bbZ/p,k)\cong k[x]\oplus\Lambda(y)$ for $|x|=2,0$.
+Note there is a notion of dimension that is an ordered pair $(a,b)$ of the graded degree and cohomology degree.
+THis give us new elements of degree $(1,1)$
+
+Look up Koszul duality! It is in the representations and local cohomology book!
+
+\subsection{Week 10}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\newpage
+\section{Lie Algebras}
+We're planning on digging into the category $\O$, the category of (reasonably finite) $\mathcal{U}(\g)$ modules.
+To begin, however, we are going to make sure we have our fundamentals vis a vis (semi) simple representations of Lie algebras.
+
+\subsection{$\sl_2$}
+In the case when $L=\sl_2$, everything is relatively nice. We read (and discussed) the fact that any simple, finite-dimensional
+$\sl_2$-module is isomorphic to $V(\lambda)$ for some integer $\lambda$ (highest weight).
+
+The action of $L$ on these modules are precisely as one would hope: $x$ and $y$ (the off-diagonal generators) act by raising 
+and lowering the weight of a vector by 2 and the central generator fixes each eigenspace.
+
+\subsection{The Serre Relations}
+Let $L$ be a semisimple Lie algebra and $\h$ a Cartan subalgebra with corresponding roots $\Phi$ with simple root system $\Delta$.
+Recall that 
+\[\langle\alpha_i,\alpha_j\rangle=\frac{2(\alpha_i,\alpha_j)}{(\alpha_j,\alpha_j)}=\alpha_i(h_{\alpha_j})\]
+Fix generators $x_i$ and $y_i$ of $L_{\alpha_i}$ and $L_{-\alpha_i}$, respectively, so that $[x_i,y_i]=h_i=h_{\alpha_i}$
+Recall that the $x_i$ and $y_i$ generate $L$ (as a Lie algebra).
+
+Then in particular we have the relations:
+\begin{align*}
+	[h_i,h_j]&=0\tag{S1}\\
+	[x_i,y_j]&=\delta_{ij}h_i\tag{S2}\\
+	[h_i,x_j]=\langle\alpha_j,\alpha_i\rangle x_j\text{ and }& [h_i,y_j]=-\langle\alpha_j,\alpha_i\rangle y_j\tag{S3}
+\end{align*}
+\begin{defn}[Serre Relations]
+	With the notation above, the relations S1-3 above are called the \textbf{Serre relations}.
+\end{defn}
+
+These relations end up being crucial in making the connection between root systems and semisimple Lie algebras concrete:
+taking the abstract free Lie algebra on generators satisfying these properties, we get a finite dimensional semisimple Lie algebra
+with roots corresponding exactly to the $\alpha_i$. This also gets us uniqueness of such a root system.
+
+\subsection{The more general case.}
+This primarily borrows from \S 21 in \cite{humphreys1} but also relies on some results from earlier that I will cite as 
+necessary.
+
+Throughout this discussion, we rely on the theory developed in sections 15 and 16 regarding Cartan and Borel subalgebras:
+\begin{defn}[Cartan Subalgebra (CSA)]
+	If $L$ is a Lie algebra, then $C\subseteq L$ is called a \textbf{Cartan subalgebra} if $C$ is
+	\begin{enumerate}
+		\item Nilpotent; and
+		\item $N_L(C)=C$.
+	\end{enumerate}
+\end{defn}
+\begin{defn}[Borel Subalgebra]
+	If $L$ is a Lie algebra, $B\subseteq L$ is called a \textbf{Borel subalgebra} if $B$ is a
+	maximal solvable subalgebra.
+\end{defn}
+\begin{rmk}
+	Since every nilpotent Lie algebra is solvable, each CSA lies in some Borel subalgebra $B$ of $L$.
+	then by showing that any two Borel subalgebras are conjugate under $\mathscr{E}(L)$ (the subalgebra of $\Int L$ generated
+	by the strongly $\ad$-nilpotent elements of $L$), it suffices to 
+	show that two CSAs are $\mathscr{E}(L)$-conjugate in any \textit{semisimple} Lie algebra.
+\end{rmk}
+
+In all that follows, we've fixed a CSA $\h$ of $L$ and root system $\Phi$ with simple root system $\Delta=\alpha_1,\dots,\alpha_l$
+of positive roots with $\mathcal{W}$ the Weyl group.
+
+\subsection{Representations}
+Since $\h$ is semisimple (since $L$ is), it acts diagonally on any finite dimensional $L$-module. Thus we can always
+do the same trick of decomposing $V$ into eigenspaces $V=\sqcup_{\lambda\in \h^*}V_\lambda$ that we do for $\sl_2$.
+Slightly more generally (when $V$ is infinite-dimensional), we can consider the sum of weight spaces $V_\lambda$ of $V$, 
+which is necessarily  direct\footnote{Think: vectors can't have multiple eigenvalues.}, so there always
+exists a direct sum of simple modules contained in $V$. When $\dim V<\infty$, this is equal to $V$ itself.
+
+\subsubsection{Standard Cyclic Modules}
+Here
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\newpage
+\section{Quantum Groups}\label{sec-quantum}
+I am mostly working out of Jantzen's \textit{Lectures on Quantum Groups} (\cite{janzten-quantum}) with the goal of reading through Ginzburg \& Kumar's
+(apparently influential!) paper \textit{Cohomology of Quantum Groups at Roots of Unity} (\cite{ginzburg-kumar}).
+
+\subsection{What they are}
+The core idea to keep in mind is that quantum groups (or quantum enveloping algebras) are ``deformations'' (or perhaps $q$-analogues) of
+regular enveloping algebras. Well, at least that is the kind we're mostly working with here. A general definition eludes me (and Jantzen).
+
+Mostly we will be interested in $U=\mathcal{U}_q(\sl_2)$. In this setting, we get the definition
+\begin{defn}[Quantum Enveloping Algebra of $\sl_2$]
+	$\mathscr{U}_q(\sl_2)$ is the associative algebra $k\langle E,F,K,K^{-1}\rangle$ under the following relations:
+	\begin{align*}
+		KK^{-1}&=K^{-1}K=1\tag{R1}\\
+		KEK^{-1} &= q^2 E\tag{R2}\\
+		KFK^{-1} &= q^{-2}F\tag{R3}\\
+		EF-FE &= \frac{K-K^{-1}}{q-q^{-1}}\tag{R4}
+	\end{align*}
+	where $q$ is some nonzero element of $k$ such that $q^2\ne 1$.
+\end{defn}
+
+Some notation that may be helpful in the future:
+\begin{itemize}
+	\item For $a\in\bbZ$, define $[a]=\frac{v^a-v^{-a}}{v-v^{-1}}\in\bbQ(v)$ -- the $v$ analogue of $a$. Equivalently, (and perhaps suggestively):
+	\[[a]=v^{a-1}+v^{a-3}+\cdots+v^{-a+3}+v^{-a+1}.\]
+	\item the $v$ analogue of the factorial is
+	\[[a]^!=[a][a-1]\cdots[1]\]
+	\item Then you get the binomial coefficients:
+	\[\qbinom{a}{n}=\frac{[a]^!}{[n]^![a-n]^!}\]
+	\item \[[K;a]=\frac{Kq^a-K^{-1}q^{-a}}{q-q^{-1}}\]
+\end{itemize}
+
+\subsection{Getting a basis}
+The first part of the first chapter focuses on getting a PBW-style basis for $U$, which comes in the form $F^s,K^n,E^r$ for $r,s\in\bbN$, $n\in\bbZ$.
+This is mostly standard but I thought this proof had a neat idea in it:
+
+\begin{prf}
+	Let $A=k[X,Y,Z]$ and define $f,e,h\in \Hom_k(A,A)$ such that 
+	\begin{align*}
+		f(Y^sZ^nX^r)&=Y^{s+1}Z^nX^r,\\
+		e(Y^sZ^nX^r)&=q^{-2n}Y^sZ^nX^{r+1}+[s]Y^{s-1}[Z;1-s]Z^n X^r,\\
+		h(Y^sZ^nX^r)&=q^{-2s}Y^sZ^{n+1}X^r
+	\end{align*}
+	One can check that these vector space homomorphisms exactly mirror multiplication of these monomials
+	by $F, E,$ and $H$, respectively, so we get a map $\varphi:U\to\Hom_k(A,A)$. Then take any linear combination of 
+	$F^sK^nE^r$, mapped though to $A$ via $\varphi$ and then evaluation at 1, gives us a linear combination of the $Y^sZ^nX^r$,
+	so it must be trivial. Sick.
+\end{prf}
+
+\brk
+
+After that, the book focuses on $U_0=k\langle Z,Z^{-1}\rangle$ as well as $U^+=k\langle E\rangle$ and $U^-=k\langle F\rangle$,
+the latter of two which are isomorphic to $k[x]$ by the linear independence above.
+
+\brk
+
+Here's another cool bit: By studying products of monomials (or just defining degree outright) we can impose a graded structure on $U$:
+let $\deg(E)=1$, $\deg(F)=-1$ and $\deg(K)=\deg(K^{-1})=0$ and notice (R1-4) are homogeneous according to this degree.
+Thus the quotient by this homogeneous ideal gives a graded structure to $U$.
+
+Furthermore if $u\in U$ is homogeneous of degree $i$, 
+\[KuK^{-1}=q^{2i}u\]
+so when the powers of $q$ are distinct ($|q|\ne 1$), the graded pieces are precisely the 
+ of $K$ under the adjoint action.
+
+\subsection{Representations of $U$}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\newpage
+\section{Reading Ginzburg \& Kumar}
+This section contains my notes and thoughts about \cite{ginzburg-kumar}. As a high-level summary, this paper establishes
+methods for computing (Hochschild) cohomology of the small quantum groups. It begins by establishing how they are constructed
+as quotients or subalgebras of quantized universal enveloping algebras (see section~\ref{sec-quantum}), and then proceeds to compute
+the cohomology using a filtration (this is vital!) and spectral sequences.
+
+This paper is important because it ends up that this general formula is the same as is applied later in \cite{julia10-pointed} to prove
+a much more general result. This also extends a method originally created by \cite{FP} for computing the cohomology of Lie algebras and
+algebraic groups.
+
+\subsection{The Main Result}
+In this paper they prove the following (see \ref{sec-defs} for notation):
+\begin{thm}[GK '93]
+	If $k=\bbQ(\xi)$ for a primitive root of unity of odd order (not 1) and $u_\xi$ the restricted enveloping algebra of some $\g$. Then
+	\[H^{odd}(u_\xi,k)=0\]
+	and there is a natural graded algebra isomorphism
+	\[H^{ev}(u_\xi,k)\cong k^\bullet[\mathcal N].\]
+\end{thm}
+
+\subsection{The Objects in Question}
+We begin by considering the objects in question and how they are constructed. One thing we haven't seen (yet!) in Lie algebras
+but comes up in the structure theory is the 
+\begin{defn}[Kostant $\bbZ$-form]\label{def-kostant}
+	Let $U(\g)$ be the universal enveloping algebra of $\g$, spanned (over $k$) by the $x_\alpha,y\alpha,$ and $h_\alpha$ for all $\alpha\in\Phi$. 
+	Then the \textbf{Kostant $\bbZ$-form} is the $\bbZ$-span of the elements
+	\[X^{(r)}_\alpha=\frac{x_\alpha^r}{r!},\quad Y^{(r)}_\alpha=\frac{y_\alpha^r}{r!},\quad H^{(r)}_\alpha=\frac{h_\alpha^r}{r!},\quad \]
+	the so-called \textbf{divided powers} of $x_\alpha, y_\alpha,$ and $h_\alpha$. This ($\bbZ$-)algebra is denoted $U_\bbZ$.
+\end{defn}
+
+Before we use it, we had better write down the definiton (via \cite{janzten-quantum}) of a
+\begin{defn}[Quantized Enveloping Algebra]
+	Fix some odd root of unity $q$ in the field\footnote{Here Jantzen also requires that $\ch k\ne2$ and that $\ch k\ne3$ if $\Phi$ has an irreducible component of type $G_2$.
+These can be treated, but not in quite the same manner.} $k$.
+	Let $\g$ be a Lie algebra over $k$ and let $U_q(\g)$ denote the $k$ algebra with generators $E_\alpha,F_\alpha, K_\alpha,$ and $K_\alpha^{-1}$
+	for all $\alpha$ in a basis $\Pi$ of $\Phi$, subject to the relations for all $\alpha,\beta\in\Pi$
+	\begin{align*}
+		K_\alpha K_\alpha^{-1}=&1=K^{-1}_\alpha K_\alpha\\
+		K_\alpha K_\beta &= K_\beta K_\alpha\\
+		K_\alpha E_\beta K_\alpha^{-1} &= q^{(\alpha,\beta)}E_\beta\\
+		K_\alpha F_\beta K_\alpha^{-1} &= q^{-(\alpha,\beta)}F_\beta\\
+		E_\alpha F_\beta - F_\beta E_\alpha &= \delta_{\alpha\beta}\frac{K_\alpha-K_\alpha^{-1}}{q_\alpha-q_\alpha^{-1}}\\
+		\sum_{s=0}^{1-a_{\alpha\beta}}(-1)^s\sqbinom{1-a_{\alpha\beta}}{s}_{q_\alpha} &E_\alpha^{1-a_{\alpha\beta}-s}E_\beta E_\alpha^s = 0\\
+		\sum_{s=0}^{1-a_{\alpha\beta}}(-1)^s\sqbinom{1-a_{\alpha\beta}}{s}_{q_\alpha} &F_\alpha^{1-a_{\alpha\beta}-s}F_\beta F_\alpha^s = 0
+	\end{align*}
+	where $q_\alpha=q^{d_\alpha}=q^{(\alpha,\alpha)/2}$ and $a_{\alpha,\beta}=\langle\alpha,\beta\rangle=2(\alpha,\beta)/(\alpha,\alpha)$.
+\end{defn}
+
+\begin{rmk}
+	In \cite{janzten-quantum}, he mentions that he will often work in the algebra that includes all the above relations
+	but \textbf{omitting the last two}. Apparently this will be done for computational simplicity and then studying $U_q(\g)$ 
+	itself will amount to taking a quotient. This may not be relevant to me, but it's good to keep in mind for this reference.
+\end{rmk}
+
+\brk
+
+For the case of quantum groups, we are extending all of our notions to quantized enveloping algebras in the case when our field $k$ is the rational function field
+$\bbQ(v)$.
+\begin{defn}[$q$-analog of the Kostant $\bbZ$-form]
+	Let $U=U_q(\g)$ be the (Hopf!) algebra described above. Then we define the $q$-analog of the Kostant $\bbZ$-form
+	to be the $\bbZ[v,v^{-1}]$-span of the $q$-divided powers the $E_\alpha,F_\alpha$ and $K_\alpha^{\pm 1}$, which are the same as
+	the divided powers in definiton~\ref{def-kostant} except the factorial is replaced by the $q$-analog thereof.
+\end{defn}
+
+\subsubsection{Notation and Related Algebras}\label{sec-defs}
+But this isn't what we're studying yet! (This are the so-called ``big'' quantum groups: what Drinfeld studies in \cite{drinfeld}). 
+We are interested in taking particular quotients and subalgebras. Here we define the notation used throughout and how they are related to $U=U_q(\g)$.
+
+We will be interested primarily in the case when $v=\xi$, that is, when $k$ is the cyclotomic field over $\bbQ$ found by adjoining
+an \textbf{odd} primitive $l^{th}$ root of unity ($l$ not 1). For the rest of this paper, fix such a $k$. Then define
+\[U_\varepsilon :=(k\otimes_{\bbZ[v,v^{-1}]}U_\bbZ)/(K_\alpha^l-1)=:U_k/(K_\alpha^l-1)\]
+where we defined $U_k$ by extension of scalars from $U_\bbZ$ and we are using here the fact that any $\bbZ[\xi]$-module 
+becomes a $\bbZ[v,v^{-1}]$-module by letting $v$ act by $\xi$.
+
+\brk
+
+Another construction is the so called \textbf{restricted enveloping algebra} $u_\xi$ of $U_\xi$ generated by the 
+degree one generators $F_\alpha, E_\alpha$, and $K^{\pm 1}_\alpha$. This is a finite-dimensional algebra of dimension $l^{\dim \g}$.
+This is actually the object that we are studying here. It amounts to restricting coefficients in $U_\xi$ to polynomials in $\xi$ with integer coefficients.
+
+\brk
+
+Define $\mathcal N$ to be the \textbf{nilpotent cone} of $\g$, consisting of all $\ad$-nilpotent elements in $\g$. This admits a 
+natural graded structure over the nonnegative integers by recognizing it as the ring of regular functions on a cone.
+
+\subsubsection{Another formulation}
+
+There is a parallel construction of $u_\xi$ that highlights a ``duality'' property of these objects. Set $\mathscr U_\bbQ$ to be the 
+$\bbQ[v,v^{-1}]$-span of $E_\alpha,F_\alpha,K^{\pm 1}_\alpha$ (notice we are not using divided powers here, but allowing coefficients in $\bbQ$) 
+and extend scalars again to $\mathscr U_k$. Again let $\mathscr U_\xi$ be the quotient
+where we identify $K_\alpha^l=1$. Define $\mathscr L$ to be the subalgebra of $\mathscr U_\xi$ generated by $E_\alpha^l$ and $F_\alpha^l$.
+
+Then it is a result of Lusztig that $\mathscr L$ and $u_\xi$ are normal (algebras) in $\mathscr U_\xi$ and $U_\xi$, respectively,
+and that 
+\[\mathscr U_\xi//\mathscr L\cong u_\xi\qquad\text{and}\qquad U_\xi//u_\xi\cong U(\g)\]
+where
+\begin{defn}
+	Given an augmented algebra $B$ with augmentation ideal $B_+$, a subalgebra $A\subseteq B$ is called \textbf{normal} in $B$ if 
+	\[B\cdot (A\cap B_+)=:B\cdot A_+=A_+\cdot B.\]
+\end{defn}
+
+(My) intuition for augmented algebras come from group algebras $kG$, where you have the augmentation map $\epsilon(\sum_G c_gg)=\sum_G c_g$.
+Here the augmentation ideal is $\ker\varphi$. In fact, any map $\epsilon:A\to k$ you have induces such a structure. Usually in the case of Hopf algebras
+we pick the counit (as we do in group algebras).
+
+\begin{defn}
+	If $A$ is normal in $B$, then we write 
+	\[B//A:=B/(B\cdot A_+)=B/(A_+\cdot B).\]
+\end{defn}
+
+The duality we mentioned earlier comes from the fact that $\mathscr L$ is a commutative Hopf algebra that is ``dual'' to (the cocommutative) $U(\g)$.
+
+\subsubsection{Triangular Decompositions}
+We use superscripts $+,-,$ and $0$ to denote the subalgebras generated, respectively, by the $E$'s, $F$'s, and $K$'s\footnote{with the exception of $U_0$ which has another set of generators due to the divided powers}.
+
+Then for $U_\xi$, $\mathscr U_\xi$, and $u_\xi$ we get the standard triangular decompositions; e.g.
+\[u_\xi = u_\xi^+\oplus u_\xi^0\oplus u_\xi^-.\]
+
+For each, define the \textbf{Borel part} to be (e.g.) $\mathscr B_\xi=\langle K_\alpha, E_\alpha\rangle$ (i.e. the nonnegative part).
+
+\subsection{The Argument}
+\subsubsection{Filtrations}
+The core idea (according to Julia) is the existence of certain filtrations such that the associated graded algebras have nice structure.
+
+In this paper, we are using an earlier result in \cite{de-concini} which tells us that $\mathscr B_\xi$ admits such a filtration:
+(Recall that we are using roots $\alpha$ in some basis $\Pi$ of $\Phi$)
+\begin{thm}[DK '90]
+	The algebra $\mathscr B_\xi$ has a multiplicative filtration such that $\Gr \mathscr B_\xi$ is generated by the homogeneous elements $\{E_\alpha,K_\beta\}_{\alpha\in \Delta_+,\beta\in \Pi}$
+	subject to the relations
+	\[K_iK_j=K_jK_i\quad K_i^l=1\quad K_iE_\alpha=\xi^{\langle \alpha,\alpha_i\rangle} E_\alpha K_i\]
+	and
+	\[E_\alpha E_\beta=\xi^{\langle\alpha,\beta\rangle}E_\beta E_\alpha\]
+	if $\alpha\succ\beta$.
+\end{thm}
+Then we can use this filtration to get filtrations on all quotients and subalgebras, most importantly
+$\mathscr B_\xi,\mathscr U_\xi^+, b_\xi,$ and $u_\xi^+$.
+\begin{qst}
+	What is the filtration here? 
+\end{qst}
+I ended up finding the book in which \cite{de-concini} is published (in the library) and the answer is that you look at 
+monomials 
+\[M_{k,r;u}=F^kuE^r\]
+where $F^k=\prod_\alpha F_\alpha^{k_{\alpha}}$ and similar for $E^r$ and $u\in\mathcal U^0$.
+
+Then you can define the \textbf{height:}
+\[\operatorname{ht}(M_{k,r;u})=\prod_\alpha (k_\alpha+r_\alpha)\operatorname{ht}\alpha\in\bbZ_+\]
+(What is $\operatorname{ht}(\alpha)?$) and then finally define \textbf{degree:}
+\[d(M_{k,r;u})=(k_N,k_{N-1},\dots,k_1,r_1,\dots,r_N,\operatorname{ht}(M_{k,r;u}))\in\bbZ_+^{2N+1}\]
+which is totally ordered via the lexicographic ordering and therefore gives us a $\bbZ^{2N+1}$ grading.
+
+\brk
+
+Then we proceed to prove that $H^\bullet(\Gr\mathcal U_\xi^+)\cong\Lambda_\xi$, which follows from some arguments that can be found 
+in \cite{priddy}.
+\begin{qst}
+	What role do augmented algebras play in this theory? I have noted on several occasions that augmentations
+	have been used to define normal subalgebras. It also popped up in the discussion of Koszul algebras in \cite{priddy}.
+\end{qst}
+
+The idea here is to give $k$ the structure of an $A$-module. This is non-trivial for the folloiwng reason:
+consider $A=k[x_i]$ and localize at a non-maximal prime $p$. Then the residue field will have positive transcendence degree
+over $k$, and since $p$ will now be the (only!) maximal ideal, this will be the smallest field to which there is such a map.
+
+\begin{qst}
+	In the computation of cohomology for $\mathcal L^+$, why do the terms in the Koszul complex
+	look like $\mathcal L^+\otimes_k \Lambda^i(\mathfrak N)$? Is this exploiting somehow that 
+	all Hopf modules are trivial (e.g. $\mathcal L^+$) determines the entire action by 
+\end{qst}
+
+Let's take a regular sequence $(x_1,\dots,x_n)$. Then $R\xrightarrow{x_1} R$ is $K(x_1)$, the 
+Koszul complex for this regular subsequence. If we have two elements, then $K(x_1)\otimes K(x_2)$
+is the complex $R\to R\oplus R\to R$. Corollary 4.5.5 in Weibel says that whenever the sequence is regular, if you 
+construct the entire Koszul complex to get a resolution of the quotient $R/(x_i)$.
+
+Note that this essentially boils down to the computation that $\Lambda^iS^n\cong \Lambda^i(k^n\otimes_k S)=(\lambda^i k^n)\otimes S$.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\newpage
+\section{Julia \& Dave Benson's preprint}
+As an aside, in response to me saying I needed to practice doing computations with spectral sequences,
+Julia recommended I read her current work \cite{julia-dave} with Dave Benson wherein there is a supposedly
+straightforward spectral computation in support of their larger work \cite{detecting-nilpotence}.
+
+I have already spent some time reading this paper, but the core objects in question are \textbf{supergroup schemes}
+which have coordinate Hopf algebras algebras with a $\bbZ/2$ grading that are furthermore (graded) cocommutative (just 
+as group schemes correspond to finite dimensional cocommutative Hopf algebras). The fundamental example is $\bbG_a^-$,
+which has as its coordinate ring $k[x]/(x^2)$, or equivalently the exterior algebra on a single element $x$,
+where $x$ is taken to be a primitive element.
+
+Then if we let $A$ be the coordinate algebra of $\bbG_a^-\times\bbG_a^-$ and $H$ that of $\bbG_{a(r)}\times(\bbZ/p)^s$ where $\bbG_{a(r)}$ is the
+$r^{th}$ Frobenius kernel of $\bbG_a$, we can define the \textbf{smash product} or in this case, since 
+$A$ is graded cocommutative, the \textbf{semidirect product} $A\#H=A\rtimes H$ where we define the action
+to be (if $\tau:H\otimes A\to A$ is the action of $H$ on $A$)
+\[(a\otimes h)\cdot(a'\otimes h')=\sum (-1)^{|h_1||b|}a\tau(h_1\otimes b)\otimes h_2g.\]
+
+So we use that $A\cong \Lambda(k[u,v])$, the exterior algebra on two (primitive) generators $u$ and $v$
+and that, if $t_i$ is the generator for the coordinate algebra of the $i^{th}$ copy of $\bbZ/p$, and if 
+we let $g_i=t_i-1$, these along with the $s_1,\dots s_r$ generating the Frobenius kernel, generate $H$.
+
+One of the first propositions (3.1) of the paper computes the cohomology of $\bbG_a^-$ with 
+coefficients in $k$ (actually, refers to another paper) and computes how the ($\bbZ/2$-graded)
+Steenrod operations act on it.
+
+\begin{defn}[Steenrod Operations]
+	If $p$ is odd and $A$ is a $\bbZ$-graded cocommutative Hopf algebra, then there are natural maps/operations
+	\[\mathcal P^i:H^{s,t}(A,k)\to H^{s+(2i-t)(p-1),pt}(A,k)\]
+	\[\beta \mathcal P^i:H^{s,t}(A,k)\to H^{s+1+(2i-t)(p-1),pt}(A,k)\]
+	satisfying the properties
+	\begin{enumerate}
+		\item $\mathcal P^i=0$ whenever $2i<t$ or $2i>s+t$,
+		\item $\beta\mathcal P^i=0$ whenever $2i<t$ or $2i\ge s+t$.
+		\item $\mathcal P^i(x)=x^p$ when $2i=s+t$
+		\item $\mathcal P^j(xy)=\sum_i\mathcal P^i(x)\mathcal P^{j-i}(y)$
+		\item $\beta\mathcal P^j(xy)=\sum_i(\beta\mathcal P^i(x)\mathcal P^{j-i}(y)+\mathcal P^i(x)\beta\mathcal P^{j-i}(y))$
+		\item $\mathcal P$ and $\beta\mathcal P$ satisfy the Adem relations.
+		\item $\mathcal P^0(\lambda u)=\lambda^p\mathcal P^0(u)$ for all $\lambda \in k$ ($\ch k= p$)
+	\end{enumerate}
+\end{defn}
+
+In the course of proving things here, the authors adapt these operations to the $\bbZ/2$ grading
+by reindexing, calling these new operations $\mathscr P_i$ and $\beta\mathscr P^i$ (they use \texttt{mathscr} but I've
+reassigned mine to a font I like better. :P)
+
+Okay, time is running out, so I am pulling together some questions:
+\begin{qst}
+	Is the cohomology you're using here group cohomology? It seems to me the Lyndon-Hochschild-Serre
+	applies to group cohomology.
+
+	If so, I will need to focus some time on getting familiar with it.
+\end{qst}
+
+Yes it is group cohomology, or more specifically group \textit{scheme} cohomology. Here we 
+are computing $\Ext_G^{i,\varepsilon}(k,k)$.
+
+We begin with our group algebra (here $\Lambda(v_1,v_2)\#k\bbG_{a(1)}$) and take 
+a projective resolutions and computing $H^*(\Hom_A(p^\bullet,k))$.
+
+Recall the syzygys, $\Omega^i k$ that occur in a projective resolution of $k$. Then we can compute the cohomology
+via the stable module category:
+\[\underline{\Hom}_G(\Omega^ik,k)=\Ext_G^{i,\varepsilon}(k,k)\]
+
+There is a result by Jeremy Rickard (rep theory) and by Buchweitz (more generally) that proves
+
+\begin{qst}
+	Where is $H^{*,*}(\bbG_a^-,k)$ computed? The papers claim it is $S^{*,*}(V^\sharp)\cong k[\zeta,\eta]$
+	where $\zeta$ and $\eta$ are the dual generators corresponding to the generators $u$ and $v$ of $V$,
+	the augmentation ideal of $k(\bbG_a^-\times \bbG_a^-)$.
+\end{qst}
+
+Again, this is an example of Koszul duality -- use that $k\bbG_a^-=\Lambda(v)$ where $v$ is primitive in degree 1.
+But then $k(\bbG_a^-\times\bbG_a^-)\cong\Lambda(v_1,v_2)$. Then applying Koszul duality you get the 
+symmetric algebra. Also you need the K\"unneth formula, where as algebras
+\[H^*(A\otimes B,k)\cong H^*(A,k)\otimes H^*(A,k).\]
+
+Evans has a book on The groups of cohomology. Here there are computations and resolutions of (cyclic)
+groups. There is also a discussion of spectral sequences which is relatively terse (although it defers
+somewhat to other sources).
+
+\begin{qst}
+	What is the importance of suspensions and what role do they play in this theory? They 
+	have come up at several points and I am realizing they are fairly pervasive.
+
+	In particular, I understand the topological definition, but I don't see why they correspond
+	to what we want. I am looking at the formula from \cite{detecting-nilpotence}:
+	\[H^{1,1}(G,k)=\Hom_{sGr/k}(G^0,\mathbb{G}_a^-)^\pi\]
+	where $\pi$ is the group of connected components of $G$.
+\end{qst}
+
+\begin{qst}
+	In $G=(\bbG_a^-\times\bbG_a^-)\rtimes G_{a(1)}$, where the action is said to be ``nontrivial'',
+	why does this seem to imply that the action on one of the copies of $\bbG_a^-$ is trivial?
+
+	The fact that there is a LHS spectral sequence 
+	\[H^{*,*}(\bbG_a^-\times \bbG_{a(1)},H^{*,*}(\bbG_a^-,k))\Rightarrow H^{*,*}(G,k)\]
+	seems to imply that there is a short exact sequence 
+	\[1\to \bbG_a^-\to G\to \bbG_a^-\times \bbG_{a(1)}\to 1\]
+	but since the product is direct on the right, doesn't this imply that the action of $\bbG_{a(1)}$ on 
+	$\bbG_a^-$ is trivial?
+\end{qst}
+
+Path 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\newpage
+\section{Affine Group Schemes (Waterhouse and Jantzen)}
+I have quite a bit of information to process before this, so I will get started!
+\begin{defn}[Closed Embedding]
+	If $G$ and $H$ are affine group schemes represented by $A$ and $B$, respectively and
+	if $\psi:H\to G$ is a homomorphism of affine group schemes (locally a group homomorphism)
+	then if the corresponding algebra map $A\to B$ is surjective, then $\psi$ is called
+	a \textbf{closed embedding.}
+\end{defn}
+
+As its name suggests, this means that $\psi$ is an isomorphism onto a \textbf{closed subgroup}
+$H'$ of $G$. This is, in fact, a definition of this property. One can also think about 
+it in the following ways: a group scheme $H$ is closed in $G$ if
+\begin{itemize}
+	\item $H$ is defined by the relations imposed by $G$ plus some additional ones.
+	\item $H=V(I)$ for some ideal $I\subset k[G]$. 
+\end{itemize}
+Thinking back to our algebraic geometry, these are not too hard to see as equivalent. 
+For instance, there is a closed embedding of $\mun n$ in $\Gm$ (simply adding in $x^n-1$)
+and of $\SL{n}$ in $\GL{n}$ (adding $\det = 1$).
+
+\subsection{Hopf Ideals}
+One problem with the above characterization that one cannot choose $I$ arbitrarily and end
+up with a group scheme. This is equivalent to arbitrarily adding relations to a group, which is
+not always guaranteed to work out well (think adding $\det = 2$ to $\GL n$).
+
+Actually, we can exactly categorize the closed embeddings of subgroups in $G$ by considering certain
+ideals of the algebra $A$ which represents it. 
+\begin{defn}
+	Let $A$ be an algebra and $I\lhd A$. Then if 
+	\begin{itemize}
+		\item $\Delta(I)$ goes to zero under the map $A\otimes A\to A/I\otimes A/I$,
+		\item $S(I)\subseteq I$
+		\item $\varepsilon(I)=0$
+	\end{itemize}
+	then $I$ is called a \textbf{Hopf Ideal of $A$.}
+\end{defn}
+
+\begin{rmk}
+	That these ideals exactly characterize closed subgroups follows since any group (scheme) represented
+	by $A'$ has the property that $\Delta'(A')\subseteq A'\otimes A'$. Since the new comultiplication map
+	$\Delta'$ is derived from the old one $\Delta$, we have the diagram
+	\begin{figure}[h]
+		\begin{tikzcd}
+			A \ar[r,"\Delta"]\ar[d] & A\otimes A\ar[d]\\
+			A/I\ar[r,"\Delta'"] & A/I\otimes A/I
+		\end{tikzcd}
+	\end{figure}
+
+	\noindent from which we see that $\Delta(I)$ must go to zero in the map on the right.
+	Similarly, we want that $S$ and $\varepsilon$ satisfy similar diagrams.
+
+	Note that an example of such an ideal is $I=\ker(\varepsilon)$.
+\end{rmk}
+
+\begin{defn}
+	A \textbf{character} is a scheme morphism $G\to \Gm$.
+\end{defn}
+
+Let $\Phi$ be a character of $G$ and notice that the corresponding hopf algebra map 
+$\varphi:k[X,1/X]\to A$ is defined by $\varphi(X)=b$, which
+is automatically invertible in $A$. Furthermore since for any $a,b\in G(R)$ we have
+\[\Phi\circ m(a,b)=\Phi(ab)=\Phi(a)\Phi(b)=m\circ(\Phi,\Phi)(a,b)\]
+we get the diagram
+\begin{figure}[h]
+	\begin{tikzcd}
+		G\times G\ar[r,"\Phi\times\Phi"]\ar[d,"m"] & \Gm\times \Gm\ar[d,"m"]\\
+		G\ar[r,"\Phi"] & \Gm
+	\end{tikzcd}
+\end{figure}
+
+\noindent and then by dualizing we get that
+\newpage
+\begin{figure}[h]
+	\begin{tikzcd}
+		A\otimes A & k[X,X^{-1}]\otimes k[X,X^{-1}]\ar[l,"\varphi\otimes\varphi",swap]\\
+		A\ar[u,"\Delta"] & k[X,X^{-1}]\ar[u,swap,"\Delta"]\ar[l,"\varphi"]
+	\end{tikzcd}
+\end{figure}
+
+\noindent where in both diagrams we are abusing notation by using the same symbols for 
+(co)multiplication in the different schemes. So we conclude that 
+\[\Delta(b)=\Delta\circ\varphi(X)=(\varphi\otimes\varphi)\circ \Delta(X)=(\varphi\otimes\varphi)(X\otimes X)=b\otimes b.\]
+
+Then we can easily compute that $\varepsilon(b)=1$ and $S(b)=b^{-1}$. This is precisely:
+\begin{defn}
+	Let $A$ be a Hopf algebra and $a\in A$ such that
+	\begin{itemize}
+		\item $a$ is invertible
+		\item $\Delta(a)=a\otimes a$
+		\item $\varepsilon(a)=1$
+		\item $S(a)=a^{-1}$
+	\end{itemize}
+	then $a$ is called a \textbf{group-like} element of $A$.
+\end{defn}
+\begin{rmk}
+	As we saw above, every character of an affine group scheme $G$ corresponds to a 
+	group-like element of its representing algebra.
+\end{rmk}
+\begin{rmk}
+	Using a parallel construction with any morphism $\Phi:G\to\Ga$, we get an element $b\in A$
+	such that $\Delta(b)=b\otimes 1+1\otimes b$, $\varepsilon(b)=0$ and $S(b)=-b$. These elements
+	are called \textbf{primitive}
+\end{rmk}
+
+\begin{defn}
+	A group scheme $G$ represented by $A$ that consists entirely of group-like elements is
+	called \textbf{diagonalizable.}
+\end{defn}
+\begin{rmk}\label{rmk-diag}
+	Notice that an alternative definition is to begin with an Abelian group $M$ and to 
+	define $S,\Delta,$ and $\varepsilon$ such that each element is group-like. The resulting
+	Hopf algebra represents a diagonalizable group scheme.
+\end{rmk}
+
+Consider the group algebra $k[M]$ on the group in Remark \ref{rmk-diag}. If this algebra
+is finitely generated, we get the following:
+\begin{thm}
+	Let $G$ be diagonalizable and represented by $A$ and assume $A$ is finitely generated
+	as a $k$-algebra. Then $G$ is a finite product of copies of $\Gm$ and $\mun n$.
+\end{thm}
+\begin{prf}
+	Let $x_1,\dots, x_n$ be generators for $k[M]=A$. Each $x_i$ can be written as a (finite!)
+	$k$-linear combination of elements in $M$, so we can instead use the finitely many $m_i$
+	which generate these generators, and this gives us a $k$-algebra basis for $k[M]$. Call this
+	new generating set $U$.
+
+	Let $M'$ be the abelian group generated by $U$ and notice that $k[M']$ is a subalgebra of
+	$k[M]$ containing $U$ so $k[M']=k[M]$ and therefore $M'=M$. This establishes that $M$ is 
+	a finitely generated abelian group, so we can split up the algebra into a tensor product
+	of $k[\bbZ]$ and $k[\bbZ/n\bbZ]$.
+
+	When $M=\bbZ$, $k[M]\cong k[X,X^{-1}]$, so $G= \Gm$\footnote{Notice that here Waterhouse also
+	uses that the basis elements $e_i$ are group-like and so $\Delta(e_i)=e_i\otimes e_i$. I am not
+	quite sure why this is necessary.} Similarly if $M=\bbZ/n$ then $G\cong \mun{n}$
+\end{prf}
+
+Looking at my progress over the last several days, I am thinking that I am getting
+too much in the weeds writing out all the lemmas and proofs. Perhaps it is better to read more quickly
+and only take note of theorems as I need/use them.
+
+\subsection{Cartier Duals}
+The gist here is that if $G$ is represented by $A$, then we can define the dual 
+$A^D=\Hom(A,k)$ and then $A^D$ represents a new scheme $G^D$ called the \textbf{Cartier Dual.}
+
+We do some work to show that, in fact, elements of $G^D$ are in correspondence with the 
+group-like elements of $A$, or equivalently the character group of $G$, $X_G$.
+
+It is easy enough to evaluate $G^D(k)\cong\Hom(A^D,k)$, but luckily one shows that
+``dualizing'' commutes with tensor products (and thus with extension of scalars) and $\Hom$. Thus
+if $G_R$ is the scheme represented by $A\otimes_k R$, then using that $\Hom_k(A^D,R)\cong\Hom_R(A^D\otimes R,R)$,
+\[G^D(R)=(G^D)_R(R)\]
+which is represented by $(A\otimes R)^D=A^D\otimes R$, so finally 
+\[G^D(R)=(G_R)^D(R)=\{\text{group-like elements in }A\otimes_k R\}\cong \Hom(G_R,(\Gm)_R)\]
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\newpage
+\section{Algebraic Geometry}
+
+\newpage
+\section{Problems}
+
+\subsection{Waterhouse}
+\begin{wprob}{2.1}
+	\begin{enumerate}
+		\item Show that there are no nontrivial homomorphisms from $\Gm$ to $\Ga$.
+		\item If $k$ is reduced, show that there are no nontrivial homomorphisms from $\Ga$ to $\Gm$.
+		\item For each nonzero $b\in k$ with $b^2=0$, find a nontrivial homomorphism $\Ga\to\Gm$.
+	\end{enumerate}
+\end{wprob}
+Hmm, there must be a small bug in my macro but I can't find it. I get an error about \texttt{missing `\textbackslash item'}
+in the definition below. It seems to only occur when I use \texttt{\textbackslash subsubsection*} and the like.
+
+\begin{sol}
+	\subsubsection*{(a)}
+	Let $\Phi:\Gm\to\Ga$ and let $\varphi:k[X]\to k[X,X^{-1}]$ be the corresponding Hopf
+	algebra map. But then we have that 
+	\[\varphi\otimes\varphi \circ \Delta_a(X)=\Delta_m\circ\varphi(X)\]
+	and so using that $\Delta_m(a)=a\otimes a$ and $\Delta_a(b)=b\otimes1+1\otimes b$, we get
+	\[\varphi(X)\otimes\varphi(1)+\varphi(1)\otimes\varphi(X)=\varphi(X)\otimes\varphi(X)\]
+\end{sol}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\part{Winter 2019}
-\section{Preparation for the Quarter}
-This is my first official quarter as Julia's student! My plan for now is to continue working on Waterhouse as well
-as learn about algebraic geometry (alongside my usual classes, of course).
-To give a sense of direction, Julia recommended that I take a look at the following regularity theorem:
-\begin{thm}[Smoothness Theorem]\label{smooth}
-	Let $G$ be an algebraic affine group scheme over a field $k$. Then $k[G]\otimes \bar k$ is reduced if and
-	only if $\dim G = \rank \Omega_{k[G]}$.
-\end{thm}
-
-\section{Overarching Ideas and Notes}
-\subsection{Week 2}
-Consider the representations of A over a field $k$ (or more generally $A$ modules, which we can make into a tensor category).
-Recall that in this category we have enough projectives and (with finitely generated Hopf algebras) we get that all projectives are injective.
-
-The object of study here:
-\begin{defn}
-	The \textbf{Hochschild cohomology} of a Hopf algebra $A$ over the field $k$ is 
-	\[\Ext_A^*(k,k)=H^*_A(A,k)\].
-\end{defn}
-
-The conjecture here is
-\begin{conj}
-	when $A$ is a finite dimensional Hopf algebra, $H^*(A,k)$ (the Hochschild cohomology) 
-	is finitely generated as a $k$-algebra.
-\end{conj}
-This is known in some special cases:
-\begin{itemize}
-	\item Finite groups
-	\item Finite group schemes (in positive characteristic) due to Friedlander and Suslin.
-	\item In characteristic zero we have
-	\begin{itemize}
-		\item Quantum groups
-		\item Hopf algebras that come from Nichols algebras
-	\end{itemize}
-\end{itemize}
-
-Papers to look at:
-\begin{itemize}
-	\item Ginzburg and Kumar '94 ``Cohomology of Quantum Groups at Roots of Unity'' \cite{ginzburg-kumar}
-	\item Mastnak, Pevtsova, Shauenburg, and Witherspoon '09 ``Cohomology of Finite Dimensional Pointed Hopf Algebras''\cite{julia10-pointed}
-	\item Witherspoon '17 ``Varieties for Modules of Finite Dimensional Hopf Algebras''\cite{witherspoon-expository}
-\end{itemize}.
-
-Sarah Witherspoon has a book she is writing on Hochschild cohomology that could be very good. It's on her website (Texas A\& M). Check out chapter 9 and appendix A.
-
-\subsection{Week 3}
-We primarily discussed support varieties in the context of Sarah Witherspoon's \textit{Varieties for Modules of Finite Dimensional Hopf Algebras.}\cite{witherspoon-expository}
-
-\subsubsection{Questions}
-\begin{enumerate}
-	\item On page 5, in the definition of the support variety of a Hopf module $M$ (see Definition~\ref{def-support-var}),
-	the construction seems to depend rather heavily on the choice (and existence!) of a graded subalgebra $H\le \HH^*(A)$
-	satisfying (fg1) and (fg2). But the notation seems to imply there is no dependence there. What's going on?
-	\item On page 9, why is $\mathcal{Z}(kG)=\HH^0(kG)$ a local ring (for $G$ a finite group)?
-	\item Directly after the last question, why would this ring being local imply that, on the variety level,
-	we can use $H^{ev}(G,k)$ instead of $H^{ev}(G,k)\cdot \HH^0(kG)$?
-\end{enumerate}
-
-\subsubsection{Answers/Hints}
-\begin{enumerate}
-	\item Actually the definition of support variety, when properly generalized to triangulated tensor categories,
-	doesn't rely on the choice of this algebra. But to see this, we need to dive deeper into that subject. Julia gave 
-	me a book by Dave Benson, Srikanth Iyengar, and Henning Krause on local cohomology and support. It was written
-	as a compendium from a course at Oberwolfach that they gave. It is available on the Arxiv.
-	\item Julia suggests that this is a ``standard'' result that one might find in Webb's book (or another book where the author has taken care to be thorough).
-	I couldn't locate it there immediately, but I found a page online that gave me the maximal ideal, so I am going to try to prove it as lemma~\ref{lem-local}.
-	\item The take-away here was that when computing the varieties the nilpotent elements are irrelevant. Since 
-	the only maximal ideal in $\HH^0(kG)$ is composed of nilpotents, it doesn't contribute to the support variety.
-\end{enumerate}
-
-\begin{defn}[Support Variety]\label{def-support-var}
-	Let $A$ be a finite dimensional algebra and assume $_AA$ is injective (notice that finite dimensional Hopf algebras are
-	Frobenius whence self-injective). Assume further that there is a graded subalgebra $H\le \HH^*(A)$ that satisfies the following 
-	finiteness conditions:
-	\begin{itemize}
-		\item \textbf{(fg1):} $H$ is finitely generated, commutative, and $H^0:=H\cap \HH^0(A)=\HH^0(A)$;
-		\item \textbf{(fg2):} for all finite dimensional $A$-modules $M$, $\Ext^*_A(M,M)$ is finitely generated as an $H$-module.
-	\end{itemize}
-
-	Then if $I_A(M)$ is the annihilator of $\Ext_A^*(M,M)$ in $H$, the \textbf{support variety} is
-	\[V_A(M)=\operatorname{MaxSpec(H/I_A(M))}\]
-\end{defn}
-\begin{rmk}
-	In taking with Julia about this definition she mentioned that the use of the max spectrum instead of the entire
-	spectrum is a bit watered-down and, while simpler, perhaps hearkens back to the beginnings of AG where that is what one considered.
-\end{rmk}
-
-\begin{lem}\label{lem-local}
-	If $G$ is a finite $p$-group and $k$ a field of characteristic $p$, $\mathcal{Z}(kG)$ is local.
-\end{lem}
-\begin{prf}
-	I claim that the maximal ideal is
-	\[I=\left\{\sum_{g\in G}a_gg:\sum_ga_g=0\right\}\]
-	To see this, let $A=\mathcal{Z}(kG)$, and notice that $A=\ker\varphi$ where $\varphi(\sum a_g g)=\sum a_i$.
-	Since $\varphi(kG)=k$, this gives us that $I$ is a maximal ideal in $A$.
-
-	Notice that if we prove that $I$ consists entirely of nilpotents, we will be done. This is because
-	the nilradical of $A$ is the intersection of its prime ideals. But if $I$ is contained in the nilradical, 
-	it must, in particular, be contained in every maximal ideal of $A$. Thus it must be the only maximal ideal.
-
-	To see this, we proceed by induction on $n$ where $|G|=p^n$. When $n=0$ this 
-	is obvious and when $n=1$, this means $G=\bbZ/p\bbZ=\langle\alpha\rangle$. In this case each element is nilpotent of 
-	degree $p$ due to the fact that (since in this case our ring is commutative):
-	\[\left(\sum_1^p a_i\alpha^i\right)^p=\sum_1^p a_i^p\alpha^{pi}=\sum_1^p a_i1_G=0.\]
-
-	Now assume that $I$ consists of nilpotents for all $n\le k$ and let $|G|=p^{k+1}$. Let $Z\le G$ be the
-	(necessarily nontrivial) center of $G$. Extend the canonical quotient map $q:G\to G/Z$ to a map 
-	between group rings $\tilde q:kG\to k(G/Z)$. Let $I_Z$ be the augmentation ideal of $k(Z)$. Then
-	$\tilde q$ is surjective with kernel $I_Z\cdot kG$. That this ideal is contained within the kernel is clear
-	since if $y$ is a $kG$-linear sum of elements of the form $\sum_Z a_zz$ where $\sum_Z a_z=0$, then
-	the image is just the sum of zeros.
-
-	To see that every element of the kernel is of this form, let $y=\sum k_i g_i$ and notice
-	\[\tilde q\left(\sum k_ig_i\right)=\sum k_i\tilde q(g_i)=\sum_j\left(\sum_{g_i-g_j\in Z}k_i\right)g_j\]
-	and this is zero precisely when all the sums are. But then the sum of the coefficients of $y$ restricted to any coset of $Z$
-	are zero.
-
-	Thus in particular $kG/(I_Z)\cong k(G/Z)$.
-
-	Then look at the augmentation ideal $I_{G/Z}\lhd k(G/Z)$ which, by the induction hypothesis, consists of nilpotents.
-	Pull this back through $\tilde q$ to the ideal $I=(I_{G/Z})+ (I_Z)$. Since some power of $p$ kills all elements of the two 
-	summands, this implies that some power of $p$ kills each element in $I$, completing the proof.
-\end{prf}
-
-\subsection{Week 4}
-Recall that $\mathcal{U}:\Lie\to \Ass$ is a functor that assigns to each 
-Lie algebra the universal enveloping algebra: the tensor algebra over $\g$ modulo the relations
-$x\otimes y-y\otimes x-[x,y]$.
-
-Then we have the PBW theorem:
-\begin{thm}[Poincar\'e-Birkhoff-Witt]
-	If $\g$ is a finite dimensional Lie algebra, with basis $x_1,\dots,x_n$, then $\mathcal{U}(\g)$ has a basis
-	in the form $\prod x_i^{a_i}$ for $a_i\in\bbZ_{\ge 0}$.
-\end{thm}
-\begin{rmk}
-	A fancier way of saying this is to say that $\mathcal{U}(\g)$ admits a filtration such 
-	that $\operatorname{gr}\mathcal{U}(\g)\cong S^*(\g)$.
-\end{rmk}
-
-Furthermore, there is a more general class of algebras called \textbf{PBW algebras} where you get commutation
-modulo lower-order terms. This idea is also what forms the basis of the proof for the above remark.
-
-The idea here, then, is to take the center and see how it acts on the universal enveloping algebra.
-
-\subsubsection{The BGG category $\mathcal{O}$}
-The representations of finite dimensional Lie algebras are in correspondence with the dominant integral weights. This
-will be worth looking into, perhaps.
-
-The general question we want to answer is what he structure of finite dimensional representations of $\g$. In the case
-when $\g=\sl_2$, we have the representation $V=ke_1\oplus ke_2$.
-
-Here I need to work out some of the notation, but $k[x,y]$ is a representation and $S^d(V)=k_d[x,y]$ is as well. (F\&H lecture 11 for more details there).
-
-So we end up realizing that we can generate any such Lie algebra by taking a single highest-weight vector and 
-applying $\mathcal{U}(\mathfrak{n}^-)$ to get the whole thing. This help motivate:
-\begin{defn}[BGG Category $\mathcal{O}$]
-	$\mathcal{O}$ is the full subcategory of $\mathcal{U}(\g)$ modules such that
-	\begin{itemize}
-		\item $M$ is finitely generated over $\mathcal{U}(\g)$ -- this is necessary since we want free modules
-		and $\mathcal{U}(\g)$ itself to be representations under study and these are excluded from finite ($k$-) dimensional algebras.
-		\item $M$ is $\mathfrak{h}$ semisimple.
-		\item $M$ is $\n$-locally finite -- for all $v\in M$, $\mathcal{U}(\n)v$ is finite dimensional.
-	\end{itemize}
-\end{defn}
-
-Ideas for things to understand/cover for next week:
-\begin{itemize}
-	\item First things first, understand the representations of $\sl(2,k)$.
-	\item Then in Humphreys(I), read/do section \S 6.21 to understand the universal enveloping algebra.
-	\item Read Humphreys(II) chapter 1 as much as possible.
-\end{itemize}
-
-\subsection{Week 5}
-We spent most of the time deriving the fact that any simple $\sl_2$ module is isomorphic to one of the ``standard''
-highest weight modules.
-
-Our assignment for this upcoming week:
-Write down the Serre relations. Then do some computations and read section 21 in Humphreys. Read those lemmas and theorems and start to describe decomposition of general modules of Lie algebras.
-
-\subsection{Week 6}
-We skipped Lie algebras this week because of snow (and then Ale couldn't make our meeting). So instead
-Julia just met up to talk about my reading.
-
-My primary question (regarding the Ginzburg and Kumar paper): why are we factoring out the toral elements in quantum groups? 
-The answer here is that these are called the ``small quantum groups''. So certainly having a finite-dimensional algebra makes certain things
-more tractable, although when computing cohomology that isn't always the case. Sometimes when you get smaller 
-objects you end up getting more nontrivial elements in the cohomology. There's a clear analog to topology here.
-
-So then maybe a refinement of the question is this: why is it necessary at all to have this larger object serving as the ambient 
-space for the object we truly want to study? Here the answer more-or-less boils down to the fact that it end up playing a central
-role in computations involving these objects.
-
-Having the big algebra $U$ over $u$ (modulo something central) gives you that thePBW basis for $U$ descends to 
-$u$. In general (e.g. Nichols algebras) you fail to have a PBW basis for the cover, but somehow (for a technical reason that 
-will become apparent as we read more) you still can use similar techniques to do the computations (the more general theory arises in the 2010 paper).
-
-Julia is going to sent me a paper that is a survey of finite dimensional algebras that will give some sense for how the small quantum groups fit in.
-
-\subsubsection{This week's assignment}
-Continue reading through GK until we get to the spectral sequence nonsense. Then this points to Jantzen (AG). Try to recreate the
-argument for constructing the first page of the spectral sequence, but for the GZ case (note, you'll use projectives instead of injectives)
-After that, go to Weibel and try to read the section of spectral sequences over filtered complexes \S 5.4 (what we are usually interested in).
-Don't worry too much about issues of convergence -- we are mostly interested in the ``first quadrant'' case.
-
-David Anick is the name of an algebraic topologist who started off created (free) resolutions that we will end up looking at. 
-His paper \cite{anick-resolution} will probably be useful at some point.
-
-\subsection{Week 8}
-The last paper on Julia's page is a preprint with her and Dave Benson. There is a spectral sequence computation that might be illustrative.
-There is also an older paper...
-
-One of the things Julia is interested in at the moment is a super group scheme. These arise by having a $\bbZ_2$ grading on your Hopf algebra.
-For instance we can look at exterior algebra $\Lambda(x)$, which represents the super group scheme $G_a^-$.
-
-The first non-elementary example is $G_a^-\times G_a^-$ with a $\bbZ_p$ action on it that ``folds'' the second factor
-onto the first. To write this out explicitly, we consider the group algebra $\Lambda(u,v)$ where $uv=-vu$ and the coordinate algebra of $\bbZ_p$
-is $k[t]/t^p$ where $t\cdot u=0$ and $t\cdot v=\alpha u$ for some $\alpha$. This is a $4p$ dimensional algebra.
-
-Then we want to compute coholomology. We know $H^{**}(k\bbZ/p,k)\cong k[x]\oplus\Lambda(y)$ for $|x|=2,0$.
-Note there is a notion of dimension that is an ordered pair $(a,b)$ of the graded degree and cohomology degree.
-THis give us new elements of degree $(1,1)$
-
-Look up Koszul duality! It is in the representations and local cohomology book!
-
-\subsection{Week 10}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\newpage
-\section{Lie Algebras}
-We're planning on digging into the category $\O$, the category of (reasonably finite) $\mathcal{U}(\g)$ modules.
-To begin, however, we are going to make sure we have our fundamentals vis a vis (semi) simple representations of Lie algebras.
-
-\subsection{$\sl_2$}
-In the case when $L=\sl_2$, everything is relatively nice. We read (and discussed) the fact that any simple, finite-dimensional
-$\sl_2$-module is isomorphic to $V(\lambda)$ for some integer $\lambda$ (highest weight).
-
-The action of $L$ on these modules are precisely as one would hope: $x$ and $y$ (the off-diagonal generators) act by raising 
-and lowering the weight of a vector by 2 and the central generator fixes each eigenspace.
-
-\subsection{The Serre Relations}
-Let $L$ be a semisimple Lie algebra and $\h$ a Cartan subalgebra with corresponding roots $\Phi$ with simple root system $\Delta$.
-Recall that 
-\[\langle\alpha_i,\alpha_j\rangle=\frac{2(\alpha_i,\alpha_j)}{(\alpha_j,\alpha_j)}=\alpha_i(h_{\alpha_j})\]
-Fix generators $x_i$ and $y_i$ of $L_{\alpha_i}$ and $L_{-\alpha_i}$, respectively, so that $[x_i,y_i]=h_i=h_{\alpha_i}$
-Recall that the $x_i$ and $y_i$ generate $L$ (as a Lie algebra).
-
-Then in particular we have the relations:
-\begin{align*}
-	[h_i,h_j]&=0\tag{S1}\\
-	[x_i,y_j]&=\delta_{ij}h_i\tag{S2}\\
-	[h_i,x_j]=\langle\alpha_j,\alpha_i\rangle x_j\text{ and }& [h_i,y_j]=-\langle\alpha_j,\alpha_i\rangle y_j\tag{S3}
-\end{align*}
-\begin{defn}[Serre Relations]
-	With the notation above, the relations S1-3 above are called the \textbf{Serre relations}.
-\end{defn}
-
-These relations end up being crucial in making the connection between root systems and semisimple Lie algebras concrete:
-taking the abstract free Lie algebra on generators satisfying these properties, we get a finite dimensional semisimple Lie algebra
-with roots corresponding exactly to the $\alpha_i$. This also gets us uniqueness of such a root system.
-
-\subsection{The more general case.}
-This primarily borrows from \S 21 in \cite{humphreys1} but also relies on some results from earlier that I will cite as 
-necessary.
-
-Throughout this discussion, we rely on the theory developed in sections 15 and 16 regarding Cartan and Borel subalgebras:
-\begin{defn}[Cartan Subalgebra (CSA)]
-	If $L$ is a Lie algebra, then $C\subseteq L$ is called a \textbf{Cartan subalgebra} if $C$ is
-	\begin{enumerate}
-		\item Nilpotent; and
-		\item $N_L(C)=C$.
-	\end{enumerate}
-\end{defn}
-\begin{defn}[Borel Subalgebra]
-	If $L$ is a Lie algebra, $B\subseteq L$ is called a \textbf{Borel subalgebra} if $B$ is a
-	maximal solvable subalgebra.
-\end{defn}
-\begin{rmk}
-	Since every nilpotent Lie algebra is solvable, each CSA lies in some Borel subalgebra $B$ of $L$.
-	then by showing that any two Borel subalgebras are conjugate under $\mathscr{E}(L)$ (the subalgebra of $\Int L$ generated
-	by the strongly $\ad$-nilpotent elements of $L$), it suffices to 
-	show that two CSAs are $\mathscr{E}(L)$-conjugate in any \textit{semisimple} Lie algebra.
-\end{rmk}
-
-In all that follows, we've fixed a CSA $\h$ of $L$ and root system $\Phi$ with simple root system $\Delta=\alpha_1,\dots,\alpha_l$
-of positive roots with $\mathcal{W}$ the Weyl group.
-
-\subsection{Representations}
-Since $\h$ is semisimple (since $L$ is), it acts diagonally on any finite dimensional $L$-module. Thus we can always
-do the same trick of decomposing $V$ into eigenspaces $V=\sqcup_{\lambda\in \h^*}V_\lambda$ that we do for $\sl_2$.
-Slightly more generally (when $V$ is infinite-dimensional), we can consider the sum of weight spaces $V_\lambda$ of $V$, 
-which is necessarily  direct\footnote{Think: vectors can't have multiple eigenvalues.}, so there always
-exists a direct sum of simple modules contained in $V$. When $\dim V<\infty$, this is equal to $V$ itself.
-
-\subsubsection{Standard Cyclic Modules}
-Here
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\newpage
-\section{Quantum Groups}\label{sec-quantum}
-I am mostly working out of Jantzen's \textit{Lectures on Quantum Groups} (\cite{janzten-quantum}) with the goal of reading through Ginzburg \& Kumar's
-(apparently influential!) paper \textit{Cohomology of Quantum Groups at Roots of Unity} (\cite{ginzburg-kumar}).
-
-\subsection{What they are}
-The core idea to keep in mind is that quantum groups (or quantum enveloping algebras) are ``deformations'' (or perhaps $q$-analogues) of
-regular enveloping algebras. Well, at least that is the kind we're mostly working with here. A general definition eludes me (and Jantzen).
-
-Mostly we will be interested in $U=\mathcal{U}_q(\sl_2)$. In this setting, we get the definition
-\begin{defn}[Quantum Enveloping Algebra of $\sl_2$]
-	$\mathscr{U}_q(\sl_2)$ is the associative algebra $k\langle E,F,K,K^{-1}\rangle$ under the following relations:
-	\begin{align*}
-		KK^{-1}&=K^{-1}K=1\tag{R1}\\
-		KEK^{-1} &= q^2 E\tag{R2}\\
-		KFK^{-1} &= q^{-2}F\tag{R3}\\
-		EF-FE &= \frac{K-K^{-1}}{q-q^{-1}}\tag{R4}
-	\end{align*}
-	where $q$ is some nonzero element of $k$ such that $q^2\ne 1$.
-\end{defn}
-
-Some notation that may be helpful in the future:
-\begin{itemize}
-	\item For $a\in\bbZ$, define $[a]=\frac{v^a-v^{-a}}{v-v^{-1}}\in\bbQ(v)$ -- the $v$ analogue of $a$. Equivalently, (and perhaps suggestively):
-	\[[a]=v^{a-1}+v^{a-3}+\cdots+v^{-a+3}+v^{-a+1}.\]
-	\item the $v$ analogue of the factorial is
-	\[[a]^!=[a][a-1]\cdots[1]\]
-	\item Then you get the binomial coefficients:
-	\[\qbinom{a}{n}=\frac{[a]^!}{[n]^![a-n]^!}\]
-	\item \[[K;a]=\frac{Kq^a-K^{-1}q^{-a}}{q-q^{-1}}\]
-\end{itemize}
-
-\subsection{Getting a basis}
-The first part of the first chapter focuses on getting a PBW-style basis for $U$, which comes in the form $F^s,K^n,E^r$ for $r,s\in\bbN$, $n\in\bbZ$.
-This is mostly standard but I thought this proof had a neat idea in it:
-
-\begin{prf}
-	Let $A=k[X,Y,Z]$ and define $f,e,h\in \Hom_k(A,A)$ such that 
-	\begin{align*}
-		f(Y^sZ^nX^r)&=Y^{s+1}Z^nX^r,\\
-		e(Y^sZ^nX^r)&=q^{-2n}Y^sZ^nX^{r+1}+[s]Y^{s-1}[Z;1-s]Z^n X^r,\\
-		h(Y^sZ^nX^r)&=q^{-2s}Y^sZ^{n+1}X^r
-	\end{align*}
-	One can check that these vector space homomorphisms exactly mirror multiplication of these monomials
-	by $F, E,$ and $H$, respectively, so we get a map $\varphi:U\to\Hom_k(A,A)$. Then take any linear combination of 
-	$F^sK^nE^r$, mapped though to $A$ via $\varphi$ and then evaluation at 1, gives us a linear combination of the $Y^sZ^nX^r$,
-	so it must be trivial. Sick.
-\end{prf}
-
-\brk
-
-After that, the book focuses on $U_0=k\langle Z,Z^{-1}\rangle$ as well as $U^+=k\langle E\rangle$ and $U^-=k\langle F\rangle$,
-the latter of two which are isomorphic to $k[x]$ by the linear independence above.
-
-\brk
-
-Here's another cool bit: By studying products of monomials (or just defining degree outright) we can impose a graded structure on $U$:
-let $\deg(E)=1$, $\deg(F)=-1$ and $\deg(K)=\deg(K^{-1})=0$ and notice (R1-4) are homogeneous according to this degree.
-Thus the quotient by this homogeneous ideal gives a graded structure to $U$.
-
-Furthermore if $u\in U$ is homogeneous of degree $i$, 
-\[KuK^{-1}=q^{2i}u\]
-so when the powers of $q$ are distinct ($|q|\ne 1$), the graded pieces are precisely the 
- of $K$ under the adjoint action.
-
-\subsection{Representations of $U$}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\newpage
-\section{Reading Ginzburg \& Kumar}
-This section contains my notes and thoughts about \cite{ginzburg-kumar}. As a high-level summary, this paper establishes
-methods for computing (Hochschild) cohomology of the small quantum groups. It begins by establishing how they are constructed
-as quotients or subalgebras of quantized universal enveloping algebras (see section~\ref{sec-quantum}), and then proceeds to compute
-the cohomology using a filtration (this is vital!) and spectral sequences.
-
-This paper is important because it ends up that this general formula is the same as is applied later in \cite{julia10-pointed} to prove
-a much more general result. This also extends a method originally created by \cite{FP} for computing the cohomology of Lie algebras and
-algebraic groups.
-
-\subsection{The Main Result}
-In this paper they prove the following (see \ref{sec-defs} for notation):
-\begin{thm}[GK '93]
-	If $k=\bbQ(\xi)$ for a primitive root of unity of odd order (not 1) and $u_\xi$ the restricted enveloping algebra of some $\g$. Then
-	\[H^{odd}(u_\xi,k)=0\]
-	and there is a natural graded algebra isomorphism
-	\[H^{ev}(u_\xi,k)\cong k^\bullet[\mathcal N].\]
-\end{thm}
-
-\subsection{The Objects in Question}
-We begin by considering the objects in question and how they are constructed. One thing we haven't seen (yet!) in Lie algebras
-but comes up in the structure theory is the 
-\begin{defn}[Kostant $\bbZ$-form]\label{def-kostant}
-	Let $U(\g)$ be the universal enveloping algebra of $\g$, spanned (over $k$) by the $x_\alpha,y\alpha,$ and $h_\alpha$ for all $\alpha\in\Phi$. 
-	Then the \textbf{Kostant $\bbZ$-form} is the $\bbZ$-span of the elements
-	\[X^{(r)}_\alpha=\frac{x_\alpha^r}{r!},\quad Y^{(r)}_\alpha=\frac{y_\alpha^r}{r!},\quad H^{(r)}_\alpha=\frac{h_\alpha^r}{r!},\quad \]
-	the so-called \textbf{divided powers} of $x_\alpha, y_\alpha,$ and $h_\alpha$. This ($\bbZ$-)algebra is denoted $U_\bbZ$.
-\end{defn}
-
-Before we use it, we had better write down the definiton (via \cite{janzten-quantum}) of a
-\begin{defn}[Quantized Enveloping Algebra]
-	Fix some odd root of unity $q$ in the field\footnote{Here Jantzen also requires that $\ch k\ne2$ and that $\ch k\ne3$ if $\Phi$ has an irreducible component of type $G_2$.
-These can be treated, but not in quite the same manner.} $k$.
-	Let $\g$ be a Lie algebra over $k$ and let $U_q(\g)$ denote the $k$ algebra with generators $E_\alpha,F_\alpha, K_\alpha,$ and $K_\alpha^{-1}$
-	for all $\alpha$ in a basis $\Pi$ of $\Phi$, subject to the relations for all $\alpha,\beta\in\Pi$
-	\begin{align*}
-		K_\alpha K_\alpha^{-1}=&1=K^{-1}_\alpha K_\alpha\\
-		K_\alpha K_\beta &= K_\beta K_\alpha\\
-		K_\alpha E_\beta K_\alpha^{-1} &= q^{(\alpha,\beta)}E_\beta\\
-		K_\alpha F_\beta K_\alpha^{-1} &= q^{-(\alpha,\beta)}F_\beta\\
-		E_\alpha F_\beta - F_\beta E_\alpha &= \delta_{\alpha\beta}\frac{K_\alpha-K_\alpha^{-1}}{q_\alpha-q_\alpha^{-1}}\\
-		\sum_{s=0}^{1-a_{\alpha\beta}}(-1)^s\sqbinom{1-a_{\alpha\beta}}{s}_{q_\alpha} &E_\alpha^{1-a_{\alpha\beta}-s}E_\beta E_\alpha^s = 0\\
-		\sum_{s=0}^{1-a_{\alpha\beta}}(-1)^s\sqbinom{1-a_{\alpha\beta}}{s}_{q_\alpha} &F_\alpha^{1-a_{\alpha\beta}-s}F_\beta F_\alpha^s = 0
-	\end{align*}
-	where $q_\alpha=q^{d_\alpha}=q^{(\alpha,\alpha)/2}$ and $a_{\alpha,\beta}=\langle\alpha,\beta\rangle=2(\alpha,\beta)/(\alpha,\alpha)$.
-\end{defn}
-
-\begin{rmk}
-	In \cite{janzten-quantum}, he mentions that he will often work in the algebra that includes all the above relations
-	but \textbf{omitting the last two}. Apparently this will be done for computational simplicity and then studying $U_q(\g)$ 
-	itself will amount to taking a quotient. This may not be relevant to me, but it's good to keep in mind for this reference.
-\end{rmk}
-
-\brk
-
-For the case of quantum groups, we are extending all of our notions to quantized enveloping algebras in the case when our field $k$ is the rational function field
-$\bbQ(v)$.
-\begin{defn}[$q$-analog of the Kostant $\bbZ$-form]
-	Let $U=U_q(\g)$ be the (Hopf!) algebra described above. Then we define the $q$-analog of the Kostant $\bbZ$-form
-	to be the $\bbZ[v,v^{-1}]$-span of the $q$-divided powers the $E_\alpha,F_\alpha$ and $K_\alpha^{\pm 1}$, which are the same as
-	the divided powers in definiton~\ref{def-kostant} except the factorial is replaced by the $q$-analog thereof.
-\end{defn}
-
-\subsubsection{Notation and Related Algebras}\label{sec-defs}
-But this isn't what we're studying yet! (This are the so-called ``big'' quantum groups: what Drinfeld studies in \cite{drinfeld}). 
-We are interested in taking particular quotients and subalgebras. Here we define the notation used throughout and how they are related to $U=U_q(\g)$.
-
-We will be interested primarily in the case when $v=\xi$, that is, when $k$ is the cyclotomic field over $\bbQ$ found by adjoining
-an \textbf{odd} primitive $l^{th}$ root of unity ($l$ not 1). For the rest of this paper, fix such a $k$. Then define
-\[U_\varepsilon :=(k\otimes_{\bbZ[v,v^{-1}]}U_\bbZ)/(K_\alpha^l-1)=:U_k/(K_\alpha^l-1)\]
-where we defined $U_k$ by extension of scalars from $U_\bbZ$ and we are using here the fact that any $\bbZ[\xi]$-module 
-becomes a $\bbZ[v,v^{-1}]$-module by letting $v$ act by $\xi$.
-
-\brk
-
-Another construction is the so called \textbf{restricted enveloping algebra} $u_\xi$ of $U_\xi$ generated by the 
-degree one generators $F_\alpha, E_\alpha$, and $K^{\pm 1}_\alpha$. This is a finite-dimensional algebra of dimension $l^{\dim \g}$.
-This is actually the object that we are studying here. It amounts to restricting coefficients in $U_\xi$ to polynomials in $\xi$ with integer coefficients.
-
-\brk
-
-Define $\mathcal N$ to be the \textbf{nilpotent cone} of $\g$, consisting of all $\ad$-nilpotent elements in $\g$. This admits a 
-natural graded structure over the nonnegative integers by recognizing it as the ring of regular functions on a cone.
-
-\subsubsection{Another formulation}
-
-There is a parallel construction of $u_\xi$ that highlights a ``duality'' property of these objects. Set $\mathscr U_\bbQ$ to be the 
-$\bbQ[v,v^{-1}]$-span of $E_\alpha,F_\alpha,K^{\pm 1}_\alpha$ (notice we are not using divided powers here, but allowing coefficients in $\bbQ$) 
-and extend scalars again to $\mathscr U_k$. Again let $\mathscr U_\xi$ be the quotient
-where we identify $K_\alpha^l=1$. Define $\mathscr L$ to be the subalgebra of $\mathscr U_\xi$ generated by $E_\alpha^l$ and $F_\alpha^l$.
-
-Then it is a result of Lusztig that $\mathscr L$ and $u_\xi$ are normal (algebras) in $\mathscr U_\xi$ and $U_\xi$, respectively,
-and that 
-\[\mathscr U_\xi//\mathscr L\cong u_\xi\qquad\text{and}\qquad U_\xi//u_\xi\cong U(\g)\]
-where
-\begin{defn}
-	Given an augmented algebra $B$ with augmentation ideal $B_+$, a subalgebra $A\subseteq B$ is called \textbf{normal} in $B$ if 
-	\[B\cdot (A\cap B_+)=:B\cdot A_+=A_+\cdot B.\]
-\end{defn}
-
-(My) intuition for augmented algebras come from group algebras $kG$, where you have the augmentation map $\epsilon(\sum_G c_gg)=\sum_G c_g$.
-Here the augmentation ideal is $\ker\varphi$. In fact, any map $\epsilon:A\to k$ you have induces such a structure. Usually in the case of Hopf algebras
-we pick the counit (as we do in group algebras).
-
-\begin{defn}
-	If $A$ is normal in $B$, then we write 
-	\[B//A:=B/(B\cdot A_+)=B/(A_+\cdot B).\]
-\end{defn}
-
-The duality we mentioned earlier comes from the fact that $\mathscr L$ is a commutative Hopf algebra that is ``dual'' to (the cocommutative) $U(\g)$.
-
-\subsubsection{Triangular Decompositions}
-We use superscripts $+,-,$ and $0$ to denote the subalgebras generated, respectively, by the $E$'s, $F$'s, and $K$'s\footnote{with the exception of $U_0$ which has another set of generators due to the divided powers}.
-
-Then for $U_\xi$, $\mathscr U_\xi$, and $u_\xi$ we get the standard triangular decompositions; e.g.
-\[u_\xi = u_\xi^+\oplus u_\xi^0\oplus u_\xi^-.\]
-
-For each, define the \textbf{Borel part} to be (e.g.) $\mathscr B_\xi=\langle K_\alpha, E_\alpha\rangle$ (i.e. the nonnegative part).
-
-\subsection{The Argument}
-\subsubsection{Filtrations}
-The core idea (according to Julia) is the existence of certain filtrations such that the associated graded algebras have nice structure.
-
-In this paper, we are using an earlier result in \cite{de-concini} which tells us that $\mathscr B_\xi$ admits such a filtration:
-(Recall that we are using roots $\alpha$ in some basis $\Pi$ of $\Phi$)
-\begin{thm}[DK '90]
-	The algebra $\mathscr B_\xi$ has a multiplicative filtration such that $\Gr \mathscr B_\xi$ is generated by the homogeneous elements $\{E_\alpha,K_\beta\}_{\alpha\in \Delta_+,\beta\in \Pi}$
-	subject to the relations
-	\[K_iK_j=K_jK_i\quad K_i^l=1\quad K_iE_\alpha=\xi^{\langle \alpha,\alpha_i\rangle} E_\alpha K_i\]
-	and
-	\[E_\alpha E_\beta=\xi^{\langle\alpha,\beta\rangle}E_\beta E_\alpha\]
-	if $\alpha\succ\beta$.
-\end{thm}
-Then we can use this filtration to get filtrations on all quotients and subalgebras, most importantly
-$\mathscr B_\xi,\mathscr U_\xi^+, b_\xi,$ and $u_\xi^+$.
-\begin{qst}
-	What is the filtration here? 
-\end{qst}
-I ended up finding the book in which \cite{de-concini} is published (in the library) and the answer is that you look at 
-monomials 
-\[M_{k,r;u}=F^kuE^r\]
-where $F^k=\prod_\alpha F_\alpha^{k_{\alpha}}$ and similar for $E^r$ and $u\in\mathcal U^0$.
-
-Then you can define the \textbf{height:}
-\[\operatorname{ht}(M_{k,r;u})=\prod_\alpha (k_\alpha+r_\alpha)\operatorname{ht}\alpha\in\bbZ_+\]
-(What is $\operatorname{ht}(\alpha)?$) and then finally define \textbf{degree:}
-\[d(M_{k,r;u})=(k_N,k_{N-1},\dots,k_1,r_1,\dots,r_N,\operatorname{ht}(M_{k,r;u}))\in\bbZ_+^{2N+1}\]
-which is totally ordered via the lexicographic ordering and therefore gives us a $\bbZ^{2N+1}$ grading.
-
-\brk
-
-Then we proceed to prove that $H^\bullet(\Gr\mathcal U_\xi^+)\cong\Lambda_\xi$, which follows from some arguments that can be found 
-in \cite{priddy}.
-\begin{qst}
-	What role do augmented algebras play in this theory? I have noted on several occasions that augmentations
-	have been used to define normal subalgebras. It also popped up in the discussion of Koszul algebras in \cite{priddy}.
-\end{qst}
-
-The idea here is to give $k$ the structure of an $A$-module. This is non-trivial for the folloiwng reason:
-consider $A=k[x_i]$ and localize at a non-maximal prime $p$. Then the residue field will have positive transcendence degree
-over $k$, and since $p$ will now be the (only!) maximal ideal, this will be the smallest field to which there is such a map.
-
-\begin{qst}
-	In the computation of cohomology for $\mathcal L^+$, why do the terms in the Koszul complex
-	look like $\mathcal L^+\otimes_k \Lambda^i(\mathfrak N)$? Is this exploiting somehow that 
-	all Hopf modules are trivial (e.g. $\mathcal L^+$) determines the entire action by 
-\end{qst}
-
-Let's take a regular sequence $(x_1,\dots,x_n)$. Then $R\xrightarrow{x_1} R$ is $K(x_1)$, the 
-Koszul complex for this regular subsequence. If we have two elements, then $K(x_1)\otimes K(x_2)$
-is the complex $R\to R\oplus R\to R$. Corollary 4.5.5 in Weibel says that whenever the sequence is regular, if you 
-construct the entire Koszul complex to get a resolution of the quotient $R/(x_i)$.
-
-Note that this essentially boils down to the computation that $\Lambda^iS^n\cong \Lambda^i(k^n\otimes_k S)=(\lambda^i k^n)\otimes S$.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\newpage
-\section{Julia \& Dave Benson's preprint}
-As an aside, in response to me saying I needed to practice doing computations with spectral sequences,
-Julia recommended I read her current work \cite{julia-dave} with Dave Benson wherein there is a supposedly
-straightforward spectral computation in support of their larger work \cite{detecting-nilpotence}.
-
-I have already spent some time reading this paper, but the core objects in question are \textbf{supergroup schemes}
-which have coordinate Hopf algebras algebras with a $\bbZ/2$ grading that are furthermore (graded) cocommutative (just 
-as group schemes correspond to finite dimensional cocommutative Hopf algebras). The fundamental example is $\bbG_a^-$,
-which has as its coordinate ring $k[x]/(x^2)$, or equivalently the exterior algebra on a single element $x$,
-where $x$ is taken to be a primitive element.
-
-Then if we let $A$ be the coordinate algebra of $\bbG_a^-\times\bbG_a^-$ and $H$ that of $\bbG_{a(r)}\times(\bbZ/p)^s$ where $\bbG_{a(r)}$ is the
-$r^{th}$ Frobenius kernel of $\bbG_a$, we can define the \textbf{smash product} or in this case, since 
-$A$ is graded cocommutative, the \textbf{semidirect product} $A\#H=A\rtimes H$ where we define the action
-to be (if $\tau:H\otimes A\to A$ is the action of $H$ on $A$)
-\[(a\otimes h)\cdot(a'\otimes h')=\sum (-1)^{|h_1||b|}a\tau(h_1\otimes b)\otimes h_2g.\]
-
-So we use that $A\cong \Lambda(k[u,v])$, the exterior algebra on two (primitive) generators $u$ and $v$
-and that, if $t_i$ is the generator for the coordinate algebra of the $i^{th}$ copy of $\bbZ/p$, and if 
-we let $g_i=t_i-1$, these along with the $s_1,\dots s_r$ generating the Frobenius kernel, generate $H$.
-
-One of the first propositions (3.1) of the paper computes the cohomology of $\bbG_a^-$ with 
-coefficients in $k$ (actually, refers to another paper) and computes how the ($\bbZ/2$-graded)
-Steenrod operations act on it.
-
-\begin{defn}[Steenrod Operations]
-	If $p$ is odd and $A$ is a $\bbZ$-graded cocommutative Hopf algebra, then there are natural maps/operations
-	\[\mathcal P^i:H^{s,t}(A,k)\to H^{s+(2i-t)(p-1),pt}(A,k)\]
-	\[\beta \mathcal P^i:H^{s,t}(A,k)\to H^{s+1+(2i-t)(p-1),pt}(A,k)\]
-	satisfying the properties
-	\begin{enumerate}
-		\item $\mathcal P^i=0$ whenever $2i<t$ or $2i>s+t$,
-		\item $\beta\mathcal P^i=0$ whenever $2i<t$ or $2i\ge s+t$.
-		\item $\mathcal P^i(x)=x^p$ when $2i=s+t$
-		\item $\mathcal P^j(xy)=\sum_i\mathcal P^i(x)\mathcal P^{j-i}(y)$
-		\item $\beta\mathcal P^j(xy)=\sum_i(\beta\mathcal P^i(x)\mathcal P^{j-i}(y)+\mathcal P^i(x)\beta\mathcal P^{j-i}(y))$
-		\item $\mathcal P$ and $\beta\mathcal P$ satisfy the Adem relations.
-		\item $\mathcal P^0(\lambda u)=\lambda^p\mathcal P^0(u)$ for all $\lambda \in k$ ($\ch k= p$)
-	\end{enumerate}
-\end{defn}
-
-In the course of proving things here, the authors adapt these operations to the $\bbZ/2$ grading
-by reindexing, calling these new operations $\mathscr P_i$ and $\beta\mathscr P^i$ (they use \texttt{mathscr} but I've
-reassigned mine to a font I like better. :P)
-
-Okay, time is running out, so I am pulling together some questions:
-\begin{qst}
-	Is the cohomology you're using here group cohomology? It seems to me the Lyndon-Hochschild-Serre
-	applies to group cohomology.
-
-	If so, I will need to focus some time on getting familiar with it.
-\end{qst}
-
-Yes it is group cohomology, or more specifically group \textit{scheme} cohomology. Here we 
-are computing $\Ext_G^{i,\varepsilon}(k,k)$.
-
-We begin with our group algebra (here $\Lambda(v_1,v_2)\#k\bbG_{a(1)}$) and take 
-a projective resolutions and computing $H^*(\Hom_A(p^\bullet,k))$.
-
-Recall the syzygys, $\Omega^i k$ that occur in a projective resolution of $k$. Then we can compute the cohomology
-via the stable module category:
-\[\underline{\Hom}_G(\Omega^ik,k)=\Ext_G^{i,\varepsilon}(k,k)\]
-
-There is a result by Jeremy Rickard (rep theory) and by Buchweitz (more generally) that proves
-
-\begin{qst}
-	Where is $H^{*,*}(\bbG_a^-,k)$ computed? The papers claim it is $S^{*,*}(V^\sharp)\cong k[\zeta,\eta]$
-	where $\zeta$ and $\eta$ are the dual generators corresponding to the generators $u$ and $v$ of $V$,
-	the augmentation ideal of $k(\bbG_a^-\times \bbG_a^-)$.
-\end{qst}
-
-Again, this is an example of Koszul duality -- use that $k\bbG_a^-=\Lambda(v)$ where $v$ is primitive in degree 1.
-But then $k(\bbG_a^-\times\bbG_a^-)\cong\Lambda(v_1,v_2)$. Then applying Koszul duality you get the 
-symmetric algebra. Also you need the K\"unneth formula, where as algebras
-\[H^*(A\otimes B,k)\cong H^*(A,k)\otimes H^*(A,k).\]
-
-Evans has a book on The groups of cohomology. Here there are computations and resolutions of (cyclic)
-groups. There is also a discussion of spectral sequences which is relatively terse (although it defers
-somewhat to other sources).
-
-\begin{qst}
-	What is the importance of suspensions and what role do they play in this theory? They 
-	have come up at several points and I am realizing they are fairly pervasive.
-
-	In particular, I understand the topological definition, but I don't see why they correspond
-	to what we want. I am looking at the formula from \cite{detecting-nilpotence}:
-	\[H^{1,1}(G,k)=\Hom_{sGr/k}(G^0,\mathbb{G}_a^-)^\pi\]
-	where $\pi$ is the group of connected components of $G$.
-\end{qst}
-
-\begin{qst}
-	In $G=(\bbG_a^-\times\bbG_a^-)\rtimes G_{a(1)}$, where the action is said to be ``nontrivial'',
-	why does this seem to imply that the action on one of the copies of $\bbG_a^-$ is trivial?
-
-	The fact that there is a LHS spectral sequence 
-	\[H^{*,*}(\bbG_a^-\times \bbG_{a(1)},H^{*,*}(\bbG_a^-,k))\Rightarrow H^{*,*}(G,k)\]
-	seems to imply that there is a short exact sequence 
-	\[1\to \bbG_a^-\to G\to \bbG_a^-\times \bbG_{a(1)}\to 1\]
-	but since the product is direct on the right, doesn't this imply that the action of $\bbG_{a(1)}$ on 
-	$\bbG_a^-$ is trivial?
-\end{qst}
-
-Path 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\newpage
-\section{Affine Group Schemes (Waterhouse and Jantzen)}
-I have quite a bit of information to process before this, so I will get started!
-\begin{defn}[Closed Embedding]
-	If $G$ and $H$ are affine group schemes represented by $A$ and $B$, respectively and
-	if $\psi:H\to G$ is a homomorphism of affine group schemes (locally a group homomorphism)
-	then if the corresponding algebra map $A\to B$ is surjective, then $\psi$ is called
-	a \textbf{closed embedding.}
-\end{defn}
-
-As its name suggests, this means that $\psi$ is an isomorphism onto a \textbf{closed subgroup}
-$H'$ of $G$. This is, in fact, a definition of this property. One can also think about 
-it in the following ways: a group scheme $H$ is closed in $G$ if
-\begin{itemize}
-	\item $H$ is defined by the relations imposed by $G$ plus some additional ones.
-	\item $H=V(I)$ for some ideal $I\subset k[G]$. 
-\end{itemize}
-Thinking back to our algebraic geometry, these are not too hard to see as equivalent. 
-For instance, there is a closed embedding of $\mun n$ in $\Gm$ (simply adding in $x^n-1$)
-and of $\SL{n}$ in $\GL{n}$ (adding $\det = 1$).
-
-\subsection{Hopf Ideals}
-One problem with the above characterization that one cannot choose $I$ arbitrarily and end
-up with a group scheme. This is equivalent to arbitrarily adding relations to a group, which is
-not always guaranteed to work out well (think adding $\det = 2$ to $\GL n$).
-
-Actually, we can exactly categorize the closed embeddings of subgroups in $G$ by considering certain
-ideals of the algebra $A$ which represents it. 
-\begin{defn}
-	Let $A$ be an algebra and $I\lhd A$. Then if 
-	\begin{itemize}
-		\item $\Delta(I)$ goes to zero under the map $A\otimes A\to A/I\otimes A/I$,
-		\item $S(I)\subseteq I$
-		\item $\varepsilon(I)=0$
-	\end{itemize}
-	then $I$ is called a \textbf{Hopf Ideal of $A$.}
-\end{defn}
-
-\begin{rmk}
-	That these ideals exactly characterize closed subgroups follows since any group (scheme) represented
-	by $A'$ has the property that $\Delta'(A')\subseteq A'\otimes A'$. Since the new comultiplication map
-	$\Delta'$ is derived from the old one $\Delta$, we have the diagram
-	\begin{figure}[h]
-		\begin{tikzcd}
-			A \ar[r,"\Delta"]\ar[d] & A\otimes A\ar[d]\\
-			A/I\ar[r,"\Delta'"] & A/I\otimes A/I
-		\end{tikzcd}
-	\end{figure}
-
-	\noindent from which we see that $\Delta(I)$ must go to zero in the map on the right.
-	Similarly, we want that $S$ and $\varepsilon$ satisfy similar diagrams.
-
-	Note that an example of such an ideal is $I=\ker(\varepsilon)$.
-\end{rmk}
-
-\begin{defn}
-	A \textbf{character} is a scheme morphism $G\to \Gm$.
-\end{defn}
-
-Let $\Phi$ be a character of $G$ and notice that the corresponding hopf algebra map 
-$\varphi:k[X,1/X]\to A$ is defined by $\varphi(X)=b$, which
-is automatically invertible in $A$. Furthermore since for any $a,b\in G(R)$ we have
-\[\Phi\circ m(a,b)=\Phi(ab)=\Phi(a)\Phi(b)=m\circ(\Phi,\Phi)(a,b)\]
-we get the diagram
-\begin{figure}[h]
-	\begin{tikzcd}
-		G\times G\ar[r,"\Phi\times\Phi"]\ar[d,"m"] & \Gm\times \Gm\ar[d,"m"]\\
-		G\ar[r,"\Phi"] & \Gm
-	\end{tikzcd}
-\end{figure}
-
-\noindent and then by dualizing we get that
-\newpage
-\begin{figure}[h]
-	\begin{tikzcd}
-		A\otimes A & k[X,X^{-1}]\otimes k[X,X^{-1}]\ar[l,"\varphi\otimes\varphi",swap]\\
-		A\ar[u,"\Delta"] & k[X,X^{-1}]\ar[u,swap,"\Delta"]\ar[l,"\varphi"]
-	\end{tikzcd}
-\end{figure}
-
-\noindent where in both diagrams we are abusing notation by using the same symbols for 
-(co)multiplication in the different schemes. So we conclude that 
-\[\Delta(b)=\Delta\circ\varphi(X)=(\varphi\otimes\varphi)\circ \Delta(X)=(\varphi\otimes\varphi)(X\otimes X)=b\otimes b.\]
-
-Then we can easily compute that $\varepsilon(b)=1$ and $S(b)=b^{-1}$. This is precisely:
-\begin{defn}
-	Let $A$ be a Hopf algebra and $a\in A$ such that
-	\begin{itemize}
-		\item $a$ is invertible
-		\item $\Delta(a)=a\otimes a$
-		\item $\varepsilon(a)=1$
-		\item $S(a)=a^{-1}$
-	\end{itemize}
-	then $a$ is called a \textbf{group-like} element of $A$.
-\end{defn}
-\begin{rmk}
-	As we saw above, every character of an affine group scheme $G$ corresponds to a 
-	group-like element of its representing algebra.
-\end{rmk}
-\begin{rmk}
-	Using a parallel construction with any morphism $\Phi:G\to\Ga$, we get an element $b\in A$
-	such that $\Delta(b)=b\otimes 1+1\otimes b$, $\varepsilon(b)=0$ and $S(b)=-b$. These elements
-	are called \textbf{primitive}
-\end{rmk}
-
-\begin{defn}
-	A group scheme $G$ represented by $A$ that consists entirely of group-like elements is
-	called \textbf{diagonalizable.}
-\end{defn}
-\begin{rmk}\label{rmk-diag}
-	Notice that an alternative definition is to begin with an Abelian group $M$ and to 
-	define $S,\Delta,$ and $\varepsilon$ such that each element is group-like. The resulting
-	Hopf algebra represents a diagonalizable group scheme.
-\end{rmk}
-
-Consider the group algebra $k[M]$ on the group in Remark \ref{rmk-diag}. If this algebra
-is finitely generated, we get the following:
-\begin{thm}
-	Let $G$ be diagonalizable and represented by $A$ and assume $A$ is finitely generated
-	as a $k$-algebra. Then $G$ is a finite product of copies of $\Gm$ and $\mun n$.
-\end{thm}
-\begin{prf}
-	Let $x_1,\dots, x_n$ be generators for $k[M]=A$. Each $x_i$ can be written as a (finite!)
-	$k$-linear combination of elements in $M$, so we can instead use the finitely many $m_i$
-	which generate these generators, and this gives us a $k$-algebra basis for $k[M]$. Call this
-	new generating set $U$.
-
-	Let $M'$ be the abelian group generated by $U$ and notice that $k[M']$ is a subalgebra of
-	$k[M]$ containing $U$ so $k[M']=k[M]$ and therefore $M'=M$. This establishes that $M$ is 
-	a finitely generated abelian group, so we can split up the algebra into a tensor product
-	of $k[\bbZ]$ and $k[\bbZ/n\bbZ]$.
-
-	When $M=\bbZ$, $k[M]\cong k[X,X^{-1}]$, so $G= \Gm$\footnote{Notice that here Waterhouse also
-	uses that the basis elements $e_i$ are group-like and so $\Delta(e_i)=e_i\otimes e_i$. I am not
-	quite sure why this is necessary.} Similarly if $M=\bbZ/n$ then $G\cong \mun{n}$
-\end{prf}
-
-Looking at my progress over the last several days, I am thinking that I am getting
-too much in the weeds writing out all the lemmas and proofs. Perhaps it is better to read more quickly
-and only take note of theorems as I need/use them.
-
-\subsection{Cartier Duals}
-The gist here is that if $G$ is represented by $A$, then we can define the dual 
-$A^D=\Hom(A,k)$ and then $A^D$ represents a new scheme $G^D$ called the \textbf{Cartier Dual.}
-
-We do some work to show that, in fact, elements of $G^D$ are in correspondence with the 
-group-like elements of $A$, or equivalently the character group of $G$, $X_G$.
-
-It is easy enough to evaluate $G^D(k)\cong\Hom(A^D,k)$, but luckily one shows that
-``dualizing'' commutes with tensor products (and thus with extension of scalars) and $\Hom$. Thus
-if $G_R$ is the scheme represented by $A\otimes_k R$, then using that $\Hom_k(A^D,R)\cong\Hom_R(A^D\otimes R,R)$,
-\[G^D(R)=(G^D)_R(R)\]
-which is represented by $(A\otimes R)^D=A^D\otimes R$, so finally 
-\[G^D(R)=(G_R)^D(R)=\{\text{group-like elements in }A\otimes_k R\}\cong \Hom(G_R,(\Gm)_R)\]
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\newpage
-\section{Algebraic Geometry}
-
-\newpage
-\section{Problems}
-
-\subsection{Waterhouse}
-\begin{wprob}{2.1}
-	\begin{enumerate}
-		\item Show that there are no nontrivial homomorphisms from $\Gm$ to $\Ga$.
-		\item If $k$ is reduced, show that there are no nontrivial homomorphisms from $\Ga$ to $\Gm$.
-		\item For each nonzero $b\in k$ with $b^2=0$, find a nontrivial homomorphism $\Ga\to\Gm$.
-	\end{enumerate}
-\end{wprob}
-Hmm, there must be a small bug in my macro but I can't find it. I get an error about \texttt{missing `\textbackslash item'}
-in the definition below. It seems to only occur when I use \texttt{\textbackslash subsubsection*} and the like.
-
-\begin{sol}
-	\subsubsection*{(a)}
-	Let $\Phi:\Gm\to\Ga$ and let $\varphi:k[X]\to k[X,X^{-1}]$ be the corresponding Hopf
-	algebra map. But then we have that 
-	\[\varphi\otimes\varphi \circ \Delta_a(X)=\Delta_m\circ\varphi(X)\]
-	and so using that $\Delta_m(a)=a\otimes a$ and $\Delta_a(b)=b\otimes1+1\otimes b$, we get
-	\[\varphi(X)\otimes\varphi(1)+\varphi(1)\otimes\varphi(X)=\varphi(X)\otimes\varphi(X)\]
-\end{sol}
-
-\section{To be filed:}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2019-05-16 Triangulated Categories/Tri Cat.tex
Haven't pushed in a while

\documentclass[12pt]{article}

\usepackage{setspace}

\usepackage{amsmath, amsfonts, amssymb, graphicx, color, fancyhdr, lipsum, scalerel, stackengine, mathrsfs, tikz-cd, mdframed, enumitem, framed, adjustbox, bm, upgreek, x	color}
\usepackage[framed,thmmarks]{ntheorem}

%Replacement for the old geometry package
\usepackage{fullpage}

%Input my definitions
\input{/home/nico/latex-includes/mydefs.tex}

%header stuff
\setlength{\headsep}{24pt}  % space between header and text
\pagestyle{fancy}     % set pagestyle for document
\lhead{Notes on Hopf Algebras} % put text in header (left side)
\rhead{Nico Courts} % put text in header (right side)
\cfoot{\itshape p. \thepage}
\setlength{\headheight}{15pt}
\allowdisplaybreaks

% Document-Specific Macros
\newcommand*{\dgVectk}{\mathbf{dgVect}_k}
\newcommand*{\dgmodA}{\mathbf{dgmod}\text{-}A}
\newcommand*{\modA}{\mathbf{mod}\text{-}A}
\newcommand*{\Amod}{A\text{-}\mathbf{mod}}

\begin{document}
%make the title page
\title{Cohomology Theories, Triangulated Categories, and Applications\vspace{-1ex}}
\author{A course by: Prof. James Zhang\\
Notes by: Nico Courts}
\date{Spring 2019}
\maketitle

\renewcommand{\abstractname}{Introduction}
\begin{abstract}
	These are the notes I took in class during the Winter 2019 topics course
	\textit{Math 583H - Cohomology Theories, Triangulated Categories, and Applications} at University of Washington, Seattle.

	The course description follows:

	\brk

	This covers some (co)homology theories and structures of triangulated categories recently used in noncommutative algebra and 
	noncommutative algebraic geometry. In addition, we will discuss some latest developments and applications in representation 
	theory and noncommutative algebraic geometry. Some basics (see first two topics) can be found in the book \textit{An Introduction to Homological Algebra}
	by Charles Weibel. Here is a list of possible topics:

\begin{itemize}
	\item Hochschild cohomology and group cohomology.
	\item Homotopy categories and derived categories.
	\item Dualizing complexes over noncommutative algebras and applications.
	\item Artin-Schelter regular algebras and noncommutative projective spaces.
	\item Semiorthogonal decomposition of derived categories.
\end{itemize}
\end{abstract}

\section{April 1, 2019}

The topic for this week includes dg-algebras and -categories. There is no great 
textbook for these subjects (for the next topics you can use Weibel), but there is 
a paper by B. Keller ``Deriving dg-categories'' which is pretty abstract. We might 
need a little time to understand it, however.

\subsection{What we are working with}
Let $k$ be a field throughout the course. As a matter of notation, we let $\dgVectk$
denote the category of differential graded vector spaces over $k$. Today we will discuss two ways to
view this, as a \textit{category} and as a \textit{dg-category.}

\subsection{As a category}
Recall that $\Vectk$ is the category with $k$-vector spaces with $k$-linear maps. Now $\dgVectk$ consists of
objects that are (cochain) complexes of $k$-modules $(M^i)_{i\in\bbZ}$ with differential
$\partial_M=(\partial_M^i)_{i\in\bbZ}$ where each 
\[\partial_M^i:M^i\to M^{i+1}\]
is a differential in the usual sense in homological algebra.

To describe the morphisms in this category: Let $M,N$ be two objects in $\dgVectk$. Then 
a (degree zero) morphism $f:M\to N$ are chain morphisms (the components commute with the differentials).

\begin{ex}
	$M^i=kx^i\oplus ky^i$ and $\partial_M^i(x^i)=0$, $\partial_M^i(y^i)=x^{i+1}$. This clearly
	works with everything above, so $M\in\dgVectk$.

	Now let $N^i=kx^i$ and $\partial_N^i(x^i)=0$. Then there is a morphism $f:M\to N$ sending $y^i\mapsto x^i$ and $x^i\mapsto 0$.
\end{ex}

\begin{prob}
	Check that $\dgVectk$ is an abelian category.
\end{prob}

\begin{defn}
	If $M,N\in \dgVectk$ and $f:M\to N$ such that $f^i:M^i\to N^{i+d}$ for all $i$, 
	we say $\deg f=|f|=d$.
\end{defn}

\subsection{As a symmetric monoidal category}
The first thing to do is to define the tensor product. Let $M,N\in \dgVectk$. Then we define
$M\otimes N$ to be
\[(M\otimes N)^i=\oplus_{s+t=i} M^s\otimes N^t\]
with differential (if $m\in M^s$ and $n\in N^t$)
\[\partial_{M\otimes N}^i(m\otimes n)=\partial_M^s(m)\otimes n+(-1)^sm\otimes \partial_N^t(n)\]
where we will often use the more universal notation $(-1)^s=(-1)^{|m|}$.

\begin{rmk}
	Notice that $|\partial|=1$, so the above sign can be more suggestively written as $(-1)^{|m||\partial|}$
	which is a nod to the Koszul sign rule below.

	To be clear, the Koszul sign rule states, if $f$ and $g$ are graded (that is, chain) maps, then
	\[(f\otimes g)(m\otimes n)=(-1)^{|m||g|}f(m)\otimes g(n)\]
	and so the map we are considering above is $\id_M\otimes \partial_N+\partial_M\otimes \id_N$
\end{rmk}
\begin{prob}
	Check that $\partial_{M\otimes N}$ is a differential.
\end{prob}

We want this monoidal category to be \textit{symmetric}, however. So to see this, define the twist functor
\[\tau:M\otimes N\to N\otimes M\]
such that $\tau(m\otimes n)=(-1)^{|m||n|}n\otimes m$ (Koszul sign rule).
\begin{lem}
	\begin{enumerate}
		\item $\tau\in\Hom_{\dgVectk}(M\otimes N,N\otimes M)$. 
		\item $\tau\circ\tau|_{M\otimes N}=\id_{M\otimes N}$.
	\end{enumerate}
\end{lem}

Then since we have the twist functor $\tau$ satisfying the above properties, we get that $\dgVectk$ is 
symmetric monoidal. Notice that the unit object for the monoidal aspect is the complex
\[\cdots\to 0\to k\to 0\to\cdots\]

\subsection{Differential Graded Algebras}
\begin{defn}
	A differential graded algebra $A$ is an algebra object in $\dgVectk$.
\end{defn}
\begin{defn}
	A differential graded algebra $A$ is a $\bbZ$-graded algebra $A=\oplus_{i\in \bbZ}A_i$ with differential
	$\partial_A^i:A^i\to A^{i+1}$ satisfying:
	\begin{enumerate}
		\item $1\in A_0$
		\item $A_iA_j\in A_{i+j}$
		\item $\partial_A(1_A)=0$
		\item $\partial_A(ab)=\partial_A(a)b+a\partial_A(b)$
	\end{enumerate}
\end{defn}

\section{April 3, 2019}
At the end of the last lecture, we wondered why we want to do homological algebra (and thus why
are we interested in dg-structures)?

\subsection{Some History}
Suppose we have a class $\mathcal C$ of objects. Then around 1900, Poincar\'e noted we can construct 
a class of complexes consisting of these objects. Then one can compute (co)homologies and their dimensions (Betti numbers).
Then the alternating sum of the Betty numbers gives us the Euler characteristic.

In topology, around 1925, people did this with topological spaces and singular (co)homology. Later, in the 1930's, 
de Rham introduced de Rham complexes for manifolds and continued similarly. Also around this time, \v Cech developed 
\v Cech complexes for schemes for \v Cech cohomology. In 1945, Hochschild developed Hochschild complexes.
In 1948, Chevalley-Eilenberg did Lie algebras and Harrison did commutative algebras.

\subsection{Commutative DGAs}
Last time we saw the definition of a differential graded algebra. 
\begin{defn}
	A DGA $A$ is called \textbf{commutative} if 
	\[ab=(-1)^{|a||b|}ba\]
	for homogeneous elements $a,b\in A$.
\end{defn}

\subsection{Returning to $\dgVectk$}
Recall that the objects in this category are complexes $M^\bullet$ with differential $\partial^i:M^i\to M^{i+1}$.
\begin{defn}
	For $M,N\in \dgVectk$, define
	\[\Hom_{\dgVectk}(M,N)=(\Hom_{\dgVectk}(M,N)^i,\partial^i)\]
	where 
	\[\Hom_{\dgVectk}(M,N)^i:= \prod_{s\in\bbZ}\Hom(M^s,N^{s+i})\]
	and for each $i$, $\partial^i_{\Hom}:\Hom_{\dgVectk}(M,N)^i\to \Hom_{\dgVectk}(M,N)^{i+1}$, where
	\[\partial^i_{\Hom}((f^s)_{s\in\bbZ})=(\partial_N^{s+i}\circ f^s-(-1)^{|f||partial|}f^{s+1}\circ\partial_M^s)_{s\in\bbZ}\]
\end{defn}
\begin{prob}
	Show that $\partial_{\Hom}$ is a differential. Hint: use that $\partial_M$ and $\partial_M$ are.
\end{prob}
\begin{rmk}
	The composition in $\dgVectk$ is defined as: if $f\in\Hom_{\dgVectk}(M,N)^i$ and $g\in\Hom_{\dgVectk}(K,M)^j$,
	then $f\circ g\in\Hom_{\dgVectk}(K,N)^{i+j}$ where $(f\circ g)=(f^{s+j}\circ g^s)_{s\in\bbZ}$.

\end{rmk}
\begin{lem}
	\begin{enumerate}
		\item $1_M\in\Hom_{\dgVectk}(M,M)^0.$
		\item Composition has the following properties:
		\begin{enumerate}
			\item $\circ$ is associative.
			\item $\Hom_{\dgVectk}(N,P)\otimes \Hom_{\dgVectk}(M,N)\xrightarrow{\circ} \Hom_{\dgVectk}(M,P)$
			is a morphism of complexes. 
		\end{enumerate}
	\end{enumerate}
\end{lem}

\subsection{dg Categories}
\begin{defn}
	A \textbf{dg Category} is a category $\mathcal C$ enriched over $\dgVectk$.
\end{defn}
Okay, that is slightly cheating although it is quite nice. 
\begin{defn}
	Let $\mathcal C$ be a $k$-linear category. We say $\mathcal C$ is a \textbf{dg category} if:
	\begin{enumerate}
%		\item For all objects in $M,N\in\mathcal C$, $\Hom_{\mathcal C}(M,N)$ is a complex $(\Hom_{\mathcal C}(M,N)^i,\partial^i_\Hom )_{i\in\bbZ}$
		\item $1_M\in\Hom_{\dgVectk}(M,M)^0$
		\item The other stuff we talked about in the above lemma.
	\end{enumerate}
\end{defn}
\begin{rmk}
	So a natural question to arise is whether a specific category admits an ``enhancement'', or enrichment 
	over $\dgVectk$. If so, we can do homological algebra!

	Next lecture will be more concrete. :)
\end{rmk}

\section{April 5, 2019}
Today we will just focus on examples of dg categories! No definitions today. :)

\subsection{Examples of dg categories}
\begin{ex}
	$\dgVectk$ is a dg category. We also use this notation to refer to the category without the dg structure.

	Some notation: when we write 
	\[\hom(M,N)=\{\text{maps of degree zero}\}\]
	we are taking about the regular category . Then we call the same category where the maps 
	have arbitrary degree as the dg enchancement.
\end{ex}
\begin{ex}
	A dg algebra $A$ is a dg category with one object. It has a single object $\ast$ and $\Hom_A(\ast,\ast)=A$.
\end{ex}
\begin{rmk}
	This gives us a third definition for a dg category: a dg algebra with several objects. This parallels the fact
	that an algebra over $k$ can be considered a $k$-linear category with a single object and a $k$-algebra with several
	objects is a $k$-linear category. This is a bit funny.
\end{rmk}
In this last example, the differential on $A$ is $\Hom(\ast,\ast)$ and composition respects grading 
and the differential. Furthermore, composition is a homomophism of degree zero ($\deg(a\circ b)=\deg a+\deg b$)
and finally it is a morphism of complexes of degree zero.
\begin{ex} 
	Let $R$ be a algebra, then the category of complexes of right $R$-modules $\Ch=\Ch(\mathbf{mod}\text{-}R)$
	is a dg category whose objects are complexes of right $R$-modules and whose morphisms are themselves complexes.

	That is, $\Hom_{\Ch}(M,N)^i=\prod_{s\in\bbZ}\Hom_R(M^s,N^{s+i})$ with differential
	\[d_{\Hom}(f)=d_N\circ f-(-1)^{|f|}f\circ d_M.\]
\end{ex}
\begin{rmk}
	Usually when one studies $\Ch$, one only considers the regular (non dg) category structure. Actually 
	this is the case in books like Weibel and Rotman.

	When one is looking at representation theory, on considers $R$-modules and so it is natural to consider
	$\mathbf{D}(\modR)$, the derived category where objects are complexes. Then $\Ch(R)$
	has complexes for both objects and morphisms.

	Recovering $R$ from the modules is Morita theory. Getting $\mathbf{mod}\text{-}R$ from its derived category
	is derived Morita theory. 
\end{rmk}

\subsection{He lied about no definitions}
\begin{defn}
	Let $A=(A,m,u,\partial)$ be a dg algebra. $M$ is called a \textbf{dg right $A$-module} if 
	\begin{enumerate}
		\item $M$ is a complex with differential $\partial_M$. 
		\item There is a right action on $M$ given by $m_M:M\otimes A\to M$ such that $m_M$ is a morphism
		of complexes of degree zero (i.e. $\partial_M(ma)=\partial_M(m)a+(-1)^{|m|}m\partial_A(a)$). Furthermore
		$m_M$ satisfies the usual diagrams for a right action.
	\end{enumerate}
\end{defn}
\begin{rmk}
	We write $\dgmodA$ to denote the right $A$-modules. As usual, we can define it as a dg category
	or regular category.

	Let $M$ and $N$ be two dg right $A$-modules. A morphism of degree zero from $M$ to $N$ maps
	$M^i$ to $N^i$ for all $i$ and commutes with the differential and is $A$-linear.

	When scaling up to the dg structure, we get $\Hom_{\dgmodA}(M,N)$ consists of complexes where each 
	component consists of maps of the appropriate degree that are all still (right) $A$-linear.

	Notice that in the general case we don't get commutativity with differentials.
\end{rmk}
\begin{rmk}
	Notice that these are nonempty since $\id_M\in\Hom(M,M)^0$ and if $I$ is any dg-ideal  of $A$
	(graded ideal closed under $\partial$), then the quotient map is in $\Hom(A,A/I)^0$.
\end{rmk}
The differential on $\Hom(M,N)$ is the usual thing with the sign trick.
\begin{prob}
	Prove that $\partial_{\Hom}\circ \partial_{\Hom}=0$ and that $\dgmodA$ is a dg category.
\end{prob}

Note that if we let $A=(\cdots\to 0\to R\to 0\to \cdots)$ then $\dgmodA$ is the same as $\Ch(\modR)$.

\section{April 8, 2019}
Today we will mostly be playing with morphisms. We expect it to be quite easy. On Wednesday we will play with the objects.

\subsection{Morphisms in a dg category}
Notice that if $A$ is a dg algebra (or just an algebra) -- recall that this is a (dg) Category with one object.
Then we can define $B\subseteq A$ subalgebras, $I\subseteq A$ ideals, and quotients $A/I$ as usual.

It is important to notice that properties can change from $A\to B$ and $A\to A/I$. The set of units may change, for instance.

Today we will let $\mathcal C$ be a dg category and then analyze 6 (dg) categories and ideals associated to $\mathcal C$. Along
the way, it is good to keep $\mathbf{dgAlg}$ and $\dgmodA$ in mind as examples.

Recall we had one definition of a dg category $\mathcal C$ as an Algebra with several objects. Then we 
are really studying $\mathcal B,\mathcal I\subseteq\mathcal C$ and $\mathcal C/\mathcal I$. Here we are 
assuming that the objects in all related categories are the same.
\begin{prob}
	Write down the definitions of subcategory, ideal, and quotient that work with these definitions.
\end{prob}

\subsection{Homological Categories?!}
Let $\mathcal C$ be a dg category. Then we will define analogs of the homological algebra objects 
we look for usually.
\begin{defn}
	Define $Z(\mathcal C)$ to be the category with the same objects as $\mathcal C$ with 
	\[Z\Hom_{\mathcal C}(M,N)=\Hom_{Z(\mathcal C)}(M,N)=\{f\in \Hom_{\mathcal C}(M,N)|\partial_{\Hom_{\mathcal C}}(f)=0\}\]
\end{defn}
\begin{lem}
	$Z(\mathcal C)$ is a dg subcategory of $\mathcal C$ (with zero differentials).
\end{lem}
\begin{prf}
	It is clear $Z(\mathcal C)\subseteq \mathcal C$. Recall that an axiom of a dg category is that $\partial 1_M=0.$
	Thus $1_M\in \Hom_{Z(\mathcal C)}(M,M). $

	Let $f\in\Hom(M,N)$ and $g\in\Hom(N,P)$. Then 
	\[\partial(g\circ f)=\partial (g)\circ f+(-1)^{|g|}g\circ \partial(f)=0\]
	and we are done. :)
\end{prf}
\begin{defn}
	$B(\mathcal C)$ again consists of the same objects as $\mathcal C$ but now 
	\[B\Hom_{\mathcal C}(M,N)=\Hom_{B(\mathcal C)}(M,N)=\{f\in \Hom(M,N)|f=\partial_{\Hom}s f'\}\]
\end{defn}
\begin{lem}
	$B(\mathcal C)$ is an ideal of $Z(\mathcal C)$.
\end{lem}
\begin{rmk}
	To show that $B(\mathcal C)$ is an ideal in $Z(\mathcal C)$, we need to check that for all $f\in Z(\mathcal C)$ and $g\in B(\mathcal C)$
	that (if $f$ and $g$ can be composed) that $f\circ g$ and/or $g\circ f$ are back in $B(\mathcal C)$.
\end{rmk}
\begin{prf}
	Let $f\in \Hom_Z(M,N)$ and $g\in\Hom_B(L,K)$. Then by definition $g=\partial g'$. If $f\circ g$ 
	is defined, then 
	\begin{align*}
		(-1)^{|f|}f\circ g &= f\circ \partial(g')\\
		&= \partial(f)\circ g +(-1)^{|f|}f\circ\partial(g')\\
		&=\partial(f\circ g')\in B(\mathcal C)
	\end{align*}
	reversing the composition, if allowable, can be checked similarly.
\end{prf}
\begin{defn}
	$H(\mathcal C)$ is defined to have the objects of $\mathcal C$ and
	\[H\Hom(M,N)=\Hom_{H(\mathcal C)}(M,N)=\Hom_Z(M,N)/\Hom_B(M,N).\]
\end{defn}
James uses the book \textit{Rational Homotopy Theory} when he is talking about dgas and dg categories.
He says it's a good reference for this part of the class.

\begin{defn}
	Suppose $\mathcal C$ is a dg category with zero differential. Then define
	$\mathcal C^0$ to be the category with the same objects and 
	\[\Hom_{\mathcal C^0}(M,N)=\Hom_{\mathcal C}(M,N)^0\]
\end{defn}
\begin{lem}
	$\mathcal C^0$ is a category.
\end{lem}
\begin{prf}
	``It's kinda obvious, right?''
\end{prf}

\subsection{Homotopy categories}
We can define in the more concrete case the category $\mathbf{K}(\modR):=H^0(\Ch(\modR))$.

\section{April 10, 2019}
Today is going to be a three-star day. We are going to be doing the next natural thing: ``playing with objects \textit{and} morphisms.''

\subsection{Playing with Objects}
Last time we defined the categories we were talking about without changing the underlying objects of the category.
We need to define:
\begin{itemize}
	\item Shift or suspension
	\item Mapping cone
	\begin{itemize}
		\item To do this we will need some motivation.
	\end{itemize}
\end{itemize}
Recall that last time we discussed the process of creating $H^0(\dgmodA)$. Here the $\hom$ sets 
are just vector spaces (instead of dg vector spaces) so we lose some structure. Then the shift and the
cone operations will help recover (some) of the lost information.

\begin{rmk}
	From last time: let $A$ be a (dg) algebra. Then we can define a subalgebra $B$ of $A$
	and ideals $I\lhd A$ and quotients $A/I$. We did everything in categories on Monday: subcategories, ideals, and quotients.

	One important process in algebras is localization. The analog also exists in category theory: we can invert 
	a collection of morphisms. This is more complicated (as expected) than in the case of algebras, but analogous.
\end{rmk}
\begin{rmk}
	$Z^0(\dgmodA)$ is an Abelian category. $H^0(\dgmodA)$ is usually not Abelian. So we can't really
	talk about subobjects and quotients, etc, in the homotopy category although we can talk about them in the 
	cycle category and remember the correspondence.
\end{rmk}
\begin{rmk}
	Consider $\Ch(\modR)$. We can define three different functors:
	\begin{itemize}
		\item $P:\modR\to H^0(\Ch(\modR))$, which takes the projective resolution of a module.
		\item $i:\modR\to H^0(\Ch(\modR))$, which embeds $M$ as a complex $\cdots\to 0\to M\to 0\to \cdots$.
		\item $I:\modR\to H^0(\Ch(\modR))$, which assigns to each module an injective resolution.
	\end{itemize}
	The first and third functors are called the \textbf{projective} and \textbf{injective deriving functors,} respectively.

	We have a natural transformations $P\to i\to I$ in $H^0(\Ch(\modR))$, but when we pass to $D(\modR)$,
	these become natural isomorphisms.
\end{rmk}

\subsection{Shift/Suspension}
\begin{defn}
	Let $\mathcal C=\dgmodA$. Let $M$ be a right dg $A$ module (complex). Then the \textbf{shift}
	of $M$, denoted by $\Sigma M$, is another right dg $A$-module such that:
	\begin{itemize}
		\item $\Sigma M=M$ as a right $A$-module.
		\item $(\Sigma M)^i=M^{i+1}$ for all $i$.
		\item $\partial_{\Sigma M}=\partial_M$
	\end{itemize}
\end{defn}
\begin{rmk}
	Elements in $(\Sigma M)^i=M^{i+1}$ are denoted by $sm$, where $m\in M^{i+1}$. Then 
	\[\partial_{\Sigma M}(sm)=(-1)s(\partial_M(m))\in (\Sigma M)^{i+1}=M^{i+2}.\]
\end{rmk}
\begin{rmk}
	Actually, $\Sigma$ is a functor! Just using that $|\partial|=1$ got us the sign, but we 
	can generalize completely using $\sigma f=(-1)^|f|f.$
\end{rmk}
\begin{ex}
	Suppose $A=(\cdots\to 0\to R\to 0\to \cdots)$ with $R$ in the zeroth position. Then $\Sigma M$ is the same complex
	with $R$ shifted to the $(-1)^{st}$ position.
\end{ex}
\begin{lem}
	$\Sigma$ is invertible and the inverse is denoted $\Sigma^{-1}$ in $\dgmodA$ and all cycle, cocycle, and homology categories.
\end{lem}
\begin{rmk}
	It is a bit troubling talking about functors on $B^0(\mathcal C)$, which is an ideal category, but not 
	a category proper. Basically the idea here is that we still get that it preserves morphisms, but identities may 
	not exist and therefore can't be preserved.
\end{rmk}
\begin{lem}
	\[\Hom_{\dgmodA}(M,\Sigma N)=\Sigma|_{\dgVectk} \Hom_{\dgmodA}(M,N)\] 
	Hint: $\Hom(M,\Sigma N)^i=(\Sigma\Hom(M,N))^i$.
\end{lem}
\subsection{Mapping Cone}
As a construction, we first consider it as a map of sets
\[\cone:\Hom_{Z^0(\dgmodA)}(M,N)\to \dgmodA.\]
\begin{defn}
	If $f:M\to N$ is a cycle in $\dgmodA$, then we define \textbf{the cone of $f$}, denoted $\cone(f)$,
	to be a right dg $A$ module such that
	\begin{itemize}
		\item $(\cone(f))^i=(\Sigma M\oplus N)^i=M^{i+1}\oplus N^i$
		\item $(sm+n)\cdot a=s(m\cdot a)+n\cdot a$
		\item $\partial_{\cone(f)}(sm+n)=-s\partial_M^{i+1}(m)+[f(m)+\partial_N^i(n)]$
	\end{itemize}
\end{defn}
\begin{lem}
	$\partial^{i+1}\circ\partial^i=0$.
\end{lem}
\begin{prf}
	\[\partial(sm+n)=-s\partial_M(m)+[f(m)+\partial_N(n)]\]
	so applying the differential again:
	\[\partial^2(sm+n)=s\partial_M^2(m)+-s\partial\]
	I am missing something here but I believe it is clear.
\end{prf}
\begin{thm}
	$\cone:\operatorname{Mor}(Z^0(\dgmodA))\to \dgmodA$ descends to a map to $H^0(\dgmodA).$
\end{thm}
Here is some homework or just some facts if you are feeling lazy:
\begin{itemize}
	\item $0\to N\to \cone(f)\to \Sigma M\to 0$ is short exact (in $\dgmodA$).
	\item If $K\in \dgmodA$, then 
	\[\Hom_{\dgmodA}(K,\cone(f))\cong\cone_{\dgVectk}(\Hom_{\dgmodA}(K,f))\]
\end{itemize}

\section{April 12, 2019}
Next week's topic will be Hochschild cohomology, which you can find in Weibel (chapter 9). Today we are going to finish
the lecture we began on Wednesday. We will also review projective resolutions.

\subsection{Mapping Cone (again)}
Recall that if we have $f:M\to N$ in $Z^0(\dgmodA)$, we can construct $\cone(f)\in\dgmodA$ where
\[\cone(f)^i=(\Sigma M\oplus N)^i\]
and
\[\partial_{\cone(f)}=\partial_{\Sigma M}\oplus(f+\partial_N)\]
and more specifically
\[\partial_{\cone(f)}(sm+n)=-s\partial_M^{i+1}(m)+(f(m)+\partial_N^i(n)).\]

\begin{rmk}
	Where does this formula come from? Consider the diagram

	\begin{center}
		\begin{tikzcd}
			M^{i-1}\ar[r,"\partial_M^{i-1}"]\ar[d,"f^{i-1}"] & M^i\ar[r,"\partial^i_M"]\ar[d,"f^i"]\ar[dl] & M^{i+1}\ar[d,"f^{i+1}"]\ar[dl]\\
			N^{i-1}\ar[r,"\partial_N"] & N^i\ar[r,"\partial N"] & N^{i+1}
		\end{tikzcd}
	\end{center}

	\noindent where the diagonal arrow indicate ``pushing down'' a copy of $M^i$ to direct sum with $N^{i-1}$.
	But then using the maps you have in the diagrams, you get the maps $\partial_M^{i+1},f^{i+1}$ and $\partial_N^i$
	between $M^{i+1}\oplus N^i$ and $M^{i+1}\oplus N^{i+1}$, and $\partial_{\cone(f)}$ is just their sum!

	I need to think more to see whether this fact is actually meaningful or just a mnemonic.
\end{rmk}
\begin{ex}
	Consider a projective resolution $P_M$ of $M$ and the map $f:P_0\to M$. Then considering that $M$ is 
	a complex with zeros everywhere except the zero position and consider the chain map $f$ induced by $f_0$
	(and zeros elsewhere) and then we can compute
	\[\cone(f)=\cdots \to P^{-1}\to P^0\to M\to 0. \]
\end{ex}
\subsection{Facts about $\cone(f)$}
\begin{rmk}
	There exists a short exact sequence in $\dgmodA$
	\[0\to N\to\cone(f)\to\Sigma M\to 0\]
	where $n\mapsto 0+n$ under the first man and $sm+n\mapsto sm$ in the second.

	This is called the \textbf{mapping cone sequence.}
\end{rmk}
\begin{rmk}
	Let $K$ be another dg right $A$ module. Then
	\[Z^0(\dgVectk)\ni\Hom_{\dgmodA}(K,f):\Hom_{\dgmodA}(K,M)\to\Hom_{\dgmodA}K,N)\]
	and furthermore
	\[\cone(\Hom_{\dgmodA}(K,f))\cong\Hom_{\dgmodA}(K,\cone(f))\]
	and
	\[\cone(\Hom_{\dgmodA}(\Sigma f,K))\cong\Hom_{\dgmodA}(\cone(f),K)\]
\end{rmk}
\begin{rmk}
	Let $K$ be a left dg $A$ module. Then the map
	\[f\otimes_A \id_K:M\otimes_AK\to N\otimes_AK\]
	is such that
	\[\cone(f\otimes \id_K)=\cone(f)\otimes K.\]

	We can also switch ``left'' and ``right.''
\end{rmk}
\subsection{A ``problem''}
The map
\[\cone:\operatorname{Mor}(Z^0(\dgmodA))\to \dgmodA\]
is \textbf{not} functorial!

However, suppose that $f,g:M\to N$ and $[f]=[g]\in H^0(\dgmodA)$. Whenever this is true, there exists
a map $\Phi:\cone(f)\to \cone(g)$ that fits into the diagram
\begin{center}
\begin{tikzcd}
	0\ar[r] & N\ar[d,"\cong"]\ar[r]& \cone(f)\ar[r]\ar[d,dashed,"\Phi"] & \Sigma M\ar[d,"\cong"]\ar[r] & 0\\
	0 \ar[r] & N\ar[r] & \cone(g)\ar[r] & \Sigma M\ar[r] & 0
\end{tikzcd}
\end{center}

\noindent which is clearly an isomorphism via the five lemma. In fact, such a map exists 
if and only if these two maps are homotopic!

\subsection{Classical Homological Algebra}
Recall that we can always construct projective resolutions of a (right) $R$-module by taking
a free module that maps onto $M$ and then iterate. Thus such a resolution always exists.

Let $f:M\to N$ be a morphism of right $R$-modules. Then there exits $F$, the lift of $f$,
between any projective resolutions $P_M$ and $P_N$. Notice that the resolutions \textit{don't}
include the module themselves. Furthermore any two lifts of $f$ differ only by a boundary in $\Ch(R)$,
and thus are equal in the homotopy category.

From this we can conclude that any $M$ in $H^0(\Ch(R))$, there exists a unique projective resolution,
up to (a unique) isomorphism. In this way, we can see that ``taking a projective resolution'' defines a 
functor from $\modR$ to $H^0(\Ch(R))$.

For any two $R$ modules $M$ and $N$, 
\[\Ext_R^i(M,N):=H^i(\Hom_{\Ch(R)}(P_M,P_N)\left(=H^i(\Hom_{\Ch(R)}(P_M,N)\right).\]

If $M=N$, then $\Hom_{\Ch(R)}(P_M,P_M)$ is a dg algebra. As a consequence,
\[\Ext_R^*(M,M)=\oplus_i H^i(\Hom(P_M,P_M))\]
is a graded algebra.

\begin{defn}
	Then we can define \textbf{the Hochschild cohomology} as
	\[\oplus_{i\in \bbZ}\Ext_{R\otimes R^{op}}^i(R,R)\]
\end{defn}
\begin{rmk}
	A nickname of Hochschild cohomology is ``the derived functor of the center.'' 
\end{rmk}

\section{April 15, 2019}
Today we will focus on understanding $\HH^0(A)$, the zeroth Hochschild cohomology. Recall that 
Sarah Witherspoon has a preprint (I have it) of a book she is going to release on the subject.

Define the Hochschild cohomology as we saw last week. Here james spent some extra time defining $A^{op}$ 
as well as going over why it suffices to define it over $A^e$.

\subsection{Hochschild Cohomology}
For clarity:
\begin{defn}[Hochschild Cohomology]
	For every $i\in\bbZ_{\ge 0}$, the $i^{th}$ Hochschild cohomology of $A$ is defined to be
	\[\HH^i(A)=\Ext_{A^e}^i(A,A)=H^i(\Hom_{\Ch(A^e)}(P_A,P_A))\]
	where $A$ is any associative algebra.
\end{defn}
\begin{rmk}
	In particular, $\HH^0(A)=\Ext_{A^e}^0(A,A)=\Hom_{A^e}(A,A)$.
\end{rmk}
\begin{rmk}
	``\textbf{Slogan 1}'': Hochschild cohomology is the derived functor of $\Hom_{A^e}(A,A)$.
\end{rmk}
\begin{rmk}
	In fact, we can (and will!) define \textit{Hochschild cohomology of an algebra with coefficents
	in an $A^e$ module $M$} as $\HH^i(M)=\Ext_{A^e}^i(A,M).$
\end{rmk}
\begin{lem}
	$\Hom_{A^e}(A,A)=Z(A)$.
\end{lem}
\begin{prf}
	Given any $f\in \Hom(A,A)$, set $c=f(1)$. We claim that $c\in Z(A)$. To see this, notice that for 
	every $a\in A$, $ac=af(1)=f(a\cdot 1)=f(1\cdot a)=f(1)a=ca.$

	Thus we can define a map $f\mapsto f(1)$. To define the inverse, just sent $c\in Z(A)$ to 
	$f$ mapping $a\mapsto ca(=ac)$. Then a standard verification shows what we want.
\end{prf}
\begin{rmk}
	``\textbf{Slogan 2}'': Hochschild cohomology is the derived functor of ``the center.''
\end{rmk}
This leads us to a really ``wild'' conjecture:
\begin{conj}
	$\HH^*(A)$ is a commutative algebra.
\end{conj}
\begin{rmk}
	We really have only proved this for the zeroth piece, which is why this is so wild. In fact, it is false.
\end{rmk}
Instead,
\begin{thm}
	$\HH^*(A)$ is \textit{graded} commutative.
\end{thm}
\begin{rmk}
	We do not have the machinery we need to prove this yet. For today we will prove a weaker result.
\end{rmk}
\begin{thm}
	$\HH^*(A)$ is a graded algebra.
\end{thm}
\begin{prf}
	$\HH^*(A)=\Ext_{A^e}^{\ge 0}(A,A)$ by definition. This can be computed as
	\[H^*(\Hom_{\Ch(A^e)}(P_A,P_A))=H^*(B)\]
	where $B$ is a dg algebra. But then $H*(B)$ is a graded algebra.
\end{prf}

\subsection{Some more categories}
Let $\calC$ be an abelian category and let $\mathbf{Rex}(\calC)$ be the category of functors $\calC\to \calC$
that are right exact and commute with $\oplus.$
\begin{rmk}
	Notice that whenever $\operatorname{Fun}(\calC,\calC)$ can be a replacement, but it isn't always abelian.
\end{rmk}
\begin{thm}[Watts]
	If $\calC=R\text{-}\mathbf{mod}$, then $\mathbf{Rex}(\calC)=A^e\text{-}\mathbf{mod}$.
\end{thm}
\begin{rmk}
	More generally, any right exact functor from $A$ modules to $B$ modules that commutes with direct sums
	is isomorphic to $M\otimes_A -$ for some $(B,A)$-bimodule.
\end{rmk}
\begin{lem}
	Let $\calC$ be $A$-$\mathbf{mod}$. Then $\id_\calC=A\otimes_A-$.
\end{lem}
\begin{defn}
	Let $\calC$ be an abelian category. For every $i\in\bbZ$, the $i^{th}$ Hochschild cohomology
	of $\calC$ is defined to be
	\[\HH^i(\calC)=\Ext^i_{\mathbf{Rex}(\calC)}(\id_\calC,\id_\calC).\]
\end{defn}
\begin{rmk}
	If $\calC$ is $A$-$\mathbf{mod}$, then this agrees with $\HH^i(A)$. We have a map $\Phi:\mathbf{Rex}(\calC)\to A^e$-$\mathbf{mod}$
	sending $\id_\calC$ to $A$, and when $i=0$,
	\[\HH^0(\calC)=\Hom_{\mathbf{Rex}}(\id_\calC,\id_\calC)=\operatorname{Mor}_\calC(\id_\calC,\id_\calC)\]
	where $\operatorname{Mor}$ denotes the natural transformations. This forms a commutative ring where the multiplication
	is composition and the commutativity follows from the naturality squares.
\end{rmk}
\begin{defn}
	With $\calC$ as above, $\operatorname{Mor}_\calC(\id_\calC,\id_\calC)$ is called the \textbf{center of $\calC$}. This definition
	arises from the analogy with algebras.
\end{defn}
\begin{rmk}
	Hochschild cohomology is the derived functor of $\Hom_{\mathbf{Rex}}(\id_\calC,\id_\calC)$.
\end{rmk}
Next time we will consider $\HH^1$!

\section{April 17, 2019}
The lectures this week will cover three apparently distinct topics: the center, derivations, and deformations. The 
underlying math that connects these is the Hochschild complex, which we won't define until next week.

Some homework (for those who haven't seen it):
\begin{ex}
	Find the definition of a Lie algebra and work out some examples.
\end{ex}
\subsection{Derivations and $\HH^1(A)$}
Recall that $\HH^i(A)=\Ext_{A^e}^i(A,A)$.
\begin{defn}
	Let $A$ be an algebra. A $k$-linear map $f:A\to A$ is called a \textbf{derivation} if 
	\[f(1)=0\]
	and for all $a,b\in A$,
	\[f(ab)=f(a)b+af(b).\]

	The set of all derivations of $A$ is denoted $\Der(A)$.
\end{defn}
\begin{defn}
	Let $x\in A$. Define $I_x:A\to A$ by $I_x(a)=xa-ax$. One can easily check that $I_x\in\Der(A)$.
	This derivation is very special: it is called an \textbf{inner derivation.} Any derivation which is 
	not inner is \textbf{outer.}
\end{defn}
\begin{ex}
	$A=k[x]$ with derivation $\frac{d}{dx}$. Furthermore $\Der(A)=\{f(x)\frac{d}{dx}|f(x)\in A\}$. The only
	inner derivation is zero since $A$ is commutative.
\end{ex}
\begin{rmk}
	Notice that $\Der(A)$ is a Lie algebra with $[f,g]=fg-gf\in\Der(A)$. Furthermore $\text{Inn}(A)$ is a 
	Lie ideal, and $\Der(A)/\text{Inn}(A)$ is a quotient Lie algebra.
\end{rmk}
And finally we get to another definition of $\HH^1$:
\begin{defn}
	The first Hochschild cohomology group of $A$ is defined to be
	\[\HH^1(A)=\Der(A)/\text{Inn}(A)\]
	which we sometimes write as $\text{Out}(A)$, the outer derivations of $A$.
\end{defn}
\begin{rmk}
	(\textbf{Slogan 4}): $\oplus\HH^{i+1}(A)$ is the derived functor of $\text{Out}(A)$.
\end{rmk}
\begin{thm}
	$\oplus\HH^{i+1}(A)$ is a graded Lie algebra.
\end{thm}
\begin{prf}
	Nope.
\end{prf}
\begin{rmk}
	Notice that if $\HH^1(A)=0$, then every derivation is inner. For example, when 
	$A=M_n(k)$, every derivation of $A$ is inner. To see this, notice that $A\otimes A^{op}=M_n(k)\otimes M_n(k)=M_{n^2}(k)$,
	which has global dimension zero. So $\Ext_{A^e}^1(A,A)=0$.
\end{rmk}

\subsection{Using Categories}
Recall that $\HH^i(A)=\Ext_{\mathbf{Rex(\calC)}}^i(\id_{\calC},\id_\calC)$. It ends up that if we pick two 
algebras $A$ and $B$ which are Morita equivalent (that is, their module categories are equivalent) and let $\calC=A\text{-}\mathbf{mod}$ or $B\text{-}\mathbf{mod}$, this construction
yields the same thing! So $\HH^i(A)=\HH^i(B)$. This is not at all obvious on the level of complexes, 
so it is a vote in favor of abstract nonsense.

\section{April 19, 2019}
Today we are talking about \textbf{infinitesimal deformations of algebras} and how they will relate to $\HH^2(A)$.

It ends up that ``\textit{every deformation (moduli) problem is related to some cohomology.}'' We will use the letter 
``i'' to stand for ``infinitesimal.'' Recall that we have usually been considering algebras $A$ over a field $k$, but
we also want to consider algebras over a ring $R=k[t]/(t^2)\cong k\oplus kt$
\subsection{Infinitesimal Deformations and $\HH^2$}
\begin{defn}
	Let $(A,m)$ be a $k$ algebra. An \textbf{infinitesimal deformation} of $A$ is a $k[t]/t^2$ algebra
	structure $m_t=\ast_t$ (this is the multiplication) on $A\otimes k[t]/t^2$ (note this is free over $R$, using the $k$-basis of $A$).

	The multiplication on the deformation is defined as 
	\[a\ast_t b=ab+\mu(a,b)t\in A\oplus At\cong A\otimes R\]
	for some $\mu:A\otimes A\to A$. This only defines the multiplication for elements in $A$, but this extends by linearity to $A\oplus At:$
	\[(a+a't)\ast_t(b+b't)=a\ast_t b+(a'\ast_t b)t+(a\ast_t b')t+0\]
	since in $R$, $t^2=0$.
\end{defn}
\begin{rmk}
	Notice that it is actually part of the definition of an infinitesimal deformation is \textbf{associative.}
	I didn't see this on the board, but rather than having this family of algebras freely parameterized by $\mu$, 
	we only take those $\mu$ such that the deformation is associative.
\end{rmk}
\begin{ex}
	The \textbf{trivial infinitesimal deformation of $A$} is determined by $\mu=0$.
\end{ex}
\begin{ex}
	Let $A=k[x,y]$ (works for arbitrary variables) where we define, for all $f,g\in A$
	\[f\ast_t g=fg+\left(\frac{\partial f}{\partial x}\frac{\partial g}{\partial y}-\frac{\partial f}{\partial y}\frac{\partial g}{\partial x}\right)t.\]
	It ends up that this algebra is associative! You can show it. :) This version of Nico doesn't feel like typing it.
\end{ex}
\begin{lem}[Condition 2-C]\label{lem-2-C}
	$m_t$ is associative if and only if for all $a,b\in A$,
	\[a\mu(b,c)-\mu(ab,c)+\mu(a,bc)-\mu(a,b)c=0.\]
\end{lem}
\begin{lem}
	An $i$-deformation is completely determined by $\mu:A\otimes A\to A$ satisfying the 2-C condition in lemma~\ref{lem-2-C}.
\end{lem}
\begin{prf}
	Suppose $a,b,c\in A$ and that $(a\ast b)\ast c=a\ast(b\ast c)$. Then computing using the definitions gets us
	\[(ab)c+\mu(ab,c)t+\mu(a,b)ct=a(bc)-\mu(a,bc)t+a\mu(b,c)t.\]

	To go in the opposite direction, you basically use this identity to recover the equality for associativity.
\end{prf}

\begin{cor}
	The collection of infinitesimal deformations is in one-to-one correspondence with the collection of $k$-linear maps $\mu A\otimes A\to A$ 
	satisfying 2-C.
\end{cor}
\begin{cor}
	The collection of infinitesimal deformations is a $k$ vector space.
\end{cor}
\begin{rmk}
	Just formally define each infinitesimal deformation to be a vector corresponding to the $\mu$ which defines it.
	The fact that the set of such $\mu$ is a $k$ vector space means the structure can be ``pushed over'' to the other collection.
\end{rmk}

\begin{defn}
	Let $m_t^1$ and $m_t^2$ be two infinitesimal deformations of $A$. We say they are \textbf{equivalent} if 
	there is an $R=k[t]/t^2$-algebra isormorphism
	\[f_t:(A\otimes R,m_t^1)\to (A\otimes R, m_t^2)\]
	such that $f(a)=a+\zeta(a)t$ for all $a\in A$. In this case we write $m_t^1\sim m_t^2$ or $\mu^1\sim \mu^2$.
\end{defn}
\begin{defn}
	$m_t$ is called \textbf{quasi-trivial} if $m_t$ is equivalent to the trivial infinitesimal deformation.
\end{defn}

\begin{lem}[Condition 2-B]
	Let $m_t^1$ and $m_t^2$ be two infinitesimal deformations of $A$. Then $m_t^1\sim m_t^2$ if and only if 
	there exists $\zeta:A\to A$ such that $\mu^1(a,b)-\mu^2(a,b)=a\zeta(b)-\zeta(ab)+\zeta(a)b$ for all $a,b\in A$.
	This is called condition 2-B.
\end{lem}
\begin{lem}
	Every equivalence between $m_t^1$ and $m_t^2$ is completely determined by $\zeta$ in 2-B.
\end{lem}
\begin{defn}
	$\mu:A\otimes A\to A$ is called a \textbf{2-coboundary} if there exists $\zeta:A\to A$
	such that
	\[\mu(a,b)=a\zeta(b)-\zeta(ab)+\zeta(a)b\]
	(that is if $\mu$ is quasi-trivial).
\end{defn}
\begin{defn}
	$\mu$ is called a \textbf{2-cocycle} if it satisfies 2-C.
\end{defn}
\begin{rmk}
	This gives us a vector subspace of coboundaries (quasi-trivial infinitesimal deformations) of the
	space of all infinitesimal deformations, so we can do cohomology!
\end{rmk}
\begin{defn}
	The \textbf{second Hochschild cohomology of $A$} is
	\[\HH^2(A)=\frac{\text{{infinitesimal deformations}}}{\text{{quasi-trivial i-defs}}}=\frac{\text{{2-cocycles}}}{\text{{2-coboundaries}}}\]
\end{defn}

\section{April 22, 2019}
Today we will focus on connecting the notions that we have seen in the past several lectures (center/derivations/infinitesimal deformations) 
via the Hochschild complex. On Wednesday we are going to prove that the cohomology associated with the Hochschild complex are what we have computed so far.

Recall the definitions and constructions of $\HH^i(A)$ for $i=0,1,2$. Furthermore we asserted that $\HH^*(A)$
is a graded commutative ring. Furthermore we showed that $\HH^1(A)$ is a Lie algebra and that $\HH^{*+1}(A)$ is a 
graded Lie algebra.

A natural question one may ask is if there is a ``multiplication'' operation on $\HH^2$ that we can apply, analogous to the commutative ring/Lie algebra
in the first two examples. In that case, can we extend this structure to $\HH^{*+2}(A)$?
\begin{rmk}
	Weibel has a different way to talk about $\HH^2$. In fact, James seems to imply there are several different ways to think about it.
\end{rmk}

\subsection{The Hochschild Complex}
\begin{defn}
	Let $A$ be an algebra over $k$. Define the \textbf{Hochschild complex} $C(A)$ of $A$ to be a cochain complex 
	such that 
	\[C(A)^n=\Hom_k(A^{\otimes n},A)\]
	for $n\ge 0$ and $0$ for $n<0$. Note that $A^{\otimes 0}=k$.

	The differential is defined to be a map $\partial^n:C(A)^n\to C(A)^{n+1}$ where for every $f\in C(A)^n$,
	\begin{align*}
		\partial^n f(a_1\otimes \cdots\otimes a_{n+1})&=a_1f(a_2\otimes a_3\otimes\cdots\otimes a_{n+1})\\
		&\qquad +\sum_1^n(-1)^if(a_1\otimes\cdots\otimes a_ia_{i+1}\otimes\cdots \otimes a_n)\\
		&\qquad + (-1)^{n+1}f(a_1\otimes\cdots\otimes a_n)a_{n+1}
	\end{align*}
\end{defn}
\begin{rmk}
	The idea here is the first and last terms actually arise from the fact that $\Hom_k(A^{\otimes n},A)$ has a left- and right 
	$A$-module structure while the interior terms arise more from the chain complex.
\end{rmk}
\begin{lem}
	$C(A)$ is a cochain complex.
\end{lem}
\begin{rmk}
	We omit the proof for the sake of time. James assures us it is true. :P
\end{rmk}

\subsection{Double-checking our work}
Let's compute! 
\begin{defn}
	Define the $i$-cocycles and $i$-coboundaries as we do with any chain complex.
\end{defn}
Then we can compute $H^0(C(A))=Z^0(C(A))/B^0(C(A))=Z^0(C(A))$. But $C(A)^0=\Hom_k(k,A)\cong A$, 
so we can determine when $\partial(f)=0$:
\[\partial(f)(a)=af(1)-f(1)a=0\]
for all $a$, so this means that $f(1)$ is in the center of $A$, so $Z^0(C(A))=H^0(C(A))\cong\cal Z(A)$.

\brk

What about $\HH^1$? Now $Z^1$ is the set of $f\in \Hom(A,A)$, such that
\[\partial(f)(a\otimes b)=af(b)-f(ab)+f(a)b=0\]
and so rewriting this, we get $f(ab)=af(b)+f(a)b$, so $f$ is a derivation!

To compute the coboundaries, say $f=\partial(g)\in C(A)^1$. Then 
\[f(a)=\partial(g)(a)=ag(1)-g(1)a=-I_{g(1)}(a)\]
so $f$ is an inner derivation of $A$. So we recover $\HH^1$.

\brk

Now $Z^2$ is the set of $f\in C(A)^2$ such that $\partial(f)=0$ -- so
\[\partial(f)(a\otimes b\otimes c)=af(b\otimes c)-f(ab\otimes c)+f(a\otimes bc)-f(a\otimes b)c=0\]
which, via the universal property of the tensor product gets us a $\mu$ satisfying condition 2-C, giving us
an infinitesimal derivation.

Similarly unfurling definitions gets us that $B^2$ are the 2-coboundaries.

\section{April 24, 2019}
Today we are going to talk about the equivalence (in all degrees) of $H^i(C(A))$ and $\Ext_A^i(A,A)=\HH^i(A)$.
On Friday, we well begin working on triangulated and derived categories and will spend about two weeks there. 
For a reference, you can read Weibel, but the lecture will likely not follow his exposition for the most part.

\subsection{The bar resolution}
\begin{defn}
Let $A$ be an algebra over $k$. The \textbf{bar resolution} of $A$ is a chain complex
($\deg\partial=-1$) which is defined as follows:
\[B(A)=B(A;A)=\left(\cdots A^{\otimes 4}\to A^{\otimes 3}\to A^{\otimes 2}\to A\to 0\to \cdots\right)\]
where the tensor power in the $i^{th}$ term is $i+2$, whenever that makes sense.

The differential is zero eventually, and otherwise is defined to be $\partial_N:A^{\otimes n+2}\to A^{\otimes n+1}$ via
\[\partial(a_0\otimes a_1\otimes \cdots\otimes a_{n+1})=\sum_{i=0}^n(-1)^ia_0\otimes\cdots\otimes a_ia_{i+1}\otimes\cdots\otimes a_{n+1}\]
\end{defn}
\begin{rmk}
	We could actually define this as a cochain, but it is pretty standard to use a chain instead.
\end{rmk}
\begin{rmk}
	The name ``bar resolution'' comes from the fact that the notation for these tensor was (traditionally)
	\[a_0\otimes\cdots\otimes a_{n+1}\sim[a_0|a_1|\cdots|a_{n+1}].\]
\end{rmk}
\begin{ex}
	Some computations: $\partial_0(a_0\otimes a_1)=a_0a_1$ and
	\[\partial(a_0\otimes a_1\otimes a_2)=a_0a_1\otimes a_2-a_0\otimes a_1a_2\]
	so the idea is we are ``generalizing multiplication'' to higher tensor powers.
\end{ex}
\begin{lem}
	$B(A)$ is a complex.
\end{lem}
\begin{lem}
	Consider each $A^{\otimes n+2}$ as an $(A,A)$-bimodule using the natural multiplication on each 
	factor. Then $\partial_n$ is an $(A,A)$-bimodule morphism.
\end{lem}
\begin{cor}
	This gives us that $B(A)$ is a complex of $(A,A)$-bimodules.
\end{cor}
\begin{prf}
	For any $x\in A$, we need to show that 
	\[\partial_n(x\cdots(a_0\otimes\cdots\otimes a_{n+1}))=x\cdots\partial_n(a_0\otimes\cdots\otimes a_{n+1})\]
	but this is pretty much immediate once you define it the right way:

	Basically, define the left action of $A$ on $A^{\otimes n}$ by multiplication (in $A$ on the leftmost factor)
	and the right action by multiplication on the rightmost factor. There is your structure and it is easy enough 
	to confirm is compatible with your differential.
\end{prf}
\begin{thm}
	There exist maps (note: not bimodule maps) $s_n:B(A)_n\to B(A)_{n+1}$ such that for all $n$,
	\[\id_{B(A)_n}=s_{n-1}\circ\partial_n+\partial_{n+1}\circ s_n\]
\end{thm}
\begin{prf}
	It is easy enough to confirm that the maps
	\[s_n(a_0\otimes\cdots\otimes a_{n+1})=1\otimes a_0\otimes \cdots\otimes a_{n+1}\]
	work.
\end{prf}
\begin{lem}
	$B(A)$ is exact.
\end{lem}
\begin{rmk}
	We've just shown that $\id_{B(A)}\sim 0$, so we're basically done. For a reminder: say $f\in \ker\partial_n$. 
	But then 
	\begin{align*}
		f&=\id_{B(A)_n}(f)\\
		&= s_{n-1}\partial_n(f)+\partial_{n+1}s_n(f)\\
		&= 0 + \partial_{n+1}(s_n(f))\in\operatorname{Im}\partial_{n+1}.
	\end{align*}
\end{rmk}
\begin{thm}
	$B(A)_{n\ge 0}$ is a free/projective resolution of the $A^e$ module $A$.
\end{thm}
\begin{prf}
	We have already seen that the $\delta$ are $A^e$ module maps and that this is an exact sequence.
	It remains to see that each $A^{\otimes n+2}$ is free. But
	\[A^{\otimes n+2}=A\otimes A^{\otimes n}\otimes A\cong A\otimes A\otimes A^{\otimes n}\cong A^e\otimes A^{\otimes n}\]
	and so this is free over the $k$-basis for the $A^{\otimes n}$ term.
\end{prf}

\begin{cor}
	$\Ext_{A^e}^i(A,A)=H^i(\Hom_{\Ch(A^e\text{-}\mathbf{mod}}(B(A)_{n\ge 0},(B(A)_{n\ge 0})))=H^i(\Hom_{\Ch}(B(A)_{n\ge 0},A))$
\end{cor}
\begin{thm}
	$\Hom_{\Ch}(B(A)_{n\ge 0},A)\cong C(A)$
\end{thm}
\begin{prf}
	The tricky part here is defining $\partial_{\Hom_{\Ch}(B(A),A)}:f\mapsto f\circ\partial_{B(A)}$ and then doing the computation. But recall that
	$\partial_{B(A)}$ and $\partial_{C(A)}$ are almost the same except the first and last terms. Well, use the
	$(A,A)$-bimodule structure to pull out $a_1$ and $a_{n+1}$ and you get what you want!
\end{prf}
\begin{cor}
	$H^i(C(A))\cong\Ext_{A^e}^i(A,A)$
\end{cor}

\subsection{The more general problem}
\begin{lem}
	Let $M$ be a (left) $A$ module. Then $B(A)_{n\ge 0}\otimes_A M$ is a free resolution of $M$.
\end{lem}
\begin{rmk}
	We could have just as easily have used right module structure by swapping the tensor.
\end{rmk}
\begin{rmk}
	This gives us a very nice thing -- we can get a free resolution of any module! But unfortunately this 
	is quite a terrible resolution to use if you want to do any computations.
\end{rmk}

\section{April 26, 2019}
Finally we are talking about the thing I named these notes after: triangulated categories!
I was late, but I don't believe that James has done the axioms yet.
\begin{rmk}
	It is an open question (Julia also mentioned this) about whether the octahedral axiom is really required for a triangulated category.
\end{rmk}
\subsection{Triangulated Categories}
Some motivation (from Grothendieck): Let $X$ be a projective scheme of dimension $d$ and dualizing 
sheaf $\omega_X$. Let $M$ be a coherent sheaf on $X$. Then Serre duality gets us that
\[H_X^i(M)^*\cong H^{d-i}_X(\omega_x\otimes M)\]
for all $i$ (as long as $X$ is Cohen-Macaulay).

But when $X$ is NOT CM, what can we do? Well there is a complex $\Omega_X$ in the bounded derived category of coherent sheaves over $X$
such that (in what is now called Grothendieck duality)
\[H_X^i(M)^*\cong H^{d-i}_X(\Omega_X\otimes M).\]

A triangulated (or derived) category is the place where such an $\Omega_X$ lives.
\begin{defn}
	A \textbf{triangle} in $\cal C$ is a diagram in $\cal C$:
	\[X\to Y\to Z\to\Sigma X.\]
\end{defn}
\begin{defn}
	A \textbf{morphism of triangles} is a commutative diagram
	\begin{center}
		\begin{tikzcd}
			X\ar[r,"u"]\ar[d,"f"]& Y\ar[r,"v"]\ar[d,"g"] & Z\ar[r,"w"]\ar[d,"h"] & \Sigma X\ar[d,"\Sigma f"]\\
			X'\ar[r] & Y'\ar[r] & Z'\ar[r] & \Sigma X'
		\end{tikzcd}
	\end{center}
\end{defn}
\begin{lem}
	The collection of all triangles in $\cal C$ with triangle morphisms form a $k$-linear additive category denoted $\Delta(\cal C)$
\end{lem}
\begin{rmk}
	By forgetting all the information except the $X\to Y$ (resp. just $X$) we can define a forgetful
	functor $\Delta(\cal C)\to\mathbf{Mor}(\cal C)$ (resp. $\Delta(\cal C)\to\cal C$). Our goal is to define a map in 
	the reverse direction (an adjoint?).
\end{rmk}
\begin{defn}
	The \textbf{rotation} of a triangle $X\xrightarrow{u} Y\to Z\to \Sigma X$ is $Y\to Z\to \Sigma X\xrightarrow{-\Sigma u}\Sigma Y.$
	We may be using $\mathscr R(\Delta)$ for the rotation of $\Delta$.
\end{defn}

\begin{defn}
	Let $\cal C$ be a $k$-linear additive category with a \textbf{suspension} or \textbf{shift} functor (auto-equivalence, actually)
	\[\Sigma:\cal C\to\cal C.\]

	Then $({\cal C}, \Sigma ,D)$ is \textbf{pre-triangulated} category where $D$ is a full, nonempty subcategory $D\subseteq\Delta(\cal C)$
	with shift functor $\Sigma$ of $D$ and we have the following axioms:
	\begin{itemize}
		\item (\textbf{TR0}) $X\xrightarrow{\id} X\to 0$ is in $D$ for each $X\in \cal C$ and furthermore $D$ is closed under both shifts and triangle isomorphisms.
		\item (\textbf{TR1}) [\textit{Mapping Cone Axiom}] For any $f:X\to Y$ in $\cal C$, there is a triangle
		\[X\xrightarrow{f} Y\to Z\to\Sigma X\]
		\item (\textbf{TR2}) [\textit{Rotation Axiom}] If $F\in D$, then $\mathscr R(F),\mathscr R^{-1}(F)\in D$
		\item (\textbf{TR3}) [\textit{Morphism Axiom}] Given two triangles 
		\begin{center}
			\begin{tikzcd}
				X\ar[r]\ar[d,"f"] & Y\ar[r]\ar[d,"g"] & Z\ar[r] & \Sigma X\ar[d,"\Sigma f"]\\
				X'\ar[r] & Y'\ar[r] & Z'\ar[r] & \Sigma X'
			\end{tikzcd}
		\end{center}
		with maps $f$ and $g$, there exists an $h:Z\to Z'$. I am unclear about whether the resulting diagram should commute.
	\end{itemize}
\end{defn}
\begin{defn}
	A triple $(\cal C,\Sigma, D)$ is a \textbf{triangulated category} if it is pre-triangulated and, in addition,
	satisfies the axiom \textbf{TR4} (Verdier/octahedral axiom):

	Suppose we are given three triangles: $X\xrightarrow{u}Y\to Z'\to\Sigma X$, $Y\xrightarrow{v}Z\to X'\to\Sigma Y$
	and $X\xrightarrow{v\circ u} Z\to Y'\to\Sigma X$. Then there is a triangle $Z'\to Y'\to X'\to\Sigma Z'$ such that
	\begin{center}
		\begin{tikzcd}
			X\ar[dr,"u"]\ar[rr,"v\circ u",bend left] && Z\ar[rr,bend left]\ar[dr] && X'\ar[rr,dashed,bend left]\ar[dr] && \Sigma Z'\\
			& Y\ar[ur,"v"]\ar[dr] && Y'\ar[ur,dashed]\ar[dr] && \Sigma Y\ar[ur] &\\
			&& Z'\ar[rr,bend right]\ar[ur,dashed] && \Sigma X\ar[ur] &&
		\end{tikzcd}
	\end{center}
\end{defn}
\begin{rmk}
	Notice that the map constructing a triangle for any morphism is not well-defined but it exists. Furthermore
	it is far from being functorial. This is one of the qualms people have with the definition of these categories.
	Grothendieck attempted to fix this in an unpublished paper on ``derivators'' but it is significantly more complicated.
\end{rmk}
\begin{rmk}
	Topologists generally drop the $k$-linearity in $\cal C$. James is unsure if it actually ever comes up.
\end{rmk}
\begin{rmk}
	The model category you should be thinking about is $H^0(\Ch(\modR))$.
\end{rmk}

\section{April 29, 2019}
Recall (notice?) that, given maps $f:X\to X'$ and $g:Y\to Y'$, using the morphism axiom,
we can define a morphism of triangles by just this data.

We have a couple other diagrams equivalent to the octahedral diagram (with slightly different information):
\begin{center}
	\begin{tikzcd}
		X\ar[r,"u"]\ar[d,"\sim"] & Y\ar[r]\ar[d,"v"] & Z'\ar[r]\ar[d,dashed] &\Sigma X\ar[d,"\sim"]\\
		X\ar[r,"v\circ u"] & Z\ar[r]\ar[d] & Y'\ar[r]\ar[d,dashed] &\Sigma X\ar[d,"\Sigma u"]\\
		& X'\ar[r,"\sim"] \ar[d] & X\ar[r]\ar[d,dashed] & \Sigma Y\\
		&\Sigma Y\ar[r] & \Sigma Z' &
	\end{tikzcd}
\end{center}

\begin{center}
	\begin{tikzcd}
		& & X\ar[dl,"u"]\ar[dd,"v\circ u"] &\\
		& Y\ar[dr,"v"]\ar[ddl] & &\\
		& & Z\ar[d]\ar[dr] &\\
		Z'\ar[rr,dashed] & & Y'\ar[r,dashed] & X'
	\end{tikzcd}
\end{center}

\begin{ex}
	Let ${\cal C}=K(\modR)=H^0(\Ch(\modR))$. This is a triangulated category. It is easy to check that $\cal C$ is $k$-linear additive.

	$\Sigma$ is given by shifting of complexes and is an auto-equivalence. $D$ is the full subcategory containing all triangles that are
	isomorphic to $\cone(f)$ where 
	\[X\xrightarrow{f} Y\to \cone(f)\to \Sigma X\]
	is an object such that $\cone(f)^i=(\Sigma X\oplus Y)^i$ and 
	\[\partial_{\cone}:(sm)+(n)\mapsto (-s\partial (m))+(f(m)+\partial(m)).\]
	Notice that there is some collision of notation: $\cone(f)$ can refer either to the complex or the triangle above.

	Then we need to find a chain homotopy to establish that $X\to X\to \cone(\id_X)\to\Sigma X$ is isomorphic to $X\to X\to 0\to\Sigma X$ 
	to show that $D$ satisfies (TR0).

	To prove (TR1), let $u:X\to Y$ be any morphism. Then we have the triangle
	\[X\xrightarrow{u} Y\to\cone(u)\to\Sigma(X)\]
	which gives us the axiom.

	For (TR2): start with a triangle $X\to Y\to Z\to \Sigma X=F$ we want to show that $F\in D$ if and only if $\mathcal{R}(F)\in D$.
	Since all our triangles are isomorphic to the mapping cone, we can just assume that $F$ is the mapping cone of $X\xrightarrow{u}Y$.
	To show $\mathcal{R}(F)$ is in $D$, we want to show that 
	\[(Y\xrightarrow{i}\cone(u)\to\Sigma X\to \Sigma Y)\cong (Y\xrightarrow{i}\cone(u)\to\cone(i)\to \Sigma Y)\]
	The idea here is that you notice that $\cone(\id_Y)=\Sigma Y\oplus Y$ is nullhomotopic. Then you get your copy of $\Sigma X$
	since $\cone(i)=\Sigma Y\oplus(\Sigma X\oplus Y)$ (with the appropriate differential).

	For (TR3), if we have $a:X\to X'$ and $b:Y\to Y'$ and $u:X\to Y$ and $u':X'\to Y'$, then we can define the 
	map $\Sigma a+b:\cone(u)\to\cone(u')$ by 
	\[\Sigma a+b(sm+n)=sa(m)+b(n).\]
	This map is natural and makes the diagram commute so we're golden.

	(TR4) is a bit of a pain. But given the triangles 
	\[X\xrightarrow{u}Y\to Z'\to \Sigma X,\quad X\xrightarrow{v\circ u}Z\to Y'\to \Sigma X,\quad Y\xrightarrow{v}Z\to X'\to\Sigma Y\]
	we can again assume that these are mapping cones. That this works is slightly less trivial because of the dimensionality of the 
	diagrams necessary, but can be done.

	So here we assume $Z'=\cone(u)$, $Y'=\cone(v\circ u)$ and $X'=\cone(v)$. This gives us the diagram
	\begin{center}
		\begin{tikzcd}
			x\ar[r,"u"]\ar[d,"\sim"] & Y\ar[r]\ar[d,"v"] & \cone(u)\ar[r]\ar[d,dashed,"\id_{\Sigma X}\oplus v"] &\Sigma X\ar[d,"\sim"]\\
			X\ar[r,"v\circ u"] & Z\ar[r]\ar[d] & \cone(v\circ u)\ar[r]\ar[d,dashed,"\Sigma u+\id_Z"] &\Sigma X\ar[d,"\Sigma u"]\\
			& \cone(v)\ar[r,"\sim"] \ar[d] & \cone(v)\ar[r]\ar[d,dashed,"\pi_{\Sigma Y}"] & \Sigma Y\\
			&\Sigma Y\ar[r] & \Sigma Z'=\Sigma^2X\oplus \Sigma Y &
		\end{tikzcd}
	\end{center}
\end{ex}
\begin{prob}
	Fill in the details of the above proof for (TR2) (which is surprisingly similar to the ``fake'' proof above.)
\end{prob}

\begin{defn}
	We define $\Ch^b(\modR)$, $\Ch^+(\modR)$ and $\Ch^-(\modR)$ to be the subcategories of $\Ch(\modR)$ consisting 
	of the (totally) bounded, bounded below, and bounded above chains.
\end{defn}

\begin{prop}
	$H^0(\Ch^x(\modR))=K^x(\modR)$ are triangulated categories for $x\in\{b,+,0\}$. 
\end{prop}
\begin{rmk}
	People refer to these as 
	subcategories. Even though this isn't strictly true, it most works out.
\end{rmk}
\begin{rmk}
	It ends up that you can just pop different properties on your modules (e.g. finite dimensional, projective injective, flat)
	are all preserved by your shift functor and other operations, so these are also examples. ``It works with any 
	property you can think of.'' --James
\end{rmk}
\subsection{Properties of Triangulated Categories}
\begin{prop}
	IF $X\xrightarrow{u}Y\xrightarrow{v} Z\xrightarrow{w}\Sigma X\in D$ then $v\circ u=0$.
\end{prop}
\begin{prf}
	Consider the diagram
	\begin{center}
		\begin{tikzcd}
			X\ar[r,"\id_X"]\ar[d,"\id_X"] & X\ar[r]\ar[d,"u"] & 0\ar[d,dashed,"0"]\ar[r] & \Sigma X\ar[d,"\Sigma(\id_X)"]\\
			X\ar[r,"u"] & Y\ar[r,"v"] & Z\ar[r] &\Sigma X
		\end{tikzcd}
	\end{center}
	where both the above triangles are in $D$, so we can apply (TR3) to get the existence of the dashed map above, 
	and then by commutativity $v\circ u=0.$
\end{prf}
\begin{prop}
	If $(\cal C,\Sigma, D)$ is a triangulated category, then so is $(\cal{C}^{op},\Sigma^{-1},D^{op})$.
\end{prop}

\section{May 4, 2019}
Missed this day!

\section{May 6, 2019}
Today we will discuss 3 kinds of localizations. We will construct the derived category $D(\modR)$ via localization.

\subsection{Localization}
\begin{defn}
	Let $\cal C$ be a category. Let $S$ be a collection of morphisms in $\cal C$. The localization of $\cal C$ with respect
	to $S$ is a category $S^{-1}\cal C$ together with a quotient map $\pi:\cal{C}\to S^{-1}\cal C$ such that 
	\begin{enumerate}
		\item The objects in $S^{-1}\cal C$ are the same as those in $\cal C$.
		\item For any $s\in S$, $\pi(s)$ is invertible (an isomorphism).
		\item For any functor $F:\cal{C}\to\cal D$ such that $F(s)$ is invertible for all $s\in S$, $F$ factors through $S^{-1}\cal C$.
	\end{enumerate}
\end{defn}
\begin{ex}
	If $G$ is a semigroup and $\cal C$ is a one-object category whose morphisms are the elements in $G$, then if 
	we let $S=G$, then $S^{-1}\cal C$ is the category with one object whose morphisms are $\langle G\rangle$.
\end{ex}
\begin{ex}
	If $\cal C$ is the one-object category for $R$, a $k$-algebra, then we can pick a (multiplicative) subset $S$ of $R$ 
	to get $S^{-1}\cal C$ is essentially $R_S$.
\end{ex}
\begin{ex}
	$K(\modR)$ is a localization of $\Ch(\modR)$.
\end{ex}
\begin{rmk}
	Ignoring set-theoretic issues, $S^{-1}\cal C$ always exists. If we want to fix any such problems, we 
	can just work over a (locally) small category. Generally if $S$ is a set we are okay, but we will see 
	soon that we will want to include all our identity maps! So we had better have a small $\cal C$.
\end{rmk}

\subsection{Ore Localization}
\begin{defn}
	A collection $S$ of morphisms in $\cal C$ is called a \textbf{multiplicative system} 
	if
	\begin{itemize}
		\item $S$ is closed under composition and contains $\id_X$ for all $X\in\cal C$.
		\item (\textbf{Right Ore Condition}) If $t:Z\to Y$ is in $S$ and $g:X\to Y$ in $\cal C$,
		then there are maps $f$ and $s$ and a commutative diagram
		\begin{center}
			\begin{tikzcd}
				W\ar[r,"f",dashed]\ar[d,"s",dashed] & Z\ar[d,"t"]\\
				X\ar[r,"g"] & Y
			\end{tikzcd}
		\end{center}
		such that $s\in S$.
		\item (\textbf{Right Ore Condition}) $\mathcal{C}^{op}$ has the right Ore condition.
		\item (\textbf{Cancellation}) If $f,g:X\to Y$ are two morphisms, then the following are equivalent:
		\begin{itemize}
			\item $sf=sg$ for some $s:Y\to Z$ in $S$
			\item $ft=gt$ for some $t:V\to X$ in $S$
		\end{itemize}
	\end{itemize}
\end{defn}
\begin{defn}
	If $S$ is a multiplicative system, then $S^{-1}\cal C$ is called an \textbf{Ore localization} of $\cal C$.
\end{defn}
\begin{thm}
	Let $S$ be a multiplicative system in $\cal C$. Then the $\Hom$ in $S^{-1}\cal C$ can be described 
	as
	\[\Hom_{S^{-1}\cal C}(X,Y)=\{X\xleftarrow{s}Z\xrightarrow{f}Y| s\in S\}/\sim=\{(s,f)|s\in S\}/\sim\]
	where the equivalence $\sim$ is defined as follows: $(s_1,f_1)\sim(s_2,f_2)$ if there exists $(s_3,f_3)$
	and $a_1,a_2\in S$ such that the following commutes:
	\begin{center}
		\begin{tikzcd}
			& Z_1\ar[dl,"s_1"]\ar[dr,"f_1"] &\\
			X & Z_3\ar[u,"a_1"]\ar[d,"a_2"]\ar[l,"s_3"]\ar[r,"f_3"] & Y\\
			& Z_2\ar[ul,"s_2"]\ar[ur,"f_2"] &
		\end{tikzcd}
	\end{center}
	In this case, the equivalence class of $(s,f)$ is denoted $[fs^{-1}]$ or $[f/s]$.
\end{thm}
\begin{rmk}
	Dually we can characterize $\Hom$ using diagrams $X\xrightarrow{g} Z\xleftarrow{t} Y$ for $t\in S$. These are
	equivalent but we have to use the left Ore condition instead of right to do composition.
\end{rmk}

So then we get a functor $\pi:\mathcal{C}\to S^{-1}\mathcal{C}$ that sends $X\mapsto X$ and $(f:X\to Y)\mapsto (X\xleftarrow{\id_X}X\xrightarrow{f}Y)$.
Composition is just what you'd expect, using the ore conditions. Checking that composition is well-defined (modulo the equivalence relation)
seems like a bit of a chore, but maybe worth checking.

\subsection{Verdier Localization}
Let $(\mathcal{C}, \Sigma, D)$ be a triangulated category. let $\cal T$ be a full triangulated subcategory of $\cal C$.
Then $\mathcal{C}/\mathcal{T}$ is the category such that $X\simeq 0$ for all $X\in\mathcal{T}$.

Recall
\begin{prop}
	If $X\xrightarrow{f} Y\to Z\to\Sigma X$ is a triangle, $f$ is an isomorphism if and only if $Z\simeq 0$.
\end{prop}

Set $S_{\cal T}$ to be $\{f|\cone(f)\in\mathcal{T}\}$. Then 
\begin{lem}
	$\Sigma S_{\mathcal{T}}=S_{\mathcal{T}}$.
\end{lem}
\begin{lem}
	$S_{\mathcal{T}}$ is a multiplicative system.
\end{lem}
\begin{thm}
	$S_{\mathcal{T}}^{-1}\mathcal{C}$ is a triangulated category.
\end{thm}
\begin{rmk}
	James may or may not prove this on Wednesday. Some sketches: $\Sigma_{S^{-1}\cal C}$ sends $X\to\Sigma X$ and $[f/s]$ to $[(\Sigma f)/(\Sigma s)]$.

	Then $D_{S^{-1}\cal C}$ is the set of all triangles in $\Delta(S^{-1}\cal C)$ that are isomorphic to some $X\to Y\to Z\to\Sigma X$ in $D(\mathcal{C})$.
\end{rmk}

\section{May 8, 2019}
Today we will discuss examples of Verdier localization. First some motivation for it: in commutative algebra or in algebraic 
geometry, there is always a concept of localization and this (along with ``local-global principles'') gives us a nice 
way to get a handle on things.

In a more general case, we no longer have localization! So we don't have hope of chopping up things into local pieces.
So instead we can move to a triangulated category! So if we are trying to study $\modR$ for some algebra $R$,
we can instead consider $D(\modR)$ where we can localize by any subcategory we choose!

\subsection{Examples of Verdier Localization}
As a convention, let $(\mathcal{C},\Sigma,D)$ be a triangulated category and let $\mathcal{T}$ be a full triangulated subcategory
such that if $Y\in\mathcal T$ and $X\cong Y$, then $X\in\mathcal T$ ($\mathcal T$ is closed under isomorphism). Furthermore, $\mathcal T$ is 
triangulated via $(\mathcal{T}, \Sigma|_{\mathcal T}, D\cap\Delta(\mathcal{T}))$.

Now let $P$ be some property. First, we will say $P$ is a property on objects.

\begin{ex}
	Let $\mathcal C=K(\modR)$. Let $P(C)$ mean that $C$ is a (bounded/bounded above/bounded below) complex of 
	(projective/injective/flat/f.g./some mixture). Then take $\mathcal T=\{X|\exists Y: Y\cong X,P(Y)\}$. We need to check that $P$ is 
	stable under suspension and that the set of $P$-objects is closed under cone constructions.

	In this case since the cone is just a coproduct, it suffices to show that this preserved under coproducts.
\end{ex}

Next $P$ is a property on morphisms:
\begin{ex}
	Now let $X$ be a $P$ object if $\Hom_{\calC}(\Sigma^nX,-)$ commutes with coproducts for all $n\in\bbZ$. Alternatively, we could 
	let the $P$ objects be those $Y$ such that $\Hom_\calC(\Sigma^nX,Y)=0$ for all $n\in\bbZ$ for some fixed $X$.

	Then we can take the objects of $\calT$ to be the objects in $\calC$ isomorphic to a $P$ object. Once one proves that these  
	objects are stable under suspension and cone constructions, $\calC/\calT$ exists.
\end{ex}
\begin{lem}
	$\Hom_{\Ch(\modR)}(R,X)=X$.
\end{lem}
\begin{prf}
	Look at the $i^{th}$ component of each:
	\[\Hom_{\Ch(\modR)}(R,X)^i=\prod_{s\in\bbZ}\Hom_R(R^s,X^{s+i})=\Hom_R(R^0,X^i)=X^i\]
	It remains to show the differentials are the same, but it can be shown (HW).
\end{prf}
\begin{lem}
	If $X$ and $Y$ are acyclic complexes and $X\xrightarrow{f}Y\to Z\to\Sigma X$ is a triangle, then $Z$ is acyclic.
\end{lem}
\begin{rmk}
	To prove this, one applies $\Hom_\calC(\Sigma^n R,-)$ to the triangle and uses that $\Hom_\calC(\Sigma^n R,\Sigma X)=\Hom_{\calC}(\Sigma^{n-1}R,X)=0$.
\end{rmk}
\begin{ex}
	Let $\calC=K(\modR)$. Let the $P$ objects to be the $X$ such that $\Hom_\calC(\Sigma^n R,X)=0$. Now notice 
	\[\Hom_\calC(\Sigma^0R,X)=\Hom_{H^0(\Ch(\modR))}(R,X)=H^0(\Hom_{\Ch(\modR)}(R,X))=H^0(X)\]
	and (a claim) $\Hom_{\calC}(\Sigma^nR,X)=H^n(X)=0$.

	Thus $X$ has $P$ if and only if $H^n(X)=0$ for all $n$ -- that is, if $X$ is \textbf{acyclic.}
\end{ex}
\begin{defn}
	The \textbf{derived category of complexes of right $R$ modules} is 
	\[D(\modR):=K(\modR)/K_{\text{acyclic}}(\modR)\]
\end{defn}
\begin{rmk}
	There are some useful subcategories $D^b(\modR)$, the bounded derived category, as well as $D^+(\modR)$ and $D^-(\modR)$,
	where are the bounded (resp.) below and above complexes. Here one applies the property to the homotopy category before taking the quotient. 
	One also considers $D_{f.g.}^b(\modR)$ where one puts the finite generation 
	condition on $\modR$ first, then takes the quotient of the appropriate homotopy (sub) categories.
\end{rmk}

\begin{ex}
	Let $\scrD=D^b_{f.d.}(\modR)$ and let $P$ be the complexes of finitely generated projective modules (i.e. \textbf{perfect complexes}).
	Then $\calT_P:=D_{per}^b(\modR)$, a sub triangulated category of $\scrD$. Then 
	\[D_{sing}(\modR):=D^b_{f.g.}(\modR)/D_{per}^b(\modR)\]
	is called the \textbf{singular category} and corresponds to the singularities of $\operatorname{spec} R$.
\end{ex}
\begin{rmk}
	If $\gldim R <\infty$ and $R$ is Noetherian, then $D_{sing}(R)=0$.
\end{rmk}

\section{May 10, 2019}
Today will be largely random remarks. There will continue to be holes in proofs. :)

Last time, we let $R$ be an algebra and considered the derived category $D(\modR)$, which is a Verdier localization of $K(\modR)$
or (somehow equivalently) a quotient. I think he is playing a bit fast and loose here.

Note that the mapping cones (the triangles) are precisely the quasi-isomorphisms. This can be seen by applying $H^0=\Hom_K(R,-)$
to $\Sigma^{-1}\cone(f)\to X\xrightarrow{f} Y\to \cone(f)$
\begin{rmk}
	If $S$ is the set of quasi-isomorphisms in $\Ch(\modR)$, then 
	\[S^{-1}\Ch(\modR)\cong D(\modR),\] 
	but this is just (for now?) as $k$-linear categories, since $\Ch(\modR)$ is not triangulated (it's abelian!).
\end{rmk}
\begin{rmk}
	There exists a fully faithful embedding $\modR$ into $D(\modR)$ where we have $\Hom_R(M,N)=\Hom_{D(\modR)}(M,N)$.
\end{rmk}
\begin{rmk}
	$\Ext_R^i(M,N)=\Hom_{D(\modR)}(M,\Sigma^iN)$. This can be useful for computations, but recall that composition in the Verdier localization is 
	all about them roof constructions.
\end{rmk}
\begin{rmk}
	Let $P_\bullet\to N$ be a projective resolution. Then $P_\bullet$ is quasi-isomorphic to $N$, so they become (honestly) isomorphic in $D(\modR)$
	even though $P_\bullet\not\cong N$ in $K(\modR)$.
\end{rmk}
\begin{rmk}
	Suppose $X$ is a bounded-above complex of projectives. Then 
	\[\Hom_{D(\modR)}(X,Y)=\{X\to Y\}/\sim = \Hom_{K(\modR)}(X,Y)=H^0(\Hom_{\Ch(\modR)}(X,Y)).\]
	This motivates somewhat why we often want to have projective resolutions of our objects -- they become more computationally feasible 
	in this case.
\end{rmk}
\begin{rmk}
	$K^{-}(\mathbf{projmod}\text{-}R)\cong D^-(\modR)$. In other words, (recall we have a fully faithful embedding of $\modR$ into $D(\modR)$)
	there is a fully faithful embedding of $\modR$ into $K^-(\mathbf{projmod}\text{-}R)$ sending each $M$ to 
	a projective resolution of $M$.

	Dually, $D^+(\modR)$ is equivalent to $K^+(\mathbf{projmod}\text{-}R)$, so $\modR$ embeds fully faithfully into it.
\end{rmk}
\begin{rmk}
	In particular, $D^b(\mod-R)$ embeds into either $K^+(\mathbf{injmod}\text{-}R)$ or $K^-(\mathbf{injmod}\text{-}R)$.
\end{rmk}

\subsection{Derived Functors}
Consider algebras $A$ and $B$. Let $F$ be an additive $k$-linear functor from $A$ modules to $B$ modules. This easily extends 
to a functor between the chain and homotopy categories of these categories. But \textbf{there is not a clear extension of $F$ to the derived category.}

Consider the composition of maps (and embeddings):
\[D^-(\mathbf{mod}\text{-}A)\hookrightarrow K^-(\mathbf{projmod}\text{-}A)\xrightarrow{F} K^-(\mathbf{mod}\text{-}B)\xrightarrow{\nu} D^-(\mathbf{mod}\text{-}B)\]
where $\nu$ is the Verdier localization.

Then the composition of these maps is called the \textbf{left derived functor of $F$}, which is written $LF$.
\begin{ex}
	Let $F=-\otimes_R N:\modR\to\Vectk$. Then $H^i(LF(M))=\Tor_{-i}^A(M,N)$.
\end{ex}
Next week we are going to discuss dualizing complexes. From these you can prove some very nice results 
that otherwise require some complicated argument.

\section{May 13, 2019}
\subsection{Dualizing Complexes}
Today we will start with definitions and an introduction to the idea.

Grothendieck introduced dualizing complexes (over schemes) in the 1960's. Using these complexes, 
he proved what was called ``Grothendieck duality'', a generalization of Serre duality. In 1990(ish)
Yekutieli introduced them in the noncommutative case.

This is a powerful tool in noncommutative/homological algebra. Even the existence of dualizing 
complexes has many consequences:
\begin{thm}
	Take any simple factor ring of the universal enveloping algebra $\calU(\frakg)$ ($\frakg$ finite dimensional) is Auslander-Gorenstein and Cohen-MacCaulay.
\end{thm}
\begin{thm}[Bass Theorem (for Hopf algebras)]
	Let $H$ be a Noetherian Hopf algebra satisfying a polynomial identity. let $M$ be a finitely generated (left) $H$-module.
	Then $\injdim_H M$ is either infinite or equal to $\injdim_H H$.
\end{thm}
\begin{thm}[Auslander-Buchsbaum]
	Let $A$ be a local Noetherian algebra with balanced dualizing complex. Let $M$ be a finitely generated (left) $A$-module.
	Then $\projdim M$ is either infinite or is $\operatorname{depth}(A)-\operatorname{depth}(M)\le \operatorname{depth}(A)$.
\end{thm}

\subsection{Conventions}
We will be working a lot in $\Algk,\modR,$ and $\Ch$ over the next couple weeks. We should fix 
some conventions to work with.

\begin{itemize}
	\item Let $A,B,C$ be Noetherian $k$-algebras.
	\item Usually $R$ will stand for a dualizing complex (Except directly below where my notation varied).
	\item Let $M$ and $N$ be left (or right) modules over $A$.
	\item Let $X,Y,Z$ be complexes.
	\item The category of $(A,B)$-bimodules is identified with left $A\otimes B^{op}$ modules.
	\item Usually we work with left $A$-modules. To study the right $A$-modules, we study the left $A^{op}$ modules.
	\item We want to identify $\Db(\Rmod)$ with a subcategory of $\D(\modR)$, so notice
	\[\Db(\Rmod)\cong \{X\in \D(\modR)|X\simeq\text{ some object in } \Db(\Rmod)\}\]
	where the isomorphism there is on the level of triangulated categories.
\end{itemize}
\begin{prop}
	If $X\simeq Y$ for some bounded complex, then $H^n(X)=0$ for all $n>>0$. The converse is also true.
\end{prop}

This is a slightly nonstandard definition of projective dimension, but it is the one we will use in this class:
\begin{defn}
	Let $X\in\Db(\Rmod).$ Then 
	\[\projdim X=\max\{i|\Ext_R^i(X,M)\ne 0,\text{ some } M\in\Rmod\}\]
\end{defn}
The upshot here is
\begin{lem}
	$\projdim \Sigma X=\projdim X+1$ and therefore $\projdim\Sigma^n X=\projdim X+n.$
\end{lem}
\begin{lem}
	Let $d=\projdim X<\infty$. Then 
	\[X\simeq (\cdots 0\to P^{-d}\to\cdots\to P^s\to 0)\]
	with all $P^i$ projective.
\end{lem}

Then define $\injdim X$ dually.

Recall that for bounded complexes $X$ and $Y$ of $R$-modules, 
\[R\Hom_R(X,Y):=\Hom_{\Ch(\Rmod)}(X,I_Y)=\Hom_{\Ch(\Rmod)}(P_X,Y)\in \D(\Vectk)\]
\begin{lem}
	Let $M$ be a left $R=A\otimes B^{op}$ module.
	\begin{enumerate}
		\item If $M$ is free over $R$, then it is free over both $A$ and $B$.
		\item If $M$ is projective over $R$, then it is projective over both $A$ and $B$.
		\item If $M$ is flat over $R$, then it is flat over both $A$ and $B$.
		\item If $M$ is injective over $R$, then it is injective over both $A$ and $B$.
	\end{enumerate}
\end{lem}
\begin{prf}
	(a): If $M\cong (A\otimes B^{op})^n$, then $A\otimes B^{op}\cong A^{\dim B^{op}}$, so $M\cong A^{n\dim B^{op}}$.

	(b): pretty easy.

	(c) and (d): a little more tough. Use adjoints.
\end{prf}
\begin{lem}
	Let $M$ and $N$ be $(A,B)$-bimodules. Then $\Hom_A(M,N)$ is a $(B,B)$-bimodule. Furthermore, $\Hom_{B^{op}}(M,N)$ is an $(A,A)$-bimodule.
\end{lem}
\begin{lem}
	Let $X,Y$ be complexes of $(A,B)$-bimodules. Then $\Hom_{\Ch(\Amod)}\in\Ch(B^e\text{-}\mathbf{mod})$.
	Also flip it and reverse it.
\end{lem}
\begin{lem}
	Let $X$ and $Y$ be complexes of $(A,B)$-bimodules. Then $R\Hom_A(X,Y)\in \D(B\otimes B^{op}\text{-}\mathbf{mod})$ and flip it.
\end{lem}

\subsection{Dualizing Complexes}
\begin{defn}
	Let $A$ and $B$ be two $k$-algebras and let $R$ be a complex in $\Db(A\otimes B^{op}\text{-}\mathbf{mod})$.
	Then $R$ is called a \textbf{dualizing complex over $(A,B)$} if 
	\begin{itemize}
		\item $H^i(R)\in A\otimes B^{op}\text{-}\mathbf{mod}$ is finitely generated on both sides.
		\item $\injdim_AR,\injdim_BR<\infty$.
		\item The natural maps $A\xrightarrow{\Phi_A}R\Hom_B^{op}(R,R)$ and $B\xrightarrow{\Phi_B} R\Hom_A(R,R)$
		are isomorphisms in $\D(A^e\text{-}\mathbf{mod})$ and $\D(B^e\text{-}\mathbf{mod})$, respectively.
	\end{itemize}
\end{defn}
\begin{rmk}
	The map Recall that $R\Hom_{B^{op}}(R,R)=\Hom_{\Ch(B^{op}\text{-}\mathbf{mod})}(P_R,P_R)$. Then $\Phi_A$ sends 
	$a\mapsto l_a:P_R\to P_R$.

	One can verify this more easily when $A=B=R=k$.
\end{rmk}

\end{document}

### 2019-05-22 Homological Algebra Notes/Hom Alg.tex
update

@@
 
+\section{May 17, 2019}
+Recall that today we will have extra class at 5pm today. Cody is going to give his talk!
+
+\subsection{The Bar Resolution}
+We begin with a non-standard presentation of the Bar resolution. Consider 
+\[\cdots\to (\bbZ G)^2\to \bbZ G\to \bbZ\]
+where the differential takes $(g_0,\dots,g_n)\in(\bbZ G)^{n+1}$ to $\sum_i(-1)^i(g_0,\cdots,\widehat{g_i},\cdots,g_n)$.
+\begin{prop}
+	This is a free resolution of $\bbZ$ as a $\bbZ G$ module.
+\end{prop}
+\begin{prf}
+	It suffices to show that this is a resolution (i.e. that it is contractible). Thus we want to 
+	construct the homotopy:
+	\[s(g_0,\cdots,g_n)=(e,g_0,\cdots,g_n)\]
+	which amounts to taking an $n$-simplex and embedding it into an $(n+1)$-simplex as a face.
+
+	Then we can compute 
+	\begin{align*}
+		ds+sd(g_0,\cdots,g_n)&=d(e,g_1,\cdots,g_n)+s\left(\sum_i(-1)^i(g_0,\cdots,\widehat{g_i},\cdots,g_n)\right)\\
+		&= (g_0,\cdots,g_n)-\sum_{i}(-1)^i(e,g_1,\cdots,\widehat{g_i},\cdots,g_n)+s\left(\sum_i(-1)^i(g_0,\cdots,\widehat{g_i},\cdots,g_n)\right)\\
+		&= (g_0,\cdots,g_n)
+	\end{align*}
+	proving the statement.
+\end{prf}
+\begin{rmk}
+	Notice that $(\bbZ G)^{n+1}$ is a free $\bbZ G$ module of rank $|G|^n$ with basis 
+	\[[g_1|\cdots|g_n]=(e,g_1,g_1g_2,\dots,g_1\cdots g_n).\]
+\end{rmk}
+\begin{rmk}
+	We can define the differential on the basis elements:
+	\[d:=\sum_0^{n}(-1)^id_i:(\bbZ G)^{n+1}\to(\bbZ G)^n\]
+	where $d_i:\bbZ G^{n+1}\to\bbZ G^n$ is given by
+	\[d_i([g_1|\dots|g_n])=\left\{\begin{array}{lr}
+		g_1[g_2|\dots|g_n],& i=0\\
+		\left[g_1|\dots |g_ig_{i+1}|\dots|g_n \right], & i = n\\
+		\left[g_1|\dots|g_{n-1}\right], & i = n.
+	\end{array}\right.\]
+\end{rmk}
+\begin{defn}
+	If we set $B_n=\bbZ G^{n+1}$, then $(B_\bullet, d)$ is called the \textbf{Bar resolution} of $\bbZ$.
+\end{defn}
+\begin{defn}
+	If $D_\bullet$ is the subcomplex of $B_\bullet$ generated by all elements $[g_1|\dots|g_n]$ such that some $g_i=e$,
+	then the quotient $B_\bullet/D_\bullet=\overline{B}_\bullet$ is called the \textbf{normalized Bar complex.}
+\end{defn}
+\begin{prob}
+	Write down both the Bar and normalized Bar resolutions of $\bbZ$ using $G=\bbZ/2$ and compare.
+\end{prob}
+
+\subsection{Hochschild Complex for $H^\ast(G,M)$}
+Begin with the Bar resolution $B_\bullet\to \bbZ$ and apply $\Hom_G(-,M)$, giving us a cochain complex. It will help to notice 
+that $\Hom_G(B_n,M)\cong\Hom_{\bbZ G}(\bbZ G^{n+1},M)\cong\Hom_\bbZ(\bbZ G^n,M)$, which is equal to the set of 
+(set!) functions $f:G^n\to M$ extended $\bbZ$-linearly. 
+\begin{defn}
+	$C^n(G,M)$ is the set of all linear combinations of set maps $f:G^n\to M$.
+\end{defn}
+\begin{rmk}
+	Notice $C^n(G,M)\cong \Hom_G(\bbZ G^{n+1},M)$.
+\end{rmk}
+Then we can compute the induced differential $d_n:C^n(G,M)\to C^{n+1}(G,M)$ to be 
+\begin{align*}
+	(df)(g_1,\dots,g_{n+1})&- g_1f(g_2,\dots,g_{n+1})+\cdots+(-11)^if(g_1,\dots,g_ig_{i+1},\dots,g_{n+1})\\
+	&\qquad\qquad+\cdots+(-1)^nf(g_1,\dots,g_n)
+\end{align*}
+
+\begin{cor}
+	The Hochschild complex $C^\bullet$ computes $H^\ast(G,M).$
+\end{cor}
+
+\begin{defn}
+	Let $R$ be a ring, $M$ an $(R,R)$-bimodule. Then a map of abelian groups $d:R\to M$ is called 
+	a \textbf{derivation} if
+	\[d(rs)=rd(s)+d(r)s.\]
+\end{defn}
+\begin{defn}
+	If $M$ is a $\bbZ G$ (left) module and $d:\bbZ G\to M$ is such that 
+	\[rd(s)+d(r)\]
+	$d$ is called a \textbf{crossed homomorphism.}
+\end{defn}
+\begin{rmk}
+	This essentially arises by considering a derivation on a module where the right action is defined to be trivial! This is sort of what is happening
+	with the derivation in the Hochschild complex.
+\end{rmk}
+\begin{rmk}
+	Notice that $d^1f(g_1,g_2)=g_1f(g_2)-f(g_1g_2)+f(g_1)$. If $d^1f=0$ then we call $f$ a cocycle (as per usual).
+	So then the \textbf{cocycle condition on $C^1(G,M)$} is 
+	\[f(g_1g_2)=g_1f(g_2)+f(g_1)\]
+	or exactly the crossed homomorphisms/derivations.
+
+	This gives us a nice characterization:
+	\[H^1(G,M)\cong\{\text{crossed homomorphisms}\}/\{\text{principal crossed homomorphisms}\}.\]
+\end{rmk}
+
+\section{Talk: Lie Algebra Cohomology}
+\subsection{Geometric Notions}
+Let $G$ be a Lie group, a smooth manifold with smooth multiplication and inverse maps. Let $\frakg$ be $\operatorname{Lie}(\frakg)$
+And consider the complex $\Omega^\bullet(G),$ the complex of $k$-forms on $G$ with differential given by the exterior derivative.
+
+\begin{defn}
+	$H^\ast_{dR}(G)$ is just the homology of $\Omega^\bullet(G)$.
+\end{defn}
+
+\begin{defn}
+	Let $\Omega^\bullet_L(G)$ be the complex of left-invariant forms, $\{\omega\in\Omega^\ast(G)|L_g^\ast\omega=\omega\}$.
+\end{defn}
+\begin{rmk}
+	A key fact to showing that this is indeed a complex is that the exterior derivative commutes with pullbacks.
+\end{rmk}
+
+\begin{thm}
+	$\iota:\Omega_L^\bullet(G)\to\Omega^\bullet(G)$ is a quasi-isomorphism when $G$ is compact and connected.
+\end{thm}
+\begin{prf}
+	(Sketch) For injectivity, suppose $\iota^\ast[\omega]=0\in H_{dR}^\ast(G)$. Then $\omega=d\mu$ for some $\mu$.
+	Set $I(\omega)=\int_G L_g^\ast\omega dg$ (where this is called the Haar measure). But then after checking lots 
+	of fiddly facts, you can show 
+	\[\omega=I(\omega)=I(d\mu)=dI(\mu)\]
+
+	For sujectivity, notice that the natural pairing $\langle[z],[w]\rangle=\int_z w$ where $[z]$ is a class in the homology  and $[w]$ is in the cohomology,
+	then since any other representative for $[z]$ differs by a boundary the integral in unaffected and since any other representative in $[w]$ 
+	differs by a coboundary (exact form), the integral is also unafected by this.
+
+	Then there is some integration computations that go through lots of replacing things with other representatives (since $[z]=[L_g(z)]$ and $\iota^\ast[I(\omega)]=[\omega]$).
+\end{prf}
+
+For the next step we want to restrict to the identity. Notice that $\varepsilon:\Omega_L^\ast(G)\to(\wedge^\ast\frakg)^\ast$
+is a well-defined map (here we send each form to its restriction to the identity: $\varepsilon(\omega)=\omega_e$).
+
+Notice 
+\begin{align*}
+	\varepsilon(d\omega)(X_1|_e,\dots,X_n|_e)&=d\omega(X_1,\dots,X_n)(e)\\
+	&=\sum_1^nX_i(\omega(X_1,\dots,\hat X_i,\dots,X_n))(e)+\sum_{i<j}([X_i,X_j],X_1,\dots,\hat X_i,\dots,\hat X_j,\dots,X_n)(e)
+\end{align*}
+and since the first term in the sum is zero (why?) the latter part is exactly the differential we want:
+\[\delta\alpha(X_1,\dots,X_n)=\sum_{i<j}([X_i,X_j]_e,X_1|_e,\dots,\hat X_i|_e,\dots,\hat X_j|_e,\dots,X_n|_e)\]
+
+\begin{thm}
+	$H_L^\ast(G)\cong H^\ast(\frakg)$. Furthermore $H_{dR}^\ast(G)\cong H^\ast(\frakg)$.
+\end{thm}
+\begin{rmk}
+	Note that $H^\ast(\frakg)$ is the cohomlogy of the $(\wedge^\ast\frakg)^\ast$
+\end{rmk}
+\subsection{Algebraic Notions}
+Let $k$ be a commutative ring, $\frakg$ be a Lie algebra, and $\calU(\frakg)$ its universal enveloping algebra.
+
+Notice that there is a functor from $k$ modules to $\frakg$ modules that gives any module $M$ a trivial $\frakg$ action. Recall the definition 
+of invariants $(-)^\frakg$ ($M^\frakg=\{m\in M|xm=0\}$) and coinvariants $(-)_\frakg$ ($M_\frakg=M/\frakg M$).
+
+The plan of attack is to show that these functors have adjoints and thus are exact on one side, and then we can talk about derived functors.
+
+\begin{lem}
+	$(-)^\frakg$ is right adjoint to the trivial $\frakg$-module functor. $(-)_\frakg$ is left adjoint to the same functor.
+\end{lem}
+\begin{cor}
+	Thus $(-)^\frakg$ and $(-)_\frakg$ are left and right exact, respectively.
+\end{cor}
+\begin{lem}
+	Let $\frakg$ be a Lie algebra. Then we have an equivalence of categories (isomorphism actually!)
+	\[\frakg\text{-}\mathbf{mod}\cong \mathcal U\frakg\text{-}\mathbf{mod}\]
+\end{lem}
+\begin{rmk}
+	Since $\mathcal U\frakg$-$\mathbf{mod}$ has enough projectives, so does $\frakg$-$\mathbf{mod}$.
+\end{rmk}
+\begin{defn}
+	We write $H_\ast(\frakg,M)$ to be the left erived functor $L_\ast(-)_\frakg(M)$ and $H^\ast(\frakg,M)=R^\ast(-)^\frakg(M)$.
+	These are the (co)homology groups of $\frakg$ with coefficients in $M$.
+\end{defn}
+
+How does this relate to $\Ext$ and $\Tor$?
+\begin{rmk}
+	If $M$ is a $\frakg$-module, then we can give $M$ a $\calU\frakg$ action in the following way:
+	\[(x_1\otimes\cdots\otimes x_n)\cdot m=(x_1(x_2(\cdots(x_nm))))\]
+\end{rmk}
+\begin{thm}
+	sIf $M$ is a $\frakg$-module, then 
+	\[H_\ast(\frakg,M)=\Tor_\ast^{\calU\frakg}(k,M)\]
+	and 
+	\[H^\ast(\frakg,M)\cong\Ext_{\calU\frakg}^\ast(k,M).\]
+\end{thm}
+
+Here we need the PBW theorem, so we need to assume now that $\frakg$ is free as a $k$-module.
+\begin{defn}
+	Let $V_p(\frakg)=\calU\frakg\otimes_k\wedge^p\frakg$, where $V_0=\calU\frakg$ and $V_1=\calU\frakg\otimes\frakg$.
+
+	Let $\varepsilon:\calU\frakg\to k$ be the augmentation map. Let $d:V_1(\frakg)\to V_0(\frakg)$
+	to be the multiplication map $d(u\otimes x)=ux$ and more generally, 
+	\[d(u\otimes x_1\wedge\cdots\wedge x_n)=\theta_1+\theta_2\]
+	where
+	\[\theta_1=\sum_1^n(-1)^{i+1}(ux_1\wedge\cdots\wedge\hat x_i\wedge\cdots\wedge x_n)\]
+	and 
+	\[\theta_2=\sum_{i<j}(-1)^{i+j}u\otimes[x_i,x_j]\wedge x_1\wedge\cdots\wedge\hat x_i\wedge\cdots\wedge \hat x_j\wedge\cdots\wedge x_n\]
+
+	Then $V_\bullet(\frakg)$ with the above differential is called the \textbf{Chevalley-Eilenberg complex} for $\frakg$
+\end{defn}
+
+\begin{thm}
+	$V_\bullet(\frakg)\to k$ is a projective resolution.
+\end{thm}
+\begin{rmk}
+	This needs spectral sequences as well as the PBW theorem.
+\end{rmk}
+\begin{thm}
+	If $M$ is a left $\frakg$-module, then the cohomology modules $H^\ast(\frakg,M)$ are the cohomology of the complex 
+	\[\Hom_\frakg(V(\frakg),M)=\Hom_\frakg(\calU\frakg\otimes\wedge^\ast\frakg,M)\cong\Hom)k(\wedge^\ast\frakg,M)\]
+\end{thm}
+
+\section{May 20, 2019}
+Recall that we hat just said that $H^1(G,M)$ was crossed homomorphisms module principal crossed homomorphisms.
+
+Notice that when $M$ is trivial, a crossed homomorphism is precisely a group homomorphism and since the 
+principle ones are zero, 
+\[H^1(G,M)=\Hom_\Grp(G,M)=\Hom_\Ab(G/[G,G],M)\]
+\begin{cor}
+	If $G$ is perfect ($G=[G,G]$) then $H^1(G,\bbZ)=0$.
+\end{cor}
+\begin{rmk}
+	In fact, $H_1(G,\bbZ)=G/[G,G]$.
+\end{rmk}
+
+\subsection{$H^2$}
+Now let $A$ be an ableian group and $G$ module. Recall that 
+\[0\to A\to E\to G\to 1\]
+(a short exact sequence in $\Grp$) is called an extension of $G$ by $A$. Then we can show that (or let Weibel or Dummit \& Foote show us)
+\[H^2(G,A)\cong \{0\to A\to E\to G\to 1\}/\sim\]
+Check out Weibel 6.6.3.
+
+\subsection{Triangulated Categories}
+Recall the definition of triangulated categories from James' class or from my notes o
+n Brown Representability:
+\begin{defn}
+	Let $\cal C$ be a $k$-linear additive category with a \textbf{suspension} or \textbf{shift} functor (auto-equivalence, actually)
+	\[\Sigma:\cal C\to\cal C.\]
+
+	Then $({\calC}, \Sigma ,D)$ is \textbf{triangulated} category where $D$ is a full, nonempty subcategory $D\subseteq\Delta(\cal C)$
+	with shift functor $\Sigma$ of $D$ and we have the following axioms:
+	\begin{itemize}
+		\item (\textbf{TR0}) $0\to X\xrightarrow{\id} X\to 0$ is in $D$ for each $X\in \cal C$ and furthermore $D$ is closed under both shifts and triangle isomorphisms.
+		\item (\textbf{TR1}) [\textit{Mapping Cone Axiom}] For any $f:X\to Y$ in $\cal C$, there is a triangle
+		\[X\xrightarrow{f} Y\to Z\to\Sigma X\]
+		\item (\textbf{TR2}) [\textit{Rotation Axiom}] If $F\in D$, then $\mathscr R(F),\mathscr R^{-1}(F)\in D$
+		\item (\textbf{TR3}) [\textit{Morphism Axiom}] Given two triangles 
+		\begin{center}
+			\begin{tikzcd}
+				X\ar[r]\ar[d,"f"] & Y\ar[r]\ar[d,"g"] & Z\ar[r]\ar[d,"h",dashed] & \Sigma X\ar[d,"\Sigma f"]\\
+				X'\ar[r] & Y'\ar[r] & Z'\ar[r] & \Sigma X'
+			\end{tikzcd}
+		\end{center}
+		with maps $f$ and $g$, there exists an $h:Z\to Z'$ such that the above diagram commutes.
+		\item (\textbf{TR5}) [\textit{Octahedral Axiom}] Given three triangles: $X\xrightarrow{u}Y\to Z'\to\Sigma X$, $Y\xrightarrow{v}Z\to X'\to\Sigma Y$
+		and $X\xrightarrow{v\circ u} Z\to Y'\to\Sigma X$. Then there is a triangle $Z'\to Y'\to X'\to\Sigma Z'$ such that
+		\begin{center}
+			\begin{tikzcd}
+				X\ar[dr,"u"]\ar[rr,"v\circ u",bend left] && Z\ar[rr,bend left]\ar[dr] && X'\ar[rr,dashed,bend left]\ar[dr] && \Sigma Z'\\
+				& Y\ar[ur,"v"]\ar[dr] && Y'\ar[ur,dashed]\ar[dr] && \Sigma Y\ar[ur] &\\
+				&& Z'\ar[rr,bend right]\ar[ur,dashed] && \Sigma X\ar[ur] &&
+			\end{tikzcd}
+		\end{center}
+	\end{itemize}
+\end{defn}
+\begin{rmk}
+There are two other representations of the octahedral axiom that are sometimes more helpful:
+\begin{center}
+	\begin{tikzcd}
+		X\ar[r,"u"]\ar[d,"\sim"] & Y\ar[r]\ar[d,"v"] & Z'\ar[r]\ar[d,dashed] &\Sigma X\ar[d,"\sim"]\\
+		X\ar[r,"v\circ u"] & Z\ar[r]\ar[d] & Y'\ar[r]\ar[d,dashed] &\Sigma X\ar[d,"\Sigma u"]\\
+		& X'\ar[r,"\sim"] \ar[d] & X'\ar[r]\ar[d,dashed] & \Sigma Y\\
+		&\Sigma Y\ar[r] & \Sigma Z' &
+	\end{tikzcd}
+\end{center}
+
+\begin{center}
+	\begin{tikzcd}
+		&& & X\ar[dl,"u"]\ar[dd,"v\circ u"] &\\
+		&& Y\ar[dr,"v"]\ar[ddll] & &\\
+		&& & Z\ar[d]\ar[dr] &\\
+		Z'\ar[rrr,dashed] && & Y'\ar[r,dashed] & X'
+	\end{tikzcd}
+\end{center}
+
+\begin{center}
+	\begin{tikzcd}
+		A\ar[rr,"f"]\ar[rrrr,bend left,"g\circ f"] & & B\ar[rr,"g"]\ar[dl] & & C\ar[ddll,bend left,dotted]\ar[dl]\\
+		& C'\ar[ul,dotted]\ar[dr] & & A'\ar[ul,dotted]\ar[ll,dotted] &\\
+		& & B'\ar[ur]\ar[uull,dotted, bend left] & &
+	\end{tikzcd}
+\end{center}
+Above dotted lines refer to shifting.
+\end{rmk}
+\begin{lem}
+	Say that $A\xrightarrow{u}B\xrightarrow{v}C\to\Sigma A$ is exact. Then $v\circ u=0$.
+\end{lem}
+\begin{prf}
+	Just complete the morphism $(\id_A,u)$ to a triangle morphism. This gets it!
+\end{prf}
+\begin{ex}
+	$\K(\Ch(\Rmod))$ is triangulated with triangles being all those in $\Delta(\calC)$ isomorphic to $A\xrightarrow{f}B\to \cone(f)\to \Sigma A$.
+\end{ex}
+
+\section{May 22, 2019}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2019-05-22 Triangulated Categories/Tri Cat.tex
update

@@
 \newcommand*{\Amod}{A\text{-}\mathbf{mod}}
+\DeclareMathOperator{\depth}{depth}
 
@@
 
+\section{May 17. 2019}
+I missed a day but they talked about dualizing complexes some more and what they imply.
+
+Let's do an example of the day:
+\begin{ex}
+	Let $A$ be a finite dimensional algebra over $k$. Then $R=A^\ast=\Hom_k(A,k)$ is finite dimensional over $k$. Then we claim that $R$
+	is a dualizing complex over $(A,A)$. To see this, we must check:
+	\begin{itemize}
+		\item $H^i(R)=A^\ast$ for $i=0$ and is zero otherwise.
+		\item $\injdim_AR=\injdim_AA^\ast=0$
+		\item $R\Hom_A(R,R)=Hom_A(A^\ast,A^\ast)=\Hom_A(A,A)=A$.
+	\end{itemize}
+\end{ex}
+
+\subsection{Proving the Dualizing Theorem}
+\begin{defn}
+	$R$ is a dualizing complex over $(A,B)$ if we have functors 
+	\[\bbD=R\Hom_A(-,R):\Db_{fg}(\Amod)\to \Db_{fg}(\mathbf{mod}\text{-}B)\]
+	and $\bbD^\circ=R\Hom_{B^{op}}(-,R)$ in the other direction.
+\end{defn}
+\begin{lem}
+	Let $W$ be a complex of $(A,B)$ bimodules and let $X$ be a complex of $A$-modules and $Y$ a complex of $B^{op}$-modules. Then 
+	\[\Hom_{\Ch(\Amod)}(X,\Hom_{\Ch(B\text{-}\mathbf{mod})}(Y,W)\cong\Hom_{\Ch(\mathbf{mod}\text{-}B)}(Y,\Hom_{\Ch(\Amod)}(X,W))\]
+\end{lem}
+\begin{thm}[Dualizing Theorem]\label{thm-dualizing}
+	$(\bbD,\bbD^\circ)$ induces a contravariant equivalence between $\Db_{fg}(\Amod)$ and $\Db_{fg}(\mathbf{mod}\text{-}B)$
+\end{thm}
+\begin{rmk}
+	We are looking to construct a natural isomorphism between the identity functor on $\Db_{fg}(\Amod)$ to $\bbD^\circ\bbD$.
+\end{rmk}
+\begin{lem}
+	Let $W$ be a complex of $(A,B)$-bimodules and let $X$ and $Y$ be chains in $\Amod$ and $\mathbf{mod}\text{-}B$, respectively. Then 
+	\[R\Hom_A(X,R\Hom_{B^{op}}(Y,W))\cong R\Hom_{B^{op}}(Y,R\Hom_A(X,W)).\]
+\end{lem}
+\begin{lem}
+	Under the same conditions as above,
+	\[\Hom_{\D(\Amod)}(X,R\Hom_{B^{op}}(Y,W))\cong\Hom_{\D(\mathbf{mod}\text{-}B)}(Y,R\Hom_A(X,W))\]
+\end{lem}
+\begin{cor}
+	$R\Hom_{B^{op}}(-,W)$ and $R\Hom_A(-.W)$ is an adjoint pair. This gives us the map $\eta:1\to\bbD^\circ\bbD$.
+\end{cor}
+\begin{lem}
+	Let $F$ and $G$ be two triangulated functors from $\calT_1\to\calT_2$ and $\eta:F\to G$ be a natural transformation. Then 
+	if we let $S=\{X\in\calT|\eta_X:F(X)\to G(X)\text{ is an isomorphism}\}$, then $S$ is a triangulated subcategory.
+\end{lem}
+\begin{cor}
+	It is a fact that the triangulated subcategory of $\D(\Amod)$ generated by $A$ is $\D_{perf}(\Amod)$. Thus, $\D_{perf}(]Amod)\subseteq S$.
+	Thus in particular $\eta_X:X\mapsto\bbD^\circ\bbD(X)$ is an isomorphism for all $X\in\D_{perf}(\Amod)$.
+\end{cor}
+\begin{cor}
+	$\bbD$ and $\bbD^\circ$ induce a contravariant equivalence between $\D_{perf}(\Amod)$ and $\D_{perf}(\mathbf{mod}\text{-}B).$
+\end{cor}
+
+\begin{defn}
+	Let $X$ be a complex of $A$-modules. Then define $X_{[a,b]}$ to be 
+	\[X^i_{[a,b]}=X^i\]
+	whenever $i\in[a,b]$ and zero otherwise. Same for the differentials. This is a complex (obvious).
+\end{defn}
+
+\begin{lem}
+\begin{itemize}
+	\item If $X_{[b,\infty)}=0$, then there exists an ``embedding'' morphism $X_{[a,b]}\to X$.
+	\item If $X_{(-\infty,a]}=0$ then there exists a ``projection'' morphism $X\to X_{[a,b]}$.
+\end{itemize}
+\end{lem}
+\begin{lem}
+	let $X$ be  a complex over $A$ such that $X=X_{[-b,b]}$. Let $Y$ and $Z$ be two complexes such that 
+	\[Y_{[-c,c]}=Z_{[-c,c]}.\]
+	Then 
+	\[\Hom_{\Ch(\Amod)}(X,Y)_{[-(c-b),c-b]}=\Hom_{\Ch(\Amod)}(X,Z)_{[-(c-b),c-b]}\]
+
+	This still holds if you switch the arguments in $\Hom$.
+\end{lem}
+\begin{cor}
+	If $X$ and $Y$ ``agree locally'' as above, then $\bbD(Y)$ and $\bbD(Z)$ agree locally. This is slightly confusing since 
+	we're working in the derived category here, but this means that they have isomorphic cohomology in some range.
+\end{cor}
+
+\begin{cor}
+	$\bbD^\circ\bbD(Y)$ and $\bbD^\circ\bbD(Z)$ agree locally.
+\end{cor}
+
+\begin{prf}[of~\ref{thm-dualizing}]
+	We need to show that $H^i(P)\cong H^i(\bbD^\circ\bbD(P))$ for all $i$. Fix $i$.
+
+	Then call $Q=P_{[-c,c]}$ for $c>> i+2b$ and then $H^i(Q)\cong H^i(P)$ and looking at the naturality square everything is an isomoprhism and we're done.
+\end{prf}
+
+\section{May 20, 2019}
+Today we will discuss homological identities.
+
+\begin{defn}
+	If $A$ is an algebra over $k$, it is called \textbf{local} if $A/J(A)\cong k$, where $J(A)$ is the Jacobson radical.
+\end{defn}
+
+\begin{thm}[Bass]
+	Let $A$ be a local Noetherian algebra with balanced dualizing complex. Let $M$ be a finitely genreated left $A$-module with finite injective 
+	dimensions. Then 
+	\[\injdim M=\operatorname{depth}A<\infty.\]
+\end{thm}
+\begin{rmk}
+	We haven't yet defined a balanced dualizing complex but we will. Later we will also see some theorems regarding 
+	existence of such complexes.
+\end{rmk}
+
+\subsection{(Non)Example of the Day}
+\begin{ex}
+	There is an Artinian algebra $A$ such that for any algebra $B$, there is \textbf{no} dualizing complex over $(A,B)$. The construction 
+	is quite complicated, but it works. Note that in particular $A$ cannot be finite dimensional since we saw that 
+	the dual of $A$  is a dualizing complex for $A$ when it is.
+\end{ex}
+\begin{ex}
+	There is an Artinian algebra $A$ such that
+	\begin{itemize}
+		\item There is no dualizing complex over $(A,A)$
+		\item There is a dualizing complex over $(A,B)$ for some $B$.
+	\end{itemize}
+\end{ex}
+\begin{ex}
+	More concretely, let $A=(\begin{smallmatrix}k&k\\0 &k\end{smallmatrix})$ for some field $k$. This is not local since $A/J(A)\cong k\oplus k$.
+	Furthermore, $\gldim A=1$. Now let $e_{11}$ and $e_{22}$ refer to the standard $2x2$ matrices with ones in the appropriate spots and zeros elsewhere.
+
+	Then if $S_1=Ae_{11}$ and $S_2=\binom{0}{k}$, then we get a sequence 
+	\[0\to S_1\to\binom{k}{k}\to S_2\to 0\]
+	and so $\projdim S_1=0=\injdim S_2$ and $\projdim S_2=1=\injdim S_1$, giving us a counterexample when we don't have locality.
+\end{ex}
+\begin{ex}
+	More simply, if $A=k[[t]]$, $\projdim A=0\ne1\projdim k.$
+\end{ex}
+
+\subsection{Some Review}
+Let $X$ be a bounded (this can be defined more generally but James' definiton requires this) complex of left $A$-modules.
+\begin{defn}
+	The projective dimension is defined to be 
+	\[\projdim X=\sup\{i|\Ext^i(X,M)\ne 0,\text{ some M}\in\Amod\}\]
+	or equivalently
+	\[\projdim X=\inf_P\sup\{i|P^{-i}\ne 0\}\]
+	where $P$ ranges over all projective resolutions of $X$.
+\end{defn}
+\begin{defn}
+	The injective dimension is defined to be 
+	\[\injdim X=\sup\{i|\Ext^i(M,X)\ne 0,\text{ some M}\in\Amod\}\]
+	or equivalently
+	\[\injdim X=\inf_I\sup\{i|I^{-i}\ne 0\}\]
+	where $I$ ranges over all injective resolutions of $X$.
+\end{defn}
+
+Now let $A$ be local and write ${_A}k=A/J(A)$.
+\begin{defn}
+	Let $X$ be a bounded complex of $A$ modules. Then the \textbf{depth of $X$} is
+	\[\operatorname{depth}X=\inf\{i|\Ext^i(k,X)\ne 0\}\]
+\end{defn}
+\begin{rmk}
+	Notice that, in general, $\depth X\le \injdim X$.
+\end{rmk}
+\begin{rmk}
+	In commutative algebra, the more natural definition is in terms of lengths of regular sequences. This ends up being 
+	equivalent but requires a little work.
+\end{rmk}
+\begin{defn}
+	Let $X$ be a bounded complex. Then 
+	\[\inf X=\inf\{i|H^i(X)\ne 0\}\]
+	and 
+	\[\sup X=\sup\{i|H^i(X)\ne 0\}\]
+\end{defn}
+\begin{rmk}
+	So then $\depth X=\inf(R\Hom_A(k,X)).$
+\end{rmk}
+\begin{lem}
+	Let $A$ be a Noetherian local algebra and let $X\in\Dneg_{fg}(\Amod)$. Let $P$ be a minimal free resolution of $X$.
+	Then 
+	\[\projdim X=\max\{i|P^{-i}\ne 0\}\]
+\end{lem}
+\begin{prf}
+	We always have that $\projdim X$ is less than or equal to this by the definition. So it suffices to show the reverse inequality.
+	That is, we want to show that $\projdim X\ge i$ if $P^{-i}\ne 0.$
+
+	Since $P$ is minimal, the complex $\Hom_{\Ch(\Amod)}(P,k)$ has zero differential (since $\partial_P^i$ maps into $J(A)P^{i+1}$).
+	But then computing:
+	\[\Ext_A^i(X,k)=H^i(\Hom(P,k))=\Hom(P,k)^i=\Hom_A(P^{-i},k)\ne 0\]
+	whence we can conclude that $\projdim X\ge i$.
+\end{prf}
+
+\subsection{Canonical Truncations}
+\begin{defn}
+	Let $X$ be a complex in $\Amod$. The \textbf{$n^{th}$ bounded above truncation of $X$ is}
+	\[\calT^{\le n}X=\cdots\to X^i\to\cdots\to X^{n-1}\to\ker(\partial_X^n)\to 0\to \cdots\]
+	and the \textbf{$n^{th}$ bounded below truncation} is
+	\[\calT^{\ge n}X=\cdots 0\to\coker(\partial_X^{n-1})\to X^{n+1}\to X^{n+2}\to\cdots\]
+\end{defn}
+\begin{thm}
+	Let $A$ be a Noetherian local algebra and let $X,Y\in\Db_{fg}(\Amod)$. Suppose $\projdim X<\infty$.
+	Then 
+	\[\sup(R\Hom_A(X,Y))=\projdim X+\sup Y.\]
+\end{thm}
+\begin{prf}
+	Not enough time for the full thing. So we prove ``$\le$''. Let $P$ be a minimal free resolution of $X$. Then 
+	$P=0\to P^{-d}\to\cdots\to P^b\to 0\to\cdots$. Then $d=\projdim X$.
+
+	Say $Y=0\to Y^a\to\cdots\to Y^s\to 0$, so $\sup Y=s$. But notice $R\Hom(X,Y)=\Hom_{\Ch(\Amod)}(P,Y)$
+	so for $i>d+s$,
+	\[\Hom(P,Y)^i=\prod_{s\in\bbZ}\Hom_A(P^s,Y^{s+i})=0\]
+	so therefore $\sup \Hom(P,Y)\le d+s$.
+\end{prf}
+\begin{cor}
+	Let $Y$ be a finitely generated module and $X$ a module with finite projective dimension. Then 
+	\[\projdim X=\sup(R\Hom_A(M,N)).\]
+\end{cor}
+
 \end{document}

### 2019-05-22 Homological Algebra Notes/Hom Alg.tex
update

@@
 \DeclareMathOperator{\sExt}{\mathscr{E}xt}
+\DeclareMathOperator{\stmod}{stmod}
 
@@
 \section{May 22, 2019}
-
+Some examples of triangulated categories: we have already seen that $\K(\Rmod)$ and $\D(R)$ are. But we also have
+that $\stmod G,$ the category of all $G$ modules and 
+\[\Hom_{\stmod G}(M,N)=\underline{\Hom}(M,N)=\Hom_G(M,N)/P\Hom_G(M,N)\]
+where $P\Hom_G(M,N)$ is the collection of all morphisms that factor through a projective object $P$.
+
+In essence, $\stmod G$ is the same as considering $G$ modules ``up to projective summand''. Recall that this makes the syzygies 
+uniquely defined.
+
+\begin{prop}
+	$\stmod G$ is a tensor triangulated category.
+\end{prop}
+\begin{prf}
+	The key idea here is that the category of $G$-modules is Frobenius, so all injectives are projective and vice versa.
+
+	Then the triangulated structure comes from short exact sequences of $G$-modules and now our shift $\Omega$ is the syzygy operator.
+
+	Recall that we have seen the dimension shifting identity: for $i>1$:
+	\[\Ext_G^i(M,N)\cong\Ext^{i-1}(\Omega M,N)\cong\Ext^{i+1}(M,\Omega^{-1}N)\]
+	the reason this doesn't actually extend to $i=1$ is that this doesn't work with $\Hom$.
+
+	But it ends up that this actually \textit{does} work if we work with stable $\Hom$. But then we get $\Ext^1(M,N)\cong\underline{\Hom}(M,\Omega^{-1}N)$!
+	so somehow this gives us a notion of a map $M\to \Omega^{-1}N$, so given the triangle $N\to P \to M$ of $G$ modules,
+	\[N\to P\to M\to \Omega^{-1}N\]
+	is a triangle in $\stmod G$.
+\end{prf}
+
+\subsection{Singularity category}
+Let $R$ be a ring and consider $\Db(R)$, the bounded derived category. There is a subcategory $\Dbperf(R)$,
+which is the bounded derived category of projective modules.
+
+Now since the tensor of a projective module with anything is again projective, this forms a (categorical) ideal within $\Db(R)$. Taking the 
+(categorical) quotient, we get $\Dsing(R)$, the \textbf{singuarity category.}
+
+In the case that $R$ is regular (think finite global dimension), we get that any bounded complex is quasi-isomorphic to (via the Cartan-Eilenberg resolution)
+to a perfect complex! So then $\Dsing(R)$ is trivial.
+
+\begin{thm}[Rikard, Buchweitz, Orlov]
+	Without assumptions on $k$ or $G$,
+	\[\Dbperf(kG)\hookrightarrow \Db(kG)\to\stmod G\]
+	is an exact sequence.
+\end{thm}
+\begin{rmk}
+	Notice that this (essentially) says that semisimplicity (from the world of representation theory) is
+	effectively the same as regularity (from algebraic geometry).
+\end{rmk}
+
+\subsection{Cohomological Functors}
+\begin{defn}
+Let $\calK$ be a triangulated category and let $\calA$ be abelian. Then $H:\calK\to \calA$ is called \textbf{cohomological}
+if it takes triangles in $calK$ to long exact sequences in $\calA$.
+\end{defn}
+\begin{prop}
+	$\Hom_\calK(X,-):\calK\to\Ab$ is cohomological.
+\end{prop}
+\begin{prf}
+	Look at $A\xrightarrow{u}B\xrightarrow{v}C\to\Sigma A$. Now we want to show that $v_\ast u\ast=(vu)_\ast=0$
+	or that $\ker v_\ast= \Im u_\ast$. Basically you can find a map to show it. Use the morphism axiom!
+\end{prf}
+
+\subsection{Yoneda Lemma}
+The idea here is that objects can be determined (up to isomorphism) by the maps to or from them.
+
+Let $\calA$ be a locally small category. Then $\mathbf{Funct}(\calA,\mathbf{Set})$ is the category of $\mathbf{Set}$-valued functors. Then there is a functor 
+$\calA\to\mathbf{Funct}(\calA,\mathbf{Set})$ sending $A$ to $h_A=\Hom(A,-)$ (or $h^A=\Hom(-,A)$). Then the lemma says 
+\begin{lem}[Yoneda]
+	There exist bijections
+	\[\Hom_\calA(A,A')\cong\mathbf{Nat}(h_{A'},h_A)\cong \mathbf{Nat}(h^A,h^{A'})\]
+\end{lem}
+\begin{cor}
+	There is a fully faithful embedding of $\calA$ into $\mathbf{Funct}(\calA,\mathbf{Set})$.
+\end{cor}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2019-06-03 Homological Algebra Notes/Hom Alg.tex
Update!

@@
 \usepackage[mathscr]{euscript}
+\usepackage{hyperref}
 
@@
 \end{cor}
+
+\section{May 29, 2019}
+\subsection{Motivation}
+As some motivation for today, recall the problem on the first homework assignment where we were asked to 
+say whether $\K(\Ch R)$ is abelian. The answer was no. Recall the 5-lemma (applied to triangles):
+
+\begin{prop}
+	If the following is a commutative diagram:
+	\begin{center}
+		\begin{tikzcd}
+			A\ar[r]\ar[d,"\sim"] & B \ar[r]\ar[d,"\sim"] & C\ar[r]\ar[d,"h"] & \Sigma A\ar[d]\\
+			A'\ar[r] & B'\ar[r] & C'\ar[r] & \Sigma A'
+		\end{tikzcd}
+	\end{center}
+	is a morphism of triangles, then $h$ is an isomorphism.
+\end{prop}
+\begin{prf}
+	Basically use Yoneda.
+\end{prf}
+
+\begin{rmk}
+	In any triangulated category every monomorphism splits. To see this, take a triangle $A\xrightarrow{i}B\to C\to\Sigma A$.
+	There is also an exact triangle $A\to A\oplus C\to C\xrightarrow{0}\Sigma A$ in $\K(R)$. Then use the five lemma and $B\cong A\oplus C$.
+\end{rmk}
+\begin{ex}
+	We want to show that $\K(\bbZ)$ is not abelian. Let $C=\bbZ/4$ in degree zero and let $C'=\bbZ/2$ in degree zero. Let 
+	$f:C\to C'$ be a morphism. If this category is abelian, then there exists a kernel of $f$, 
+	so $\ker f\hookrightarrow C\xrightarrow{f}C'$ and furthermore this splits. 
+	\[\bbZ/4\cong H(C)\not\cong H(C')\oplus H(\ker f)\] 
+	but since $H(C')\cong\bbZ/2$, this is a contradiction.
+\end{ex}
+\subsection{Localization and the calculus of fractions}
+Note that here set theory and size begins to be a problem, and this is discussed (e.g.) in Weibel. We 
+won't delve into it much.
+\begin{defn}
+	Let $\calC$ be a category and $S$ a collection of morphisms in $\calC$. Then $S$ is called a multiplicative system of $\calC$ if 
+	\begin{itemize}
+		\item $S$ is closed under composition and contains all identity morphisms.
+		\item (\textit{Ore Condition}) For any $s:Y\to X$ and $f:Z\to X$ are maps with $s\in S$,
+		there are maps $t:W\to Z$ and $g:W\to Y$ with $t\in S$ such that 
+		\begin{center}
+			\begin{tikzcd}
+				W\ar[r,"g"]\ar[d,"t"] & Y\ar[d,"s"]\\
+				Z\ar[r,"f"] & X
+			\end{tikzcd}
+		\end{center}
+		\item If $f,g:X\to Y$ are maps, then there exists a map $s\in S$ such that $s\circ f=s\circ g$ if 
+		and only if there exists $t\in S$ such that $f\circ t=g\circ t$.
+	\end{itemize}
+\end{defn}
+
+Then you get a new category $S^{-1}\calC$, the localizaton of $\calC$ at $S$. There is a functor $F:\calC\to S^{-1}\calC$ and it satisfies 
+a universal property that if $G:\calC\to\calD$ is a functor for which $G(s)$ is an isomorphism for all $s\in S$, then $G$
+factors through $S^{-1}\calC$.
+
+We can do the usual ``roof construction'' on the $\Hom$ sets to complete the structure of this category.
+\begin{thm}[Gabriel-Zisman]
+	Let $S$ be a locally-small multiplicative system and let $S^{-1}\calC$ consist of the objects of $\calC$
+	and whose morphisms are the roofs modulo the appropriate equivalence relation (see James' notes). This is a category and 
+	satisfies the universal property.
+\end{thm}
+\begin{rmk}
+	If $\calC$ is additive, then $S^{-1}\calC$ is additive.
+\end{rmk}
+\begin{rmk}
+	The fact that all things in $S$ go to isomorphisms is a matter of playing with diagrams (roofs).
+\end{rmk}
+
+Using this, one can construct the derived category! Just localize $\K^\ast(R)$ a the quasi-isomorphisms.
+
+\begin{defn}
+	Let $\calA$ be an abelian category. Then $\calB$ is a Serre subcategory of $\calA$ if it is closed under
+	\begin{itemize}
+		\item subobjects
+		\item quotients
+		\item extensions
+	\end{itemize}
+\end{defn}
+Then if we fix a subcategory $\calB$, let $S_\calB$ be the set of morphisms in an abelian category $\calA$ such that
+both $\ker f$ and $\coker f$ lie in $\calB$, then 
+\[\calA/\calB\cong S_\calB^{-1}\calA\]
+is the quotient category.
+
+\section{May 31, 2019}
+Today I presented my talk on Brown representability. You can find the repository containing my notes for it \href{https://github.com/NicoCourts/Brown-Representability}{here}.
+
+\section{June 3, 2019}
+Today we're talking about spectral sequences! The primary references are Vakil's ``The Rising Sea'' (specifically section 1.7)
+as well as Weiel and May's ``A primer on spectral sequences.''
+
+\subsection{Crush Course in Spectral Sequences}
+There are two usual treatments that are followed here: either filtered complexes or exact couples. In this lecture we will do neither.
+Instead we will focus on a subcase of a filtered complex: double complexes. Recall
+\begin{defn}
+	A \textbf{double complex} $E_{\bullet\bullet}$ of a $\bbZ\times\bbZ$ indexed collection of objects $E_{pq}$ in some 
+	abelian category, with differentials 
+	\[d^v_{pq}:E_{pq}\to E_{p,q-1}\quad d^h_{p,q}:E_{p,q}\to E_{p-1,q}\]
+	such that $d^v\circ d^v=d^h\circ d^h=0$ and 
+	\[d^h\circ d^v+d^v\circ d^h=0\]
+	(anticommutativity).
+\end{defn}
+So a diagram for a double complex (specifically a first-quadrant complex which we will be focusing on) is 
+\begin{center}
+\begin{tikzcd}
+	later
+\end{tikzcd}
+\end{center}
+\begin{defn}
+	Recall that $\operatorname{Tot}(E_{\bullet\bullet})$ is the total complex where the $n^{th}$ piece is given by 
+	\[\oplus_{i+j=n}E_{i,j}\]
+	with differential $d=d^h+d^v$.
+\end{defn}
+The goal here is to compute information about the homology of this complex.
+\subsection{Spectral Sequences}
+\begin{defn}
+	A \textbf{spectral sequence} is a sequence of \textbf{pages} $E_{p,q}^0,E_{p,q}^1,\dots$ for all $p$ and $q$ and all $r\ge 0$
+	with differentials
+	\[d_{p,q}^r:E_{p,q}^r\to E_{p-r,q+r-1}^r\]
+	and fixed isomorphisms 
+	\[E_{p,q}^{r+1}\cong H_{p,q}^r\]
+	where $H_{p,q}^r$ is the homology on the $r^{th}$ page of $d^r$ at $E_{p,q}^r$.
+\end{defn}
+
+The zeroth page has vertical differentials, the first page has horizontal differentials. In general, the slope of the 
+differential on the $r^{th}$ page is $\frac{r-1}{r}$.
+
+\begin{rmk}
+	For each $p,q$, there is an $r$ (large enough) such that any location stablizizes:
+	that is, eventually the differential looks like 
+	\[0\xrightarrow{d^r}E_{p,q}^r\xrightarrow{d^r}0\]
+	and so $E_{p,q}^r\cong E_{p,q}^{r+1}$.
+
+	This always happens when we are using a first quadrant complex, but whenever it happens, we will call 
+	this stabilization point $E_{p,q}^\infty$. Notice where this happens is dependent on position, in general.
+\end{rmk}
+
+\begin{prop}
+	$d^h$ induces $d^1$, the differential on the first page.
+\end{prop}
+\begin{prf}
+	This amounts to showing that the obvious map is well-defined. :)
+\end{prf}
+\begin{prob}
+	Show that $d^2$ is well-defined. This seems not completely obvious, but apparently it is a diagram chase away.
+\end{prob}
+\begin{rmk}
+	Vakil apparently has an idea of ``$r$-closed $(p,q)$ strips'' that somehow obviate the need for such diagram chasing.
+\end{rmk}
+
+\subsection{Filtrations}
+\begin{thm}
+	$H_k(\operatorname{Tot}_\bullet(E_{\bullet\bullet}))$ is filtered by $E_{p,q}^\infty$ for $p+q=k$.
+
+	More precisely, there is a filtration 
+	\[0\subseteq F_0\subseteq \cdots\subseteq F_k\subseteq H_k(\operatorname{Tot}_\bullet(E_{\bullet\bullet}))\]
+	such that $F_i/F_{i-1}\cong E_{i,k-i}^\infty$.
+\end{thm}
+\begin{rmk}
+	In such a case, we say that $E_{p,q}^r$ \textbf{converges to} (or \textbf{abuts}) $H_{p+q}$ and write $E_{p,q}^r\Rightarrow H_{p+q}$.
+\end{rmk}
+\begin{rmk}
+	We can ``flip'' across $y=x$ and use this spectral sequence. Notice the specific entries may be different.
+\end{rmk}
+
+\subsection{Applications}
+We can use it to prove the snake lemma! Consider the diagram (which anticommutes):
+\begin{center}
+	\begin{tikzcd}
+		0 & D\ar[d,"\alpha"]\ar[l] & E\ar[l]\ar[d,"-\beta"] & F\ar[l]\ar[d,"\gamma"] & 0\ar[l]\\
+		0 & A\ar[l] & B\ar[l] & C\ar[l] & 0\ar[l]
+	\end{tikzcd}
+\end{center}
+
+Then if we compute the horizontal direction first, we get $E_{pq}^1=E_{pq}^\infty=0$ so $H_k(\operatorname{Tot}_\bullet(E_{\bullet\bullet}))=0$ for all $k$.
+
+Then consider the other direction (vertical first). From here you can extract the fact that you have an exact map between the kernels/cokernels 
+and you get an isomorphism $d^2:E_{2,0}^2\to E_{0,1}^2$ on the second page, which basically gives you your connecting homomorphism.
+
+\subsection{Further Reading and Applications}
+The original context that these arose were in Leray's spectral sequence: Given an orientable fibration $F\hookrightarrow E\to B$, 
+we get 
+\[H^p(B,H^q(F))\Rightarrow H^\ast(E).\]
+
+Another example is the Grothendieck spectral sequence. Given some ``nice'' functors $\calF$ and $\calG$, 
+\[R^p\calF\circ R^q\calG(X)\Rightarrow R^{p+q}(\calF\circ \calG)(X)\]
+
+Finally, (very recently!) Karim Adiprasito proved the ``g-conjecture.'' Along the way, he used certain spectral sequences 
+of tensor products of chain complexes, and this motivated Sean's original research.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2019-06-03 Triangulated Categories/Tri Cat.tex
Update!

@@
 
+\section{May 22, 2019}
+\subsection{Example of the day}
+This is a combination of some results of Artin, Van den Bergh, Yehutieli, Wu, Chan, and many others.
+
+Suppose $A$ is a Noetherian algebra satisfying the following conditions:
+\begin{itemize}
+	\item $A$ is either connected graded or complete local
+	\item $A$ is a finite module over it center.
+\end{itemize}
+Then there exists a dualizing complex $R$ over $(A,A)$, such that 
+\begin{itemize}
+	\item $R\Hom_A(k,R)\cong k$ and $R\Hom_{A^{op}}(k,R)\cong k$. This is sometimes called the \textit{pre-balanced condition.}
+\end{itemize}
+
+\begin{ex}
+	Let $q$ be a root of unity. Let $A=k_q[[x_1,\cdots,x_n]]$ where we say any $x_i$ and $x_j$ $q$-commute, for $i\ne j$. Then this satisfies the hypotheses and $R$ exists.
+	There is a method of constructing it via locak cohomology that is perhaps too much for what we have left in our course. We will, however, discuss local 
+	cohomology and the $\chi$-condition next time.
+\end{ex}
+
+\subsection{Back to the future (Monday)}
+We're going to tweak things a little.
+\begin{thm}
+	Let $A$ be a Noetherian local algebra. Let $X\in\Db_{fg}(\Amod)$ with $\projdim X<\infty$.
+	Then there is a $Y\in\Db(\Amod)$ (here's where we weaken slightly) such that $\Hom_A(H^s(Y),k)\ne 0$ where 
+	$s=\sup(Y)$. Then 
+	\[\sup(R\Hom_A(X,Y))=\projdim X+\sup Y.\]
+\end{thm}
+\begin{prf}
+	Last time we saw ``$\le$''. Now let $d=\projdim X$ and $s=\sup Y$. It suffices to show that
+	\[H^{d+s}(R\Hom_A(X,Y))=H^{d+s}(\Hom_{\Ch(\Amod)}(P,Y))\ne 0\]
+	where we take $P$ to be any (in this case minimal) free resolution of $X$. But if you consider the $d+s$ morphisms from $P$ to $Y$,
+	there is only one map between nonzero things! So 
+	\[\Hom_{\Ch(\Amod)}(P,Y)^{d+s}=\Hom_A(P^{-d},Y^s).\]
+
+	We want to construct an $\tilde f\in\Hom_{\Ch}(P,Y)^{d+s}$ (notice it is already in the kernel) such that $\tilde f$ isn't in the image. To see this, 
+	set $N=H^s(Y)=Y^s/\Im(\partial_Y^{s-1})\ne 0$ (we can assume $Y^{s+1}=0$). Then by the hypothesis, $\Hom_A(N,k)\ne 0$,
+	but this is equal to 
+	\[\Hom_A(N,A/J(A))=\Hom_k(A/J(A)\otimes N, A/J(A))\ne 0.\]
+
+	This implies that $A/J(A)\otimes N\ne 0$ or $N/J(A)N\ne 0$. Notice that
+	\[N/J(A)N=Y^s/(\Im(\partial_Y)+J(A)Y^s)\ne 0.\]
+
+	Now since $N/J(A)N$ is a direct sum of simple modules, there is a nonzero morphism $f:P^{-d}\to N/J(A)N$. Since $P^{-d}$ is 
+	(in particular) projective, we can lift to a morphism $\tilde f:P^{-d}\to Y^s$. Furthermore, $\Im(\tilde f)$ is not contained within
+	$\Im(\partial_Y^{s-1})+J(A)Y^s$ (since it wasn't the zero map on $N/J(A)N$).
+
+	Now consider the chain complex $\Hom_{\Ch(\Amod)}(P,Y)^\ast$ at the $d+s-1^{st}$ term. The differential maps from 
+	$\Hom_A(P^-d,Y^{s-1})\oplus\Hom_A(P^{-d+1},Y^s)\to \Hom_A(P^{-d},Y^s)$ where 
+	\[\partial(\alpha+\beta)=\partial_Y\circ\alpha+\beta\circ\partial_P\]
+
+	But then by checking maps (and using that since $P$ is minimal, $\partial_P^{-d}$ maps into $J(A)P^{-d+1}$), we can 
+	show that the image of any map under this differential must map into this object, so $\tilde f$ must not lie in the image, 
+	so it is a nonzero element of the homology.
+\end{prf}
+
+\subsection{Tensor Product of Complexes}
+Now let $X\in\Ch(\modA)$ and $Y\in\Ch(\Amod)$. Then $X\otimes_A Y\in\Ch(\Vectk)$ is
+\[(X\otimes_A Y)^i=\bigoplus_{i+j=i}X^i\otimes Y^j\]
+and the differential is given by the Koszul sign rule $\partial(x\otimes y)=\partial_X(x)\otimes y+(-1)^|x|x\otimes\partial_Y(y)$
+
+The derived functor of $-\otimes_A-$ is 
+\[X\otimes_A^LY\cong P_X\otimes_A y\cong X\otimes_A P_Y\cong P_x\otimes_AP_Y\]
+where, for example, $A\otimes_A^LY=A\otimes_AY=Y$ and $X\otimes_A^LA=X$.
+
+\begin{thm}
+	Let $A$ be a Noetherian local algebra. Let $X\in\Db_{fg}(\Amod)$ with $\projdim X<\infty$. Let $Y\in\Db(\modA)$
+	such that $\Hom_A(k,H^i(Y))\ne 0$ where $i=\inf Y$. Then 
+	\[\inf(Y\otimes_A^LX)=-\projdim X+\inf Y\]
+\end{thm}
+
+\begin{lem}
+	Let $A$ be Noetherian and $X\in\Db_{fg}(\Amod)$, $Y\in\Db(A^e\text{-}\mathbf{mod})$ and $Z\in \Dbperf(\Amod)$. Then 
+	\[R\Hom_A(XY\otimes_A^LZ)\cong R\Hom_A(X,Y)\otimes_A^LZ\]
+\end{lem}
+\begin{prf}
+	Let $F(-)=R\Hom_A(X,Y\otimes_A^L-)$ and $G(-)=R\Hom_A(X,Y)\otimes_A^L-$. We want to construct 
+	a natural isomorphism between them. There is a map $\eta:F\to G$ that sends 
+	\[f\otimes z\mapsto \{x\mapsto f(x)\otimes z\}.\]
+	Then it only remains to check it is natural.
+
+	The next step is to show that it holds (trivially!) for $Z=A$ and then since $A$ generates $\Dbperf(\Amod)$, we get that $G\cong F$ for all 
+	things in $\Dbperf(\Amod)$.
+\end{prf}
+
+\section{May 24, 2019}
+Had to miss this day.
+
+\section{May 29, 2019}
+
+\subsection{Example of the Day}
+Let $A$ be a Noetherian algebra that is a finitely generated module over its center. Furthermore, suppose that 
+$\calZ(A)\cong k[x_1,\cdots,x_n]/I$. Then $A$ has a dualizing complex $R$ where $R=R\Hom_{k[x_1,\dots,x_n]}(A,k[x_1,\dots,x_n][n])$
+
+Then $\injdim R_A=\injdim {_A}R=0$, $\Ext^i_A(S,R)=\delta_{i=0}T$ for all simple left $A$-modules $S$.
+
+\begin{defn}
+	$R$ is called balanced if 
+	\begin{itemize}
+		\item $\injdim {_A}R=\injdim R_A=0$
+		\item $R\Hom_A(k,R)=k$ and $R\Hom_{A^{op}}(k,R)=k$
+	\end{itemize}
+\end{defn}
+
+So the idea we wrote down above is very similar to the idea of a balanced dualizing complex (it is stronger).
+
+\subsection{Bass' Theorem}
+\begin{thm}[Bass]
+	Let $A$ be a local Noetherian algebra with balanced dualizing complex. Let $X\in\Dbperf(\Amod)$ with finite injective dimension. Then 
+	\[\injdim X=\sup X+\depth A\]
+\end{thm}
+\begin{cor}
+	Under the same hypotheses, let $M$ be a finitely-generated left $A$-module with finite injective dimension. Then 
+	\[\injdim M=\depth A\]
+\end{cor}
+\begin{rmk}
+	Notice that this also implies that the depth of $A$ is finite.
+\end{rmk}
+
+We begin the proof of the theorem with a lemma or two:
+\begin{lem}
+	Suppose $Y\in\Db_{fg}(\Amod)$. Then 
+	\[\projdim Y=\sup\{R\Hom_A(Y,k)\}\]
+	and 
+	\[\sup Y=-\inf\{R\Hom_A(Y,k)\}\]
+\end{lem}
+\begin{prf}
+	To see this, let $P$ be a minimal projective resolution of $Y$. Write $S=\sup(U)$ and $d=\projdim Y$.
+
+	The idea for completing this proof again leverages that $\partial(P^i)\subseteq J(A)P^{i+1}$ for all $i$. Then notice that 
+	everything is in the kernel of $\partial^\ast$, the induced map from the functor $\Hom_A(-,k)$, which then implies that
+	$\sup(\Hom_A(P,k))=d$ and that $\inf \Hom_A(P,k)=-s$ is clear. Then the result follows from using that $R\Hom_A(Y,k)=\Hom_A(P,k)$.
+\end{prf}
+Let's get another lemma!
+\begin{lem}
+	Suppose $X\in\Db_{fg}(A)$. Then 
+	\[\injdim\bbD(X)\le \projdim X\]
+	and 
+	\[\projdim \bbD(X)\le\injdim X.\]
+\end{lem}
+\begin{prf}
+	Recall here that $\bbD(X)=R\Hom_A(X,R)$ and $\bbD^\circ(X_A)=R\Hom_{A^{op}}(X,R).$ Then suppose that $d=\projdim X<\infty.$ Let $P$ be a minimal 
+	free resolution of $X$. Write $I_R$ to denote a resolution of $R$ as an $A$-bimodules such that each term is injective on the right. Since $\injdim R=0$, $I_R^{>0}=0.$
+
+	But now $\bbD(X)=R\Hom_A(X,R)=\Hom_A(P,I_R)$. Now note that 
+	\[\Hom_A(P,I_R)^i=\oplus_{w\in\bbZ}\Hom(P^w,I_R^{i+w})=\oplus (I_r^{i+w})^{\operatorname{rank}(P^w)}\]
+	so in particular, it is injective. Furthermore, $\Hom_A(P,I_k)^i=0$ whenever $i>d.$
+
+	But then $\bbD(X)$ has an injective resolution that is bounded above at $d$.
+
+	But then we know that 
+	\[\projdim \bbD(X)=\sup\{R\Hom_{A^{op}}()\bbD(X),k)\}=\sup\{R\Hom_A(\bbD^\circ(k),\bbD^\circ\bbD(X))\}\]
+	where the last equality is by the (earlier) duality theorem. But $\bbD^\circ(k)=k$ and $\bbD^\circ\bbD(X)=X$,
+	and so since $\sup\{R\Hom_A(k,X)\}\le\injdim X$, we get
+	\[\projdim\bbD(X)\le\injdim X\]
+
+	I think I missed the other one?!
+\end{prf}
+\begin{cor}\label{cor-equalities}
+	\[\projdim \bbD(X)=\injdim X\] 
+	and 
+	\[\injdim X=\sup\{R\Hom_A(k,X)\}\].
+\end{cor}
+\begin{lem}
+	If $R$ is the balanced dualiznig complex of $A$,
+	\[\depth A=-\sup(R)\ge 0\]
+\end{lem}
+\begin{prf}
+	\begin{align*}
+		\depth A &= \inf\{R\Hom_A(k,A)\}\\
+		&= \inf\{R\Hom_{A^{op}}(\bbD(A),\bbD(k))\}\\
+		&=\inf\{R\Hom_{A^{op}}(R,k)\}\\
+		&= -\sup R.
+	\end{align*}
+\end{prf}
+Next recall some identities from last week:
+\begin{lem}
+	Let $X\in\Dbperf(\Amod)$. Then if $\Hom(H^s(Y),k)\ne 0$ where $s=\sup Y$, then 
+	\[\sup\{R\Hom(X,Y)\}=\projdim X+\sup Y.\]
+\end{lem}
+
+Finally we get around to proving the theorem:
+\begin{prf}[of Bass' theorem]
+	Suppose $\injdim X<\infty.$  Then by corollary~\ref{cor-equalities} above, $\projdim\bbD(X)<\infty$. Then
+	\begin{align*}
+		\sup X &=\sup\{R\Hom_A(A,X)\}\\
+		&=\sup\{R\Hom_{A^{op}}(\bbD(X),\bbD(A))\}\\
+		&=\sup\{R\Hom_{A^{op}}(\bbD(X),R)\}\\
+		&=\projdim\bbD(X)+\sup R\\
+		&=\injdim X-\depth A.
+	\end{align*}
+	And that's what we wanted to prove!
+\end{prf}
+\begin{rmk}
+	The preceding proof was adapted by Peter Jorgensen's paper. He apparently has very clear exposition 
+	and arguments where each step is carefully justified. The paper in question here is entitled something like \textit{Homological Identiies}
+	but apparently all of his papers have a similar level of polish and James recommends reading them all.
+\end{rmk}
+
+\subsection{No Holes Theorem}
+\begin{thm}[No Holes]
+	For every $i$ between $\depth M$ and $injdim M$,
+	\[\Ext_A^i(k,M)\ne 0.\]
+\end{thm}
+
+\section{May 31, 2019}
+Today we begin our discussion of Artin-Schelter regular algebras.
+\subsection{Some History/Comments}
+In the 1980's ('86), Artin and Schelter started a project of classifying connected graded regular algebras of global dimension 3. 
+This is somehow comparible to quantum or noncommutative projective planes $q\bbP^2$ or $\bbP^2_{nc}$.
+
+There were actualy some examples of such algebras discovered befor this, however. In 1982, 
+Sklyanin found examples of regular algebras of global dimension 4 in the suty of they Yang-Baxter equation and quantum groups.
+
+In 1991, the ``AS'' project was finished by Artin, Schelter, Tate, and Ven den Bergh. All together they constitute about 
+150 pages and the result was that there were 14 families. Later on these were named Artin-Shelter regular algebras.
+
+\subsection{Examples}
+\begin{ex}
+	Let $d=\gldim A$. Then the following are AS-regular:
+	\begin{itemize}
+		\item When $d=0$, only $k$ is AS regular.
+		\item When $d=1$, $k[x]$ is AS regular when $\deg x>0$.
+		\item For $d=2$, if $k=\bar k$, then $k_q[x,y]:=k\langle x,y\rangle/(yx-qxy)$ and $k_J[x,y]=k\langle x,y\rangle/(yx-xy-x^2)$ are examples.
+	\end{itemize}
+\end{ex}
+\subsection{Definitions}
+Recall that $A$ is called connected graded if $A=\oplus_\bbN A_i$ with $1_A\in A_0=k$ and $A_iA_j\subseteq A_{i+j}$. Then $\bbN$ is called 
+an internal (or Adams) grading, which is NOT the complex grading -- in particular the Koszul sign rule does \textbf{not} apply to this grading.
+\begin{defn}
+	A connected graded algebra $A$ is called \textbf{Artin-Schelter regular} (AS regular) if 
+	\begin{itemize}
+		\item $\gldim A=d<\infty$
+		\item Let $k=A/A_{\ge 1}$ be the trivial module. Then 
+	\[\Ext_A^i(k,A)=\Ext_{A^{op}}^i(k,A)=\left\{\begin{array}{lr}0, & i\ne d\\ k,& i=d\end{array}\right.\]
+	\end{itemize}
+\end{defn}
+\begin{rmk}
+	Notmally one requires that the algebra also has finite GK dimension (no idea what this is), but we will assume that our algebras are Noetherian, 
+	which obviates this restriction.
+\end{rmk}
+\begin{prop}
+	If $A$ is connected graded and commutative, then the following are equivalent:
+	\begin{itemize}
+		\item $A$ is AS regular.
+		\item $A$ has finite global dimension
+		\item $A=k[x_1,\cdots,x_n]$.
+	\end{itemize}
+\end{prop}
+\begin{ex}
+	A \textbf{non-example} is that $k\langle x_1,\cdots,x_n\rangle$ is \textit{not} AS-regular.
+\end{ex}
+\begin{rmk}
+	A connected graded algebra is a graded version of a complete local algebra. That is, $A=\varprojlim A/A_{\ge n}$.
+\end{rmk}
+\begin{rmk}\label{rmk-AB}
+	\textsc{(Auslander-Buchsbaum formula)}: Suppose $A$ satisfies the $\chi$ condition (that is, for all finitely generated $A$-modules $M$, $\dim\Ext_A^i(k,M)<\infty$ for all $i$)
+	If $M$ has finite projective dimension then 
+	\[\projdim M=\depth A-\depth M.\]
+\end{rmk}
+
+\subsection{Today's Main Result}
+\begin{thm}
+	Let $A$ be a connected graded Notherian algebra with finite global dimension. Then the following are equivalent:
+	\begin{itemize}
+		\item $A$ is AS regular
+		\item $A$ satrisfies the $\chi$ condition
+		\item For any $X\in\Db(\Amod_{fg})$ (can also switch sides) then $\dim\Ext_A^i(k,X)<\infty,\forall i$.
+		\item $A$ has a balanced dualizing complex.
+	\end{itemize}
+\end{thm}
+To prove this, we will first need two lemmas.
+\begin{lem}
+	Suppose $A$ satisfies $\chi$  and $\gldim A=d<\infty$. Then 
+	\begin{itemize}
+		\item $\depth A=d$
+		\item $\projdim k=d$.
+	\end{itemize}
+\end{lem}
+\begin{prf}
+	Proving this uses the Auslander-Buchsbaum formula from remark~\ref{rmk-AB}. For any finitely-generated module $M$,
+	\[\depth A=\projdim M=\depth M\ge \projdim M\]
+	and so by definition, $\depth A\ge \gldim A=d$.
+
+	To prove the other inequality we need to unpack the definition of depth:
+	\[\depth A=\inf(R\Hom(k,A))\le \sup(R \Hom(k,A))=\projdim k\le \gldim A=d.\]
+	Thus $\depth A=d$. But we also havev $\projdim k$ in there, so the second equality also holds.
+\end{prf}
+\begin{lem}
+	Let $F:\Amod_{fd}\to \modA_{fd}$ and $G:\modA_{fg}\to\Amod_{fg}$ be inverse functors. Then $F({_A}k)=k_A$
+	and $G(k_A)={_A}k.$
+\end{lem}
+\begin{prf}
+	This is almost trivial if you think hard about where a simple trivial module has to be sent.
+\end{prf}
+\begin{prf}[of theorem]
+	(1)$\Rightarrow$(3): By definition we know what $\Ext_A^i(k,A)$ is. Define $\calT$ to be the triangulated subcategory of 
+	$\Db_{fg}(\Amod)$ of all objects $X$ such that $\dim\Ext_A^i(k,X)<\infty$ for all $i$.
+
+	It is easy to check that this is a triangulated subcategory. By the definition of AS-regularity, $A$ is in $\calT$. And we know $\Db_{fg}(\Amod)=\Dbperf(\Amod)=\langle A\rangle$, so $\calT$ is 
+	the entire category $\Db_{fg}(\Amod).$
+
+	\brk
+
+	(3)$\Rightarrow$(2) is left as an exercise, but we can do it by taking $X=M$.
+
+	\brk
+
+	(2)$\Rightarrow$(1) By the first lemma, we have $\Ext_A^i(k,A)=0$ when $i\ne d$ and is finite dimensional otherwise.
+	Let $F(-)=\Ext_A^d(-,A)$ and $G(-)=\Ext^d_{A^{op}}(-,A)$. Now note that 
+	\[F(M)=R\Hom_A(M,A)[-d]=R\Hom_A(M,A[d])\]
+	and let $R=A[d]$. The claim is that $R$ is a dualizing complex. This implies that $F(M)=\bbD(M)$ and $G(N)=\bbD^\circ(N)$,
+	and we can apply the lemma to show that $\Ext_A^d(k,A)=F(k)=k$ and $\Ext_{A^{op}}^d(k,A)=k$, so we're done.
+\end{prf}
+
 \end{document}

### 2019-06-05 Homological Algebra Notes/Hom Alg.tex
Almost the end!

@@
 \section{June 3, 2019}
-Today we're talking about spectral sequences! The primary references are Vakil's ``The Rising Sea'' (specifically section 1.7)
-as well as Weiel and May's ``A primer on spectral sequences.''
+Today we're talking about spectral sequences! Sean Griffin is presenting. The primary references are Vakil's ``The Rising Sea'' (specifically section 1.7)
+as well as Weibel and May's ``A primer on spectral sequences.''
 
@@
 of tensor products of chain complexes, and this motivated Sean's original research.
+
+\section{June 5, 2019}
+Today we are going to discuss Galois cohomology. The Thomases Carr and Browning are presenting.
+
+\subsection{Getting Started}
+The set up here is that we have an extension $L/K$ of fields with $G=\operatorname{Gal}(L/K)$. It ends up that really Galois cohomology is 
+really just group cohomology, transported to this context. There are two modules one may consider: $L^+$ and $L^\times$. Unfortunately,
+\begin{thm}
+	$H^i(G,L^+)=0$ for all $i>0$.
+\end{thm}
+So instead we ocus on using $L^\times.$ Here we still get a nice result:
+\begin{thm}[Hilbert Theorem 90]
+	$H^1(G,L^\times)=1$ 
+\end{thm}
+\begin{rmk}
+	Notice that in general the higher $H^i(G,L^\times)$ may or may not be trivial.
+\end{rmk}
+\subsection{Some Definitions}
+\begin{defn}
+	An abelian extension $L/K$ with Galois group $G$ is said to be of \textbf{exponent $n$} if $\sigma^n=1$ for all $\sigma\in G$.
+\end{defn}
+\begin{defn}
+	A \textbf{Kummer extension} $L/K$ is an abelian extension of exponent $n$ where $k$ contains all $n^{th}$ roots of unity.
+\end{defn}
+\begin{ex}
+	When $n=2$, $L=K(\sqrt{a})$ where $a$ is a nonsquare in $K$ is a Kummer extension.
+
+	When $n=3$, we can let $L$ be the splitting field of $x^3-\alpha$ where $\alpha$ is a non-cube. Then we set $K$ to be the subfield containing the third roots of unity.
+\end{ex}
+
+\subsection{Good Stuff}
+Let $L/K$ be Kummer. Consider 
+\[1\to \mu_n\to L^\times\xrightarrow{n^{th}\text{power}}(L^\times)^n)\to 1\]
+and using Galois cohomology, we get the long exact sequence (leveraging that $H^1(G,L^\times)=1$
+\[1\to \mu_n\to K^\times\to K^\times\cap(L^\times)^n\twoheadrightarrow H^1(G,\mu_n)\to 1\]
+and so since $H^1(G,\mu_n)$ ($\mu_n$ is a trivial module and $G$ is abelian although the latter is not necessary), 
+\[K^\times\cap(L^\times)^n/(K^\times)^n\cong \Hom(G,\mu_n).\]
+
+But then we can use a form of Pontryagin duality,
+\[\Hom(G,\mu_n)\cong\Hom(G,\bbZ/\bbZ)\]
+There is a
+
+And we get an overall correspondence between subgroups of $K^\times\cap (L^\times)^n$ containing $(K^\times)^n$ and 
+Kummer extensions $L/K$ of exponent $n$. The upshot here is that if we pass to some maximal Kummer extension of $K$, (call it $N$), 
+then $K^\times\cap (N^\times)^n=K^\times$, so subgroups of $K^\times$ containing $(K^\times)^n)$ are in bijection with 
+Kummer extensions of $K$ of exponent $n$.
+
+More explicitly the correspondence takes a field $L$ to a group $(L^\times)^n\cap K^\times$ and a group $\Delta$ to $K(\sqrt{\Delta})=K(\sqrt[n]{\alpha},\alpha\in\Delta)$.
+
+\subsection{What this means}
+As an application, we get a correspondence between the subgroups of $\bbQ^\times/(\bbQ^\times)^2$ and Kummer extensions of exponent 2.
+
+\subsection{Global Stuff}
+Let $K$ be a number field (a finite extension over $\bbQ$). Then we get a SES 
+\[1\to K^\times\to I_K\to C_K\to 1\]
+where $I_K$ is the so-called \textbf{idele group} and $C_K$ is the \textbf{idele class group}.
+
+To define these we need some fancy stuff so I am going to ignore it. Something about restricted products. :)
+
+Now if $L/K$ is Galois, we get $K^\times\subseteq L^\times$. This naturally gives us $I_K\subseteq I_L$ and some kind of inclusion $C_K\to C_L$.
+If $G$ is the Galois group for $L/K$, then 
+\[(L^\times)^G=K^\times\quad (I_L)^G=I_K\quad (C_L)^G=C_K.\]
+
+Introduce the following notation:
+\[H^n(L/K)=H^n(\operatorname{Gal}(L/K),C_L)\]
+Then we get the following generalization of Hilbert 90:
+\begin{thm}
+	$H^2(L/K)=1$
+\end{thm}
+\begin{prf}
+	This is at least the interesting bit. Induct of $[L:K]$. Let $\operatorname{Gal}(L/F)$ be a Sylow $p$-subgroup 
+	of the bigger Galois group. Then we get restriction and corestriction maps:
+	\[H^1(L/K)\xrightarrow{res}H^1(L/F)\xrightarrow{cores}H^1(L/K)\]
+	and in particular $res$ is injective on the $p$-subgroup. So our goal will be to show that $H^1(L/F)$ is trivial, 
+	proving that the $p$ part of $H^1(L/K)$ is trivial, and letting it vary over primes.
+
+	Now assume that $\operatorname{Gal}(L/K)$ to be a $p$-group and $\operatorname{Gal}(L/F)$ be normal of index $p$.
+	Then you get an inflation restriction sequence 
+	\[1\to H^1(F/K)\xrightarrow{inf}H^1(L/K)\xrightarrow{res}(L/F)\]
+	and then you can do some number theory to prove what you want.
+\end{prf}
+\begin{rmk}
+	Notice that the above sequence is, in fact, the first three terms of the five term exact sequence you get from the Lyndon-Hochschild-Serre
+	spectral sequence in group cohomology.
+\end{rmk}
+
+\begin{thm}
+	$H^2(L/K)\cong\left(\frac{1}{[L:K]}\bbZ\right)/\bbZ\subseteq \bbQ/\bbZ$ is cyclic of order $[L:K]$
+\end{thm}
+
+\subsection{A slight aside: Tate cohomology}
+Let $G$ be a finite group and $A$ a $G$-module. Let $N=\sum_{g\in G} g\in\bbZ G$. Then we get a map 
+$A_G\xrightarrow{N_G}A^G$. Then define
+
+\[\hat H^n(G,A)=\left\{\begin{array}{lr} H^n(G,A),& n\ge 1\\
+	\coker N_G, & n=0\\
+	\ker N_G & n=-1\\
+H_{-n-1}(G,A),& n\le -2\end{array}\right.\]
+
+It just happens that this ends up working together very nicely!
+
+\begin{prop}
+	Given $0\to A\to B\to C\to 0$ an SES, we get our long exact sequence in Tate cohomology.
+\end{prop}
+\begin{prop}
+	When $G$ is cyclic, $\hat H^n(G,A)\cong\hat H^{n+2}(G,A)$
+\end{prop}
+\begin{prop}
+	We have a nice cup product
+	\[\hat H^m(G,A)\otimes \hat H^n(G,B)\to \hat H^{n+m}(G,A\otimes B)\]
+\end{prop}
+\begin{prop}
+	For all $m$ there exists a ``shift module'' $S_m$ such that 
+	\[\hat H^n(G,A)\cong\hat H^{n-m}(G,A\otimes S_m).\]
+\end{prop}
+\begin{thm}[Tate]
+	Let $G$ be finite, $A$ a $G$ module such that 
+	\begin{itemize}
+		\item $\forall H\le G,\hat H^1(H,A)=0$
+		\item $\forall H\le G, \hat H^2(H,A)$ is cyclic of order $|H|$
+	\end{itemize}
+	Then if $A$ generates $\hat H^2(G,A)$ then we get an isomorphism
+	\[a\cup -:\hat H^n(G,\bbZ)\cong \hat H^{n+2}(G,A)\]
+\end{thm}
+Then setting $n=-2$, we get that 
+\[\hat H^{-2}(G,\bbZ)\cong\hat H^0(G,A)=\coker N_G=A^G/NA\]
+but then 
+\[A^G/NA\cing \hat H^0(G,A)\cong H_2(G,\bbZ)=G^{ab}\]
+where the last thing is the abelianization of $G$.
+
+$C_L^G/NC_L\cong\operatorname{Gal}(L/K)^{ab}$
 

### 2019-06-05 Triangulated Categories/Tri Cat.tex
Almost the end!

@@
 \end{prf}	
+\begin{rmk}
+	In all honesty I lost track of the proof near the end, but I am pretty confused about whether that last part actually proved the equivalence of 4 or not.
+	In class the next day, James mentioned that (4)$\Rightarrow$(2) is a result of Yekutieli in his thesis.
+\end{rmk}
+
+\section{June 3, 2019}
+Today we are talking about Poincar\'e duality and something about Nakayama.
+\subsection{Examples of AS regular algebras}
+Examples abound, but this result gives us several:
+\begin{thm}[Brown]
+	Let $A$ be a connected graded algebra over a field of characteristic zero with finite GK dimension (think noncommutative Krull dimension). If $A$ admits a Hopf 
+	algebra structure, then $A$ must be AS regular.
+
+	Furthermore, $A$ satisfies 
+\[\Ext_{A^e}(A,A^e)=\left\{\begin{array}{rl}
+	0, & \text{GK dim}\ne i\\
+	A,&\text{otherwise}\end{array}\right.\]
+\end{thm}
+
+\subsection{Some Major Theorems}
+This one is courtesy of James and our guest Dan Rogalski!
+\begin{thm}[RRZ]
+	Let $A$ be connected graded. Then the following are equivalent:
+	\begin{itemize}
+		\item $A$ is AS regular.
+		\item $A$ is skew Calabi-Yau.
+	\end{itemize}
+\end{thm}
+Another from Michelle Van den Bergh:
+\begin{thm}[Van den Bergh]\label{thm-VdB}
+	Let $A$ be connected graded and AS regular (or equivalently skew Calabi-Yau). Then for any bimodule $M$,
+	and for some automorphism $\mu$ of $A$,
+	\[\HH^i(M)=\Ext^i_{A^e}(A,M)\cong \Tor_{d-i}^{A^{e}}(A,{^\mu}M)=\HH_{d-i}({^\mu}M)\]
+\end{thm}
+\begin{rmk}
+	This gives us a form of Poincar\'e duality that is similar to the one that comes from topology.
+\end{rmk}
+\begin{defn}
+	Let $M$ be an $A$ bimodule and $\mu$ an automorphism of $A$. Then $'M^\mu$ is a new $A$ bimodule where $A$ 
+	has the same action on the left as $M$ but the right action is twisted by $\mu$.
+\end{defn}
+\begin{defn}
+	An algebra $A$ is called skew Calabi-Yau (CY) if 
+	\begin{itemize}
+		\item $A\in\Dbperf(A^{e}\text{-}\mathbf{mod})$
+		\item There is an integer $d$ and algebra automorphism $\mu$ such that 
+	\[\Ext_{A^{e}}(A,A^e)=\left\{\begin{array}{rl}
+	0, & i\ne d\\ 'A^\mu, & \text{otherwise}\end{array}\right.\]
+	\end{itemize}
+	Here $\mu$ is called a \textbf{Nakayama automorphism} of $A$.
+\end{defn}
+\begin{rmk}
+	In 1988 Kansvitch introduced CW categories. Later in 2006 Gingzburg defined CYalg, the Calabi-Yau category. This 
+	was done in a (mysteriously unpublished) paper that James highly recommends we read.
+\end{rmk}
+\begin{lem}\label{lem-useful}
+	If $X\in\Dbperf(B)$, add $Y\in D(B)$, then 
+	\[R\Hom_B(X,Y)\cong R\Hom_B(X,B)\otimes_B^LY.\]
+\end{lem}
+\begin{prf}
+	The above equation holds when we set $X=B$. This is more-or-less self-evident.
+
+	Then we can take $\calT$ to be the set of objects satisfying the above equation, show it is a triangulated subcategory, then show it's everything 
+	because $\langle B\rangle=\Dbperf(B)$.
+\end{prf}
+\begin{prf}[of theorem~\ref{thm-VdB}]
+	\begin{align*}
+		\HH^i(M)&=H^i(R\Hom_{A^e}(A,M))\\
+		&=H^i(R\Hom(A,A^e)\otimes_{A^e}^LM)\\
+		&=H^i('A^\mu[-d]\otimes_{A^e}^LM)\\
+		&=H^{i-d}('A^\mu\otimes_{A^e}^LM)\\
+		&=H^{i-d}(A\otimes_{A^e}^L(A\otimes 'A^\mu)\otimes_{A^e}^LM)\\
+		&=H^{i=d}(A\otimes_{A^e}^L({^\mu}M))\\
+		&=\HH_{d-i}(^{\mu}M).
+	\end{align*}
+\end{prf}
+\subsection{Another neat result}
+\begin{thm}[S. P. Smith]
+	Let $A$ be a connected graded algebra. Then $A$ is AS regular if and only if $\Ext_A^\ast$ is Frobenius.
+\end{thm}
+
+Next lecture will be our last class and we will discuss the proof to James' theorem mentioned above.
+
+\section{June 5, 2019}
+Today we will finish up establishing that AS regular algebras are precisely the skey CY ones.
+
+As a special calse, say that $L$ is an $N_{\ge 0}$-graded Lie algebra what is finite dimensional over $k$. Then $\calU(L)$ is connected graded and 
+furthermore is AS regular and CY.
+
+\begin{defn}
+	An algebra $A$ is called \textbf{skew Calabi-Yau} if 
+	\begin{itemize}
+		\item $A\in\Dbperf(A^{e}\text{-}\mathbf{mod})$
+		\item $\HH^i(A^e)=\Ext_{A^e}^i(A,A^e)=\delta_{i=d}{'}A^\mu$.
+	\end{itemize}
+\end{defn}
+
+To prove [RRZ], we need several lemmas. (10 in total although we only need four to prove (skew CY)$\Rightarrow$(AS)).
+\begin{lem}
+	\[\gldim A\le \projdim_{A^e}A\]
+\end{lem}
+\begin{lem}
+	\begin{itemize}
+		\item $\gldim A\otimes B\le \projdim_{A^e\otimes B^e}(A\otimes B)$
+		\item $\projdim_{A^e}(A)\le \gldim A^e\le \projdim_{(A^e)^e)}A^e=2\projdim_{A^e}A.$
+	\end{itemize}
+\end{lem}
+Then we also have lemma~\ref{lem-useful} from yesterday, as well as 
+\begin{lem}
+	Let $\sigma$ be an automorphism of $A$. Then 
+	\begin{itemize}
+		\item $('A^\sigma\otimes k)\otimes_{A^e}^LA\cong k$
+		\item $A\otimes_{A^e}^L('A^\sigma\otimes k)\cong k$.
+	\end{itemize}
+\end{lem}
+\begin{prf}
+	Let $P$ be a projective resolution of ${_{A^e}}A$. Then 
+	\[('A^\sigma\otimes k)\otimes_{A^e}^LA=('A^\sigma\otimes k)\otimes_{A^e}P='A^\sigma\otimes_AP\otimes_Ak={^{\sigma^{-1}}}P\otimes_A k={^{\sigma^{-1}}}A\otimes_Ak={^{\sigma^{-1}}}k\]
+\end{prf}
+
+\subsection{Koszul Duality}
+\begin{defn}
+	Let $A$ be a connected graded algebra that is generated in degree 1. Then $A$ is called \textbf{Koszul} if 
+	(there are several equivalent definitions but this is the easiest) 
+	\[E:=\oplus_{i\ge 0}\Ext_A^i(k,k)\] 
+	is generated by $\Ext_A^1(k,k)$.
+\end{defn}
+\begin{thm}[Koszul Duality: Beilinson-Ginzburg-Soergel]
+	let $A$ be a Koszul AS-regular algebra. Then we have an equivalence of categories
+	\[\Db_{fg}(\Amod)\simeq \Db_{fg}(E\text{-}\mathbf{mod}).\]
+\end{thm}
+\begin{cor}
+	$\HH^\ast(A)\cong\HH^\ast(E)$.
+\end{cor}
+\begin{rmk}
+	Notice that we \textbf{do not} get isomorphism on the individual graded pieces above: in general, $\HH^i(A)\not\cong\HH^i(E)$.
+	In fact, there are only finitely many $\HH^i(A)$ nonzero, while this is not true for $\HH^i(E)$.
+\end{rmk}
+
+\subsection{The right context for noncommutative algebraic geometry}
+\begin{defn}
+	$\mathbf{proj} A$ is the category of graded modules over $A$ modulo the finitely generated ones. This is an abelian category.
+\end{defn}
+\begin{thm}[Beilinson-Gelfand-Gelfand]
+	$\Db(\mathbf{proj} A)\cong \Db_{fg}(E\text{-}\textbf{Grmod})/\Dbperf(E\text{-}\mathbf{mod}).$
+\end{thm}
+
+\begin{thm}[Beilinson]
+	$\Db(\mathbf{proj} A)\cong\Db_{fd}(B\text{-}\mathbf{mod})$ where $B\subseteq M_{d\times d}(A)$
+	where $d=\gldim A$, where the matrices look like elements of $k$ down the diagonal and when $j>i$, the element in 
+	the $(i,j)^{th}$ position comes from $A_{j-i}.$
+\end{thm}
+
+\begin{thm}[Orlov]
+	Let $\bbV=\Db_{fd}(\Vectk)$. Then 
+	\[\Db(\mathbf{proj}A)\cong (\bbV\rtimes \bbV)\rtimes\bbV\cdots\rtimes \bbV\]
+	where the above is $d=\gldim A$ copies of $A$. We haven't defined the semidirect product of a category, but it can be done. This is called a semi-orthogonal decomposition of $\Db(\mathbf{proj}A)$.
+\end{thm}
 

### 2019-06-18 Homological Algebra Notes/Hom Alg.tex
Update and create new file

@@
 \input{/home/nico/latex-includes/mydefs.tex}
+
+%Shade definitions
+\theoremindent0cm
+\theoremheaderfont{\normalfont\bfseries} 
+\def\theoremframecommand{\colorbox[rgb]{1,0.9,.55}}
+\newshadedtheorem{defn}[thm]{Definition}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%% Customize Below %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 but then 
-\[A^G/NA\cing \hat H^0(G,A)\cong H_2(G,\bbZ)=G^{ab}\]
+\[A^G/NA\cong \hat H^0(G,A)\cong H_2(G,\bbZ)=G^{ab}\]
 where the last thing is the abelianization of $G$.
@@
 $C_L^G/NC_L\cong\operatorname{Gal}(L/K)^{ab}$
+
+Bah I missed something here.
+
+\section{June 7, 2019}
+Today Christine and Tejas are going to talk about \textit{Abelian Sheaf Cohomology}.
+
+\subsection{Sheaves}
+\begin{ex}
+	The idea here is that we are generalizing functions on spaces. For instance, consider $X=S^2$, the 2 sphere.
+	Then take our fuctions to be ones satisfying $f\in\bbR[x,y,z]/(x^2+y^2+z^2-1)$ and open sets to be the usal ones on the spheres.
+\end{ex}
+\begin{defn}
+	A \textbf{presheaf} of ableiang groups on a toplological space $X$ is a functor 
+	\[\calF:\mathbf{Ob}(X)^{op}\to \Ab\]
+	where $\mathbf{Op}(X)$ is the poset category on the open sets of $X$.
+
+	Generally we suppres the space and the fact we are using abelian groups.
+\end{defn}
+\begin{defn}\label{defn-sheaf}
+	A \textbf{sheaf} of abelian groups on $X$ is a presheaf such that the following are satisfied:
+	\begin{itemize}
+		\item (\textbf{Locality}) If $\{U_i\}$ is an open cover of an open set $U$ and if $s,t\in\calF(U)$ such that 
+		\[s|_{U_i}=t|_{U_i},\forall i\]
+		then $s=t$.
+		\item (\textbf{Gluing}) If $\{U_i\}$ is an open cover of $U$ and if for each $i$ there is a $s_i\in\calF(U_i)$
+		such that $s_i|_{U_i\cap U_j}=s_j|_{U_i\cal U_j}$, then there is a \textbf{global section} $s\in \calF(U)$ such taht $s|_{U_i}=s_i$.
+	\end{itemize}
+\end{defn}
+If we like categories more (WE DO!) then we can equivalently define it thusly:
+\begin{defn}\label{defn-sheaf-2}
+	A sheaf is a presheaf $\calF$ such that for any open $U$ with cover $\{U_i\}$,
+	\[\calF(U)\to \prod_i\calF(U_i)\stackrel{\to}{\to}\prod_{i,j}\calF(U_i\cap U_j)\]
+	where, more generally, we could write this using $U_i\times_UU_j=U_i\cap U_j.$
+\end{defn}
+
+What is a morphism of sheaves, then?
+\begin{defn}
+	Let $\calF$ and $\calG$ be sheaves on $X$ a morphism $\varphi:\calF\to\calG$ is just a natural transformation.
+\end{defn}
+\begin{rmk}
+	We say that $\varphi$ is injective (resp. surjective) if for all $x\in X$, 
+	\[\varphi_x:\colim_{U\ni x}\calF(U)\to colim_{U\ni x}\calG(U)\]
+	is injective (resp. surjective).
+\end{rmk}
+
+\begin{prop}
+	The category of abelian sheaves form an Abelian category. :)
+\end{prop}
+
+\subsection{Operations with Sheaves}
+Notice first that every sheaf is a presheaf. Then for any presheaf $\calF$, we can associate to it a sheaf 
+by ``sheafifying.'' You can peek at Hartshorne for the actual definiton, but basically you add in the global sections that are missing (where local sections glue together to form them)
+and remove global sections that are locally trivial.
+
+More concretely, we get a functor:
+\[()^+:\mathbf{Presh}_X\to\mathbf{Sh}_X\]
+such that $()^+$ is left adjoint to the forgetful functor.
+
+\begin{defn}
+	The \textbf{global sections functor} is a functor 
+	\[\Gamma(X,-):\mathbf{Sh}_X\to\Ab\]
+	defined to be $\Gamma(X,\calF)=\calF(X)$.
+\end{defn}
+\begin{prop}
+	$\Gamma(X,-)$ is left but not right exact.
+\end{prop}
+\begin{prf}
+	That it is left exact can be proved a few ways, but Isaac mentioned that limits commute with limits which seems to be pretty slick.
+
+	The primary counter-example that it is not right exact, consider the space $X=\bbC\setminus \{0\}$ and the sequence:
+	\[0\to\underline\bbZ\hookrightarrow \calO\to \calO^\ast\to 0\]
+	where $\underline\bbZ$ is the constant sheaf, $\calO$ is the holomorphic functions on $X$ and $\calO^\ast$ is the nonzero holomorphic functions on $X$. But then letting 
+	the surjection be $f\mapsto e^{2\pi i f}$, there is clearly no global logarithm function in all cases. 
+\end{prf}
+
+\begin{thm}[Grothendieck `57]
+	$\mathbf{Sh}_X$ has enough injectives.
+\end{thm}
+
+\subsection{Sites}
+The motivation here is that there is a correspondence between affine schemes and commutative rings, and every scheme is glued together in such a way.
+The problem here, though, is that he standard (Zariski) topology is just too coarse to glean anything useful.
+
+The goal here is to put a topology on the category of schemes, forgetting completely about what the underlying topological space is. This motivates \'Etale coverings.
+
+\begin{defn}
+	Let $\calC$ be any category. A \textbf{Grothendieck topology} is a collection of distinguished collections indexed by objects in $\calC$ (think a collection of all coverings of a topological space)
+	such that 
+	\begin{itemize}
+		\item Isomorphisms are coverings
+		\item Coverings of coverings are coverings 
+		\item Base changes of coverings are coverings: If $\{U_i\to X\}$ is a covering and if $V\to X$ is any morphism,
+		then the fiber products (pullbacks) $\{U_i\times_XV\to V\}$ exist and are a covering.
+	\end{itemize}
+	Then $\calC$ along with its Grothendieck topology is called a \textbf{site.}
+\end{defn}
+\begin{ex}
+	If $X$ is a topological space, $\mathbf{Op}(X)$ is a site, called the \textbf{small site} of $X$.
+
+	$\mathbf{Top}$ is a site with jointly surjective open embeddings. Then $\mathbf{Top}/X$ is the \textbf{big site} of $X$.
+\end{ex}
+
+We're not going to talk about schemes here, but we will talk more about schemes.
+
+\begin{defn}
+	A \textbf{sheaf on a site} is a functor $\calF:\calC^{op}\to\mathbf{Set}$ satisfying the equalizer 
+	diagram I wrote in definition~\ref{defn-sheaf-2}.
+
+	An \textbf{abelian sheaf} is a sheaf that factors as $\calF:\calC^{op}\to\Ab$.
+\end{defn}
+\begin{ex}
+	$\Hom_{\mathbf{Top}}(-,X):\mathbf{Top}^{op}\to\mathbf{Set}$. That this is a sheaf (on the big site of $X$) basically follows from the fact that 
+	we have the gluing property in topological spaces.
+\end{ex}
+\begin{rmk}
+	Notice that on the category of topological spaces, this tells us that all representable functors are sheaves!
+\end{rmk}
+\begin{ex}
+	If $X$ is an abelian group object in $\mathbf{Top}$, then $\Hom(-,X)$ is an abelian functor.
+\end{ex}
+\begin{defn}
+	$\Ab(\calC)$ is the category of abelian sheaves on $\calC$. 
+\end{defn}
+\begin{thm}
+	$\Ab(\calC)$ is an abelian category with enough injectives.
+\end{thm}
+\begin{prop}
+	For any object $X\in \calC$, we can define 
+	\[\Gamma(X,-):\Ab(\calC)\to \Ab.\]
+
+	Furthermore $\Gamma(X,-)$ is left exact, but not right exact.
+\end{prop}
+
+Then the goal is to measure failure of exactness of $\Gamma(T,-)$ where $T$ is a terminal object. Since we are in an abelian category we can do this in the usual way.
+
+\begin{defn}
+	If $\calF\in\Ab(\calC)$, $V\in \calC$, then the $i^{th}$ cohomology group of $\calF$ over $V$ is 
+	\[H^i(V,\calF):=R^i\Gamma(V,\calI^\bullet)\]
+	where $\calI^\bullet$ is an injective resolution of your $\calF$.
+\end{defn}
 
@@
 \newpage
-\section{HW2 -- Due XXXXX}
+\section{HW2 -- Due May}
 \begin{prob}
@@
 \begin{prob}
-	Let $(R, \frakm, k)$ be a commutative local ring and $M$ be a finitely generated $R$-module. Prove that the following are equivalent: 
+	Let $(R, \frakm, k)$ be a Noetherian commutative local ring and $M$ be a finitely generated $R$-module. Prove that the following are equivalent: 
 	\begin{enumerate}
@@
 \begin{sol}
-	(a)$\Leftrightarrow$(b): First let $M$ be free (and finitely generated). Then $M\cong R^k$ for some finite $k$. 
-	Then take an injection
+	(a)$\Rightarrow$(b): First let $M$ be free (and finitely generated). Then $M\cong R^k$ for some finite $k$. 
+	Then take an injection $\iota:A\hookrightarrow B$ and consider the induced map $A\otimes_R M\to B\otimes_R M$. Since $M\cong R^k$, we get 
+	that $A\otimes M\cong A^k$ and $B\otimes M\cong B^k$ with the induced map between them being $\iota$ on each summand. This is clearly injective.
+
+	(b)$\Rightarrow$(c) follows since if $M$ is flat, $-\otimes_RM$ is exact, so it preserves injections. Thus in particular 
+	the image of $\frakm\hookrightarrow R$ is also an injection.
+
+	(c)$\Rightarrow$(d) Consider the short exact sequence 
+	\[0\to\frakm\to R\to k\to 0\]
+	and apply the long exact sequence in $\Tor$. But since the induced map from (c) is still injective, the following is exact:
+	\[0=\Tor_1(R,M)\to \Tor_1(k,M)\to \frakm\otimes M\to R\otimes M\to k\otimes M\to 0\]
+	where the equality on the left is because free modules are flat (see above).
+	But then by the assumption in (c), this implies that $\Tor_1(k,M)$ injects into the trivial module, so it must be 0 as well.
+
+	(d)$\Rightarrow$(a) Let $\{m_1,\dots,m_n\}$ be a minimal set of generators for $M$. Then define the map $f:R^n\to M$ via 
+	$f(r_1,\dots,r_n)=r_1m_1+\cdots+r_nm_n.$ Let $I=\ker f$. Then we have the short exact sequence 
+	\[0\to I\to R^n\to M\to 0\]
+	and taking the long exact sequence associated to $-\otimes k$, we get 
+	\[\cdots \Tor_1(k,M)\to I\otimes k\to R^n\otimes k\to M\otimes k\to 0\]
+	and since $\Tor_1(k,M)=0$ and leveraging a couple more isomorphisms,
+	\[0\to I/\frakm I\to k^n\to M/\frakm M\to 0\]
+	is exact.
+
+	But now by Nakayama, a basis for $M/\frakm M$ lifts to a minimal set of generators for $M$, but then $M/\frakm M\cong k^n$. Then by a dimension counting 
+	argument, this forces $I/\frakm I=0.$ Since we assume that $R$ was Noetherian, $I$ is finitely generated, so again we can leverage Nakayama: since the quotient is 
+	trivial, so is $I$ itself. But then $M\cong R^n$ and we're done. \textit{Thanks to Cody for giving me a hint to use Nakayama here.}
 \end{sol}
@@
 \begin{sol}
-
+	\textbf{(a):} Define the form 
+	\[\omega\left(\sum_{g\in G}a_gg,\sum_{g\in G}b_g g\right)=\sum_{g\in G}a_gb_g.\]
+
+	One can easily check that this is a (symmetric!) nondegenerate form on $R\otimes R$. Note that I came up with this 
+	by just guessing, but my first thought was to mirror the natural pairing we get with a vector space and its dual (linear functionals). Using this,
+	one can construct (essentially) the same form. Perhaps this is what was intended, although my guesswork paid off this time. :)
 \end{sol}

### 2019-06-18 Local Duality/local_duality.tex
Update and create new file

\documentclass[12pt]{article}

\usepackage{setspace}

\usepackage{amsmath, amsfonts, amssymb, graphicx, color, fancyhdr, lipsum, scalerel, stackengine, mathrsfs, tikz-cd, mdframed, enumitem, framed, adjustbox, bm, upgreek, xcolor, hyperref}
\usepackage[framed,thmmarks]{ntheorem}
\usepackage[style=alphabetic]{biblatex}
%Set the bibliography file
\bibliography{sources}

%Replacement for the old geometry package
\usepackage{fullpage}

%Input my definitions
\input{./mydefs.tex}

%Shade definitions
\theoremindent0cm
\theoremheaderfont{\normalfont\bfseries} 
\def\theoremframecommand{\colorbox[rgb]{0.9,1,.8}}
\newshadedtheorem{defn}[thm]{Definition}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%% Customize Below %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%header stuff
\setlength{\headsep}{24pt}  % space between header and text
\pagestyle{fancy}     % set pagestyle for document
\lhead{Notes on Local Duality} % put text in header (left side)
\rhead{Nico Courts} % put text in header (right side)
\cfoot{\itshape p. \thepage}
\setlength{\headheight}{15pt}
\allowdisplaybreaks

% Document-Specific Macros

\begin{document}
%make the title page
\title{Local Duality Theorems \vspace{-1ex}}
\author{Nico Courts}
\date{Summer 2019}
\maketitle

\renewcommand{\abstractname}{Introduction}
\begin{abstract}
	These notes are my attempt to understand the current state of the art in ``local duality'' theorems, especially 
	of the kind investigated by Benson, Iyengar, Krause, and Pevtsova (herein abbreviated BKIP) in \cite{BIKPgroupschemes} and \cite{BKIPgorenstein}.

	Furthermore I will continue my investigation into quantum groups and other Hopf algebras and see if some of the methods developed earlier 
	can be applied to this new area.
\end{abstract}

\section{Current State of the Art}
\subsection{Ideas to flush out}
From meeting with Julia, the idea that I have gathered is that these two papers constitute two (slightly) different approaches 
to showing that the notion of Serre duality (which by itself is a ``global'' phenomenon) restricts ($\frakp$)-locally (more on this later)
to an analogous result.

The original result in \cite{BIKPgroupschemes} proves that such a duality exists for finite group schemes, but the objects of study here are 
much simpler (for instance the algebras that arise are Frobenius). Furthermore (this part may be sketchy), the results that are proven 
rely on the (relatively simply) monoidal structure of $G$-modules (where $G$ is a finite group scheme).

The newer result in \cite{BKIPgorenstein} reconstructs this result in the context of Gorenstein rings, which are decidedly less degenerate than 
the case of finite group schemes. What is important here is that this is done although the structure theory of these algebras 
is rather poorly understood, so what this represents is the formation of the idea that this duality somehow can be understood at
a higher level, meaning that it may apply to broader classes of algebras.

\subsection{Background results and definitions}
I have seen a good deal of these results already, but it will be useful to put everything into one place so I can reference them as needed.

First, Serre duality, which comes from \cite{hartshorneAG}:
\begin{thm}
	Let $XS$ be a projective Cohen-Macaulay scheme of equidimension $n$ over $k$. Then for any locally free sheaf $\scrF$ on $X$, 
	there are natural isomorphisms
	\[H^i(X,\scrF)\cong H^{n-i}(X,\scrF^\vee\otimes \omega_X^\circ)'\]
	where $(-)'$ indicates taking a vector space dual, $(-)^\vee$ indicates taking the dual of a locally free sheaf:
	$\mathscr Hom(-,\calO_X)$ (sheaf hom) and $\omega_X^\circ$ denotes a dualizing sheaf on $X$.
\end{thm}

\begin{defn}
	Let $X$ be a proper scheme of dimension $n$. Recall that a dualizing sheaf is a (coherent) sheaf $\omega_X^\circ$ such that there is an isomorphism 
	\[\Hom(\scrF,\omega_X^\circ)\cong H^n(X,\scrF)'.\]
\end{defn}
\begin{rmk}
	In fact, there is something more: that this isomorphism is induced from a natural pairing of $\Hom$ and $H^n$. I won't worry 
	about this too much for now, but it can be found in \cite[p. 240]{hartshorneAG}
\end{rmk}

\subsection{BKIP Paper 1---Finite Group Schemes}
Let's start off with the primary result. We will need some definitions and other results to get us there, but it will give us a sense of 
where we're headed.
\begin{thm}[BKIP) `18]

\end{thm}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%  Bibliography %%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\medskip

\printbibliography

\end{document}

### 2019-07-27 Local Duality/local_duality.tex
Update!

@@
 % Document-Specific Macros
+% Primes and Maximals for the lazy person.
+\newcommand{\p}{\frakp}
+\newcommand{\m}{\frakm}
 
@@
 much simpler (for instance the algebras that arise are Frobenius). Furthermore (this part may be sketchy), the results that are proven 
-rely on the (relatively simply) monoidal structure of $G$-modules (where $G$ is a finite group scheme).
+rely on the (relatively simple) monoidal structure of $G$-modules (where $G$ is a finite group scheme).
 
@@
 First, Serre duality, which comes from \cite{hartshorneAG}:
-\begin{thm}
+\begin{thm}[Serre Duality]\label{thm-serre-duality}
 	Let $XS$ be a projective Cohen-Macaulay scheme of equidimension $n$ over $k$. Then for any locally free sheaf $\scrF$ on $X$, 
@@
 	\[H^i(X,\scrF)\cong H^{n-i}(X,\scrF^\vee\otimes \omega_X^\circ)'\]
-	where $(-)'$ indicates taking a vector space dual, $(-)^\vee$ indicates taking the dual of a locally free sheaf:
-	$\mathscr Hom(-,\calO_X)$ (sheaf hom) and $\omega_X^\circ$ denotes a dualizing sheaf on $X$.
+	where $(-)'$ indicates taking a vector space dual, $(-)^\vee$ indicates taking the dual of a locally free sheaf via the functor
+	$\mathscr Hom(-,\calO_X)$ (sheaf hom) and $\omega_X^\circ$ denotes a dualizing sheaf (\ref{defn-dualizing-sheaf}) on $X$.
 \end{thm}
 
-\begin{defn}
-	Let $X$ be a proper scheme of dimension $n$. Recall that a dualizing sheaf is a (coherent) sheaf $\omega_X^\circ$ such that there is an isomorphism 
+\begin{defn}\label{defn-dualizing-sheaf}
+	Let $X$ be a proper scheme of dimension $n$. Recall that a \textbf{dualizing sheaf} is a (coherent) sheaf $\omega_X^\circ$ such that there is an isomorphism 
 	\[\Hom(\scrF,\omega_X^\circ)\cong H^n(X,\scrF)'.\]
@@
 
+I will also include the definition and explanation of compact generation that I wrote down for my Brown representability presentation.
+\begin{defn}[Compact Object]\label{defn-compact-obj}
+	Let $X\in\calT$ be an object in a triangulated category. Then $X$ is called \textbf{compact} if for every coproduct of objects in $\calT$
+	\[\Hom_\calT\left(X,\coprod_{\lambda\in\Lambda}t_\lambda\right)=\coprod_{\lambda\in\Lambda}\Hom_\calT(X,t_\lambda)\]
+\end{defn}
+\begin{rmk}
+	The idea here is that when the codomain is a product, this equality always exists. The projections give you 
+	a product of maps into the factors and the universal property of products gives you a (unique!) map into the product if you
+	have maps into each of the factors.
+
+	This does not hold in general for coproducts, however! Recall that in additive categories 
+	one has a canonical isomorphism between finite products and coproducts. Since triangulated 
+	categories are additive, this is akin to saying that $X$ is well-behaved in that the maps 
+	from $X$ still split even when the codomain is an arbitrarily large coproduct.
+\end{rmk}
+\begin{defn}[Compactly Generated Triangulated Category]\label{defn-compact-tricat}
+	Let $\calT$ be a triangulated category. Then $\calT$ is called \textbf{compactly generated by a set $G$} of its elements 
+	if $\calT$ is closed under small coproducts and $G$ consists of compact objects of $\calT$ such that for all $X\in\calT$,
+	\[\Hom_\calT(G,X)=0\quad\Rightarrow\quad X=0.\]
+\end{defn}
+\begin{rmk}
+	Here the generators of $\calT$ are elements that have ``sufficient complexity'' to detect all nonzero elements.
+
+	In the more general context where $\calT$ is not necessarily additive, we can generalize this definition to a set of objects 
+	such that for every generator $E\in G$ and any $X,Y\in\calC$, if for every $f,g\in\Hom(X,Y)$ and $h\in\Hom(E,X)$ we have 
+	\[f\circ e=g\circ e\]
+	then 
+	\[f=g\]
+	and here we would say that $G$ has ``sufficient complexity'' to differentiate between any two non-identical maps.
+\end{rmk}
+
+\begin{defn}\label{defn-R-linear-cat}
+	Recall that with such an $R$, an \textbf{$R$-linear category} $\calC$ is one such that the graded ring $\End^\ast_\calC(X)$ admits
+	an $R$ module structure via a graded ring homomorphism $\varphi_X:R\to \End^\ast_\calC(X)$.
+\end{defn}
+
+The following comes from \cite{nlab-serre-functor}. 
+\begin{defn}\label{defn-serre-func}
+	Let $\calS$ be an endofunctor $\calS:\calA\to\calA$ on a $k$-linear category with finite dimensional $\Hom$ objects,
+	where $k$ is any field. Then $\calS$ is called a \textbf{Serre functor} if it is an additive equivalence admitting 
+	bi-functorial isomorphisms
+	\[\phi_{A,B}:\Hom_\calA(A,B)\xrightarrow{\sim}\Hom_\calA(B,\calS(A))^\ast\]
+	for all $A$ and $B$ in $\calA$.
+\end{defn}
+
 \subsection{BKIP Paper 1---Finite Group Schemes}
@@
 where we're headed.
-\begin{thm}[BKIP) `18]
+\subsubsection{Main Results}
+\begin{thm}[BKIP `18]\label{thm-BKIP18}
+	Let $G$ be a finite group scheme over $k$. Fix some homogeneous prime ideal $\frakp\lhd H^\ast(G,k)$ not containing $H^{\ge 1}(G,k)$. Let $C=\gamma_\frakp(\stmod G)$
+	and let $d$ be the Krull dimension of $H^\ast(G,k)/\frakp$. Then the assignment
+	\[M\mapsto \Omega^d\delta_G\otimes_k M\]
+	induces a Serre functor (defn~\ref{defn-serre-func}) for $C$. Thus for all $M$ and $N$ in $C$,
+	\[\Hom_{H^\ast(G,k)}(\Hom_C^\ast(M,N),I(\frakp))\cong\Hom_C(N,\Omega^d\delta_G\otimes M)\]
+	where $I(\frakp)$ is the injective hull of the graded $H^\ast(G,k)$ module $H^\ast(G,k)/\frakp$.
+\end{thm}
 
+\noindent This result follows from a more general result:
+\begin{thm}[BKIP `18]\label{thm-BKIP18-ext}
+	For any $G$ module $M$ and $I\in\bbZ$, there is a natural isomorphism
+	\[\widehat{\Ext}^i_G(M,\Gamma_\frakp(\delta_G))\cong\Hom_{H^\ast(G,k)}(H^{\ast-d-i}(G,M),I(\frakp)).\]
 \end{thm}
+\begin{rmk}
+	Notice that here $\Gamma_\frakp(\delta_G)$ plays the part of the dualizing object in the second theorem, as Serre duality takes 
+	on a similar form when $\scrF$ is a coherent sheaf on a nonsingular projective variety $X$ of dimension $n$:
+	\[\Ext_X^i(\scrF,\omega_X)\cong \Hom_k(H^{n-i}(X,\scrF),k)\]
+\end{rmk}
+
+\subsubsection{Conventions}
+
+Throughout we are fixing $R$, a \textbf{graded commutative, Noetherian} ring and $\calT$, a \textbf{compactly-generated (\ref{defn-compact-obj}, \ref{defn-compact-tricat}) 
+$R$-linear (\ref{defn-R-linear-cat}) triangulated category.} Let $\calT^c$ be the full subcategory of $\calT$ comprising of compact objects.
+
+\subsubsection{Torsion and Localization}
+We should brush up on a bit of how torsion and localization from rings passes to such a $\calT$: We say that $M\in \Rmod$ is \textbf{$\fraka$-torsion} if 
+$M_\frakq=0$ whenever $\fraka\not\subseteq\frakq$. In other words, whenever we invert something in $\fraka$, we kill everything. Analogously,
+we say that $X\in\calT$ is $\fraka$-torsion if $\Hom^\ast_\calT(C,X)$ is $\fraka$-torsion for all $C\in\calT^c$.
+
+Now if $\p$ is a (homogeneous!) prime in $R$,  we say that $M\in \Rmod$ is $\p$-local if the localization map $M\mapsto M_\frakp$ is invertible.
+This means that ``all of the action is at $\p$'' -- recall that in the case of rings, this would mean that $\frakp$ is maximal and contains all non-units,
+so $R$ is local. We say that $X\in\calT$ is $\p$-local if $\Hom^\ast_\calT(C,X)$ is $\p$-local for all $C\in\calT^c$.
+
+We will be devoting significant attention to $\Gamma_\p(\calT)$, the (full, localizing) subcategory of $\calT$ consisting 
+of the $\p$-local $\p$-torsion elements. As far as I can tell, we should be thinking of $R/\p^i$ as an analogy in $\Rmod$.
+
+One of the primary functors in question (the one that extracts the ``local'' nature) is $\Gamma_\p$, which assigns to $\calT$ the 
+set (actually it's a localizing subcategory in $\calT$)
+\[\Gamma_\p(\calT)=\{X\in\calT| X\text{is $\p$-local and $\p$-torsion\}}.\]
+
+The paper points this out as an important result:
+\begin{lem}
+	$\Gamma_\p(\StMod G)$ is finitely generated as a triangulated category by the family of objects $(M//\p)_\p$.
+\end{lem}
+Here $M//\p$ is defined as $M//p_1\otimes\cdots\otimes M//p_k$ where $p_1,\dots,p_k$ generate $\p$ and $M//a$ is defined to 
+be the mapping cone of the map $f_a:k\to \Omega^{-d}k$ that comes from the element $a\in H^d(G,k)$.
 

### 2019-07-27 Workshop Notes/workshop_notes.tex
Update!

\documentclass[12pt]{article}

\usepackage{setspace}

\usepackage{amsmath, amsfonts, amssymb, graphicx, color, fancyhdr, lipsum, scalerel, stackengine, mathrsfs, tikz-cd, mdframed, enumitem, framed, adjustbox, bm, upgreek, xcolor, hyperref}
\usepackage[framed,thmmarks]{ntheorem}

%Replacement for the old geometry package
\usepackage{fullpage}

%Input my definitions
\input{./mydefs.tex}

%Shade definitions
\theoremindent0cm
\theoremheaderfont{\normalfont\bfseries} 
\def\theoremframecommand{\colorbox[rgb]{0.9,1,.8}}
\newshadedtheorem{defn}[thm]{Definition}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%% Customize Below %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%header stuff
\setlength{\headsep}{24pt}  % space between header and text
\pagestyle{fancy}     % set pagestyle for document
\lhead{Workshop Notes} % put text in header (left side)
\rhead{Nico Courts} % put text in header (right side)
\cfoot{\itshape p. \thepage}
\setlength{\headheight}{15pt}
\allowdisplaybreaks

% Document-Specific Macros
% Primes and Maximals for the lazy person.
\newcommand{\p}{\frakp}
\newcommand{\m}{\frakm}
\DeclareMathOperator{\Spec}{Spec}
\DeclareMathOperator{\supp}{supp}
\newcommand{\SH}{\mathcal{SH}}
\DeclareMathOperator{\DMack}{DMack}
\DeclareMathOperator{\Sing}{Sing}
\DeclareMathOperator{\SpcBal}{Spc_{Bal}}
\newcommand{\Rep}{\mathbf{Rep}\,}

%stable module category stuff
\DeclareMathOperator{\uHom}{\underline{Hom}}
\DeclareMathOperator{\uExt}{\underline{Ext}}
\DeclareMathOperator{\uEnd}{\underline{End}}

%group schemes
\newcommand{\Ga}{\bbG_a}

\begin{document}
%make the title page
\title{Workshop Notes\vspace{-1ex}}
\author{Nico Courts}
\date{June 24-28, 2019}
\maketitle

\renewcommand{\abstractname}{Introduction}
\begin{abstract}
	These notes were the ones I took while attending the ``Triangulated Categories in Geometry and Representation Theory'' Workshop 
	at the University of Sydney during the week beginning June 23, 2019.
\end{abstract}

\section{The spectrum of the category of derived Mackey functors}
This talk was given by Beren Sanders on a joint work alongside Irakli Patchkoria and Christian Wimmer.

First we are going to talk about tensor triangulated geometry. Essentially we are looking at 
an essentially small tensor triangulated category and then we want to talk about the ``geometry''.
Analogous to the way we define the spectrum of a ring, we can define 
\begin{defn}
	The \textbf{spectrum} of a tensor-triangulated category $\calK$ is 
	\[\operatorname{Spec}{\calK}=\{\p\subsetneq \calK|\p\text{is a prime tensor triangulated ideal}\}\]
\end{defn}

Then if $X\in\calK$ is an object, then we can define the support of $X$ to be 
\[\operatorname{supp}(X)=\{\p\in\operatorname{Spec}(\calK)|x\notin\p\}\]

We define the closure of a prime to be the set of primes that are contained in it. There are some great universal properties.

\begin{lem}
	\begin{itemize}
		\item $\Spec(D(T)^c)\cong\Spec(R)$
		\item $\Spec(\Db(kG-\mathbf{mod}))\cong \Spec(H^\ast(G,k))$.
	\end{itemize}
	The second result is was generalized by (among others) Julia to group schemes.	
\end{lem}

\subsection{The Stable Homotopy Category}
This is the context for most of the talk for today. Let $\SH$ be the stable homotopy category.
There are a couple sources for objects here. We can consider $\sigma^\infty X$, the infinite suspension of a pointed space, 
or a cohomology theory $\bbE$ (e.g. singular cohomology).

$\SH$ is a tensor-triangulated category although I missed the tensor structure.

\begin{thm}[Balmer `10]
	There is a \textbf{comparison map}, a continuous map
	\[\rho_\calK:\Spec(\calK)\to \Spec(\End_\calK(1))\]
	where $1$ is the monoidal unit (since I haven't set up the macro for the bb font.)
\end{thm}
The idea here is that the object on the right is more easy to understand in general.

So for the example of $\SH^c$, we can compute that the endomorphism ring is $\bbZ$. 
Then if we look at any prime $\p\in\Spec(\bbZ)$, the fiber contains all the extraordinary cohomology theories 
that correspond to computing cohomology over characteristic $p$ fields.

\subsection{The spectrum of \texorpdfstring{$\SH(G)$}{SH(G)}}
If $G$ is a finite group, then $\SH(G)$ is the $G$-equivariant stable homotopy group.
In this case, the endomorphism ring is known as the Burnside ring $A(G)$.

Dress in `69 computed the spectrum of $A(G)$: For each $H\le G$, we ahve a ring homomorphism 
\[f^H:A(G)\to \bbZ\]
where $[X]$ is sent to $|X^H|.$ These induce embeddings:
\[(f^H)^\ast:\Spec(\bbZ)\to\Spec(A(G))\]
where $(p)\mapsto \p(H,p)$ and $(0)\mapsto \p(H,0)$.

\begin{defn}
	Let $O^p(G)$ be the smallest normal subgroup of $G$ such that $G/O^p(G)$ is a $p$-group.
\end{defn}
\begin{thm}[Dress '69]
	$\p(H,p)=\p(K,p)$ if and only if $O^p(H)\sim_G O^p(K).$
\end{thm}
\begin{rmk}
	For example if $p\nmid |G|$ then $O^p(H)=H$ for all $H\le G$. If $G$ is a$p$ group, then $O^p(H)=1$ for all $H$.
\end{rmk}

\subsection{Categorifying Dress' Work}
FOr each $H\le G$, we have the \textbf{geometric $H$-fixed points functor}
\[\Phi^H:\SH(G)^c\to \SH^c\]
that induce embeddings 
\[(\Phi^H)^\ast:\Spec(\SH^c)\hookrightarrow \Spec(\SH(G)^c)\]
where $\calC_{p,n}\mapsto \calP(H,p,n)$. Furthermore the copies of $\Spec(\SH^c)$ cover $\Spec(\SH(G)^c)$. T
\textbf{They do not overlap.}

The speaker, together with Paul Blamer, computed the spectrum of $\SH(G)^c$. The picture is interesting, but there is a topological connecting
them: In particular when $G=C_p$, $\calP(1,p,n)\not\subseteq \calP(G,p,n)$ but $\calP(1,p,n+1)$ is.

\begin{defn}
	For a finite $p$-group $G$, define the ``blue shift'' to be 
	\[\beta(G,n)\]
	which is the smallest $i$ such that $\calP(1,p,n+i)\subseteq\calP(G,p,n)$
\end{defn}
It was shown that for $p$ groups, $\beta(G,n)$ is the free rank of $G$.
\begin{rmk}
	For instance, $\beta((C_p)^r,n)=r$ and $\beta(C_{p^r},n)=1$
\end{rmk}

The bottom line here is that the spectrum of the $G$-equivariant can be understood ``locally''
as $\Spec(A(G))$.

\subsection{``Linearization'' of stable homotopy theories}
Some examples of what we're talking about: We can in some way think of the linearization of $\SH$ as being $\D(\bbZ)$.
Furthermore, the motivic derived category of $k$ could be the linearization of $\SH^{\bbA^1}(k)$.

In our case, we may want to consider the derived category of Mackey functors on $G$ as the linearization of $\SH(G)$. Mackey functors 
are functors that behave well with respect to the group/representation theory (e.g. induction and restriction).

One may think this category is the derived category of the abelian category of $G$-Mackey functors, but 
Kaledin (2008) argues that it should be this other category $\DMack(G)$. The speaker computed the spectrum of this 
category as well as developing a different construction.

One interpretation of the linearization in the earlier cases is that $\SH$ is the derived category of a circle 
and $D(\bbZ)$ is the derived category of the Eilenberg-Maclane spectrum $H\bbZ$. For our example, we can define 
$H\bbZ_G$ to be $H\bbZ$ considered as a trivial $G$-module in $\SH(G)$. Then we say 
\[\D(H\bbZ_G)=\DMack_{Kal}(G)\]
and compute the spectrum.

\begin{thm}
	The extension of scalars functor 
	\[\SH(G)\to D(H\bbZ_G)\]
	induces an embedding 
	\[\Spec(D(H\bbZ_G))\hookrightarrow \Spec(\SH(G))\]
	into the top and bottom layers. 
\end{thm}

The main takeaway her is that for $C_p$ the spectrum of the derived category of Mackey functors lies between that 
of $A(C_p)$ and of $\Spec(\SH(C_p))$ where the first step splits points into two and the second introduces this chromatic splitting.

\section{Supports and Cosupports in the Stable Module Category}
This series of talks was given by Julia Pevtsova.

\subsection{Tensor triangulated geometry for finite group schemes}
Let $A$ be a finite dimensional Hopf algebra and associate to $A$ the category $\Sing(A)$, the singularity category of $A$ which measures how non-semisimple
$A$ is. One of the reasons we ways to approach tensor-triangulated geometry is to develop what points are. As a disclaimer, Julia 
says that there might be some misunderstandings she has here, but that it has been developed by Balmer and Krause.

For some motivation, let $R$ be a commutative ring, $\p\in\Spec R$ and let $M$ be an $R$-module. We can gain local information 
by looking at $M_\p$ or $M\otimes_R k_p$.

Now ler $\calT=\D(R)$ where $\calT^c=\Dbperf(R)$. Take $Y\in\calT$ and $\p\in\Spec(R)$ and we can define $Y_\p$ or 
$Y\otimes_R^Lk_p$. Then we define 
\[\supp Y=\{\p\in\Spec R|Y\otimes_R^Lk_\p\ne 0\}.\]

This has some nice properties:
\begin{itemize}
	\item (Defection Property) $Y\simeq 0\Leftrightarrow \supp Y=\varnothing$.
	\item Invariant under $\otimes, \oplus,$ cones, and syzygies.
\end{itemize}
\begin{thm}[Neeman '92]
	The class of localizing subcategories in $\D(R)$ (here $R$ is Noetherian) are in one to one correspondence via $\supp$ to the subsets in $\Spec R$.

	Furthermore restricting to $\calT^c$, the thick subcategories in $\Dbperf(R)$ correspond to specialization closed subsets in $\Spec R$.
\end{thm}

\subsection{General Definitions and \texorpdfstring{$\SpcBal(\calT^c)$}{SpcBal(Tc)}}
\begin{defn}
	Let $T$ be a tensor triangulated category. We say that $\calC\subseteq\calT$ is \textbf{localizing} subcategory if $\calC$ is a full triangulated 
	subcategory that is closed under $\oplus$.

	$\calC\subseteq \calT^c$ is \textbf{thick} if it satisfies the same conditions.

	$\calC\subseteq\calT$ is a tensor ideal if it absorbs tensor products.
\end{defn}

\begin{rmk}
	In this talk $\calT$ is symmetric but this works in braided categories.
\end{rmk}
\begin{defn}
	The \textbf{Balmer Spectrum of $\calT^c$} is defined via a map $T^c\mapsto \SpcBal(T^c)$ where we send 
	$M\mapsto\supp_{Bal}M$.
\end{defn}

Now knowing $\SpcBal(T^c)$ is equivalent ot classifying thick tensor ideals in $T^c$. When $T^c=\Dbperf(R)$, 
and $\p\in\Spec R$, the map $R\to k_\p$ induces maps 
\[\Dbperf(R)\to \Dbperf(k_\p)=\D\mathbf{Vect}_{k_\p}\]
and thus 
\[\SpcBal \D\mathbf{Vect}_{k_\p}\to \SpcBal\Dbperf(R).\]

\subsection{Doing this in modular representation theory}
Classically we are doing modular representation theory when we have a finite group $G$ and a field $k$ where $\ch k| |G|$.
In essence this is just saying that our representation theory is with non-semisimple rings.

We can do this theory on a series of objects:
\begin{itemize}
	\item Finite groups (as above)
	\item Finite group schemes
	\item Finite supergroup schemes
	\item Small quantum groups
	\item Algebraic group $G$ and algebra $A$ over $G$. Then considering $\operatorname{Bimon}(G,A)$ is one possibility.
	\item Finite dimension pointed Hopf algebras
	\item Lie superalgebras (characteristic zero).
	\item One can continue the list. The idea here is that you have a big enough endomorphism ring.
\end{itemize}

Today we are focusing on finite group schemes (FGS) although we might get to say something about supergroup schemes and quantum groups.

\subsection{Affine Group Schemes}
\begin{defn}
	An \textbf{affine groups scheme $G$ over $k$} is a representable functor $G$ from commutative $k$-algebras to groups.
\end{defn}

\begin{rmk}
	$G(R)\cong \Hom_{\Algk}(k[G],R)$ is how the representation works. $k[G]$ is a commutative Hopf algebra. $G$ is finite if $\dim_kk[G]<\infty$.
\end{rmk}

\begin{defn}
	\[kG:=k[G]^\vee=\Hom_k(k[G],k)\]
	is the \textbf{group algebra} corresponding to $G$. This is a finite dimension cocommutative Hopf algebra.
\end{defn}

\begin{rmk}
	There is an equivalence of categories between finite group schemes and finite dimensional cocommutative Hopf algebras using $G\mapsto kG$. 
	Furthermore, the category $\mathbf{Rep}_kG$ is equivalent to $kG$-$\mathbf{Mod}$ (here uppercase will denote non-finite-dimensional), which we will call $G$-$\mathbf{Mod}$.
\end{rmk}

Then we can define the cohomology for $G$ via the Hochschild cohomology
\[\Ext_G^\ast(k,k)=H^\ast(G,k)=H^\ast(kG,k)\]

Some examples of these are finite groups, restricted Lie algebras (Lie algebra over fields of positive characteristic that also have a $p^{th}$ power operation). 
To construct a restricted Lie algebra, One cal let $\calG$ be some algebraic group. Then $\operatorname{Lie}\calG=\frakg$ is a restricted Lie algebra.
The representation theory of such a $\frakg$ is precisely the representation theory of the (finite-dimensional, cocommutative) Hopf algebra $u(\frakg)$,
the quotient of the universal enveloping algebra that identifies $p^{th}$ powers (in the PBW basis) with the internal $p^{th}$ powers in $\frakg$.

More generally one can consider \textbf{Frobenius kernels}. Let $F:\calG\to\calG$ be the Frobenius map. Then 
$\calG_{(1)}=\ker F$ and more generally $\calG_{(r)}=\ker F^(r)$, the $r^{th}$ iteration of $F$. These are connected
finite group schemes. These things have only a single point! However the group algebra and coordinate algebra really still capture 
all the representation theory of the original group.

If $\calG$ is an algebraic group, then we can compute $\calG_{(1)}$ and $\operatorname{Lie}\calG$ and 
\[k\calG_{(1)}\cong u(\operatorname{Lie}(\calG)).\]

When $\calG=\Ga$, then $k[\Ga]=k[T]$ where $\Delta T)=T\otimes 1+1\otimes T$ and 
\[{\Ga}_{(r)}(R)=\{a\in R|a^{p^r}=0\}.\]
Now $k[{\Ga}_{(r)}]=k[T]/(T^{p^r})$ and $k{\Ga}_{(r)}=k[u_0\cdots,u_{r-1}]/(u_0^p,\dots,u_{r-1}^p)$

Recall that $\StMod G$ is the category of $G$ representations where homs are considered module factoring through projective objects. This becomes a tensor triangulated category 
where the triangles comes from short exact sequences in $\Rep G$ and tensor is the usual one. The shift functor is the syzygy functor $\Omega^{-1}$ and the monoidal unit is $k$.

\begin{rmk}
	We have an exact sequence 
	\[\D(G)\to \K(\mathbf{Inj}(G))\to\StMod G\]
	and 
	\[\Dbperf(G)\to \Db(G)\to\mathbf{stmod} G\cong\Dsing(G).\]
	so in a way this measures the lack of regularity in the sense of commutative algebra.
\end{rmk}

Now here we can talk about $\uEnd^\ast(k)$ as well as 
\[\uHom^\ast(M,N)=\oplus_n\uHom(M,\Omega^{-n}N)\]
and 
\[\uEnd^\ast(k)\cong\widehat\Ext_G^\ast(k,k)=\hat H^\ast(G,k)\]
where the hat denotes Tate cohomology.

Now $\uEnd^\ast(k)$ acts on $\calT$ (on $\calT^c$) by acting on $\uHom^\ast(M,N)$.

THis leads to to Benson-Iyengar-Krause support theory where 
\[R=H^\ast(G,k)\to\hat H^\ast(G,k)\]
acts on $\calT$.

We can try setting $X=\Spec R=\Spec H^\ast(G,k)$ for the space where the support lives.

\subsection{\texorpdfstring{$\pi$}{pi}-points and local cohomological functors}
Recall last time we lookedd at $R=H^\ast(G,k)=\Ext_G^\ast(k,k)$ acting (via $\uEnd^\ast(k)$ on $\StMod G$, where $G$ was a finite group scheme.

For examples we considered $\frakg=\operatorname{Lie}\frakG$ where $\frakG$ is a reducitive algebraic group. Here 
$H^{odd}(\frakg,k)=0$ and $H^{ev}(\frakg,k)\simeq k[\calN]$ where $\calN$ are the set of nilpotents in $\frakg.$

\begin{thm}
	When $\frakg$ is a restricted Lie algebra,
	\[\Spec H^\ast(\frakg,k)=\calN^{[p]}=\{x\in\frakg|x^{[p]}=0\}.\]
\end{thm}
\begin{thm}[Suslin-Friedlander-Bendel '97]
	\[\Spec H^\ast(GL_{n(r)},k)\]
	is the variety of $r$ duples of $p$-nilpotent commuting matrices.
\end{thm}

Some facts about cohomology:
\begin{itemize}
	\item $H^\ast(G,k)$ is a graded commutative algebra.
	\item Friedlander and Suslin in '95 proved $H^\ast(G,k)$ is finite generated (has FG).
\end{itemize}

The idea to develop the support theory is to look in $X=\operatorname{Proj}(R)$ and then proceed via either $\pi$-points or via local cohomological functors (developed by Bens0n, Iyengar, and Krause).

\begin{ex}
	Another motivational example: let $\frakg$ be a restricted Lie algebra. Define $X=\operatorname{Proj}H^\ast(\frakg,k)\simeq\calN^{[p]}$
\end{ex}

\begin{rmk}
	$\frakg_a=\operatorname{Lie}\Ga$ is an algebra that is very small since $u(\frakg_a)=k[x]/x^p$. Now $\StMod k[x]/x^p$ is called a ``tt-field'', and for any $x\in\calN^{[p]}\subseteq\frakg$,
	setting $\frakg_a=\langle x\rangle$ gives us a functor 
	\[\calF_x:\StMod\frakg\to\StMod\frakg_a\]
	induced by inclusion.

	Now $\calF_x$ is a tensor triangulated functor and this functor exists for all $x$ in the nilcone (or its image in the projective closure $X$).
\end{rmk}
\begin{defn}
	Let $M\in\StMod G$. Then the support is 
	\[\supp M=\{x\in\calN^{[p]}|M|_{\langle x\rangle}=\calF_x(M)\neq 0\}\subseteq X=\operatorname{Proj}\calN^{[p]}\]
\end{defn}

The support satisfies all the good properties. (?)

if $G$ is a finite group scheme then as we have seen $kG$ is a finite dimensional cocommutative Hopf algebra.
\begin{defn}
	If $K/k$ is an extension, then 
	\[\alpha:K[t]/t^p\to KG=kG\otimes_k K\]
	is a $\pi$-point if it's flat and there is $\calU\subseteq G_K$ such that $\alpha$ factors through $K\calU$ where $\calU$ is 
	a unipotent abelian subgroup scheme of $G_K$.
\end{defn}

Given such a map $\alpha:K[t]/t^p\to KG$, we get a functor 
\[\calF_\alpha:\StMod G\to\StMod K[t]/t^p\]
where the codomain is a tt-field.
\begin{rmk}
	$\calF_\alpha$ is not a tensor functor, but 
	\[\calF_\alpha(M\otimes N)\cong 0\Leftrightarrow \calF_\alpha(M)\otimes\calF_\alpha(N)\cong 0\]
	and either of these happen if and only if $\calF_\alpha$ is zero on $M$ or $N$.
\end{rmk}
\begin{rmk}
	Note that this is a place where things break down for quantum groups. We no longer have factoring through an abelian subgroup.
\end{rmk}

Now return again to $X=\operatorname{Proj}H^\ast(G,k)$. Then given a $\pi$-point $\alpha$, we get maps 
\[H^\ast(G,k)\xrightarrow{-\otimes_R K}H^\ast(G,K)\xrightarrow{\alpha^\ast}H^\ast(K[t]/t^p,K)\cong K[x]\otimes\bigwedge(\lambda)\]
and this map gives us $P_\alpha$, which is the radical of the kernel of this map. This gives us a map $\alpha\mapsto P_\alpha\in X$.

The nice property here is that $\pi$-points realize all points on $X$:
\begin{thm}
	For all $\p\in X$, there exists an extension $K/k$ and $\pi$-point $\alpha$ such that $P_\alpha=\p$.
\end{thm}

\begin{defn}
	Let $M\in\StMod G$. Then the $\pi$ support is 
	\[\pi-\supp M=\{\p\in X|\calF_{\alpha_\p}(M)=\alpha^\ast_\p(M\otimes_k K)\ne 0\}\]
	notice that the construction here doesn't depent on choice of $\pi$ point, but that is something to show.
\end{defn}
\begin{defn}
	\[\pi-\operatorname{cosupp}(M)=\{\p\in X|\calF^{\alpha_\p}:=\alpha_\p^\ast(\Hom_k(K,M))\ne 0\}\]
	where this is computed ve the corestriction functor (this needs clarification).
\end{defn}
\begin{thm}
	$\pi$-$\supp(M\otimes_k N)=\pi$-$\supp(M)\cap\pi$-$\supp(N)$.
\end{thm}
\begin{thm}[BKIP `18]
	$\pi$-$\supp(M)=\varnothing$ if and only if $M=0$.
\end{thm}

\subsection{Local Cohomology Functors}
Let's take a moment to talk about the classical definition of support. When $R=H^\ast(G,k)=\Ext^\ast_G(k,k)$, we can act on 
$\uHom^\ast(M,M)$. When $M\in\stmod G$, we let $I_M=\operatorname{Ann}_R\uHom^\ast(M,M)$. Then $V(I_M)\subseteq X$ is the \textbf{support of 
$M$,} denoted $\supp M$.

To generalize this, we are going to let $\calT=\StMod G$ and associate to each $\p\in X$ a functor 
\[\Gamma_\p:\calT\to\calT\]
which selects the $\p$-local, $\p$-torsion objects.
\begin{defn}
	If $M\in\StMod G$, we can define the \textbf{localization map} $M\to M_\p$ where 
	\[\uHom^\ast(C,M_\p)=\uHom^\ast(C,M)_\p\]
	for each $C\in\calT^c$. (recall $\calT$ is $R$ linear so hom can be localized regularly.)
\end{defn}
\begin{defn}
	There exists a functorial map 
	\[\Delta\Gamma_{V(\p)}M\to M\to L_{V(\p)}M\]
	where the left part is universal $\p$-torsion and the right (I believe) is $\p$-torsion in the usual way (check the paper).
\end{defn}

Then $\Gamma_\p$ is found by taking the $\p$-local and $\p$-torsion operations in either order.

Now 
\[\Gamma_\p M:=\Gamma_p k\otimes_k M\]
and so
\begin{defn}
	$\supp_{BIK}M=\{p\in X:\Gamma_\p k\otimes M\ne 0\}$ and $\operatorname{cosupp}_{BIK}=\{\p\in X|\Hom_k(\Gamma_\p k,M)\ne 0\}$
\end{defn}

And then zooming out, we define 
\[\Gamma_\p(\calT)=\Gamma_\p(\StMod G)=\{M\in\calT|\supp_{BIK}(M)\subseteq\p\}\]

\begin{prop}
	\begin{itemize}
		\item $\pi$-$\supp(\Gamma_\p k)=\p$
		\item $\pi$-$\supp=\supp_{BIK}$ (formal from detection, tensor product)
	\end{itemize}
\end{prop}
\begin{cor}[Balmer, Friedlander-P]
	$\SpcBal(\stmod G)\cong\operatorname{Proj}H^\ast(G,k)$ (even on the level of schemes).
\end{cor}

\subsection{Localizing ideals in \texorpdfstring{$\StMod G$}{StMod G}}
Here we use the BIK local-global principle: to classify localizing $\otimes$ ideals, we need to know that $\Gamma_\p\calT$ are minimal. The ``wishful thinking''
here (which we'll flush out next time) is as follows: if we know $\pi$-$\operatorname{cosupp}=\operatorname{cosupp}_{BIK}$,
then we can use ``Neeman's minimality lemma'' which says that $\calC\subseteq\calT$ is minimal if for all $M$ and $N$ in $\calC$,
$\Hom^\ast_k(M,N)\ne 0$.

Next time we will see some more and a proof!

\subsection{Day 3}
Recall the context: we were letting $X=\operatorname{Proj}H^\ast(G,k)$ and $\calT=\StMod G$ and then looking at
\[\Gamma_p=\{M\in\calT|\supp M\subseteq \p\}\]

\begin{lem}
	$\m\in X$ is a closed point. Then $\Gamma_\m \calT$ is minimal.
\end{lem}
\begin{prf}
	We begin with a fact: $\m\in\pi$-$\supp M$ if and only if $\m\in \pi$-$\operatorname{cosupp} M$.

	To show that $\Gamma_\m T$ is minimal, it suffices to show that $\Hom_k(M.N)\ne 0$ for all $M$ and $N$. But 
	\[\pi\text{-}\operatorname{cosupp}\Hom_k(M,N)=\pi\text{-}\supp M\cap \pi\text{-}\operatorname{cosupp}N\]
	But since the $\pi$ support of $M$ is $\m$ and since $\m$ lies in the $\pi$-cosupport of $N$, then the intersection is nonempty.
	Thus since the cosupport is nontrivial, the module itself is nontrivial.
\end{prf}

\subsection{Reduction to closed points}
We want a Koszul object (in rep theory a Carlsson object): for every $b\in H^d(G,k)\cong\uHom^d(k,\Omega^{-d}k)$, we get an object
$k//b$ formed by the cone of $b:k\to\Omega^{-d}k$. Then for any sequence $\underline{b}=(b_1,\dots,b_n)$, we set 
\[k//\underline{b}=k//b_1\otimes\cdots\otimes k//b_n.\]

Now if we have any finite field extension $K/k$ (the idea here is we would like to take $k$ itself, but ), we can choose a $\m\in X_K=\operatorname{Proj}H^\ast(G,K)$ lying over $\p\in H^\ast(G,k)$.
Then by taking the additional generators for $\m$, we can pick a $\underline b$ such that $\m=\sqrt{\tilde\p+\underline b}$ where $\tilde\p$ is the lifting of $\p$ in the bigger ring.

Then we can consider the (Sylow?) functor for restriction of scalars $\operatorname{Res}^{G_K}_G:\StMod_K G\to \StMod G$
and look at the image of $\Gamma_\m(K//\underline b)$.
\begin{lem}
	$\Gamma_\m(K//\underline b)\downarrow_G\simeq \Gamma_\p k$
\end{lem}
\begin{cor}
	the image $\operatorname{Res}_G^{G_K}$ of $\Gamma_\m\StMod_K G$ in $\Gamma_\p\StMod_k G$ is dense.
\end{cor}
\begin{cor}
	$\Gamma_\p$ is minimal for all $\p$.
\end{cor}

\subsection{Support theory}
The following is tied closely to the following results:
\begin{prop}
	$\pi$-$\supp M=\varnothing$ if and only if $M\cong 0$.

	That is, the $\pi$ support has sufficient strength to detect things (``detection'').
\end{prop}
\begin{prop}
	For all $\p\in X$ there is an $\alpha$ with $\calF_\alpha$ realizing $\p$.
\end{prop}

I am unsure about how this fits in, but she also mentioned the detection of nilpotents in $H^\ast(G,k)$.
\begin{thm}[Quillen '71]
	If $G$ is a finite group, $g\in H^\ast(G,k)$ is nilpotent if and only if for all elements in an abelian $p$-group $E\subseteq G$,$g\downarrow_E$ is nilpotent.
\end{thm}

This means $H^\ast(E,k)\cong k[x_1,\dots,x_n]\otimes \bigwedge^\ast(\tau_1,\dots,\tau_n)$
\begin{thm}[Chouinard '76]
	if $G$ is a finite group, then $M$ is projective if any only if for all elementary abelian $E\le G$.
\end{thm}

Though these results we get that Jarrod's work on the lattice of elementary abelian subgroups of specific groups correspond directly to the lattice (modulo conjugacy and some finite 
group actions) of affine subspaces in spec of the ring.

Now if $\frakg$ is a Lie algebra (say $\frakg=\operatorname{Lie}\calG$ where $\calG$ is a reductive group), 
$H^\ast(\frakg,k)=k[\calN]$. Then fix some Borel subgroup $\calB\subseteq \calG$. Take any $\lambda\in \calG/\calB$ (which is apparently a flag variety?)
and extract from this a subalgebra $b_\lambda\subseteq\frakg$. Then extract a functor 
\[\calF_\lambda:\StMod\frakg\to\StMod b_\lambda\]
\begin{thm}
	Then if $M\in\StMod\frakg$, $M\cong 0$ if and only if $\calF_\lambda(M)=0$ for all $\lambda.$
\end{thm}
Let $G=\calG_{(r)}.$
\begin{defn}
	A \textbf{one-parameter subroup} is
	\[\bbG_{a_(r)}\to\calG_{(r)}\]
	where $k\bbG_{a_{(r)}}\cong k[u_0,\dots,u_{r-1}]/(u_0^p,\dots, u_{r-1}^p)$.
\end{defn}
\begin{thm}
	Nilpotence and projectivity are detected on one-parameter subgroups (up to scalar extension).
\end{thm}
Here we define the $\pi$-point:
\[K[t]/t^p\xrightarrow{t\mapsto u_{r-1}} k\bbG_{a_{(r)}}\to K\calG\]

\brk

But then we get an isogeny 
\[\operatorname{Mor}(\bbG_{a_{(r)}},G)\simeq\operatorname{Spec}H^\ast(G,k)\]

\begin{defn}
	A finite group scheme is elementary if 
	\[\calE=\bbG_{a_{(r)}}\times(\bbZ/p)^s\]
\end{defn}
\begin{thm}[Suslin '76]
	$\xi\in H^\ast(G,k)$ is nilpotent if any only if for all $K/k$ and for any elementary abelian subgroup $\calE\subseteq G_K$,
	the restriction $\xi_K\downarrow_\calE$ is nilpotent.
\end{thm}

\section{Smoothness and Properness}
Greg Stevenson gave this series of lectures.

\subsection{Beginning definitions}

Today we will be working up to dg categories. In what follows let $k$ be a fixed base field (although it works without being a field, this 
is a nice assumption). Recall the definition of a chain complex $\cdots\to X^i\xrightarrow{d^i} X^{i+1}\to\cdots$.

Recall that a chain morphism is a sequence of maps indexed by $\bbZ$ such that the obvious diagrams commute. Denote by $\Ch(k)$ the category of chain complexes of $k$
vector spaces. We have the \textbf{suspension} or \textbf{shift functor} in this category, denoted $\Sigma$. This is an autoequivalence of $\Ch(k)$.

Given a chain map $f:X\to Y$, we define $H^i(X)=\ker d^i/\Im d^{i-1}$. This gives us a functor $H^i:\Ch(k)\to \Vectk$. We say that $f$ is a quasi-isomorphisms if it induces
isomorphisms on cohomology.

We can form the mapping cone of $f$, $\cone f$ (think this is kind of a kernel and kind of a cokernel)
\[\cone(f)=(\Sigma X\oplus U,(\begin{smallmatrix}d_{\Sigma X} & 0\\ f & d_Y\end{smallmatrix}))\]

Recall the definition of acyclic objects. Then 
\begin{lem}
	$f:X\to Y$ is a quasi-isomorphism if and only if $\cone(f)$ is acyclic.
\end{lem}
\begin{rmk}
	This supports the idea that this is in fact a kernel and cokernel in a way.
\end{rmk}

Begin $\Ch(k)$ admits a symmetric monoidal structure from the total complex of the associated double 
complex (with some signs flipped). This defines a functor $-\otimes -$ into $\Ch(k)$.

We also have internal hom (aka the $\hom$ complex). Here 
\[\hom(X,Y)^n=\prod_{i\in\bbZ}\Hom_k(X^i,Y^{i+n})\]
with differential given by the graded commutator:
\[d(f)=d_Y^{i+n} f^i-(-1)^nf^{i+1}d_X^i\]

\begin{rmk}
	This is measuring, in a graded way, 
\end{rmk}
\begin{ex}
	$Z^0\hom(X,Y)=\ker d^0_{\hom(X,Y)}=\{\text{degree zero maps} f| fd=df\}$. Then we can compute that 
	\[H^0\hom(X,Y)\]
	which is precisely the homotopy classes of maps from $X$ to $Y$.
\end{ex}

\brk

We have a tensor hom adjunction (from here let $\Ch=\Ch(k)$)
\[\Ch(X\otimes Y,Z)\cong \Ch(X,\hom(Y,Z))\]
for all $X,Y,Z$ so we have the counit in the form of coevaluation
\[\varepsilon:X\otimes\hom(X,Y)\to Y\]
and given a $Z\in\Ch$, we can consider 
\[X\otimes\hom(X,Y)\otimes\hom(Y,Z)\xrightarrow{\varepsilon_Y\otimes 1}Y\otimes \hom(Y,Z)\xrightarrow{\varepsilon_Z} Z\]
and so via the adjunction we get a composition map 
\[\hom(X,Y)\otimes\hom(Y,Z)\xrightarrow{\circ}\hom(X,Z).\]

This composition map is associative. We also get unit maps $1_X:k\to\hom(X,X)$ determined by the image of $\id_X\in\Ch(X,X)$ in $\hom(X,X)$.
These maps act as units for $\circ$, using the fact that $k$ is the unit of $\otimes_k$.

So using $1_X$ and $\circ$, we get something that looks like a category! This is the primordial example of a dg-category.

\subsection{Examples and extensions}
Give $X\in\Ch$, we get a monoid $\hom(X,X)\in\Ch$. What does that mean? Well, we have a map $\hom(X,X)\otimes\hom(X,X)\to\hom(X,X)$
which is just composition, and a unit $1_X:k\to\hom(X,X)$. This is an example of a differential graded algebra. I.e. we have an algebra $A$ equipped with 
a degree 1 differential $d$ which is a graded derivation.
This is also an example of a dg-category.

\subsection{dg-categories}
\begin{defn}
	$\calA$ is a dg-category if it is given by a collection of objects along with a complex 
	$\calA(a,b)\in\Ch$ for each $a,b\in\calA$ as well as unit maps 
	\[1_a:k\to\calA(a,a)\]
	and composition maps 
	\[\circ:A(b,c)\otimes A(a,b)\to A(a,c)\]
	such that composition is associative and unital with respect to the unit maps.
\end{defn}
\begin{rmk}
	The slick way of saying this is that a dg-category is a category enriched over chain complexes. :)
\end{rmk}

The $\hom$ complexes make $\Ch$ into a dg-category. I will call this $\scrC$. If $A$ is a dga, consider the dg-cat $BA$, 
which is the category of a single object with hom set $A$.

If $B$ is any $k$-linear category we can view it as a dg category $\tilde B$ with $\tilde B(b,c)=B(b,c)$ in degree zero.

Given a dg category $\calA$, we can do a couple things to get a boring category: first, you can take the underlying category $Z^0\calA$ that has the same objects 
and the maps are the zeroth cocycles of $A(a,b)$ for each $a$ and $b$. For example $Z^0\scrC=\Ch$.

We can also compute the homotopy category of $\calA$, denoted $H^0\calA$ which again has teh same objects, but 
\[H^0\calA(a,b)=H^0(\calA(a,b)).\]
Now $H^0\scrC=K(k)$, the homotopy category of chain complexes.

\begin{ex}
	If $R$ is a $k$ algebra, then we get a dg category over $k$ $\scrC(R)$ with objects chain complexes of $R$-modules 
	and where the hom complexes are $\hom_R(X,Y)$.
\end{ex}

\subsection{dg modules and derived categories}
Given a pair of dg categories $\calA$ and $\calB$, a dg functor $F:\calA\to \calB$ consists of an object assignment
along with a map of $\hom$s that preserve units and composition. There is also a notion of being a dg-natural transformation 
and dg-equivalence (this is the part of this we will need), and so on.

Let's finish off today with dg modules. 
\begin{defn}
	Given a small dg-category $\calA$ (notice they are always locally small!), a \textbf{right $A$-dg-module} $X$ is 
	a dg functor $\calA^{op}\to\scrC$.

	Equivalently, for all $a\in\scrA$, $X(a)\in\Ch$ and for every $a,b\in\calA$ a map 
	\[\calA(a,b)\to\scrC(Xa,Xb)\]
	which is equivalent via the tensor hom adjunction to a map 
	\[\calA(b,a)\otimes Xa\to X_b\]
	which looks a lot more like a module action. :)
\end{defn}
Let $\mathbf{Mod}$-$\calA$ be the category of dg modules and dg natural transformations.

\brk

Recall that we had $\mathbf{Mod}$-$\calA=[\calA^{op},\Ch]$. The objects are dg functors and maps are natural transformations with a 
hexagonal naturality diagram (I don't know what htis is).

This is a honest category, but we can upgrade it to a dg category as follows: for each $X,Y\in\mathbf{Mod}$-$\calA$, we define $\mathscr{M}od\calA(X,Y) = \operatorname{eq}(D)$ where $D$ is the diagram 
\begin{center}
	\begin{tikzcd}
		\prod_{c\in\calA}\Ch(Xc,Yc)\ar[r,"\lambda",bend right]\ar[r,"\rho",swap,bend left]& \prod_{a,b\in\calA}\Ch(\calA(b,a),\scrC(Xa,Xb))
	\end{tikzcd}
\end{center}

$\lambda$ is a way to define a right action by precompositon and $\rho$ is the same but right action and postcomposition. So in a way this is saying the two actions are compatible.
This actually ends up being exactly what we usually ask for when we want naturality.

\begin{rmk}
	$Z^0\scrM od\calA(X,Y)=\mathbf{Mod}$-$\scrA$.
\end{rmk}
\begin{rmk}
	$\scrM od\calA(X,Y)=\int_{a\in\calA}\Ch(Xa,Ya)$, if the notation suits you.
\end{rmk}

We can extend notions of cones, suspension, quasi-isomorphisms, etc, pointwise to $\scrM od \calA$. For instance, if $X\in\scrM od\calA$, we define $\Sigma X$ to be the dg functor 
$A^{op}\xrightarrow{X}\Ch\xrightarrow{\Sigma}\Ch$. Again, just define the cone pointwise and since it is functorial (note: this is not true in the 
derived theory, but it still is here.)

We will call $X$ in $\scrM od\calA$ \textbf{acyclic} if $X(a)$ is acyclic for all $a$.

\subsection{Some special dg modules}
Whenever $a\in\calA$, there is a corresponding representable dg module $\hat a=\calA(-,a)$. The dg functor structure here is given by composition.
In this case we get 
\begin{lem}[Strong Yoneda]
	There is a dg functor $\calA\to\scrM od\calA$ sending $a\mapsto \hat a$ which is dg fully faithful. In fact for every $X\in\scrM od\calA$,
	\[\scrM od\calA(\hat a,X)\cong X(a).\]
\end{lem}

In particular, $\scrM od\calA(\hat a,-)$ preserves all things defined pointwise (including quasi-isos, acyclics, etc).

\begin{defn}
	The derived category $\D(\calA)$ is 
	\[\D(\calA)=\mathbf{Mod}\text{-}\calA[\text{quasi isomorphisms}]^{-1}=H^0\scrM od\calA/{\text{acyclic complexes}}\]
	which are analogous to what we usually do with abelian categries.
\end{defn}
\begin{prop}
	$\D(\calA)$ is locally small.
\end{prop}

Inside of $\D(\calA)$ we can consider the images of the representable dg functors $\hat a$. We want to define a thing that sits over this 
\[\mathbf{Perf}\calA\]
to be the smallest full dg subcategory of $\scrM od\calA$ containing the image of the Yoneda embedding, closed under cones, suspensions, and homotopy retracts.

\begin{thm}
	Let $\calA$ be a small dg category. Then $\D(\calA)$ is a compactly generated triangulated category and the compacts are $\D(\calA)^c=H^0\mathbf{Perf}\calA$.
\end{thm}

In a sense, you want to remember that $\mathbf{Perf}\calA$ determines $\D(\calA)$. In particular, fif $\calA$ and $\calB$ are 
small dg categories and we have a \textbf{quasi-equivalence} $F:\mathbf{Perf}\calA\to\mathbf{Perf}\calB$ (that is, $F$ induces a quasi-isomorphism on $\scrH om$s)
and $F$ descends to an essential surjection on the homotopy categories of the $\mathbf{Perf}$ objects, then $\D(\calA)\xrightarrow{\sim}\D(\calB)$.

In such a case we say that $\calA$ and $\calB$ are \textbf{derived Morita equivalent}.
\begin{ex}
	Consider the path algebra of the $A_2$ quiver $kA_2\cong (\begin{smallmatrix}
		k&k\\0&k
	\end{smallmatrix})$. Then $\mathbf{Perf}kA_2\cong\scrC^b(\operatorname{proj} kA_2)$.

	This is derived Morita equivalent (in fact just equivalent) to $B\cong R\Hom_{kA_2}(S_2\oplus S_2,S_1\oplus S_2)$.
\end{ex}

In general, if we let $\calK$ be a small triangulated category with $\tilde\calK$ a dg category such that $\calK\cong H^0\tilde\calK$ as triangulated categories 
where $H^0\tilde\calK$ inherits is triangulated structure from $H^0\scrM od\hat\calK.$ This is called an \textbf{enhancement of $\calK$.}

Call $g\in\calK$ a \textbf{generator} if the smallest thick subcategory containing $g$ in $\calK$ is all of $\calK$. We call $g$ a \textbf{strong generator} if there exists a uniform bound on the number of cones needed to build
any object (there is a global bound on the number of cones we have to take to build any object). In this latter case we call $\calK$ \textbf{regular} (in the sense of algebraic geometry).

\begin{thm}[Keller]
	If $g\in\calK$ is a generator, then $\tilde\calK$, an enhancement of $\calK$, is quasi equivalent to $\mathbf{Perf}\tilde\calK(g,g).$
\end{thm}
\begin{rmk}
	What this is saying is that if $\calK$ (i.e. $H^0\tilde\calK$) ahs a generator then $\tilde \calK$ is derived morita equivalent to a dg algebra.
\end{rmk}

In practice, most triangulated categories we care about (at least the algebraic ones -- the topological ones don't work here) come from dg algebras.

\subsection{Examples, properness, and smoothness}
\begin{ex}
	Let $R$ be a finitely generated $k$ algebra viewed as a dg algebra concentrated in degree zero. Then we have 
	\[\Ch(R)=\scrM od R\supseteq \Ch^{-,b}(\operatorname{proj}R)=\underline\D^b(R)\supseteq \mathbf{Perf} R=\Ch^b(\operatorname{proj} R)\]
	which is analogous to the inclusion 
	\[\D(R)\supseteq \D^b(R)\supseteq\Dbperf(R)\]
\end{ex}

\begin{ex}
	If $X$ is a scheme of finite tipe over $k$, then we can define $\operatorname{Perf} X$ to be the collection of injective resolutions of chain complexes of 
	quasicoherent sheaves on $X$ such that these complexes are perfect in $\D_{qc}(X)$. The point here is that 
	\[H^0\operatorname{Perf}X\cong\Dbperf(X).\]
\end{ex}

\subsection{Properness}
Suppose $\Lambda$ is a finite dimensional $k$ algebra. If $M,N\in\mathbf{Perf}\Lambda$, then $\mathbf{Perf}\Lambda(M,N)\in\mathbf{Perf}k$ up to quasi-isomorphism.
That is $H^\ast\mathbf{Perf}\Lambda(M,N)$ is finite deimensional over $k$.

If $X$ is a proper $k$ scheme, then for $E,F\in\operatorname{Perf}X$, we have $\operatorname{Perf}X(E,F)\in\mathbf{Perf}k$. That is, the cohomology on a proper scheleme is finite dimensional.

\begin{defn}
	We say that a dg category over $k$ $A$ is \textbf{proper} if for all $a,b\in\calA,$ $\calA(a,b)\in\mathbf{Perf}k$ up to quasi isomorphism.
\end{defn}

An example of a proper category is $\operatorname{Perf}\Lambda$ where $\Lambda=k[\varepsilon]/\varepsilon^2$. But on the other hand, $\underline\D^b(\Lambda)$ is not proper. For instance,
notice that a resolution is of $\Lambda$ with $\varepsilon$ maps everywhere, but $\Hom(P,P)$ computes $\Ext^\ast(k,k)$, which in every degree is $k$.

\begin{lem}
	Properness is a derived Morita invariant. That is a dg category over $k$ $\calA$ is proper if and only if $\operatorname{Perf}\calA$ is proper.
\end{lem}
\begin{rmk}
In partucular an algebra $\Lambda$ is proper (i.e. finite dimensional) if and only if $\operatorname{Perf}\Lambda$ is.
\end{rmk}
\begin{thm}[Lipmon?]
	If $X$ is a scheme of finite type over $k$, then $X$ is proper iff $\operatorname{Perf}X$ is.
\end{thm}

\subsection{Smoothness}
Let $L=\bbF_p(t)$ and $R=\bbF_p(t^{1/p})$. Then $R$ is a field so $\gldim R=0$, but 
$R\otimes_LR$ is nologingler finite global dimension since $t^{1/p}\otimes 1-1\otimes t^{1/p}$ is nilpotent.

\begin{defn}
	Recall that a $k$ algebra $R$ is \textbf{smooth} if $R\in\operatorname{Perf}R^e$.
\end{defn}
\begin{lem}
	If $R$ is smooth over $k$, then $\gldim R<\infty$.
\end{lem}

Some geometry: $X$ of finite type over $k$ is smooth if $X\times_k \bar k$ is regular (where $\bar k$ is the algebraic closure of $k$). If $X$ has enough locally free sheaves (resolution property),
then this is equivalent to 
\[\Delta_\ast\calO_X\in\operatorname{Perf}(X\times_k X)\]
where $\Delta$ is the diagonal map $X\to X\times X$.

In an attempt to make a common generalization:
\begin{defn}
	Let $\calA$ and $\calB$ to be two dg-categories over $k$. Define $\calA\otimes_k \calB$ to be the category of pairs of 
	objects from $\calA$ and $\calB$, written $a\otimes b$ and make the maps 
	\[\calA\otimes \calB(a\otimes b,a'\otimes b')=\calA(a,a')\otimes \calB(b,b').\]
\end{defn}
Then we set $\calA^e=\calA^{op}\otimes_k\calA$. Then we consider $\scrM od\calA^e$ to be the categories of $\calA,\calA$-bimodules.

\begin{ex}
THe diagonal $\calA^e$ modules $\Delta:\calA^e\to\Ch$ is the map $\calA(-,-)$, sending a pair $a\otimes a'$ to $\calA(a,a')$.
\end{ex}
\begin{rmk}
	This gives us another definition: $\calA$ is smooth over $k$ if $\Delta=\calA(-,-)\in\operatorname{Perf}\calA^e$.
\end{rmk}
\begin{lem}
	Smoothness is a derived Morita invariant. Thus it agrees with the notion for algebras.
\end{lem}
\begin{thm}[Lunts-Schn\"urer]
	If $X$ is finite type, separable, and has enough locally free sheaves, then $X$ is smooth iff $\operatorname{Perf}X$ is smooth.
\end{thm}

There are lots more exotic examples including things you wouldn't really expect to be smooth.
\begin{thm}[Auslander, Elogin-Lunts-Schn\"urer]
	If $\Lambda$ is a finite dimensional algebra over $k$ such that $\Lambda/\operatorname{rad}\Lambda$ is separable over $k$ (e.g. when $k$ is algebraically closed), then 
	$\underline\Db(\mathbf{mod}\,\Lambda)$ is smooth.
\end{thm}
\begin{rmk}
	Notice that by contrast that $\operatorname{Perf}\Lambda$ is smooth if and only if $\Lambda$ is. This is a smaller object, so violates our geometric intuition.

	The upshot here is that we can regard $\operatorname{Perf}\Lambda\hookrightarrow\underline\Db(\Lambda)$ to be a kind of resolution of singularities.
\end{rmk}
\begin{thm}[Lunts]
	If $X$ is a separable scheme of finite type over a perfect field, then $\underline\Db(X)$ is smooth.
\end{thm}
By Morita invariance, if $E$ is a dg algebra and $\operatorname{Perf}E$ is smooth, then $E$ is smooth. For instance, take a prime $p\ge 3$, let $k=\bbF_p$
and consider $\underline\Db(k C_p)\cong\operatorname{Perf}E$ where $E=R\Hom_{kC_p}(k,k)$.

But now notice that
\[H^\ast(E)=Ext_{k C_p}^\ast(k,k)=H^\ast(C_p,k)=k\langle\tau,\theta\rangle\]
where $|\tau|=1$ and $|\theta|=2$ and the above is the free graded commutative algebra on these things. Now since $|\tau|=1$, $\tau^2=0$, so $H^\ast E$ has infinite global dimension, so $H^\ast E$ is not smooth.

In general, one can't hope to say much about $H^\ast A$ when $A$ is smooth. However,
\begin{thm}[Raedschelder-S.]
	If $A$ is a smooth dg algebra is smooth and connective ($A^{\ge 1}=0$), then $H^0A$ is smooth.
\end{thm}

\subsection{Regularity and smoothness}
What exactly does smoothness buy us?
\begin{thm}
	If $A$ is a smooth dg algebra then $A$ is regular in the sense that $A$ strongly generates $H^0\operatorname{Perf}A$.
\end{thm}
Regularity buys one quite a lot: we get representability results for cohomological (dg) functors. The idea here is that even if we don't have the infinite coproducts we need for Brown,
but somehow the ``needing uniformly finite cones'' of strong generation takes care of this for us.

\begin{rmk}
	It is not always true that for any Noetherian scheme that $\Db(X)$ is even regular.

	There are many examples of dg categories that are regular but not smooth. For instance, $\operatorname{Perf}k[[x]]$ is regular, but not smooth.
\end{rmk}

Smoothness is also a strong finiteness condition on $\calA$. 
\begin{thm}[To\"en]
	Let $\calA$ be a smooth dg category. Then there exists a generator $g\in\operatorname{Perf}\calA$ so 
	$\calA$ is derived Morita equivalent to $\operatorname{Perf}\calA(g,g)$, a dg algebra.
\end{thm}

\subsection{Ideas for proofs}
Say $\calA$ is a smooth dg category, so $\Delta\in\operatorname{Perf}\calA^e$. What are the perfects over the enveloping algebra? For $a\otimes b\in \calA^e$, 
write $\widehat{a\otimes b}$ for $\calA(a,-)\otimes\calA(-,b)\in\operatorname{Perf}\calA$. But then 
\[\operatorname{Perf}\calA^e=\operatorname{thick}(\widehat{a\otimes b}|a\otimes b\in\calA^e)\]
then the upshot here is that computing the thick subcategory containing something is built out of finitary pieces. That is, 
for anything in $\operatorname{Perf}\calA^e$, $\Delta\in\operatorname{thick}_{\calA^e}(\widehat{a_i\otimes b_i}|i\in\calI)$
where $\calI$ is a finite set.

So if $X\in\operatorname{Perf}\calA$, consider $X_\calA\otimes_\calA {_\calA}\Delta_\calA\cong X\in\scrM od\calA$
but then 
\[X=X\otimes_\calA \Delta\in\operatorname{thick}_\calA(X\otimes_\calA \widehat{a_i\otimes b_i}|i\in\calI)\]
and futhermore 
\[X\otimes_\calA\widehat{a_i\otimes b_i}= X(a_i)\otimes_k \hat b_i\]
so this is a big direct sum of shifts of $\calA(-,b_i)$.

Then by Thomason's this implies that $X\in\operatorname{thick}(b_i|i\in\calI)$, so the $b_i's$ genreate $\operatorname{Perf}\calA$, 
whence $\calA$ is morita equivalent to $\oplus_{i,j\in\calI}\calA(b_i,b_j)$.

Finally since $\Delta$ can be built in some finite number of cones, $X$ can be build in the same (or fewer) number of cones.

\begin{rmk}
	In general, smoothness and properness are often dual to one another. That is they often come in pairs. FOr example, if $\Lambda$ is a finite dimensional dg algebra over $k$, then $\operatorname{Perf}\Lambda$ is proper.

	Then $\operatorname{Perf}k$ is the monoidal unit for a closed monoidal structure on a dg category over $k$ $[\operatorname{Perf}\Lambda,\operatorname{Perf}k]=[\Lambda,\operatorname{Perf}k]\cong\underline\Db(\Lambda^{op})$ and the thing on the right is smooth.
	This is not an exact duality because there are counterexamples but it's mostly correct.
\end{rmk}

\section{(Bounded) t-structures and approximable categories}
Let $R$ be a ring. Consider the category $\D(R)$, the derived category of $R$ modules. This is a triangulated category with 
coproducts. Write $\Sigma$ for the suspension.

We have two (full) subcategories $\Dpos(R)$ and $\Dneg(R)$, which are elements such that cohomology 
vanishes below and above zero. Notice that $\Dneg(R)$ is stable under $\Sigma$ and $\Dpos(R)$ is stable 
under $\Sigma^{-1}$. Sometimes we write $\D^{\ge 1}(R)$ (e.g.).

It is not too hard to check that $\Hom(\Dneg(R),\D^{\ge 1}(R))=0$. For each $x\in\D(R)$, there is 
a triangle 
\[Y\to X\to Z\]
where $Y\in\Dneg(R)$ and $Z\in\D^{\ge 1}(R).$

Finally $\Dneg(R)\cap\Dpos(R)=\Rmod$.

\subsection{t-structures}
\begin{defn}
	A \textbf{t-structure} on a triangulated category $\calT$ is a pair of full subcategories $\calT^{\ge 0},\calT^{\le 0}$ satisfying 
	\begin{itemize}
		\item $\Sigma\calT^{\le 0}\subset \calT^{\le 0}$
		\item other stuff I missed.
	\end{itemize}
\end{defn}

The \textbf{heart of $\calT$} is $\calT^{\ge 0}\cap\calT^{\le 0}$ which 
is an abelian subcategory of $\calT.$ We say $\calT^{\le 0}$ and $\calT^{\ge 0}$ are the \textbf{aisle and coaisle} of $\calT$,
respectively.
\begin{rmk}
	The t-structure is determined entirely by the (co) aisle. This is because
	\[\calT^{\ge 0}=(\calT^{\le 0})^\perp:=\{Y\in\calT|\Hom(\calT^{\le -1})\}\]
\end{rmk}

We have truncation functors that are right and left adjoints to $\calT^{\le 0}\hookrightarrow\calT$ and $\calT^{\ge 0}\hookrightarrow\calT$, respectively.

\subsection{Aisles vs Pre-Aisles}
Every aisle $\calT^{\le 0}$ is closed under positive suspensions, taking summands, and extensions. If $\calT$ has coproducts, then $\calT^{\le 0}$ is closed under coproducts.
\begin{defn}[Keller-Vossieck]
	A subcategory $\calS\subseteq\calT$ is a \textbf{pre-aisle} if it is closed under extensions, positive suspensions, and taking summands.
\end{defn}

A question one may ask is when is a pre-aisle an aisle? Keller-Vossieck said that a pre-aisle $\calS\subseteq \calT$ is an aisle if and only if $\calS\hookrightarrow\calT$
has a right adjoint.

Let $S\subseteq\calT$ be a set of objects. Write 
\begin{itemize}
	\item $\langle S\rangle$ to be the smallest thick triangulated subcategory of $\calT$ contatining $S$.
	\item $\langle\overline{S}\rangle$ is the same as above but closed under coproducts.
	\item $\langle S\rangle^{(-\infty,0]}$ is the smallest pre-aisle in $\calT$ containing $S$
	\item $\langle\overline{S}\rangle^{(-\infty,0]}$ is what you'd think.
\end{itemize}

\begin{thm}[Neeman]
	Let $\calT$ be a well-generated (whatever that means) category and $S\subseteq\calT$ a set of objects. Then 
	there exists a t-structure with aisle $\langle \overline{S}\rangle^{(-\infty,0]}$
\end{thm}

As an example, when $\calT=\D(R)$, then $\calS=\langle \overline{R}\rangle^{(-\infty,0]}$ is the aisle of the standard t-structure.

Another example: Let $\calT=\Db(R)$ where $R$ is Noetherian. Let $S=\langle R\rangle^{(-\infty,0]}=\calT_c^{b\le 0}$ is the aisle of the restricted standard t-structure.
This structure is bounded. That is, $\calT$ is the intersection of the shifted versions of the usual structures.

Why do we care about bounded t-structures? One reason is that if $\calT$ has a bounded t-structure, then $\calT$ is generated by its heart. Another reason is that 
it is an ingredient in Bridgeland stability. Finally is something to do with K theory that will show up in David's talk.

\subsection{A nontrivial pre-aisle}
Bondal found the following example of a pre-aisle which is not an aisle:
\begin{ex}
	Let $\calD=\Db_{coh}(\bbP^2(k))$ where $k$ is algebraically closed of characteristic zero. Let 
	\[D_1^{\le 0}\langle \calO(1),\calO(2)\rangle\]
	and 
	\[D_2^{\le 0}=\langle\calO(-1),\calO(-2)\rangle.\]

	These are both aisles as well, but $D_1^{\le 0}\cap D_2^{\le 0}$ is a pre-aisle but it can be shown (though significant computation)
	to not be an aisle.
\end{ex}

\subsection{Neeman's Great Idea}
Let $\calS=\calT^{\le 0}$ for some t-structure on $\calT$. Let $x,y\in\calT$. We get as surjective map 
\[\left\{x\to x\to y|s\in\calS\right\}\twoheadrightarrow\Hom(x,y^{\le 0})\]
by virtue of the fact that any map from a truncated thing to an object factors through its truncation. Neeman showed that under 
a certain equivalence relation we can make this into an isomorphism.

The quotient $H_\calS(x,y)\cong\Hom(x,y^{\le 0})$ can be defined for any pre-aisle $\calS$, so we get a homological functor 
$H_\calS(-,y):\calT\to\Ab$.

\begin{prop}[Neeman]
	Let $\calS$ be a pre-aisle. $\calT-\calT^{\le 0}$ for some t-structure on $\calT$ if and only if $H_\calS(-,y)$ is representable for all $y\in\calT$.
\end{prop}
\begin{rmk}
	Notice that here is where we run into set theoretic issues. One must show that such a functor is set-valued to apply Brown representability.
\end{rmk}

\subsection{Approximating Objects}
\begin{defn}
	A \textbf{metric} on a category is a function that assigns a positive real number (length) to every morphism in a way that the triangle inequality is satisfied.
\end{defn}

Now every t-structure on $|calT$ gives rise to a metric on $\calT$. To understand it, it suffices to specify the balls 
\[B_n=\{x\in\calS|0\to X\text{ has length}\le 1/n\}\]
Taking the balls to be $B_n=\calT^{\le -n}$ gives rise to a ``good metric''. I missed the discussion here unfortunately.

Recall the definition of compact generators.

\begin{defn}
	Let $\calT$ be a triangulated category with coproducts. Then $\calT$ is \textbf{approximable} if there exists a compact 
	generator $G\in\calT$ and a t-structure and an integer $A>0$ such that 
	\begin{itemize}
		\item $G\in T^{\le A}$ and $\Hom(G,T^{\le -A})=0$.
		\item For every object $y\in\calT^{\le 0}$, there exists a triangle
		\[x\to y\to x\]
		with $z\in\calT^{\le -1}$ and $x\in\calT^{\ge 0}$ (I think).
	\end{itemize}
\end{defn}

Some examples due to Neeman: $\D(R)$ is approximable. IF $X$ is quasicompact and separated, then $\D_{qc}(X)$ is approximable. Finally the homotopy category of spectra is approximable.

\subsection{\texorpdfstring{$\calT^b_c$}{Tcb}}
Suppose that $\calT$ has a compact generator $G$. Consider the t-structure generated by $G$ (the aisle is generated by $G$). Call $\calT_c^-$
the full subcategory of $\calT$ such that for all $n>0$ there exists a triangle $x\to y\to z$ with $x$ compact and $z\in\calT^{\le -n-1}$. That is, $y$ is being approximated by compact objects.

\begin{defn}
\[\calT_c^b=\calT^b\cap\calT^-c\]
\end{defn}
\begin{rmk}
	The categories above do not depend on the choice of generator.
\end{rmk}
\begin{rmk}
	If $\calT$ is approximable, then $\calT_c^-$ and $\calT_c^b$ are thick.
\end{rmk}

For some examples, when $\calT=\D(R)$, $\calT^b=\Db(R)$ and $\calT_c^-=\Dneg(R)$. If you look at quasicoherent sheaves you also get something nice.

\section{Cluster tilting modules for mesh algebras}

This talk was given by Sira Gratz in conjunction with Erdmann and Lamberti.

The motivation here is that the clustering aspect comes from combinatorics (e.g. cluster algebras) and the tilting portion 
refers to tilting theory/modules.

Thoughout let $\Lambda$ be a finite dimensional algebra over $k$. If $\Lambda$ has a CT-modules, this implies that the representation dimension of $\Lambda$ is less than or equal to three. What does this mean?
That $\Lambda$ is close to being ``representation finite''. Auslander proved that $\operatorname{repdim} \Lambda =2$ iff it is representation finite.

That is $0\to T''\to T'\to M\to 0$ exists for any $M$, where $T'$ and $T''$ are indecomposibles. So in some way if you understand the indecomposable theory
you can reconstruct the structure of any module using only two ``layers'' of this information.

\subsection{What are CT-modules?}
\begin{defn}
	Let $\calT$ be a triangulated or Abelian category. Then $\calC\subseteq\calT$ is \textbf{cluster tilting}
	if it is functorially finite (This means you have pre-covers and pre-envelopes for every object in your category -- roughly $\calC$ approximates $\calT$) and the folowing are equivalent for all $M\in\calT$:
	\begin{itemize}
		\item $M\in \calC$
		\item $\Ext^1(C,M)=0$ for all $C\in\calC$
		\item $\Ext^1(M,C)=0$ for all $C\in\calC$
	\end{itemize}
\end{defn}

If $\calT=\mathbf{mod}$-$\Lambda$, then $A$ in this category is a CT-modules if $\operatorname{add}(A)$ is CT.
A good place to look to start is $\Lambda$ that are self-injective.

\begin{thm}[Erdmann-Holm]
	If $\Lambda$ is self-injective and has a CT-module, then for all $\Lambda$ modules $M$ have complexity less than 1. 
\end{thm}
\begin{rmk}
	Basically this talks about how quickly the dimension of the elements in a ``minimal'' projective resolution grows. Complexity zero means that a module has finite 
	projective dimenson and complexity one means that the dimension of the modules in the resolution is bounded.
\end{rmk}

\subsection{Mesh Algebras (of Dynkin type)}
The benefit of these algebras is that they have periodic resolutions, so we can apply the previous theorem.

One example comes from the quiver 
\begin{center}
	\begin{tikzcd}
		0\ar[r,"\bar\alpha",bend right,swap] & 0 \ar[l,"\alpha",bend right]\ar[r,"\beta",bend right] & 2 \ar[l,"\bar\beta",bend right,swap]
	\end{tikzcd}
\end{center}

The idea here is you look at a Galois cover of this diagram and then use graph automorphisms to introduce a ``twist'' to 
create a mesh algebra.

\begin{thm}[Geiss-Leckers-Schr\"oer]
	Mesh algebras of type $A,D,$ and $E$ have CT-modules.
\end{thm}
\begin{thm}[Gratz-Erdmann-Lamberti]
	All mesh algebras (of Dynkin type) have CT-modules.
\end{thm}

Another result:
\begin{thm}[Darpo-Iyama]
	If $\calC$ is $k$-linear and locally bounded such that $\operatorname{add}(\calC)$ is Krull-Schmidt $G$ admissibly,
	there is a one-t-one correspondence between $G$-equivariant CT-subacategories of $\calC$ and CT-subcategories of $\calC/G$.
\end{thm}
\subsection{Mutation}
Alnother result from the original group:
\begin{thm}[GLS]
	If $\Lambda$ is a mesh algebra of type $A,D,$ or $E$, and $T$ is a CT-module in $\mathbf{mod}$-$\Lambda$. Let $X$ be a non-projective indecomposable 
	summand  then there exists a unique $Y\not\cong X$ such that $T/X\oplus Y$ is CT. There are also some short exact sequences.
\end{thm}

The theory here works on the idea that $\Ext^1(X,Y)\cong D\Ext^1(Y,X)$ which is not in general true. But in our case 
we get that (if $\gamma$ is the automorphism of $\Lambda$ induced by $\sigma$) that $\Ext^1(X,Y)\cong D\Ext^1(Y,{_\gamma}X)$, where 
${_\gamma}X$ is the twisting of the $\Lambda$ action on $X$ by $\gamma$.

A futher result is that each of the CT modules are $\gamma$-equivariant except in the exception case of $P(G_2)$, the mesh algebra 
corresponding to $D_4$, that has an order 3 automorphism. In this case, there exists such a CT module, but they are not all (something).

\section{Triangulated Categories (and syzygies) associated to singularities of algebraic varieties.}
This talk was given by Jesse Burke.

Here let $P=k[x_0,\dots,x_n]$ and let $P_j$ be the homogenous degree $j$ polynomials. An ideal $I$ of $P$ is \textbf{homogeneous} if there 
exists a set of homogeneous generators. A $P$-module is homogeneous if $M=\oplus M_i$ and $P_j\otimes M_i$ maps to $M_{j+i}$.

Write $P(m)_j=P_{m+j}$.

\subsection{Hilbert and Syzygies}
Hilbert was one of the first to systematically use syzygies. As an example, let $P=k[x_0,\dots,x_3]$ and let
\[R=\frac{P}{(x_0x_2-x_1^2,x_0x_3-x_1x_2,x_1x_3-x_2^2)}\]
which gives us an embedding of $\bbP^1\hookrightarrow\bbP^3$.

A free resolution here is 
\[0\leftarrow R\leftarrow P^1\leftarrow P(-2)^3\leftarrow P(-3)^2\leftarrow 0\]
In general it is difficult to look at abunct of equations and talk about what properties a variety has. One thing Hilbert did was 
write down the Hilbert series $H(R)=\sum_n \dim_k R_nt^n$. Hilbert proved that this is always rational with a particular denominator.

Let $R$ be a graded quotient ring of polynomial ring $P$. A graded free resolution of a graded $R$ module is what you'd expect.
A graded free resolution is minimal if all differentials have entries in $(x_0,\dots,x_n)$.

\subsection{Examples}
\begin{ex}
	Let $R=k[x,y]$ and $M=k=R/(x,y)$, the residue field. Then we have resolution 
	\[0\to R(-2)\xrightarrow{\binom{-y}{x}} R(-1)^2\xrightarrow{(x,y)} R^1\to k\to 0\]
	which is the same as the Koszul complex.
\end{ex}
\begin{ex}
	Now let $R=k[x,y]/(x^2)$. Then a resolution of $k$ looks like 
	\[\cdots R(-3)^2\to R(-2)^2\to R(-1)^2\to R\to k\to 0\]
	where eventually the maps become periodic (so the syzygies are periodic).
\end{ex}
\begin{ex}
	When $R=k[x,y]/(x^2,y^2)$, the $k^{th}$ syzygy is generated by $k$ elements.
\end{ex}
\begin{ex}
	When $R=k[x,y]/(x^2,xy)$, You get that the rank grows like the Fibonacci numbers.
\end{ex}

\section{Modular realisations of derived equivalences in representation theory}
This talk was given by Daniel Chan on work with Tarig Abdelgadir and Boris Lerner.

We always work over an algebraically closed field $k$ of characteristic zero. The motto here is that Moduli stacks are a fruitful way to study a finite dimension algebra because 
they are essentially a machine to construct functors.

As usualy, we should fix some discrete invariants of a finite dimensional $A$ module. You can start by 
fixing $\dim M$. Slightly more subtlely, you can fix an idempotent $e$ and fix $\dim Me$.

Recall the path algebra: start with a quiver $Q$ (directed graph) without oriented cycles.
Let $kQ$ be the path algebra with $k$-basis all paths of length $\ge 0$. Multiplication is concatenation or zero.

For every vertex $v$, there is a path $e_v$ of length zero which is idempotent.

\subsection{King's Interpretation of Beilinson's derived equivalence}
Let $Q$ be the Kronecker quiver on two vertices with two arrows between them in the same direction. $A=kQ$. Fix the dimension of our modules to be $(1,1)$.

This got pretty interesting but also complicated so I just listened. :) If I ever get into stacks this seems like a good guy to read.

\section{Groups, Spherical Twists and Stability Conditions}
This was a three-part lecture given by Asilata Bapat, Anand Deopurkar, and Anthony Licata.

The goal of this first talk is to paint a picture of the aim of this ``movement''. More specifically, one goal is to use homological algebra
to prove some things about groups. This will hopefully cement this theory of a thing that group theorists care about.

\subsection{What kinds of groups?}
One kind are Artin-Tits braid groups. We will not use full generality but will suffice here. Fix some finite connected graph (no loops) $\Gamma$.
Then this graph yeilds $B_{r_\Gamma}$, a group given by generators $\sigma_i$ for each vertex. Then we add the relations 
$\sigma_i\sigma_j=\sigma_j\sigma_i$ if $i$ and $j$ are not connected by an edge and the braid relation $\sigma_i\sigma_j\sigma_i=\sigma_j\sigma_i\sigma_j$ if they are.

Very little is known about these groups. For instance, there are no known solutions to the word problem (when are two words the same) and the conjugacy problem (when are two elements conjugate). 
People also don't know the what the centers of these groups are or whether they have finite representation dimension.

It is not known whether $B_{r_\Gamma}$ are linear. However, these groups do act on ``nice'' triangulated categories. For example: 
\begin{ex}
	Let $\scrP$ denote the $\bbC$-linear additive category generated by $\{P_i\}_{i\in \Gamma_0}$ (indexed by the vertices). Then the hom sets are given by 
	\[\Hom_\scrP(P_i,P_j)=\left\{\begin{array}{lr}
		\bbC\cdot 1\oplus \bbC \cdot x_i, & i=j\\
		\bbC\cdot y_{ij}, & (i,j)\in\Gamma_1\\
		0, & \text{otherwise}.
	\end{array}\right.\]
	To define the composition, one can think of there as being a graded structure on $\Hom^\ast$ given by $\deg(x_i)=2$ and $\deg(y_{ij})=2$ and composition respects grading.
\end{ex}

Now let $\calT_\Gamma=K^b(\scrP_\Gamma)$. 
\begin{thm}[Huerfano-Khovanov]
	$B_{r_\Gamma}$ acts on $\calT_\Gamma$.

	The generator $\sigma_i$ acts by twisting in $P_i$:
	\[\sigma_i(X):=\cone(P_i\otimes_\bbC \Hom(P_i,X)\xrightarrow{ev}X)\]
\end{thm}

We have a couple conjectures:
\begin{conj}
	The action above is faithful.
\end{conj}
and even more strongly (and importantly):
\begin{conj}
	The moduli space of Bridgeland stability conditions(this word may be wrong) $\operatorname{Stab}(\calT)$ is contractible.
\end{conj}

\subsection{What kind of picture are we trying to paint?}
As a small (seeming) aside: How does one study mapping class groups of surfaces?

Let $\Sigma$ be a surface, possibly with boundary and punctures. Then 
\[\operatorname{MCG}(\Sigma)=\frac{\operatorname{Diff}_+(\Sigma)}{\operatorname{Diff}_0(\Sigma)}\]
of oriented diffeomorphisms of the $\Sigma$ to itself modulo the ones isotopic to the identity. As an example, if $\Sigma$ is a disk with 
$n$ punctures, the mapping class group is the braid group.

Now the MCG of $\Sigma$ acts on:
\begin{itemize}
	\item $\operatorname{Teich}(\Sigma)$, the Teichm\"uller space, or the moduli space of hyperbolic metrics on $\Sigma.$
	\item The isotopy classes of simple closed (multi-)curves on $\Sigma$.
\end{itemize}

Thurston explained how to unify these two actions. The first thing to notice is that the Teichm\"uller space comes as a manifold while the other set (call it $S$)
doesn't have a clear topology. So in unifying the two, Thurston developed a method of understanding it through charts. The charts are ``train tracks'' which are a bit confusing to me.
They each have three distinguished points and they (through four symmetries) are supposed to categorize each of the possible (maybe simple closed) curve in your space.
I see there being a problem if the distinguished points are not punctures (e.g. if a curve goes through them).

Then Thurston defined a compactification of the Teichm\"uller space. You can embed them into $\bbR^S$, the (nonzero) maps from $S$ to $\bbR$, which itself embeds (obviously)
into its projective closure.

Then the closure of $\operatorname{Teich}$ is the closure of its image in $\bbP(\bbR^S)$. It ends up that the interior is precisely Teich, and the 
boundary has an interpretation as the moduli space of projective measured foliations on $\Sigma.$

So in particular, MCG acts on $\overline{\operatorname{Teich}}=\operatorname{Teich}\sqcup \text{PMF}$. One can show that Teich is homeomorphic to $\bbR^n$ and PMF is homeomorphic to $S^{n-1}$, 
so the closure of Teich is homeomorphic to the closed Euclidean ball. This gives us solutions to various algorithmic problems.

For instance, each $g\in$MCG$(\Sigma)$ gives rise to a dynamical system by studying repeated action of $g$ on the closed ball. In particular, one is generally interested in fixed points which yields 
Neelson-Thurston classification of mapping classes into periodic, reduced, and pseudo-Arasov.

\subsection{Chainging gears}
\textbf{Our goal is to try to recreate this construction using as our group the autoequivalence group of a triangulated category $\Aut(\calT)$.}

We need a bit of a starting point here to even give us reason to believe this is the correct approach. Luckily it becomes clear that (Bridgeland) stability conditions naturally play the role of the Teichm\"uller space.
So then we need to decide what plays the role of $S$, and here we will use spherical (stable) objects.

That is, we aim to study an action of $\Aut(\calT)$ on a compactification of $\operatorname{Stab}(\calT)$. We will talk about these in more detail in talks 2 and 3.
In particular, talk two will include some (rather technical) definitions for $\operatorname{Stab}(\calT)$ and the compactification. In talk three, we will 
describe an example that we understand, fully developed in the world of homological algebra.

\subsection{Definitions and compactification}
FIx a triangualted category $\calT$. Recall that 
\begin{defn}
	A Brigeland stability condition on $\calT$ is $(Z,P)$ where 
	\[P=\{P(\phi)|\phi\in\bbR\}\]
	(notice that $P(\phi)$ is a full subcategory of $\calT$) satisfying 
	\begin{itemize}
		\item $P(\phi+1)=P(\phi)[1]$
		\item If $\phi_1>\phi_2$, then $\Hom(P(\phi_1),P(\phi_2))=0$
		\item If $E\in\calT$ then there exists a unique $\phi_1>\phi_2>\cdots>\phi_n$ and a uniqu fitration 
		\[0=E_0\to E_1\to\cdots E_n=E\]
		where each $E_i\to E_{i+1}$ admits $A_{i+1}\in P(\phi_{i+1})$ and maps 
		\[E_{i+1}\to A_{i+1}\to E_i\]
		this is called the Harder-Narasimhan filtration.
	\end{itemize}

	Furthermore $Z:K_0(\calT)\to\bbC$ such that if $A\in P(\phi)$, then $Z(A)=m(A)\cdot e^{i\pi\phi}$ where $m(A)\in\bbR_{>0}.$
\end{defn}

As a matter of notation:
\begin{itemize}
	\item $P$ is called a \textbf{slicing} of $\calT$
	\item $Z$ is called the \textbf{central charge}
	\item If $\phi\in\bbR$, the objects of $P(\phi)$ are called \textbf{semistable of phase $\phi$} (simple objects of $P(\phi)$ are called \textbf{stable})
	\item If $A\in P(\phi)$, then $m(A)$ is called the \textbf{mass} of $A$. In general $m(E):=\sum m(A_i)$ where $A_i$ are as in the filtration.
\end{itemize}

A slicing gives a bounded t-structure on $\calT$: for any $\phi\in\bbR$, $(P(\ge\phi),P(<\phi+1))$
form your t structure whose heardt is $P([\phi,\phi+1])$. Given any (Z,P), we'll say that $P([0,1])$ is the \textbf{standard t-structure for $(Z,P)$.}

\begin{prop}[Bridgeland]
	Specifing $(Z,P)$ on $\calT$ is equivalent to giving 
	\begin{itemize}
		\item A bounded t structure on $\calT$ with heart $\calA$
		\item Two functions $m:\calA\setminus\{0\}\to \bbR_{\ge 0}$ and $\text{ph}:\calA\setminus\{0\}\to [0,1)$
		such that $Z(A):= m(A)e^{i\pi\text{ph}(A)}$ is a group homomorphism $K_0(\calA)\to\bbC$.
		\item (HN Condition) This is automatic if objects of $\calA$ are of finite length.
	\end{itemize}
\end{prop}
\begin{rmk}
	We can recover $P$ as follows: if $\phi\in[0,1),$ then 
	\[P(\phi)=\{A\in\calA|\text{ph}(A)=\phi\text{ and } \forall 0\ne B\subseteq A, \text{ph}(B)\le\text{ph}(A)\}\]
	otherwise if $\phi\in[n,n+1)$, $P(\phi)=P(\phi-n)[n].$
\end{rmk}

Set the space $\operatorname{Stab}\calT$ to be the set of all stability conditions $(Z,P)$ on $\calT$.
\begin{thm}[Bridgeland]
	$\operatorname{Stop}(\calT)$ is a (complex) manifold.
\end{thm}
In fact, there is a $\bbC$-action on $\operatorname{Stab}\calT$ where if $w=x+iy$ and if $(Z,P)\in\operatorname{Stab}(\calT)$,
then $Z\mapsto e^\omega Z$ (``scaling mass'') and $P(\phi)\mapsto P(\phi+\frac{y}{\pi})$ (``rotating phase''). Then we want to look at $\operatorname{Stab}\calT/\bbC.$

\subsection{Plugging things into yesterday's talk}
We will be replacing Teich with $\operatorname{Stab}\calT/\bbC$, MCG with $\Aut\calT$, and the (honest) sphere with the set of spherical objects of $\calT$ modulo shifts.

\begin{defn}
	An object $A$ is spherical if
	\[\Hom^i(A,A)=\left\{\begin{array}{lr}
		k, & i=0,2\\
		0,& \text{otherwise}
	\end{array}\right.\]
\end{defn}

\subsection{Construction the closure of our space}
We consider the map $\operatorname{Stab}/\bbC\xrightarrow{\iota} \bbP\bbR^S$ that sends $\tau$ to the map $c\mapsto m_\tau(c)$. We will also use the morphism 
$S\xrightarrow{\delta} \bbP\bbR^S$ that sends $A\in S$ to the map $B\mapsto \overline\hom^\ast(A,B)$ where the over line denotes we set $\overline\hom(A,A)=0$.

Set $\overline{\operatorname{Stab}/\bbC}$ to be the closure of the image of $\iota$ in $\bbP\bbR^S$.

\subsection{Some conjectures}
These are pretty loose/wishful thinking. 
\begin{conj}
	$\iota$ is an embedding.
\end{conj}
\begin{conj}
	$\delta$ is an embedding.
\end{conj}
\begin{conj}
	If $\overline{\operatorname{Stab}/\bbC}=\operatorname{Stab}\sqcup \text{Bdy}$, then $S$ is in the boundary and is dense in the boundary.
\end{conj}
\begin{conj}
	$\overline{\operatorname{Stab}/\bbC}$ is a closed ball which is the union of an open ball $\operatorname{Stab}/\bbC$ with boundary a sphere.
\end{conj}

This was really cool. I especially enjoyed the third talk. :)

\section{Bounded t-structures and negative K-theory of stable infinity categories}
This was given by David Gepner. It is joint work with Ben Antieau and Jeremiah Heller.

We've heard a lot about t structures over this conference. Here we will talk about some obstructions to getting t structures that arise from K theory.
First some conjectures:
\begin{conj}[Marco Schlicting]
	If $\calA$ is a small abelian category then $K_n(\calA)=0$ for $n<0$.
\end{conj}
\begin{conj}[A-G-H]
	If $\calC$ is a small stable $\infty$-category with a bounted t structure, then $K_n(\calC)=0$ for all $n<0$.
\end{conj}
\begin{conj}
	If $\calC$ is a small stable infinity category with bounded t-structure, then $K_n(\calC^\heartsuit)\simeq K_n(\calC)$ is an equivalence for all $n\in\bbZ$.
\end{conj}
\begin{rmk}
	Actually for non-negative $n$, this is the so-called ``theorem of the heart'' that may be due to Neeman. Notice that $K(\calA)=K(\Db(\calA))$ where the bounded 
	derived category is not the regular category, but the category where we remember the dg structure.
\end{rmk}

The second conjecture above implies both the first and second.
\begin{rmk}
	$K_0$ is an invariant of the triangulated homotopy category of a dg or stable infinity category. That being said, 
	the entire K theory $K(\calC)$ depends on a dg (or stable) enhancement.
\end{rmk}

\subsection{Background on (stable) \texorpdfstring{$\infty$}{infty}-categories}
The idea here is that an $\infty$-category is a category enriched in spaces, where we actually mean that it is enriched up to coherent homotopy.
So for every $A,B\in\calC$, there is a ``mapping space'' $\operatorname{Map}(A,B)$ with a composition operator that is only well-defined up to homotopy.

\begin{defn}
	An infinity category $\calC$ is \textbf{stable} if it has finite limits and colimits, a zero object, and a square is a pushout iff it is a pullback.
\end{defn}
\begin{rmk}
	In case you need help wrapping your head around limits in infinity categories, it is useful to remember that $\mathbf{Cat}\hookrightarrow\mathbf{Cat}_\infty$
	fully faithfully via a map called the \textbf{nerve}. Furthermore this functor commutes with limits and colimits, so these things kind of what you think they should be.
\end{rmk}
\begin{rmk}
	If $\calC$ is stable, then $\operatorname{Ho}(\calC)$ (the ordinary category with the same objects, but $\Hom_{\text{Ho}(\calC)}(A,B)=\pi_0\operatorname{Map}(A,B))$ is canonically triangualted.
\end{rmk}

You just define $\Sigma A$ to be the pushout of $0\leftarrow A\to 0$ and $\Omega A$ is defined dually. Here distinguished triangles are the ones that come from just doing what you'd hope: compute the cofibration of the 
map $A\to B$ to get $C=B/A$ and then $C/B$ will be the same as $\Sigma A$.

Then $\calC$ is stable if and only if it has finite limits, colimits, 0 and $\Omega=\Sigma^{-1}$.

\begin{defn}
	A t-structure on a stable infinity category is jsut a t-structure on Ho$(\calC)$.
\end{defn}
\begin{rmk}
	If $\calC$ is a stable infinity category, then $\calC^\heartsuit$ is already an abelian (1-)category.
\end{rmk}

\subsection{Alebraic K theory and t-structures}
K-theory $K:\mathbf{Cat}_\infty^{st}\to\mathbf{Sp}$ is a functor from stable infinity categories with exact functors (ones that preserve finite (co)limits) to spectra (one can 
also think graded groups, etc.)

There are lots of places that algebraic and geometric objects give rise to stable $\infty$ categories (e.g. $X$, $\operatorname{Perf}X$, $\Db(X)$)

\begin{defn}
	$K$ is the unit of the convolutional symmetric monoidal structure on the subcollection of $\mathbf{Func}(\mathbf{Cat}_\infty^{st},\mathbf{Sp})$ that are localizing.

	Furthermore, notice that $\mathbf{Cat}_\infty^{st}\hookrightarrow \mathbf{Func}^{loc}(\mathbf{Cat}_\infty^{st},\mathbf{Sp})$ via the Yoneda embedding and K-theory 
	is corepresented by $\operatorname{Perf}(S)$ (lost me here).
\end{defn}

Here a functor $F:\mathbf{Cat}_\infty^{st}\to\mathbf{Sp}$ is localizing if, given a Verdier localization sequence 
\[\calA\hookrightarrow\calB\twoheadrightarrow\calC\cong(\calB/\calA)^{idem}\]
then $F(\calA)\to F(\calB)\to F(\calC)$ is an exact triangle of spectra (and $F$ preserves filtered colimits).

\begin{thm}[Schlichting]
	If $\calA$ is a small abelian category, then $K_{-1}\calA=0$ and if $\calA$ is Noetherican, then $K_{-n}\calA=0$ for all $n>0.$
\end{thm}

\subsection{Results of A-G-H}
\begin{thm}
	If $\calC$ is a small, stable $\infty$-category with bounded t-structure, then $K_{-1}(\calC)=0$.
\end{thm}
\begin{thm}
	If $\calC$ is small, stable $\infty$-category with bounded t-structure with Noetherian heart, then $K_{-n}(\calC)=0$ for all $n>0$.
\end{thm}
\begin{thm}[Nonconnective theorem of the Noetherian heart]
	IF $\calC$ is a stable infinty category with bounded t-structure such that $\calC^\heartsuit$ is Noetherian, then $K(\calC^\heartsuit)\simeq K(\calC)$.
\end{thm}	
\begin{thm}
	Let $R$ be a commutative ring and $A$ a (cohomologically) graded dg $R$-algebra such that $H^0(A)$ is semisimple and $H^i(A)$ is finite generated as a right $H^0(A)$-module,
	then $K_{-n}(A)=0$ for all $n>0$.
\end{thm}
\begin{thm}
	If $i:\calA\to \calB$ is a fully fairthful exact functor of small stable infinity categories equipped with compatible bounded t-structures. Then set $\calC$ to be the Verdier quotient $\calB/\calA$,
	and $\operatorname{Ind}(\calC)$ inherits a t-structure which restricts to a bounded t-structure on $\calC$ if and only if $\calA^\heartsuit\hookrightarrow \calB^\heartsuit$ is the inclusion of a Serre subcategory.
\end{thm}

A major question/conjecture: Given an abelian category $\calA$, can one construct an abelian category $\calB$ containing $\calA$ as a Serre subcategory such that $K(\calB)=0$?

A positive answer to this would give tools necessary to answer the conjectures given in the affirmative.

\section{Local duality for Gorenstein algebras}
This talk was given by Henning Krause on recent work by Benson, Iyengar, and Pevtsova.

Much of Julia's talk sets the scene for this talk! Let $G$ be a finite group, let $k$ be a field. Then if $M$ is a $kG$-module, 
we can compute $H^\ast(G,M)=\Ext^\ast_{kG}(k,M)$, which is a graded module over $R:=H^\ast(G,k)$ and this is a finitely generated algebra.

Pick any $\p\in\Spec R$ and let $I(\p)$ be an injective envelope of $R/\p$. Then the theorem says 
\begin{thm}[BIKP, '19]
	Let $d$ be the Krull dimension of $R/\p$ where $\p\lhd R$ is prime and is not all of $R^{\ge 0}$. Then 
	\[\Hom_R(H^{\ast-d-i}(G,M),I(\p))\cong\widehat\Ext_{kG}^i(M,\Gamma_\p(k))\]
	for all $i\in\bbZ$.
\end{thm}

\subsection{Serre Duality}
\begin{thm}
Let $X$ be a projective nonsingular scheme over a field $k$, and let $n=\dim X$. Then 
\[\Hom_k(H^{n-i}(X,\calF),k)\cong\Ext_X^i(\calF,\omega_X)\]
for all coherent sheaves $\calF$ on $X$ and $i\ge 0$.
\end{thm}

A more modern formulation is that $\Db(\operatorname{coh}X)$ has a Serre functor, $F=-\otimes_X^L\omega_X[n]$.

\begin{defn}[Bendal-Kapraha, '89]
	Let $\calT$ be a $k$-linear triangular category, hom-finite (what does this mean?) $\calF:\calT\to\calT$ is a \textbf{Serre functor}
	if $D\Hom_\calT(X,Y)\cong \Hom_\calT(Y,\calF(X))$ which is natural for all $X,Y\in\calT$. (here he wrote $D=\Hom_k(-,k)$).
\end{defn}

Let $T\in\operatorname{coh}X$ be a tilding object and let $\Lambda=\End_X(T)$. Then we have an equivalence 
\[R\Hom_X(T,-):\Db(\operatorname{coh}X)\xrightarrow{\sim}\Db(\mathbf{mod}\Lambda)\]
That is, $\Db(\mathbf{mod}\Lambda)$ has a Serre functor.

\begin{thm}[Reiten-van den Bergh '01, Happel '87]
	Let $\Lambda$ be a finite dimensional $k$ algebra. Then the following are equivalent:
	\begin{itemize}
		\item $\Db(\mathbf{mod}\Lambda)$ has a Serre functor
		\item $\Db(\mathbf{mod}\Lambda)$ has Auslander-Reiten triangles
		\item $\gldim \Lambda<\infty$.
	\end{itemize}
\end{thm}
\subsection{Gorenstein Algebras}
\begin{defn}
	Let $\Lambda$ be a finite dimensional $k$-algebra. Then $\Lambda$ is (Iwanaga) Gorenstein if $\injdim(\Lambda_\Lambda)$ and $\injdim({_\Lambda}\Lambda)$ are both finite.
\end{defn}
Some examples of such algebras:
\begin{itemize}
	\item If $\gldim \Lambda<\infty$, $\Lambda$ is Gorenstein.
	\item If $\Lambda$ is self-injective, $\Lambda$ is Gorenstein.
	\item If $\Lambda$ and $\Gamma$ are both Gorenstein, then $\Lambda\otimes_k\Gamma$ is Gorenstein. For instance the path algebra of a quiver $kQ$ tensored with $k[\varepsilon]$ the ``dual numbers''.
	The former (usually) has Gorenstein dimension 1 and the latter has Gorenstein dimension 0 (it is self-injective).
\end{itemize}

Now let $\mathbf{Mod}\Lambda\supset \mathbf{mod}\Lambda\supset \mathbf{proj}\Lambda$ be the $\Lambda$ modules, finitely generated $\Lambda$ modules and finitely-generated projectives, respectively.

Let $\mathbf{GProj}\Lambda=\{X\in\mathbf{Mod}\Lambda|\Ext^i(X,\Lambda)=0\text{ for }i>0\}$ and let $\mathbf{Gproj}\Lambda=\mathbf{GProj}\Lambda\cap\mathbf{mod}\Lambda$.

Finally $\uHom_\Lambda(X,Y)=\Hom_\Lambda(X,Y)/P\Hom_\Lambda(X,Y)$ as usual.

Notice that if $\gldim\Lambda$ is finite, then $\mathbf{GProj\Lambda}=\mathbf{Proj}\Lambda$ and if $\Lambda$ is self-injective, 
$\mathbf{GProj}\Lambda=\mathbf{Mod}\Lambda$, giving us (in a way) two extremes.

\begin{lem}
	$\mathbf{GProj}\Lambda$ form a Froebenus category and $\underline{\mathbf{GProj}}\Lambda$ is a compactly generated trianglulated category equivalent to the subcategory of compact objects.
\end{lem}

Buchwertz and Orlov ('87 and '04) did work to show that $\underline{\mathbf{Gproj}}\Lambda$ was equivalent to the singularity category of finitely generated $\Lambda$ modules.

We have a Nakayama functor $\nu:-\otimes_\Lambda D(\Lambda):\mathbf{mod}\Lambda\to \mathbf{mod}\Lambda$ (it takes projectives to injectives).

\subsection{Auslander-Reiten Duality}
We have a diagram 
\begin{center}
	\begin{tikzcd}
		\Db(\mathbf{proj}\Lambda)\ar[r,hookrightarrow]\ar[d] & \Db(\mathbf{mod}\Lambda)\ar[d,"\sim"]\ar[r,two heads,"\nu"] & \Dsing(\lambda)\ar[d,"\bar\nu"]\\
		\Db(\mathbf{proj}\Lambda)\ar[r,hookrightarrow] & \Db(\mathbf{mod}\Lambda)\ar[r,two heads] & \Dsing(\Lambda)
	\end{tikzcd}
\end{center}
\begin{thm}
	$\nu$ is a Serre functor for $\Db(\mathbf{proj}\Lambda)$ and $\Sigma^{-1}\circ\bar\nu$ is aSerre functor for $\Dsing(\Lambda)$.
\end{thm}
\begin{rmk}
	$\calF:\underline{\mathbf{Gproj}}\Lambda\xrightarrow{\sim}\underline{\mathbf{Gproj}}\Lambda$
	is an autoequivalence where $\calF$ sends 
	\[X\mapsto \Omega^{-1}\text{GP}(D\operatorname{Tr}X)\]
	where GP is the Gorenstein projective approximation functio and $D$Tr is the dual of the transpose. This is suppsoed to explain 
	how this is a statement of duality.
\end{rmk}

\subsection{Hochschild cohomology}
$\HH^\ast(\Lambda)=\Ext^\ast_{\Lambda^e)}(\Lambda,\Lambda)$ is a graded commutative ring.
Then $\uHom_\Lambda^\ast(X,Y)=\oplus\uHom_\Lambda(X,\Omega^{-i}Y)$ where $X$ and $Y$ are Gorenstein projectives. Then $\HH^\ast(\Lambda)$ acts on $\underline{\mathbf{GProj}}$ 
via 
\[\varphi_X:\HH^\ast(\Lambda)\xrightarrow{-\otimes_\Lambda X}\Ext^\ast_\Lambda(X,X)\xrightarrow{tx\to px}\uHom^\ast(X,X)\]
where I have no idea what that last map was.

Thus $\uHom^\ast(X,Y)$ are graded $\HH^\ast(\Lambda)$-modules for all $X$ and $Y$ in $\underline{\mathbf{GProj}}\Lambda$. As an assumption, 
fix $R\subseteq\HH^\ast(\Lambda)$ to be a homogeneous $k$ subalgebra such that
\begin{itemize}
	\item $R$ is connected ($R^0=k$)
	\item $R$ is a finitely generated $k$ algebra.
\end{itemize}
Then for $\p$ be a homogeneous prime in $R$ and we get an endofunctor $\Gamma_\p$ on $\underline{\mathbf{GProj}}\Lambda$ that extracts the $\p$-local $\p$-torsion.
\begin{thm}[BIKP, '19]
	For $X,Y\in\mathbf{GProj}\Lambda$ where $X$ is finite dimensional and $\p$ a homogeneous prime with $R/\p$ of Krull dimension $d$,
	\[\Hom_R(\Ext_\Lambda^\ast(X,Y),I(\p))\cong\uHom_\Lambda(Y,\Omega^d\,\Gamma_p \,\text{GP}\,\nu(X))\]
\end{thm}
\begin{prf}
	Passage to closed points via $k\subseteq K$, similar to what we saw in Julia's talk.
\end{prf}

\subsection{Local duality for \texorpdfstring{$\Db(\mathbf{mod}\Lambda)$}{Db(mod Lambda)}}
Let $\calD:=\Db(\mathbf{mod}\Lambda)$ where $\Lambda$ is Gorenstein, $\p$ is homogenous prime in $R\subseteq\HH^\ast(\Lambda)$. Let $\calD_\p$ be the $\p$-localization
of $\calD$, whre the objects are the same and $\Hom_{\calD_\p}^\ast(X,Y)=\Hom^\ast_\calD(X,Y)_\p$.

Then the $\p$-localization functor is exact.

Now let
\[\gamma_\p(\calD):=\{X\in\calD_\p|\End^\ast_{\calD_\p}(X)\text{ is $\p$-torsion}\}\]
which is a thick subcategory of $\calD_\p$.
\begin{rmk}
	if $\calT$ is a compactly generated $R$-linear triangualted category. Then 
	\[\gamma_\p(\calT)\xrightarrow{\sim}(\Gamma_\p\calT)^c\]
	and we get from $\nu$ a local Nakayama endofunctor $\nu_\p$ on $\gamma_\p(\calD)$.
\end{rmk}
\begin{thm}
	$\Sigma^{-d}\nu_\p$ is a Serr functor for $\gamma_\p(\calD)$ and $\Hom_{R_\p}(-,I(\p))$ plays the role of duality.
\end{thm}

For $\m= R^{>0}$ the maximal ideal, $\Db(\mathbf{proj}\Gamma)\subseteq\gamma_\m(\calD)$ with equaility if (FG) holds.

Thus we get Serre duality for $\Db(\mathbf{proj}\Lambda)$ via $\nu_\p$.

\end{document}

### 2019-07-29 MSRI ML Notes/MSRI_ML_notes.tex
Add MSRI notes

--- 2019-07-27 Workshop Notes/workshop_notes.tex
@@
 \pagestyle{fancy}     % set pagestyle for document
-\lhead{Workshop Notes} % put text in header (left side)
+\lhead{``Mathematics of Machine Learning'' Notes} % put text in header (left side)
 \rhead{Nico Courts} % put text in header (right side)
@@
 % Document-Specific Macros
-% Primes and Maximals for the lazy person.
-\newcommand{\p}{\frakp}
-\newcommand{\m}{\frakm}
-\DeclareMathOperator{\Spec}{Spec}
-\DeclareMathOperator{\supp}{supp}
-\newcommand{\SH}{\mathcal{SH}}
-\DeclareMathOperator{\DMack}{DMack}
-\DeclareMathOperator{\Sing}{Sing}
-\DeclareMathOperator{\SpcBal}{Spc_{Bal}}
-\newcommand{\Rep}{\mathbf{Rep}\,}
-
-%stable module category stuff
-\DeclareMathOperator{\uHom}{\underline{Hom}}
-\DeclareMathOperator{\uExt}{\underline{Ext}}
-\DeclareMathOperator{\uEnd}{\underline{End}}
-
-%group schemes
-\newcommand{\Ga}{\bbG_a}
 
@@
 %make the title page
-\title{Workshop Notes\vspace{-1ex}}
+\title{MSRI's \textit{Mathematics of Machine Learning} Summer School\vspace{-1ex}}
 \author{Nico Courts}
-\date{June 24-28, 2019}
+\date{July 29-Aug 9, 2019}
 \maketitle
@@
 \begin{abstract}
-	These notes were the ones I took while attending the ``Triangulated Categories in Geometry and Representation Theory'' Workshop 
-	at the University of Sydney during the week beginning June 23, 2019.
+	These notes were the ones I took while attending MSRI's ``Mathematics of Machine Learning'' summer school
+	at the University of Washington during the two weeks above.
 \end{abstract}
 
-\section{The spectrum of the category of derived Mackey functors}
-This talk was given by Beren Sanders on a joint work alongside Irakli Patchkoria and Christian Wimmer.
-
-First we are going to talk about tensor triangulated geometry. Essentially we are looking at 
-an essentially small tensor triangulated category and then we want to talk about the ``geometry''.
-Analogous to the way we define the spectrum of a ring, we can define 
-\begin{defn}
-	The \textbf{spectrum} of a tensor-triangulated category $\calK$ is 
-	\[\operatorname{Spec}{\calK}=\{\p\subsetneq \calK|\p\text{is a prime tensor triangulated ideal}\}\]
-\end{defn}
-
-Then if $X\in\calK$ is an object, then we can define the support of $X$ to be 
-\[\operatorname{supp}(X)=\{\p\in\operatorname{Spec}(\calK)|x\notin\p\}\]
-
-We define the closure of a prime to be the set of primes that are contained in it. There are some great universal properties.
-
-\begin{lem}
-	\begin{itemize}
-		\item $\Spec(D(T)^c)\cong\Spec(R)$
-		\item $\Spec(\Db(kG-\mathbf{mod}))\cong \Spec(H^\ast(G,k))$.
-	\end{itemize}
-	The second result is was generalized by (among others) Julia to group schemes.	
-\end{lem}
-
-\subsection{The Stable Homotopy Category}
-This is the context for most of the talk for today. Let $\SH$ be the stable homotopy category.
-There are a couple sources for objects here. We can consider $\sigma^\infty X$, the infinite suspension of a pointed space, 
-or a cohomology theory $\bbE$ (e.g. singular cohomology).
-
-$\SH$ is a tensor-triangulated category although I missed the tensor structure.
-
-\begin{thm}[Balmer `10]
-	There is a \textbf{comparison map}, a continuous map
-	\[\rho_\calK:\Spec(\calK)\to \Spec(\End_\calK(1))\]
-	where $1$ is the monoidal unit (since I haven't set up the macro for the bb font.)
-\end{thm}
-The idea here is that the object on the right is more easy to understand in general.
-
-So for the example of $\SH^c$, we can compute that the endomorphism ring is $\bbZ$. 
-Then if we look at any prime $\p\in\Spec(\bbZ)$, the fiber contains all the extraordinary cohomology theories 
-that correspond to computing cohomology over characteristic $p$ fields.
-
-\subsection{The spectrum of \texorpdfstring{$\SH(G)$}{SH(G)}}
-If $G$ is a finite group, then $\SH(G)$ is the $G$-equivariant stable homotopy group.
-In this case, the endomorphism ring is known as the Burnside ring $A(G)$.
-
-Dress in `69 computed the spectrum of $A(G)$: For each $H\le G$, we ahve a ring homomorphism 
-\[f^H:A(G)\to \bbZ\]
-where $[X]$ is sent to $|X^H|.$ These induce embeddings:
-\[(f^H)^\ast:\Spec(\bbZ)\to\Spec(A(G))\]
-where $(p)\mapsto \p(H,p)$ and $(0)\mapsto \p(H,0)$.
-
-\begin{defn}
-	Let $O^p(G)$ be the smallest normal subgroup of $G$ such that $G/O^p(G)$ is a $p$-group.
-\end{defn}
-\begin{thm}[Dress '69]
-	$\p(H,p)=\p(K,p)$ if and only if $O^p(H)\sim_G O^p(K).$
-\end{thm}
-\begin{rmk}
-	For example if $p\nmid |G|$ then $O^p(H)=H$ for all $H\le G$. If $G$ is a$p$ group, then $O^p(H)=1$ for all $H$.
-\end{rmk}
-
-\subsection{Categorifying Dress' Work}
-FOr each $H\le G$, we have the \textbf{geometric $H$-fixed points functor}
-\[\Phi^H:\SH(G)^c\to \SH^c\]
-that induce embeddings 
-\[(\Phi^H)^\ast:\Spec(\SH^c)\hookrightarrow \Spec(\SH(G)^c)\]
-where $\calC_{p,n}\mapsto \calP(H,p,n)$. Furthermore the copies of $\Spec(\SH^c)$ cover $\Spec(\SH(G)^c)$. T
-\textbf{They do not overlap.}
-
-The speaker, together with Paul Blamer, computed the spectrum of $\SH(G)^c$. The picture is interesting, but there is a topological connecting
-them: In particular when $G=C_p$, $\calP(1,p,n)\not\subseteq \calP(G,p,n)$ but $\calP(1,p,n+1)$ is.
-
-\begin{defn}
-	For a finite $p$-group $G$, define the ``blue shift'' to be 
-	\[\beta(G,n)\]
-	which is the smallest $i$ such that $\calP(1,p,n+i)\subseteq\calP(G,p,n)$
-\end{defn}
-It was shown that for $p$ groups, $\beta(G,n)$ is the free rank of $G$.
-\begin{rmk}
-	For instance, $\beta((C_p)^r,n)=r$ and $\beta(C_{p^r},n)=1$
-\end{rmk}
-
-The bottom line here is that the spectrum of the $G$-equivariant can be understood ``locally''
-as $\Spec(A(G))$.
-
-\subsection{``Linearization'' of stable homotopy theories}
-Some examples of what we're talking about: We can in some way think of the linearization of $\SH$ as being $\D(\bbZ)$.
-Furthermore, the motivic derived category of $k$ could be the linearization of $\SH^{\bbA^1}(k)$.
-
-In our case, we may want to consider the derived category of Mackey functors on $G$ as the linearization of $\SH(G)$. Mackey functors 
-are functors that behave well with respect to the group/representation theory (e.g. induction and restriction).
-
-One may think this category is the derived category of the abelian category of $G$-Mackey functors, but 
-Kaledin (2008) argues that it should be this other category $\DMack(G)$. The speaker computed the spectrum of this 
-category as well as developing a different construction.
-
-One interpretation of the linearization in the earlier cases is that $\SH$ is the derived category of a circle 
-and $D(\bbZ)$ is the derived category of the Eilenberg-Maclane spectrum $H\bbZ$. For our example, we can define 
-$H\bbZ_G$ to be $H\bbZ$ considered as a trivial $G$-module in $\SH(G)$. Then we say 
-\[\D(H\bbZ_G)=\DMack_{Kal}(G)\]
-and compute the spectrum.
-
-\begin{thm}
-	The extension of scalars functor 
-	\[\SH(G)\to D(H\bbZ_G)\]
-	induces an embedding 
-	\[\Spec(D(H\bbZ_G))\hookrightarrow \Spec(\SH(G))\]
-	into the top and bottom layers. 
-\end{thm}
-
-The main takeaway her is that for $C_p$ the spectrum of the derived category of Mackey functors lies between that 
-of $A(C_p)$ and of $\Spec(\SH(C_p))$ where the first step splits points into two and the second introduces this chromatic splitting.
-
-\section{Supports and Cosupports in the Stable Module Category}
-This series of talks was given by Julia Pevtsova.
-
-\subsection{Tensor triangulated geometry for finite group schemes}
-Let $A$ be a finite dimensional Hopf algebra and associate to $A$ the category $\Sing(A)$, the singularity category of $A$ which measures how non-semisimple
-$A$ is. One of the reasons we ways to approach tensor-triangulated geometry is to develop what points are. As a disclaimer, Julia 
-says that there might be some misunderstandings she has here, but that it has been developed by Balmer and Krause.
-
-For some motivation, let $R$ be a commutative ring, $\p\in\Spec R$ and let $M$ be an $R$-module. We can gain local information 
-by looking at $M_\p$ or $M\otimes_R k_p$.
-
-Now ler $\calT=\D(R)$ where $\calT^c=\Dbperf(R)$. Take $Y\in\calT$ and $\p\in\Spec(R)$ and we can define $Y_\p$ or 
-$Y\otimes_R^Lk_p$. Then we define 
-\[\supp Y=\{\p\in\Spec R|Y\otimes_R^Lk_\p\ne 0\}.\]
-
-This has some nice properties:
-\begin{itemize}
-	\item (Defection Property) $Y\simeq 0\Leftrightarrow \supp Y=\varnothing$.
-	\item Invariant under $\otimes, \oplus,$ cones, and syzygies.
-\end{itemize}
-\begin{thm}[Neeman '92]
-	The class of localizing subcategories in $\D(R)$ (here $R$ is Noetherian) are in one to one correspondence via $\supp$ to the subsets in $\Spec R$.
-
-	Furthermore restricting to $\calT^c$, the thick subcategories in $\Dbperf(R)$ correspond to specialization closed subsets in $\Spec R$.
-\end{thm}
-
-\subsection{General Definitions and \texorpdfstring{$\SpcBal(\calT^c)$}{SpcBal(Tc)}}
-\begin{defn}
-	Let $T$ be a tensor triangulated category. We say that $\calC\subseteq\calT$ is \textbf{localizing} subcategory if $\calC$ is a full triangulated 
-	subcategory that is closed under $\oplus$.
-
-	$\calC\subseteq \calT^c$ is \textbf{thick} if it satisfies the same conditions.
-
-	$\calC\subseteq\calT$ is a tensor ideal if it absorbs tensor products.
-\end{defn}
-
-\begin{rmk}
-	In this talk $\calT$ is symmetric but this works in braided categories.
-\end{rmk}
-\begin{defn}
-	The \textbf{Balmer Spectrum of $\calT^c$} is defined via a map $T^c\mapsto \SpcBal(T^c)$ where we send 
-	$M\mapsto\supp_{Bal}M$.
-\end{defn}
-
-Now knowing $\SpcBal(T^c)$ is equivalent ot classifying thick tensor ideals in $T^c$. When $T^c=\Dbperf(R)$, 
-and $\p\in\Spec R$, the map $R\to k_\p$ induces maps 
-\[\Dbperf(R)\to \Dbperf(k_\p)=\D\mathbf{Vect}_{k_\p}\]
-and thus 
-\[\SpcBal \D\mathbf{Vect}_{k_\p}\to \SpcBal\Dbperf(R).\]
-
-\subsection{Doing this in modular representation theory}
-Classically we are doing modular representation theory when we have a finite group $G$ and a field $k$ where $\ch k| |G|$.
-In essence this is just saying that our representation theory is with non-semisimple rings.
-
-We can do this theory on a series of objects:
-\begin{itemize}
-	\item Finite groups (as above)
-	\item Finite group schemes
-	\item Finite supergroup schemes
-	\item Small quantum groups
-	\item Algebraic group $G$ and algebra $A$ over $G$. Then considering $\operatorname{Bimon}(G,A)$ is one possibility.
-	\item Finite dimension pointed Hopf algebras
-	\item Lie superalgebras (characteristic zero).
-	\item One can continue the list. The idea here is that you have a big enough endomorphism ring.
-\end{itemize}
-
-Today we are focusing on finite group schemes (FGS) although we might get to say something about supergroup schemes and quantum groups.
-
-\subsection{Affine Group Schemes}
-\begin{defn}
-	An \textbf{affine groups scheme $G$ over $k$} is a representable functor $G$ from commutative $k$-algebras to groups.
-\end{defn}
-
-\begin{rmk}
-	$G(R)\cong \Hom_{\Algk}(k[G],R)$ is how the representation works. $k[G]$ is a commutative Hopf algebra. $G$ is finite if $\dim_kk[G]<\infty$.
-\end{rmk}
-
-\begin{defn}
-	\[kG:=k[G]^\vee=\Hom_k(k[G],k)\]
-	is the \textbf{group algebra} corresponding to $G$. This is a finite dimension cocommutative Hopf algebra.
-\end{defn}
-
-\begin{rmk}
-	There is an equivalence of categories between finite group schemes and finite dimensional cocommutative Hopf algebras using $G\mapsto kG$. 
-	Furthermore, the category $\mathbf{Rep}_kG$ is equivalent to $kG$-$\mathbf{Mod}$ (here uppercase will denote non-finite-dimensional), which we will call $G$-$\mathbf{Mod}$.
-\end{rmk}
-
-Then we can define the cohomology for $G$ via the Hochschild cohomology
-\[\Ext_G^\ast(k,k)=H^\ast(G,k)=H^\ast(kG,k)\]
-
-Some examples of these are finite groups, restricted Lie algebras (Lie algebra over fields of positive characteristic that also have a $p^{th}$ power operation). 
-To construct a restricted Lie algebra, One cal let $\calG$ be some algebraic group. Then $\operatorname{Lie}\calG=\frakg$ is a restricted Lie algebra.
-The representation theory of such a $\frakg$ is precisely the representation theory of the (finite-dimensional, cocommutative) Hopf algebra $u(\frakg)$,
-the quotient of the universal enveloping algebra that identifies $p^{th}$ powers (in the PBW basis) with the internal $p^{th}$ powers in $\frakg$.
-
-More generally one can consider \textbf{Frobenius kernels}. Let $F:\calG\to\calG$ be the Frobenius map. Then 
-$\calG_{(1)}=\ker F$ and more generally $\calG_{(r)}=\ker F^(r)$, the $r^{th}$ iteration of $F$. These are connected
-finite group schemes. These things have only a single point! However the group algebra and coordinate algebra really still capture 
-all the representation theory of the original group.
-
-If $\calG$ is an algebraic group, then we can compute $\calG_{(1)}$ and $\operatorname{Lie}\calG$ and 
-\[k\calG_{(1)}\cong u(\operatorname{Lie}(\calG)).\]
-
-When $\calG=\Ga$, then $k[\Ga]=k[T]$ where $\Delta T)=T\otimes 1+1\otimes T$ and 
-\[{\Ga}_{(r)}(R)=\{a\in R|a^{p^r}=0\}.\]
-Now $k[{\Ga}_{(r)}]=k[T]/(T^{p^r})$ and $k{\Ga}_{(r)}=k[u_0\cdots,u_{r-1}]/(u_0^p,\dots,u_{r-1}^p)$
-
-Recall that $\StMod G$ is the category of $G$ representations where homs are considered module factoring through projective objects. This becomes a tensor triangulated category 
-where the triangles comes from short exact sequences in $\Rep G$ and tensor is the usual one. The shift functor is the syzygy functor $\Omega^{-1}$ and the monoidal unit is $k$.
-
-\begin{rmk}
-	We have an exact sequence 
-	\[\D(G)\to \K(\mathbf{Inj}(G))\to\StMod G\]
-	and 
-	\[\Dbperf(G)\to \Db(G)\to\mathbf{stmod} G\cong\Dsing(G).\]
-	so in a way this measures the lack of regularity in the sense of commutative algebra.
-\end{rmk}
-
-Now here we can talk about $\uEnd^\ast(k)$ as well as 
-\[\uHom^\ast(M,N)=\oplus_n\uHom(M,\Omega^{-n}N)\]
-and 
-\[\uEnd^\ast(k)\cong\widehat\Ext_G^\ast(k,k)=\hat H^\ast(G,k)\]
-where the hat denotes Tate cohomology.
-
-Now $\uEnd^\ast(k)$ acts on $\calT$ (on $\calT^c$) by acting on $\uHom^\ast(M,N)$.
-
-THis leads to to Benson-Iyengar-Krause support theory where 
-\[R=H^\ast(G,k)\to\hat H^\ast(G,k)\]
-acts on $\calT$.
-
-We can try setting $X=\Spec R=\Spec H^\ast(G,k)$ for the space where the support lives.
-
-\subsection{\texorpdfstring{$\pi$}{pi}-points and local cohomological functors}
-Recall last time we lookedd at $R=H^\ast(G,k)=\Ext_G^\ast(k,k)$ acting (via $\uEnd^\ast(k)$ on $\StMod G$, where $G$ was a finite group scheme.
-
-For examples we considered $\frakg=\operatorname{Lie}\frakG$ where $\frakG$ is a reducitive algebraic group. Here 
-$H^{odd}(\frakg,k)=0$ and $H^{ev}(\frakg,k)\simeq k[\calN]$ where $\calN$ are the set of nilpotents in $\frakg.$
-
-\begin{thm}
-	When $\frakg$ is a restricted Lie algebra,
-	\[\Spec H^\ast(\frakg,k)=\calN^{[p]}=\{x\in\frakg|x^{[p]}=0\}.\]
-\end{thm}
-\begin{thm}[Suslin-Friedlander-Bendel '97]
-	\[\Spec H^\ast(GL_{n(r)},k)\]
-	is the variety of $r$ duples of $p$-nilpotent commuting matrices.
-\end{thm}
-
-Some facts about cohomology:
-\begin{itemize}
-	\item $H^\ast(G,k)$ is a graded commutative algebra.
-	\item Friedlander and Suslin in '95 proved $H^\ast(G,k)$ is finite generated (has FG).
-\end{itemize}
-
-The idea to develop the support theory is to look in $X=\operatorname{Proj}(R)$ and then proceed via either $\pi$-points or via local cohomological functors (developed by Bens0n, Iyengar, and Krause).
-
-\begin{ex}
-	Another motivational example: let $\frakg$ be a restricted Lie algebra. Define $X=\operatorname{Proj}H^\ast(\frakg,k)\simeq\calN^{[p]}$
-\end{ex}
-
-\begin{rmk}
-	$\frakg_a=\operatorname{Lie}\Ga$ is an algebra that is very small since $u(\frakg_a)=k[x]/x^p$. Now $\StMod k[x]/x^p$ is called a ``tt-field'', and for any $x\in\calN^{[p]}\subseteq\frakg$,
-	setting $\frakg_a=\langle x\rangle$ gives us a functor 
-	\[\calF_x:\StMod\frakg\to\StMod\frakg_a\]
-	induced by inclusion.
-
-	Now $\calF_x$ is a tensor triangulated functor and this functor exists for all $x$ in the nilcone (or its image in the projective closure $X$).
-\end{rmk}
-\begin{defn}
-	Let $M\in\StMod G$. Then the support is 
-	\[\supp M=\{x\in\calN^{[p]}|M|_{\langle x\rangle}=\calF_x(M)\neq 0\}\subseteq X=\operatorname{Proj}\calN^{[p]}\]
-\end{defn}
-
-The support satisfies all the good properties. (?)
-
-if $G$ is a finite group scheme then as we have seen $kG$ is a finite dimensional cocommutative Hopf algebra.
-\begin{defn}
-	If $K/k$ is an extension, then 
-	\[\alpha:K[t]/t^p\to KG=kG\otimes_k K\]
-	is a $\pi$-point if it's flat and there is $\calU\subseteq G_K$ such that $\alpha$ factors through $K\calU$ where $\calU$ is 
-	a unipotent abelian subgroup scheme of $G_K$.
-\end{defn}
-
-Given such a map $\alpha:K[t]/t^p\to KG$, we get a functor 
-\[\calF_\alpha:\StMod G\to\StMod K[t]/t^p\]
-where the codomain is a tt-field.
-\begin{rmk}
-	$\calF_\alpha$ is not a tensor functor, but 
-	\[\calF_\alpha(M\otimes N)\cong 0\Leftrightarrow \calF_\alpha(M)\otimes\calF_\alpha(N)\cong 0\]
-	and either of these happen if and only if $\calF_\alpha$ is zero on $M$ or $N$.
-\end{rmk}
-\begin{rmk}
-	Note that this is a place where things break down for quantum groups. We no longer have factoring through an abelian subgroup.
-\end{rmk}
-
-Now return again to $X=\operatorname{Proj}H^\ast(G,k)$. Then given a $\pi$-point $\alpha$, we get maps 
-\[H^\ast(G,k)\xrightarrow{-\otimes_R K}H^\ast(G,K)\xrightarrow{\alpha^\ast}H^\ast(K[t]/t^p,K)\cong K[x]\otimes\bigwedge(\lambda)\]
-and this map gives us $P_\alpha$, which is the radical of the kernel of this map. This gives us a map $\alpha\mapsto P_\alpha\in X$.
-
-The nice property here is that $\pi$-points realize all points on $X$:
-\begin{thm}
-	For all $\p\in X$, there exists an extension $K/k$ and $\pi$-point $\alpha$ such that $P_\alpha=\p$.
-\end{thm}
-
-\begin{defn}
-	Let $M\in\StMod G$. Then the $\pi$ support is 
-	\[\pi-\supp M=\{\p\in X|\calF_{\alpha_\p}(M)=\alpha^\ast_\p(M\otimes_k K)\ne 0\}\]
-	notice that the construction here doesn't depent on choice of $\pi$ point, but that is something to show.
-\end{defn}
-\begin{defn}
-	\[\pi-\operatorname{cosupp}(M)=\{\p\in X|\calF^{\alpha_\p}:=\alpha_\p^\ast(\Hom_k(K,M))\ne 0\}\]
-	where this is computed ve the corestriction functor (this needs clarification).
-\end{defn}
-\begin{thm}
-	$\pi$-$\supp(M\otimes_k N)=\pi$-$\supp(M)\cap\pi$-$\supp(N)$.
-\end{thm}
-\begin{thm}[BKIP `18]
-	$\pi$-$\supp(M)=\varnothing$ if and only if $M=0$.
-\end{thm}
-
-\subsection{Local Cohomology Functors}
-Let's take a moment to talk about the classical definition of support. When $R=H^\ast(G,k)=\Ext^\ast_G(k,k)$, we can act on 
-$\uHom^\ast(M,M)$. When $M\in\stmod G$, we let $I_M=\operatorname{Ann}_R\uHom^\ast(M,M)$. Then $V(I_M)\subseteq X$ is the \textbf{support of 
-$M$,} denoted $\supp M$.
-
-To generalize this, we are going to let $\calT=\StMod G$ and associate to each $\p\in X$ a functor 
-\[\Gamma_\p:\calT\to\calT\]
-which selects the $\p$-local, $\p$-torsion objects.
-\begin{defn}
-	If $M\in\StMod G$, we can define the \textbf{localization map} $M\to M_\p$ where 
-	\[\uHom^\ast(C,M_\p)=\uHom^\ast(C,M)_\p\]
-	for each $C\in\calT^c$. (recall $\calT$ is $R$ linear so hom can be localized regularly.)
-\end{defn}
-\begin{defn}
-	There exists a functorial map 
-	\[\Delta\Gamma_{V(\p)}M\to M\to L_{V(\p)}M\]
-	where the left part is universal $\p$-torsion and the right (I believe) is $\p$-torsion in the usual way (check the paper).
-\end{defn}
-
-Then $\Gamma_\p$ is found by taking the $\p$-local and $\p$-torsion operations in either order.
-
-Now 
-\[\Gamma_\p M:=\Gamma_p k\otimes_k M\]
-and so
-\begin{defn}
-	$\supp_{BIK}M=\{p\in X:\Gamma_\p k\otimes M\ne 0\}$ and $\operatorname{cosupp}_{BIK}=\{\p\in X|\Hom_k(\Gamma_\p k,M)\ne 0\}$
-\end{defn}
-
-And then zooming out, we define 
-\[\Gamma_\p(\calT)=\Gamma_\p(\StMod G)=\{M\in\calT|\supp_{BIK}(M)\subseteq\p\}\]
-
-\begin{prop}
-	\begin{itemize}
-		\item $\pi$-$\supp(\Gamma_\p k)=\p$
-		\item $\pi$-$\supp=\supp_{BIK}$ (formal from detection, tensor product)
-	\end{itemize}
-\end{prop}
-\begin{cor}[Balmer, Friedlander-P]
-	$\SpcBal(\stmod G)\cong\operatorname{Proj}H^\ast(G,k)$ (even on the level of schemes).
-\end{cor}
-
-\subsection{Localizing ideals in \texorpdfstring{$\StMod G$}{StMod G}}
-Here we use the BIK local-global principle: to classify localizing $\otimes$ ideals, we need to know that $\Gamma_\p\calT$ are minimal. The ``wishful thinking''
-here (which we'll flush out next time) is as follows: if we know $\pi$-$\operatorname{cosupp}=\operatorname{cosupp}_{BIK}$,
-then we can use ``Neeman's minimality lemma'' which says that $\calC\subseteq\calT$ is minimal if for all $M$ and $N$ in $\calC$,
-$\Hom^\ast_k(M,N)\ne 0$.
-
-Next time we will see some more and a proof!
-
-\subsection{Day 3}
-Recall the context: we were letting $X=\operatorname{Proj}H^\ast(G,k)$ and $\calT=\StMod G$ and then looking at
-\[\Gamma_p=\{M\in\calT|\supp M\subseteq \p\}\]
-
-\begin{lem}
-	$\m\in X$ is a closed point. Then $\Gamma_\m \calT$ is minimal.
-\end{lem}
-\begin{prf}
-	We begin with a fact: $\m\in\pi$-$\supp M$ if and only if $\m\in \pi$-$\operatorname{cosupp} M$.
-
-	To show that $\Gamma_\m T$ is minimal, it suffices to show that $\Hom_k(M.N)\ne 0$ for all $M$ and $N$. But 
-	\[\pi\text{-}\operatorname{cosupp}\Hom_k(M,N)=\pi\text{-}\supp M\cap \pi\text{-}\operatorname{cosupp}N\]
-	But since the $\pi$ support of $M$ is $\m$ and since $\m$ lies in the $\pi$-cosupport of $N$, then the intersection is nonempty.
-	Thus since the cosupport is nontrivial, the module itself is nontrivial.
-\end{prf}
-
-\subsection{Reduction to closed points}
-We want a Koszul object (in rep theory a Carlsson object): for every $b\in H^d(G,k)\cong\uHom^d(k,\Omega^{-d}k)$, we get an object
-$k//b$ formed by the cone of $b:k\to\Omega^{-d}k$. Then for any sequence $\underline{b}=(b_1,\dots,b_n)$, we set 
-\[k//\underline{b}=k//b_1\otimes\cdots\otimes k//b_n.\]
-
-Now if we have any finite field extension $K/k$ (the idea here is we would like to take $k$ itself, but ), we can choose a $\m\in X_K=\operatorname{Proj}H^\ast(G,K)$ lying over $\p\in H^\ast(G,k)$.
-Then by taking the additional generators for $\m$, we can pick a $\underline b$ such that $\m=\sqrt{\tilde\p+\underline b}$ where $\tilde\p$ is the lifting of $\p$ in the bigger ring.
-
-Then we can consider the (Sylow?) functor for restriction of scalars $\operatorname{Res}^{G_K}_G:\StMod_K G\to \StMod G$
-and look at the image of $\Gamma_\m(K//\underline b)$.
-\begin{lem}
-	$\Gamma_\m(K//\underline b)\downarrow_G\simeq \Gamma_\p k$
-\end{lem}
-\begin{cor}
-	the image $\operatorname{Res}_G^{G_K}$ of $\Gamma_\m\StMod_K G$ in $\Gamma_\p\StMod_k G$ is dense.
-\end{cor}
-\begin{cor}
-	$\Gamma_\p$ is minimal for all $\p$.
-\end{cor}
-
-\subsection{Support theory}
-The following is tied closely to the following results:
-\begin{prop}
-	$\pi$-$\supp M=\varnothing$ if and only if $M\cong 0$.
-
-	That is, the $\pi$ support has sufficient strength to detect things (``detection'').
-\end{prop}
-\begin{prop}
-	For all $\p\in X$ there is an $\alpha$ with $\calF_\alpha$ realizing $\p$.
-\end{prop}
-
-I am unsure about how this fits in, but she also mentioned the detection of nilpotents in $H^\ast(G,k)$.
-\begin{thm}[Quillen '71]
-	If $G$ is a finite group, $g\in H^\ast(G,k)$ is nilpotent if and only if for all elements in an abelian $p$-group $E\subseteq G$,$g\downarrow_E$ is nilpotent.
-\end{thm}
-
-This means $H^\ast(E,k)\cong k[x_1,\dots,x_n]\otimes \bigwedge^\ast(\tau_1,\dots,\tau_n)$
-\begin{thm}[Chouinard '76]
-	if $G$ is a finite group, then $M$ is projective if any only if for all elementary abelian $E\le G$.
-\end{thm}
-
-Though these results we get that Jarrod's work on the lattice of elementary abelian subgroups of specific groups correspond directly to the lattice (modulo conjugacy and some finite 
-group actions) of affine subspaces in spec of the ring.
-
-Now if $\frakg$ is a Lie algebra (say $\frakg=\operatorname{Lie}\calG$ where $\calG$ is a reductive group), 
-$H^\ast(\frakg,k)=k[\calN]$. Then fix some Borel subgroup $\calB\subseteq \calG$. Take any $\lambda\in \calG/\calB$ (which is apparently a flag variety?)
-and extract from this a subalgebra $b_\lambda\subseteq\frakg$. Then extract a functor 
-\[\calF_\lambda:\StMod\frakg\to\StMod b_\lambda\]
-\begin{thm}
-	Then if $M\in\StMod\frakg$, $M\cong 0$ if and only if $\calF_\lambda(M)=0$ for all $\lambda.$
-\end{thm}
-Let $G=\calG_{(r)}.$
-\begin{defn}
-	A \textbf{one-parameter subroup} is
-	\[\bbG_{a_(r)}\to\calG_{(r)}\]
-	where $k\bbG_{a_{(r)}}\cong k[u_0,\dots,u_{r-1}]/(u_0^p,\dots, u_{r-1}^p)$.
-\end{defn}
-\begin{thm}
-	Nilpotence and projectivity are detected on one-parameter subgroups (up to scalar extension).
-\end{thm}
-Here we define the $\pi$-point:
-\[K[t]/t^p\xrightarrow{t\mapsto u_{r-1}} k\bbG_{a_{(r)}}\to K\calG\]
-
-\brk
-
-But then we get an isogeny 
-\[\operatorname{Mor}(\bbG_{a_{(r)}},G)\simeq\operatorname{Spec}H^\ast(G,k)\]
-
-\begin{defn}
-	A finite group scheme is elementary if 
-	\[\calE=\bbG_{a_{(r)}}\times(\bbZ/p)^s\]
-\end{defn}
-\begin{thm}[Suslin '76]
-	$\xi\in H^\ast(G,k)$ is nilpotent if any only if for all $K/k$ and for any elementary abelian subgroup $\calE\subseteq G_K$,
-	the restriction $\xi_K\downarrow_\calE$ is nilpotent.
-\end{thm}
-
-\section{Smoothness and Properness}
-Greg Stevenson gave this series of lectures.
-
-\subsection{Beginning definitions}
-
-Today we will be working up to dg categories. In what follows let $k$ be a fixed base field (although it works without being a field, this 
-is a nice assumption). Recall the definition of a chain complex $\cdots\to X^i\xrightarrow{d^i} X^{i+1}\to\cdots$.
-
-Recall that a chain morphism is a sequence of maps indexed by $\bbZ$ such that the obvious diagrams commute. Denote by $\Ch(k)$ the category of chain complexes of $k$
-vector spaces. We have the \textbf{suspension} or \textbf{shift functor} in this category, denoted $\Sigma$. This is an autoequivalence of $\Ch(k)$.
-
-Given a chain map $f:X\to Y$, we define $H^i(X)=\ker d^i/\Im d^{i-1}$. This gives us a functor $H^i:\Ch(k)\to \Vectk$. We say that $f$ is a quasi-isomorphisms if it induces
-isomorphisms on cohomology.
-
-We can form the mapping cone of $f$, $\cone f$ (think this is kind of a kernel and kind of a cokernel)
-\[\cone(f)=(\Sigma X\oplus U,(\begin{smallmatrix}d_{\Sigma X} & 0\\ f & d_Y\end{smallmatrix}))\]
-
-Recall the definition of acyclic objects. Then 
-\begin{lem}
-	$f:X\to Y$ is a quasi-isomorphism if and only if $\cone(f)$ is acyclic.
-\end{lem}
-\begin{rmk}
-	This supports the idea that this is in fact a kernel and cokernel in a way.
-\end{rmk}
-
-Begin $\Ch(k)$ admits a symmetric monoidal structure from the total complex of the associated double 
-complex (with some signs flipped). This defines a functor $-\otimes -$ into $\Ch(k)$.
-
-We also have internal hom (aka the $\hom$ complex). Here 
-\[\hom(X,Y)^n=\prod_{i\in\bbZ}\Hom_k(X^i,Y^{i+n})\]
-with differential given by the graded commutator:
-\[d(f)=d_Y^{i+n} f^i-(-1)^nf^{i+1}d_X^i\]
-
-\begin{rmk}
-	This is measuring, in a graded way, 
-\end{rmk}
-\begin{ex}
-	$Z^0\hom(X,Y)=\ker d^0_{\hom(X,Y)}=\{\text{degree zero maps} f| fd=df\}$. Then we can compute that 
-	\[H^0\hom(X,Y)\]
-	which is precisely the homotopy classes of maps from $X$ to $Y$.
-\end{ex}
-
-\brk
-
-We have a tensor hom adjunction (from here let $\Ch=\Ch(k)$)
-\[\Ch(X\otimes Y,Z)\cong \Ch(X,\hom(Y,Z))\]
-for all $X,Y,Z$ so we have the counit in the form of coevaluation
-\[\varepsilon:X\otimes\hom(X,Y)\to Y\]
-and given a $Z\in\Ch$, we can consider 
-\[X\otimes\hom(X,Y)\otimes\hom(Y,Z)\xrightarrow{\varepsilon_Y\otimes 1}Y\otimes \hom(Y,Z)\xrightarrow{\varepsilon_Z} Z\]
-and so via the adjunction we get a composition map 
-\[\hom(X,Y)\otimes\hom(Y,Z)\xrightarrow{\circ}\hom(X,Z).\]
-
-This composition map is associative. We also get unit maps $1_X:k\to\hom(X,X)$ determined by the image of $\id_X\in\Ch(X,X)$ in $\hom(X,X)$.
-These maps act as units for $\circ$, using the fact that $k$ is the unit of $\otimes_k$.
-
-So using $1_X$ and $\circ$, we get something that looks like a category! This is the primordial example of a dg-category.
-
-\subsection{Examples and extensions}
-Give $X\in\Ch$, we get a monoid $\hom(X,X)\in\Ch$. What does that mean? Well, we have a map $\hom(X,X)\otimes\hom(X,X)\to\hom(X,X)$
-which is just composition, and a unit $1_X:k\to\hom(X,X)$. This is an example of a differential graded algebra. I.e. we have an algebra $A$ equipped with 
-a degree 1 differential $d$ which is a graded derivation.
-This is also an example of a dg-category.
-
-\subsection{dg-categories}
-\begin{defn}
-	$\calA$ is a dg-category if it is given by a collection of objects along with a complex 
-	$\calA(a,b)\in\Ch$ for each $a,b\in\calA$ as well as unit maps 
-	\[1_a:k\to\calA(a,a)\]
-	and composition maps 
-	\[\circ:A(b,c)\otimes A(a,b)\to A(a,c)\]
-	such that composition is associative and unital with respect to the unit maps.
-\end{defn}
-\begin{rmk}
-	The slick way of saying this is that a dg-category is a category enriched over chain complexes. :)
-\end{rmk}
-
-The $\hom$ complexes make $\Ch$ into a dg-category. I will call this $\scrC$. If $A$ is a dga, consider the dg-cat $BA$, 
-which is the category of a single object with hom set $A$.
-
-If $B$ is any $k$-linear category we can view it as a dg category $\tilde B$ with $\tilde B(b,c)=B(b,c)$ in degree zero.
-
-Given a dg category $\calA$, we can do a couple things to get a boring category: first, you can take the underlying category $Z^0\calA$ that has the same objects 
-and the maps are the zeroth cocycles of $A(a,b)$ for each $a$ and $b$. For example $Z^0\scrC=\Ch$.
-
-We can also compute the homotopy category of $\calA$, denoted $H^0\calA$ which again has teh same objects, but 
-\[H^0\calA(a,b)=H^0(\calA(a,b)).\]
-Now $H^0\scrC=K(k)$, the homotopy category of chain complexes.
-
-\begin{ex}
-	If $R$ is a $k$ algebra, then we get a dg category over $k$ $\scrC(R)$ with objects chain complexes of $R$-modules 
-	and where the hom complexes are $\hom_R(X,Y)$.
-\end{ex}
-
-\subsection{dg modules and derived categories}
-Given a pair of dg categories $\calA$ and $\calB$, a dg functor $F:\calA\to \calB$ consists of an object assignment
-along with a map of $\hom$s that preserve units and composition. There is also a notion of being a dg-natural transformation 
-and dg-equivalence (this is the part of this we will need), and so on.
-
-Let's finish off today with dg modules. 
-\begin{defn}
-	Given a small dg-category $\calA$ (notice they are always locally small!), a \textbf{right $A$-dg-module} $X$ is 
-	a dg functor $\calA^{op}\to\scrC$.
-
-	Equivalently, for all $a\in\scrA$, $X(a)\in\Ch$ and for every $a,b\in\calA$ a map 
-	\[\calA(a,b)\to\scrC(Xa,Xb)\]
-	which is equivalent via the tensor hom adjunction to a map 
-	\[\calA(b,a)\otimes Xa\to X_b\]
-	which looks a lot more like a module action. :)
-\end{defn}
-Let $\mathbf{Mod}$-$\calA$ be the category of dg modules and dg natural transformations.
-
-\brk
-
-Recall that we had $\mathbf{Mod}$-$\calA=[\calA^{op},\Ch]$. The objects are dg functors and maps are natural transformations with a 
-hexagonal naturality diagram (I don't know what htis is).
-
-This is a honest category, but we can upgrade it to a dg category as follows: for each $X,Y\in\mathbf{Mod}$-$\calA$, we define $\mathscr{M}od\calA(X,Y) = \operatorname{eq}(D)$ where $D$ is the diagram
-\begin{center}
-	\begin{tikzcd}
-		\prod_{c\in\calA}\Ch(Xc,Yc)\ar[r,"\lambda",bend right]\ar[r,"\rho",swap,bend left]& \prod_{a,b\in\calA}\Ch(\calA(b,a),\scrC(Xa,Xb))
-	\end{tikzcd}
-\end{center}
-
-$\lambda$ is a way to define a right action by precompositon and $\rho$ is the same but right action and postcomposition. So in a way this is saying the two actions are compatible.
-This actually ends up being exactly what we usually ask for when we want naturality.
-
-\begin{rmk}
-	$Z^0\scrM od\calA(X,Y)=\mathbf{Mod}$-$\scrA$.
-\end{rmk}
-\begin{rmk}
-	$\scrM od\calA(X,Y)=\int_{a\in\calA}\Ch(Xa,Ya)$, if the notation suits you.
-\end{rmk}
-
-We can extend notions of cones, suspension, quasi-isomorphisms, etc, pointwise to $\scrM od \calA$. For instance, if $X\in\scrM od\calA$, we define $\Sigma X$ to be the dg functor 
-$A^{op}\xrightarrow{X}\Ch\xrightarrow{\Sigma}\Ch$. Again, just define the cone pointwise and since it is functorial (note: this is not true in the 
-derived theory, but it still is here.)
-
-We will call $X$ in $\scrM od\calA$ \textbf{acyclic} if $X(a)$ is acyclic for all $a$.
-
-\subsection{Some special dg modules}
-Whenever $a\in\calA$, there is a corresponding representable dg module $\hat a=\calA(-,a)$. The dg functor structure here is given by composition.
-In this case we get 
-\begin{lem}[Strong Yoneda]
-	There is a dg functor $\calA\to\scrM od\calA$ sending $a\mapsto \hat a$ which is dg fully faithful. In fact for every $X\in\scrM od\calA$,
-	\[\scrM od\calA(\hat a,X)\cong X(a).\]
-\end{lem}
-
-In particular, $\scrM od\calA(\hat a,-)$ preserves all things defined pointwise (including quasi-isos, acyclics, etc).
-
-\begin{defn}
-	The derived category $\D(\calA)$ is 
-	\[\D(\calA)=\mathbf{Mod}\text{-}\calA[\text{quasi isomorphisms}]^{-1}=H^0\scrM od\calA/{\text{acyclic complexes}}\]
-	which are analogous to what we usually do with abelian categries.
-\end{defn}
-\begin{prop}
-	$\D(\calA)$ is locally small.
-\end{prop}
-
-Inside of $\D(\calA)$ we can consider the images of the representable dg functors $\hat a$. We want to define a thing that sits over this 
-\[\mathbf{Perf}\calA\]
-to be the smallest full dg subcategory of $\scrM od\calA$ containing the image of the Yoneda embedding, closed under cones, suspensions, and homotopy retracts.
-
-\begin{thm}
-	Let $\calA$ be a small dg category. Then $\D(\calA)$ is a compactly generated triangulated category and the compacts are $\D(\calA)^c=H^0\mathbf{Perf}\calA$.
-\end{thm}
-
-In a sense, you want to remember that $\mathbf{Perf}\calA$ determines $\D(\calA)$. In particular, fif $\calA$ and $\calB$ are 
-small dg categories and we have a \textbf{quasi-equivalence} $F:\mathbf{Perf}\calA\to\mathbf{Perf}\calB$ (that is, $F$ induces a quasi-isomorphism on $\scrH om$s)
-and $F$ descends to an essential surjection on the homotopy categories of the $\mathbf{Perf}$ objects, then $\D(\calA)\xrightarrow{\sim}\D(\calB)$.
-
-In such a case we say that $\calA$ and $\calB$ are \textbf{derived Morita equivalent}.
-\begin{ex}
-	Consider the path algebra of the $A_2$ quiver $kA_2\cong (\begin{smallmatrix}
-		k&k\\0&k
-	\end{smallmatrix})$. Then $\mathbf{Perf}kA_2\cong\scrC^b(\operatorname{proj} kA_2)$.
-
-	This is derived Morita equivalent (in fact just equivalent) to $B\cong R\Hom_{kA_2}(S_2\oplus S_2,S_1\oplus S_2)$.
-\end{ex}
-
-In general, if we let $\calK$ be a small triangulated category with $\tilde\calK$ a dg category such that $\calK\cong H^0\tilde\calK$ as triangulated categories 
-where $H^0\tilde\calK$ inherits is triangulated structure from $H^0\scrM od\hat\calK.$ This is called an \textbf{enhancement of $\calK$.}
-
-Call $g\in\calK$ a \textbf{generator} if the smallest thick subcategory containing $g$ in $\calK$ is all of $\calK$.
-We call $g$ a \textbf{strong generator} if there exists a uniform bound on the number of cones needed to build
-any object (there is a global bound on the number of cones we have to take to build any object). In this latter case we call $\calK$ \textbf{regular} (in the sense of algebraic geometry).
-
-\begin{thm}[Keller]
-	If $g\in\calK$ is a generator, then $\tilde\calK$, an enhancement of $\calK$, is quasi equivalent to $\mathbf{Perf}\tilde\calK(g,g).$
-\end{thm}
-\begin{rmk}
-	What this is saying is that if $\calK$ (i.e. $H^0\tilde\calK$) ahs a generator then $\tilde \calK$ is derived morita equivalent to a dg algebra.
-\end{rmk}
-
-In practice, most triangulated categories we care about (at least the algebraic ones -- the topological ones don't work here) come from dg algebras.
-
-\subsection{Examples, properness, and smoothness}
-\begin{ex}
-	Let $R$ be a finitely generated $k$ algebra viewed as a dg algebra concentrated in degree zero. Then we have 
-	\[\Ch(R)=\scrM od R\supseteq \Ch^{-,b}(\operatorname{proj}R)=\underline\D^b(R)\supseteq \mathbf{Perf} R=\Ch^b(\operatorname{proj} R)\]
-	which is analogous to the inclusion 
-	\[\D(R)\supseteq \D^b(R)\supseteq\Dbperf(R)\]
-\end{ex}
-
-\begin{ex}
-	If $X$ is a scheme of finite tipe over $k$, then we can define $\operatorname{Perf} X$ to be the collection of injective resolutions of chain complexes of 
-	quasicoherent sheaves on $X$ such that these complexes are perfect in $\D_{qc}(X)$. The point here is that 
-	\[H^0\operatorname{Perf}X\cong\Dbperf(X).\]
-\end{ex}
-
-\subsection{Properness}
-Suppose $\Lambda$ is a finite dimensional $k$ algebra. If $M,N\in\mathbf{Perf}\Lambda$, then $\mathbf{Perf}\Lambda(M,N)\in\mathbf{Perf}k$ up to quasi-isomorphism.
-That is $H^\ast\mathbf{Perf}\Lambda(M,N)$ is finite deimensional over $k$.
-
-If $X$ is a proper $k$ scheme, then for $E,F\in\operatorname{Perf}X$, we have $\operatorname{Perf}X(E,F)\in\mathbf{Perf}k$. That is, the cohomology on a proper scheleme is finite dimensional.
-
-\begin{defn}
-	We say that a dg category over $k$ $A$ is \textbf{proper} if for all $a,b\in\calA,$ $\calA(a,b)\in\mathbf{Perf}k$ up to quasi isomorphism.
-\end{defn}
-
-An example of a proper category is $\operatorname{Perf}\Lambda$ where $\Lambda=k[\varepsilon]/\varepsilon^2$. But on the other hand, $\underline\D^b(\Lambda)$ is not proper. For instance,
-notice that a resolution is of $\Lambda$ with $\varepsilon$ maps everywhere, but $\Hom(P,P)$ computes $\Ext^\ast(k,k)$, which in every degree is $k$.
-
-\begin{lem}
-	Properness is a derived Morita invariant. That is a dg category over $k$ $\calA$ is proper if and only if $\operatorname{Perf}\calA$ is proper.
-\end{lem}
-\begin{rmk}
-In partucular an algebra $\Lambda$ is proper (i.e. finite dimensional) if and only if $\operatorname{Perf}\Lambda$ is.
-\end{rmk}
-\begin{thm}[Lipmon?]
-	If $X$ is a scheme of finite type over $k$, then $X$ is proper iff $\operatorname{Perf}X$ is.
-\end{thm}
-
-\subsection{Smoothness}
-Let $L=\bbF_p(t)$ and $R=\bbF_p(t^{1/p})$. Then $R$ is a field so $\gldim R=0$, but 
-$R\otimes_LR$ is nologingler finite global dimension since $t^{1/p}\otimes 1-1\otimes t^{1/p}$ is nilpotent.
-
-\begin{defn}
-	Recall that a $k$ algebra $R$ is \textbf{smooth} if $R\in\operatorname{Perf}R^e$.
-\end{defn}
-\begin{lem}
-	If $R$ is smooth over $k$, then $\gldim R<\infty$.
-\end{lem}
-
-Some geometry: $X$ of finite type over $k$ is smooth if $X\times_k \bar k$ is regular (where $\bar k$ is the algebraic closure of $k$). If $X$ has enough locally free sheaves (resolution property),
-then this is equivalent to 
-\[\Delta_\ast\calO_X\in\operatorname{Perf}(X\times_k X)\]
-where $\Delta$ is the diagonal map $X\to X\times X$.
-
-In an attempt to make a common generalization:
-\begin{defn}
-	Let $\calA$ and $\calB$ to be two dg-categories over $k$. Define $\calA\otimes_k \calB$ to be the category of pairs of 
-	objects from $\calA$ and $\calB$, written $a\otimes b$ and make the maps 
-	\[\calA\otimes \calB(a\otimes b,a'\otimes b')=\calA(a,a')\otimes \calB(b,b').\]
-\end{defn}
-Then we set $\calA^e=\calA^{op}\otimes_k\calA$. Then we consider $\scrM od\calA^e$ to be the categories of $\calA,\calA$-bimodules.
-
-\begin{ex}
-THe diagonal $\calA^e$ modules $\Delta:\calA^e\to\Ch$ is the map $\calA(-,-)$, sending a pair $a\otimes a'$ to $\calA(a,a')$.
-\end{ex}
-\begin{rmk}
-	This gives us another definition: $\calA$ is smooth over $k$ if $\Delta=\calA(-,-)\in\operatorname{Perf}\calA^e$.
-\end{rmk}
-\begin{lem}
-	Smoothness is a derived Morita invariant. Thus it agrees with the notion for algebras.
-\end{lem}
-\begin{thm}[Lunts-Schn\"urer]
-	If $X$ is finite type, separable, and has enough locally free sheaves, then $X$ is smooth iff $\operatorname{Perf}X$ is smooth.
-\end{thm}
-
-There are lots more exotic examples including things you wouldn't really expect to be smooth.
-\begin{thm}[Auslander, Elogin-Lunts-Schn\"urer]
-	If $\Lambda$ is a finite dimensional algebra over $k$ such that $\Lambda/\operatorname{rad}\Lambda$ is separable over $k$ (e.g. when $k$ is algebraically closed), then 
-	$\underline\Db(\mathbf{mod}\,\Lambda)$ is smooth.
-\end{thm}
-\begin{rmk}
-	Notice that by contrast that $\operatorname{Perf}\Lambda$ is smooth if and only if $\Lambda$ is. This is a smaller object, so violates our geometric intuition.
-
-	The upshot here is that we can regard $\operatorname{Perf}\Lambda\hookrightarrow\underline\Db(\Lambda)$ to be a kind of resolution of singularities.
-\end{rmk}
-\begin{thm}[Lunts]
-	If $X$ is a separable scheme of finite type over a perfect field, then $\underline\Db(X)$ is smooth.
-\end{thm}
-By Morita invariance, if $E$ is a dg algebra and $\operatorname{Perf}E$ is smooth, then $E$ is smooth. For instance, take a prime $p\ge 3$, let $k=\bbF_p$
-and consider $\underline\Db(k C_p)\cong\operatorname{Perf}E$ where $E=R\Hom_{kC_p}(k,k)$.
-
-But now notice that
-\[H^\ast(E)=Ext_{k C_p}^\ast(k,k)=H^\ast(C_p,k)=k\langle\tau,\theta\rangle\]
-where $|\tau|=1$ and $|\theta|=2$ and the above is the free graded commutative algebra on these things.
-Now since $|\tau|=1$, $\tau^2=0$, so $H^\ast E$ has infinite global dimension, so $H^\ast E$ is not smooth.
-
-In general, one can't hope to say much about $H^\ast A$ when $A$ is smooth. However,
-\begin{thm}[Raedschelder-S.]
-	If $A$ is a smooth dg algebra is smooth and connective ($A^{\ge 1}=0$), then $H^0A$ is smooth.
-\end{thm}
-
-\subsection{Regularity and smoothness}
-What exactly does smoothness buy us?
-\begin{thm}
-	If $A$ is a smooth dg algebra then $A$ is regular in the sense that $A$ strongly generates $H^0\operatorname{Perf}A$.
-\end{thm}
-Regularity buys one quite a lot: we get representability results for cohomological (dg) functors. The idea here is that even if we don't have the infinite coproducts we need for Brown,
-but somehow the ``needing uniformly finite cones'' of strong generation takes care of this for us.
-
-\begin{rmk}
-	It is not always true that for any Noetherian scheme that $\Db(X)$ is even regular.
-
-	There are many examples of dg categories that are regular but not smooth. For instance, $\operatorname{Perf}k[[x]]$ is regular, but not smooth.
-\end{rmk}
-
-Smoothness is also a strong finiteness condition on $\calA$. 
-\begin{thm}[To\"en]
-	Let $\calA$ be a smooth dg category. Then there exists a generator $g\in\operatorname{Perf}\calA$ so 
-	$\calA$ is derived Morita equivalent to $\operatorname{Perf}\calA(g,g)$, a dg algebra.
-\end{thm}
-
-\subsection{Ideas for proofs}
-Say $\calA$ is a smooth dg category, so $\Delta\in\operatorname{Perf}\calA^e$. What are the perfects over the enveloping algebra? For $a\otimes b\in \calA^e$, 
-write $\widehat{a\otimes b}$ for $\calA(a,-)\otimes\calA(-,b)\in\operatorname{Perf}\calA$. But then 
-\[\operatorname{Perf}\calA^e=\operatorname{thick}(\widehat{a\otimes b}|a\otimes b\in\calA^e)\]
-then the upshot here is that computing the thick subcategory containing something is built out of finitary pieces. That is, 
-for anything in $\operatorname{Perf}\calA^e$, $\Delta\in\operatorname{thick}_{\calA^e}(\widehat{a_i\otimes b_i}|i\in\calI)$
-where $\calI$ is a finite set.
-
-So if $X\in\operatorname{Perf}\calA$, consider $X_\calA\otimes_\calA {_\calA}\Delta_\calA\cong X\in\scrM od\calA$
-but then 
-\[X=X\otimes_\calA \Delta\in\operatorname{thick}_\calA(X\otimes_\calA \widehat{a_i\otimes b_i}|i\in\calI)\]
-and futhermore 
-\[X\otimes_\calA\widehat{a_i\otimes b_i}= X(a_i)\otimes_k \hat b_i\]
-so this is a big direct sum of shifts of $\calA(-,b_i)$.
-
-Then by Thomason's this implies that $X\in\operatorname{thick}(b_i|i\in\calI)$, so the $b_i's$ genreate $\operatorname{Perf}\calA$, 
-whence $\calA$ is morita equivalent to $\oplus_{i,j\in\calI}\calA(b_i,b_j)$.
-
-Finally since $\Delta$ can be built in some finite number of cones, $X$ can be build in the same (or fewer) number of cones.
-
-\begin{rmk}
-	In general, smoothness and properness are often dual to one another.
-That is they often come in pairs.
-FOr example, if $\Lambda$ is a finite dimensional dg algebra over $k$, then $\operatorname{Perf}\Lambda$ is proper.
-
-	Then $\operatorname{Perf}k$ is the monoidal unit for a closed monoidal structure on a dg category over $k$ $[\operatorname{Perf}\Lambda,\operatorname{Perf}k]=[\Lambda,\operatorname{Perf}k]\cong\underline\Db(\Lambda^{op})$ and the thing on the right is smooth.
-	This is not an exact duality because there are counterexamples but it's mostly correct.
-\end{rmk}
-
-\section{(Bounded) t-structures and approximable categories}
-Let $R$ be a ring. Consider the category $\D(R)$, the derived category of $R$ modules. This is a triangulated category with 
-coproducts. Write $\Sigma$ for the suspension.
-
-We have two (full) subcategories $\Dpos(R)$ and $\Dneg(R)$, which are elements such that cohomology 
-vanishes below and above zero. Notice that $\Dneg(R)$ is stable under $\Sigma$ and $\Dpos(R)$ is stable 
-under $\Sigma^{-1}$. Sometimes we write $\D^{\ge 1}(R)$ (e.g.).
-
-It is not too hard to check that $\Hom(\Dneg(R),\D^{\ge 1}(R))=0$. For each $x\in\D(R)$, there is 
-a triangle 
-\[Y\to X\to Z\]
-where $Y\in\Dneg(R)$ and $Z\in\D^{\ge 1}(R).$
-
-Finally $\Dneg(R)\cap\Dpos(R)=\Rmod$.
-
-\subsection{t-structures}
-\begin{defn}
-	A \textbf{t-structure} on a triangulated category $\calT$ is a pair of full subcategories $\calT^{\ge 0},\calT^{\le 0}$ satisfying 
-	\begin{itemize}
-		\item $\Sigma\calT^{\le 0}\subset \calT^{\le 0}$
-		\item other stuff I missed.
-	\end{itemize}
-\end{defn}
-
-The \textbf{heart of $\calT$} is $\calT^{\ge 0}\cap\calT^{\le 0}$ which 
-is an abelian subcategory of $\calT.$ We say $\calT^{\le 0}$ and $\calT^{\ge 0}$ are the \textbf{aisle and coaisle} of $\calT$,
-respectively.
-\begin{rmk}
-	The t-structure is determined entirely by the (co) aisle. This is because
-	\[\calT^{\ge 0}=(\calT^{\le 0})^\perp:=\{Y\in\calT|\Hom(\calT^{\le -1})\}\]
-\end{rmk}
-
-We have truncation functors that are right and left adjoints to $\calT^{\le 0}\hookrightarrow\calT$ and $\calT^{\ge 0}\hookrightarrow\calT$, respectively.
-
-\subsection{Aisles vs Pre-Aisles}
-Every aisle $\calT^{\le 0}$ is closed under positive suspensions, taking summands, and extensions. If $\calT$ has coproducts, then $\calT^{\le 0}$ is closed under coproducts.
-\begin{defn}[Keller-Vossieck]
-	A subcategory $\calS\subseteq\calT$ is a \textbf{pre-aisle} if it is closed under extensions, positive suspensions, and taking summands.
-\end{defn}
-
-A question one may ask is when is a pre-aisle an aisle? Keller-Vossieck said that a pre-aisle $\calS\subseteq \calT$ is an aisle if and only if $\calS\hookrightarrow\calT$
-has a right adjoint.
-
-Let $S\subseteq\calT$ be a set of objects. Write 
-\begin{itemize}
-	\item $\langle S\rangle$ to be the smallest thick triangulated subcategory of $\calT$ contatining $S$.
-	\item $\langle\overline{S}\rangle$ is the same as above but closed under coproducts.
-	\item $\langle S\rangle^{(-\infty,0]}$ is the smallest pre-aisle in $\calT$ containing $S$
-	\item $\langle\overline{S}\rangle^{(-\infty,0]}$ is what you'd think.
-\end{itemize}
-
-\begin{thm}[Neeman]
-	Let $\calT$ be a well-generated (whatever that means) category and $S\subseteq\calT$ a set of objects. Then 
-	there exists a t-structure with aisle $\langle \overline{S}\rangle^{(-\infty,0]}$
-\end{thm}
-
-As an example, when $\calT=\D(R)$, then $\calS=\langle \overline{R}\rangle^{(-\infty,0]}$ is the aisle of the standard t-structure.
-
-Another example: Let $\calT=\Db(R)$ where $R$ is Noetherian. Let $S=\langle R\rangle^{(-\infty,0]}=\calT_c^{b\le 0}$ is the aisle of the restricted standard t-structure.
-This structure is bounded. That is, $\calT$ is the intersection of the shifted versions of the usual structures.
-
-Why do we care about bounded t-structures? One reason is that if $\calT$ has a bounded t-structure, then $\calT$ is generated by its heart. Another reason is that 
-it is an ingredient in Bridgeland stability. Finally is something to do with K theory that will show up in David's talk.
-
-\subsection{A nontrivial pre-aisle}
-Bondal found the following example of a pre-aisle which is not an aisle:
-\begin{ex}
-	Let $\calD=\Db_{coh}(\bbP^2(k))$ where $k$ is algebraically closed of characteristic zero. Let 
-	\[D_1^{\le 0}\langle \calO(1),\calO(2)\rangle\]
-	and 
-	\[D_2^{\le 0}=\langle\calO(-1),\calO(-2)\rangle.\]
-
-	These are both aisles as well, but $D_1^{\le 0}\cap D_2^{\le 0}$ is a pre-aisle but it can be shown (though significant computation)
-	to not be an aisle.
-\end{ex}
-
-\subsection{Neeman's Great Idea}
-Let $\calS=\calT^{\le 0}$ for some t-structure on $\calT$. Let $x,y\in\calT$. We get as surjective map 
-\[\left\{x\to x\to y|s\in\calS\right\}\twoheadrightarrow\Hom(x,y^{\le 0})\]
-by virtue of the fact that any map from a truncated thing to an object factors through its truncation. Neeman showed that under 
-a certain equivalence relation we can make this into an isomorphism.
-
-The quotient $H_\calS(x,y)\cong\Hom(x,y^{\le 0})$ can be defined for any pre-aisle $\calS$, so we get a homological functor 
-$H_\calS(-,y):\calT\to\Ab$.
-
-\begin{prop}[Neeman]
-	Let $\calS$ be a pre-aisle. $\calT-\calT^{\le 0}$ for some t-structure on $\calT$ if and only if $H_\calS(-,y)$ is representable for all $y\in\calT$.
-\end{prop}
-\begin{rmk}
-	Notice that here is where we run into set theoretic issues. One must show that such a functor is set-valued to apply Brown representability.
-\end{rmk}
-
-\subsection{Approximating Objects}
-\begin{defn}
-	A \textbf{metric} on a category is a function that assigns a positive real number (length) to every morphism in a way that the triangle inequality is satisfied.
-\end{defn}
-
-Now every t-structure on $|calT$ gives rise to a metric on $\calT$. To understand it, it suffices to specify the balls 
-\[B_n=\{x\in\calS|0\to X\text{ has length}\le 1/n\}\]
-Taking the balls to be $B_n=\calT^{\le -n}$ gives rise to a ``good metric''. I missed the discussion here unfortunately.
-
-Recall the definition of compact generators.
-
-\begin{defn}
-	Let $\calT$ be a triangulated category with coproducts. Then $\calT$ is \textbf{approximable} if there exists a compact 
-	generator $G\in\calT$ and a t-structure and an integer $A>0$ such that 
-	\begin{itemize}
-		\item $G\in T^{\le A}$ and $\Hom(G,T^{\le -A})=0$.
-		\item For every object $y\in\calT^{\le 0}$, there exists a triangle
-		\[x\to y\to x\]
-		with $z\in\calT^{\le -1}$ and $x\in\calT^{\ge 0}$ (I think).
-	\end{itemize}
-\end{defn}
-
-Some examples due to Neeman: $\D(R)$ is approximable. IF $X$ is quasicompact and separated, then $\D_{qc}(X)$ is approximable. Finally the homotopy category of spectra is approximable.
-
-\subsection{\texorpdfstring{$\calT^b_c$}{Tcb}}
-Suppose that $\calT$ has a compact generator $G$. Consider the t-structure generated by $G$ (the aisle is generated by $G$). Call $\calT_c^-$
-the full subcategory of $\calT$ such that for all $n>0$ there exists a triangle $x\to y\to z$ with $x$ compact and $z\in\calT^{\le -n-1}$. That is, $y$ is being approximated by compact objects.
-
-\begin{defn}
-\[\calT_c^b=\calT^b\cap\calT^-c\]
-\end{defn}
-\begin{rmk}
-	The categories above do not depend on the choice of generator.
-\end{rmk}
-\begin{rmk}
-	If $\calT$ is approximable, then $\calT_c^-$ and $\calT_c^b$ are thick.
-\end{rmk}
-
-For some examples, when $\calT=\D(R)$, $\calT^b=\Db(R)$ and $\calT_c^-=\Dneg(R)$. If you look at quasicoherent sheaves you also get something nice.
-
-\section{Cluster tilting modules for mesh algebras}
-
-This talk was given by Sira Gratz in conjunction with Erdmann and Lamberti.
-
-The motivation here is that the clustering aspect comes from combinatorics (e.g. cluster algebras) and the tilting portion 
-refers to tilting theory/modules.
-
-Thoughout let $\Lambda$ be a finite dimensional algebra over $k$.
-If $\Lambda$ has a CT-modules, this implies that the representation dimension of $\Lambda$ is less than or equal to three.
-What does this mean?
-That $\Lambda$ is close to being ``representation finite''. Auslander proved that $\operatorname{repdim} \Lambda =2$ iff it is representation finite.
-
-That is $0\to T''\to T'\to M\to 0$ exists for any $M$, where $T'$ and $T''$ are indecomposibles. So in some way if you understand the indecomposable theory
-you can reconstruct the structure of any module using only two ``layers'' of this information.
-
-\subsection{What are CT-modules?}
-\begin{defn}
-	Let $\calT$ be a triangulated or Abelian category. Then $\calC\subseteq\calT$ is \textbf{cluster tilting}
-	if it is functorially finite (This means you have pre-covers and pre-envelopes for every object in your category -- roughly $\calC$ approximates $\calT$) and the folowing are equivalent for all $M\in\calT$:
-	\begin{itemize}
-		\item $M\in \calC$
-		\item $\Ext^1(C,M)=0$ for all $C\in\calC$
-		\item $\Ext^1(M,C)=0$ for all $C\in\calC$
-	\end{itemize}
-\end{defn}
-
-If $\calT=\mathbf{mod}$-$\Lambda$, then $A$ in this category is a CT-modules if $\operatorname{add}(A)$ is CT.
-A good place to look to start is $\Lambda$ that are self-injective.
-
-\begin{thm}[Erdmann-Holm]
-	If $\Lambda$ is self-injective and has a CT-module, then for all $\Lambda$ modules $M$ have complexity less than 1. 
-\end{thm}
-\begin{rmk}
-	Basically this talks about how quickly the dimension of the elements in a ``minimal'' projective resolution grows. Complexity zero means that a module has finite 
-	projective dimenson and complexity one means that the dimension of the modules in the resolution is bounded.
-\end{rmk}
-
-\subsection{Mesh Algebras (of Dynkin type)}
-The benefit of these algebras is that they have periodic resolutions, so we can apply the previous theorem.
-
-One example comes from the quiver 
-\begin{center}
-	\begin{tikzcd}
-		0\ar[r,"\bar\alpha",bend right,swap] & 0 \ar[l,"\alpha",bend right]\ar[r,"\beta",bend right] & 2 \ar[l,"\bar\beta",bend right,swap]
-	\end{tikzcd}
-\end{center}
-
-The idea here is you look at a Galois cover of this diagram and then use graph automorphisms to introduce a ``twist'' to 
-create a mesh algebra.
-
-\begin{thm}[Geiss-Leckers-Schr\"oer]
-	Mesh algebras of type $A,D,$ and $E$ have CT-modules.
-\end{thm}
-\begin{thm}[Gratz-Erdmann-Lamberti]
-	All mesh algebras (of Dynkin type) have CT-modules.
-\end{thm}
-
-Another result:
-\begin{thm}[Darpo-Iyama]
-	If $\calC$ is $k$-linear and locally bounded such that $\operatorname{add}(\calC)$ is Krull-Schmidt $G$ admissibly,
-	there is a one-t-one correspondence between $G$-equivariant CT-subacategories of $\calC$ and CT-subcategories of $\calC/G$.
-\end{thm}
-\subsection{Mutation}
-Alnother result from the original group:
-\begin{thm}[GLS]
-	If $\Lambda$ is a mesh algebra of type $A,D,$ or $E$, and $T$ is a CT-module in $\mathbf{mod}$-$\Lambda$. Let $X$ be a non-projective indecomposable 
-	summand  then there exists a unique $Y\not\cong X$ such that $T/X\oplus Y$ is CT. There are also some short exact sequences.
-\end{thm}
-
-The theory here works on the idea that $\Ext^1(X,Y)\cong D\Ext^1(Y,X)$ which is not in general true. But in our case 
-we get that (if $\gamma$ is the automorphism of $\Lambda$ induced by $\sigma$) that $\Ext^1(X,Y)\cong D\Ext^1(Y,{_\gamma}X)$, where 
-${_\gamma}X$ is the twisting of the $\Lambda$ action on $X$ by $\gamma$.
-
-A futher result is that each of the CT modules are $\gamma$-equivariant except in the exception case of $P(G_2)$, the mesh algebra 
-corresponding to $D_4$, that has an order 3 automorphism. In this case, there exists such a CT module, but they are not all (something).
-
-\section{Triangulated Categories (and syzygies) associated to singularities of algebraic varieties.}
-This talk was given by Jesse Burke.
-
-Here let $P=k[x_0,\dots,x_n]$ and let $P_j$ be the homogenous degree $j$ polynomials. An ideal $I$ of $P$ is \textbf{homogeneous} if there 
-exists a set of homogeneous generators. A $P$-module is homogeneous if $M=\oplus M_i$ and $P_j\otimes M_i$ maps to $M_{j+i}$.
-
-Write $P(m)_j=P_{m+j}$.
-
-\subsection{Hilbert and Syzygies}
-Hilbert was one of the first to systematically use syzygies. As an example, let $P=k[x_0,\dots,x_3]$ and let
-\[R=\frac{P}{(x_0x_2-x_1^2,x_0x_3-x_1x_2,x_1x_3-x_2^2)}\]
-which gives us an embedding of $\bbP^1\hookrightarrow\bbP^3$.
-
-A free resolution here is 
-\[0\leftarrow R\leftarrow P^1\leftarrow P(-2)^3\leftarrow P(-3)^2\leftarrow 0\]
-In general it is difficult to look at abunct of equations and talk about what properties a variety has. One thing Hilbert did was 
-write down the Hilbert series $H(R)=\sum_n \dim_k R_nt^n$. Hilbert proved that this is always rational with a particular denominator.
-
-Let $R$ be a graded quotient ring of polynomial ring $P$. A graded free resolution of a graded $R$ module is what you'd expect.
-A graded free resolution is minimal if all differentials have entries in $(x_0,\dots,x_n)$.
-
-\subsection{Examples}
-\begin{ex}
-	Let $R=k[x,y]$ and $M=k=R/(x,y)$, the residue field. Then we have resolution 
-	\[0\to R(-2)\xrightarrow{\binom{-y}{x}} R(-1)^2\xrightarrow{(x,y)} R^1\to k\to 0\]
-	which is the same as the Koszul complex.
-\end{ex}
-\begin{ex}
-	Now let $R=k[x,y]/(x^2)$. Then a resolution of $k$ looks like 
-	\[\cdots R(-3)^2\to R(-2)^2\to R(-1)^2\to R\to k\to 0\]
-	where eventually the maps become periodic (so the syzygies are periodic).
-\end{ex}
-\begin{ex}
-	When $R=k[x,y]/(x^2,y^2)$, the $k^{th}$ syzygy is generated by $k$ elements.
-\end{ex}
-\begin{ex}
-	When $R=k[x,y]/(x^2,xy)$, You get that the rank grows like the Fibonacci numbers.
-\end{ex}
-
-\section{Modular realisations of derived equivalences in representation theory}
-This talk was given by Daniel Chan on work with Tarig Abdelgadir and Boris Lerner.
-
-We always work over an algebraically closed field $k$ of characteristic zero. The motto here is that Moduli stacks are a fruitful way to study a finite dimension algebra because 
-they are essentially a machine to construct functors.
-
-As usualy, we should fix some discrete invariants of a finite dimensional $A$ module. You can start by 
-fixing $\dim M$. Slightly more subtlely, you can fix an idempotent $e$ and fix $\dim Me$.
-
-Recall the path algebra: start with a quiver $Q$ (directed graph) without oriented cycles.
-Let $kQ$ be the path algebra with $k$-basis all paths of length $\ge 0$. Multiplication is concatenation or zero.
-
-For every vertex $v$, there is a path $e_v$ of length zero which is idempotent.
-
-\subsection{King's Interpretation of Beilinson's derived equivalence}
-Let $Q$ be the Kronecker quiver on two vertices with two arrows between them in the same direction. $A=kQ$. Fix the dimension of our modules to be $(1,1)$.
-
-This got pretty interesting but also complicated so I just listened. :) If I ever get into stacks this seems like a good guy to read.
-
-\section{Groups, Spherical Twists and Stability Conditions}
-This was a three-part lecture given by Asilata Bapat, Anand Deopurkar, and Anthony Licata.
-
-The goal of this first talk is to paint a picture of the aim of this ``movement''. More specifically, one goal is to use homological algebra
-to prove some things about groups. This will hopefully cement this theory of a thing that group theorists care about.
-
-\subsection{What kinds of groups?}
-One kind are Artin-Tits braid groups. We will not use full generality but will suffice here. Fix some finite connected graph (no loops) $\Gamma$.
-Then this graph yeilds $B_{r_\Gamma}$, a group given by generators $\sigma_i$ for each vertex. Then we add the relations 
-$\sigma_i\sigma_j=\sigma_j\sigma_i$ if $i$ and $j$ are not connected by an edge and the braid relation $\sigma_i\sigma_j\sigma_i=\sigma_j\sigma_i\sigma_j$ if they are.
-
-Very little is known about these groups. For instance, there are no known solutions to the word problem (when are two words the same) and the conjugacy problem (when are two elements conjugate). 
-People also don't know the what the centers of these groups are or whether they have finite representation dimension.
-
-It is not known whether $B_{r_\Gamma}$ are linear. However, these groups do act on ``nice'' triangulated categories. For example: 
-\begin{ex}
-	Let $\scrP$ denote the $\bbC$-linear additive category generated by $\{P_i\}_{i\in \Gamma_0}$ (indexed by the vertices). Then the hom sets are given by 
-	\[\Hom_\scrP(P_i,P_j)=\left\{\begin{array}{lr}
-		\bbC\cdot 1\oplus \bbC \cdot x_i, & i=j\\
-		\bbC\cdot y_{ij}, & (i,j)\in\Gamma_1\\
-		0, & \text{otherwise}.
-	\end{array}\right.\]
-	To define the composition, one can think of there as being a graded structure on $\Hom^\ast$ given by $\deg(x_i)=2$ and $\deg(y_{ij})=2$ and composition respects grading.
-\end{ex}
-
-Now let $\calT_\Gamma=K^b(\scrP_\Gamma)$. 
-\begin{thm}[Huerfano-Khovanov]
-	$B_{r_\Gamma}$ acts on $\calT_\Gamma$.
-
-	The generator $\sigma_i$ acts by twisting in $P_i$:
-	\[\sigma_i(X):=\cone(P_i\otimes_\bbC \Hom(P_i,X)\xrightarrow{ev}X)\]
-\end{thm}
-
-We have a couple conjectures:
-\begin{conj}
-	The action above is faithful.
-\end{conj}
-and even more strongly (and importantly):
-\begin{conj}
-	The moduli space of Bridgeland stability conditions(this word may be wrong) $\operatorname{Stab}(\calT)$ is contractible.
-\end{conj}
-
-\subsection{What kind of picture are we trying to paint?}
-As a small (seeming) aside: How does one study mapping class groups of surfaces?
-
-Let $\Sigma$ be a surface, possibly with boundary and punctures. Then 
-\[\operatorname{MCG}(\Sigma)=\frac{\operatorname{Diff}_+(\Sigma)}{\operatorname{Diff}_0(\Sigma)}\]
-of oriented diffeomorphisms of the $\Sigma$ to itself modulo the ones isotopic to the identity. As an example, if $\Sigma$ is a disk with 
-$n$ punctures, the mapping class group is the braid group.
-
-Now the MCG of $\Sigma$ acts on:
-\begin{itemize}
-	\item $\operatorname{Teich}(\Sigma)$, the Teichm\"uller space, or the moduli space of hyperbolic metrics on $\Sigma.$
-	\item The isotopy classes of simple closed (multi-)curves on $\Sigma$.
-\end{itemize}
-
-Thurston explained how to unify these two actions. The first thing to notice is that the Teichm\"uller space comes as a manifold while the other set (call it $S$)
-doesn't have a clear topology. So in unifying the two, Thurston developed a method of understanding it through charts. The charts are ``train tracks'' which are a bit confusing to me.
-They each have three distinguished points and they (through four symmetries) are supposed to categorize each of the possible (maybe simple closed) curve in your space.
-I see there being a problem if the distinguished points are not punctures (e.g. if a curve goes through them).
-
-Then Thurston defined a compactification of the Teichm\"uller space. You can embed them into $\bbR^S$, the (nonzero) maps from $S$ to $\bbR$, which itself embeds (obviously)
-into its projective closure.
-
-Then the closure of $\operatorname{Teich}$ is the closure of its image in $\bbP(\bbR^S)$. It ends up that the interior is precisely Teich, and the 
-boundary has an interpretation as the moduli space of projective measured foliations on $\Sigma.$
-
-So in particular, MCG acts on $\overline{\operatorname{Teich}}=\operatorname{Teich}\sqcup \text{PMF}$. One can show that Teich is homeomorphic to $\bbR^n$ and PMF is homeomorphic to $S^{n-1}$, 
-so the closure of Teich is homeomorphic to the closed Euclidean ball. This gives us solutions to various algorithmic problems.
-
-For instance, each $g\in$MCG$(\Sigma)$ gives rise to a dynamical system by studying repeated action of $g$ on the closed ball. In particular, one is generally interested in fixed points which yields 
-Neelson-Thurston classification of mapping classes into periodic, reduced, and pseudo-Arasov.
-
-\subsection{Chainging gears}
-\textbf{Our goal is to try to recreate this construction using as our group the autoequivalence group of a triangulated category $\Aut(\calT)$.}
-
-We need a bit of a starting point here to even give us reason to believe this is the correct approach.
-Luckily it becomes clear that (Bridgeland) stability conditions naturally play the role of the Teichm\"uller space.
-So then we need to decide what plays the role of $S$, and here we will use spherical (stable) objects.
-
-That is, we aim to study an action of $\Aut(\calT)$ on a compactification of $\operatorname{Stab}(\calT)$. We will talk about these in more detail in talks 2 and 3.
-In particular, talk two will include some (rather technical) definitions for $\operatorname{Stab}(\calT)$ and the compactification. In talk three, we will 
-describe an example that we understand, fully developed in the world of homological algebra.
-
-\subsection{Definitions and compactification}
-FIx a triangualted category $\calT$. Recall that 
-\begin{defn}
-	A Brigeland stability condition on $\calT$ is $(Z,P)$ where 
-	\[P=\{P(\phi)|\phi\in\bbR\}\]
-	(notice that $P(\phi)$ is a full subcategory of $\calT$) satisfying 
-	\begin{itemize}
-		\item $P(\phi+1)=P(\phi)[1]$
-		\item If $\phi_1>\phi_2$, then $\Hom(P(\phi_1),P(\phi_2))=0$
-		\item If $E\in\calT$ then there exists a unique $\phi_1>\phi_2>\cdots>\phi_n$ and a uniqu fitration 
-		\[0=E_0\to E_1\to\cdots E_n=E\]
-		where each $E_i\to E_{i+1}$ admits $A_{i+1}\in P(\phi_{i+1})$ and maps 
-		\[E_{i+1}\to A_{i+1}\to E_i\]
-		this is called the Harder-Narasimhan filtration.
-	\end{itemize}
-
-	Furthermore $Z:K_0(\calT)\to\bbC$ such that if $A\in P(\phi)$, then $Z(A)=m(A)\cdot e^{i\pi\phi}$ where $m(A)\in\bbR_{>0}.$
-\end{defn}
-
-As a matter of notation:
-\begin{itemize}
-	\item $P$ is called a \textbf{slicing} of $\calT$
-	\item $Z$ is called the \textbf{central charge}
-	\item If $\phi\in\bbR$, the objects of $P(\phi)$ are called \textbf{semistable of phase $\phi$} (simple objects of $P(\phi)$ are called \textbf{stable})
-	\item If $A\in P(\phi)$, then $m(A)$ is called the \textbf{mass} of $A$. In general $m(E):=\sum m(A_i)$ where $A_i$ are as in the filtration.
-\end{itemize}
-
-A slicing gives a bounded t-structure on $\calT$: for any $\phi\in\bbR$, $(P(\ge\phi),P(<\phi+1))$
-form your t structure whose heardt is $P([\phi,\phi+1])$. Given any (Z,P), we'll say that $P([0,1])$ is the \textbf{standard t-structure for $(Z,P)$.}
-
-\begin{prop}[Bridgeland]
-	Specifing $(Z,P)$ on $\calT$ is equivalent to giving 
-	\begin{itemize}
-		\item A bounded t structure on $\calT$ with heart $\calA$
-		\item Two functions $m:\calA\setminus\{0\}\to \bbR_{\ge 0}$ and $\text{ph}:\calA\setminus\{0\}\to [0,1)$
-		such that $Z(A):= m(A)e^{i\pi\text{ph}(A)}$ is a group homomorphism $K_0(\calA)\to\bbC$.
-		\item (HN Condition) This is automatic if objects of $\calA$ are of finite length.
-	\end{itemize}
-\end{prop}
-\begin{rmk}
-	We can recover $P$ as follows: if $\phi\in[0,1),$ then 
-	\[P(\phi)=\{A\in\calA|\text{ph}(A)=\phi\text{ and } \forall 0\ne B\subseteq A, \text{ph}(B)\le\text{ph}(A)\}\]
-	otherwise if $\phi\in[n,n+1)$, $P(\phi)=P(\phi-n)[n].$
-\end{rmk}
-
-Set the space $\operatorname{Stab}\calT$ to be the set of all stability conditions $(Z,P)$ on $\calT$.
-\begin{thm}[Bridgeland]
-	$\operatorname{Stop}(\calT)$ is a (complex) manifold.
-\end{thm}
-In fact, there is a $\bbC$-action on $\operatorname{Stab}\calT$ where if $w=x+iy$ and if $(Z,P)\in\operatorname{Stab}(\calT)$,
-then $Z\mapsto e^\omega Z$ (``scaling mass'') and $P(\phi)\mapsto P(\phi+\frac{y}{\pi})$ (``rotating phase''). Then we want to look at $\operatorname{Stab}\calT/\bbC.$
-
-\subsection{Plugging things into yesterday's talk}
-We will be replacing Teich with $\operatorname{Stab}\calT/\bbC$, MCG with $\Aut\calT$, and the (honest) sphere with the set of spherical objects of $\calT$ modulo shifts.
-
-\begin{defn}
-	An object $A$ is spherical if
-	\[\Hom^i(A,A)=\left\{\begin{array}{lr}
-		k, & i=0,2\\
-		0,& \text{otherwise}
-	\end{array}\right.\]
-\end{defn}
-
-\subsection{Construction the closure of our space}
-We consider the map $\operatorname{Stab}/\bbC\xrightarrow{\iota} \bbP\bbR^S$ that sends $\tau$ to the map $c\mapsto m_\tau(c)$. We will also use the morphism 
-$S\xrightarrow{\delta} \bbP\bbR^S$ that sends $A\in S$ to the map $B\mapsto \overline\hom^\ast(A,B)$ where the over line denotes we set $\overline\hom(A,A)=0$.
-
-Set $\overline{\operatorname{Stab}/\bbC}$ to be the closure of the image of $\iota$ in $\bbP\bbR^S$.
-
-\subsection{Some conjectures}
-These are pretty loose/wishful thinking. 
-\begin{conj}
-	$\iota$ is an embedding.
-\end{conj}
-\begin{conj}
-	$\delta$ is an embedding.
-\end{conj}
-\begin{conj}
-	If $\overline{\operatorname{Stab}/\bbC}=\operatorname{Stab}\sqcup \text{Bdy}$, then $S$ is in the boundary and is dense in the boundary.
-\end{conj}
-\begin{conj}
-	$\overline{\operatorname{Stab}/\bbC}$ is a closed ball which is the union of an open ball $\operatorname{Stab}/\bbC$ with boundary a sphere.
-\end{conj}
-
-This was really cool. I especially enjoyed the third talk. :)
-
-\section{Bounded t-structures and negative K-theory of stable infinity categories}
-This was given by David Gepner. It is joint work with Ben Antieau and Jeremiah Heller.
-
-We've heard a lot about t structures over this conference. Here we will talk about some obstructions to getting t structures that arise from K theory.
-First some conjectures:
-\begin{conj}[Marco Schlicting]
-	If $\calA$ is a small abelian category then $K_n(\calA)=0$ for $n<0$.
-\end{conj}
-\begin{conj}[A-G-H]
-	If $\calC$ is a small stable $\infty$-category with a bounted t structure, then $K_n(\calC)=0$ for all $n<0$.
-\end{conj}
-\begin{conj}
-	If $\calC$ is a small stable infinity category with bounded t-structure, then $K_n(\calC^\heartsuit)\simeq K_n(\calC)$ is an equivalence for all $n\in\bbZ$.
-\end{conj}
-\begin{rmk}
-	Actually for non-negative $n$, this is the so-called ``theorem of the heart'' that may be due to Neeman. Notice that $K(\calA)=K(\Db(\calA))$ where the bounded 
-	derived category is not the regular category, but the category where we remember the dg structure.
-\end{rmk}
-
-The second conjecture above implies both the first and second.
-\begin{rmk}
-	$K_0$ is an invariant of the triangulated homotopy category of a dg or stable infinity category. That being said, 
-	the entire K theory $K(\calC)$ depends on a dg (or stable) enhancement.
-\end{rmk}
-
-\subsection{Background on (stable) \texorpdfstring{$\infty$}{infty}-categories}
-The idea here is that an $\infty$-category is a category enriched in spaces, where we actually mean that it is enriched up to coherent homotopy.
-So for every $A,B\in\calC$, there is a ``mapping space'' $\operatorname{Map}(A,B)$ with a composition operator that is only well-defined up to homotopy.
-
-\begin{defn}
-	An infinity category $\calC$ is \textbf{stable} if it has finite limits and colimits, a zero object, and a square is a pushout iff it is a pullback.
-\end{defn}
-\begin{rmk}
-	In case you need help wrapping your head around limits in infinity categories, it is useful to remember that $\mathbf{Cat}\hookrightarrow\mathbf{Cat}_\infty$
-	fully faithfully via a map called the \textbf{nerve}. Furthermore this functor commutes with limits and colimits, so these things kind of what you think they should be.
-\end{rmk}
-\begin{rmk}
-	If $\calC$ is stable, then $\operatorname{Ho}(\calC)$ (the ordinary category with the same objects, but $\Hom_{\text{Ho}(\calC)}(A,B)=\pi_0\operatorname{Map}(A,B))$ is canonically triangualted.
-\end{rmk}
-
-You just define $\Sigma A$ to be the pushout of $0\leftarrow A\to 0$ and $\Omega A$ is defined dually.
-Here distinguished triangles are the ones that come from just doing what you'd hope: compute the cofibration of the
-map $A\to B$ to get $C=B/A$ and then $C/B$ will be the same as $\Sigma A$.
-
-Then $\calC$ is stable if and only if it has finite limits, colimits, 0 and $\Omega=\Sigma^{-1}$.
-
-\begin{defn}
-	A t-structure on a stable infinity category is jsut a t-structure on Ho$(\calC)$.
-\end{defn}
-\begin{rmk}
-	If $\calC$ is a stable infinity category, then $\calC^\heartsuit$ is already an abelian (1-)category.
-\end{rmk}
-
-\subsection{Alebraic K theory and t-structures}
-K-theory $K:\mathbf{Cat}_\infty^{st}\to\mathbf{Sp}$ is a functor from stable infinity categories with exact functors (ones that preserve finite (co)limits) to spectra (one can 
-also think graded groups, etc.)
-
-There are lots of places that algebraic and geometric objects give rise to stable $\infty$ categories (e.g. $X$, $\operatorname{Perf}X$, $\Db(X)$)
-
-\begin{defn}
-	$K$ is the unit of the convolutional symmetric monoidal structure on the subcollection of $\mathbf{Func}(\mathbf{Cat}_\infty^{st},\mathbf{Sp})$ that are localizing.
-
-	Furthermore, notice that $\mathbf{Cat}_\infty^{st}\hookrightarrow \mathbf{Func}^{loc}(\mathbf{Cat}_\infty^{st},\mathbf{Sp})$ via the Yoneda embedding and K-theory 
-	is corepresented by $\operatorname{Perf}(S)$ (lost me here).
-\end{defn}
-
-Here a functor $F:\mathbf{Cat}_\infty^{st}\to\mathbf{Sp}$ is localizing if, given a Verdier localization sequence 
-\[\calA\hookrightarrow\calB\twoheadrightarrow\calC\cong(\calB/\calA)^{idem}\]
-then $F(\calA)\to F(\calB)\to F(\calC)$ is an exact triangle of spectra (and $F$ preserves filtered colimits).
-
-\begin{thm}[Schlichting]
-	If $\calA$ is a small abelian category, then $K_{-1}\calA=0$ and if $\calA$ is Noetherican, then $K_{-n}\calA=0$ for all $n>0.$
-\end{thm}
-
-\subsection{Results of A-G-H}
-\begin{thm}
-	If $\calC$ is a small, stable $\infty$-category with bounded t-structure, then $K_{-1}(\calC)=0$.
-\end{thm}
-\begin{thm}
-	If $\calC$ is small, stable $\infty$-category with bounded t-structure with Noetherian heart, then $K_{-n}(\calC)=0$ for all $n>0$.
-\end{thm}
-\begin{thm}[Nonconnective theorem of the Noetherian heart]
-	IF $\calC$ is a stable infinty category with bounded t-structure such that $\calC^\heartsuit$ is Noetherian, then $K(\calC^\heartsuit)\simeq K(\calC)$.
-\end{thm}	
-\begin{thm}
-	Let $R$ be a commutative ring and $A$ a (cohomologically) graded dg $R$-algebra such that $H^0(A)$ is semisimple and $H^i(A)$ is finite generated as a right $H^0(A)$-module,
-	then $K_{-n}(A)=0$ for all $n>0$.
-\end{thm}
-\begin{thm}
-	If $i:\calA\to \calB$ is a fully fairthful exact functor of small stable infinity categories equipped with compatible bounded t-structures. Then set $\calC$ to be the Verdier quotient $\calB/\calA$,
-	and $\operatorname{Ind}(\calC)$ inherits a t-structure which restricts to a bounded t-structure on $\calC$ if and only if $\calA^\heartsuit\hookrightarrow \calB^\heartsuit$ is the inclusion of a Serre subcategory.
-\end{thm}
-
-A major question/conjecture: Given an abelian category $\calA$, can one construct an abelian category $\calB$ containing $\calA$ as a Serre subcategory such that $K(\calB)=0$?
-
-A positive answer to this would give tools necessary to answer the conjectures given in the affirmative.
-
-\section{Local duality for Gorenstein algebras}
-This talk was given by Henning Krause on recent work by Benson, Iyengar, and Pevtsova.
-
-Much of Julia's talk sets the scene for this talk! Let $G$ be a finite group, let $k$ be a field. Then if $M$ is a $kG$-module, 
-we can compute $H^\ast(G,M)=\Ext^\ast_{kG}(k,M)$, which is a graded module over $R:=H^\ast(G,k)$ and this is a finitely generated algebra.
-
-Pick any $\p\in\Spec R$ and let $I(\p)$ be an injective envelope of $R/\p$. Then the theorem says 
-\begin{thm}[BIKP, '19]
-	Let $d$ be the Krull dimension of $R/\p$ where $\p\lhd R$ is prime and is not all of $R^{\ge 0}$. Then 
-	\[\Hom_R(H^{\ast-d-i}(G,M),I(\p))\cong\widehat\Ext_{kG}^i(M,\Gamma_\p(k))\]
-	for all $i\in\bbZ$.
-\end{thm}
-
-\subsection{Serre Duality}
-\begin{thm}
-Let $X$ be a projective nonsingular scheme over a field $k$, and let $n=\dim X$. Then 
-\[\Hom_k(H^{n-i}(X,\calF),k)\cong\Ext_X^i(\calF,\omega_X)\]
-for all coherent sheaves $\calF$ on $X$ and $i\ge 0$.
-\end{thm}
-
-A more modern formulation is that $\Db(\operatorname{coh}X)$ has a Serre functor, $F=-\otimes_X^L\omega_X[n]$.
-
-\begin{defn}[Bendal-Kapraha, '89]
-	Let $\calT$ be a $k$-linear triangular category, hom-finite (what does this mean?) $\calF:\calT\to\calT$ is a \textbf{Serre functor}
-	if $D\Hom_\calT(X,Y)\cong \Hom_\calT(Y,\calF(X))$ which is natural for all $X,Y\in\calT$. (here he wrote $D=\Hom_k(-,k)$).
-\end{defn}
-
-Let $T\in\operatorname{coh}X$ be a tilding object and let $\Lambda=\End_X(T)$. Then we have an equivalence 
-\[R\Hom_X(T,-):\Db(\operatorname{coh}X)\xrightarrow{\sim}\Db(\mathbf{mod}\Lambda)\]
-That is, $\Db(\mathbf{mod}\Lambda)$ has a Serre functor.
-
-\begin{thm}[Reiten-van den Bergh '01, Happel '87]
-	Let $\Lambda$ be a finite dimensional $k$ algebra. Then the following are equivalent:
-	\begin{itemize}
-		\item $\Db(\mathbf{mod}\Lambda)$ has a Serre functor
-		\item $\Db(\mathbf{mod}\Lambda)$ has Auslander-Reiten triangles
-		\item $\gldim \Lambda<\infty$.
-	\end{itemize}
-\end{thm}
-\subsection{Gorenstein Algebras}
-\begin{defn}
-	Let $\Lambda$ be a finite dimensional $k$-algebra. Then $\Lambda$ is (Iwanaga) Gorenstein if $\injdim(\Lambda_\Lambda)$ and $\injdim({_\Lambda}\Lambda)$ are both finite.
-\end{defn}
-Some examples of such algebras:
-\begin{itemize}
-	\item If $\gldim \Lambda<\infty$, $\Lambda$ is Gorenstein.
-	\item If $\Lambda$ is self-injective, $\Lambda$ is Gorenstein.
-	\item If $\Lambda$ and $\Gamma$ are both Gorenstein, then $\Lambda\otimes_k\Gamma$ is Gorenstein. For instance the path algebra of a quiver $kQ$ tensored with $k[\varepsilon]$ the ``dual numbers''.
-	The former (usually) has Gorenstein dimension 1 and the latter has Gorenstein dimension 0 (it is self-injective).
-\end{itemize}
-
-Now let $\mathbf{Mod}\Lambda\supset \mathbf{mod}\Lambda\supset \mathbf{proj}\Lambda$ be the $\Lambda$ modules, finitely generated $\Lambda$ modules and finitely-generated projectives, respectively.
-
-Let $\mathbf{GProj}\Lambda=\{X\in\mathbf{Mod}\Lambda|\Ext^i(X,\Lambda)=0\text{ for }i>0\}$ and let $\mathbf{Gproj}\Lambda=\mathbf{GProj}\Lambda\cap\mathbf{mod}\Lambda$.
-
-Finally $\uHom_\Lambda(X,Y)=\Hom_\Lambda(X,Y)/P\Hom_\Lambda(X,Y)$ as usual.
-
-Notice that if $\gldim\Lambda$ is finite, then $\mathbf{GProj\Lambda}=\mathbf{Proj}\Lambda$ and if $\Lambda$ is self-injective, 
-$\mathbf{GProj}\Lambda=\mathbf{Mod}\Lambda$, giving us (in a way) two extremes.
-
-\begin{lem}
-	$\mathbf{GProj}\Lambda$ form a Froebenus category and $\underline{\mathbf{GProj}}\Lambda$ is a compactly generated trianglulated category equivalent to the subcategory of compact objects.
-\end{lem}
-
-Buchwertz and Orlov ('87 and '04) did work to show that $\underline{\mathbf{Gproj}}\Lambda$ was equivalent to the singularity category of finitely generated $\Lambda$ modules.
-
-We have a Nakayama functor $\nu:-\otimes_\Lambda D(\Lambda):\mathbf{mod}\Lambda\to \mathbf{mod}\Lambda$ (it takes projectives to injectives).
-
-\subsection{Auslander-Reiten Duality}
-We have a diagram 
-\begin{center}
-	\begin{tikzcd}
-		\Db(\mathbf{proj}\Lambda)\ar[r,hookrightarrow]\ar[d] & \Db(\mathbf{mod}\Lambda)\ar[d,"\sim"]\ar[r,two heads,"\nu"] & \Dsing(\lambda)\ar[d,"\bar\nu"]\\
-		\Db(\mathbf{proj}\Lambda)\ar[r,hookrightarrow] & \Db(\mathbf{mod}\Lambda)\ar[r,two heads] & \Dsing(\Lambda)
-	\end{tikzcd}
-\end{center}
-\begin{thm}
-	$\nu$ is a Serre functor for $\Db(\mathbf{proj}\Lambda)$ and $\Sigma^{-1}\circ\bar\nu$ is aSerre functor for $\Dsing(\Lambda)$.
-\end{thm}
-\begin{rmk}
-	$\calF:\underline{\mathbf{Gproj}}\Lambda\xrightarrow{\sim}\underline{\mathbf{Gproj}}\Lambda$
-	is an autoequivalence where $\calF$ sends 
-	\[X\mapsto \Omega^{-1}\text{GP}(D\operatorname{Tr}X)\]
-	where GP is the Gorenstein projective approximation functio and $D$Tr is the dual of the transpose. This is suppsoed to explain 
-	how this is a statement of duality.
-\end{rmk}
-
-\subsection{Hochschild cohomology}
-$\HH^\ast(\Lambda)=\Ext^\ast_{\Lambda^e)}(\Lambda,\Lambda)$ is a graded commutative ring.
-Then $\uHom_\Lambda^\ast(X,Y)=\oplus\uHom_\Lambda(X,\Omega^{-i}Y)$ where $X$ and $Y$ are Gorenstein projectives. Then $\HH^\ast(\Lambda)$ acts on $\underline{\mathbf{GProj}}$ 
-via 
-\[\varphi_X:\HH^\ast(\Lambda)\xrightarrow{-\otimes_\Lambda X}\Ext^\ast_\Lambda(X,X)\xrightarrow{tx\to px}\uHom^\ast(X,X)\]
-where I have no idea what that last map was.
-
-Thus $\uHom^\ast(X,Y)$ are graded $\HH^\ast(\Lambda)$-modules for all $X$ and $Y$ in $\underline{\mathbf{GProj}}\Lambda$. As an assumption, 
-fix $R\subseteq\HH^\ast(\Lambda)$ to be a homogeneous $k$ subalgebra such that
-\begin{itemize}
-	\item $R$ is connected ($R^0=k$)
-	\item $R$ is a finitely generated $k$ algebra.
-\end{itemize}
-Then for $\p$ be a homogeneous prime in $R$ and we get an endofunctor $\Gamma_\p$ on $\underline{\mathbf{GProj}}\Lambda$ that extracts the $\p$-local $\p$-torsion.
-\begin{thm}[BIKP, '19]
-	For $X,Y\in\mathbf{GProj}\Lambda$ where $X$ is finite dimensional and $\p$ a homogeneous prime with $R/\p$ of Krull dimension $d$,
-	\[\Hom_R(\Ext_\Lambda^\ast(X,Y),I(\p))\cong\uHom_\Lambda(Y,\Omega^d\,\Gamma_p \,\text{GP}\,\nu(X))\]
-\end{thm}
-\begin{prf}
-	Passage to closed points via $k\subseteq K$, similar to what we saw in Julia's talk.
-\end{prf}
-
-\subsection{Local duality for \texorpdfstring{$\Db(\mathbf{mod}\Lambda)$}{Db(mod Lambda)}}
-Let $\calD:=\Db(\mathbf{mod}\Lambda)$ where $\Lambda$ is Gorenstein, $\p$ is homogenous prime in $R\subseteq\HH^\ast(\Lambda)$. Let $\calD_\p$ be the $\p$-localization
-of $\calD$, whre the objects are the same and $\Hom_{\calD_\p}^\ast(X,Y)=\Hom^\ast_\calD(X,Y)_\p$.
-
-Then the $\p$-localization functor is exact.
-
-Now let
-\[\gamma_\p(\calD):=\{X\in\calD_\p|\End^\ast_{\calD_\p}(X)\text{ is $\p$-torsion}\}\]
-which is a thick subcategory of $\calD_\p$.
-\begin{rmk}
-	if $\calT$ is a compactly generated $R$-linear triangualted category. Then 
-	\[\gamma_\p(\calT)\xrightarrow{\sim}(\Gamma_\p\calT)^c\]
-	and we get from $\nu$ a local Nakayama endofunctor $\nu_\p$ on $\gamma_\p(\calD)$.
-\end{rmk}
-\begin{thm}
-	$\Sigma^{-d}\nu_\p$ is a Serr functor for $\gamma_\p(\calD)$ and $\Hom_{R_\p}(-,I(\p))$ plays the role of duality.
-\end{thm}
-
-For $\m= R^{>0}$ the maximal ideal, $\Db(\mathbf{proj}\Gamma)\subseteq\gamma_\m(\calD)$ with equaility if (FG) holds.
-
-Thus we get Serre duality for $\Db(\mathbf{proj}\Lambda)$ via $\nu_\p$.
+\section{Statistical Learning}
+This series of lectures was given by Robert Shapire from Microsoft Research.
 

### 2019-07-29 MSRI ML Notes/MSRI_ML_notes.tex
End of day 1

@@
 
+The topic of the talks here are somewhere at hte nexus of supervised and statistical learning. The big idea concerning what we want o do is 
+to learn how to do better in the future based on experience from the past. The hope is that these methods can be fully automated. This is a proper subset of AI, 
+but definitely a core area of how we do it.
+
+We are looking for easy to use, flexible algorithms that give us useful results. Ideally they would be interpretable, but 
+this is usually a secondary result.
+
+He believes that ML can help us to understand how learning happens in non-machine entities (cool!). For instance, one can consider the question of 
+nature vs nuture and to ask the question ``what is simplicity and why is it useful?''
+
+\subsection{The Problem}
+In this course, we will be focusing on a single kind of (simple) learning problem: learning from example. Given a set of examples, 
+we are looking for an algorithm that can classify instances given objects.
+
+For instance, character recognition. We have a set of handwritten characters and the letter it depicts. Then we want to feed this 
+into a learning algorithm that gives us a prediction rule, that given a handwritten character, outputs the predicted character.
+
+We were then given a couple of examples to try. The particulars weren't important, but the takeaway here is 
+\begin{itemize}
+	\item We needed enough data to say something meaningful.
+	\item We looked for a rule that fit the observations. We want the rule to be consistent or to contain few mistakes.
+	\item We were looking for rules that were ``simple.'' Importantly, our notion of simplicity changed when we converted one set of numbers to binary.
+\end{itemize}
+The first condition is always the case: ``more is more.'' But the latter two fit into a tradeoff between fit and simplicity.
+
+The main questions we will be trying to answer:
+\begin{itemize}
+	\item How much data do we need?
+	\item How do we define simplicity?
+	\item How much complexity do we need to represent our problem?
+	\item What is the tradeoff mentioned above?
+\end{itemize}
+
+\subsection{Studying the Problem Mathematically}
+We have to come up with a formal model for the problem so that we can do mathematics with it. The learning model should answer some basic questions:
+\textit{What is the goal of learning?} and \textit{How is the learning happening?}.
+
+We begin the description of the model:
+\begin{defn}
+	An \textbf{instance} (or sometimes informally \textbf{example}) $x$ is an object in a space called the \textbf{instance space} or \textbf{domain} $X$.
+\end{defn}
+\begin{rmk}
+	To each instance we assign a \textbf{label} or \textbf{class}. In this course we will be simplifying to only consider two classes, denoted $0$ and $1$ or $+$ and $-$.
+\end{rmk}
+\begin{rmk}
+	We work under an assumption that there is an underlying function 
+	\[c:X\to\{0,1\}\]
+	called the (target) \textbf{concept} that we are trying to learn. Later we will relax this assumption.
+\end{rmk}
+\begin{defn}
+	We define the \textbf{hypothesis} to be (similar to the concept) a map 
+	\[h:X\to\{0,1\}\]
+	that holds our current beliefs.
+\end{defn}
+We write $\calC$ to denote the \textbf{concept class}, the class of all possible functions we are exploring (we assume we know a class to which $c$ belongs). We will further 
+assume that our data were generated \textit{independently} and at random. It is limiting to make an assumption about the distribution from which these were drawn. So we will just say there is one.
+
+\begin{defn}
+	We define 
+	\[\operatorname{err}_D(h)=\Pr_{x\sim D}[h(x)\ne c(x)]\]
+	to be the (generalization) \textbf{error}.
+\end{defn}
+
+Then we obviously ask that $\operatorname{err}_D(h)<\varepsilon$ for some small $\varepsilon$, but furthermore we want that 
+this happens with high probability:
+\[\Pr[\operatorname{err}_D(h)\le\varepsilon]\ge 1-\delta\]
+and in this case, we say that that the rule is probably approximately correct  (PAC).
+\begin{defn}
+	The class $\calC$ is \textbf{PAC-learnable} by $\calH$ (a hypothesis space) if there is an algorithm $A$ such that
+	for all $c\in\calC$ and for all distributions $D$ and for all positive $\varepsilon$ and $\delta,$ then $A$ takes ``polynomial many''
+	random variables $X_i\sim D$ which outputs $h\in\calH$ such that
+	\[\Pr[\operatorname{err}_D(h)\le\varepsilon]\ge 1-\delta\]
+\end{defn}
+\begin{rmk}
+	This is not an easy problem, per se, but it is one of the most simple classes of algorithms to work with.
+\end{rmk}
+\begin{ex}
+	Let $X=\bbR$ and let $\calC$ be of the form $c_r$ for $r\in\bbR$ where 
+	\[c(x)=\left\{\begin{array}{lr} 1, & x\ge r\\
+	 0 & \text{otherwise}\end{array}\right.\]
+	 so $\calC$ is the \textit{set of positive half lines}. Then if we have some data, we can set our hypothesis 
+	 $h=c_b$ where $b$ is the leftmost positive value. Then if $c=c_r$, then $\operatorname{err}_D(h)$ has to do with the interval $[r,b]$. Let $b-r=\varepsilon$.
+
+	 Now restrict attention to $X=[0,1]$ and notice that if we got a training point within $\varepsilon$ of $c$ (to the right), then we would have gotten a smaller difference. Thus
+	 \begin{align*}
+		 \Pr[\operatorname{err}_D(h)<\varepsilon] &\le \Pr[\text{no $x_i$ are in $[b,b+\varepsilon]$}]\\
+		 &=\prod_1^m\Pr[x_i\in R]\\
+		 &=(1-\varepsilon)^m\\
+		 &\le e^{-\varepsilon m}
+	 \end{align*}
+	 using the fact that $1+x\le e^x$ for all $x$.
+\end{ex}
+
+\subsection{Sufficient conditions for learning}
+Now we consider the case when our hypothesis space $\calH$ is finite.
+\begin{thm}
+	Let $A$ be an algorithm that finds a hypothesis $h_A\in\calH$ which is consistent with $m$ random training examples (as before)
+	where
+	\[m\ge \frac{1}{\varepsilon}(\ln|\calH|+\ln\frac{1}{\delta})\]
+	then $\Pr[\operatorname{err}_D(h_A)>\varepsilon]\le \delta$.
+\end{thm}
+\begin{rmk}
+	Note that we have completely done away with trying to get a handle on the concept class here.
+\end{rmk}
+\begin{rmk}
+	Equivalently, with probability of $1-\delta$, 
+	\[\operatorname{err}_D(h_A)\le\frac{\ln|\calH|+\ln|\frac{1}{\delta}|}{m}\]
+\end{rmk}
+The idea to take away here is that $\ln|\calH|$ is in some way measuring ``description length'' of a hypothesis: you have to describe how to distinguish the correct one from the wrong ones.
+Think popping these
+in a binary tree. Then the $m$ bound gives you some sort of way to manage the complexity of the algorithm.
+
+\begin{thm}
+	Assume we have $m\ge \frac{1}{\varepsilon}(\ln|\calH|+\ln|\frac{1}{\delta}|)$ examples.
+Then with probability $1\ge 1-\delta$, for all $h\in\calH$ if $h$ is consistent then $\operatorname{err}_D(h)\le \varepsilon$ (here we often write ``$h$ is $\varepsilon$-good.'')
+\end{thm}
+\begin{prf}
+	We prove the converse: the probability of there being an $h\in\calH$ that is consistent and $\varepsilon$-bad is less than $\delta.$
+	Notice that whether $h$ is consistent with the training set is a random variable (since it depends on choice of training set). Whether it is $\varepsilon$-bad is not random. So 
+	if $\calB$ is the set of $\varepsilon$-bad hypotheses, then we really want 
+	\[\Pr[\exists h\in\calB: h\text{ is consistent}]\le \sum_{g\in\calB}\Pr[h\text{ is consistent}]\]
+using the union bound.
+
+Now fix $g\in\calB$ and we want to compute the probability of it being consistent. That is, $h(x_i)=c(x_i)$ for all $i$. Since the samples are independent, this probability is 
+\[\prod_1^m\Pr[h(x_i)=c(x_i)]\le (1-\varepsilon)^n\]
+
+Therefore the original probability is 
+\[|\calB|(1-\varepsilon)^m\le|\calH|e^{-\varepsilon m}\le \delta\]
+\end{prf}
+
+\subsection{When does this work when we have infinite hypothesis classes?}
+We saw in the first example that infinite classes can still be PAC-learnable, but where is the line drawn?
+
+To see this, consider classifications of finite subsets. But any choice of threshold between any two points on $\bbR$ will give the same classification! So really we only have 
+something like 5 different hypothesis (in the form of a threshold function at least). Compare that with $2^m$ behaviors one could consider in general!
+
+More formally, if we are given a set $S=\langle x_1,\dots,x_m\rangle$ of instances, then we can consider the collection of all $\langle h(x_1),\dots, h(x_m)\rangle$ for all $h\in\calH$.
+We call this set
+$\Pi_\calH(S)$ and define 
+\[\Pi_\calH(m)=\max_{|S|=m}|\Pi_\calH(S)|\]
+and call this the \textbf{growth function.} We can use this function to get some sort of handle on the complexity of the problem, 
+and in fact it plays a similar role in an analogous theorem to the one we saw earlier:
+\begin{prop}
+	Give $m$ training examples, with probability $1-\delta$ for all $h\in\calH$ if $h$ is consistent then 
+	\[\operatorname{err}_D(h)\le\calO\left(\frac{\ln\Pi_\calH(2m)+\ln\frac{1}{\delta}}{m}\right)\]
+\end{prop}
+\begin{rmk}
+	The proof of this result can't quite continue in the same way as before because we need to select a training set \textit{before we can know the hypothesis space}, 
+	so the dependence gets in the way. There is a fancier proof but we won't do it here (although there will be another proof later that is similar).
+\end{rmk}
+\begin{rmk}
+	Notice that if $\Pi_\calH(m)$ is polynomial, this reduces nicely and as $m$ gets larger, the error drops. Thus learning is possible.
+\end{rmk}
+
+\subsubsection{More on the growth function}
+There are at least some cases when $\Pi_\calH(m)\in\calO(m^d)$ for some constant $d$. In the worst case, $\Pi_\calH(m)=2^m$. In fact, these are the only two cases that can happen! Furthermore,
+it ends up that these two cases correspond exactly to when learning is possible!
+
+To see why, we need a new definition:
+\begin{defn}
+	A sample $S$ of size $m$ is \textbf{shattered} by $\calH$ if all possible behaviors/labelings are possible: that is, $\Pi_\calH(S)=2^m$.
+\end{defn}
+\begin{ex}
+	Given $\calH$ to be the set of all closed intervals $[a,b]\subseteq\bbR$, any set of two points in $\bbR$ is shattered by $\calH$.
+	But you cannot shatter a set of three points in $\bbR$.
+\end{ex}
+\begin{defn}
+	The \textbf{Vapnik-Chervinenkis (or VC) dimension} is
+	\[\operatorname{VCdim}(\calH)=\max\{|S|: S\text{ is shattered by }\calH\}\]
+\end{defn}
+\begin{rmk}
+	So in the exercise above, $\operatorname{VCdim}(\text{intervals})=2$.
+\end{rmk}
+Another set of hypotheses are the linear threshold functions in $\bbR^n$, and the VC dimension here is $n+1$. If you fix your planes to be subspaces, the VC dimension is $n$.
+
+\begin{rmk}
+	A good heuristic measure for VC dimension is the number of parameters. There are pathological examples, but they are truly that.
+\end{rmk}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Convex Optimization}
+This series was given by Sebastien Bubeck from Microsoft Research.
+
+\subsection{Fundamentals}
+Let $K\subseteq \bbR^n$ be a convex set: that is one such that the straight line segment between two points in $K$ is contained entirely in $K$.
+\begin{defn}
+	The map $f:K\to\bbR$ is called \textbf{a convex function} if 
+	\[f((1-\gamma)x+\gamma y)\le (1-\gamma)f(x)+\gamma f(y).\]
+\end{defn}
+\begin{rmk}
+	Then if $f$ is differentiable, this implies
+	\[\frac{f(x+\gamma(y-x))-f(x)}{\gamma}\le f(y)-f(x)\]
+	and as $\gamma\to 0$, the LHS tends to $\nabla f(x)\cdot (y-x)$. Thus
+	\[f(y)\ge f(x)+\nabla f(x)\cdot(y-x).\]
+\end{rmk}
+
+Then the goal of this course is to ``find'' $\operatorname{argmin}_{x\in K}f(x)$. A question we ask ourselves: how are $f$ and $K$ specified? Mostly in these lectures we will be 
+focusing on \textit{simple} $K$ and functions $f$ such that $\nabla f(x)$ can be computed (for any $x$).
+
+\subsection{Examples in machine learning}
+\subsubsection{Regression}
+Here we have a dataset: $(a_i,y_i)\in\bbR^n\times\bbR$ are the elements. Then we want to find a rule that assigns $y$ to $\mathbf a$. We focus on \textit{linear} rules.
+That is, we are looking for a linear space in $\bbR^{n+1}$ that approximates the ``true'' values.
+
+Tentatively we let our function be $a\mapsto x\cdot a$, parameterized by $x\in\bbR^n$. Then for each $x$ we can evaluate how the tentative function
+fits our data. More specifically:
+\[\frac{1}{m}\sum_1^m l(x\cdot a_i,y_i)\]
+where $l$ is some loss function.
+
+There are several loss functions one might consider:
+
+\noindent\textbf{Least Squares}: 
+\[l(u,v)=(u-v)^2\] 
+The upshot here, however, is that the evaluation function above in this case is convex!
+The idea here is that we are applying a linear function to a convex function. The other nice thing is that if you make a modeling assumption about the 
+distribution underlying the label distribution, nice things happen.
+
+Specifically, assume that $y\sim\calN(a\cdot \underline x,\sigma^2)$ where $\underline x$ is the \textit{true value}.
+Notice we haven't put any assumption on the draw distribution, just the relationship between the (fixed) $a_i$ and $y$ given the truth $x$. One can compute 
+\[\frac{1}{2\pi\sigma^2}\exp(-\frac{1}{2\sigma}(y_i-a_i\cdot x)^2).\]
+
+Thus the likelihood of the entire data set is the product:
+\[\frac{1}{(2\pi\sigma^2)^{m/2}}\exp(-\frac{1}{2\sigma^2}\sum_1^m(y_i-a_i\cdot x)^2)\]
+and so we recover the least squares loss function as the objective we want to minimize. That is, the maximum likelihood estimator minimizes least squares fit.
+
+\subsubsection{Classification}
+Now restrict $y\in\{\pm 1\}$. A natural loss is 
+\[l(u,v)=\delta_{\operatorname{sign}(u)\ne\operatorname{sign}(v)}.\]
+This is a problem for this class since it is non-convex.
+
+So instead we define \textbf{support vector machines} using the loss 
+\[l(u,v)=\max(0,1-uv)\]
+which is convex
+
+Another example is the \textbf{logistic loss}
+\[l(u,v)=\log(1-e^{-uv})\]
+which you can see is a smooth upper bound on the 0-1 loss we started with. Minimizing this function is equivalent to minimizing the MLE for the \textit{logistic model:}
+
+Denote $p(a)=\bbP(+1|a)$. Then we assume that the ``scale'' of our probability is given by a linear function:
+\[\log\left(\frac{p(a)}{1-p(a)}\right)=\underline x\cdot a\Leftrightarrow p(a)=\frac{1}{1+e^{-\underline{x}\cdot a}}\]
+and then the \textbf{logistic loss} is the negative log-likelihood of the logistic model.
+
+\subsubsection{Graphical Models}
+Given $a_1,\dots,a_m\in\bbR^n$, we want to infer the correlation structure of the variables (how are they all related).
+
+Let's start from a (Gaussian) model: we assume we are drawing IID samples from $\calN(\mu,\underline\Sigma)$.
+A fact: if $(\Sigma^{-1})_{ij}=0$, then $x_i$ and $x_j$ are independent, conditioned on the rest.
+Thus the goal is to estimate $\Sigma^{-1}$. So if the truth is $\Sigma$, the density of $a_1,\dots,a_n$ is 
+\[\frac{1}{\sqrt{(2\pi)^n\det(\Sigma)}}\exp(-\frac{1}{2}(a_i-\mu)^T\Sigma^{-1}(a_i-\mu))\]
+and by taking the negative log likelihood:
+\[c+\frac{m}{2}\log\det\Sigma+\frac{1}{2}\sum_1^m(a_i-\mu)^T\Sigma^{-1}(a_i-\mu)=C+\frac{m}{2}\left(-\log\det\Sigma^{-1}+\operatorname{tr}(\Sigma^{-1}S)\right)\]
+where $c$ is some constant and $S$ is the sample covariance matrix. Then to estimate $\Sigma^{-1}$ via MLE, we want to solve 
+\[\operatorname{argmin}_{X>0}-\log\det(X)+\operatorname{tr}(XS)\]
+where this is convex in $X$ (this is not obvious but we will see it in the problem session).
+
+If you know that the true matrix is sparse, you may add a penalty term $\lambda \|X\|_1$. Similarly in regression you may add either $\|x\|_1$ or $\|x\|_2^2$.
+In either case, these add curvature to your objective function. This may speed up the optimization process.
+
+\subsubsection{Unsupervised Learning}
+Finally we want to talk about \textit{clustering}.
+Notice that we have been slowly changing our dataset.
+First we dropped labels and now we are going to represent our dataset as a graph $G=(V,E)$ of interactions.
+We would like to infer some partitioning of the vertices.
+
+As a modelling assumption, we use the stochastic block model. There is a hidden partitioning $\sigma\in \{-1,1\}^|V|.$ $G$ is generated from the distribution
+\[(i,j)\in E\text{ with probability }\left\{\begin{array}{lr}p & \sigma_i=\sigma j\\ q, \text{otherwise}\end{array}\right. \]
+where we assume $q<p$. Furthermore we assume all connections are independent.
+
+The the likelihood of $x\in\{-1,1\}^n$ is given as follows: let $A=(a_{ij})$ be the adjacency matrix of $G$. Then the likelihood is 
+\[\prod_{ij}\left[\frac{1+x_ix_j}{2}\left(pa_{ij}+(1-p)(1-a_{ij})+\frac{1-x_ix_j}{2}(qa_{ij}(1-q)(1-a_{ij}))\right)\right]\]
+this is non-convex! But we can look at a convex upper bound and we're good! Next time we'll talk about gradient descent.
+
 \end{document}

### 2019-07-30 MSRI ML Notes/MSRI_ML_notes.tex
Day 2

@@
 % Document-Specific Macros
+\DeclareMathOperator{\err}{err}
+\DeclareMathOperator{\argmin}{argmin}
 
@@
 
+We have a nice little lemma:
+\begin{lem}[Sauer]
+	If $d=\operatorname{VCdim}(\calH)$, then 
+	\[\Pi_\calH(m)\le\sum_0^d\binom{m}{i}.\]
+\end{lem}
+\begin{rmk}
+	A ``nicer'' upper bound leverages
+	\[\sum\binom{m}{i}\le\left(\frac{em}{d}\right)^d\]
+	whenever $m\ge d\ge 1$.
+\end{rmk}
+
+Plugging this into the bound on the error of a hypothesis, we get that (forgetting the relatively small log term) that the VC dimension $d$ gives us
+a complexity measure---that is, it gives a bound on the error of a consistent hypothesis.
+
+\subsection{Non-consistent Hypotheses}
+The results so far have been great, but has always assumed that there is a consistent hypothesis to find. But what if our data is noisy? Then 
+learning the concept is no longer about finding the concept that matches \textit{all} the given data.
+
+Now we consider data $(x,y)\sim D$ to be a pair jointly sampled from some distribution $D$. We do basically the same 
+thing to compute error:
+\[\err_D(h)=\Pr_{(x,y)\sim D}[h(x)\ne y].\]
+We are still trying to minimize the generalization error $\err(h)$ over $\calH$ given some sample $(x_i,y_i)$ sampled from $D$.
+\begin{defn}
+	Write the \textbf{empirical error} as 
+	\[\widehat\err(h)=\frac{1}{m}\sum_1^m\delta_{h(x_i)\ne y_i}\]
+\end{defn}
+And then we can follow the route called \textit{empirical risk minimization} (ERM) that finds an empirical solution:
+\[\hat h=\argmin_{h\in\calH}\widehat\err(h).\]
+
+We want to get to the situation in which, with probability greater than $1-\delta$, for all $h\in\calH$,
+\[|\err(h)=\widehat\err(h)|\le\varepsilon.\]
+This result is a kind of \textbf{uniform convergence result} which is the kind of thing we are after. Obviously this is useful since it allows us 
+to bridge the gap between imperfect information and the true concept.
+
+Now $\delta_{h(x_i)\ne y_i}=1$ with probability $\err(h)$ and zero otherwise.
+
+\begin{ex}
+	A quick aside: Consider IID variables $Z_1,\dots,Z_m$ with $Z_i\in[0,1]$. Let $p=\bbE[Z_i]$ and then 
+	let $\hat p=\frac{1}{m}\sum_i Z_i$ be the empirical mean.
+
+	Then there is a result called \textbf{Hoeffding's inequality}, which claims
+	\[\Pr[\hat p\ge p+\varepsilon]\le e^{-2e\varepsilon^2m}.\]
+
+	Notice that what this says is that the empirical average converges to its expected value. So we can think of this 
+	as a function converging
+	\[f(Z_1,\dots,Z_m)\to\bbE[f(Z_1,\dots,Z_m)].\]
+\end{ex}
+This idea leads us to:
+\begin{lem}[McDiarmid's Inequality]
+	Suppose $f(z_1,\dots,z_m)$ is real valued such that chainging $z_i$ changes $f$ by at most $c_i$, and furthermore the $z_i$ are independent, but not necessarily identically distributed.
+	Then you get a nice bound on the error (I didn't get it in time but I am sure it is online.)
+\end{lem}
+
+So if we let $Z_i=\delta_{h(x_i)\ne y_i}$ and let $p=\bbE Z_i=\err(h)$ and $\hat p=\widehat\err(h)$, then for any particular $h\in\calH$,
+\[\Pr[|\err(h)-\widehat\err(h)|\ge \varepsilon]\le 2e^{-2\varepsilon^2 m}.\]
+So when $\calH$ is finite, we can use the union bound to say 
+\[\Pr[\exists h\in\calH: |\err(h)-\widehat\err(h)|\ge \varepsilon]\le 2|\calH|e^{-2\varepsilon^2 m}\]
+then setting the RHS equal to $\delta$,  we get that with probability $1-\delta$, for all $h\in\calH$,
+\[|\err(h)-\widehat\err(h)|\le\calO\left(\sqrt{\frac{\ln|\calH|+\ln\frac{1}{\delta}}{m}}\right)\]
+or in another (slightly weaker but sometimes more suggestive) form:
+\[\err(h)\le\widehat\err(h)+\calO\left(\sqrt{\frac{\ln|\calH|+\ln\frac{1}{\delta}}{m}}\right)\]
+\begin{rmk}
+	Notice that this really (finally) ties together how several aspects impact learning: it relates the true error to the sample size, the complexity class of the concept, 
+	and the consistency of the hypothesis with the sample data.
+\end{rmk}
+
+\begin{rmk}
+	Notice that as the complexity of your hypotheses increase, we expect the empirical error to decrease, but if we assume that 
+	the above is truly a good bound for $\err(h)$, we would expect that the true error will begin by dropping, but eventually the growth term from $\ln|\calH|$
+	will overtake things and cause the error to rise again.
+
+	This is one way to think about the concept of \textbf{overfitting}: the complexity of your hypothesis functions can incorporate more and more pathological functions that are less likely to be realistic.
+\end{rmk}
+
+\subsection{Another Complexity Measure}
+Suppose we have a labelled sample $S$ consisting of $(x_i,y_i)$, where we can now assume the label space is $\pm 1$.
+Then if we have a hypothesis, we can compute the training error $\widehat\err(h)$:
+\[\widehat\err(h)=\frac{1}{m}\sum_i\delta_{h(x_i)\ne y_i}=\frac{1}{m}\sum_i \frac{1-y_ih(x_i)}{2}=\frac{1}{2}-\frac{1}{2}\left(\frac{1}{m}\sum_i y_ih(x_i)\right)\]
+and the term in parentheses is another quantity we can consider which is just a constant away from training error.
+
+So we can consider computing 
+\[\bbE_{\mathbf\sigma}[\max_{h\in\calH}\sum_i\sigma_ih(x_i)]\]
+where $\sigma_i$ are $\pm 1$ equally likely. In the case that $|\calH|=1$, we get that this value is zero and when $\calH$ shatters $S$ (the set of all $m$-tuples of $\pm 1$), 
+the value is 1. These serve as the two extremes, obviously.
+
+We would like to work with this towards describing convergence (and rates thereof) towards the generalization error.
+To do this, we need to broaden the context slightly. Let $f:Z\to\bbR$ (standing in for the hypothesis), a space $\calF$ ($\calH$), and define the \textbf{(empirical) Rademacher complexity} to be 
+\[\hat R_\scrF(S)=\bbE_\sigma\left[\sup_{f\in\scrF}\frac{1}{m}\sum_i\sigma_if(z_i)\right]\]
+and the expected Rademacher complexity is 
+\[R_\scrF(m)=\bbE_S[\hat R_\scrF(S)]\]
+
+Then we want to show that for all $f\in\scrF$ that $\hat\bbE_S[f]\to\bbE[f]$. Specifically,
+\begin{thm}
+	Let $\scrF$ be a family of functions $f:Z\to[0,1]$ and let $S$ be a collection of $z_i$ where $z_i\sim D$. Then with probability $1-\delta$,
+	and for all $f\in\scrF$,
+	\[\bbE[f]\le\hat\bbE_S[f]+2\cdot\Psi+\calO\left(\sqrt{\frac{\ln(1/\delta)}{m}}\right)\]
+	where $\Psi$ is either $\hat R_\scrF(S)$ or $R_\scrF(m)$.
+\end{thm}
+\begin{prf}
+	Let $\Phi(S)=\sup_{f\in\scrF}(\bbE[f]=\bbE_S[f])$. Then first we show (via McDiarmid) that 
+	\[\Phi(S)\le \bbE_S[]\Phi(S)]+\calO\left(\sqrt{\frac{\ln 1/\delta}{m}}\right).\]
+	So then we have reduced the problem with finding some bound for $\bbE_S[\Phi(S)]$.
+
+	Then let $S'$ to be a hypothetical second sample (called the ``ghost sample''). Compute 
+	\[\bbE_S[\Phi(S)]=\bbE_S[\sup_f(\bbE[f]-\hat\bbE_S[f])]\le \bbE_{S,S'}[\sup(\hat\bbE_{S'}[f]-\hat\bbE_S[f])]\]
+	There is something to prove in the last line there but he didn't show us. He said there was a slide somewhere. :)
+
+	For the next step, go through the two samples $S$ and $S'$ one index at a time and for each $i$ with equal probability either swap $z_i$ and $z_i'$ or do nothing.
+	Call the resulting samples $T$ and $T'$. But then the following two expressions have the same distribution:
+	\[\hat\bbE_{S'}[f]-\hat\bbE_S[f]\quad\text{and}\quad \hat\bbE_{T'}[f]-\hat\bbE_T[f]\]
+	and so if we let $\sigma_i$ denote the random variable for whether we swapped at the $i^{th}$ step ($+1$ if we swapped $z_i$ and $z_i'$ and $-1$ otherwise),
+	the latter is 
+	\[\frac{1}{m}\sum_i\sigma_if(f(z_i')-f(z_i))\]
+	so we get 
+	\[\bbE_{S,S'}[\sup_f(\hat\bbE_{S'}-\hat\bbE_S[f])]=\bbE_{S,S'}\left[\sup_f\left(\frac{1}{m}\sum_i\sigma_if(f(z_i')-f(z_i))\right)\right]\]
+
+	Finally we can show (another slide) that this value is at most $2R_\scrF(m)$ and then use McDiarmid's once more to show that the expected Rademacher complexity is close enough to that of the empirical RC.
+\end{prf}
+
+Now the whole point was to show that $\err(h)\approx \widehat\err(h)$.
+One can show the empirical Rademacher complexity of the class of loss functions comprised of indicator functions $l(x,y)=\delta_{h(x)\ne y}$ is twice that of $\calH$ itself.
+You can easily plug this in to get $\bbE[l_h]$. Then you can get what you want.
+
+Now if $|\calH|<\infty$, (careful, he made a notation error earlier so I am switching subscripts and arguments in Rademacher complexities--they are still the same)
+\[\hat R_S(\calH)\le\sqrt{\frac{2\ln|\calH|}{m}}.\]
+
+If the space is infinite, then we choose a representative hypothesis for each behavior seen in applying all hypotheses to $S$. Call this space $\calH'$ and notice that $|\calH'|=|\Pi_\calH(S)|.$
+Then one can show that the empirical Rademacher complexity of $\calH'$ is the same as that of $\calH$!
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+\subsection{Gradient Descent}
+This is the most simple algorithm we will study (also one of the most used ones). It is an iterative method where we define 
+\[x_{t+1}=x_t-\eta\nabla f(x_t)\]
+which dates back to the mid 1800's (likely by Cauchy). Notice that this is similar to Newton's method, but the latter requires inverting an $n\times n$ matrix 
+and this was an attempt to avoid that difficult computation.
+
+Why does this work?
+By Taylor's theorem, $f(y)\approx f(x)+\nabla f(x)\cdot (y-x)$.
+The idea here is that the negative $\nabla f(x)$ direction is in the direction of the minimum, so we just follow that.
+
+Now if we have a convex function, where $f(x)-f(y)\le\nabla f(x)\cdot (x-y)$, we can set $y=x^\ast$ and define 
+\[\delta(x):=f(x)=f(x^\ast)\le\nabla f(x)\cdot (x^\ast-x)=-\nabla f(x)\cdot(x^\ast-x)\]
+then the idea is that progress in the $x^\ast-x$ direction (that is towards the target) is bounded below by $\delta(x)$.
+That is to say: \textit{the rate of decrease of the instance to OPT is lower bounded by the suboptimal gap.}
+
+There is a small problem, namely that we also take a (hopefully small) step away from the direct-line path from $x$ to $x^\ast$. To fix this, we
+use continuous time analysis:
+\[\frac{d}{dt}x(t)=-\nabla f(x(t))\]
+which we call the \textbf{gradient flow.} Then 
+\[\frac{d}{dt}\frac{1}{2}\|x(t)-x^\ast\|^2=\left(\frac{d}{dt}(x(t)-x^\ast)\right)\cdot(x(t)=x^\ast)=-\nabla f(x(t))\cdot (x(t)-x^\ast)\le -\delta(x(t))\]
+and integrating over time:
+\[\int_0^\tau \frac{1}{2}\|x(t)-x^\ast\|^2 \mathrm{d}t\le\int_0^\tau \delta(x(t))\mathrm{d}t\]
+and using the FTOC and dropping the $\tau$ term (and normalizing):
+\[\frac{1}{\tau}\int_0^\tau\delta(x(t))\mathrm dt\le \frac{1}{2\tau}\|x(0)-x^\ast\|^2:=\frac{r_0^2}{2\tau}\]
+and so finally
+\[\delta\left(\frac{1}{\tau}\int_0^\tau x(t) \mathrm{d}t\right)\le\frac{r_0^2}{2\tau}.\]
+
+Of course this isn't a discrete process, so it's useless for coding!
+So we want to run gradient descent with step size $\eta$ for $T$ steps and approximate this integral with a finite sum from $1$ to $T$.
+Then if we further impose that $f$ is Lipshitz ($\|\nabla f(x)\|\le L$) then the error in this approximation is at worst $\eta L^2$.
+So if we are trying to measure the total error of our discretized algorithm from the true value, we get an error bound 
+\[\frac{r_0^2}{T\eta}+\eta L^2\le \frac{r_0L}{\sqrt{T}}\]
+for a ``good'' choice of $\eta=\frac{r_0}{L\sqrt{T}}$.
+\subsubsection{Discrete Time Analysis}
+\begin{thm}
+	Let $g_1,\dots,g_T$ be arbitrary vectors in $\bbR^n$ and that $x_{t+1}=x_t-\eta g_t$. Then for any $x\in\bbR^n$,
+	\[\sum_1^Tg_t(x_t-x)\\le \frac{\|x_1-x\|^2}{2\eta}+\eta\sum_1^T\|g_T\|^2\]
+\end{thm}
+\begin{cor}
+	If $g_t=\nabla f(x_t)$ ($f$ convex) and $\|nabla f(x_t)\|\le L$, then
+	\[f\left(\frac{1}{T}\sum_1^Tx_t\right)-f(x^\ast)\le\frac{1}{T}\sum_1^T(f(x_t)-f(x^\ast))\le\frac{1}{T}\sum \nabla f(x_t)(x_t-x^\ast)\le\frac{r_n^2}{2\eta T}+\eta L^2\] 
+\end{cor}
+The proof here is really just one line of computation but I missed it.
+Interestingly, the proof gives us \textit{equality} rather than inequality.
+I suppose that it's nice to use it as if it were a bound.
+
+\subsubsection{Projective Gradient Descent}
+A generalization is optimization with some constraint.
+So if we have a bounded convex set $K$, we just follow each step in gradient descent with a projection onto the closest point in $K$ (should we leave it).
+This is only better for us since 
+\begin{lem}
+	$\|P_K(x)-z\|\le\| x-z\}$ for any $z\in K$.
+\end{lem}
+\begin{rmk}
+	Notice that we are assuming here that both $K$ and $f$ are convex. Without convexity we definitely don't have this.
+\end{rmk}
+
+\subsection{Stochastic Gradient Descent}
+We begin with a corollary of the theorem we just proved for standard GD:
+\begin{cor}
+	Say that $g_t$ is a random variable such that $\bbE[g_t|x_t]=\nabla x_t$ and that $\bbE[\|g_t\|^2]<B^2$.
+	Then 
+	\[\bbE[f\left(\frac{1}{T}\sum_1^T x_t\right)-f(x^\ast)]\le\frac{r_1^2}{2\eta}+\frac{\eta}{2}TB^2\le\frac{r_1B}{\sqrt{T}}\]
+	given an optimal choice of $\eta$.
+\end{cor}
+
+So then we can use this result to define SGD, Stochastic Gradient Descent. Let $f(x)=\bbE_{(a,y)}l(x,(a,y))$.
+The idea here is that we don't have the actual gradient of the loss function, but we have access (under the statistical learning framework) to a continuous stream of IID data.
+So we define the \textbf{stochastic gradient}:
+\[g_t=\nabla_x l(x_t,(a,t,y_t))\]
+that is, compute the $x$ gradient of $l(x,(a_t,y_t))$ and evaluate at $x_t$. Then basically this pos right into the above corollary.
+
+\begin{rmk}
+	here the notation varies from statistical learning, so think of $x$ as our hypothesis (the thing we're optimizing over) and $(a,y)$ denotes a sample from the dataset.
+\end{rmk}
+
+Note that here you only get one pass on the data (although here you can directly control the generalization error as opposed to bounding it with testing error in the other class).
+
+Instead, we can do multi-pass SGD where at each step we sample uniformly data points from your dataset (possibly having much more steps than there are data points) but then your function converges to an unbiased estimator for the test error rather than the generalization error.
+
 \end{document}

### 2019-07-31 MSRI ML Notes/MSRI_ML_notes.tex
Day 3

@@
 \DeclareMathOperator{\argmin}{argmin}
+\DeclareMathOperator{\argmax}{argmax}
 
@@
 
+\subsection{Weak Learning}
+At the end of the last lecture we discussed the ideal of weak vs.
+strong learning.
+In weak learning, we only require that there is an algorithm yielding a hypothesis $h$ such that $\err_R(h)\le\frac{1}{2}-\gamma$ for some $\gamma>0$.
+But surprise!
+\begin{thm}
+	A class is strongly learnable if and only fif it is weakly learnable.
+\end{thm}
+Today we will discuss why.
+We want to convert any algorithm that weakly learns a class into one that strongly learns the class.
+The idea here is that a class $\calC$ is either completely (PAC) learnable or else
+there is no algorithm that learns the concept any better than random chance.
+
+\subsubsection{The Ada Boost Algorithm}
+We will assume we are given a weak learning algorithm $A$ and data $(x_1,y_1),\dots,(x_m,y_m)\sim\scrD$. We want to find a hypothesis $H$
+whose error $\err_\scrD(H)$ is arbitrarily small.
+\begin{defn}
+	A theorem that can take a weak learning algorithm and convert it into an $\varepsilon$-good hypothesis is called a \textbf{boosting algorithm.}
+\end{defn}
+
+The basic idea is what one would expect; run $A$ a total of $T$ times, extracting weak hypotheses $h_1,\dots,h_T$ and somehow combine these in a way to get a strong hypothesis $H$.
+There is a problem with just feeding the same data in
+to $A$ over and over (e.g.
+$A$ may be deterministic).
+So instead we need to find a way of creating distributions $D_1,\dots, D_T$ that ``focus in'' on the bad/difficult examples that the weak hypotheses fail on.
+
+\begin{rmk}
+	Notice that the definition of learnability means that there has to be an algorithm that works \textbf{for all true distributions}. So for instance, the fact that 
+	we can learn something about the weather, but can't get it $\varepsilon$-good. This is because our algorithms only work well (enough) on a particular distribution.
+\end{rmk}
+
+Let $\varepsilon_t=\err_{D_t}(h_t)=\frac{1}{2}-\gamma_t$ and notice $\gamma_t\ge\gamma>0$.
+Then define the following distributions: $D_1$ will be the uniform distribution, so the weight on the $i^{th}$ training
+datum: $D(i)=\frac{1}{m}$. Then define the rest iteratively:
+\[D_{t+1}(i)=\frac{D_t(i)}{Z_t}\cdot \left\{\begin{array}{lr}e^{-\alpha_t},& h_t(x_i)=y_i\\
+e^{\alpha t}, & \text{otherwise}\end{array}\right.=\frac{D_t(i)e^{-\alpha_ty_ih_t(x_i)}}{Z_t}\]
+where $Z_t$ is a normalization constant and 
+\[\alpha_t=\frac{1}{2}\ln\left(\frac{1-\varepsilon_t}{\varepsilon_t}\right)>0\]
+
+Then we output $H$ which is defined as 
+\[H(x)=\operatorname{sign}\left(\sum_1^T\alpha_t h_t(x)\right).\]
+That is, we let a weighted sum (according to the error on the provided distribution) tells us what to select.
+
+\begin{thm}
+	\[\widehat\err(h)\le \prod_1^T2\sqrt{\varepsilon_t(1-\varepsilon_t)}=\prod\sqrt{1-4\gamma_t^2}\le\exp\left(-2\sum\gamma_t^2\right)\]
+	So if $\gamma_t>\gamma$, we get 
+	\[\widehat\err(h)\le e^{-2\gamma^2T}\]
+\end{thm}
+\begin{prf}
+	He went too fast. Slides are somewhere supposedly.
+\end{prf}
+
+So we get a bound on training error, great! Let's turn it into something about generalization error:
+if $h_t\in\calH$ and $d=\operatorname{VCdim}(\calH)$, then with probability $1-\delta$ (see problem session)
+\[\err(H)\le\widehat\err(H)+\tilde\calO\left(\sqrt{\frac{Td+\ln 1/\delta}{m}}\right)\]
+where $\tilde\calO$ means that we ignore log factors.
+
+Then something that may be surprising is that as $T$ grows, we may expect that eventually we will overfit our training data, but in practice you can often run this very far out and just keeping better results!
+in fact, your output $H$ will continue to get better even after it achieves zero error on the training set (since the underlying $h_t$ are still incorrect).
+
+But this points to the issue: the test error doesn't tell the whole story.
+Instead we are interested in measuring the \textit{confidence} the algorithm has in the combined classifier.
+So how do we do that?
+Recall that to determine $H$ we held an ``election'' for the best output.
+What we are interested in here is the \textbf{margin} (the difference between the fraction of votes for and against an output):
+we call the margin the \textit{weighted fraction of the $h_t$'s correct minus the weighted fraction of the $h_t$'s incorrect.}
+
+Doing some math, you get 
+\[\text{margin}=y\cdot\frac{\sum \alpha_t h_t(x)}{\sum\alpha_t}\]
+
+Since we are about out of time, we are sprinting to the finish line: Doing an analysis, we can show that this algorithm tends to push the margins towards the positive. Then you prove that 
+this phenomenon tends to create better generalization performance. Note that we can actually remove the dependence on $T$ by doing the following:
+\[\err(H)\le S+\tilde\calO\left(\frac{d/\theta^2+\ln 1.\delta}{m}\right)\]
+where $S$ is the fraction of training examples with margin $\le\theta$.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Bandits}
+This series was given by Kevin Jamieson, a professor in the CSE department here at UW.
+
+\subsection{Introduction}
+The core idea here (motivated very well by his description of the WSJ front page (ads, content, etc) and google maps (that does bandit-like testing to find new good routes) that balances
+the exploration (data-gathering) phase and the application phase and the transiton between them. It more resembles a conversation than a one-time transaction.
+
+The name comes from slot machines, obviously. :) We are thinking of these problems as having a slot machine with $n$ ``arms'', which can be pulled to yield some data.
+So the input to the algorithm is the number $n$ and for $t=1,2,\dots$, the algorithm pulls some arm $I_t\in[n]$ and nature reveals a corresponding reward $r_t\sim P_{I_t}$ where 
+$\bbE[r_t|I_t]=\theta_{I_t}^\ast$.
+
+\subsubsection{Regret Minimization}
+We define the \textbf{regret} as follows:
+\begin{defn}
+	After $T$ timesteps, the regret is 
+	\[R_T=\max_i\theta_i^\ast T-\bbE\left[\sum_1^T r_t\right]\]
+\end{defn}
+Then the goal is to achieve ``sublinear regret'', that is get that $R_T\in\mathbf{o}(T)$.
+\begin{lem}
+	\begin{align*}
+		R_T&=\max_i\theta_i^\ast T-\bbE[\sum_1^T\sum_1^n r_t\delta_{I_t=i}]\\
+		&=\max\theta_i^\ast T - \sum_1^n\bbE[\sum+1^T\theta_i^\ast\delta_{I_t=i}]\\
+		&=\max\theta_i^\ast T-\sum_1^n\theta_i^\ast\bbE[T_i]\\
+		&=\sum_2^n(\theta_1^\ast-\theta_i^\ast)\bbE[T_i]\\
+		&=\sum_2^n\Delta_i\bbE[T_i].
+	\end{align*}
+\end{lem}
+
+\subsection{Another game}
+This leads us to the concept of \textbf{best-arm identification}: we fix some $\delta\in(0,1)$ and we identify $\argmax_i\theta_i^\ast$ with probability $1=\delta$ with as few as possible total pulls.
+This time we don't put an upper bound on the time you're given, and instead your job is to identify what the best strategy is.
+
+\begin{defn}
+	A random variable $X$ is \textbf{$R$-sub-Gaussian} if 
+	\[\bbE[\exp(\lambda(x))]\le\exp(\lambda R^2/2)\]
+\end{defn}
+\begin{rmk}
+	Then by Hoeffding, if $X\in[a,b]$  and letting $R^2=\frac{(b-a)^2}{8}$ and notice when $X\sim \calN(0,1)$, we get $R^2=1$.
+\end{rmk}
+
+Then we can develop the Chernoff bound: Suppose $X_i$ are IID and $1$-sub-Gaussian. Then 
+\begin{align*}
+	\bbP(\sum_1^n X_i >\varepsilon)&=\bbP(\exp(\lambda\sum X_i) >\exp(\lambda\varepsilon))\\
+	&\le e^{-\lambda\varepsilon}\bbE[\exp(\lambda\sum X_i)]\\
+	&= e^{-\lambda\varepsilon}\prod\bbE[e^{\lambda X_i}]\\
+	&\le e^{-\lambda\varepsilon}(e^{\lambda^2/2})^n\\
+	&\le e^{-n\varepsilon^2/2}
+\end{align*}
+where the last step ``comes from calculus''. So by letting $\delta=e^{-n\varepsilon^2/2}$, one can solve for epsilon
+and get that with probability $1-\delta$, $\frac{1}{n}\sum X_i\le\sqrt{\frac{2\ln 1/\delta}{n}}$.
+
+If $n=2$, say $\theta_1^\ast\theta_2^\ast$ and $\Delta=\theta_1^\ast-\theta_2^\ast$. If both arms are pulled $\tau$ times, let the event
+\[\calE_i=\left\{\left|\theta_i^\ast-\frac{1}{\tau}\sum_{t=1}^\tau X_{i,t}\right|\le\sqrt{\frac{2\ln (4/\delta)}{\tau}}\right\}\]
+for $i=1,2$ (these are the ``good events''), then $\bbP(\calE_i^C)\le\delta/2$, so 
+\[\bbP(\calE_1^C\cup\calE_2^C)\le\sum\bbP(\calE_i^C)\le\delta.\]
+Then if $\tau=8\Delta^{-2}\ln(4/\delta)$, we can say 
+\[\hat\theta_2-\hat\theta_2=\hat\theta^1-\theta_1^\ast-\hat\theta_2+\theta_2^\ast+\theta_1^\ast-\theta_2^\ast>0\]
+with probability $1-\delta$. This of course means that we have accurately determined which arm to pull.
+
+Now consider the regret:
+\[R_T=\theta_1^\ast t=\bbE[\theta_1^\ast T_1+\theta_2^\ast T_2]=(\theta_1^\ast-\theta_2^\ast)\bbE[T_2]=\Delta\bbE[T_2]\]
+from which we can compute 
+\[\Delta\bbE[T_2]=\Delta\bbE[T_2\delta_{\calE_1\cap\calE_2}+T_2\delta_{\calE_1^C\cup \calE_2^C}]\le\Delta\tau+\Delta T\bbP[\calE_1^C\cup\calE_2^C]\]
+and so if we set $\delta=\frac{1}{T}$,
+\[R_T\le 8\Delta^{-1}\log(8T)+\Delta\]
+But notice that if we just always pick the bad arm, the worst regret we can get is $R_T\le \Delta T$. So 
+\[R_T\le1+\min{8\Delta^{-1}\log(8T),\Delta T}\le 1+\sqrt{8T\ln(8T)}\]
+
+If you were to graph the minimum function we had there, you'd notice that the regret tends towards zero near zero and towards infinity.
+The idea here is that if there is only a small difference between the optimal and suboptimal choice,
+then guessing randomly will only incur a tiny regret.
+Likewise, if the means are very far apart, you will quickly detect this through testing.
+There is a spike at about $1/\sqrt{T}$ that is the place where both ideas miss the mark.
+
+\subsection{Extending to more Arms}
+We are going to start by writing down an algorithm and then trying to analyze it:
+
+\subsubsection{Algorithm}
+We input $n$ arms and confidence $\delta$. Initialize $\hat X_1=\{1,\dots,n\}.$ Then for $l=1,2,\dots$ we let $\calE_l=2^{-l}$ and 
+\[\tau_l=\left\lceil 2\calE_l^{-2}\ln\frac{4 l^2 n}{\delta}\right\rceil\]
+Then pull each arm in $\hat X_l$ $\tau_l$ times and compute $\hat\theta_{i,l}$. Then define 
+\[\hat X_{l+1}=\hat X_l\setminus\left\{i\in \hat X_l:\max_{j\in\hat X_l}\hat\theta_{j,l}-\hat\theta_{i,l}>2\varepsilon_l\right\}\]
+
+\subsubsection{Analysis}
+Define $\calE_{i,l}=\{|\hat\theta_{i,l}-\theta_i^\ast|\le\varepsilon_l\}$ and let $\calE=\cap_{i,l}\calE_{i,l}$. But then 
+\[\bbP(\calE_{i,l}^C)\le 2\bbP(\hat\theta_{i,l}-\theta_i^\ast>\varepsilon_l)\le 2\exp{-\tau_l\varepsilon^2_l/2)}\le\frac{\delta}{2l^2n}\]
+and so $\bbP(\calE^c)\le\delta.$
+
+\begin{defn}
+	Let $\Delta_i=|\theta_i^\ast-\theta_1^\ast|$ for all $i>1$ and define $\Delta_1=\Delta_2$ (the minimum (nonzero) gap), where we assume for notational simplicity that $1\ge\theta_1^\ast>\theta_2^\ast\ge\cdots\ge\theta_n^\ast\ge 0$.
+\end{defn}
+\begin{lem}
+	With probability greater or equal to $1-\delta$, we have $1\in\hat X_l$ and 
+	\[\max_{i\in\hat X_l}\Delta_i\le 8\varepsilon_l\]
+	for all $l$.
+\end{lem}
+\begin{rmk}
+	The key thing to notice in the proof here is the fact that for any $j\in\hat X_l$,
+	\[\hat\theta_{j,l}-\hat\theta_{1,l}=(\hat\theta_{j,l}-\theta_j^\ast)-(\hat\theta_{1,l}-\theta_1^\ast)-\Delta_j\le 2\varepsilon_l-\Delta_j\le 2\varepsilon_l\]
+\end{rmk}
+
+Now let's talk about regret:
+\[R_T=\bbE\left[\sum_1^n\Delta_i T_i\right]\]
+and so for any $\nu>0$,
+\[\sum_1^n\Delta_iT_i\le \nu T+\sum_{l:8\varepsilon_l\ge \nu}8\varepsilon_l\tau_l|\hat X_l|\le \nu T+\sum_{l-1}^{\lceil\lg\max(\Delta,\nu)^{-1}\rceil}8\varepsilon_l\tau_l\delta_{\Delta_i\le 8\varepsilon_l}\]
+and then doing some more massaging, we get 
+\[\sum\Delta_iT_i\le \nu T+\sum_1^n c(\max(\Delta_i,\nu)^{-1})\log\left(\frac{n\log(\max(\Delta_i,\nu)^{-1})}{\delta}\right)\]
+
+\begin{rmk}
+	Notice there are some serious limitations here. In particular, it requires prior knowledge of $T$, which we don't always have.
+\end{rmk}
+
 \end{document}

### 2019-08-01 MSRI ML Notes/MSRI_ML_notes.tex
Day 4

@@
 \DeclareMathOperator{\err}{err}
-\DeclareMathOperator{\argmin}{argmin}
-\DeclareMathOperator{\argmax}{argmax}
+\DeclareMathOperator*{\argmin}{argmin}
+\DeclareMathOperator*{\argmax}{argmax}
 
@@
 
+\subsection{Optimality}
+Recall that we had two kind of bounds on expectation: $\frac{r_0L}{\sqrt{T}}$ and $\frac{r_0B}{\sqrt{T}}$ where $r_0=\|x_0-x^\ast\|_2$,
+$L=\sup_x\|\nabla f(x)\|$ and $B\ge \bbE[\|g_t\|^2]$ where $g_t$ is an unbiased estimator for $\nabla f(x_t)$, which we think of as a random oracle.
+
+\subsubsection{Basic minimax principle}
+We want to show that for all (possible randomized) algorithms such that there exists a function/oracle such that the error rate of gradient descent is optimal.
+It suffice to find a \textbf{random} function such that any deterministic algorithm on this function has error rate bigger than $\frac{\square}{\sqrt{T}}$ in expectation.
+
+This is a consequence of the fact that $\min\max\ge\max\min$ (these are read in the opposite order than you'd think).
+
+\subsubsection{Random oracle lower bound} 
+Let $K=[-1,1]$ and $f(x)=\xi x$ where $\xi=\pm\varepsilon$ randomly with equal probability.
+Then our random oracle is $g_t\sim \calN(\xi,1)$ (equivalently $g_t=\xi+Z_t$ for $Z_i$ IID $\calN(0,1)$). Intuitively, after $T$ queries the best estimator for $\xi$ is the 
+empirical mean $\frac{1}{T}\sum_1^T g_t\sim\calN(\xi,\frac{1}{T})=\xi+\frac{1}{\sqrt{T}}Z$ where $Z$ is $\calN(0,1)$.
+
+So if $\varepsilon<\frac{1}{\sqrt{T}}$, there is a constant probability to incorrectly estimate the sign of $\xi$. Formally,
+\begin{lem}
+	For any $\Psi:\bbR^T\to [-1,1],$
+	\[\bbP(\operatorname{sign}(\Psi(g_1,\dots,g_T))=\operatorname{sign}(\xi))\le\frac{1}{2}+\sqrt{T\varepsilon^2}\]
+\end{lem}
+
+\subsubsection{Deterministic oracle model} 
+let $v_1,\dots,v_{T+1}\in\bbR^n$ be a random orthonormal set. Let $f(x)=\max_{i\in[T+1]}v_i\cdot x$. We also let $K=\{x\in\bbR^n:\|x\|_2=1\}$. 
+Then $\nabla f(x)=v_i$ for $i=\argmax_{j} v_j\cdot x$ (except for a set of measure zero, where we can use subgradients).
+
+After $T$ queries, one of the $v_i$ remains unknown, say $T+1$. So for any $x, v_1,\dots, v_T$,
+\[\bbE_f[f(x)|v_1,\dots,v_T]\ge\bbE_f[v_{T+1}\cdot x]|v_1,\dots,v_T]=0\]
+And so $f^\ast=\min_{x\in B}f(x)\le-\frac{1}{\sqrt{T+1}}$. 
+\begin{rmk}
+	It is very important that $T<n$, since we are looking for an orthonormal set! But the bound is not dependent on the dimension! So although we got a similar lower bound on $T$, 
+	we still may be able to do better when we have lower dimension.
+\end{rmk}
+
+To see this, we need 
+\begin{prop}[Gr\"unbaum's Inequality]
+Let $K$ be a convex body. For any half-space $H$ that contains the center of gravity of $K$. Then 
+\[\operatorname{vol}(K]\cap M)\ge\frac{1}{e}\operatorname{vol}(K).\]
+\end{prop}
+
+How do we get half spaces by making gradient queries? Well by convexity, $f(x)\ge f(x_t)+\nabla f(x_t)(x-x_t)\ge f(x_t)$ for any $x\in H_t=\{x:\nabla(x_t)\cdot(x-x_t)\ge 0\}$.
+
+Then we have the \textbf{center of gravity algorithm}: let $K_0=K$ and for any $T\ge 0$, $x_t=cg(K_t)$ and set $K_{t+1}=K_t\setminus H_t$.
+Then by Gr\"unbaum, $\operatorname{vol}(K_t)\le(1-\frac{1}{e})\operatorname{vol}(K_{t-1})\le(1-\frac{1}{e})^t\operatorname{vol}(K)$
+\begin{thm}
+	Say $f:K\to [-1,1]$. After $\calO(n\log(1/\varepsilon))$ steps, $f(x_t)-f(x^\ast)\le\varepsilon$.
+\end{thm}
+\begin{prf}
+	After this many steps, $\operatorname{vol}(K_t)\le\varepsilon^n\operatorname{vol}(K)$.
+We claim $\exists \Omega\subseteq K$ such that $\forall x\in\Omega$, $f(x)-f(x^\ast)\le\varepsilon$ and $\operatorname{vol}(\Omega)\ge \varepsilon^n$
+\end{prf}
+
+\subsubsection{Beyond \texorpdfstring{$\frac{1}{\sqrt{T}}$}{1/sqrt(T)} in high dimensions}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+The previous algorithm we described (which one?) we will denote AE1, and we ended up with regret 
+\[R_T\le\sqrt{cnT\log(nT)}\]
+and 
+\[R_T\le c\sum_2^n\Delta_i^{-1}\log(T)\]
+
+One can show that with probability $\ge 1-\delta$, AE1 give us that $\hat X_l=\{1\}$ after at most 
+\[\tau=\sum_2^n c\Delta_i^{-2}\log\left(\frac{n\log(\Delta_i^{-2})}{\delta}\right)\]
+pulls.
+
+\subsection{Theoretical Bounds}
+Then the question one can ask is whether the bounds given are \textit{tight}. A information-theoretic result we will see in the problem session.
+\begin{prop}
+	Suppose $n$ samples are taken from $\calN(\mu,1)$ where $\mu\in\{0,\Delta\}$. To identify if $\mu=0$ or $\Delta$ with probability $1-\delta$,
+	one requires $n\ge 2\Delta^{-2}\log(1/4\delta).$
+\end{prop}
+
+This really hints how difficult the multi-arm bandit problem can be. 
+\begin{thm}
+	Any $\delta$-PAC on $\{P:P_i=\calN(\theta_i,1),\theta\in[0,1]^n\}$ satisfies 
+	\[\bbE[\tau|\theta^\ast]\ge 2\log(1/2\delta)\sum_2^n\Delta_i^{-2}\]
+\end{thm}
+\begin{rmk}
+	This is a result on expectation rather than a high-probability bound, obviously, but under these conditions one can turn this into a lower-bound in the high-probability case 
+	as well (after losing a big constant factor). Going in the other direction is tricky, because there are bad cases in which this algorithm never terminates or outputs the right thing, so the 
+	expectation goes up a bit.
+\end{rmk}
+\begin{rmk}
+	If we take the partial sums $S_T=\sum_{t=1}^T Z_t$ of your random variables, you get the \textit{law of the iterated logarithm}, which 
+	says 
+	\[\limsup_T\frac{S_T}{\sqrt{2T\log\log T}}=1\]
+	so the $\log\log$ factor in the bound we found is actually tight.
+\end{rmk}
+We also get:
+\begin{thm}\label{thm-UCB}
+	Any algorithm satisfying $\bbE[T_i]\le\mathbf o(T^\alpha)$ for all $i\ne 1$ (suboptimal samples) any $\alpha\in(0,1)$ suffers 
+	\[\lim_{T\to\infty}\frac{R_T}{\log(T)}=\sum_2^n\frac{2}{\Delta_i}\]
+\end{thm}
+and a worst-case result:
+\begin{thm}
+	For any $n$ and $T$, there exists an instance ($\theta^\ast$) such that 
+	\[R_T\ge\sqrt{(n-1)T}/27.\]
+\end{thm}
+\subsection{Upper Confidence Bound}
+An algorithm that achieves the bound in thm~\ref{thm-UCB} is as follows. For $t=1,2,\dots$,
+pull $\argmax_i\hat\theta_{c,T_i(t)}+\sqrt{\frac{2\log(t)}{T_i(t)}}$ where $T_i(t)$ is the number of pulls up to $t$.
+
+This leads to a minimization of regret, but notice that this algorithm is suboptimal on the problem of identifying the best arm. But Kevin 
+created an alteration called ``Lil' UCB'' that proceeds similarly but adds in an extra $\frac{1}{\delta}$ factor (as well as some constants) to achieve optimal best arm identification.
+
+These algorithms have this phenomenon where, as time goes on, the likelihood of being selected gradually increases, but then each pull reduces it somewhat. Also notice that this algorithm 
+is very much not optimal for an overall ranking since it will stop pulling arms that have a bad run.
+
+\subsection{Linear Bandits}
+Notice that so far we still have a dependence on $n$, but if you are running (e.g.) a music recommendation software, you likely have millions of data points to sample.
+Furthermore, songs have tons of features to compare, so we would like to be able to exploit the fact that a person is likely to enjoy music that is ``close'' under some sort of metric (using genre, BPM, etc).
+
+So we develop a \textbf{linear bandits:} For $t=1,2,\dots$, we pull arm $I_t$ and nature reveals $\langle x_{I_t},\theta^\ast\rangle+\xi_t$. 
+Here we say $\xi_t\sim\calN(0,1)$. We take as known $X=\{x_1,\dots,x_n\}\subset \bbR^d$ and $\theta^\ast\in\bbR^d$ is unknown.
+We define the best arm to be $x^\ast=\argmax_{x\in X}\langle x,\theta^\ast\rangle$.
+
+Say we observe some data $\{(x_i,y_i)\}_1^\tau$ where $y_i=\langle x_i,\theta^\ast\rangle+\xi_i$. Then setting 
+\[\hat\theta=\argmin_{\theta\in\bbR^d}\sum_1^\tau(\langle x_i,\theta\rangle-y_i)^2\]
+and then if $X=(x_i^T)^T$ and $Y=(y_i)=X\theta^\ast+\xi$, then 
+\[\hat\theta=(X^TX)^{-1}X^TY.\]
+
+Then unfolding $Y$, we get 
+\[\hat\theta=\theta^\ast+(X^TX)^{-1}X^T\xi\]
+thus $\bbE[\hat\theta]=\theta^\ast$. Computing the standard deviation:
+\[\bbE[(\hat\theta-\theta^\ast)(\hat\theta-\theta^\ast)]=(X^TX)^{-1}\]
+so $\hat\theta-\theta^\ast\sim\calN(0,(X^TX)^{-1})$.
+
+Importantly, notice that the variance of $\hat\theta-\theta^\ast$ depends only on $X$ so we can ``plan ahead'' to control how good we want our estimator $\hat\theta$ to be!
+
+Now notice that for any $z\in\bbR^d$, we have $\langle z,\hat\theta-\theta^\ast\rangle\sim\calN(0,z^T(X^TX)^{-1}z)$. Then if we just call $\sigma^2=z^T(X^TX)^{-1}z$, then 
+\[\bbP\left(\langle z,\hat\theta-\theta^\ast\rangle>\sqrt{2\sigma^2\log(1/\delta)}\right)\le\delta.\]
+Writing this another way, with probability $1-\delta,$
+\[|\langle z,\hat\theta-\theta^\ast\rangle|\le\sqrt{z^T(X^TX)^{-1}z}\cdot\sqrt{2\log(2/\delta)}=\|z\|_{(X^TX)^{-1}}\sqrt{2\log(2/\delta)}\]
+where, as a matter of notation, we write $x^TAx=\|x\|_A^2$.
+
+Somehow the idea here is that you may have data lining up along a certain line, in which case your confidence interval along the line in question goes down.
+
+\brk
+
+For all $X$, there exists $\lambda\in\Delta_x$ such that 
+\[\tau\sum_{x\in X}\lambda_x xx^T=X^TX\]
+and so if we set $A_\lambda=\sum \lambda_xxx^T$, we get a bunch of optimization problems:
+\begin{itemize}
+	\item E-optimal
+	\[\min_\lambda\max_{\|u\|_2\le 1}u^TA^{-1}_\lambda u\]
+	\item A-optimal
+	\[\min_\lambda\operatorname{Tr}(A_\lambda^{-1})\]
+	\item G-optimal
+	\[\min_\lambda\max_{x\in X}x^TA^{-1}_\lambda x\]
+	\item D-optimal
+	\[\max_{\lambda}\log\det(A_\lambda)\]
+\end{itemize}
+\begin{thm}[Kiefer-Wolfowitz]
+	For any $X\in\bbR^d$ (that spans), there is $\lambda^\ast\in\Delta_X$ such that 
+	\begin{enumerate}
+		\item $\max_\lambda g_D(\lambda)=g_D(\lambda^\ast)$
+		\item $\min_\lambda f_G(\lambda)=f_G(\lambda^\ast)$
+		\item $\operatorname{support}(\lambda^\ast)\le\frac{(d+1)d}{2}$
+		\item $f_G(\lambda^\ast)=g_D(\lambda^\ast)=d$
+	\end{enumerate}
+	where $g_D(\lambda)=\log\det(A_\lambda)$ and $f_G(\lambda)=\max_{x\in X}x^TA_\lambda^{-1}x.$
+\end{thm}
+
+\begin{prop}
+	If $\lambda^\ast$ is the $\frac{(d+1)d}2$-sparse $G$-optimal solution and $x$ is pulled $\lceil\lambda_x\tau\rceil$ and we compute $\hat\theta$ (using at most $\frac{(d+1)d}{2}+\tau$ pulls)
+	then with probability $\ge 1-\delta$, for any fixed $x\in X$,
+	\[\langle x,\hat\theta-\theta^\ast\rangle\le\sqrt{\frac{2d\log(1/\delta)}{\tau}}\]
+\end{prop}
+
+\subsection{AE revisited}
+Recall AE1: for $t=1,\dots$, $\hat\lambda_l$ is $G$-optimal for $\hat X_l$ and $\varepsilon_l=2^{-l}$ and $\tau_l=2d\varepsilon_l^{-2}\log\left(\frac{4l^2|x|}{\delta}\right).$
+Then pull arm $x$ $\lceil\tau_l\hat\lambda_l\rceil$ times and construct $\hat\theta_l$. Then let 
+\[\hat X_{l+1}=\hat X_l\setminus\{x\in\hat X_l:\max_{x'\in X}\langle x'-x,\hat\theta_l\rangle>2\varepsilon_l\}\]
+then $\bbP(\cup_l\cup_{x\in X}\calE_{x,l}^c)\le\delta$ where $\calE_{x,l}$ is the event that $\langle x,\hat\theta_l-\theta^\ast\rangle\le\varepsilon_l.$
+
+Going through basically the entire argument from before,
+\[R_T\le c\sqrt{dT\log(|X|T)}.\]
+\begin{rmk}
+	Notice that part of what we wanted in all this was to get rid of dependence on the sides of in the inputs, $|X|$.
+The upshot is, in many (all?) cases, you can just a kinds of $\varepsilon$-cover of your space, which means we can replace it with a factor of $d$.
+\end{rmk}
+
 \end{document}

### 2019-08-04 MSRI ML Notes/MSRI_ML_notes.tex
Friday Update

@@
 \DeclareMathOperator*{\argmax}{argmax}
+\newcommand{\grad}{\nabla}
 
@@
 \subsubsection{Beyond \texorpdfstring{$\frac{1}{\sqrt{T}}$}{1/sqrt(T)} in high dimensions}
+The best way to get around the lower bounds we've cooked up for ourselves is to put some further restrictions on our functions:
+\begin{defn}
+	A function $f$ is $\alpha$\textbf{-strongly convex} if $f(x)=g(x)+\frac{\alpha}{2}\|x\|_2^2$, where $g$ is convex.
+\end{defn}
+\begin{rmk}
+	This property is equivalent with the fact that 
+	\[f(x)-f(y)\le \nabla f(x)\cdot (x-y)-\frac{\alpha}{2}\|x-y\|_2^2.\]
+	So we have that $f$ curves away from its tangent at that point like a parabola with coefficient $\alpha$.
+\end{rmk}
+
+Let's do the continuous-time analysis of the gradient flow $\frac{d}{dt}x(t)=-\nabla f(x(t))$. Let $r^2(t)=\|x(t)-x^\ast\|^2$ and $\delta(t)=f(x(t))-f(x^\ast)$.
+Then the calculation we get is 
+\begin{align*}
+	\frac{d}{dt} r^2(t)=-\nabla f(x(t))\cdot(x(t)-x^\ast)\\
+	&\le -\delta(t)-\frac{\alpha}{2}\|x(t)-x^\ast\|^2_2.
+\end{align*}
+Thus $\frac{d}{dt}r^2(t)\le -\alpha r^2(t)$ and so $r^2(t)\le r^2(0)e^{-\alpha t}$.
+All this to say that gradient flow gets to OPT in time $\calO(\frac{1}{\alpha}\log(r_0^2/\varepsilon))$.
+By comparison with out strong convexity,
+we had that 
+\[\frac{1}{\tau}\int_0^\tau\delta(t)\mathrm{d}t\le \frac{r^2_0(0)}{\tau}\]
+and so our time is $\calO(r_0(0)/\varepsilon)$.
+
+By switching from gradient flow to gradient descent, we pick up a discretization error! let $\eta L^2$ be the error from gradient flow at $\tau=T\eta$.
+Then through some analysis, we get a $\tilde\calO(L^2/\alpha T)$ rate.
+
+\brk
+
+Next we can talk about \textbf{smoothness}.
+\begin{defn}
+	A function $f$ is $\beta$-smooth if $\nabla f$ is $\beta$-Lipshitz.
+\end{defn}
+Intuitively, we are looking to improve the discretization error. Before we could overshoot our maximum easily because the gradient could change sharply.
+Now, we are forcing it to tend (loosely) to zero as you get closer to the optimum. Notice
+\[\|\nabla f(x+\eta\nabla f(x))-\nabla f(x)\|\le\beta\eta\|\nabla f(x)\|\]
+which suggests that $\eta\propto\frac{1}{\beta}$ is fine for the discretization. In this case we get convergence like $r^2(0)\beta/T$ (roughly $1/\varepsilon$)
+and if we combine both $\alpha$-strong convexity and $\beta$-smoothness, you get a parameter 
+\[\kappa=\frac{\beta}{\alpha}\]
+called the \textbf{Condition number} and we get growth like $\kappa\log(1/\varepsilon)$
+
+\subsubsection{A formal inequality}
+\begin{prop}
+	Let $x^+=x-\frac{1}{\beta}\nabla f(x)$. Then 
+	\[f(x^+)-f(x)\le\frac{-1}{2\beta}\|\nabla f(x)\|_2^2.\]
+\end{prop}
+\begin{prf}
+	We begin by claiming 
+	\[f(x)-f(y)\le\nabla f(y)\cdot(x-y)+\frac{\beta}{2}\|x-y\|_2^2\]
+	which implies that 
+	\[f(x^+)-f(x)\le\grad f(x)\cdot(x^+-x)+\frac{\beta}{2}\|x^+-x\|_2^2\]
+	and unfolding this 
+	\[f(x^+)-f(x)\le-\frac{1}{2\beta}\|\nabla f(x)\|^2.\]
+\end{prf}
+\begin{defn}
+	A point $x$ is called \textbf{$\varepsilon$-stationary point of $f$} if $\|\grad f(x)\|\le \varepsilon$.
+\end{defn}
+\begin{rmk}
+	The above inequality shows that gradient descent on a bounded, smooth, non-convex function finds an $\varepsilon$-stationary point in $1/\varepsilon^2$ steps.
+\end{rmk}
+\begin{ex}
+	Let $f(x)=\frac{1}{2m}\|Ax-b\|^2$ where $A$ is $m\times m$. Then the Hessian is 
+	\[\grad^2 f(x)=\frac{1}{m}A^TA=\frac{1}{m}\sum_1^ma_ia_i^T\]
+	where the $a_i$ should be thought of as feature vectors. Then the above quantity is the sample covariance matrix of the features.
+
+	In this case, $\alpha=\lambda_{min}(\frac{1}{m}A^TA$ and $\beta=\lambda_{max}(\frac{1}{m}A^TA)$, so if $\|a_i\|_2\le 1$ then $\beta\le 1$.
+\end{ex}
+\begin{rmk}
+	It is natural to ask how we can get a better value of $\beta.$ One way is to convolve your function $f$ with some noise $z$: $\hat f(x)=f(x+z)$.
+
+	Sometimes $f$ is nonsmooth, but can be written as $f=g+r$ where $g$ is smooth and $r$ is ``simple''. Typically one takes $r(x)=\|x\|_1$.
+\end{rmk}
+\subsection{Acceleration}
+It turns out that $1/\sqrt{\varepsilon}$ (for smooth) and $\sqrt{\kappa}\log(1/\varepsilon)$ for smooth and strongly convex functions are achievable!
+
+What could be better than gradient descent? Well here's a crazy idea:
+\[x_{t+1}=\argmin_{\operatorname{span}(\grad f(x_0),\dots,\grad f(x_t))}[f(x)]\]
+where $[f(x)]=\sum_{s=1}^t\lambda_s\grad f(x_s)$. THen if $f$ is quadratic, it turns out that $\lambda_s$ has a closed form formula!
+
+\subsubsection{Nemiraski's Acceleration (1982)}
+Let 
+\[x_{t+1}=\argmin_{x\in P_t} f(x)\]
+where $P_t=\operatorname{span}(x_t^+,\sum_1^t \lambda_s\nabla f(x_s))$.
+
+Then doing some analysis, notice
+\[\sum_1^T\lambda_s\delta_s\le\sum_1^T\lambda_s\grad f(x_s)\cdot(x_s-x^\ast)=\sum_1^T\lambda_s\grad f(x_s)\cdot(-x^\ast)\]
+where the last equality follows since $x_s\in P_{s-1}$ implies that $\grad f(x_s)\in P_{s-2}^\perp$. Thus 
+\[\|x^\ast\|\cdot\sqrt{\sum_1^t\lambda_s^2\|\grad f(x_s)\|^2}\]
+since $\grad f(x_{T+1})\perp \sum_1^T\lambda_s\grad f(x_s)$. But then by smoothness, 
+\[\frac{1}{2beta}\|\grad f(x_s)\|^2\le \delta_s-\delta_{s+1}\]
+and so 
+\[\sum_1^R\lambda_s\delta_s\le\|x^\ast\|\cdot\sqrt{2\beta\sum_1^T\lambda_s^2(\delta_s-\delta_{s+1})}.\]
+
+Now pick some $\lambda_s$ such that $2\beta(\lambda_s^2-\lambda_{s-1}^2)=\lambda_s$ and after some juggling things, 
+\[\delta_T\le\frac{\beta\|x^\ast\|^2}{T^2}\]
+and so we get an algorithm that runs in time like $1/\sqrt{\varepsilon}$.
+
+\subsubsection{Nesterov's AGD (1983)}
+The key breakthrough in this algorithm is that all that matters about the past is the momentum.
+\begin{defn}
+	The \textbf{momentum} if an iterative algorithm is 
+	\[d_t:=\gamma_t(x_t-x_{t-1}).\]
+\end{defn}
+
+The began with Polyak's \textit{heavy ball method.} Let $x_{t+1}=x_t^++d_t$. This method alone doesn't accelerate (can be seen on simple quadratic functions).
+Instead, Nestrov's uses
+\[x_{t+1}=(x_t+d_t)^+.\]
+
+Now for some of the analysis. Unfortuately this is notoriously ``magical'', so hold onto your pants. Denote by $g_t=-\frac{1}{\beta}\grad f(x_t+d_t)$. Thus $x_{t+1}=x_t+d_t+g_t.$
+Then 
+\[\delta_{t+1}-\delta_t=f(x_t+d_t)^+)-f(x_t)\le f(x_t+d_t)=\frac{1}{2\beta}\|\grad f(x_t+d_t)\|_2^2-f(x_t)\]
+by the smoothness of $f$. Then by convexity,
+\[\delta_{t+1}-\delta_t\le \grad f(x_t+d_t)\cdot d_t-\frac{1}{2\beta}\|\grad f(x_t+d_t)\|_2^2=-\beta g_t\cdot d_t-\frac{\beta}{2}\|g_t\|_2^2.\]
+
+But notice that 
+\[\delta_{t+1}=f((x_t+d_t)^+)-f(x^\ast)\le=\beta g_t\cdot(x_t+d_t-x^\ast)-\frac{\beta}{2}\|g_t\|_2^2.\]
+And the plan is to take a combination of these two statements to get a weighted sum of the $\delta_i$ on the left and a telescopic sum on the right.
+It wasn't that bad.
+Basically just put in some parameters and then choose them to be the ``right ones''.
+THe punchline is approximate convergence like $\delta_{T+1}\le\frac{\beta\|x_1-x^\ast\|^2}{T^2}.$
+
+\subsection{Some variance reduction techniques}
+\subsubsection{The finite sum problem}
+Let's say we have a function $f(x)=\frac{1}{m}\sum_i f_i(x)$ where $f_i$ is $\beta$-smooth (note that in particular this implies $f$ is $\beta$-smooth) and $f$ is $\alpha$-strongly convex.
+Then stochastic gradient descent does $x_{t+1}=x_t-\eta\grad f_{i_t}(x_t)$ which runs like $1/\alpha\varepsilon$ (the number of ``data operations'').
+
+In regular gradient descent, $x_{t+1}=x_t-\frac{\eta}{m}\sum_1^m\grad f_i(x_t)$, which runs with $\kappa\log(1/\varepsilon)$ data operations.
+The differnce here is that SGD has a lot of variance in its estimators, so to be ``safe'' you have to take smaller steps. On the other hand, GD needs much more data but can run with better accuracy.
+So the idea here is to somehow control the variance.
+
+The key idea is that we, from time, to time, compute a full gradient at some point $y$.
+\[g_t:=\grad f_{i_t}(x_t)-\grad f_{i_t}+\grad f(y)\]
+We can see that $g_t$ is un unbiased estimator of $\grad f_{i_t}(x_t)$ with the added bonus that the variance $\bbE[\|g_t\|^2]$ gets smaller when close to the optimum.
+\begin{thm}
+	During an epoch $T$, let $x_1\leftarrow y$ and for $t\ge 2$, $x_{t+1}=x_t-\eta g_t$. Then set $y$ to be $\frac{1}{T}\sum_1^Tx_t$ and recompute $\grad f(y)$.
+	Then by using $\eta=\frac{1}{10\beta}$,
+	\[f(\text{new } y)-f(x^\ast)\le 0.9 (f(y)-f(x^\ast)).\]
+\end{thm}
+
+\subsection{Further Topics}
+\begin{itemize}
+	\item We talked earlier about decomposing functions into smooth and non-smooth (simple) parts. 
+	\item Mini-batch SGD: write $x_{t+1}=x_t-\frac{\eta}{|B|}\sum_B\grad f_i(x)$ where $B$ is a batch. Then we get similar results for convergence up to $b=\sqrt{m}$.
+	\item Bandit convex optimization
+	\item Interior point method: tells you how to handle complicated $K.$
+	\item Constant step size least squares: no $1/\alpha$ in the bound.
+	\item Mirror descent: the equation $x-\grad f(x)$ doesn't actually make sense (one is a column, one is a row). Instead we pick a function $\phi$ and look at 
+	\[\grad\phi^{-1}(\grad\phi(x)-\grad f(x)).\]
+\end{itemize}
 
@@
 
+\subsection{Contextual Bandits}
+Another game we can play: for $t=1,\dots$ Nature reveals context $c_t\sim D$ (IID). Then the player chooses $x_t\in X$ and nature reveals the reward (with noise)
+\[r_t=v(v_t,x_t)+\xi_t\]
+where $X$ is known and we know $\bbE[\xi]=0.$
+
+\subsubsection{Linear Contextual Bandits}
+One can actually capture this in the linear bandit framework. Assume there exists a feater map $\Phi:X\times X\to \bbR^d$ and $v(c,x)=(\Phi(c,x),\theta^\ast)$ for some unknown $\theta^\ast$.
+For example, if $x\in\bbR^p$ and $c\in\bbR^q$, we could have $\Phi(c,x)=\operatorname{vec}(xc^T)\in\bbR^{pq}$.
+\begin{rmk}
+	Often one uses as $\Phi$ a deep network. One can train a deep network and then take out the last layer and use it to assign features to your data.
+\end{rmk}
+\subsubsection{General Policies}
+A policy is a map $\pi:\calC\to X$, that is it assigns an action to each observed context. Then the \textbf{value} $V(\pi)$ is 
+\[V(\pi)=\bbE_{c\sim D}[v(c,\pi(c))+\xi].\]
+To define regret, we say 
+\[R_T=T\cdot v(\pi^\ast)-\bbE\left[\sum_1^T v(\pi_t)\right]\]
+where $\pi^\ast=\argmax_{\pi\in\Pi}v(\pi)$.
+\subsubsection{Inverse Propensity Scores}
+We define an estimator for $v(\pi)$ no matter what $\pi$ we have. Suppose we have a distribution $\mu(x|c)\in\bbS^{|X|}$. Then we play action $x_t$ distributed as $\mu(x|c_t)$ for all $t$.
+Build the estimator $\hat v(c_t,x)=r_t\frac{\delta_{x_t=x}}{\mu(x|c_t)}.$ Then we define the estimator 
+\[\hat v_t(\pi)=\frac{1}{t}\sum_{s=1}^t\hat v(c_s,\pi(c_s)).\]
+
+Notice that 
+\begin{align*}
+	\bbE[\hat v_t(c_t,x)|c_t] &= \sum_{x'\in X}\mu(x'|c_t)\bbE\left[\left.r_t\frac{\delta_{x_t=x}}{\mu(x|c_t)}\right| c_t,x_t=x'\right]\\
+	&=\sum_{x'\in X}\mu(x'|c)v(c_t,x')\frac{\delta_{x_t=x}}{\mu(x|c_t)}\\
+	&=v(c_t,x)
+\end{align*}
+So we get that $\hat v(c_t|x)$ is unbiased, but could suffer from very large variance. For instance it is zero unless $x=x_t$!
+Then one can show $\bbE[\hat v_t(\pi)]=v(\pi)$.
+\begin{rmk}
+	To follow this process, you need to record as data $\{(c_t,x_t,r_t,\mu(c_t,x_t))\}$. Most important is the distribution $\mu$! This is a common mistake in practice but is useless without this data.
+\end{rmk}
+
+To compute the variance:
+\[\bbE[(\hat v_t(c_x,x)-\bbE[\cdot])^2|c_t]\le\frac{1}{\mu(x,c_t)}.\]
+
+Now let $Q\in \bbS^{\pi}$ be a distribution over policies. Then we are going to let 
+\[\mu(x|c)=\sum_{\pi\in\Pi:\pi(c)=x}Q(\pi).\]
+Then at round $l$, we choose 
+\[\hat Q_l=\argmin_{Q\in\bbS^{|\hat\Pi_l|}}\max_{\pi\in\hat\Pi_l}\bbE[(\hat v_t(\pi)-v(\pi))^2]\]
+Using this scheme, we can get a regret bound that looks like 
+\[R_T\le\sqrt{T|x|\log(|\pi|T)}\]
+\subsection{Introduction to Adversarial Bandits}
+In this scheme, an adversary chooses $x\in[0,1]^{n\times T}$. Then for $t=1,2,\dots$, the player chooses $I_t\in[n]$, and receives reward $x_{I_t,t}$.
+Then we set the regret to be 
+\[R_T=\max_{i\in[n]}\sum_1^Tx_{i,t}-\bbE\left[\sum_1^t x_{I_t,t}\right]\]
+The twist here is that your adversary will be able to choose $x$ \textit{with full knowledge of your algorithm}.
+Notice that any deterministic algorithm can be tricked into achieving linear regret.
+So we need something stochastic.
+
+Let $\hat X_{i,t}=\frac{\delta_{I_t=i}}{P_{i,t}}x_{I_t,t}$ where $I_t\sim P_t\in\bbS^n$. Then $\bbE[\hat X_{i,t}]=x_{i,t}$.
+
+The algorithm we need will be able to deal with the high variance of this estimator. It is called EXP3 (exponential something for exploration and exploitation). Playing this game, you can get 
+\[R_T\le\sqrt{nt\log n}.\]
+
+Some textbooks: Seb has one. Also check out \href{banditalgs.org}{banditalgs.org}.
+
 \end{document}

### 2019-08-05 MSRI ML Notes/MSRI_ML_notes.tex
Second Monday

@@
 
-Some textbooks: Seb has one. Also check out \href{banditalgs.org}{banditalgs.org}.
+Some textbooks: Seb has one. You can find a preprint on the arXiv. Also check out \href{banditalgs.org}{banditalgs.org}.
+
+\section{Deep Learning}
+This talk was given by Joan Bruna from NYU's Courant Institute.
+\subsection{The Puzzle}
+We are primarily going to be interested in what is called \textit{supervised learning}. Notice that this specifically omits some hot ideas that are currently being developed.
+Throughout the talks, we will be considering $\calX$, a \textit{high-dimensional} space.
+For instance, if we were working in the domain of computer vision, $\calX$ may be the space of all $d\times d$ images,
+giving us a huge space, potentially.
+
+The next ingredient we have is the \textbf{data distribution} $\nu$ on $\calX$.
+Importantly, $U$ is \textit{unknown to us}, so we are not allowed to directly access any of its properties.
+Next, we have a \textbf{target function}
+$f^\ast:\calX\to \bbR$ which we are trying to learn. We also need a loss functional $L(f)=\bbE_\nu[l(f(x),f^\ast(x))].$
+\begin{rmk}
+	In this class, we will almost always restrict to the case when $L$ is given by 
+	\[\|f-f^\ast\|^2_{L^2(\calX,\nu)}\]
+\end{rmk}
+
+Then our goal is to predict $f^\ast$ from a finite IID sample $\{(x_i,f(x_i)\}$, where $x_i\sim \nu$.
+Notice that this is basically the same as a problem we have all done at some point: fit a curve to a finite number of points. But now the dimensionality has exploded.
+\subsubsection{Empirical Risk Minimization}
+Consider a set $\calF\subseteq\{f:\calX\to\bbR\}$, which we call the \textbf{hypothesis class}. This space of functions comes along with a notion of \textit{complexity}.
+We have a value $\gamma(f)$ for each $f\in\calF$, which we call the \textbf{complexity of $f$.} One should think of this like a norm.
+The upshot here is this gives us a way to ``organize'' $\calF$ in a nice way.
+
+A natural way to proceed is to begin with the norm zero piece and gradually increase to consider more candidate functions.
+\begin{rmk}
+	Careful! Notice that assuming that $\gamma$ is honestly a norm means that such ``balls'' are already convex! In general this will not be true.
+\end{rmk}
+\begin{defn}
+	The \textbf{empirical risk} of a function $f$ is 
+	\[\hat L(f)=\frac{1}{n}\sum_1^n|f(x_i)-f^\ast(x_i)|^2.\]
+\end{defn}
+
+Now if one wants to minimize $\hat L$, there are a couple different ways to do this. The first idea is consider the \textbf{constrained form}:
+\[\min_{f\in\calF^\delta}\hat L(f)\]
+where $\calF^\delta=\{f\in\calF:\lambda(f)\le\delta\}$. This requires a bit more prior knowledge than we generally have, so we sometimes use instead a \textbf{penalized form}:
+\[\min_{f\in\calF}\hat L(f)+\lambda\cdot\gamma(f).\]
+Another idea is called the \textbf{interpolant}:
+\[\min_{f\in\calF} \gamma(f) \text{  such that  }\hat L(f)=0\]
+
+\subsubsection{The Fundamental Theorem of Machine Learning}
+Let us assume for this formulation that we are focusing on the constrained form. This was found in the seminal paper from 2008 by Bolton and Bousquet.
+\begin{thm}[FtoML]
+	Assume that we have $\hat f\in\calF^\delta$ such that $\hat L(\hat f)\le \varepsilon+\inf_{f\in\calF^\delta}\hat L(f).$ Then 
+	\[L(\hat f)-\inf_{f\in \calF}L(f)\le \inf_{f\in\calF^\delta}L(f)-\inf_{f\in\calF}L(f)+2\sup_{f\in\calF^\delta}|L(f)-\hat L(f)|+\varepsilon.\]
+\end{thm}
+\begin{rmk}
+	Here the first two terms on the right constitute an approximation error, the third term is a statistical error, and the $\varepsilon$ comes from our optimization error.
+
+	Notice that we have a free parameter $\delta$.
+If $\delta$ is small, then our approximation error gets very bad and the statistical error drops (we are working on a ``smaller'' set in terms of complexity).
+On the other hand, if $\delta$ is very big, the statistical error blows up.
+\end{rmk}
+\subsection{The Big Questions}
+We have a series of questions we want to answer with this theory:
+\begin{itemize}
+	\item \textbf{[Approx]} How can we design ``good'' spaces $\calF$ to approximate $f^\ast$ in higher dimensions?
+	\item \textbf{[Optimization]} How to design algorithms to solve the problem of empirical risk management in general (gradient descent).
+\end{itemize}
+
+\subsubsection{The Curse of Dimensionality}
+To begin to convince ourselves that dimensionality is troubling, consider the question: How many samples do I need to estimate $f^\ast$ 
+depending on its regularity assumptions?
+
+If we assume that $f^\ast$ is \textit{linear}, then $f^\ast(x)=\langle x,\theta^\ast\rangle$ for some $\theta^\ast\in\bbR^d$.
+Then $\calF$ is the set of all such functions and we need $d$ samples to get enough equations to solve the corresponding linear equations.
+
+Now assume that $f^\ast$ is ``locally linear'', that is \textit{Lipschitz.} That is, there exists a (minimal) $\beta$ such that 
+\[|f^\ast(x)-f^\ast(y)|\le\beta\|x-y\|.\]
+Then $\calF$ is the space of all Lipschitz functions.
+Then $\gamma(f)$ is determined by the Lipschitz constant as well as $|f|_\infty$.
+Now for all $\varepsilon>0$, we want to find $f\in\calF$ such that $\|f-f^\ast\|\le\varepsilon$
+from $n$ IID samples. Then in this case, the sample complexity blows up! To get an $\varepsilon$-error, we need $n\sim \varepsilon^{-d}$ samples!
+
+Let's build the upper bound here: given a set of points $(x_i,f^\ast(x_i))$, we want to find 
+\[\hat f=\argmin_f\operatorname{Lip}(f)\text{  such that  }f(x_i)=f^\ast(x_i)\]
+where $\operatorname{Lip}(f)=\min_\beta |f(x)-f(y)|\le\beta\|x-y\|$. To do this, we can, for any $x$, pick the closest data point $x_k$ to $x$ and notice 
+\[|\hat f(x)-f^\ast(x)|\le|\hat f(x)-\hat f(x_k)|+|\hat f(x_k)-f^\ast(x_k)|+|f^\ast(x_k)-f^\ast(x)|\le2\operatorname{Lip}(f^\ast)\|x-x_k\|.\]
+
+But now 
+\[\bbE_{x\sim\nu}|\hat f(x)-f^\ast(x)|^2\le\bbE_{x\sim\nu}\|x-x_k\|^2\approx\varepsilon^2.\]
+
+\begin{defn}
+	The \textbf{maximum discrepancy} of a function class $\calF$ is 
+	\[\tilde R_n(\calF)=\sup_{f\in\calF}\frac{1}{2n}\sum_1^{n/2}f(x_i)-\sum_{n/2+1}^n f(x_i))\]
+\end{defn}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Reinforcement Learning}
+This series of lectures was given by Emma Brunskill from Stanford's CS department.
+
+\subsection{Introduction}
+There is a wide set of backgrounds in this class will start broad and get more focused towards the end.
+One can think of reinforcement learning as being at the intersection of supervised learning and bandits (and beyond).
+In a sentence,
+reinforcement learning is ``learning to make good sequences of decisions under uncertainty.''
+
+\subsubsection{Examples of things you can do}
+The main applications arise in a couple different domains: learning to play games (e.g. Atari games), teaching robots tasks, and in healthcare.
+There are a plethora of applications we will hear about more.
+
+\subsubsection{Tasks}
+There are four main aspects:
+\begin{itemize}
+	\item Optimization
+	\item Generalization
+	\item Exploration 
+	\item Delayed consequences
+\end{itemize}
+
+In general \textbf{AI learning} uses all of these except exploration. Although RL was used in solving go, it wasn't needed.
+
+\textbf{Deep learning} only exploits optimization and generalization. There is also a sense of \textbf{imitation learning}, which also omits exploration.
+
+We will be following the following (rough) plan:
+\begin{enumerate}
+	\item Markov decision processes
+	\item Q-learning
+	\item RL and generalization
+	\item Exploration in sequential domains 
+	\item Batch/counterfactual reinforcement learning
+\end{enumerate}
+
+\subsection{Markov Decision Processes}
+The big idea here is that two entities: the agent and the world. The agent takes an action and the world offers a reward. In the case of bandits, this is where the scenario stops,
+but here in RL, we continue the feedback loop. Our goal here is to maximize the expected discounted sum of rewards.
+
+Some ideas that demonstrate the delayed consequences are as follows: imagine teaching a robot to empty the dishwasher---it only receives a positive reward once its task is complete.
+
+Now one of the hardest aspects here is coming up with a \textbf{reward function}. That is, a real-valued map $r(s,a)$ given a state and action. Why is this hard?
+Well, imagine you are trying to teach students math and the fact of the matter is that teaching addition is easier than subtraction, and we design our reward to be +1 if the student is correct and -1 if incorrect.
+Then the machine is naturally going to give only additon questions (and infinitely many of them). This is an example of \textit{reward hacking.}
+
+A problem follows \textbf{Markov dynamics} if 
+\[\bbP(s_t|s_{t-1},a_{t-1},s_{t-2},a_{t-2},\cdots)=\bbP(s_t|s_{t-1},a_{t-1}).\]
+\begin{defn}
+	A \textbf{Markov decision process} is a set $\langle S,A,R,T,\gamma\rangle$ where 
+	\begin{itemize}
+		\item $S$ is a set of states,
+		\item $A$ is a discrete action space,
+		\item $R$ is a set of rules of the form $r(s,a)$ or $r(s)$ or $r(s,a,s')$,
+		\item $T$ is a collection of distributions $\bbP(s'|s,a)$ and $\gamma\in(0,1)$ is a parameter.
+	\end{itemize}
+
+	Then we are trying to learn a \textbf{policy} $\pi:S\to A$. Note this could also be stochastic.
+\end{defn}
+
+There are a couple ways to do this: first, is the \textbf{model-based approach}. Here we directly estimate and use $R$ and $T$. Another is \textbf{value-based:}
+here we consider the value function 
+\[V^\pi=\bbE_{s'\sim pi}\left[\sum_0^\infty\gamma^tr_t|s_{t=0}=s_0\right].\]
+Finally we can perform a \textbf{policy-based approach} by computing $\argmax_{\pi\in\Pi}V^\pi$.
+
+Another parameter we have is $H$, the \textbf{time horizon} on the process. This tells us the number of steps that we may take before the system is reset to $s_0$.
+
+Then this process will return
+\[G^\pi=r_{0}+\gamma r_{1}+\gamma^2r_2+\cdots+\gamma^{H-1}r_H.\]
+Notice that in general this is going to be a random variable! That;'s why $V^\pi(s_0)=\bbE[G^\pi|s_0]$.
+
+\subsubsection{Estimation}
+Now say we know all the parameters of our MDP. Then we can use $S$ as a lookup table to approximate $V^\pi$ by ``running simulations.'' That is, we essentially 
+have a black box that outputs specific measurements of $G$. We don't need the Markov assumption here since we aren't touching the distributions directly.
+
+Using the Markov assumption yields additional structure, however. Consider that 
+\[V^\pi(s)= r(s,\pi(s))+\gamma\sum_{s'}p(s|s,\pi(s))V^\pi(s')\]
+giving us our immediate and discounted future rewards.
+
+Now let $V^\ast=\max_{\pi}V^\pi$ and $\pi^\ast=\argmax_\pi V^\pi$. But then using the above equation,
+\[V^\ast(s)=\max_a r(s,a)+\gamma_{s'}p(s'|s,a)V^\ast(s')\]
+which induces a kind of dynamical programming structure on the problem. This is called the \textbf{Bellman equation.}
+
+Then we proceed by a process called \textbf{value iteration}: let $V_0(s)=0$ for all $s$. Then for all $k$, set 
+\[V_k(s)=\max_a r(s,a)+\gamma\sum_{s'}p(s'|s,a)V_{k-1}(s')\]
+Doing this is called a ``Bellman backup.'' Then we break once you converge: $\|V_k-V_{k-1}\|_\infty\le\varepsilon$.
+
+\begin{lem}
+	Assuming all rewards are valued in $[0,1]$, the Bellman backup is a contraction operator as long as the state and action spaces are discrete (i.e. the above process converges).
+\end{lem}
+\begin{prf}
+	Throughout we will be considering $\|V-V'\|_\infty=\max_s|V(s)-V'(s)|$. So we compute 
+	\begin{align*}
+		\|BV_k-BV_j\|&=\|\max_a(r(s,a)+\gamma\sum_{s'}p(s'|s,a)V_{k-1}(s'))-\max_{a'}(r(s,a')\\
+		&\qquad\qquad\qquad\quad +\gamma\sum_{s'}p(s'|a',s)V_{j-1}s)\|\\
+		&\le\|\max_a \gamma\sum p(s'|s,a)\|V_k-V_j\|\\
+		&\le \gamma\|V_k-V_j\|
+	\end{align*}
+	so as long as $\gamma<1$, our functions have gotten closer.
+\end{prf}
+\begin{rmk}
+	Notice this also implies that $V^\ast$ must be unique.
+\end{rmk}
+
+\subsubsection{MDP Policy Iteration}
+Value iteration is great, but we have to consider the set of functions $A^S$, exponential in the set of states. Instead we can do \textbf{Policy Iteration}:
+Set $i=0$ and initialize $\pi_0(s)$ randomly for all states $s$. Then while $i=0$, or $\|\pi_i-\pi_{i-1}\|_1\>0$, set $V^{\pi_i}$, the MDP
+value function policy evaluation of $\pi_i$ and let $\pi_{i+1}$ be the policy improvement.
+
+Note that we can find an analytic solution whenever you have a finite set of states and actions, otherwise you have to do the dynamic programming-type argument using the Bellman operator.
+\begin{defn}
+	The $Q$ value of a policy is 
+	\[Q^\pi(s,a)=r(s,a)+\gamma+\sum_{s'} p(s'|s,a)V^\pi(S')\]
+\end{defn}
+so now that we have computed $V^{\pi_i}$ using analytical or dynamic programming methods, this is a relatively simple step. Then we just pick
+\[\pi_{i+1}(s)=\argmax_aQ^{\pi_i}(s,a).\]
+
+\begin{rmk}
+	It ends up that $V^{\pi_{i+1}}\ge V^{\pi_i}$ with strict inequality whenever $\pi_i$ is suboptimal. Notice that this is not trivial! We are getting better optimization over all time, 
+	but since we are switching policies greedily at each step, there is something to prove.
+\end{rmk}
+
+\subsection{Reinforcement Learning and Q-learning}
+Now we are performing reinforcement learning in a contex where $R$ and $T$ are unknown, but we are still interested in finding $V^\ast$ and $\pi^\ast$.
+
+\subsubsection{Policy Evaluation}
+Here we are given $\pi$ and want to compute $V^\pi$ and $Q^\pi$.
+
+If \textbf{our horizon $H$ is finite}, then we can do \textbf{Monte Carlo Evaluation}. If we let (again)
+\[G^i=r_1+\gamma r^2+\cdots\]
+where $s,r\sim \pi$ starting with $s_0$. Then the value is 
+\[V^\pi(s_0)=\bbE_{s\sim \pi}[G|s_0]\]
+where the above does not require the Markov assumption and doesn't use the ``bootstrapping'' of MDP value iteration (where make incremental improvements to an imperfect hypothesis).
+Then the average of all runs acts as our estimator. This is unbiased but has high variance (although it converges by the law of large numbers).
+
+\subsubsection{Temporal Difference Learning}
+This is called ``model free'' in the sense that we are not directly estimating the dynamics and reward models. This combined ideas in Monte carlo and dynamic programming.
+It is episodic (finite horizon) or infinite $H$. Then each tuple $(s,a,r,s')$ can update $\hat V^\pi$ (instead of having to wait for an entire run).
+
+The MC estimator updates at each $s_0$:
+\[\hat V(s_0)+\alpha(G^i-\hat V^\ast(s_0))\]
+where $\alpha\in(0,1)$ is a learning rate. Then the idea we do here is 
+\[\hat V_{k+1}(s)=r(s,\pi(s))+\gamma\hat V_k(s')\]
+Notice that this uses bootstrapping  of $\hat V$ as well as doing a sampling MC to get the data to update with.
+This new estimator will be biased, but wil have lower variance of what we did before.
+
+So overall the TD learning we are doing looks like what follows:
+\[\hat V^\pi(s)=\hat V^\pi(s)+\alpha\left[r+\gamma\hat V^\pi(s')-\hat V^\pi(s)\right]\]
+where $r+\gamma\hat V^\pi(s')$ is called the \textbf{TD target} and this quantity minus $\hat V^\pi(s)$ is called the \textbf{TD error.}
+
+For mean estimation, let $X\in[0,1]$ is a random variable and let 
+\begin{thm}
+	Assume that $\alpha_n\in[0,1]$ and $\sum \alpha_i=\infty$ and $\sum \alpha_i^2<\infty$, and let $X_0,x_1,\dots$ be iid samples of $X$. THen define $\mu_n$ for each $n\in\bbN$ such that 
+	\[\mu_0=X_0,\quad \mu_{m+1}=(1-\alpha_m)\mu_m+\alpha_mx_m\]
+	then $\mu_m$ almost surely approaches $\bbE[X]$.
+\end{thm}
+\begin{rmk}
+	I can't really read and my probability isn't strong enough to figure out what she meant.
+\end{rmk}
+
+\subsubsection{Control}
+Compute $\pi^\ast$ and $V^\ast$.
+
+\subsection{Reinforcement Learning and Generalization}
+
+\subsection{Exploration in Sequential Domains}
+
+\subsection{Batch/Counterfactual RL}
 

### 2019-08-06 MSRI ML Notes/MSRI_ML_notes.tex
Tues the 2nd

@@
 
-\usepackage{amsmath, amsfonts, amssymb, graphicx, color, fancyhdr, lipsum, scalerel, stackengine, mathrsfs, tikz-cd, mdframed, enumitem, framed, adjustbox, bm, upgreek, xcolor, hyperref}
+\usepackage{amsmath, amsfonts, amssymb, graphicx, color, fancyhdr, lipsum, scalerel, stackengine, mathrsfs, tikz-cd, mdframed, enumitem, framed, adjustbox, bm, upgreek, xcolor, hyperref,bbm}
 \usepackage[framed,thmmarks]{ntheorem}
@@
 
+\subsection{Approximation}
+Now we ask ourselves: How can we disgn notions of regulatity that enable learning with $n<<\varepsilon^{-d}$?
+
+\subsubsection{Invariance/Symmetry}
+One of the key properties of these functions that helps us towards answering this questions are invariance and symmetry. Notice that for all $x$,
+\[|f^\ast(x)-f^\ast(x+\delta)\le\beta\|\delta\|\]
+since $f$ is assumed to be Lipschitz. But then if we let $T_\delta(x)=x+\delta$, this gets us that 
+\[f^\ast\approx f^\ast\circ T_\delta.\]
+This gets us a kind of stability under additive noise.
+
+Let $f^\ast:\calX\to\bbR$ where $\calX$ is ahigh dimensional space.
+Usually this is $\bbR^d$, but what about $L^2(\Omega)$?
+If we are considering the space of all images, we have things like translation in the (now two dimensional!) space.
+These don't require working in the high-dimensional system to understand.
+Some examples of transformations on a 2-D image are:
+\begin{itemize}
+	\item Translation
+	\item Rotation
+	\item other things
+\end{itemize}
+Then we can think of each $x\in\calX$ as a function of a simple variable $u$ and then $Tx(u)=x(tu).$ Of course these transformations come together to form groups in different ways:
+we have the translation group, for instance.
+
+Notice that for a well-chosen $x$, we can get that $\|x-Tx\|$ is on the order of $\|x\|$.
+So this is adding something that we didn't consider with Lipschitz functions.
+But if you do some dimensional analysis, you can see
+that really this gives us (at most) a 6-dimensional surface (affine transformations) of our high-dimensional space under which our objects should be stable.
+
+Instead we can sonsider non-rigid motions (non-affine) that should still preserve the meaning of the image. So we could define $x_\tau(u):=x(u-\tau(u))$
+under a reasonably nice tau. Define $\|\tau\|=\sup_u|\nabla\tau(u)|$ to be the \textbf{cost} of a deformation.
+\begin{defn}
+	A function $f^\ast$ is said to be \textbf{geometrically stable} if 
+	\[|f^\ast(x)=f^\ast(x_\tau)\le \|\tau\|\]
+	where the inequality above is approximate (omitting a constant perhaps).
+\end{defn}
+\begin{rmk}
+	The idea here is that for ``reasonably small'' deformations, the value of $f^\ast$ shouldn't change too much.
+\end{rmk}
+Something to take away: it is not the case that all problems we want to slolve are local ones.
+But if we expect that our functios are geometrically stable, it is more likely that a point's neighbors are closely related to it than arbitrary long-range ones.
+This notion is called \textbf{scale separation.}
+
+Then we can ask: \textit{How can we leverage geometric stability?} One (empirical) answer is to use something like a \textbf{convolutional neural network}.
+Here we take an element $x\in L^2(\Omega)$ and an $h\in L^2(\Omega)$ with local support.
+Then we compute the convolution:
+\[(x\ast h)(u)=\int_\Omega x(v)\cdot h(u-v)dv\]
+For instance, if $h$ is zero everywhere except a circle where it is 1, the convolution $x\ast h$ looks like the same image as $x$ except it has been ``smeared out'' in a circle of every point.
+
+Then if we have a series of filters $h_1,\dots,h_k$, we can define $y^i=\sigma(x\ast h_i(u))$ for all $i$ and for some activation function $\sigma$. This forms a single layer.
+Then by composing these into (say) $L$ layers, we get a neural net with parameters $h^1_1,\dots,h^1_k,h^2_1,\dots$.
+
+It ends up that these do a very nice job at detecting things that are similar under geometric transformations. But they are not provably so.
+
+\subsubsection{Convolutional Kernel Networks}
+Mairal, Harchasomi(?) and Bietti developed another method of doing this: we want to develop a kernel $K(x,x')$ such that $K(x,Tx)$ and $K(x,x-\tau)$ is large.
+See the paper for more details there.
+
+\subsubsection{Scattering Representation}
+B. and Mallat basically created a CNN with wavelet filters, which are objects from functional analysis. Wavelets are related to, but not the same as, Fourier transforms of functions.
+
+\subsection{Single Hidden Layer Neural Nets}
+Now we start with a vector $x\in\bbR^d$ and a $K$-dimensional hidden layer. Then we define 
+\[f_\theta(x):=\sum_{k=1}^Kc_k\sigma(\langle x,a_k\rangle+b_k)\]
+where $a_k\in\bbR^d$ and $b_k,c_k\in\bbR$ and $\sigma:\bbR\to \bbR$ is the activation function and the set of parameters is $\theta=(a_k,b_k,c_k)$.
+
+Define $H_\sigma=\{f_\theta\}\subseteq\{f:\bbR^d\to\bbR\}$.
+One may ask how expressive this set is, and the universal approximation theorem tells us that $H_\sigma$ is dense in the class of continuous functions as soon as $\sigma$ is not polynomial.
+The idea is as follows.
+Let $g_k(\langle x,a_k\rangle)$ be the $k^{th}$ summand in $f_\theta$ above.
+Such a function $\bbR^d\to\bbR$ that operates on a projection is called a \textbf{ridge function} (it dies off as you get farther away from the direction of $x$).
+Now if we let $\sigma(z)=e^{iz}$, then $g_k(x_=e^{i\langle x,a_k\rangle})$ is an element of the Fourier basis. Then we can write 
+\[f(x)=\sum_\xi \hat f(\xi)e^{i\langle x,\xi\rangle}\]
+where $\hat f(\xi)=\langle f,g_\xi\rangle$. The upshot here is that when $f$ is continuous we get that the higher weights $\hat f(\xi)$ for large $|\xi|$ die off,
+so we can take an approximation by truncation. Then one continues to approximate $\sigma_f$ with $\sigma$.
+
+\begin{rmk}
+	This is in a way a not very informative result in that it is not terribly surprising that we should be able to approximate continuous functions with finitely many parts.
+And really this doesn't answer questions about the amount of space or time we need to do such a thing practically.
+\end{rmk}
+
+\subsection{Reproducting Kernel Hilbert Spaces and Neural Networks}
+We are getting rather low on time here, so we will be doing a high-level approach to this topic.
+
+Consider $\phi(x,\theta)=\sigma(\langle x,a\rangle+b)$ to be a \textbf{neuron} where $\theta\in\nu$, a compact space. Let $\tau$ be a probability measure on $\nu$. Then define the set 
+\[\calF=\left\{f:\bbR^d\to\bbR:f(x)=\int_\nu\phi(x,v)\cdot p(v)\cdot\tau(dv)\right\}\]
+with $p\in L^2(\nu,dt).$ That is $\int_\nu |p(v)|^2\tau(dv)<\infty$.
+\begin{prop}[Bach '17]
+	$\calF$ above is a reproducting kernel Hilbert space with kernel 
+	\[K(x,x')=\int_\nu\phi(x,v)\phi(x',v)\tau(dv)\]
+\end{prop}
+\begin{defn}
+	A reproducting kernel Hilbert space is a functional space $\calF:X\to\bbR$ where function evaluation for $f\in\calF$ and $x\in X$ is 
+	\[f(x)=\langle f,\Phi_x\rangle\]
+	and the kernel $K(x,x')=\langle\Phi_x,\Phi_{x'}\rangle.$
+\end{defn}
+\subsubsection{Defining Complexity}
+We would like to get a notion of complexity for $f\in \calF$. One (natural) choice is to define 
+\[\gamma(f)^2=\inf\left\{\|p\|^2_{L^2}\text{ such that } f(x)=\int \phi(x,v)p(v)\tau(dv)\right\}\]
+
+Then Rahini and Recht (in 2008) showed that if $v_1,\dots,v_K$ are sampled iid from $\tau$, then we can adjust our model to look at functions of the form 
+$f(x)=\sum_1^Kc_i\phi(x,v_i)$ and the resulting function can be sucessfully used to approximate functions of this type.
+
+\subsubsection{Changing the scenario}
+Now consider the set of function on the $d$-sphere $\bbS^d$ and let $\sigma$ be ReLU ($\max(0,z)$) and let $\tau$ be the uniform distribution. Then 
+$G_2$ is the set of functions $f:S^d\to\bbR$ with $f(x)=\int_{\bbS^d}p(v)\sigma(\langle v,x\rangle)\tau(dv)$ where $p$ is $L^2$. Then we can write the kernel explicitly (Le Roux et al or Bach 17).
+
+Alternatively, we caon consider the functions that can be written as $\int\sigma \mu(dv)$ for any random measure with finite total variance ($\|\mu\|_{TV}=\sup_{\|\phi\|\le 1}\int\phi(v)\mu(dv))$.
+Then we can define the complexity $\gamma_1(f)$ to be the $\inf$ of $\|\mu\|_{TV}$ for all such ways to write $f$.
+Notice the first scenario is a subclass of this one.
+
+Consider the function $f(x)=\sigma(\langle x,w\rangle)$ for some fixed $w$.
+Notice that this function can't be found in the first scenario, since we don't allow point masses.
+That's why we need the flexibility of the latter case.
+More on this can be found in Bach's 2017 paper ``Breaking the curese of dimensionality with convex neural networks''.
+That opens us up to asking which questions actually require this depth.
+
+\subsubsection{When we need depth}
+In the literature, this is known as the \textit{depth separation problem}. In particular, the issue which is addressed is to find $f$ such that 
+\begin{enumerate}
+	\item $f$ cannot be well-approximated by by a shallow network with polynomial in $d$ nodes.
+	\item $f$ can be approximated by a deeper network with poly(d) nodes.
+\end{enumerate}
+
+Some of the main papers that address this question are: (Eldan and Shamir), (Safran and Shamir), (Telgarski), and (Daniely).
+We already know that if the information that is low-energy (in terms of the Fourier atoms with high coefficients) then
+the universal approximation theorem says that we can learn it with a shallow network. So the idea is to write a function $f(x)=g(\|x\|)$, which is not very illuminating practically.
+
+This question is not new. If we were to replace continuous functions with boolean functions on $\{0,1\}^d$, this question is related to the ``circuit lower bound'', 
+studied by Shannon, Alon, and Hastad. One of the current progras of research is translating results from one to the other.
+
+Finally, an answer to the question of depth separation could be related to adding a time parameter.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \begin{thm}
-	Assume that $\alpha_n\in[0,1]$ and $\sum \alpha_i=\infty$ and $\sum \alpha_i^2<\infty$, and let $X_0,x_1,\dots$ be iid samples of $X$. THen define $\mu_n$ for each $n\in\bbN$ such that 
+	Assume that $\alpha_n\in[0,1]$ and $\sum \alpha_i=\infty$ and $\sum \alpha_i^2<\infty$, and let $X_0,x_1,\dots$ be iid samples of $X$. Then define $\mu_n$ for each $n\in\bbN$ such that 
 	\[\mu_0=X_0,\quad \mu_{m+1}=(1-\alpha_m)\mu_m+\alpha_mx_m\]
@@
 \end{thm}
-\begin{rmk}
-	I can't really read and my probability isn't strong enough to figure out what she meant.
-\end{rmk}
 
 \subsubsection{Control}
-Compute $\pi^\ast$ and $V^\ast$.
-
+Now our goal is to compute $\pi^\ast$ and $V^\ast$ and then act according to our optimal policy.
+Note that algorithms for learning $\pi^\ast$ do not necessarily require a person to act according to that policy.
+This is called \textbf{off-policy learning}: learning about $V^{\pi_1}$ while acting using $\pi_2\ne\pi_1$.
+
+\begin{defn}
+	The process of \textbf{Q-learning} is as follows: set $Q(s,a)=0$ for all $s$ and $a$ and let $s_0$ be our initial state. Then for $t\in\bbN$,
+	select an action $a$ for the current state $s$. Then sample a reward $r\leftarrow r(s,a)$ and a new state $s'\sim p(s'|s,a)$ and then perform the ``update operation''
+	\[Q(s,a)\leftarrow Q(s,a)+\alpha(s,a)[r+\gamma\max_{a'}Q(s',a')-Q(s,a)].\]
+	Finally set $s$ to be $s'$.
+\end{defn}
+\begin{rmk}
+	$\alpha(s,a)$ is the \textbf{learning rate} and we allow this to depend on the current state and action.
+\end{rmk}
+\begin{rmk}
+	The whole goal here is to show that $Q(s,a)$ converges eventually to $Q^\ast(s,a)$.
+The fact that we use an approximation for $Q^\ast$ and that we sample instead of computing the expectation above shows that this is a stochasitic optimization problem.
+\end{rmk}
+
+In the field of stochastic optimization, we are trying to find stable points $H(x)=x$ where $H(x)$ can't be directly computed, but we have iid samples of $m$ noisy observations $H(x_i)+\omega_i$ with $\bbE[\omega_i]=0$.
+
+In \textit{The Foundations of Machine Learning}, we have the following theorem:
+\begin{thm}[FoML thm 17.17]
+	Let $H$ be a mapping $\bbR^N\to \bbR^N$. Let $\omega_t$ be a sequence of random variables in $\bbR^N$ and $\alpha_t$ a sequence of real numbers and $x_t$ a sequence defined by 
+	\[\forall s, 1\le s\le N:\quad x_{t+1}(s)=s_t(s)+\alpha_t(s)[H(x_t)(s)-x_t(s)+\omega_t(s)]\]
+	for some $x_0\in \bbR^N$.
+
+	Then define $\calF_t$ to be $\{(x_{t'})_{t'\le t},(\omega_{t'})_{t'\le t-1}, (\alpha_{t'})_{t'\le t}\}$.
+
+	Assume there are $K_1,K_2\in\bbR$ such that 
+	\[\bbE[\|\omega_t\|^2(s)|\calF_t]\le K_1+K_2\|x_t\|^2\]
+	for some $\|\cdot\|$ and that $\bbE[\omega_t|\calF_t]=0$ and for all $s\in[N]$ $\sum_{t=0}^\infty \alpha_t(s)=\infty$ and $\sum_0^\infty \alpha_t^2(s)<\infty$ and finally that $H$ is a $\|\cdot\|_\infty$ contraction with fixed point $X^\ast$.
+
+	Then $X_t\to X^\ast$ almost surely.
+\end{thm}
+They do not offer a proof of this in the book although it does exist out there. However we also have the next theorem:
+\begin{thm}[FoML thm 17.18]
+	Consider a finte MDP. Assum that for all $s,a$ $\sum_0^\infty \alpha(s,a)=\infty$ and $\sum_0^\infty \alpha_t(s,a)^2<\infty$. Then $Q$ learning converges to $Q^\ast$ with probability 1.
+\end{thm}
+\begin{rmk}
+	Notice that this is only a sufficient condition for convergence, but in this case the condition on the $\alpha_t(s,a)$ imply that you must ``visit'' each state-action pair infinitely often (even though you may know it is a bad action).
+	In practice you often don't do this.
+\end{rmk}
+
+So this is great---we can learn about the optimal policy without actually exploiting it, but how to we get to using $\pi^\ast$? Here we use Greedy in the Limit of Infinite Exploration (GLIE): say
+\[\lim_{i\to\infty}N_i(s,a)\to\infty\]
+where $N_i(s,a)$ is the number of visits to $(s,a)$ by time $i$ (this is baked into the above theorem). Then the behavior we want is a sequence of policies $\pi_i$ such that
+\[\lim_{i\to\infty}\pi_i(a|s)\to\argmax_aQ(s,a).\]
+
+We usually do this with a $\varepsilon$-greedy algorithm: with probability $1-\varepsilon$, set $\pi_i(a|s)=\argmax_aQ(s,a)$ and otherwise choose $a$ randomly.
+The $\varepsilon$-greedy is GLIE if we use $\varepsilon_i=\frac{1}{i}.$
+
+\subsubsection{Making Q-learning better}
+Notice that we ahve a small problem: $Q$ learning only updates $Q$ for the previous state, so propagating these changes back through a long string of decisions requires runing over the data several times.
+
+Say we have data $(s,a,r,s')$ for $i\in[n]$. Then set $y_i=r_i+\gamma\max_{a_i}\hat Q_{k-1}(s_i',a_i)\approx[B\hat Q_{k-1}](s_i,a_i).$
+Then if we say $D=[(s_i,a_i),y_i]$ and we can fit $Q_k(\theta)$ to minimize the loss on this dataset.
+
+The nice thing here is we can use linear regression or a deep network to fit this function to the data, but the problem is you lose convergence guarantees.
+This dissuaded people from using this technique in the past, but it has been shown in practice that this often gets you a decent approximation (e.g.
+Atari).
 \subsection{Reinforcement Learning and Generalization}
+We define PAC RL to be the case where for some $\varepsilon$ and $\delta$,
+\[\bbP(\sum_t \mathbbm{1}(V^{\pi_t}(s_t)<V^\ast(s_t)-\varepsilon)\le F(\delta, S, A, T))\ge 1-\delta\]
+for some bounding function $F$.
+
+In RL, regret is computed via
+\[\sum_t r(s_t,\pi^\ast(s_t))-\sum_tr(s_t,\pi_t(s_t))\]
+which we also would want to say is bounded by $F(\delta,S,A,T)$ with high probability.
+
+Now for \textbf{tabular episodic RL} we want to say that our state and action spaces are finite and we are operating for some time horizon $H$ and running one epsiode at a time (think doing one trial on each patient).
+
+\subsubsection{Model Based Interval Estimation}
+This was developed first by Strohl and Littman in 08.
+Then given $\varepsilon, S,$ and $m$.
+Then initialize $n(s,a,s')=0$ for all $s,a,s'$ and set $\hat r(s,a)=0$ and $\tilde Q(s,a)=\frac{1}{1-\gamma}$ for all $s,a,$.
+Finally let $s_t=s_0$. Then for $t\in\bbN$, set $a_t=\argmax \tilde Q(s_t,a).$ Then observe $r_t$ and $s_{t+1}$. Then set 
+\[\hat r(s,a)=\frac{1}{\sum_{s'}n(s,a,s')+1}\left(r_t+\sum_{s'}n(s,a,s')\hat r(s,a)\right)\]
+to be our empirical $r$. Finally we compute the empirical dynamics:
+\[\hat p(s'|s,a)=\frac{n(s,a,s')}{\sum_{s'}n(s,a,s')}.\]
+Then while we haven't converged, update 
+\[\tilde Q(s,a)=\hat r(s,a)+\gamma\sum_{s'}\hat p(s'|s,a)\max_{a'}\tilde Q(s,a')+\frac{\beta}{\sqrt{\sum_{s'}n(s,a,s')}}\]
+
+Here $\beta=\frac{1}{1-\gamma}\sqrt{(1/2)\ln(|S||A|m^2/\delta)}$. So the last term artificially inflates $\tilde Q$ if we haven't visited the state many times to encourage intelligent exploration.
+
+What are sufficient set of conditionf for PAC RL? Well,
+\begin{itemize}
+	\item \textbf{Optimism:} $\tilde Q_t(s,a)\ge Q^\ast(s,a)-\varepsilon$ for all $s,a,t$.
+	\item \textbf{Accuracy:} $V^{\pi_t}(s)=V_M^{\pi_t}(s)\le \varepsilon$ where $M$ is an MDP we'll talk about below.
+	\item \textbf{Bounded Learning Complexity:} We bound the total number of updates to $Q$ and the number of times we visit an ``unknown'' $(s,a)$ pair (bounded by $\theta(\varepsilon,\delta)$).
+\end{itemize}
+In the case that all three of these hold, then an RL algorithm that is executed on any MDP will follow an $\calO(\varepsilon)$-optimal on all but 
+\[\calO\left(\frac{\theta(\varepsilon,\delta)}{\varepsilon(1-\gamma)^2}\ln\frac{1}{\delta}\ln\frac{1}{\varepsilon(1-\gamma)}\right)\]
+steps with probability at least $1-2\delta$.
+
+The last item above has largely been removed in a lot of cases because they are pretty restrictive although they make the analysis significantly easier.
 

### 2019-08-07 MSRI ML Notes/MSRI_ML_notes.tex
Last day

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\pagebreak
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\pagebreak
 
@@
 
+\subsection{Optimization}
+Let $L$ be aloss function.
+Then if we construct our neurons $\phi_i$ and compose them to a neural net $\Phi$, $L\circ\Phi$ is a map from the space of inputs to $\bbR$ and we can solve the empirical regret minimization problem.
+\[\min_{\theta\in\Theta}E(\theta):=L(\phi|\theta)\]
+where $E=L\circ \Phi$.
+
+When $\Phi$ is linear, we are doing least squares in a functional space so we can just do kernel learning (we are in the convex scenario).
+What about when $\Phi$ is nonlinear?
+Generaically we have a non-convex problem---unlike in the convex setting, the result is now dependent on the starting state.
+Overall the results are brittle.
+Saddle points and poor local minima are in our way now.
+
+The rest of the section (and of today's lecture) will be focused on gradient descent. We will not spend time here focusing on the step size we need to take for good convergence. 
+\[\theta_{k+1}=\theta_k-\eta\nabla_\theta E(\theta_k)\]
+where here we are allowing this to be pretty general: this could be SGD and $E$ could be $\hat E_k$, a noisy version of loss.
+Recall from the convex optimization talks that as $\eta\to 0,$ we transition from gradient descent to gradient flow.
+To leverage the intuition in the continuous case, we are going to be assuming the learning rate $\eta$ is very small.
+
+Let $\Phi:\Theta\to \calF$ be a function from parameter space to function space. Say $\dot \theta(t)=-\nabla_\theta E(\theta(t))$. Then if $\phi(t)= \phi(\theta(t))$, we can write 
+\[\frac{d}{dt} \phi(t)=-D\phi^T|_{\theta(t)}\cdot D\phi|_{\theta(t)}\cdot L'(\phi(t))\]
+and our kernel is $k_t(x,x')=\langle\nabla_\theta\phi(\theta)(t),x),\nabla_\theta \phi(\theta(t),x')\rangle$. We can call this the \textbf{tangent kernel.}
+
+\subsubsection{Overparameterized Regimes}
+One case where we simplify our job is in a so-called overparameterized regime. Here we have a single hidden layer NN with neurons 
+\[\phi(\theta x)=\frac{1}{\sqrt{m}}\sum_1^m g(\theta_i,x)\]
+where $\theta_i$ is sampled iid from some $\mu_0$. Then we are considering when $m$ grows: If we compute the kernel:
+\[\langle \nabla\phi(\theta,x),\nabla\phi(\theta,x')\rangle=\frac{1}{m}\sum_1^m\nabla_\theta g(\theta_i,x)\cdot\nabla_\theta g(\theta_i,x')\]
+which is a random variable (due to $\theta_i$), but as $m\to\infty$ we get that this converges in expectation:
+\[K_\infty(x,x')=\bbE_{\mu_0}[\langle \nabla\phi(\theta,x),\nabla\phi(\theta,x')\rangle]\]
+and the analysis here was done by Rahimi and Recht in '08 as well as Bach in '17 in \textit{On the equivalence between random kernel expansions and quadrature rules.}
+
+\begin{thm}[Jacot, Ouglen, Wycot '18]
+	For all $T>0$ as $m\to\infty$, 
+	\[k_t(x,x')=\langle \nabla\phi(\theta(t),x),\nabla\phi(\theta(t),x')\rangle\to K_\infty(x,x')\]
+	uniformly on compact sets.
+\end{thm}
+\begin{rmk}
+	The surprising observation here is that something that is ostensibly time-dependent eventually forgets about time.
+\end{rmk}
+\begin{rmk}
+	The consequence here is that under the ``$1/\sqrt{m}$'' scaling, $\phi(\theta)$ behaves like a linear model.
+\end{rmk}
+An important consequence comes from injecting this observation back into gradient descent:
+\[\dot\phi(t)=-K\cdot(\phi(t)-f^\ast)\]
+and so by making sure our $K$ is nonsingular we can get something like linear convergence. We can make the kernel nonsingular by increasing the amount of data we use.
+This makes deep learning look easy!
+
+In a paper by Chizat, Oyallon, and Bach in `19, they defined what is called a \textbf{lazy training regime}: we take our function $\phi(\theta)$ (a non-linear model) and let $\theta_0$ be some initial condition.
+Then the \textbf{linearlized tangent model} is
+\[\bar\phi(\theta,x)=\phi(\theta_0,x)+\langle\theta-\theta_0,\nabla_\theta\phi(\theta_0,x)\rangle\]
+and we want to ask ourselves when these two models are similar.
+
+Let $E(\phi)=L(\phi(\theta))$. Assume that $E(\theta_0)>0$. Then a gradient step will give 
+\[\theta_1=\theta_0=\eta\nabla E(\theta_0)\]
+and then we want to compare the change in the objective relative to that in the linear model. In our case, we have 
+\[\Delta(E)=\frac{E(\theta_1)-E(\theta_0)}{E(\theta_0)}\approx \eta\langle\nabla E(\theta_0),\nabla E(\theta_1)\rangle\]
+using Taylor approximations.
+
+The relative change in the tangent model is computed as
+\[\Delta (D\phi)=\frac{\|D\phi(\theta_1)-D\phi(\theta_0)\|}{\|D\phi(\theta_0)\|}\le\approx \frac{\eta\|D^2\phi(\theta_0)\|\|\nabla E(\theta_0\|)}{\|D\phi(\theta_0)}\|\]
+and so 
+\[\Delta(D\phi)<<\Delta(E)\quad\Leftrightarrow\quad \frac{\|D^2\phi(\theta_0)}{\|D\phi(\theta_0)\|} << \frac{\|\nabla E(\theta_0)\|}{E(\theta_0)}.\]
+
+In particular when $L(f)=\|f-f^\ast\|^2$, we get that $\|\nabla E(\theta_0)\|\approx \|\phi(\theta_0)-f^\ast\|\cdot\|D\phi(\theta_0)\|$.
+
+\begin{defn}
+	Then the \textbf{relative scale of $\phi$ at $\theta$} is 
+	\[\kappa_\phi(\theta)=\|\phi(\theta)-f^\ast\|\cdot\frac{\|D^2\phi(\theta_0)\|}{\|\phi(\theta_0)\|^2}\]
+\end{defn}
+and then we are in the lazy regime if and only if $\kappa_\phi(\theta_0)<< 1.$
+
+\subsection{Mean Field and Beyond}
+From the last section (and lecture) we saw that were likely to see good results from the lazy regime in the case that this relative scale of $\phi$ was very small. In fact,
+\begin{thm}[Chizat O B `19 thm 2.3]
+	Assume that $\phi$ and $D\phi$ are Lipshitz around the initialization $\theta_0$.
+Let $\theta(t)$ be the trajectory of the nonlinear model and $\bar\theta(t)$ be the trajectory for the linearized model $\bar\phi(\theta)$.
+	Then for $t<c_\phi$, the following holds:
+	\[\frac{\|\phi(\theta(t))-\bar\phi(\bar\theta(t))\|}{\|\phi(\theta_0)-f^\ast\|}\le\approx t^2\cdot \kappa_\phi(\theta_0)\]
+\end{thm}
+\begin{rmk}
+	One thing we get from this is that the trajectory of the nonlinear model closely tracks that of the linearized model. And we get linear convergence in the latter, so this tells us something globally:
+\end{rmk}
+\begin{thm}[Ibid 2.4]
+	In the overparameterized regime, this can be made to be uniform over time.
+\end{thm}
+
+Define the neurons
+\[\phi_m(\theta,x)=\alpha(m)\sum_{i=1}^mg(\theta_i,x)\]
+where $\theta_i\sim  \mu_0$ are iid and $\bbE_{\mu_0}g(\theta)=0$. Set $\kappa_m=\bbE_\theta[\kappa_{\phi_m}(\theta)]$.
+
+Then we have some observations:
+\begin{itemize}
+	\item $\|\phi_m(\theta_0)-f^\ast\|\to\|f^\ast\|$
+	\item $\|D\phi_m(\theta)\|^2\simeq \alpha(m)^2\cdot m\bbE_\theta\|Dg(\theta)\|^2$
+	\item $\|D^2\phi_m(\theta)\|\simeq \alpha_m\bbE_\theta\|D^2 g(\theta)\|$
+\end{itemize}
+Using this, we see that $\kappa_m$ grows roughly like $\frac{1}{\alpha(m)\cdot m}\kappa_g$. Thus if $\alpha(m)=\frac{1}{m}$, we get back to the lazy regime. Notice a problem with this, however: 
+in practice when we train an NN we see that the features change a lot in the process. But this is not accounted for in the lazy regime.
+
+\subsubsection{Mean Field Model}
+Now we write 
+\[\phi(\theta)=\frac{1}{m}\sum_1^mg(\theta_i)\]
+where $\theta=(\theta_i)\in D^m$ is the set of parameters.
+In general, $L(\phi(\theta))$ has poor local minima for only finite $m$.
+Furthermore arriving at such poor local minima has nontrivial probability.
+But now we can ask what happens in the ``typical'' case when $m$ is sufficiently large.
+
+Let $L(f)=\|f-f^\ast\|^2$ and say $f$ is as above. Then 
+\[E(\theta_1,\dots,\theta_m)=\|f\|^2-\frac{2}{m}\sum_{i=1}^m\langle f(\theta_i),f^\ast\rangle +\frac{1}{m^2}\sum_{i,j}\langle g(\theta_i),g(\theta_j)\rangle\]
+where $\langle f,g\rangle=\bbE_{x\sim D}[f(x)\cdot g(x)]$. THen calling $F(\theta)=\langle g(\theta), f^\ast\rangle$,
+\[E(\theta_i)= C_{f^\ast}-\frac{2}{m}\sum_1^m f(\theta_i)+\frac{1}{m^2}\sum_{i.j}\kappa(\theta_i,\theta_j)\]
+
+Then there is an interesting paradigm shift where we can think of neurons $g(\theta)$ as particles $\theta\in D$. In this way we can think of us having particles in our parameter space $D$
+and look at their interactions using statistical physics!
+
+Define the measure on $D$: $\hat\mu_m=\frac{1}{m}\sum_1^n\delta_{\theta_i}$, just a sum of point masses. Then we can write 
+\[\phi(\theta,x)=\int g(\theta,x)\hat\mu_m(d\theta)\]
+but we can generalize this by picking any $\mu$ in the space of measures on $D$:
+\[f(x)=\int g(\theta,x)\mu(d\theta)\]
+and then $E[\mu]=L(f)=\|f-f^\ast\|^2$, but then 
+\[E[\mu]=C_f-2\int F(\theta)\mu(d\theta)+\iint\kappa(\theta,\theta')\mu(d\theta)\mu(d\theta')\]
+Notice that this looks like a nice convex function, but the problem is that we actually need to work with a metric that is non-Euclidean, and so this will still be non-convex in general.
+
+\subsubsection{Dynamics}
+Now we can think of the ``particle'' gradient via the Lagrangian:
+\[\dot\theta_i(t)=-\nabla_{\theta_i}E(\theta_1,\dots,\theta_m)\]
+From the Eulerian(?) perspective, 
+\[\frac{m}{2}\nabla_{\theta_i}E(\theta_1,\dots,\theta_m)=\nabla V(\theta_i;\mu_m)\]
+where $V$ is a potential field defined by 
+\[V(\theta;\mu)=-F(\theta)+\int\kappa(\theta,\theta')\mu(d\theta').\]
+
+Then a classical question in physics is to ask about what happens as the measure $\mu$ evolves under a time-varying velocity filed $\nabla V$. This is governed by the \textit{coninuity equation:}
+\[\partial_t\mu_t=\operatorname{div}(\nabla V(\theta;\mu_t)\mu_t)\]
+We can think of this as an incarnation of the divergence theorem. [Mei, Montamari, Nguyen] and [Chizat, Bach] and [Sirignan, Spilopoulos] all discussed this using 
+what is known as the \textbf{Wasserstein gradient flow.} \textit{Is the Wasserstein metric a solution to the above PDE?}
+
+\subsubsection{Final Questions}
+Finally we ask ourselves a few more questions. First, do we get convergence of the above PDE to global minimizers of $E$? 
+\begin{thm}[CB, REVE]
+	If $\mu_0$ is sufficiently smooth/diverse and $\mu_t\to \mu_\infty$, then $\mu_\infty$ is a global minimizer of $E$.
+\end{thm}
+Now notice that if we sample $\mu_0$ and get $m$ iid samples and construct $\hat\mu_{0,m}$ and then run time for both, getting $\mu_t$ from the continuity equation and doing the same for the empirical mean.
+But we should not expect in general that sampling from $\mu_t$ will get us the same thing (that is, the diagram doesn't necessarily commute).
+
+However, there is a well-studied phenomenon in molecular dynamics which says, in essense, the diagram \textit{does} commute! That is,
+\begin{thm}[CB, MMN]
+	In the limit $m,t\to\infty$, under the previous assumptions,
+	\[E[\hat\mu_{t,m}]\to \min_{\mu}E[\mu].\]
+\end{thm}
+
+\subsubsection{Takeaways}
+THe big thing to notice here is that the mean-field framework more closely mirrors what we see in real learning. The features are tuned during learning, the relative scales is large, etc.
+So one current topic of research is how to extend mean field to deeper networks.
+Also people are interesed in understanding the cause in overfitting, and in particular investigating what the impact of overparameterization is.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\pagebreak
 
@@
 
-\subsection{Exploration in Sequential Domains}
-
-\subsection{Batch/Counterfactual RL}
+\subsection{Counterfactual/Batch Reinforcement Learning}
+In this scenario, we say we have as data a set of $n$ trajectories, where the steps are taken according to some (unknown) probability distribution. 
+There has been a lot more intersection between the causal inference community and machine learnin community and this is an exciting application for such a thing!
+
+Then we can define \textbf{batch policy optimization} (finding a good policy that will perform well in the future): we solve the problem
+\[\argmax_{\pi\in\calH_i}\max_{\calH_i\in\{\calH_1,\calH_2,\dots\}}\int_{s\in S_0}\hat V^\pi(s,\calD)ds\]
+where the integral in the right comes from policy evaluation (e.g. Q-learning) and the left is optimizing the policy.
+Notice we could define our hypothesis space in a lot of different ways, so this is part of the difficulty.
+
+\subsubsection{Policy Evaluation}
+If we can get ghighly accurate estimates of the true value of $\pi$ ($V^\pi$), then this would help us greatly with our policy optimization.
+
+Most of the current literature focuses on a (single) binary decision. Here we are studying heterogeneous treatment effects, which we can think of as a two-armed bandit kind of problem.
+
+In MBIE(-EB), we computed estimators for reward and the distributions for state change.
+We then tried to go on to estimate the optimal policy, but we could use these kinds of methods to estimate the parameters of the problem.
+Unfortunately, these estimators could be biased and may be asymptotically biased.
+\[V^\pi\approx (1-\gamma\hat P^\pi)^{-1}\hat R^\pi\]
+
+We alsoe discussed model-free value function approximation, e.g. TD learning for the tabular cases or fitted $Q$ iteration. Here we set 
+\[\tilde Q^\pi(s_i,a_i)=r_i+\gamma V_\theta^\pi(s_{i+1})\]
+but unfortunately this has convergence issues although hthere have been some nice results in practice.
+
+Recall the idea of \textbf{importance sampling}: 
+\[\bbE_p[r]=\sum_x p(x)r(x)=\sum_x\frac{p(x)q(x)}{q(x)}r(x)\approx\frac{1}{N}\sum_{i=1,x\sim q}^N\frac{p(x_i)}{q(x_i)}r(x_i)\]
+and then applying this to RL, we want to evaluate
+\[V^\pi(s)=\sum_\tau p(\tau|\pi,s) \sum_1^H r_\tau(t)=\sum_\tau p(\tau|\pi,s)R(\tau)\]
+where $r_\tau(t)$ is the reward earned on the trajectory $\tau$ at the $t^{th}$ step. Then applying the above analysis to this problem, we get 
+\[V^\pi(s)\approx\frac{1}{N}\sum_{\tau\sim \pi_b}\frac{p(\tau|\pi,s)}{p(\tau|\pi_b,s)}R(\tau)=\frac{1}{N}\sum_{\tau\sim\pi_b}\frac{\prod_1^{H-1}p(s_{t+1}|s_t,a_t)p(a_t|\pi,s)}{\prod_1^{H-1}p(s_{t+1}|s_t,a_t)p(a_t|\pi_b,s)}R(\tau)\]
+and then some terms cancel! So adding in inticies more carefully:
+\[V^\pi(s)\approx \frac{1}{N}\sum_{i=1}^N\prod_{t=1}^{H-1}\frac{p(a^i_t|\pi, s^i)}{p(a^i_t|\pi_b,s^i)}R(\tau^i)\]
+\begin{rmk}
+	Notice that we have dropped any knowledge about the state changes and furthermore makes no use of the Markov assumption, so can be used more generally!
+\end{rmk}
+
+We can also consider \textbf{stationary importance sampling}: 
+\[V^\pi(s_0)=\int_{(s,a)}d_\pi(s,a)r(s,a)\]
+where $d_\pi$ is some stationary distribution over $(s,a)$ for the Markov chain induced by $\pi$ (of course now we need the Markov assumption). But then doing a similar trick,
+\[V^\pi(s_0)\approx\frac{1}{N}\sum_{s,a\sim\pi_b}\frac{d_\pi(s,a)}{d_{\pi_b}(s,a)}r(s,a)\]
+
+The problems of variance and bias discussed above can be remidated by taking a blend of these techinques, called \textbf{Doubly Robust (DR) estimation.} Here we take a model and IS-based estimator  of the form
+\[V_{DR}\hat V(s)=\rho(r-\hat R(s,a))\]
+where $\hat R$ is a model for the reward, $r$ is the reward received, and $\rho$ is $\frac{\pi_1(a|s)}{\pi_0(a|s)}$, the important sampling parameter. And of course 
+\[\hat V(s):=\sum_a\pi_1(a|s)\hat R(s,a).\]
+
+Relatively recently this was transplanted to RL by introducing a ``covariant term.''
+\[DR(D):=\sum_{i=1}^n\sum_{t=0}^\infty \gamma^tw_t^i R_t^{H_i}-\sum_{1}^n\sum_0^\infty\gamma^t(w_t^i\hat q^{\pi_i}(S_t^{H^i},A_t^{H_i})-w_{t-1}^i\hat v^{\pi_e})\]
+I think I missed a term at the end but you can look it up in a paper by Brunskill in 2016.
+
+Things are speeding up at the end so this is becoming less and less reasonable to keep up. :)
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Presentation}
+Our group (Angel, Ashwin, and I) are presenting on the last day of class on the paper \textit{Learnability Can Be Undecidable} by S.
+Ben-David, P.
+Hrube\v s, S.
+Moran, A.
+Shpilka, A.
+Yehudayoff.
+I will begin
+by introducing the concept of EMX learning, (monotone) compression schemes, and dimension and discuss the main result and its connection to the continuum hypothesis.
+After that, Angel will prove that weak
+EMX learning is equivalent to there being a weak compression scheme.
+Finally Ashwin will bring us home by discussing the connection between monotone compression and cardinality and discuss why there is no
+good\footnote{i.e. gives a criteria for ability to learn like VC dimension in PAC learning.} notion of dimension that holds for all modes of learning.
+
+\subsection{Introduction and Overview}
+This work makes connections between certain modes of learnability and cardinalities in order to leverage the undecidability of the continuum hypothesis in order to show that 
+there is some amount of uncertainty inherent in whether a specific problem is able to be suitably learned. The main result of this paper is as follows:
+\begin{thm}[BHMSY 19]
+	The EMX learnability of the family $\calF^\ast$ with respect to $\calP^\ast$ is independent of ZFC.
+\end{thm}
+\begin{rmk}
+	The definitions of EMX learnability, $\calF^\ast$ and $\calP^\ast$ are all below.
+\end{rmk}
+
+Furthermore, the authors investigate the idea of ``dimensions for learning''. 
+\begin{defn}
+	A function $D:\calF\to\bbN\cup\{\infty\}$ is a \textbf{dimension for learning} if it satisfies both of the following:
+	\begin{enumerate}
+		\item[1.] \textbf{Characterizes Learnability:} $\calF$ is learnable if and only if $D(\calF)<\infty$.
+		\item[2.] \textbf{Of finite character:} For every $d\in\bbN$ and $\calF$, $D(\calF)\ge d$ can be demonstrated by finitely many $f\in\calF$ and finitely many domain points.  
+	\end{enumerate}
+\end{defn}
+\begin{rmk}
+	Every known notion of dimension for learning (e.g. VC dimension) satsifies these conditions.
+\end{rmk}
 

### 2019-08-08 MSRI ML Notes/MSRI_ML_notes.tex
Update notes

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\pagebreak
 
@@
 
-\subsection{Introduction and Overview}
+\subsection{Overview of Results}
 This work makes connections between certain modes of learnability and cardinalities in order to leverage the undecidability of the continuum hypothesis in order to show that 
@@
 
+\subsection{Definitions and Preliminaries}
+To motivate and enable the discussion that follows, we need to remind you of/introduce you to some of the concepts we will be using.
+Recall that $\aleph_0$ is the cardinality of the integers and $\aleph_1$ is the next-largest cardinality.
+Furthermore, it can be shown that the powerset $2^{\aleph_0}$ is the cardinality of the reals.
+The \textbf{continuum hypothesis} states that
+\[2^{\aleph_0}=\aleph_1\]
+G\"odel and Cohen showed that the continuum hypothesis was \textit{independent of ZFC}, and therefore became one of the most well-known examples of 
+a problem that is \textbf{undecidable} (assuming that ZFC is consistent).
+
+\subsubsection{Our Toy Problem}
+We will work with a specific problem to generate our proofs.
+Throughout the rest of the discussion, let $X$ be some domain set and let $\calF^\ast:X\to\{0,1\}$ be a family of functions with finite support.
+Note that we can (and often will) leverage the
+equivalence between functions $X\to\{0,1\}$ and finite subsets of $X$. 
+\begin{rmk}
+	In all that follows one can use the family $\calF$ of (not-necessarily-finitely-supported) functions $X\to\{0,1\}$, but we will be using $\calF^\ast$ for our arguments in the next parts, so we might as well restrict our attention now.
+\end{rmk}
+\begin{defn}
+	A \textbf{learner} on $\calF^\ast$ is a function $G:\cup_1^\infty X^i\to \calF$ that outputs a set in $X$ given a finite sequence of points in $X$.
+\end{defn}
+Then the goal of our game is to find a learner that takes an iid sample of points from our data distribution and returns a set that maximizes mass under this distribution.
+That is, we are looking to find $\operatorname{Opt}_P(\calF^\ast)=\sup_{f\in\calF^\ast}\bbE_P(f)$.
+\begin{rmk}
+	Notice that here we have to assume that all our distibutions are finitely-supported in order for this to make sense for a finite set.
+\end{rmk}
+\begin{defn}
+	We say $\calF^\ast$ is \textbf{$(\varepsilon,\delta)$-EMX-learnable} if there exists a learner $G$ that is an \textbf{$(\varepsilon,\delta)$-EMX-learner} for $\calF^\ast$; that is, such that 
+	for some $d=d(\varepsilon,\delta)$
+	\[\operatorname*{Pr}_{S\sim P^d}[\bbE_P(G(S))\le\mathrm{Opt}_P(\calF^\ast)-\varepsilon]\le\delta\]
+	for every (finitely-supported) distribution $P$ on $X$.
+\end{defn}
+
+\subsubsection{Monotone Compression Schemes}
+Compression schemes have long been studied in relation to learnability, and the results in this paper leverages that significantly.
+The intuition for a compression scheme is to find maps $\kappa$ and $\eta$, the first of which downsamples a sequence of data and the latter of which reconstructs the label for any point in the original sample.
+
+In this we restrict our attention here to \textbf{monotone compression schemes:}
+\begin{defn}
+	An $m\to d$ monotone compression scheme for $\calF^\ast$ is a function $\eta:X^d\to\calF^\ast$ such that for every $f\in\calF$ and $x_1,\dots,x_m$ there exists 
+	a subset $x_{i_1},\dots,x_{i_d}$ such that 
+	\[\{x_1,\dots,x_m\}\subseteq \eta(x_{i_1},\dots,x_{i_d})\]
+\end{defn}
+
 \end{document}

### 2019-09-27 Algebraic Geometry/Algebraic-Geometry.tex
update

\documentclass[12pt]{article}

\usepackage{setspace}

\usepackage{amsmath, graphicx, color, fancyhdr, tikz-cd, mdframed, enumitem, framed, adjustbox, bbm, upgreek, xcolor, hyperref, manfnt}
\usepackage[framed,thmmarks]{ntheorem}
\usepackage[style=alphabetic, bibencoding=utf8]{biblatex}
%Set the bibliography file
\bibliography{sources}

\usepackage[T1]{fontenc}
\usepackage[urw-garamond]{mathdesign}
\usepackage{garamondx}

%Replacement for the old geometry package
\usepackage{fullpage}

%Input my definitions
\input{../mydefs.tex}

%Shade definitions
\theoremindent0cm
\theoremheaderfont{\normalfont\bfseries} 
\def\theoremframecommand{\colorbox[rgb]{0.9,1,.8}}
\newshadedtheorem{defn}[thm]{Definition}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%% Customize Below %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%header stuff
\setlength{\headsep}{24pt}  % space between header and text
\pagestyle{fancy}     % set pagestyle for document
\lhead{Notes on the Schur-Weyl Functor} % put text in header (left side)
\rhead{Nico Courts} % put text in header (right side)
\cfoot{\itshape p. \thepage}
\setlength{\headheight}{15pt}
%\allowdisplaybreaks

% Document-Specific Macros
\DeclareMathOperator{\1}{\mathbbm{1}}
\DeclareMathOperator{\GL}{GL}
\DeclareMathOperator{\Sym}{Sym}

\begin{document}
%make the title page
\title{Algebraic Geometry\vspace{-1ex}}
\author{A course by Max Lieblich\\
Notes by Nico Courts}
\date{Autumn 2019/ Winter and Spring 2020}
\maketitle

\begin{abstract}
	A three-quarter sequence covering the basic theory of affine and projective 
	varieties, rings of functions, the Hilbert Nullstellensatz, localization, and 
	dimension; the theory of algebraic curves, divisors, cohomology, genus, and the 
	Riemann-Roch theorem; and related topics. 
\end{abstract}

\section{September 25, 2019}
The first thing that one asks is ``what is geometry?'' One needs to be able to answer this 
question before they define AG. One idea is that geometry is topology + structure.

\subsection{What is Geometry?}

\begin{ex}
	Exotic differentiable structures on a sphere. There are many different smooth structures,
	all of which are independent of the topology,

	$S^1\times S^1$ has infinitely many complex structures (remember the parallelograms)!
\end{ex}

How to you go about defining the geometry of a thing? One idea from manifolds: charts. These 
describe the local models and the interesting part is how this comes together to a whole space.

There is another idea to capture the ``local'' model of geometry that underlies modern algebraic geometry:
consider the map $\varphi:W\to W'\in\bbC\bbP^n$ and then say that this map is $C^\infty$ if and only if its coordinate
functions are. But the coordinate functions are problematic, so we can replace it with the following idea:

$\phi:W\to W'$ is $C\infty$ if and only if for all $C^\infty$ functions $f:W'\to \bbR$,
the composition
\[\varphi^\ast f=f\circ \varphi:W\to \bbR\]
is $C^\infty$.

To capture the manifold structure on $M$, it is equivalent to know the set of $C^\infty$ fucnctions $U\to \bbR$ for every open $U\subseteq M$.

\subsection{The Big Idea}
So then the idea we are talking away here is that \textit{geometry is in the functions} that exist on a particular space!

Fix a field $k$. 
\begin{defn}
	A \textbf{space with functions} is a topological spae $X$ along with a collection (a $k$-algebra!) $\calO(U)$ of maps $U\to k$ for each open 
	$U\subseteq X$.

	$\calO(U)$ are called \textbf{regular functions} and must satisfy:
	\begin{itemize}
		\item Given an open cover $U_\alpha$ of $U$, a function is regular if and only if its restrictions to each element of the cover is regular.
		\item If $f:U\to k$ is regular, then $D(f)=\{u\in U|f(u)\ne 0\}$ is an open set and $\frac{1}{f}\in\calO(D(f))$.
	\end{itemize}
\end{defn}

For the next time, try to think of as many examples of this as you can. Next time will be a mind blowing example of a variety.

\end{document}

### 2019-09-27 Algebraic Groups/Algebraic-Groups.tex
update

\documentclass[12pt]{article}

\usepackage{setspace}

\usepackage{amsmath, graphicx, color, fancyhdr, tikz-cd, mdframed, enumitem, framed, adjustbox, bbm, upgreek, xcolor, hyperref, manfnt}
\usepackage[framed,thmmarks]{ntheorem}
\usepackage[style=alphabetic, bibencoding=utf8]{biblatex}
%Set the bibliography file
\bibliography{sources}

\usepackage[T1]{fontenc}
\usepackage[urw-garamond]{mathdesign}
\usepackage{garamondx}

%Replacement for the old geometry package
\usepackage{fullpage}

%Input my definitions
\input{../mydefs.tex}

%Shade definitions
\theoremindent0cm
\theoremheaderfont{\normalfont\bfseries} 
\def\theoremframecommand{\colorbox[rgb]{0.9,1,.8}}
\newshadedtheorem{defn}[thm]{Definition}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%% Customize Below %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%header stuff
\setlength{\headsep}{24pt}  % space between header and text
\pagestyle{fancy}     % set pagestyle for document
\lhead{Algebraic Groups} % put text in header (left side)
\rhead{Notes by Nico Courts} % put text in header (right side)
\cfoot{\itshape p. \thepage}
\setlength{\headheight}{15pt}
%\allowdisplaybreaks

% Document-Specific Macros
\DeclareMathOperator{\1}{\mathbbm{1}}
\DeclareMathOperator{\GL}{GL}
\DeclareMathOperator{\Sym}{Sym}

\begin{document}
%make the title page
\title{Algebraic Groups\vspace{-1ex}}
\author{A course by Jarrod Alper and Julia Pevtsova\\
Notes by Nico Courts}
\date{Autumn 2019/ Winter and Spring 2020}
\maketitle

\begin{abstract}
	The topic of algebraic groups is a rich subject combining both group-theoretic and algebro-geometric-theoretic techniques. Examples include the general linear group $GL_n$, 
	the special orthogonal group $SO_n$ or the symplectic group $Sp_n$. Algebraic groups play an important role in algebraic geometry, representation theory and number theory.

	In this course, we will take the functorial approach to the study of linear algebraic groups (more generally, affine group schemes) equivalent to the study of Hopf algebras. 
	The classical view of an algebraic group as a variety will come up as a special case of a smooth algebraic group scheme. Our algebraic approach will be independent (even complementary) to the analytic approach taken in the course on Lie groups.
\end{abstract}

\section{September 25, 2019}
\subsection{Group objects}
Let $\calC$ be a category with a final object and finite products.
\begin{defn}
	A \textbf{group object $G$ in $\calC$} is an object in $\calC$ along with multiplication, identity, and inverse morphisms
	satisfying the usual axioms.
\end{defn}

One thing is that we are using that there is a final object $\ast$ along with our identity morphism $e:\ast\to G$.
Here Jarrod explictly used the fact that there is a unique map to $\ast$.

\begin{ex}
	If $\calC$ is $\operatorname{Set}$, then $G$ is a group. If $\calC=\operatorname{Top}$, then $G$ is a topological group, smooth manifolds give Lie groups, and finally (interesting to us):
\end{ex}
\begin{defn}
	Let $S$ be a scheme and let $\calC$ be the category of schemes over $S$. Then a group object $G$ in $\calC$ is 
	a \textbf{group scheme over $S$.}
\end{defn}

WHen $k$ is a field and $\calC$ is schemes of finite type over $k$, we get a group scheme of finite type over $k$. There is not a great consensus on what makes an \textbf{algebraic group}, 
but this is what we will use.

When we instead restrict to \textit{affine schemes} we get an affine groupe scheme of finite tipe over $k$, or a \textbf{linear algebraic group.}

\subsection{Examples}
$\bbG_m=\operatorname{Spec} k[t]_t$ is one.

If we consider the map $f:\bbG_m\to \bbG_m$ which on the level of elements sends $t\mapsto t^p$, the kernel is 
\[\mu_p=\ker(f)=\operatorname{Spec}k[t]/(t^p-1)\]
and that's great, but when $\ch k=p$, this causes the group scheme to be \textbf{unreduced}. This is (apparently) a case when you need to use schemes.

\subsection{The Functorial Approach}
Let $\calC$ be a category with object $X$. Define the functor $h_X:\calC^{op}\to \mathbf{Set}$ where 
\[h_X(Y)=\Hom_\calC(Y,X).\]

Then we have 
\begin{lem}[Yoneda]
	Let $G:\calC^{op}\to\mathbf{Set}$ be a functor. There is a natural bijection
	\[G(X)\simeq \operatorname{Nat}(h_X,G).\]
\end{lem}
\begin{prop}
	A group object $G$ in $\calC$  is the same as an aobject $X\in\calC$ together with a choice of factorization of 
	$h_X:\calC\to\mathbf{Set}$ through $\mathbf{Grp}$.
\end{prop}

\subsection{Exercises}
\begin{enumerate}
	\item Spell out all the details of the proof of the above propositon.
	\item Given a group object $G$, define in two ways what it means for it to act on another object. (In coordinates and functorially).
\end{enumerate}

\subsection{Some Interesting Facts}
If we had to write down five results that we'd like to get out of this class:
\begin{prop}
	Every affine group scheme of finite type over a field embeds into $GL_n$ as a closed subgroup.
\end{prop}
\begin{thm}[Cheballey's Theorem]
	Let $G$ be a finite type group scheme over a field. Then it factors as 
	\[1\to H\to G\to A\to 1\]
	where $A$ is abelian and $H$ is affine (linear algebraic).
\end{thm}
\begin{prop}
	If $G$ is an affine group scheme of finite type over $k$, then we have af actorization
	\[1\to U\to G\to R\to 1\]
	where $U$ is unipotent and $R$ is reductive.
\end{prop}
\begin{prop}
	$H\subseteq G$  a subgroup scheme. Then $G/H$ is a projective scheme.
\end{prop}
Finally we want to talk about Tanakka duality and how the representations of $G$ define $G$ itself.

\end{document}

### 2019-09-27 Algebraic Geometry/Algebraic-Geometry.tex
Bump

@@
 \pagestyle{fancy}     % set pagestyle for document
-\lhead{Notes on the Schur-Weyl Functor} % put text in header (left side)
-\rhead{Nico Courts} % put text in header (right side)
+\lhead{Algebraic Geometry} % put text in header (left side)
+\rhead{Notes by Nico Courts} % put text in header (right side)
 \cfoot{\itshape p. \thepage}
@@
 \begin{defn}
-	A \textbf{space with functions} is a topological spae $X$ along with a collection (a $k$-algebra!) $\calO(U)$ of maps $U\to k$ for each open 
+	A \textbf{space with functions} is a topological space $X$ along with a collection (a $k$-algebra!) $\calO(U)$ of maps $U\to k$ for each open 
 	$U\subseteq X$.
@@
 
+\section{September 27, 2019}
+\begin{prob}
+	Do all the exercises in Kempf chapter 1!
+\end{prob}
+
+For now we assume that $k$ is algebraically closed.
+
+\subsection{Examples of spaces with functions}
+There were lots of suggestions, but here are a couple:
+\begin{ex}
+	Let $X=\bbS^2$ and let $\calO_X^{cts}$ be the continuous $\bbC$-valued functions. Alternatively
+	we could consider a different sheaf $\calO_X^{an}$, the holomorphic functions. Or we could 
+	consider $\calO_X^\infty$, the $C^\infty$ functions (under some smooth structure).
+\end{ex}
+
+\begin{defn}
+	A \textbf{morphism} of spaces with functions between $(X,\calO_X)$ and $(Y,\calO_Y)$ is a continuous map 
+	$\varphi:X\to Y$ such that for all $U\subseteq Y$ open and $f\in\calO_Y(U)$, the function 
+	\[\phi^\ast f=f\circ\phi|_{\phi^{-1}(U)}:\phi^{-1}(U)\to k\in \calO_X(\phi^{-1}(U))\]
+
+	In other words, a morphism of spaces with functions is a map of spaces that \textit{respects the regular functions.}
+\end{defn}
+\begin{ex}
+	Let $X,Y$ be topological spaces and let $\calO_X$ and $\calO_Y$ be the continuous functions. Then morphisms are just continuous maps.
+\end{ex}
+\begin{ex}
+	When $X$ and $Y$ are manifolds and $\calO_\bullet$ are complex-valued functions, then the maorphisms are maps 
+	of manifolds.
+\end{ex}
+
+So now we return to the examples we saw before: $(\bbS^2,\calO^\infty)$, $(\bbS^2,\calO^{cts})$, and $(\bbS^2,\calO^{an})$.
+A natural question to ask is when we have morphisms between these spaces to see if there exist ones that are the identity on $\bbS^2$.
+
+Consider the identity map from the continuous to the analytic functions. Then take any map $f\in\calO^{an}$ and consider that 
+\[f=f\circ id_{id^{-1}(U)}:U\to k\in\calO^{cts}(U)\]
+and there is no map in the other direction.
+\begin{rmk}
+	Notice that since we are pulling functions back, the maps go in the opposite direction as you may think at first.
+\end{rmk}
+
+We can also talk about \textbf{open subspaces}. If $V\subseteq X$ is an open subset, we can let the induced space with functions 
+be $(V,\calO_V)$ where if $U\subseteq V$ then $\calO_V(U):=\calO_X(U)$.
+
+\subsection{Varieties}
+\begin{defn}
+	An \textbf{affine $k$-variety} is a space with functions $(Y,\calO_Y)$ such that for every space with functions $(X,\calO_X)$,
+	the natural map 
+	\[\Hom((X,\calO_X),(Y,\calO_Y))\to\Hom_{\Alg}(\calO_Y(Y),\calO_X(X))\]
+	is a bijection and furthermore $\calO_Y(Y)=:k[Y]$ is a finitely generated $k$-algebra.
+\end{defn}
+\begin{rmk}
+	The idea here is that the algebra maps (on the right) are precisely the same as the geometry maps (on the left).
+	Algebraic geometry, baby.
+\end{rmk}
+So then this leads to a very simple (loose) definition:
+\begin{defn}
+	A \textbf{variety} is something that is covered by affine varieties.
+\end{defn}
+
+\begin{ex}
+	$\bbA^1=k$. Give this space the cofinite topology. Then if we have $U=k\setminus\{x_1,\dots,x_n\}\subset \bbA^1$,
+	\[\calO_{\bbA^1}(U)=\{f(t)\in k(t)| \text{poles are in }\{x_i\}\}\]
+\end{ex}
+\begin{prob}
+	Show that $\bbA^1$ is an affine variety!
+\end{prob}
+\begin{rmk}
+	Notice that this staetment is equivalent to saying that any morphism of spaces with functions gives us a regular map $X\to k$.
+\end{rmk}
+
 \end{document}

### 2019-09-27 Algebraic Groups/Algebraic-Groups.tex
Bump

@@
 \DeclareMathOperator{\Sym}{Sym}
+\DeclareMathOperator{\Spec}{Spec}
 
@@
 \title{Algebraic Groups\vspace{-1ex}}
-\author{A course by Jarrod Alper and Julia Pevtsova\\
+\author{A course by Jarod Alper and Julia Pevtsova\\
 Notes by Nico Courts}
@@
 \end{prop}
-\begin{thm}[Cheballey's Theorem]
+\begin{thm}[Chevalley's Theorem]
 	Let $G$ be a finite type group scheme over a field. Then it factors as 
@@
 
+\section{September 27th, 2019}
+Last time we defined a group scheme (a group object in the category of schemes over a base scheme). We also mentioned that 
+You could define it as a map $h_G:\mathbf{Sch}/S\to \mathbf{Set}$ along with a factorization through $\mathbf{Grp}$.
+
+We defined an \textbf{algebraic group} over $k$ as a group scheme over $\operatorname{Spec} k$ of finite type and a \textbf{linear algebraic group}
+to be an \textit{affine} group scheme over $k$ of finite type.
+
+\subsection{Hopf Algebras}
+Let $G=\Spec A$ be a linear algebraic group over $k$. I have seen most of these before (see Waterhouse or my Hopf algebra notes)
+\begin{rmk}
+	One think I haven't seen explicitly before: Notice that the augmentation ideal $\ker \varepsilon$, where $\varepsilon$ is the counit,
+	is the (maximal!) ideal corresponding in the algebro-geometric sense to the identity element in $G$.
+\end{rmk}
+\begin{defn}
+	A \textbf{Hopf algebra} is ...
+\end{defn}
+\begin{defn}
+	Let $G$ be an algebraic group over $k$. Then if $h_G$ factors through $\Ab$, $G$ is called \textbf{commutative.}
+\end{defn}
+
+\subsection{Some Examples}
+\begin{rmk}
+	Note that to define a functor from schemes over $k$, is suffices to define it on affine schemes, thereby defining 
+	the (Zariski) local behavior of any such map. Thus we really only need to consider maps in $\Alg$.
+\end{rmk}
+\begin{itemize}
+	\item $\bbG_a$.
+Here we can define it as a functor that sends $S\mapsto\Gamma(S,\calO_S)$.
+Geometrically, $\bbG_a=\bbA^1$ where the multiplication is addition, inverses send $x\mapsto -x$ and the unit is the zero map.
+	The Hopf algebraic picture is the usual dual thing.
+	\item $\bbG_m$ as a scheme isthe map $S\mapsto \Gamma(S,\calO_S)^\ast$. In the geometric picture, $\bbA^1\setminus\{0\}$ and the algebra structure comes from multiplciation. Hopf is pretty easy.
+	\item $\GL_n$ is a scheme that sends
+	\[S\mapsto \left\{A=(a_{ij}): a_{ij}\in\Gamma(S,\calO_S), \det(A)\in\Gamma(S,\calO_S)^\ast\right\}\]
+	the algebra is $\bbA^{n\times n}\setminus \{\det = 0\}$ with the usual multiplication. The coalgebra structure can be seen in the book.
+\end{itemize}
+This one requires some more explaination so I am setting it apart.
+\begin{ex}
+	Let $V$ be a finite dimensionatl vector space over $k$. Then we can define the algebraic group $V_a$ which sends 
+	\[S\mapsto \Gamma(S,\calO_S)\otimes_k V.\]
+	Geometrically we are looking at $\bbA(V)=\Spec\Sym ^\ast V^\vee\simeq \Spec k[x_1,\dots,x_n]$ where $n=\dim V$.
+\end{ex}
+
+What about finite groups? As a scheme, we want $G=\sqcup_{g\in G}\Spec k$. The functor sends $S\mapsto \operatorname{Mor}_{\operatorname{Set}}(\pi_0(S),G)$,
+or maps from the connected components into $G$.
+
+\begin{ex}
+	Now consider the $n^{th}$ roots of unity: as a scheme, $\mu_n=\Spec k[t]/(t^n-1)\subseteq \bbG_m$.
+	If both $k=\bar k$ and $\ch k\nmid n$, then $\mu_n\cong\bbZ/n\bbZ$.
+
+	But if (e.g.) $k=\bbQ$, then $\mu_3$ is $\bbQ[t]/(t^3-1)=\Spec\bbQ\sqcup \Spec\bbQ(\xi)$ where $\xi$ is a primitive third root of unity.
+
+	If, on the other hand, $k=\bar \bbF_3$ and consider $\mu_3$, we get a single point with residye field $\bar\bbF_3$.
+\end{ex}
+
+\begin{ex}
+	If we are in the case of positive characteristic, then we get an algebraic group $\alpha_p$. Here the scheme is $\Spec k[x]/x^p$ and functorially it 
+	is the map $S\mapsto \{F\in\Gamma(S,\calO_S)|f^p=0\}$.
+\end{ex}
+
+\subsection{Matrix Groups}
+We already defined $\GL_n$, but we can also define 
+\[\operatorname{SL}_n:S\mapsto\{A=(a_{ij})|\det A=1\}\]
+with scheme $\Spec k[x_{ij}]/(\det-1)$.
+
+We also have the (upper) triangular matrices $T_n$ and unitary group $U_n$ and diagonal group $D_n$
+
+\begin{defn}
+	Let $G$ be a linear algebraic group. Then 
+	\begin{itemize}
+		\item $G$ is a \textbf{vector group} if $G\cong V_a$ for some finite dimensional $V$.
+		\item $G$ is a \textbf{split torus} if $G\cong \bbG_m^n$.
+		\item $G$ is a \textbf{torus} if there is a field extention $k\to k'$ such that 
+		\[G\times_{\Spec k}\Spec k'\cong \bbG^n_{m,k'}\]
+	\end{itemize}
+\end{defn}
 \end{document}

### 2019-10-01 Algebraic Geometry/Algebraic-Geometry.tex
update

@@
 \DeclareMathOperator{\Sym}{Sym}
+\DeclareMathOperator{\Spec}{Spec}
+\let\O\relax
+\newcommand*{\O}{\calO}
+\newcommand{\A}[1]{\bbA^{#1}}
+\let\P\relax
+\newcommand{\P}[1]{\bbP^{#1}}
 
@@
 
+\section{September 30th, 2019}
+One question that was asked: if we have fixed the underlying topological space in a space with functions, must there be a morphism 
+between them somehow? Might there instead be a common cover of the two?
+
+\begin{ex}
+	Let $k$ be a field with some topology on it such that every point is closed (you could do the discrete topology).
+	Let $\tilde\O(U)$ be the continuous funcitons $U\to k$. In other words, these functions are locally constant.
+
+	Locally constant functions behave nicely under restrictions to opens, of course. The other axioms are also great.
+
+	Have we really found an initial object in our category? This would be enough to establish a ``tent'' (as in 
+	localization of categories). Try this out and see what happens!
+\end{ex}
+
+\subsection{The question of affine space}
+Recall the question about whether $\A 1$ is an affine variety. The idea here is that $\phi:X\to k$ is a morphism of 
+spaces with functions if and only if it is regular (that is, in $\O_{\A 1}$).
+
+One direction is tautological (a morphism to $\A1$ has a polynomial underlying it), so let $\phi$ be regular. Then to see that $\phi$ is continuous 
+can be checked by pulling back all closed sets. The important observation is that $D(f-a)=X\setminus\phi^{-1}(a)$, which is closed (an axiom for spaces with functions).
+
+The last thing to check is where $\phi$ pulls back regular functions to regular functions. This relies on the facts that $\O_X$ is a $k$-algebra and that 
+$\phi(x)-b_j$ is regular on $U$ when $b_j\notin U$.
+
+\subsection{Algebra maps}
+Notice that since we have a condition that $\O_X(X)$ must be finitely generated as a $k$-algebra, this means that 
+\[\Hom(X,Y)=\Hom_k(\O_Y(Y),\O_X(X))=\Hom_k(k[x_1,\dots,x_n]/(f_1,\dots,f_m),O_X(X))\]
+and
+\[\Hom(X,Y)=\{(\gamma_1,\dots,\gamma_n)\in(\O_X(X))^n:f_j(\gamma_i)=0,\forall j=1,\dots,m\}\]
+In other words, we are looking at maps that factor through $Z$:
+\begin{center}
+	\begin{tikzcd}
+		(\gamma_1,\dots,\gamma_n):X\ar[r]\ar[dr] & k^n\\
+		& Z=Z(f_i)\ar[u,hook]
+	\end{tikzcd}
+\end{center}
+Now what we want to say is that $Y=Z$. That is, \textit{affine varieties are closed subsets of affine spaces.}
+
+Now this is all good, but the problem is that we had to \textit{choose} a presentation of $\O_Y(Y)$ to get this picture.
+of course we want something more canonical! We will see in this class (and in Kempf) that this can be done.
+
 \end{document}

### 2019-10-01 Algebraic Groups/Algebraic-Groups.tex
update

@@
 \end{defn}
+
+\section{September 30th, 2019}
+Another example to consider:
+\begin{ex}
+	Let $G=\operatorname{PGL}_n$, the projective linear group. Recall we want to define this as $\GL_n/k^\ast$ (from group theory). To do this for algebraic groups,
+	we define 
+	\[\operatorname{PGL}_n=\operatorname{Proj}k[x_{ij}]_{det}:= \Spec (k[x_{ij}]_{det})_0\]
+
+	The geometric picture is difficult since we haven't yet defined quotients, but as a functor we say $\operatorname{PGL}_n$ 
+	is $\Aut(\bbP^n)$, the functor that sends $S\mapsto \Aut(\bbP_S^n)$ where $\bbP^n_S=\bbP_k^n\times_{\Spec k} S$.
+\end{ex}
+
+\subsection{Non-affine group schemes}
+\begin{ex}
+	Let $\lambda\ne 0,1$ be an element in $k$. Then we can define the elliptic curve 
+	\[E_\lambda=V(y^2z-x(x-z)(x-\lambda z))\subset \bbP^2\]
+	Which gives us a double cover over $(0,1)$ and $(\lambda,\infty)$ with singleton fiber (ramified) over $0,1,$ and $\lambda$.
+
+	Then for any $\lambda\ne 0,1$, $E_\lambda$ is a \textbf{projective} group scheme.
+\end{ex}
+\begin{rmk}
+	If you look at the $\bbC$-points, you get $E_\lambda(\bbC)=\Lambda_\lambda$, giving you a torus. Recall (from e.g. complex analysis) that the 
+	moduli here is $\operatorname{SL}_2(\bbZ)$ of all elliptic curves.
+\end{rmk}
+
+\subsection{Abelian Varieties}
+\begin{defn}
+	An \textbf{abelian variety over $k$} is asmooth, geometrically connected ($A\times_{\Spec k}\Spec\bar k$ is connected), proper group scheme $A$ over $k$.
+\end{defn}
+\begin{ex}
+	Over $\bbC$, $\bbC^g/\Lambda$ where $\Lambda\cong \bbZ^{2g}\subseteq \bbC^g$ gives us a genus $g$ example.
+\end{ex}
+\begin{thm}
+	Any abelian variety over $k$ is commutative and projective.
+\end{thm}
+\begin{thm}[Chevalley]
+	If $G$ is any group scheme, then the sequence 
+	\[1\to H\to G\to A\to 1\]
+	is exact, where $H$ is a linear algebraic group (affine!) and $A$ is an abelian variety.
+\end{thm}
+\begin{ex}
+	Let $X\to \Spec k$ be a geometrically integral projective scheme (proper may suffice). The idea here is that over $\bbC$
+	the rings over every open set are integral domains.
+
+	Now consider the \textbf{Picard functor} $\operatorname{Pic}_X:\operatorname{Pic}:\mathbf{Sch/k}\to \Grp$ sending 
+	\[S\mapsto \operatorname{Pic}(X_S=X\times_k S)/p^k\operatorname{Pic(S)}\]
+\end{ex}
+\begin{thm}
+	$\operatorname{Pic}_X$ is represented by a scheme locally of finite type, thus $\operatorname{Pic}_X^0$, the connected 
+	component of the identity in $[\calO_X]\in\operatorname{Pic}_X$ is an abelian variety.
+\end{thm}
+\subsection{Relative Group Schemes}
+\begin{ex}
+	Consider $\bbG_{m,\bbZ}=\Spec \bbZ[t]_t$. Then $G_{m,S}=\bbG_{m,\bbZ}\times_{\Spec \bbZ}S$. 
+	In the case that $S=\Spec R$, $\bbG_{m,S}=\Spec R[t]_t$.
+\end{ex}
+
+\begin{ex}
+	Let $\bbA^1=\Spec k[x]$ and define $G=\Spec k[x,y]_{xy+1}\subseteq\bbA^2$. Notice this is the plane minus a hyperbola.
+
+	Define $\cdot:G\times_{\bbA^1}G\to G$ to be given by 
+	\[(x,y)\cdot(x,y')=(x,xyy'+y+y')\]
+
+	Then the thing here is the fiber (think vertical line in the plane!) over 0 is $\bbG_a$ and is isomorphic to $\bbG_m$ otherwise.
+\end{ex}
+
+\begin{ex}
+	Let $\calE_\lambda=V(y^2z-x(x-z)(x-\lambda z))$ over $\Spec k[\lambda]$. Then when $\lambda=0$, we get the nodal cubic given by 
+	$y^2z-x^2(x-z)$ (node at the origin).
+
+	Now if you look at the connected component around 0 of $\Aut(\calE_\lambda)/\bbA_\lambda$, you actually find (when $\lambda=0$) 
+	that $\bbG_m\cong\Aut(\calE_0)^0$.
+\end{ex}
+
+\subsection{Some definitions}
+\begin{defn}
+	A \textbf{homomorphism} $\phi:G\to G$ of group schemes over $S$ is a map $\phi:H\to G$ of schemes such that 
+	\begin{center}
+		\begin{tikzcd}
+			H\times_S H\ar[r,"m_H"]\ar[d,"\phi\times\phi"] & H\ar[d]\ar[d,"\phi"]\\
+			G\times_S G\ar[r,"m_G"] & G
+		\end{tikzcd}
+	\end{center}
+\end{defn}
+\begin{prob}
+	Show that this automoatically imples that the identity and inversion maps are respected as well (automatically).
+\end{prob}
+
+\begin{defn}
+	A \textbf{subgroup of $G\to S$} is a subscheme $H\subseteq G$ such that $H(T)\le G(T)$ for all $T$ over $S$.
+\end{defn}
+\begin{prob}
+	Show that $\ker(\phi)\subseteq H$ is a subgroup.
+\end{prob}
+\begin{rmk}
+	This gives you a nice way to construct new group schemes. For example, the following are exact:
+	\[1\to \operatorname{SL}_n\to GL_n\xrightarrow{\det} \bbG_m\to 1\]
+	and 
+	\[1\to \bbG_m\to \GL_n\to \operatorname{PGL}_n\to 1\]
+\end{rmk}
+
+\begin{prop}
+	Let $G\to S$ be a group scheme. Then $G\to S$ is separated if andy only if $e:S\to G$ is a closed immersion.
+\end{prop}
+\begin{prf}
+	The idea here is that $S\to G$ is a closed immersion. Then we consider the map $m\circ(\id,S):G\times_SG\to G$
+	and consider this along with the diagonal map $\Delta:G\to G\times_S G$ and this is a pullback square!
+\end{prf}
+\begin{cor}
+	Any group scheme over $k$ is separated.
+\end{cor}
+The idea is going to be that if $X$ is any scheme over $k$, then any point $X\in X(k)$ is closed.
+
 \end{document}

### 2019-10-02 Algebraic Geometry/Algebraic-Geometry.tex
Update

@@
 
-\usepackage{amsmath, graphicx, color, fancyhdr, tikz-cd, mdframed, enumitem, framed, adjustbox, bbm, upgreek, xcolor, hyperref, manfnt}
+\usepackage{amsmath, graphicx, color, fancyhdr, tikz-cd, mdframed, enumitem, framed, adjustbox, bbm, upgreek, xcolor, hyperref, manfnt, wasysym}
 \usepackage[framed,thmmarks]{ntheorem}
@@
 
+\section{October 2, 2019}
+\subsection{Questions without (complete) answers}
+\subsubsection{Morphisms and stuff}
+A question to get things started for the day. Let $X$ and $Y$ be spaces with functions and let $Y$ be an affine variety and let $f:Y\to X$ be a map of sets 
+(but with no further assumption on $f$). This naturally induces amap from $\calO_X(X)$ to the functions $\Hom_{\Set}(Y,k)$ (which 
+clearly contains the regular functions on $Y$).
+
+Further assume that there exists a $\gamma:\O_X(X)\to\O_Y(Y)$. We know that since $Y$ is affine, $\gamma$ corresponds to a morphism $\varphi:Y\to X$.
+Then the question is: when does $f=\varphi$? We've already answered this question for $\A1$, notice.
+\subsubsection{Algebraic closure}
+Where do we use algebraic closure of the base field? It has been swept under the rug a bit, but consider the function
+\[\frac{1}{x^2+1}:\bbR\to \bbR.\]
+
+This certainly seems like it should be a regular function (e.g. it is rational and defined everywhere on $\bbR$) but this 
+conflicts with the idea that we want to identify $\O_{\A1}(\A1)=k[t]$, but that is clearly not the case here. Think about this.
+\subsubsection{Yet another}
+Consider the set $R$ of all continuous maps $k\to k$ under the cofinite topology. Someone asked if $R$ is a $k$-algebra. The answer is 
+a bit convoluted, but the short answer is no. Specifically if we are using the product topology on $k\times k$, the addition map isn't continuous! 
+This also points to the question of what topology is the correct one to use on these things.
+
+\subsection{Back to affine varieties}
+Recall that we constructed a (highly-non-canonical) picutre of how any affine variety arises as a closed subset of some affine space $k^n$.
+
+We want to remove this dependence on presentation, however, and that is what we are working toward.
+
+\subsubsection{Affine Space}
+Now we focus in on $\A n=k^n$. We really want that the projection functions $x_i:k^n\to k$ should be regular. But since we want this (eventually)
+to form a $k$-algebra, we want that each $f\in[x_1,\dots,x_n]$ should be regular!
+
+The axioms of a space with functions tells us that the \textbf{vanishing locus}
+\[Z(f)=\{a|f(a)=0\}\subseteq k^n\]
+and furthermore $Z(S)$ should be closed for all $S\subseteq k[x_1,\dots,x_n]$. This leads us to a definition:
+\begin{defn}
+	A subset $Z\subseteq k^n$ is \textbf{Zariski-closed} if there exists an $S\subseteq k[x_1,\dots,x_n]$
+	such that $Z=Z(S)$.
+\end{defn}
+\begin{lem}
+	The Zariski closed sets are the closed sets of a topology (called the \textbf{Zariski Topology}).
+\end{lem}
+\begin{prf}
+	Just do it. Nike. \checked
+\end{prf}
+\begin{rmk}
+	Notice that here the set $\{(a,-a)\}\subseteq k^2$ (the pullback of zero under the addition map) is Zariski closed!
+	This fixes the problem we were running into in the third question above.
+\end{rmk}
+Now since $Z(S)=Z(I_S)$ where $I_S$ is the ideal generated by $S$, it is enough to consider vanishing loci of ideals. Furthermore we have the 
+map that extracts the ideal of functions that vanish on a set $Z\subseteq k^n$. There are a ton of great identiies you can prove here. Go to your 
+favorite algebra book (e.g. Dummit \& Foote) to see them.
+
+\subsubsection{Functions}
+What about functions on these spaces? If we take $f\in k[x_1,\dots,x_n]$ these seem like they should be regular functions $k^n\to k$.
+\begin{thm}[(Weak) Nullstellensatz]
+	Say $k=\bar k$. Then every maximal ideal $\frakm\lhd k[x_1,\dots,x_n]$ has the form $(x_1-a_1,\dots,x_n-a_n)$.
+\end{thm}
+\begin{rmk}
+	Equivalently, it is the kernel of a $k$-algebra morphis $k[x_1,\dots,x_n]\to k$.
+\end{rmk}
+\begin{cor}[Nullstellensatz]
+	Let $J$ be an ideal of $k[x_1,\dots,x_n]$. Then $I(Z(J))=\sqrt{J}$.
+\end{cor}
+\begin{prf}
+	\textit{Notice this only works when $k$ is uncountable!} Suppose that $\frakm$ is a maximal ideal with residue field 
+	$L=k[x_i]/\frakm$. This gives us a surjection of $k[x_1,\dots,x_n]\to L$. Thus $\dim_k L$ is countable!
+
+	But $\dim_k k(t)$ is uncountable! The proof here is that the $\frac{1}{t-\lambda}$ for $\lambda\in k$ is a linearly-independent collection.
+	So then $L/k$ is algebraic, and since $k=\bar k$ $L=k$.
+\end{prf}
+
 \end{document}

### 2019-10-02 Algebraic Groups/Algebraic-Groups.tex
Update

@@
 
+\section{October 2, 2019}
+Notice that a \textbf{relative group scheme} (referred to in last lecture) refers to a groups scheme over an arbitrary 
+base scheme $S$.
+
+\subsection{Properties of schemes}
+Today we are going to be talking about reducedness, connectedness, irreduciblility, regularity, and smoothness.
+
+Recall that a scheme $X$ is \textbf{reduced} if and only if $\forall x\in X$, $\calO_{X,x}$ is reduced. An example of a non-reduced 
+scheme is $\Spec k[x]/(x^2)$.
+\begin{defn}
+	We say a scheme $X$ over $k$ is \textbf{geometrically reduced} if for all field extensions $k'/k$,
+	\[X_{k'}=X\times_{\Spec k}\Spec k'\]
+	is reduced.
+\end{defn}
+\begin{rmk}
+	It is equivalent that $X_{\bar k}$ is reduced if and only if every $k'/k$ is purely inseperable (I think).
+\end{rmk}
+\begin{rmk}
+	If $k$ is perfect, then $X$ is reduced if and only if $X$ is geometrically reduced.
+\end{rmk}
+\begin{defn}
+	A local ring $(A,\frakm)$ is \textbf{regular} if $\dim _{\text{Krull}}A=\dim_{A/\frakm} \frakm/\frakm^2$
+\end{defn}
+\begin{defn}
+	A scheme $X$ is regular if for all $x\in X$, $\calO_{X,x}$ is regular.
+\end{defn}
+\begin{rmk}
+	If $X\to \Spec k$ and $x\in X(k)$, the tangent space at $x$ is 
+	\[T_{X,x}=(\frakm/\frakm^2)^\vee=\{f:\Spec k[\varepsilon]/\varepsilon^2\to X|0\mapsto x\}\]
+\end{rmk}
+\begin{rmk}
+	Notice that if $X\to \Spec k$ is regular and $k'/k$ is a field extension, then $X_{k'}$ is not necessarily regular.
+\end{rmk}
+
+\begin{defn} 
+	A Scheme $X\to \Spec k$ of finite type is \textbf{smooth} if $X_{\bar k}$ is regular.
+\end{defn}
+
+\subsection{Facts about algebraic groups}
+Then we can return to the proposition we want to prove:
+\begin{prop}
+	Let $G\to \Spec k$ be an algebraic group. Then $G$ is geometrically reduced if and only if $G$ is smooth over $\Spec k$.
+\end{prop}
+\begin{prf}
+	Smoothness over $k$ implies reducedness. Now since we are only interested in the algebraic closure of $k$, we can say $k=\bar k$. Because $G$ 
+	is reduced, there exists a nonempty open $U\subseteq G$ that is smooth. Then since $G(k)\subseteq |G|$ is dense in $G$ (as a topological space) 
+	and Then $G=\cup_{g\in G(k)}m_g(U)$ for our smooth $U$, and this gives us a smooth cover of $G$.
+\end{prf}
+
+We will see next itme that all linear algebraic groups over $k$ where $\ch k=0$ are all 
+geometrically reduced (and thus smooth).
+
+\subsection{Connectedness}
+Let $G$ be an algebraic group over $k$. Then we have our maps $e:\Spec k\to G$, so consider it as $e\in G(k)$.
+Let $G^0\subseteq G$ be the connected component of $e$. It is both open and closed.
+\begin{rmk}
+	If $X\to \Spec k$ is of finite type and $x\in X(k)$, then $X$ being connected implies that $X$ is geometrically connected.
+\end{rmk}
+This establishes that $G^0$ is actually geometrically connected! We actually will see
+\begin{prop}
+	$G^0\subseteq G$ is an (open and closed) algebraic subgroup.
+\end{prop}
+The idea here is that $G^0\times G^0$ is connected, so the image of the multipication map on this set lands in a connected component (since it is connected).
+Since $e\in G^0$, and $m(e,e)=e\in G^0$, this shows that the multiplication map restricts to a well-defined map $G^0\times G^0\to G^0$. A similar argument 
+goes through for the inverset map, etc.
+
+The upshot here is that if $G$ is an algebraic group, then there exists a factorization 
+\[1\to G^0\to G\to \pi_0(G)\to 1\]
+where $\pi_0(G)$ is given the structure of a discrete group.
+\begin{rmk}
+	Now we also have that $(G^0)_{k'}=(G_{k'})^0$ for all $k'/k$. The idea is to get a map of of one into the other and then use clopenness and connectedness 
+	to show they are equal.
+\end{rmk}
+
+\begin{prop}
+	A connected algebraic group over $k$ is irreducible.
+\end{prop}
+\begin{prf}
+	We can assume $k=\bar k$. Suppose $G= X\cup Y$, where both are closed, $X$ is irreducible, 
+	and $X\cap Y\ne\varnothing$. Thus there exists an element $g\in X\setminus Y$. That is, $g$ lies in a single irreducible component.
+
+	But then using the multiplication by $h$ map on $G$, we get to every other point in $G$, so every point 
+	is in a single irreducible component. But the intersection was nontrivial! Or something.
+\end{prf}
+
+\begin{prop}
+	If $G_{\text{red}}$ is geometrically connected, then $G_{\text{red}}\subseteq G$ is a subgroup. In particular, 
+	if $k$ is perfect, then $G_{\text{red}}$ is a subgroup of $G$.
+\end{prop}
+\begin{rmk}
+	$X$ is geometrically reduced implies that $X\times X$ is geometrically reduced.
+\end{rmk}
+
 \end{document}

### 2019-10-04 Algebraic Geometry/Algebraic-Geometry.tex
Update

@@
 \newcommand{\P}[1]{\bbP^{#1}}
+\newcommand{\eqdef}{\stackrel{\text{def}}{=}}
 
@@
 conflicts with the idea that we want to identify $\O_{\A1}(\A1)=k[t]$, but that is clearly not the case here. Think about this.
-\subsubsection{Yet another}
+\subsubsection{Yet another}\label{subsubsec:third}
 Consider the set $R$ of all continuous maps $k\to k$ under the cofinite topology. Someone asked if $R$ is a $k$-algebra. The answer is 
@@
 	Notice that here the set $\{(a,-a)\}\subseteq k^2$ (the pullback of zero under the addition map) is Zariski closed!
-	This fixes the problem we were running into in the third question above.
+	This fixes the problem we were running into in the third question (sec.~\ref{subsubsec:third}) above.
 \end{rmk}
@@
 \end{rmk}
-\begin{cor}[Nullstellensatz]
+\begin{cor}[Nullstellensatz]\label{nullstellensatz}
 	Let $J$ be an ideal of $k[x_1,\dots,x_n]$. Then $I(Z(J))=\sqrt{J}$.
@@
 
+\section{October 4, 2019}
+Today we are going to be talking a bit more about the existence of affine varieties. Max talked a bit about the philosophy of work in this course:
+he made this extended metaphor concerning butterflies but the take-away is to take learning onto ourselves. :)
+
+\subsection{Questions from last time}
+\subsubsection{Maps and elements}
+In the book we did this silly thing where we defined $\Spec A\eqdef\Hom_{\Alg}(A,k)$ and then idendified 
+$A$ with $k[\Spec A]$ by $a(f)=f(a)$. This seems a bit silly at first, but it may have something to do with the fact that 
+we are looking for a natural way to construct affine varieties without having to choose a presentation. We will hopefully see 
+something about this by the end of the day.
+
+\subsection{Back to the Nullstellensatz}
+Recall that we defined the operators $Z$ and $I$ that ``do the work'' of the Nullstellensatz. We then wrote (cor.~\ref{nullstellensatz}) $I(Z(J))=\sqrt{J}$. The idea is that 
+this will gives us the function structure on an affine variety.
+\begin{prf}[Cor.~\ref{nullstellensatz}]
+	One way is not too hard. For the more difficult direction:
+	Let $g\in I(Z(J))$. Then $Z(J)\subseteq Z(g)$. Now notice taht $D(g)$ can be naturally identified 
+	with $\Spec k[x_i][1/g]$. Then consider
+	\[J'=Jk[x_i][1/g]\]
+	and the key realization is that $J'$ cannot be contained in any maximal ideal. The idea is that you can work by contradiction:
+	this implies that $J$ is contained in an element of $D(g)$, but it isn't!
+
+	Thus $J'=(1)$. So we can write $1=\frac{f}{g^N}$. Thus $g^k(f-g^N)=0$ in $k[x_i]$ and since $g$ isn't nilpotent, $f=g^N$.
+\end{prf}
+\begin{cor}
+	There is a lattice anti-isomorphism between the radiacal ideals in $k[x_i]$ and Zariski-closed 
+	subsets $Z\subseteq k^n$ via the maps $J\to Z(J)$ and $Z\mapsto I(Z)$.
+\end{cor}
+\begin{cor}
+	For any ideal $J\subseteq k[x_i]$, 
+	\[\sqrt{J}=\bigcap_{\text{maximal } \frakm\supset J}\frakm\]
+\end{cor}
+\begin{rmk}
+	``The functions that vanish at the zero locus of $J$ are precisely those that vanish at all the points of $J$''.
+\end{rmk}
+\begin{cor}
+	$D(g)\subseteq k^n$. Then the map 
+	\[k[x_i][1/g]\to \Hom(D(g),k)\]
+	via the map 
+	\[\frac{f}{g^N}\mapsto \left(x\mapsto \frac{f(x)}{g(x)^n}\right)\]
+	is injective.
+\end{cor}
+
+\subsection{Affine space}
+Let's define $\A n\eqdef k^n$ with the Zariski topology. Let 
+\[\calO_{\A n}(U)=\{f\in k(x_1,\dots,x_n)|\operatorname{poles}(f)\subseteq\A n\setminus U\}\subseteq\Hom(U,k)\]
+Then, for example,
+\[\O_{\A n}(D(g))=k[x_1,\dots,x_n][1/g].\]
+\begin{prop}
+	$\A n$ is an affine variety.
+\end{prop}
+\begin{prf}
+	$\phi:X\to \A n$ gives us maps $\phi_1,\dots,\phi_n:X\to k$. Then that $\A n$ is affine relies on the fact:
+	$\phi$ is a morphis if and only if the $\phi_i$ are regular. One direction is not too bad since coordinate functions 
+	are regular by the axioms of morphisms. The other direction needs to be completed! DO IT!
+\end{prf}
+
 \end{document}

### 2019-10-04 Algebraic Groups/Algebraic-Groups.tex
Update

@@
 \DeclareMathOperator{\Spec}{Spec}
+\newcommand{\red}[1]{{#1}_{\text{red.}}}
 
@@
 
+\section{October 4, 2019}
+Some review. Let $G$ be an algebraic group and denote $e:\Spec k\to G$ be the identity. We saw a lot of propositions last time.
+
+Now let $k$ be a nonperfect field and take $t\in k\setminus k^p$. Then define 
+\[G\eqdef V(x^{p^2}-tx^p)\subseteq \bbG_a\]
+which Milne claims is not reduced. We can see why it's not geometrically reduced, but we're missing the details here.
+
+\subsection{Another special case}
+\begin{thm}
+	When $k=\bar k$, $G$ is smooth if and only if 
+	\[\dim T_e \red G=\dim T_e G.\]
+\end{thm}
+\begin{rmk}
+	When $G$ is smooth, it is reduced, so the equality is clear. For the other direction, we get that $k$ is pefect, so $\red G$
+	which is geoemetrically reduced if and only if $G$ is smooth. But 
+	\[\dim G\le \dim T_e G=\dim T_e\red G=\dim \red G=\dim G\]
+\end{rmk}
+\begin{thm}
+	If $G$ is a linear algebraic group over $k$ and $\ch k=0$, $G$ is smooth,
+\end{thm}
+\begin{prf}
+	We can assume $k=\bar k$. Then set $G=\Spec A$ where $A$ is a Hopf algebra. THen we get Hopf algebra maps $m^\ast$ and $e^\ast$.
+	Notice that the augmentation ideal $\frakm=\ker(e^\ast)$ is a maximal ideal.
+
+	Then we want to prove 
+	\begin{enumerate}
+		\item $A\cong\frakm\oplus k$ as a $k$-vector space (obvious).
+		\item $\forall a\in\frakm$, $m^\ast(a)-a\otimes 1-1\otimes a\in \frakm\otimes \frakm.$
+	\end{enumerate}
+	To see the second, notice that $m^\ast(a)-a\otimes 1-1\otimes a$ is in the kernel of 
+	\[e^\ast\otimes \id:A\otimes A\to k\otimes A.\]
+	This is clear from the commutative diagram
+	\begin{center}
+		\begin{tikzcd}
+			k\otimes A & A\otimes A\ar[l,"e^\ast\otimes\id"] \\
+			& A\ar[ul,"\sim"]\ar[u,"m^\ast"]
+		\end{tikzcd}
+	\end{center}
+	Then we conclude $f\in \ker(e^\ast\otimes\id)\cap\ker(\id\otimes e^\ast)=A\otimes\frakm\cap\frakm\otimes A$ by a symmetric argument.
+	Finally we notice that $A\otimes\frakm\cap\frakm\otimes A=\frakm\otimes\frakm$, and so $f$ lies in this ideal.
+
+	Now we want to show that $\dim T_e G=\dim_k\frakm/\frakm^2=\dim_k\frakm/(\sqrt{0}+\frakm^2)=\dim T_e\red G$. It suffices to show that 
+	for all $a\in\sqrt{0}$, $a\in \frakm^2$. Suppose the opposite--so let $a\in\sqrt{0}\setminus \frakm^2$. Consider the diagram 
+	\begin{center}
+		\begin{tikzcd}
+			A\ar[r]\ar[d] &A_\frakm\ar[d]\\
+			A/\frakm^2\ar[r,"\sim"] & A_\frakm/(\frakm A_\frakm)^2
+		\end{tikzcd}
+	\end{center}
+	Now the image of $a$ in $A_\frakm$ is nonzero, so there exists $n>0$ such that $a^n\in A_\frakm$ but $a^{n-1}\ne 0$ in $A_\frakm$.
+	Thus there exists $f\notin \frakm$ such that $a^nf=0\in A$. Subsititute $af$ for $a$, and thus there is an $a\in\sqrt{0}$ such that $a^n=0$ in $A$ but $a^{n-1}\ne 0$ in $A_\frakm$.
+
+	Then by fact 2,
+	\[m^\ast(a)=1\otimes a+a\otimes 1+r,\quad r\in\frakm\otimes\frakm\]
+	and since $m^\ast$ is a ring homomorphism, 
+	\[0=m^\ast(a^n)=(m^\ast(a))^n=(a\otimes 1+(1\otimes a+r))^n=a^n\otimes 1+n(a^{n-1}\otimes a+(a^{n-1}\otimes 1)r)+X\]
+	where $X\in A\otimes\frakm^2$. But since $a^n=0$, we get 
+	\[n(a^{n-1}\otimes a+(a^{n-1}\otimes 1)r)\in A\otimes \frakm^2\]
+
+	Now since $(a^{n-1}\otimes 1)r\in (a^{n-1})\frakm\otimes A$, so 
+	\[n(a^{n-1}\otimes a)\in (a^{n-1})\frakm\otimes A+A\otimes\frakm^2\]
+	and since $\ch k=0$, we get that $n$ is a unit, so 
+	\[a^{n-1}\otimes a\in (a^{n-1})\frakm\otimes A+A\otimes\frakm^2\]
+
+	Now since this lives in $A\otimes A$, consider the image of the quotient map $A\otimes A\to A\otimes A/\frakm^2$. Then 
+	\[a^{n-1}\otimes\bar a\in (a^{n-1})\frakm\otimes A/\frakm^2\subseteq A\otimes A/\frakm^2\]
+	And note that $a^{n-1}\notin a^{n-1}\frakm$ because otherwise $a^{n-1}=a^{n-1}q$ for $q\in \frakm$.
+	Then $a^{n-1}(1-q)=0\in A_\frakm$, which implies that $a^{n-1}=0\in A_\frakm$ (since $1-q$ is a unit here).
+
+	Then somehow we get that $\bar a=0\in A/\frakm^2$, so $a\in \frakm^2$.
+\end{prf}
+
 \end{document}

### 2019-10-07 Algebraic Geometry/Algebraic-Geometry.tex
update

@@
 % Document-Specific Macros
-\DeclareMathOperator{\1}{\mathbbm{1}}
-\DeclareMathOperator{\GL}{GL}
-\DeclareMathOperator{\Sym}{Sym}
-\DeclareMathOperator{\Spec}{Spec}
-\let\O\relax
-\newcommand*{\O}{\calO}
-\newcommand{\A}[1]{\bbA^{#1}}
-\let\P\relax
-\newcommand{\P}[1]{\bbP^{#1}}
-\newcommand{\eqdef}{\stackrel{\text{def}}{=}}
 
@@
 	via the map 
-	\[\frac{f}{g^N}\mapsto \left(x\mapsto \frac{f(x)}{g(x)^n}\right)\]
+	\[\frac{f}{g^N}\mapsto \left(x\mapsto \frac{f(x)}{g(x)^N}\right)\]
 	is injective.
@@
 
+\section{October 7th, 2019}
+\subsection{Questions/Discussions}
+\subsubsection{Initial and final objects}
+We asked before whether the space with functions $(X,\O^{\text{loc. constant}})$ is an initial or terminal 
+object. Adam asserts that it is a final object in the category of spaces with functions where the underlying space is $X$ (I believe this 
+is sheaves over $X$).
+
+Notice we can't use all continus maps where $k$ has the discrete topology, since $0\in k$ is not closed.
+
+\subsubsection{Subalgebras and rings of functions}
+Assume that $f:B\to A$. This gives us a nice map $\tilde f:\Spec A\to \Spec B$. One question we may have is 
+``if $f$ is injective, does this imply that $\tilde f$ is surjective?''
+
+Consider an example: Say $B=k[t]\hookrightarrow k[s,t]=A$. Then any function on $A$ looks like $(s-a,t-b)$
+and the map induced on functions is just projection, so this gives us the map $(t-b)$, which is all the maps on $B$.
+
+Another example: consider the map $\bbC[t]\to\bbC[s]$ sending $t\to s^2$. Then this induces a map $z\to z^2$ from $\bbC\to \bbC$ (why?)
+which is again surjective.
+
+Next consider $k[x,u]/(xy-1)=k[x][1/x]=A$, which is a hyperbola over $\bbR$. Then the localization map $B=k[x]\hookrightarrow A$
+induces a map that is basically the identity everywhere \textit{except zero}. So it is \textbf{not surjective.}
+
+\brk
+
+Some properties to notice: examples one and two are \textit{flat} extensions. The second is a \textbf{finite} extension. The 
+third is neither. We will investigate what is going on further later on.
+
+One idea: consider whether the map $X=\Spec B\to \A 2\setminus(\A1\setminus\{0\})$ exists. One of the things that he keeps questioning is 
+whether the target spaces is open or closed as a subset of affine space. (Note that a map is proper if it sends closed to closed).
+
+\subsection{Back to affine varieties}
+Continuing our proof/discussion from last time, we were considering $\varphi:X\to\A n$ which we said was a morphism iff each coordinate $\phi_i:X\to k$ is regualr.
+(This basically follow since $\varphi$ is continuous and sends regular functions to regular functions.
+
+For the regularity, consider $U\subseteq\A n$. Then $U$ admits a cover of $D(g)$, so it suffices to check where $\phi$
+pulls back regular maps on $D(g)$. Of course this is $k[x_1,\dots,x_n][1/g]$! So consider the image of 
+\[\frac{1}{g^N}\sum a_ix^i\]
+and its image in $\phi^{-1}(D(g))=D(\phi^\ast(g))$ is 
+\[\frac{\sum a_i(\phi^\ast x)^i}{(\phi^\ast g)^N}\]
+
+\begin{rmk}
+	Big idea: We started with the ``dream'': that there is a correspondence between the algebra and the geometry. This is 
+	our main guiding principle, so we know we've found the ``right'' topology when we have found one that supports this dream.
+	This is an answer to the question ``why is the Zariski topology not just a degenerate case?''
+\end{rmk}
+
+\subsubsection{Affine varieties in general}
+Let $J\subseteq k[x_1,\dots,x_n]$ be a radical ideal. We know this corresponds (uniquely!) 
+to a subset $Z\subseteq \A n$. Here we can consider $(Z,\O_Z)$ where $Z$ is a subspace of $\A n$.
+
+Then take any closed $W\subseteq Z$, which is the intersection (by definition of the topology)
+\[\cap_{i\in I}Z(f_i)\quad f_i\in h[x_i]/J\]
+and then $\O_Z(D(g))=\frac{k[x_i]}{J}[1/g]$ (note we used the nullstellensatz here!).
+
+So then we claim that $Z$ is an affine variety. To see this, consider a map $\varphi:X\to Z$ and the composition
+\[X\xrightarrow{\varphi} Z\hookrightarrow \A n\]
+so topologically $X$ cfactors through $Z$ if and only if $J\subseteq \O^{\A n}(\A n)$ maps to zero in $\O_X(X)$.
+
+The takeaway here is that a morphism $X\to\A n$ factors through $Z$ topologically if and only if it factors in the categorification of 
+spaces with functions.
+
 \end{document}

### 2019-10-07 Algebraic Groups/Algebraic-Groups.tex
update

@@
 % Document-Specific Macros
-\DeclareMathOperator{\1}{\mathbbm{1}}
-\DeclareMathOperator{\GL}{GL}
-\DeclareMathOperator{\Sym}{Sym}
-\DeclareMathOperator{\Spec}{Spec}
-\newcommand{\red}[1]{{#1}_{\text{red.}}}
 
@@
 
+\section{October 7, 2019}
+Today we will be primarly concerned with 
+\subsection{Group actions}
+Let $G$ be an algebraic group over $k$. 
+\begin{defn}
+	A \textbf{group action} of $G$ on a scheme $X$ over $k$ is the data of a morphism
+	\[\mu:G\times X\to X\]
+	such that the usual axioms hold. That is,
+	\begin{center}
+		\begin{tikzcd}
+			G\times G\times X\ar[r,"m\times\id"]\ar[d,"\id\times \mu"] & G\times X\ar[d,"\mu"]\\
+			G\times X\ar[r,"\mu"] & X
+		\end{tikzcd}
+		\begin{tikzcd}
+			\Spec k\times x\ar[r,"e\times\id"]\ar[rd,"\sim"] & G\times X\ar[d,"\mu"]\\
+			& X
+		\end{tikzcd}
+	\end{center}
+\end{defn}
+\begin{rmk}
+	Apparently it was an exercise already to show that this is equivalent to an action of $h_G$ on $h_X$.
+\end{rmk}
+\begin{rmk}
+	The map $(g,x)\mapsto(g,gx)$ is an automorphism of $G\times X$, so if $p_2:G\times X\to X$ is projection,
+	\begin{center}
+		\begin{tikzcd}
+			G\times X\ar[rr,"\sim"] \ar[dr,"\mu"] & & G\times X\ar[dl,"p_2"]\\
+			& X &
+		\end{tikzcd}
+	\end{center}
+	commutes.
+\end{rmk}
+\begin{defn}
+	Let $X$ and $Y$ be schemes over $k$ with a $G$ action. Then a \textbf{$G$-equivariant morphism} $f:X\to Y$ 
+	is one such that for all $g\in G$, the following commutes:
+	\begin{center}
+		\begin{tikzcd}
+			G\times X\ar[r,"\id\times f"]\ar[d,"\mu_X"] &G\times Y \ar[d,"\mu_Y"]\\
+			X\ar[r,"f"] & Y
+		\end{tikzcd}
+	\end{center}
+\end{defn}
+\subsubsection{Some examples}
+\begin{itemize}
+	\item $G$ actions on itself by multiplication or conjugation.
+	\item $\Gm$ acts on $\A1$. Geometrically, we are just looking at $k^\ast$ acting on $k$ by scaling.
+	Algebraically, we want a map $\mu\Gm\times\A1\to\A1$ given by the map of algebras:
+	\[k[x]\xrightarrow{\mu^\ast}k[t]_t\otimes k[x]\quad\text{via}\quad x\mapsto tx\]
+	Functorially, if $S$ is a scheme over $k$, then $\Gm(S)=\Gamma(S,\O_S)^\ast$ which acts on $\A1(S)=\Gamma(S,\O_S)$,
+	again by scaling.
+	\item You can consider $\GL_n$ action on $\A n$ by multiplicaiton or on $\A{n\times n}$ via multiplication or conjugation.
+\end{itemize}
+\subsubsection{Orbits and Stabilizers}
+Let $G$ be an algebraic group over $k$ action on a scheme $X$ over $k$. Let $x\in X(k)$. Then we have a map
+\[\mu_x:G\times \Spec k\xrightarrow{\id\times x}G\times X\xrightarrow{\mu}X\]
+where 
+\[g\mapsto (g,x)\mapsto gx.\]
+\begin{defn}
+	The \textbf{orbit of $x$} is $Gx=\mu_x(G)\subseteq |X|$ set-theoretically. The \textbf{stabilizer of $x$ in $G$} 
+	is $G_x=\mu_x^{-1}(x)\subseteq G$. 
+\end{defn}
+\begin{rmk}
+	$G_x$ is always a closed algebraic subgroup of $G$.
+\end{rmk}
+\begin{prop}
+	$\mu_x(G)$ is open in its closure in $|X|$.
+\end{prop}
+Recall first the following:
+\begin{thm}[Chevalley's Theorem (different?)]
+	If $f:X\to Y$ is a map of schemes of finite type over $k$, then $f(X)\subseteq Y$ is contstructible (i.e. 
+	is a disjoint union of finitely many locally closed subsets).
+
+	\noindent\textit{Recall that locally closed means closed in an open subspace.}
+\end{thm}
+\begin{cor}
+	\textit{Maybe a definition:} The orbit $Gx\subseteq \operatorname{im}(\mu_x)\subseteq X$. If 
+	$G$ is reduced, then $Gx$ is reduced.
+\end{cor}
+
+\subsubsection{Applications}
+Say that $\ch k=p$. Then $\mu_p$ acts on $\Gm$ by multiplication.
+
+\brk
+
+$\Gm$ acts on $\A1$ with two orbits: $\Gm\cdot 1=\{x\ne 0\}$ and $\Gm\cdot 0=\{0\}$. The stabilizers are 
+$G_1=1$ and $G_0\cong\Gm$.
+
+\brk
+
+Consider $G$ acting on $\A2$ via $t(x,y)=(tx,t^{-1}y)$. Then the orbits are hyperbolas! There is also a notion of closed 
+orbits that I didn't quite catch. Also apparently the orbit-stabilizer statement is easy to see in geometry via 
+a fiber bundle $G$ over $Gx$ where the fiber over $x$ is $G_x$.
+
+\begin{prop}
+	If $\phi:G\to H$ is a homomorphism of algebraic groups, then $\phi(G)\subseteq |H|$ is closed.
+\end{prop}
+\begin{rmk}
+	The proof included reducing first to $k=\bar k$. The trick here is to consider the group action induced by $\phi$
+	and then consider the map $\mu_{e_H}$ of this action. Then $\mu_{e_H}(G)=\phi(G)$ and one can prove that this is closed.
+\end{rmk}
+
+In particular, we have that \textbf{subgroups of an algebraic group are always closed.} Note that this stands in stark contrast 
+to Lie theory where you get
+
 \end{document}

### 2019-10-10 Algebraic Groups/Algebraic-Groups.tex
Update

@@
 In particular, we have that \textbf{subgroups of an algebraic group are always closed.} Note that this stands in stark contrast 
-to Lie theory where you get
+to Lie theory where you get non-closed subgroups.
+
+\section{October 9, 2019}
+Recall that last time we were considering actions of algebraic groups on schemes of finite type over $k$. We discussed the orbit and stabilizer 
+of an element $x\in X$ and showed that $G\cdot x$ is open in its closure. We also saw that $G_x$ is a closed subgroup.
+
+We also say that $\phi(H)$ (as a set) is always closed! None of these facts are true for Lie groups or relative group schemes 
+(the base scheme is not $\Spec k$ for $k$ a field).
+
+\subsection{Cartier Duality}
+Let $G\to\Spec k$ be a \textbf{finite} group scheme (so $G=\Spec A$ and $A$ is a finite dimensional Hopf algebra).
+Some examples of finite group schemes are:
+\begin{itemize}
+	\item $G$ a finite group. Then $G=\sqcup_{g\in G}\Spec k=\Spec\left(\prod_{g\in G}k\right)$
+	\item $\mu_n=\Spec k[t]/(t^n-1)$
+	\item $\ch k=p$ and $\alpha_p=\Spec k[t]/t^p$
+\end{itemize}
+\begin{rmk}
+	Recall all the maps and diagrams that $A$ has as a Hopf algebra.
+\end{rmk}
+A question one may ask: what if we apply the idea of dualizing $(-)^\vee=\Hom_\Alg(-,k)$ to $A$?
+Do we get another Hopf algebra?
+
+The short and sweet of it is yes! But notice that we are coming from the commutative world, so we expect 
+$A$ to be commutative. But in general, $A$ is not cocommutative (in fact, it is if and only if $G$ itself was commutative as a group).
+
+Thus $A^\vee$ is indeed a (cocommutative) Hopf algebra, and when $G$ is commutative, $A^\vee$ is as well. So
+\begin{prop}
+	If $G=\Spec A$ is a commutative group scheme, then the Cartier dual $G^D=\Spec A^\vee$ is a commutative group scheme as well.
+\end{prop}
+\begin{rmk}
+	The above observations gives us an anti-autoequivalence of the category of commutative affine group schemes. Furthermore $(G^D)^D=G$.
+\end{rmk}
+\begin{ex}
+	Consider $\mu_n=\Spec A-\Spec \oplus_0^{n-1}k\cdot t^i$. So then if we let $\{e_i\}$ be the basis for $A^\vee$ dual 
+	to $\{t_i\}$. we can compute comultiplication 
+	\[e_i\mapsto \sum_{j=0}^{n-1}e_j\otimes e_{i-j}\]
+	and multipication
+	\[e_i\otimes e_j\mapsto \delta_{ij}e_i\]
+
+	Then it can be shown that $G^D\cong\bbZ/n\bbZ$.
+\end{ex}
+
+Now given $G$, an algebraic group over $k$, define 
+\[\underline\Hom(G,\Gm):\mathbf{Sch}/k\to\Set\]
+which takes 
+\[T\mapsto \Hom_{\mathbf{AlgGrp}}(G_T,{\Gm}_T).\]
+\begin{thm}\label{thm:cartier}
+	If $G$ is a commutative finite group scheme over $k$, then 
+	\[G^D\cong\uHom(G,\Gm)\]
+\end{thm}
+Let $H=\Spec B\to \Spec R$ be a group scheme. Then 
+\[H_{\mathbf{GrpSch}/R}(H,{\Gm}_R)\subseteq\operatorname{Mor}_{\mathbf{Sch}/R}(H,{\Gm}_R)\]
+But the left hand side is equivalent to the grouplike elements of $B$ and the right hand side is equivalent
+to $\Hom_{\Alg_R}(R[t]_t,B)$.
+
+This leads to a proof of thm.~\ref{thm:cartier}:
+\begin{prf}
+	Let $G=\Spec A$ and $G^D=\Spec A^\vee$. First look at the $k$-points:
+	\begin{align*}
+		G^D(k)&= \Mor_{\Sch/k}(\Spec k, G^D)\\
+		&=\Hom_{\Alg_k}(A^\vee,k)=\{f\in A|m^\ast(f)=f\otimes f\}\hookrightarrow \Hom_{k}A^\vee,k)\\
+		&=\Hom(G,\Gm)\\
+		&=\uHom(G,\Gm)(k)
+	\end{align*}
+
+	If we then look at $R$ points for a general $R$, most things just change over, but we see 
+	\[\{f\in A\otimes R|m_R^\ast(f)=R\otimes R\}=\Hom_{\Alg_k}(A^\vee,R)=\Hom(G_R,\Gm)\]
+	and the rest follows.
+\end{prf}
+A question one may ask: what is $\Hom_\AlgGrp(\Gm,\Gm)$? It ends up it is $\bbZ$. You can send $t\mapsto t^n$ for all $n\in\bbZ$.
+But then $\uHom(\Gm,\Gm)$ is $\bbZ$ as a group scheme over $k$, which is not quasicompact. There was more but I am le tired.
 

### 2019-10-11 Algebraic Geometry/Algebraic-Geometry.tex
Update

@@
 
+\section{October 11, 2019}
+Today we are going to talk some more about varieties. Coming up on the horizon is a discussion of 
+the functor-of-points perspective and we'll talk about Yoneda.
+
+\subsection{Questions/Discussion}
+A group of students met up on Wednesday (which we skipped for Yom Kippur) and were talking about 
+how to prove the statement: ``points in an affine variety are closed.''
+
+The Nullstellensatz gives us that a point $(x_1,\dots,x_n)\in X$ correspinds to the vanishing locus of $(x_i-a_i)$. 
+What if we use the definition in the book, though? If $X=\Spec A=\Hom_{\Alg_k}(A,k)$ (where $A$ is a reduced finitely generated 
+$k$-algebra). But then the points are the maximal ideals $\frakm\in k[X]=A$, which are exactly the points we want!
+
+\subsection{Back to Varieties}
+Recall the definition of a variety: we are considering $(X,\O_X)$ is a space of functions. Then the idea we want is that we want to say 
+that this thing is locally affine. But if you consider infinitely many copies of $\A1$ intersecting pairwise, there is 
+an obvious cover by (affine) $\A1$'s. This shouldn't be affine. It doesn't embed in an affine space, for example.
+
+So the definition we used (and the one in Kempf) is that we must have a \textit{finite} cover by affine spaces. We wanted to discuss some examples.
+\begin{itemize}
+	\item Clearly all affine varieties are varieties.
+	\item $\P1$ is our first nontriial example. As a space it is the one-point compactification of $\A1$. The functions 
+	$\O_X(U)$ are the rational functions in $t$ with poles not in $U$. You can also construct it by taking the morphism 
+	\[\Gm\xrightarrow{t\mapsto 1/t}\Gm\]
+	and gluing along this morphism to get a copy of $\P1$! We've already shown in Kempf that it is not affine. 
+	What happens if we were to pick the identity above?! We get the line with two origins. It's a variety! But notice that 
+	(under the Euclidean topology) the space isn't Hausdorff! 
+\end{itemize}
+\begin{prob}
+	If $X$ and $Y$ are affine varieties, is $X\sqcup Y$ affine?
+\end{prob}
+\begin{prob}
+	How can we expresse the non-Hausdorffness of the line/plane with two origins in the Zariski topology?!
+\end{prob}
+
+\subsection{Varieties Glue}
+We've been throwing things around, but this is important to write down: Start with $U_i, i\in I$, where $I$ is finite 
+(although we could drop finiteness if we don't care about the thing being a variety). For all $i,j$, we have open subsets 
+$V_{ij}\subseteq U$ such that $V_{ii}=U_i$ and isomorphisms $\varphi_{ij}:V_{ij}\xrightarrow{\sim}V_{ji}$ of varieties such that 
+\begin{enumerate}
+	\item $\phi_{ii}=\id$
+	\item $\varphi_{ij}(V_{ik}\cap V_{ij})=V_{ji}\cap V_{jk}$
+	\item $\forall i,j,k,\, \varphi_{jk}\circ\varphi_{ij}=\varphi_{ik}$ on $V_{ij}\cap V_{ik}$
+\end{enumerate}
+then there exists a unique package $(X,\iota_i:U_i\hookrightarrow X)$ where $XS$ is a variety and each $\iota_i$ is an open embedding.
+
+One idea is you can rephrase this in categorical language as the colimit of a diagram in a category. Hmmm
+
 \end{document}

### 2019-10-11 Algebraic Groups/Algebraic-Groups.tex
Update

@@
 
+\section{October 11, 2019}
+Today we're doing problems and stuff. Forgot about that.
+
+\subsection{Casey's Presentation}
+Let $G$ and $H$ be objects in a category $\calC$ with finite products. Let $h:H\to G$ be a group homomorphism. That is,
+
+Then we get similar diagrams for the identity and inverse maps (they are respected by $h$). 
+Then there is a bunch of diagram work. It's too hard to do a diagram without knowing the shape ahead of time.
+
+Oh hey I used Adam's site! 
+\begin{center}
+	\begin{tikzcd}
+		\ast \arrow[rd, "{(e,e,e)}"]               &                                                                                                                                                  &                                                                                        &                                                                        &                                                                           \\
+												   & H\times H\times H \arrow[r, "(i_g\circ h)\times h\times h"] \arrow[d, "1\times m_H"']                                                            & G\times G\times G \arrow[d, "1\times m_g"]                                             &                                                                        &                                                                           \\
+												   & H\times H \arrow[r, "(i_G\circ h)\times h"]                                                                                                      & G\times G \arrow[rd, "m_G"]                                                            &                                                                        &                                                                           \\
+												   &                                                                                                                                                  &                                                                                        & G                                                                      &                                                                           \\
+	\end{tikzcd}
+\end{center}
+\begin{center}
+	\begin{tikzcd}												   & \ast \arrow[r, "e"] \arrow[rd, "{(e,e,e)}"] \arrow[rdddd, "he"', bend right=49]                                                                   & H \arrow[rd, "{(h,h)}"]                                                                &                                                                        &                                                                           \\
+												   & &H\times H\times H \arrow[d, "(i_g\circ h)\times h\times h"']                                                                                                                                                                         & G\times G \arrow[rd, "\ast\times 1"] \arrow[ld, "{(\ast,1)\times 1}"',swap] &                                                                           \\
+												   &                                                                                                                                                  & G\times G\times G \arrow[d, "1\times m"'] \arrow[rd, "m\times 1"]                      &                                                                        & \ast\times G \arrow[ld, "e\times 1"'] \arrow[lldd, "\pi_1", bend left=49] \\
+												   &                                                                                                                                                  & G\times G \arrow[d, "m"']                                                              & G\times G \arrow[ld, "m"]                                              &                                                                           \\
+												   &                                                                                                                                                  & G                                                                                      &                                                                        &                                                                           \\
+	\end{tikzcd}
+\end{center}
+\begin{center}
+	\begin{tikzcd}
+												   & \ast \arrow[rd, "he"', dashed,swap] \arrow[r, "\ast", bend left] \arrow[d, "{(e,e,1)}"'] \arrow[ld, "{(e,e,e)}"'] \arrow[dd, "{(e,e)}", bend left=60] & \ast \arrow[r, "e"]                                                                    & G                                                                      &                                                                           \\
+		H\times H\times H \arrow[rd, "1\times m"'] & H\times H\times\ast \arrow[d, "1\times \pi_2"'] \arrow[l, "1\times 1\times e"']                                                                  & G \arrow[u, "\ast"'] \arrow[d, "{(1,1)}"'] \arrow[rd, "{(i,1)}", bend left] &                                                                        &                                                                           \\
+												   & H\times H \arrow[r, "h\times h"] \arrow[rr, "(ih)\times h", bend right]                                                                          & G\times G \arrow[r, "i\times 1"]                                                       & G\times G \arrow[uu, "m"]                                              &
+		\end{tikzcd}
+\end{center}
+\begin{rmk}
+	The idea above is we want to show the first diagram commutes. That is captured in the paths of the second diagram which commutes by 
+	the axioms of a group object. The third diagram shows a similar commutativity for the unit $e$.
+\end{rmk}
+
 \end{document}

### 2019-10-14 Algebraic Geometry/Algebraic-Geometry.tex
update

@@
 
+\section{October 14th, 2019}
+We're going off-book a bit to talk about 
+\subsection{Yoneda Lemma}
+Recall that we said that $Y$ was affine if 
+\[\Hom(X,Y)\cong \Hom_k(\O_Y(Y),\O_X(X)).\]
+Now wer are going to take some time to put this into a broader context. Let $\calC$ be a category. We get naturally two functors 
+\[h_a:\calC^{op}\to\Set\qquad h^a:\calC\to\Set\]
+where 
+\[h_a(b)=\Hom_\calC(b,a)\qquad h^a(b)=\Hom(a,b).\]
+
+Now given a map $f:a\to a'$, we get natural transformations $f\circ -:h_a\to h_{a'}$ and $-\circ f:h^{a'}\to h^a$. Now notice that 
+if $\ast$ is a one-point variety, then $h_X(\ast)=|X|$, the underlying point set of $X$. Notice that in a similar way 
+$h_X(\A1)$ is something like the ``line space'' of $X$.
+
+Notice that we then get a functor
+\[h_{(-)}:\calC\to\mathbf{Func}(\calC^{op},\Set)\]
+and we should \textbf{think} that a functor $\calC^{op}\to\Set$ is a \textbf{space over $\calC$}. That is, we can think of 
+$\calC$ as the category of open sets of a topological space. We can say 
+\[\Hom(U,V)=\left\{\begin{array}{lr}
+	\varnothing, & U\not\subset V\\
+	\{\varnothing\}, & \text{otherwise}.
+\end{array}\right.\]
+
+Notice the category $\mathbf{Open}(X)$ gives us a functor $\phi:\mathbf{Open}(X)^{op}\to\Set$ where we map 
+an open $U\subseteq X$ to $\phi(U)$. For instance if $F$ is a vector bundle over $X$, we can define
+\[\phi_F(U)\]
+to be the sections over $U$ of the covering map. Cool.
+
+So we're working with the slice category $\mathbf{Top}/X=\{X\to Y,\text{cts}\}$. Here morphisms are maps $Y\to X$ satisfying 
+\begin{center}
+	\begin{tikzcd}
+		Y\ar[rr]\ar[rd] & &Z\ar[ld]\\
+		& X &
+	\end{tikzcd}
+\end{center}
+
+So now fix $F\to X$. Then the map $\phi_F(Y\to X)$ which is the collection of of diagrams of the form above ($Z=F$).
+
+Then if we look at points $x\hookrightarrow X$, we get that $\phi_F(x\to X)$ is the fiber of $F$ over $x$. One ould hole that somehow 
+you could recover all of $F$ from these maps, and that is precisely the content of 
+\begin{lem}[Yoneda]
+	The functor $h:\calC\to\mathbf{Func}(\calC^{op},\Set)$ is fully faithful.
+\end{lem}
+\begin{rmk}
+	That is, 
+	\[\Hom_\calC(a,a')\to\Hom_{\mathbf{Func}(\calC^{op},\Set)}(h_a,h_{a'})\]
+	is a bijection.
+\end{rmk}
+\begin{prf}
+	This proof is so tautological it is sometimes confusing to prove. We can look it up in any of our old notes or books but the idea is 
+	to look at the image of the identity map.
+\end{prf}
+Next time we will see a bunch of examples and exercises.
+
 \end{document}

### 2019-10-14 Algebraic Groups/Algebraic-Groups.tex
update

@@
 
+\section{October 14th, 2019}
+Let $G$ be a finite group. Recall the definiton of a \textbf{representation} (a linear action of $G$ on a vector space $V/k$).
+This is the same data as a group homomorphism to $\GL(V)$.
+
+\subsection{Representations of Algebraic Groups}
+
+Now what if $G$ is an algebraic group over $k$? Now we have some exra structure of $G$ as a variety.
+\begin{defn}
+	A (finite dimensional) \textbf{representation} of an algebraic group $G/k$ is a homomorphism $\rho:G\to \GL(V)$ of algebraic groups.
+\end{defn}
+\begin{rmk}
+	Notice that when $V$ is infinite-dimensional, $\GL(V)$ is no longer of finite type, so we have to let $\rho$ be a 
+	morphism of group schemes.
+\end{rmk}
+We have the standard representation of $\GL_n$ acting on $k^n$ in the natural way. We also have the regular representation 
+$G$ action on $\Gamma(G,\O_G)$. When $G=\Gm$, we get over $\bbC$ an action of $\Gm$ on $\GL(V)$ in the usual way (scaling by $\bbC^\ast$).
+
+Observe that 
+\[\rho:G\to \GL(V)=\Spec(\operatorname{Sym}^\ast(V\otimes V^\vee))_{\det}\]
+corresponds to a ring morphism 
+\[\Sym^\ast(V\otimes V^\vee)_{\det}\to \Gamma(G,\O_G)\eqdef \Gamma(G)\]
+which corresponds to a map 
+\[V\otimes V^\vee\to \Gamma(G)\]
+and then tensoring with $V$, this gives us a map
+\[V\xrightarrow{\sigma}\Gamma(G)\otimes V\]
+
+So any group action gives us a \textbf{coaction} of $\Gamma(G)$ on $V$.
+\begin{defn}
+	A representation of $G$ is a $k$-vector space $V$ along with a coaction 
+	\[\sigma:V\to \Gamma(G)\otimes V\]
+	satisfying the usual dual diagrams to actions.
+\end{defn}
+
+\begin{rmk}
+	As a matter of notation, recall that if $G=\Spec A$, then $A$ is a Hopf algebra. So we call $V$ an $A$-comodule.
+\end{rmk}
+
+\subsection{Reps of diagonalizable group schemes}
+Let $k$ be a field (or even a ring!) and let $A$ be a finitely-generated abelian group. Define $D(A)$ to be 
+\[D(A)=\oplus_{a\in A}k\cdot t^a\eqdef \Spec R\]
+Then we get a multiplication 
+\[R\otimes R\to R\qquad t^a\otimes t^b\mapsto t^{a+b}\]
+and comultiplication
+\[R\to R\otimes R\qquad t^a\to t^a\otimes t^a\]
+and counit $\varepsilon$ sending $t^a\to 1$ (all $t^a$ are primitive).
+\begin{prop}
+	$R$ is a Hopf algebra. In particular, $D(A)\to \Spec k$ is a linear algebraic group.
+\end{prop}
+
+As an example, consider $A=\bbZ$. Then $R\cong k[t]_t$. Thus $D(A)=\Gm$.
+
+If instead $A=\bbZ/n$, then $R\cong l[t]/(t^n-1)$, so $D(A)\cong\mu_n$.
+
+Finally when $A\cong \bbZ^r\oplus\bbZ/n_1\oplus\cdots\oplus \bbZ/n_k$, then 
+\[D(A)=\Gm^r\times\mu_{n_1}\times\cdots\times \mu_{n_k}\]
+
+\begin{defn}
+	An algebraic group over $k$ is \textbf{diagonalizable} if $G\cong D(A)$ for some $A$.
+\end{defn}
+Recall the definiton of irreduciblility.
+\begin{thm}
+	Let $A$ be af initely generated abelian group and $G=D(A)$. Then 
+	\begin{itemize}
+		\item Every irreducible representation of $G$ is one-dimensional and isomorphic to $I_a$, 
+		corresponding to $k\to \Gamma(G)\otimes k$ where $1\mapsto t^a\otimes 1$ for some $a\in A$.
+		\item Every representation decomposes as a direct sum of irreducibles.
+	\end{itemize}
+\end{thm}
+\begin{prf}
+	Let $\sigma:V\to \Gamma(G)\otimes V$ be a representation of a diagonalizable group. For $a\in A$, define 
+	\[V_a\eqdef\{v\in V|\sigma(v)=t^a\otimes v\}\subseteq V\]
+	Now the claim is that $V_a\cap V_b=0$ if $a\ne b$ and furthermore $\sum V_a=V$. The first isn't too hard to see.
+
+	The second follows by considering $v\in V$ and looking at the image of it under $\sigma$. That is, 
+	\[\sigma(v)=\sum_1^Nt^{\alpha_i}\otimes v_i\]
+	where $\alpha_i\in A$ and $v_i\in V$. Then a very simple argument shows $v=\sum v_i$ (using linearity).
+	Then it remains to show that $v_i\in V_{a_i}$, but this will make things work. (use the other axiom of a coaction).
+\end{prf}
+\begin{rmk}
+	When $A=\bbZ$, $G=D(\bbZ)=\Gm$, which tells us that representations of $\Gm$ are in bijection with $\bbZ$-gradings 
+	of $V\cong\oplus_{n\in\bbZ}V_n$!
+\end{rmk}
+\begin{defn}
+	A linear algebraic group $G\to\Spec k$ is called \textbf{linear reductive} if every representation decomposes as a direc sum of irreducibles.
+\end{defn}
+\begin{prob}
+	Show the above is equivalent to the statements
+	\begin{itemize}
+		\item for each $G$-representations $W\subseteq V$, there exists $W'\subseteq V$ subrepresentations such that $V\cong W\oplus W'$.
+		\item $0\to W\to V\to W'\to 0$ is exact.
+	\end{itemize} 
+\end{prob}
+\begin{rmk}
+	Notice that this says that $D(A)$ is linear reducible. In particular, $\Gm$ and $|mu_n$ are in \textbf{any characteristic}. This runs counter to Maschke in 
+	finite groups.
+\end{rmk}
+
+Consider $\bbZ/p$ in $\ch p$. We get an action $\bbZ/p$ on $k^2$ via 
+\[1\cdot(x,y)=(x+y,y).\]
+But notice that $k\stackrel{\hookrightarrow}{y=0}k^2$ is a subrepresentation, but has no complement! Thus this group is not linearly reductive!
+
+\brk
+
+As another example, consider
+\[\Ga\cong\left\{\begin{pmatrix}
+	1 & \alpha \\ 0 & 1
+\end{pmatrix}\right\}
+\subset\GL_2(k)\]
+where $\Ga$ acts on $k^2$ by $\alpha(x,y)=(a+\alpha y,y)$. Then it can be easily seen not to be alinear representation.
+
 \end{document}

### 2019-10-16 Algebraic Groups/Algebraic-Groups.tex
update

@@
 \begin{defn}
-	A linear algebraic group $G\to\Spec k$ is called \textbf{linear reductive} if every representation decomposes as a direc sum of irreducibles.
+	A linear algebraic group $G\to\Spec k$ is called \textbf{linear reductive} if every representation decomposes as a direct sum of irreducibles.
 \end{defn}
@@
 \subset\GL_2(k)\]
-where $\Ga$ acts on $k^2$ by $\alpha(x,y)=(a+\alpha y,y)$. Then it can be easily seen not to be alinear representation.
+where $\Ga$ acts on $k^2$ by $\alpha(x,y)=(a+\alpha y,y)$. Then it can be easily seen not to be a linear representation.
+
+\section{October 16th, 2019}
+Last time we talked about representations! Woot.
+
+Notice that if $G$ is linear (i.e. affine), then the multiplication map induces comultiplication 
+\[\Gamma(G)\to \Gamma(G)\otimes\Gamma(G)\]
+so $\Gamma(G)$ is the \textbf{regular} representation with coaction given by multiplication.
+
+We also saw some equivalent contitions similar to Machke for linear reductive groups. Finally we say some examples and diagonalizable groups.
+
+\subsection{New Stuff}
+Given a $G$-representation $V$, let $V^G$ be 
+\[\{v\in V|\sigma(v)=1\otimes v\}=\operatorname{Eq}\{V\stackrel{\xrightarrow{1\otimes -}}{\xrightarrow{\sigma}}\Gamma(G)\otimes V\}=\Hom^G(k,V)\subseteq V.\]
+\begin{rmk}
+	I need to figure out the \TeX for equalizers/parallel maps.
+\end{rmk}
+\begin{ex}
+	Given the representation $\Gm$ action on $V=\oplus V_d$, $V^G=V_0$.
+\end{ex}
+\begin{prob}
+	If $G(k)$ is dense in $G$, then $V^G=V^{G(k)}$.
+\end{prob}
+\begin{prop}
+	A linear algebraic group $G$ over $k$ is linearly reductive if and only if the functor from $G$-representations to 
+	$k$-vector spaces given by $V\mapsto V^G$ is exact.
+\end{prop}
+\begin{prf}
+	If $V\cong W\oplus W'\twoheadrightarrow W$ is a $G$ representation, then 
+	\[W^G\oplus(W')^G=V^G\twoheadrightarrow W^G\]
+	is also surjective.
+
+	Suppose that we have a short exact sequence 
+	\[0\to W'\to V\to W\to 0\]
+	and that this functor is exact. Then we want to show we get a section $\sigma:W\to V$. To do this, consider the functor 
+	$\Hom^G(W,-)=\Hom^G(k,W^\vee\otimes -)=(W^\vee\otimes -)^G$, so by the assumption this is exact and we can lift the identity on $W$ 
+	to a map in $\Hom^G(W,V)$, giving us our section.
+\end{prf}
+\begin{prop}
+	Let $G$ be a linear algebraic group over $k$ and $V$ a $G$-representation. Let $W\subseteq V$ be a finite dimensional $k$-subspace (not 
+	necessarily $G$-invariant). Then there exists $W\subseteq W'\subseteq V$ such that $W'$ is a finite dimensional representation of $G$.
+\end{prop}
+\begin{prf}
+	We can assume that $W=\langle w\rangle$ for $w\in V$. Apply $\sigma:V\to \Gamma(G)\otimes V$. Then if $\{t_i\}$ is a 
+	basis for $\Gamma(G)$, we get 
+	\[w\mapsto\sum t_i\otimes w_i.\]
+	Then we claim that $w\in\langle w_i\rangle$ and $\langle w_i\rangle\subseteq V$ is a subrepresentation.
+
+	For the first, consider the diagrams:
+	\begin{center}
+		\begin{tikzcd}
+			k\otimes V & \Gamma(G)\otimes V\ar[l,swap,"e^\ast\otimes\id"]\\
+			& V\ar[ul,"\sim"]\ar[u,"\sigma",swap]
+		\end{tikzcd}
+		\begin{tikzcd}
+			\sum e^\ast(t_i) w_i & \sum t_i\otimes w_i\ar[l]\\
+			& w\ar[u]
+		\end{tikzcd}
+	\end{center}
+	so $w$ is in the span of the $w_i$.
+
+	For the second claim, we need to show that 
+	\[\sigma(w_i)\in\Gamma(G)\otimes\langle w_i\rangle.\]
+	To see this consider the diagram 
+	\begin{center}
+		\begin{tikzcd}
+			\Gamma(G)\otimes\Gamma(G)\otimes V & \Gamma(G)\otimes V\ar[l,"m^\ast\otimes\id"]\\
+			\Gamma(G)\otimes V\ar[u,"\id\otimes\sigma"] & V\ar[l,"\sigma"]\ar[u,"\sigma"]
+		\end{tikzcd}
+	\end{center}
+	And tracing through $w\in V$, we get that 
+	\[\sum t_i\otimes\sigma(w_i)=\sum_{i,j,k}\alpha_{i,j,k}t_i\otimes t_j'\otimes w_k\]
+	and so by looking at coefficients of $t_i\otimes \Gamma(G)\otimes V=\sigma(w_i)$ (look closer here), we see it is 
+	\[\sum_{j,k}\alpha_{i,j,k}t_j'\otimes w-k\in\Gamma(G)\otimes\langle w_i\rangle.\]
+\end{prf}
+\begin{cor}
+	If $V$ is a $G$ representation, then 
+	\[W=\bigcup_{W\subset V}W\]
+	where the union is over all finite dimensional subgroups.
+\end{cor}
+\begin{cor}
+		If $G$ is a linear algebraic group (affine finite type over $k$), then for some $n$ $G\subseteq \GL_n$ is a closed subgroup.
+		In other workds, there exists a faithful representation $V$ of $G$.
+\end{cor}
+
+Now consider the regular representation $\Gamma(G)\to \Gamma(G)\otimes \Gamma(G)$. Notice that $\Gamma(G)$ is 
+a $k$-algebra of finite type. Choose generators $g_1,\dots,g_n$ for $\Gamma(G)$. Take a subrepresentation of $\Gamma(G)$
+spanned by $\langle h_1,\dots,h_N\rangle$ containing the span of the $g_i$.
+
+We have a map $G\to \GL(W)$ where $W=\langle h_i\rangle$ and we aske whether the induced map 
+\[\Sym^\ast(W\otimes W^\vee)_{\det}\]
+is surjective.
+
+Let's say that $h_i\mapsto \sum_j\gamma_{i,j}\otimes h_j$ under $\sigma$. Then using this map and the natural pairing between $W$ and $W^\vee$, we get a map 
+\[\Sym^\ast(W\otimes W^\vee)\supset W\otimes W^\vee\to\Gamma(G)\otimes W\otimes W^\vee\to \Gamma(G)\]
+where we send 
+\[h_i\otimes h_j^\ast\mapsto \gamma_{i,j}\]
+So using the counit identiy we can write 
+\[h_i=\sum_j e^\ast(\gamma_{i,j})h_j\]
+but we really want to write $h_i$ as a linear combination of the $\gamma_{i,j}$ (since we have shown they all lie in the image of this map).
+We don't know how to finish up.
 

### 2019-10-16 Algebraic Geometry/Algebraic-Geometry.tex
update

@@
 % Document-Specific Macros
+\newcommand{\SpcFun}{\mathbf{SpcFun}}
 
@@
 
+\section{October 16}
+We're going to finish up with with Yoneda! The milestone for the next week or so: we should have seen and digested some of the 
+algebra in the early sections of chapter 2. Max will assume we'll have seen it already starting Friday(ish). We should shoot for having all the problems 
+in chapter 2 done by a week from Friday (or so).
+
+\subsection{Old Questions} 
+We spoke a bit about lines glued together and the automorphisms of one of the lines that extend to the entire space. 
+This necessitates talking about what the automorphisms of $\A1$ are.
+
+We also talked about whether the coordinate axes in $\A3$ are isomorphic as varieties to the projection 
+of these axes onto a plane. We discussed that this must induce an isomorphism between the coordinate algebras since they are affine.
+
+\subsection{Back to Yoneda}
+Some useful notation for encoding Yoneda: given a functor $F:\calC^{op}\to\Set$, we say that an object $a\in\calC$
+\textit{represents} $F$ if $h_a\simeq F$. Then another formulation of Yoneda says that a representing object of a representable functor is unique up to isomorphism of
+representing objects.
+\begin{center}
+	\begin{tikzcd}
+		h_a\ar[r,"\sim"]\ar[rr,bend right,"\sim"] & F & h_{a'}\ar[l,"\sim",swap]\\
+		& & \\
+		& a\to a'&
+	\end{tikzcd}
+\end{center}
+\begin{rmk}
+	$h_a$ is called the \textit{functor of points of $a$.}
+\end{rmk}
+\begin{ex}
+Given $f_1\dots,f_n\in k[x_i]$, consider the $k$-algebra $k[x_i]/(f_i)$. Then Yoneda gives us that $A$ is uniquely
+determined by the functor 
+\[h^A:\Alg_k\to\Set\]
+via
+\[B\mapsto \Hom(A,B)=\{(b_1,\dots,b_n)\in B^n|f_i(b_1,\dots,b_n)=0\}.\]
+\end{ex}
+\begin{ex}
+	The functor $h_{\P n}:\Alg_k\to\Set$ is 
+	\[A\mapsto \{A\twoheadrightarrow L| L\text{ invertible $A$-module}\}/\cong\]
+	This is appartently very important.
+\end{ex}
+
+\subsection{Representable or not?}
+\begin{itemize}
+	\item $\Ga$ as a functor from $\SpcFun^{op}\to\Set$ is represented by $\A1$.
+	\item $\Gm$ as a functor from $\SpcFun^{op}\to\Set$ is represented by $\A1\setminus\{0\}$.
+	\item $\GL_n:\SpcFun^{op}\to\Set$
+	\item $|\cdot|:\SpcFun\to\Set$ underlying set 
+	\item $\varnothing:\SpcFun^{op}\to\Set$ sending $X\mapsto\varnothing$
+	\item $\{\varnothing\}:\SpcFun^{op}\to\Set$ sending $X\mapsto\{\varnothing\}$.
+\end{itemize}
+
 \end{document}

### 2019-10-21 Algebraic Geometry/Algebraic-Geometry.tex
update

@@
 	\item $\Gm$ as a functor from $\SpcFun^{op}\to\Set$ is represented by $\A1\setminus\{0\}$.
-	\item $\GL_n:\SpcFun^{op}\to\Set$
-	\item $|\cdot|:\SpcFun\to\Set$ underlying set 
-	\item $\varnothing:\SpcFun^{op}\to\Set$ sending $X\mapsto\varnothing$
-	\item $\{\varnothing\}:\SpcFun^{op}\to\Set$ sending $X\mapsto\{\varnothing\}$.
+	\item $\GL_n:\SpcFun^{op}\to\Set$ which is represented by $\Spec k[x_{ij}]_{\det}$
+	\item $|\cdot|:\SpcFun\to\Set$ underlying set represented by $\ast$
+	\item $\varnothing:\SpcFun^{op}\to\Set$ sending $X\mapsto\varnothing$. Not representable!
+	\item $\{\varnothing\}:\SpcFun^{op}\to\Set$ sending $X\mapsto\{\varnothing\}$. Representable by $\ast$.
+	\item $\mu_n:\SpcFun^{op}\to\Set$ sending $X\mapsto\{f\in\O_X(X)|f^n=1\}$ represented by $\Spec k[t](t^n-1)$
 \end{itemize}
 
+\section{October 18th, 2019}
+This is something that Max just mentioned: what is $\O_X(\varnothing)?$ There is a single map here!
+
+\subsection{Old Questions}
+Is $U=\A2\setminus (0,0)$ affine? Well first notice 
+\[\O_{U}(U)\subseteq \bigcap_{f\in k[x,y]\text{ irred}}k[x,y][1/f]=k[x,y]\]
+\textbf{Why is the last equality true?! We're working over a UFD.} But then of course 
+$U$ can't be affine since $U\ne \A2=\Spec k[U].$ Another way to think of this is that every isomorphism of 
+coordinate rings yields an isomorphism of spaces!
+
+\subsection{Back (again) to affine varieties}
+\begin{thm}
+	\begin{itemize}
+		\item Let $\calA\subseteq\Alg_k$ be the subcategory of finitely-generated, reduced $k$-algebras. Then $\Spec$ defines an equivalence of categories
+		\[\Spec:\calA^{op}\to \mathbf{Aff}\]
+		\item The functor
+		\[X\mapsto\Spec \O_X(X):\SpcFun^\ast\to\mathbf{Aff}\]
+		is left adjoint to the canonical inclusion $\mathbf{Aff}\subseteq\SpcFun^\ast$, where $\SpcFun^\ast$
+		is the subcategory of $\SpcFun$ such that the global sections are finitely generated.
+
+		Thus $X\to\Spec \O_X(X)$ is universal for maps to affine varieties.
+	\end{itemize}	
+\end{thm}
+\begin{prob}
+	The last thing is saying that every morphism $X\to \Spec A$ factors (uniquely) through $X\to\Spec \O_X(X)$. Show this!
+\end{prob}
+
+\section{October 21st, 2019}
+Today we are moving on from Yoneda to speak a bit about the topological properties of algebraic varieties.
+
+Later we will do a lot more examples with the functor of points perspective, but for now the idea to keep around 
+is that this enables us to study something \textit{in relation to something else}.
+
+\subsection{Questions}
+We thought about $\mu_n:\SpcFun^{op}\to\Set$. Notice that (at least when $\ch k\nmid n$), we get that 
+$k[t]/(t^n-1)$, the representing algebra, is $\prod_{\zeta\in\mu_n(k)}k$. Probably the best way to understand 
+this is that it is $n$ points, but moreso that it is embedded in a natural way in $\Gm$ via the SES:
+\[\mu_n\hookrightarrow\Ga\to\Gm\]
+
+\subsection{Topological properties of varieties}
+\begin{defn}
+	A topological space $X$ is \textbf{quasi-compact} if every open cover has a finite subcover.
+\end{defn}
+\begin{rmk}
+	In France, compactness requires a space be Hausdorff. Thus in the development of AG we used the French 
+	definition and it stuck!
+\end{rmk}
+\begin{lem}
+	An affine variety is quasi-compact.
+\end{lem}
+\begin{prf}
+	The idea was that if $X=\cup U_i$, then since $D(f)$ generate the topology for $X$, we can refine
+	the cover so that $U_i=D(f_i)$. By the Nullstellensatz, $X=\cup D(f_i)$, we know that $(f_i)\lhd \O_X(X)$
+	is the whole ring!
+
+	But then $1=\sum_{j=1}^n a_jf_{i_j}$, so we can restrict to a finite cover!
+\end{prf}
+But note that we didn't need that our ring was Noetherian! So the exact same proof shows that $\Spec \bbC[x_1,x_2,\dots]$ is 
+quasi-compact! But $\Spec \bbC[x_i]/(x_1,\dots,x_n,\dots)$ is not quasi-compact!!!
+
+\begin{rmk}
+	A very cool (``AMAZING'' according to Max) fact is the following: if $U\subset X$ is any open in an affine space, then $U$ is quasi-compact.
+	Compare a similar idea in Hausdorff spaces, where this is actually false.
+\end{rmk}
+
+Let $X=\Spec A$ where $A=k[x_1,\dots,x_n]/(f_1,\dots,f_m)$, which is Noetherian by Hilbert basis and correspondence theorem.
+Now we have an anti-isomorphism of lattices between the radical ideals in $A$ and the closed sets in $X$.
+Thus 
+\begin{defn}
+	A topological space $X$ is Noetherian if any \textbf{descending} chain of closed subspaces stabilizes.
+\end{defn}
+\begin{rmk}
+	Assuming the axiom of choice, this is equivalent to the statement ``A non-empty set of closed subsets in $X$ has 
+	a minimal element.''
+\end{rmk}
+\begin{rmk}
+	Note that $\Spec A$ is Noetherian for any (finitely-generated) $k$-algebra.
+\end{rmk}
+So let $X$ be a Noetherian topological space.
+\begin{lem}
+	$Z\subseteq X$ is closed implies that $Z$ is Noetherian and quasi-compact.
+\end{lem}
+\begin{rmk}
+	The Noetherian bit is pretty clear! The quasi-compact part needs a bit of arguing. Use the definition that $X$ is quasi-compact 
+	if and only if $\cap Z_i=\varnothing\twoheadrightarrow \cap_{j=1}^n Z_{i_j}=\varnothing$ for some $i_j$.
+
+	Then using choice and the contrapositive it falls out!
+\end{rmk}
+\begin{defn}
+	A topological space $X$ is \textbf{irreducible} if $X\ne\varnothing$ and if $X=X_1\cup X_2$ where 
+	the $X_i$ are both closed imples that either $X=X_1$ or $X=X_2$.
+\end{defn}
+\begin{rmk}
+	Equivalently, if $U_1$ and $U_2$ are open and nonempty then $U_1\cap U_2\ne\varnothing$.
+	We also have that any non-empty open set is dense.
+\end{rmk}
+
 \end{document}

### 2019-10-21 Algebraic Groups/Algebraic-Groups.tex
Thanks Sean!
update

@@
 % Document-Specific Macros
+\DeclareMathOperator{\SL}{SL}
 
@@
 \[1\cdot(x,y)=(x+y,y).\]
-But notice that $k\stackrel{\hookrightarrow}{y=0}k^2$ is a subrepresentation, but has no complement! Thus this group is not linearly reductive!
+But notice that $k\stackrel{y=0}{\hookrightarrow}k^2$ is a subrepresentation, but has no complement! Thus this group is not linearly reductive!
 
@@
 
+\section{October 18th, 2019}
+Last time we say that any linear algebraic group $G$ embeds into $\GL_n$. The argument was basically that you look at the global functions 
+$\Gamma(G)$ and doing cool stuff. Right at the end Taffy and Tuomas figured out that we just needed to use the other ``side'' of the counit diagram.
+
+\subsection{An example}
+Consider 
+\[\operatorname{PGL}_2=(\operatorname{Proj} k[a,b,c,d])_{ad-bc]}=\Spec(k[a,b,c,d]_{\det})_0\]
+Then consider the representation spanned by 
+\[\left\langle\frac{a^2}{\det},\frac{ab}{\det},\dots,\frac{d^2}{\det}\right\rangle\]
+which has dimension 10 in $\Gamma(\operatorname{PGL}_2)$.
+
+Thus we have a representation $\operatorname{PGL}_2\to\GL_{10}$. We can compute the matrix representing a matrix (whose determinant
+can be assumed to be 1 since we are modding out by scalars).
+\begin{prob}
+	Do this! In Sage or something.
+\end{prob}
+\subsection{Special Linear Groups}
+Lets discuss $\SL_2$. 
+\begin{thm}
+	If $\ch k=0$, then 
+	\begin{itemize}
+		\item $\SL_2$ is linearly reductive.
+		\item Every irreducible representation of $\SL_2$ is isomorphic to $\Sym^d k^2$ for some $d$.
+	\end{itemize}
+	where $k^2$ is the standard representation of $\SL_2$.
+\end{thm}
+\begin{prf}
+	\textit{Sketch:} Recall that in the proof of Maschke one takes a surjection $V\twoheadrightarrow W$ 
+	of $G$ representations and we want to show it has a section. We pick a section $s$ in terms of vector spaces and then ``average''
+	it:
+	\[\tilde s:W\to V\qquad w\mapsto\frac{1}{|G|}\sum_{g\in G}g\cdot s(g^{-1}w)\]
+	which is our section.
+
+	Then using the Harr measure on the group, we get from the inclusion of (compact) $\operatorname{SU}_2$ in $\SL_2$ with quotient $\bbC$
+	and we can construct the section via 
+	\[w\mapsto \int_Gg\cdot s(g^{-1}w)\,\mathrm{d}g\]
+	and we get $T_e\SL_2=(T_e\operatorname{SU}_2\otimes_\bbR\bbC)$ and then there is a bit more Lie theory needed to show this makes full sense over $\bbC$.
+\end{prf}
+
+Now consider $\Sym^d(k^2)^\vee$, the degree $d$ polynomials on two variables. Then we get an action of $\SL_2$ via 
+\[g\cdot f(x,y)=f(g^{-1}\binom{x}{y})-f(dx-by,-cx+ay)\]
+Then, as many arguments in linear algebraic groups, we can reduce to a so-called \textit{maximal torus} 
+of matrices $(\begin{smallmatrix}a & 0\\0 & a^{-1}\end{smallmatrix})\cong\Gm$. Then we can use techniques on the lie algebra 
+$T_e\SL_2=\sl_2$.
+
+Whoa coool. The short exact sequence 
+\[1\to \SL_2\to \GL_n\xrightarrow{\det}\Gm\to 1\]
+gives us that any representation of $\SL_2$ lifts via $-\otimes \det^i$ to a representation of $\GL_2$.
+Sean asked whether this actually gets all the representations or just the polynomial ones. I feel like I should know the 
+answer to this. Jarod seems to think this is all of them. I think \href{https://math.stackexchange.com/questions/694001/non-polynomial-representations-of-gl-n}{this stack post}
+says something about that.
+
+\textit{I got caught up in thinking and googling and missed an example.}
+
+Let $\ch k=p$ and let $\alpha\in k^\times$. Then define $\alpha\cdot f=\alpha^pf$. This gives us a map 
+\[\Sym^d k^n\to \Sym^{dp}k^n\]
+that is additive taking $p^{th}$ powers.
+
+So then $\Sym^N k^n$ is \textbf{not simple} if $p|N$.
+
+\section{October 21st, 2019}
+Recall that we had that any linear algebraic group over a field $k$ injects into $\GL_n$ as a closed subgroup for some $n$.
+
+An open question is as follows: if $G\to\Spec k[\varepsilon]/\varepsilon^2$ is flat affine group scheme of finite type
+Then is $G\subseteq\GL_{n,k[\varepsilon]}$ for some $n?$ This question was asked (as far as Jarod knows) by Brian Conrad on Stack Overflow 
+and is still open.
+
+Our goal is to answer the following: if $\phi:H\to H$ then $K=\ker\varphi=H\times_G\Spec k\subseteq H$ is a closed subgroup. What about its image $H/K$?
+
+\subsection{Torsors}
+Today we are going to be talking about $G$-torsors.
+\begin{defn}
+	If $G$ is a group, a \textbf{torsor under $G$} is a set $P$ with a free and transitive group action.
+\end{defn}
+\begin{rmk}
+	So then by fixing a point $p\in P$, we get $G\xrightarrow{\sim}P$ by sending $g\mapsto pg$. In this way, it is like thinking about 
+	a group without the identity.
+\end{rmk}
+An example is by taking a Galois extension $K(\alpha)/K$ with minimal polynomial $f$ of $\alpha$. Then $G=\operatorname{Gal}(K(\alpha)/K)$ acts on 
+$\{x|f(x)=0\}$, which is a $G$-torsor.
+\begin{defn}
+	A \textbf{$G$-torsor over a set $S$} is a set $P$ with a free right $G$-action such that $P\to S$ is $G$-invariant and $S\cong P/G$.
+\end{defn}
+\begin{rmk}
+	Notice that a torsor under $G$ is a specialization of this definition by requiring that $S=\{\ast\}$, the singleton set.
+\end{rmk}
+\begin{ex}
+	Let $H\subseteq G$. THen $H$ acting on $H\to H\setminus G$ (left cosets $gH$ of $H$) is an $H$-torsor.
+\end{ex}
+\subsection{Flatness}
+\begin{defn}
+	A map of rings $A\to B$ is \textbf{flat} if $-\otimes_A B$ is exact.
+\end{defn}
+\begin{rmk}
+	Equivalently: for all $p\in \Spec A$, $A_p\to B_p$ is flat. Also: for all $q\in\Spec B$. $A_{\phi^{-1}(q)}\to B_q$ is flat.
+\end{rmk}
+\begin{defn}
+	$A\to B$ is \textbf{faithfully flat} if and only if $-\otimes_AB$ is \textbf{faithfully exact} (exactness and its converse).
+\end{defn}
+\begin{rmk}
+	Other equivalence to faithful flatness: $A\to B$ is flat and $\Spec B\twoheadrightarrow\Spec A$; or $A\to B$ 
+	is flat and for any $A$-module $M$, $M=0\leftrightarrow m\otimes_AB=0$.
+\end{rmk}
+\begin{rmk}
+	If $\Spec B\twoheadrightarrow\Spec A$ is faithfully flat and finite presented, then $\Spec A$ has the quotient topology.
+\end{rmk}
+
+\begin{prop}
+	Let $S=\Spec A$ be Noetherian. Let $G\to S$ be an affine group scheme of flat and finite type over $S$. Let $P\to S$ 
+	be a scheme over $S$ with a right $G$-action $P\times_SG\xrightarrow{\sigma}P$. Then the following are equivalent:
+	\begin{enumerate}
+		\item $P\to S$ is affine, (faithfully?)\footnote{We tried to prove this in class and it seemed not to be true if we don't say this}
+		flat, finite type and $(\sigma,\pi_P):(p,g)\mapsto (pg,p)$ is an isomorphism.
+		\item There exists a faithfully flat $S'$ such that 
+		\begin{center}
+			\begin{tikzcd}
+				P_{S'}\ar[r]\ar[d] & P\ar[d]\\
+				S'\ar[r,"\sigma"] &S
+			\end{tikzcd}
+		\end{center}
+		And $P_{S'}\cong G_{S'}$ as $G_{S'}$-modules.
+	\end{enumerate}
+\end{prop}
+\begin{rmk}
+	Note that the above says exactly that $P$ is a $G$-torsor. Another name that has been mentioned and Jarod seems to 
+	like is \textbf{principal $G$ bundle.}
+\end{rmk}
+
+\subsection{Descent}
+Along the way of proving the above propositon, we use the idea of descent.
+\begin{lem}
+	Consider $X'=\Spec B'\stackrel{\text{f.flat, f.type}}{\twoheadrightarrow}X=\Spec B\to Y=\Spec A$
+	where $X',X,$ and $Y$ are Noetherian. Then 
+	\begin{enumerate}
+		\item $X'\to Y$ is flat implies that $X\to Y$ is flat.
+		\item $X'\to Y$ is faithfully flat implies that $X\to Y$ is faithfully flat.
+		\item $X'\to Y$ is finite type implies that $X\to Y$ is.
+	\end{enumerate}
+\end{lem}
+\begin{rmk}
+	The idea for the first two is just looking at the functors using that $B\to B'$ is faithfully flat. For the third, if $B=\cup_\lambda B_\lambda$,
+	then $A\to B_\lambda$ is finitely generated. Then tensoring over $B$ with $B'$ gets us $B'=\cup_\lambda B_\lambda\otimes_B B'$.
+
+	But since $A$ is finitely generated over $B_\lambda$< eventually $B_\lambda\otimes_BB'=B'$. Then 
+	consider
+	\[0\to B_\lambda\to B\to B/B_\lambda.\]
+	After tensoring with (faithfully flat!) $B'$ over $B$, since for some $\lambda$
+	\[0\to B_\lambda\otimes_B B'\xrightarrow{\sim}B'\to B/B_\lambda\otimes_BB'\to 0\]
+	is exact, forcing the rightmost term to be zero. But by faithfulness this implies $B/B_\lambda=0$ and we are done.
+\end{rmk}
+\begin{prop}
+	Consider
+	\begin{center}
+		\begin{tikzcd}
+			X'\ar[r]\ar[d] & X\ar[d]\\
+			S'=\Spec A'\ar[r,"\text{f.type, f.flat}",two heads,swap] & S=\Spec A
+		\end{tikzcd}
+	\end{center}
+	which is a Cartesian square. Then 
+	\begin{enumerate}
+		\item $X'\to S'$ is an isomorphism iff $X\to S$ is.
+		\item $X'\to S'$ is affine iff $X\to S$ is.
+	\end{enumerate}
+\end{prop}
+
 \end{document}

### 2019-10-23 Algebraic Geometry/Algebraic-Geometry.tex
update

@@
 
+\section{October 23rd, 2019}
+The plan moving forward: we are going to talk about some more ideas not from the book about topology and other topics.
+Then we will rejoin the book with chapter three (skipping the discussion of dimension theory).
+
+Here's where we're aiming: last time we talked about what it meant for a Noetherian topological space (or variety) to be irreducible.
+In particular for any Noetherian topological space.
+\begin{thm}\label{thm:decomp}
+	$X$ admits a decomposition 
+	\[X=X_1\cup\cdots\cup X_n\]
+	with each $X_i$ irreducible and for all $i\ne j$, $X_i\not\subseteq X_j$.
+
+	Furthermore, given two such decompositions 
+	\[X=X_1\cup\cdots\cup X_n=Y_1\cup\cdots\cup Y_m\]
+	then there exists a function $\iota:[n]\to[m]$ such that $X_i=Y_{\iota(i)}$.
+\end{thm}
+Recall that when $X=\Spec A$, the irreducible components $X_i$ of $X$ correspond precisely to the minimal primes $\frakp\lhd A$.
+
+\subsection{Noetherian Induction}
+Let $X$ be a Noetherian topological space and let $P$ be a set of closed subsets of $X$. Suppose that for all $Y\subseteq X$ closed, if 
+\[\text{for all closed $Z\subsetneq Y$, $Z\in P$ implies that $Y\in P$ }\]
+Then $X\in P$
+\begin{prf}
+	By contrapositive: assume that $X\in P^c.$ But now we can use axiom of choice: there exists a minimal $W\in P^c$. 
+	But then for any closed $Z\subsetneq P$, we have $Z\in P$, so $W\in P$, a contradiction. But then $P^c=\varnothing=X$.
+\end{prf}
+
+Now we can prove theorem~\ref{thm:decomp}
+\begin{prf}
+	$\varnothing$ has the empty decomposition. Thus $\varnothing\in P$. Now if $Y\subset X$ is closed, then either $Y$ is irreducible and $Y\in P$ 
+	(``admits a decomposition'') or else $Y=Z\cup W$ and assuming each are in $P$ we get that $Y\in P$ by concatenating (and eliminating). Thus by Noetherian induction we get our decomposition.
+
+	To get the last bit, take two decompositions. Then consider 
+	\[X_1=X_1\cap X=(Y_1\cap X_1)\cup\cdots\cup(Y_m\cap X_1).\]
+	Then since $X_1$ is irreducible, $X_1\subseteq X_1\cap Y_j$ for some $j$ and thus $X_1\subseteq Y_j$. But then similarly $Y_j\subseteq X_i$ for some $i$.
+	But by assumption on the $X_k$, $X_1=Y_j=X_i$, giving us our $\iota$.
+\end{prf}
+
+\subsection{Important Morphisms}
+\begin{defn}
+	A morphism $f:X\to Y$ of varieties is 
+	\begin{itemize}
+		\item \textbf{affine} if there exists an affine covering $U_i\subseteq Y$ such that $f^{-1}(U_i)\subseteq X$ is affine.
+		\item \textbf{finite} if there exists an affine covering $U_i\subseteq Y$ such that $f^{-1}(U_i)$ is affine and 
+		\[\O_Y(U_i)\to\O_X(f^{-1}(U_i))\]
+		is a finite ring extension.
+	\end{itemize}
+\end{defn}
+\begin{prop}
+	$f:X\to Y$ is affine (resp. finite) if and only if for all $U\subseteq Y$ affine, $f^{-1}(U)$ is affine (resp. $f^{-1}(U)$ is affine 
+	and $\O_X(U)\to \O_X(f^{-1}(U))$ is finite).
+\end{prop}
+
 \end{document}

### 2019-10-23 Algebraic Groups/Algebraic-Groups.tex
update

@@
 
+\section{October 23rd, 2019}
+Recall the following definition/proposition:
+\begin{defn}
+	Let $S=\Spec R$ be Noetherian. Let $G\to S$ be an affine group scheme that is flat and of finite type. Let $G$ be a scheme over $S$ with a right $G$-action.
+
+	Then the following are equiavlent:
+	\begin{itemize}
+		\item $P\to S$ is a \textbf{$G$-torsor}
+		\item $P\to S$ is faithfully flat and of finite type and $P\times_SG\xrightarrow{(\sigma,\pi_1}P\times_S P$ is an isomorphism.
+		\item There exists $S'\twoheadrightarrow S$ faithfully flat such that 
+		\begin{center}
+			\begin{tikzcd}
+				G\times_S S'\cong P\times_S S'\ar[r]\ar[d] & P\ar[d]\\
+				S'\ar[r] & S
+			\end{tikzcd}
+		\end{center}
+		commutes (where the isomorphism shown is $G\times_SS'$-equivariant.)
+	\end{itemize}
+\end{defn}
+
+\subsection{Some Examples}
+We have the trival torsor $P=G\to S$. It is a propositoon that $P\to S$ is trivial iff there exists a section $s:S\to P$.
+
+Let $L/K$ be a finite Galois extension. Then we get $\operatorname{Gal}(L/K)$ acting on $P=\Spec L\to \Spec K$ is a $G=\operatorname{Gal}(L/K)$-torsor.
+Then in the diagram in the definition above, $\Spec L$ plays the part of $S'$. Then we get that $L\otimes L\cong L[x]/f\cong \prod_{g\in G}L$ and 
+\[G\times_{\Spec K}\Spec L\cong \Spec(L\otimes L)\cong \sqcup_{g\in G}\Spec L.\]
+
+\brk
+
+Now let $X$ be a scheme and let $\Gm$ act on a line bundle $L\to X$ with section $o:X\to L$. Then $(L\setminus o(X))\to X$ is a $\Gm$ torsor.
+It is a result, although we don't have the machinery yet, that 
+\begin{prop}
+	There is a bijection between line bundles on $X$ and $\Gm$-torsors.
+\end{prop}
+\begin{rmk}
+	We can do womething similar with any vector bundle over $X$: if $V\to X$ is one, then $V_x$ over $x\in X$ is a vector space. 
+	We just send $V$ to $\operatorname{Frame}(V)$, which over any $x\in X$ we have the set of orderd bases of $V_x$. This gives us a $\GL_n$-torsor.
+\end{rmk}
+
+\brk
+
+If we have a subgroup (say of an algebraic group) $H\subseteq G$, recall that we wanted to show the existence of an $H$-torsor $G\to G/H$.
+
+We begin by talking about \textit{abstract groups}. Assume we have an exact sequence 
+\[1\to K\to G\xrightarrow{\pi}Q.\]
+Then $G\times K\xrightarrow{\sim}G\times_QG$ via the map $(g,k)\mapsto(g, gk)$. The proof isn't too bad.
+
+So now consider a geometric group. If we have the same exact sequence of \textit{algebraic groups over $k$}, we get $K=G\times_Q\Spec k$(??)
+and then evaluating at any scheme $S$, we get 
+\[1\to K(S)\to G(S)\to Q(S)\]
+and consider the map $G(S)\times K(S)\to G(S)\times_{Q(S)}G(S)$ as above. Then by Yoneda we get an isomorphism $G\times K\cong G\times_Q G$ of schemes.
+
+\begin{cor}
+	If $\pi:G\to Q$ is a faithfully flat map of linear algebraic groups over $k$, then $G\to Q$ is a torsor under $K=\ker\pi$
+\end{cor}
+Let's do even better!
+\begin{cor}
+	Let $\pi:G\to Q$ be dominant (i.e. the image is dense in $Q$) and furthermore that $Q$ is reduced. Then $G\to Q$ is faithfully flat and in particular $G\to Q$ is a $K$-torsor.
+\end{cor}
+
+\begin{prf}
+	(Of the second corollary): We use the idea of ``generic flatness''. That is there exists a $U\subseteq Q$ such that $\pi^{-1}(U)\to U$ is flat. Then we can translate this by 
+	the $G$-action (after passing to the algebraic closure of $k$ so that the points are dense) and then flat descent gives us the result we want. :)
+\end{prf}
+
+\begin{thm}
+	Let $G$ be alinearly algebraic group over $k$. Let $X$ be a scheme over $k$ of finite type.
+	Then
+	\[\{G\text{-bundles on $X$}\}\cong H_{fl}^1(X,G)\]
+	where $H_{fl}$ is flat cohomology.
+\end{thm}
+The idea here is clear when $G=\Gm$: you get connections between line bundles on $X$ (i.e.
+the Picard group of $X$) and $\Gm$-torsors and similarly between the line bundles and $H_{Zar}^1(X,\O_X^\times)$,
+using the (usual) Zariski sheaf cohomology.
+
 \end{document}

### 2019-10-26 Algebraic Geometry/Algebraic-Geometry.tex
update

@@
 		\[\O_Y(U_i)\to\O_X(f^{-1}(U_i))\]
-		is a finite ring extension.
+		is a finite ring extension (finitely generated as a module over the base).
 	\end{itemize}
@@
 
+\section{October 25th, 2019}
+Another takeaway from this class: \textit{algebraic geometry is relative and not absolute.} One way we are starting to dip our toes into this idea 
+is via the definition of affine and finite morphisms. Notice that the definition of finite is a bit clunky because it requires us 
+to restate the definition of affine withing the definiton of a finite morphism. Gross. We should be able to write it as \textit{``affine + $\square$.''}
+
+For some examples, let $Y=\Spec k$. Then $X\to \Spec k$ is affine if and only if $X$ is affine. So when is $X\to\Spec k$ finite? Well since $X$ is affine, $X=\Spec A$ 
+where each $A$ is a finite dimensional (reduced!) $k$-algebra. Therefore it is Artinian and we get a decomposition 
+\[A=A_1\times\cdots\times A_n\]
+where each $A_i$ are finite dimensional local rings with maximals $\frakm_i\subset A_i$ that are nilpotent (use Nakayama!). So by this theorem (since $A$ is reduced),
+we get $A_i=k$, so 
+\[A\cong k\times\cdots\times k.\]
+So then $\Spec A=\sqcup \Spec k$, just a handful of points!
+
+Let's look at another example we saw on the first day: $k[x]\to k[t]$ sending $x\mapsto t^2$. This corresponds to a map $\A1\to\A1$, 
+and we can talk about the fibers of this morphism. They are all (well, almost all) of size two. But they are definitely finite! Similarly if 
+we have a finite map $A\to B$, we can ask if it has finite fibers. Here if $\varphi:A\to B$ is our morphism and $f:\Spec B\to\Spec A$ is the map of rings,
+then since
+\[f^{-1}(\frakm)=\{\frakn\subseteq B|\varphi(n)=m\}\]
+we get a lemma 
+\begin{lem}
+	Given a map $f:\Spec B\to \Spec A$ and $\frakm\in\Spec A$, 
+	\[f^{-1}(\frakm)=\{\frakn\subseteq B/\frakm B,\text{ maximal}\}\]
+\end{lem}
+And so (think this through) the answer is \textbf{yes, the fibers are finite.}
+
+A natural question involves the following 
+\begin{defn}
+	A morphism $f:X\to Y$ is \textbf{quasi-finite} if for all $y\in U$, $f^{-1}(y)$ is finite.
+\end{defn}
+and we ask: is this equivalent to finiteness? The answer is no. Consider the map $\A1\setminus 0\hookrightarrow\A1$. Both of these are affine, so the map is affine. 
+But definitely the $k[t]\to k[t,t^{-1}]$ is not finite!
+
+\begin{prop}
+	Let $X\to Y$ be finite and $Z\subseteq X$ clsoed. Then $f(Z)\subseteq Y$ closed.
+\end{prop}
+Try it!
+
 \end{document}

### 2019-10-26 Algebraic Groups/Algebraic-Groups.tex
update

@@
 
+\section{October 25th, 2019}
+We are going to deviate slightly from the result promised last time, because we want to talk first a bit about 
+\subsection{Descent}
+Recall that we had two results before about descent involving faithfully flat morphisms of finite type (where the schemes are Noetherian).
+
+\begin{ex}Let $X=\Spec B$ and let $U_i$ be a Zariski-open cover. Then the map 
+\[\sqcup U_i\to X\]
+is faithfully flat and of finite type.
+\end{ex}
+So then one idea is that if we have a nice Zariski cover of $X$, flatness, faithful flatness, and finite type are all ``local on the source''
+in that you just have to check the property locally on $X$.
+
+\brk
+
+\begin{prop}
+	Let $A\to B$ be a faithfully flat ring homomorphism. Then 
+	\begin{center}
+		\begin{tikzcd}
+			A\ar[r] & B\ar[r,shift left=0.75ex, "p_1"]\ar[r,shift right=0.75ex,"p_2",swap] & B\otimes _AB
+		\end{tikzcd}
+	\end{center}
+	where $p_1(b)=b\otimes 1$ and $p_2(b)=1\otimes b$ is an exact sequence. More generally, 
+	\begin{center}
+		\begin{tikzcd}
+			M\ar[r] & M\otimes_A B\ar[r,shift left=0.75ex,"\id\otimes p_1"]\ar[r,shift right=0.75ex,swap,"\id\otimes p_2"] & M\otimes_A B\otimes_A B
+		\end{tikzcd}
+	\end{center}
+	is exact.
+\end{prop}
+
+There is some geometry that lost me a bit. Sorry. :(
+\begin{prop}
+	If $A\to B$ is faithfully flat, then the map $\rmod A\to\{(N,\alpha)|N\in\lmod B, \alpha:p_1^\ast N\cong p_2^\ast N, P(\alpha)\}$ 
+	sending
+	\[M\mapsto(M_B=M\otimes_AB,\alpha_{can})\]
+	is 
+	an equivalence of categories where $P(\alpha)$ means that $\alpha$ satisfies a cocyle condition. 
+\end{prop}
+\begin{rmk}
+	The canonical isomorphism above is 
+	\[\alpha_{can}:M_B\otimes_{B,p_1}(B\otimes_AB)\to M_B\otimes_{B,p_2}(B\otimes_A B)\]
+\end{rmk}
+
 \end{document}

### 2019-10-29 Algebraic Geometry/Algebraic-Geometry.tex
update

@@
 
+\section{October 28th, 2019}
+Max heard some things from the feedback: he is interested in seeing if he can't help us 
+put things together at the end of class. That should help!
+
+\subsection{Products}
+Today we are going to be talking about products (and fiber products) in the category of varieties. 
+\begin{defn}
+	Let $\calC$ be a category and $X,Y\in\calC$. Then the product is the usual thing with the usual universal property.
+	\begin{center}
+		\begin{tikzcd}
+			& X\prod Y\ar[dl,"\pi_X"]\ar[dr,"\pi_Y"] &\\
+			X & & Y\\
+			& A\ar[ul]\ar[ur]\ar[dashed,uu] &
+		\end{tikzcd}
+	\end{center}
+\end{defn}
+\begin{rmk}
+	Notice that we can also ``think in terms of functors.'' THen we can define the product of $h_X$ and $h_Y$ as $h_X\times h_Y$ 
+	in terms of the (Cartesian) product of sets.
+\end{rmk}
+Also recall the definition of a fiber product. Notice that if we translate back to functors, we are again trying to figure out what to 
+make of the set 
+\[h_X\times_{h_Z}h_Y.\]
+If $f$ and $g$ are maps from $h_X$ and $h_Y$, respectively, into $h_Z$, then the fiber over $x\in h_Z$ in $h_X\times_{h_Z}h_Y$
+is precisely the product of the fibers (whhaaaaat \textit{like a fibered product}) in $f$ and $g$.
+
+\begin{prob}
+	If $\calC$ has a final object $\Omega$, then $X\times Y=X\times_\Omega Y$. THink about this.
+\end{prob}
+\begin{thm}
+	Fiber products exist in the category of varieties.
+\end{thm}
+Today we are going to be discussing the affine case: given $X\to Z\leftarrow Y$ of affines, constructing 
+$X\times_Z Y$. To start, let $Z=\Spec k$. We are looking to study the functor $h_X\times h_Y$ and then show representability.
+
+Let's look at this pointwise: let $X=\Spec A$ and $Y=\Spec B$. Then
+\[h_X(S)\times h_Y(S)=\Hom(S,X)\times \Hom(S,Y)=\Hom(A,\O_S)\times\Hom(B,\O_S)=\Hom(A\otimes B,\O_S)=h_{\Spec A\otimes B}(S)\]
+\begin{prob}
+	Show that the coproduct in $\Alg_k$ is the tensor product.
+\end{prob}
+\begin{rmk}
+	For the above problem, notice that $A\otimes B$ is odd notation because it is not just the underlying vector space, but carries with it 
+	an algebra structure.
+\end{rmk}
+\begin{prob}
+	Is the tensor product of a finitely-generated algebra finitely generated? Can you do it without coordinates?
+\end{prob}
+
+But notice for what we wrote above to be ``okay'' we need that $A\otimes_k B$ is a reduced finitely generated $k$-algebra. The finite generation 
+isn't too hard to see. The reducedness comes from the following (in Kempf):
+
+We ahve a map $A\otimes B\to \{|\Spec A|\times |\Spec B|\to k\}$ as set of functions. The claim is that this map is injective.
+The idea is that if $\sum_i f_i\otimes g_i\mapsto 0$, we can assume that the $f_i$ and $g_i$ are linearly independent.
+
+But then for all $x,$ we get $\sum_i f_i(x)g_i=0:\Spec B\to k$. But then since the $g_i$ are linearly independent,
+this means the $f_i(x)$ are all zero. Thus the nullstellensatz tells us that all $f_i=0$.
+
 \end{document}

### 2019-11-01 Algebraic Geometry/Algebraic-Geometry.tex
update

@@
 
+\section{October 30th, 2019}
+Let's start by discussing the idea that for any final object $\Omega$, $X\times_\Omega Y\cong X\times Y$. One way to do this 
+is to show that $X\times Y$ also has the universal property corresponding to the limit of $X\to \Omega\leftarrow Y$. But we can 
+also do things functorially! Notice that we have a map 
+\[h_{X\times_\Omega Y}=h_X\times_{h_\Omega}h_Y\xrightarrow{\cong}h_X\times h_Y=h_{X\times Y}\]
+where we can show the isomorphism of these things on points.
+
+Let's talk about the fiber product 
+\[\Spec A\times_{\Spec C}\Spec B\]
+which intuitively should be something like $\Spec A\otimes_C B$. But is this going to be reduced still? Let's do an example. Consider $A=k[t]$ and $C=k[x]$ 
+and consider the map $x\mapsto t^2$ as a map $C\to A$. Then let $B=k$ with the map that sends $x\mapsto 0$. Then $A\otimes_C B\cong k[\varepsilon]/\varepsilon^2$.
+
+The takeaway here is that the tensor product is sometimes non-reduced! It ends up that if we take $(A\otimes_C B)_{red}$ is the algebra we want, which is the 
+appropriate object \textit{in the category of reduced $k$-algebras.}
+\begin{prob}
+	Consider the maps $\Spec(-)$ and $\O(-)$, between the categories $\Alg_k^{op}$ and varieties. These form an adjoint pair! Which object is 
+	left adjoint? Do left adjoints commute with limits? colimits?
+\end{prob}
+
+\subsection{A series of fun exercises}
+Here's some work to do!
+\begin{thm}
+	Fiber products exist in the category of varieties.
+\end{thm}
+\begin{rmk}
+	You should prove this! We will discuss the idea here. The only thing we know how to do is do this with affines! 
+	reduce to the affine case (for all three) and glue!
+	\begin{itemize}
+		\item Begin by showing: Suppose $U\subseteq Z$ is open and $X\times_Z Y$ exists. Then we can look at 
+		\begin{center}
+			\begin{tikzcd}
+				(X\times_ZY)_U\ar[d]\ar[r,hookrightarrow]& X\times_Z Y\ar[d]\\
+				U\ar[r,hookrightarrow] & Z
+			\end{tikzcd}
+		\end{center}
+		and we claim that $(X\times_ZY)_U$ representas $h_{X_u}\times_{h_u}h_{Y_u}$. This gives us a way to shrink over the base.
+		\item Now suppose $W\subseteq Y$ is open. Then looking at a similar diagram, we want to show that $(X\times_ZY)_W$ represents $h_X\times_{h_Z}h_W$. The same thing can be done with a subset of $X$.
+	\end{itemize}
+	Now we want to reduce to $Z$ being affine, then reduce to $Y$ is affine, then $X$. Then we are golden.
+	For the first, assume it holds when $Z$ is affine and then show it works for arbitrary $Z$: pick a covering $U_i$ of $Z$ by affines. Consider $U_i$ and $U_j$:
+
+	Each $(X_{U_i}\times_{U_i}Y_{U_i})_{U_i\cap U_j}\cong (X_{U_j}\times_{U_j}Y_{U_j})_{U_i\cap U_j}$ via a canonical isomorphism $\phi_{ij}$. This is the crux of it. THINK. :)
+\end{rmk}
+
+\section{November 1st, 2019}
+Let's look at some examples of things. $\A1\times\A1=\Spec k[x]\otimes k[y]\cong\Spec k[x,y]\cong \A2$. Another way to see this 
+is to think about the functors:
+\[\Hom(X,\A1\times\A1)\cong\Hom(X,\A1)\times\Hom(X,\A1)=\Hom(k[s,t],\O(X))=\Hom(X,\A2).\]
+Similarly you can do something along these lines with $\A n$.
+
+What about $\P1\times\P1$? It's not $\P2$. Why? Because there are curves $C_1$ and $C_2$ in the product that do not intersect.
+But Bezout's theorem says that any two curves in $\P2$ intersect!
+
+The problem with solving the above problem is that we really need sheaf theory to solve the above problem. 
+We will circle back around to this.
+
+Question: Is the zariski topology on $X\times Y$ the product topology? NO! For instance $\A2$: the product topology is not the cofinite one!
+
+\subsection{Projective spaces}
+\begin{defn}
+	A variety is \textbf{projective} if there exists a closed immersion $X\hookrightarrow \P n$.
+\end{defn}
+\begin{prop}
+	Products of projective varieties are projective.
+\end{prop}
+\begin{prob}
+	Classify all projective objects in the category of varieties! You can do it. :)
+\end{prob}
+\begin{rmk}
+	Notice that right now we are working in the \textit{Italian} projective space. This is the usual $\A n\setminus\{0\}$.
+	Then we say that $Z$ is closed if the closure of its inverse image in $\A n\setminus \{0\}$ is invariant under scaling.
+\end{rmk}
+Now think about the functions on $\P n$. There is a ring $R\subseteq k(x_0,\dots,x_n)$ called the field of homogeneous rational functions. These are the 
+ratio of homogeneous polynomials of the same degree. It's a field baby. Then 
+\[\O_{\P n}(U)=\{\phi\in R|\operatorname{poles}(\phi)\subseteq\P n\setminus U\}.\]
+
+\subsection{The Point}
+There is a famous map called the Segre embedding. We get a map 
+\[\P n\times\P m\hookrightarrow \P {(n+1)(m+1)}\]
+where we can just imagine taking $(v_i)$ and $(w_i)$ and map it to $\mathbf{v}\mathbf{w}^T$.
+
+This is just the projectivization of a matrix! It dends up that it identifies the product with a closed subvariety equal to \textbf{the set of matrices of rank 1.}
+There was something at the end about $2\times 2$ subdeterminants!
+
 \end{document}

### 2019-11-04 Algebraic Geometry/Algebraic-Geometry.tex
update

@@
 
+\section{November 4th, 2019}
+We're going to talk a bit more about projective varieties because we really haven't done a lot of examples and that is a shame.
+
+\subsection{What we want to do}
+By the end of the quarter (we have four more weeks, but only three more weeks worth of meetings with holidays), we want to 
+have talked a bit about sheaves and dimension/intersection theory. Continuing into next quarter, we are going to dip into schemes, although perhaps not immediately.
+
+\subsection{Some tangible examples}
+One idea is to read Harris' \textit{Algebraic Geometry: a first course}. Now we're talking to talk about $\P n$.
+
+First things first, we form $\P n$ as a set as the lines in $k^{n+1}$ or $k^{n+1}\setminus \{0\}/k^\times$. A point in $\P n$ 
+has homogeneous coordinates $[x_0:\cdots:x_n]$, which is only well-defined up to scaling. But if any $x_i$ is nonzero, we can normalize this 
+coordinate to get well-defined coordinates. The collection of sets $\{x_i\ne 0\}=\A n\subset \P n$ is called the standard affine open 
+cover of $\P n$.
+
+A thought: if we have an affine variety $Z\subseteq \A n\subseteq \P n$, how can we view the closure of $Z$ in $\P n$?
+Given $f\in k[x_0,\dots,x_n]$ homogeneous, $Z(f)\subseteq \P n$ is well-defined. More generally, if $I$ is a homogeneous ideal (generated by homogeneous polys)
+$Z(I)$ makes sense and furthermore $Z((f_1,\dots,f_k))=\cap_i Z(f_i)$. It ends up that these are precisely the closed sets in the Zariski topology.
+
+\begin{thm}[Projective Nullstellensatz]
+	$I(Z(J))=\sqrt{J}$ for all homogeneous $J$ excipt $J$ such thtat $\sqrt{J}$ is the irrelevant ideal $(x_0,\dots,x_n)$.
+\end{thm}
+
+\subsection{Homogenization/dehomogenization}
+Given a homogenous polynomial $f(x_0,\dots,x_n)$ of degree $d$, we can \textit{dehomogenize} things by setting $x_0=1$. For instance $x_0^2+x_1^2+x_2^2\mapsto 1+x_1^2+x_2^2$.
+Given a polynomial $g(x_1,\dots,x_n)$, we just look at 
+\[\tilde g(x_0,\dots,x_n)=x^{\deg g}g(x_1/x_0,\dots,x_n/x_0).\]
+One can see relatively easily that $\tilde g$ is a homogeneous polynomial of degree $d$.
+
+\begin{prob}
+	What happens when you homogenize then dehomogenize? How about the other way around?
+\end{prob}
+
 \end{document}

### 2019-11-04 Algebraic Groups/Algebraic-Groups.tex
update

@@
 
+\section{November 4th, 2019}
+Today we are going to do a bit of review as well as do a little more work with $G$-torsors.
+
+\subsection{Review}
+Recall the following setup: $S=\Spec A$ is Noetherian. Then let $G\to S$ be an affine group scheme of finite type over $S$. Let $P\to S$ 
+be a scheme with a right $G$-action:
+\[P\times_S G\xrightarrow{\sigma}P.\]
+\begin{defn}\label{defn:torsor}
+	Then $P\to S$ is a \textbf{$G$-torsor} if either of the following hold:
+	\begin{itemize}
+		\item $P\to S$ is affine, surjective, of finite type, and furthermore transitive (that is $P\times_S G\xrightarrow{(\sigma,\pi_1)}P\times_SP$ is an isomoporphism).
+		\item There exists a faithfully flat $S'\to S$ with $P\times_S S'\cong G\times_S S'$ withere this isomorphism is $G\times_S S'$-equivariant.
+	\end{itemize}
+\end{defn}
+\begin{rmk}
+	If $G$ is a linear algebraic gorup over $k$ and $S$ is defined over $k$, then we say $P\to S$ is a $G$-torsor if it is a torsor under $G\times_k S$.
+\end{rmk}
+
+Now a \textbf{trivial $G$-torsor} over $S$ is $S\times G\to S$ if and only if $P\to S$ has a section.
+
+Some examples:
+\begin{itemize}
+	\item The $\Gm$ torsors over $S$ are in bijection with the line bundles over $S$.
+	\item The $\GL_n$ torsors over $S$ are in bijection with the vector bundles over $S$.
+\end{itemize}
+Since line and vector bundles are tivialized in the Zariski topology, we see that the second condition in definition~\ref{defn:torsor} can be replaced with: 
+There exists an open cover $\{S_i\}$ such that $P|_{S_i}\cong S_i\times G$.
+
+\brk
+
+We proved some stuff with descent. Look at it.
+\begin{cor}
+There exists an equivalence between $\Alg_A$ and the category of $B$ algebras $C$ along with isomorphisms $\alpha:p_1^\ast C\to p_2^\ast C$ along with the cocyle condition.
+\end{cor}
+\begin{rmk}
+	To prove this, we descend a $(B,\alpha)$ to an $A$-module and show that the multiplication on $B$ descends nicely to $A$.
+\end{rmk}
+\begin{cor}
+	We can also descend $G$-torsors. That is, we have an equivalence between $G$-torsors $P\to S$ and $G$-torsors $P'\to S'$ with isomorphisms and the cocycle condition.
+\end{cor}
+This last result takes quite a bit of doing although we have all the machinery we need. This is the 
+one we're going to be using.
+
+\subsection{New Stuff}
+Let $X$ be a scheme. Then $\operatorname{Pic}(X)$ is the group of line bundles, or equivalently $\Gm$-torsors on $X$.
+\begin{thm}
+	$\operatorname{Pic}(X)=H^1(X,\O_X^\times)=H^1(X,\Gm)$.
+\end{thm}
+Notice that $\O_X^\times$ is a sheaf assigning to each $U$ $\O_X(U)^\times$. On Zariski opens, this is the same as $\Gm(U)$.
+
+To prove the above result, we need to discuss 
+\subsection{\texorpdfstring{\v C}{C}ech Cohomology}
+Assume $X$ is separated. Let $\calU=\{U_i\}$ be an affine covering where $U_i\cap U_j$ is also affine. Then considering the complex 
+\[\sqcup U_i\cap U_j\cap U_k\to \sqcup U_i\cap U_j\to \sqcup U_i\to X\]
+where we have several a parallel maps for each $n-1$-tuple of intersections in the previous term. Then we can applie $\O_X^\times$:
+\[\prod \O_X(U_i)^\times\to \prod\O_X(U_i\cap U_j)^\times\to\cdots\]
+where, for instance, the first map above sends a product of maps $(s_i)$ to $(s_i|_{U_i\cap U_j}-s_j|_{U_i\cap U_j})$.
+
+\begin{defn}
+	$\hat H_\calU^i(X,\O_X^\times)$ is the $i^{th}$ cohomoogy of the above complex.
+\end{defn}
+Then it can be shown that $\hat H_\calU^i(X,\O_X^\times)$ is independent of cover if the $U_i$ are affine.
+
+So what is $H^1(X,\O_X^\times)$? It is the $s_{i,j}in\O_X^\times(U_i\cap U_j)$ modulo the cocycle condition.
+
+\subsection{A new result}
+\begin{thm}
+	$H^1(X,G)$ is identified with $G$-torsors.
+\end{thm}
+What is $H^1(X,G)$? We are going to try to recover \v Cech cohomology. Let $X$ be quasi-compact and let $\cup_1^n U_i\to X$ be 
+faithfully flat. The seperatedness gets us the intersections are affine.
+
+Then we can play the same game, but this time applying $G$: (notice that $U_i\cap U_j=U_i\times_X U_j$)
+\[\prod G(U_i)\to \prod G(U_i\times_X U_j)\to\cdots\]
+where (notice now $G$ may be nonabelian!) we map 
+\[(s_i)\mapsto (s_i|_{U_i\times_X U_j}\cdot s_j|_{U_i\times_X U_j}^{-1})_{i,j}\]
+and 
+\[(s_{i,j})\mapsto (s_{i,j}\cdot s_{j,k}\cdot s_{i,k}^{-1})\]
+where each $s$ is restricted to $U_i\times_XU_j\times_X U_k$.
+
+Then we can define $\hat H_\calU^1(X,G)$ to be the first cohomology of the above chain. Then 
+\begin{defn}
+	The \textbf{flat cohomology} is 
+	\[H^1_\text{flat}(X,G)=\operatorname{colim}_\calU H^1_\calU(X,G).\]
+\end{defn}
+
+The overall result here is 
+\begin{thm}
+	The $G$-torsors on $X$ in bijection with the elements of $H^1_\text{flat}(X,G)$.
+\end{thm}
+
 \end{document}

### 2019-11-06 Algebraic Geometry/Algebraic-Geometry.tex
update

@@
 Given a polynomial $g(x_1,\dots,x_n)$, we just look at 
-\[\tilde g(x_0,\dots,x_n)=x^{\deg g}g(x_1/x_0,\dots,x_n/x_0).\]
+\[\tilde g(x_0,\dots,x_n)=x_0^{\deg g}g(x_1/x_0,\dots,x_n/x_0).\]
 One can see relatively easily that $\tilde g$ is a homogeneous polynomial of degree $d$.
@@
 
+\section{November 6, 2019}
+Today we are going to focus on the pickles and not the meat. Go Max.
+
+\subsection{Questions and Thoughts}
+First things first, we wanted to talk about the problem above. It is easy to see that homogenizing and then dehomongenizing gives you the same polynomial.
+For the other direction, it seems that a (homogeneous) polynomial comes back to itself if it is not divisible by a power of $x_0$.
+
+What is happening geometrically? Dehomogenizing is intersecting with an open affine subset. Homogenizing is taking the closure.
+\begin{rmk}
+	The idea came up that taking the closure might be the same as taking the cone containing the original thing. But consider 
+	the hyperbola $xy-1$ and its homogenization $xy-z^2$. Now $Z=Z(xy-z^2)$ is closed and contains $H=Z(xy-1)$, so it contains $\bar H$.
+
+	Conjecture: $Z\cong \P1\subseteq\P2$ as a conic. How can we realize this as a map from $\P1$ to $\P2$? Well,
+	we need a map 
+	\[(x_0,x_1)\mapsto(f_0(x_i),f_1(x_i),f_2(x_i))\]
+	where the $f_i$ are homogeneous of the same degree and have no common zeros in $\P1$.
+\end{rmk}
+\begin{rmk}
+	More generally, given varieties $X$ and $Y$, a \textbf{rational map} from $X$ to $Y$ $f:X\dashrightarrow Y$ is an equivalence class of morphisms $U\to Y$ where $U$ is open in $X$ 
+	and $U\xrightarrow{\alpha} Y$ and $V\xrightarrow{\beta}X$ are equvialent if $\alpha|_{U\cap V}=\beta|_{U\cap V}$.
+
+	Then we can choose $f_0,\dots,f_m\in k[x_0,\dots,x_n]$ that are homogeneous of degree $d$ and these coordinate functions give us a rational map (in fact, all the rational maps) from $\P n$ to $\P m$.
+\end{rmk}
+\begin{prob}
+	Find a morphism $f:\P1\to\P2$ whose image is $Z(xy-z^2)$.
+\end{prob}
+I am thinking $x_0^2$, $x_1^2$, and $x_0x_1$.
+\begin{prob}
+	Suppsoe that $Q$ is a non-singular quadratic form in three variables, so $Z(Q)\subseteq \P2$. Then show that $Z(Q)\cong \P1$.
+	\textit{Hint: Linear changes of coordinates are algebraic!}
+\end{prob}
 \end{document}

### 2019-11-06 Algebraic Groups/Algebraic-Groups.tex
update

@@
 
+\section{November 6th, 2019}
+Today we are going to prove some more things about representation theory. Down the pipe somewhere we will hope to talk about Tanakka duality, 
+which gives us that Morita equivalence (is this still what it's called?) of two algebraic groups gives us the groups are isomorphic.
+
+Given an algebraic group $G/k$, we could consider representations, which were either a group scheme morphism $G\to \GL(V)$ or else a coaction $V\xrightarrow{\sigma} \Gamma(G)\otimes_k V$.
+Then we said that $G$ was linear reductive if and only if every representation is completely reducible,
+
+\begin{ex}
+	If $G=\Gm$, then every irreducible representation is one-dimensional: $\Gm\to\Gm$ sends $t\mapsto t^\alpha$.
+\end{ex}
+A natural question one may ask: what can we say in general?
+\subsection{The regular representation}
+The regular representation is $\Gamma(G)$ with coaction given by comultiplicaton.
+\begin{prop}
+	And finite dimentional representation $V$ of $G$ embeds $V\subseteq \Gamma(G)^{\oplus n}$.
+\end{prop}
+\begin{prf}
+	The coaction gives us a map 
+	\[V\xrightarrow{\sigma}\Gamma(G)\otimes V=\Gamma(G)\otimes_k V_{\text{vs}}\cong \Gamma(G)^{\oplus\dim V}\]
+	where $V_{vs}$ is the underlying vector space of $V$. Then the claim is that the overall map is injective and that it is a map of $G$-reps.
+
+	The latter property is almost tautological by the property of a coaction. Injectivity follows from the fact that 
+	\[(e^\ast\otimes\id)\circ\sigma\]
+	is an isomorphism $V\to k\otimes V$ (again by one of the coaction axioms), so $\sigma$ is injective.
+\end{prf}
+\begin{prop}
+	If $V$ is a finite dimensional faithful representation, then every other finite-dimensional representation can be obtained from $V$ by direct sums, tensors, duals, subrepresentations, and quotients.
+\end{prop}
+\begin{rmk}
+	More precisely, $W\subseteq (V\oplus V^\vee)^{\otimes n}$.
+\end{rmk}
+\begin{prf}
+	$G\subseteq\GL(V)$ is a closed subgroup. Notice that since we have an injective map into $\GL(V)$, $G$ is already a linear algebraic group.
+\end{prf}
+\begin{thm}[Peter-Weyl]
+	If $G$ is linear reductive, then 
+	\[\Gamma(G)=\oplus_{\text{irr $V$}}(V\otimes V^\vee)\]
+\end{thm}
+\begin{rmk}
+	Notice that this is as two-sided representations, but if we only want to look at (say) left representations, we give $V^\vee$ the trivial 
+	representation structure and we get the more familiar 
+	\[\Gamma(G)=\oplus_{\text{irr $V$}} V^{\oplus\dim V}\]
+	and then the formula from finite groups:
+	\[|G|=\sum_{\text{irr $V$}} (\dim V)^2\]
+\end{rmk}
+As an example, consider $\Gm=\Spec k[t]_t$ and $k[t]_t=\oplus_{n\in\bbZ}k\langle t^n\rangle$. As an exercise, thing about what happens for $\SL_2$.
+
+\subsection{Stabilizers of subspaces}
+If $G$ is an algebraic group over $k$ and $V$ is a $G$ representation, let $W\subseteq V$ be a subspace. We want a subgroup $G_W$ which plays a similar role as the stabilizer in group theory.
+
+Consider $G$ as a functor from $\Alg_k\subseteq \Sch/k\to\Grp$ taking $T\mapsto\Hom(T,G)$. Then for a $k$-algebra $R,$ $G(R)$
+acts on $V_R=V\otimes R$.
+\begin{prop}
+	The functor $\Alg_k\to\Grp$ sending 
+	\[R\mapsto \{g\in G(R)|gW_R=W_R\}\]
+	is representable by a subgroup of $G$, which we will denote $G_W$.
+\end{prop}
+\begin{rmk}
+	The idea here is to fix a basis $e_i$ of $W$ and complete them to a basis of $V$ by appending $f_i$. Then take the coaction $V\to\Gamma(G)\otimes V$ and consider the image 
+	\[f_k\mapsto \sum_{i\in I}a_{ki}\otimes e_i+\sum_{j\in J}a_{kj}\otimes f_j\]
+	suppose we ahve $g\in G(R)=\Hom(\Spec R, G)$. This gives us a map in $\Hom(\Gamma(G),R)$. Then using the action on $V\otimes R$:
+	\[g\cdot(f_k\otimes 1)=\sum g(a_{ki})\otimes e_i+\sum g(a_{kj})\otimes f_j\in V_R\]
+	and we see that this is actually in $W_R$ exactly when $g(a_{kj})=0$ for all $j$.
+
+	But then this means that $g$ lands in $V(a_{kj})$, the vanishing of the functions $a_{kj}\otimes 1\in\Gamma(G)_R$, so we get that 
+	the closed subscheme $V(a_{kj})$ represents the functor $G_W$.
+\end{rmk}
+
+For the converse, let $G$ be an algebraic group over $k$ and let $H\subseteq G$ be a subgroup. Then there exists a finite dimensional representation $V$ of $G$ and 
+a $L\subseteq V$ a one-dimensional subspace such that $H=G_L$.
+
+As an example, look at $G=\SL_2$ acting on $k^2$. Let $L=\langle 1,0\rangle$. We can see easily that $L$ is preserved by $g\in G$ if and only if the lower-left coordinate of $g$ is zero.
+Now if we take $\Gm=\operatorname{diag}(t,t^{-1})\subseteq\SL_2$, consider the representation $\langle xy\rangle\subseteq \operatorname{Sym}^2k^2$. This isn't quite it but maybe you can work it out!
+
 \end{document}

### 2019-11-08 Algebraic Geometry/Algebraic-Geometry.tex
update

@@
 \end{prob}
+
+\section{November 8th, 2019}
+Last time we were thinking about conic curves in $\P2$, that is $Z(Q)$ where $Q$ is a nondegenerate quadratic form in $x_0,x_1,x_2$. One thing that people tried 
+was to classify the ways that $C=Z(Q)$ intersects with copies of $\A2$. Then using these options, you can proceed. This leverages automorphisms of 
+$\P2$ to reduce everything to a very finite set of cases.
+
+Another idea was to do a linear change of coordinates that takes $C$ to $Z(xy-z^2)$. This amounts to doing some work on the matrix representing $Q$. Then answer it for a hyperbola.
+
+For the other question, we were trying to find a good morphism $\P1\to\P2$ with image the hyperbola. The idea was the one I said. It can be seen relatively easily to be bijective.
+\begin{rmk}
+	Is a bijective morphism an isomorphism?! (No) The idea to consider is $R=k[x,y]/(y^2-x^3)\to k[t]$ where $k[t]$ is the normal closure of the former. Then if we are looking at the fiber over $0$, 
+	we compute $R/(x,y)\cong k\to k[t]/(t^2,t^3)\cong k[t]/t^2$
+\end{rmk}
+\begin{rmk}
+	Another idea that came up: if we are thinking about the geometry behind the homogenization/dehomogenization maps, notice that this tells us something about how 
+	dehomogenizing is not a unique process!
+\end{rmk}
+
+\subsection{The Veronese Embedding}
+Given $n$ and $d$, there is a (somewhat) canonical map 
+\[\P n\to \P{\binom{n+d}{d}-1}\]
+where the coordinate functions are the degree $d$ monomials in $x_0,\dots, x_n$. The image here is ``cut out by quadrics.'' These quadrics 
+come from $MN=PQ$ for (not necessarily distinct) coordinate functions.
+
+Recall the Segre embedding. When $n=m=1$, we get an embedding 
+\[\P1\times \P1\to\P3\]
+where 
+\[(x_0:x_1),(y_0,y_1)\mapsto (x_0y_0,x_0y_1,x_1y_0,x_1y_1)=(z_0,z_1,z_2,z_3).\]
+Here the relation is $z_0z_3=z_1z_2$.
+
+Some problems:
+\begin{prob}
+	Show that the image of $\P1\times\P1$ in $\P3$ above is $Z(z_0z_3=z_1z_2)$.
+\end{prob}
+\begin{prob}
+		Show that $\P1\times \P1\not\cong \P2$.
+\end{prob}
+\begin{prob}
+	For any non-degenerate quadratic form $z_0,\dots,z_3$, do we have that $Z(Q)\cong \P1\times\P1$?
+\end{prob}
+\begin{prob}
+	Are all the smooth quadratic hypersurfaces of a fixed dimension isomorphic? Are they all isomorphic to $(\P1)^d$?
+\end{prob}
+\begin{prob}
+	What about cubic hypersurfaces?
+\end{prob}
+
 \end{document}

### 2019-11-08 Algebraic Groups/Algebraic-Groups.tex
update

@@
 
+\section{November 8th, 2019}
+We started by talking elections. Go Andrew Lewis. You can do it.
+
+Recall that last time we had a result that said that we had an analog of the stabilizer $G_W\subseteq G$ 
+representing the functor 
+\[R\mapsto \{g\in G(R)|gW_R=W_R\}.\]
+The idea was to take a basis for $W$ and complete it to one of $V$ and then to find a group that fixes $W$. We found 
+that $G_W=V(\{a_{kj}\})$.
+\begin{thm}[Chevalley]
+	If $H\subseteq G$ is a subgroup of an algebraic group over $k$, then there exists a $G$ representation $V$ and a line $L\subseteq V$ such that $H=G_L$.
+\end{thm}
+\begin{prf}
+	Let $\pi:\Gamma(G)\twoheadrightarrow\Gamma(H)$ be the map induced by $H\hookrightarrow G$. Let $q_1,\dots,q_n$ be
+	generators of $\ker\pi=I\subseteq\Gamma(G)$, the regular representation. Take a finite dimensional representation $V$ containing $I$ in $\Gamma(G)$
+	and pick a basis $e_1,\dots,e_s$ of $W=V\cap I$. Extend this basis to one of $V$ with the additional vectors denoted $f_1,\dots,f_t$.
+
+	Now the image of the coaction gives us 
+	\[e_i\mapsto\sum_k a_{ik}\otimes e_k+\sum_k b_{jk}\otimes f_k\]
+	and let $I'=(b_{jk})$. We claim that $I=I'$. If this is true, then $H=G_W$ and $W\subseteq V$, so we set
+	\[L=\bigwedge^{\dim W}W\subseteq \bigwedge^{\dim W}V\]
+	and now $H= G_L$.
+
+	To see the claim, consider that since $e_i\in I$, we get $\sum_ka_{ik}\otimes e_i\in \Gamma(G)\otimes I$.
+	But then the comultiplication on $\Gamma(G)$ gives us that 
+	\[m^\ast(I)\subseteq I\otimes \Gamma(G)+\Gamma(G)\otimes I\subseteq \Gamma(G)\otimes\Gamma(G)\]
+	so we get that 
+	\[\sum b_{ik}\otimes f_k\in I\otimes \Gamma(G)+\Gamma(G)\otimes I\]
+	but since $f_{ij}\notin I$, this forces $b_{ik}\in I$. Thus $I'\subseteq I$.
+
+	For the other containment I got behind.
+\end{prf}
+\subsection{Quotients}
+Let $H\subseteq G$ be normal. Now the goal is to construct $G/H$ as an algebraic group. As a preview, we are going to use the last theorem 
+to get a representation $V$ of $G$ such that $H=G_L$ such that $G\to \GL(V)$ descends to a representation $G\to \GL(V^H)$. Then $G/H$ will be the image of this map in $\GL(V^H)$.
+
+\begin{defn}
+	A subgroup $H\subseteq G$ of an algebraic group over $k$ is \textbf{normal} if for all $k$-schemes $S$, $H(S)\le G(S)$ is anormal subgroup.
+\end{defn}
+\begin{rmk}
+	As an exercise one can show that if $G(k)\subseteq G$, then $H\subseteq G$ is normal if and only if for all $g\in G(k)$, $gHg^{-1}\subseteq H$.
+\end{rmk}
+\begin{lem}
+	If $\pi:H\to G$ is a homomorphism of algebraic groups, then $\ker\pi\subseteq H$ is normal.
+\end{lem}
+To see the above, consider the kernel is the pullback:
+\begin{center}
+	\begin{tikzcd}
+		S\ar[dr,dashed]\ar[drr,bend left]\ar[ddr,bend right] & &\\
+		& \ker\pi\ar[r]\ar[d] & \Spec k\ar[d,"e_G"]\\
+		& H\ar[r,"\pi"] & G
+	\end{tikzcd}
+\end{center}
+And then the kernel $\ker \pi(S)=K(S)$ is normal in $H$. Think about this one some more.
+\begin{defn}
+	A homomorphism of algebraic groups $G\to Q$ is a \textbf{quotient map} if $G\to Q$ is faithfully flat.
+\end{defn}
+\begin{prop}
+	A quotient map of linear algebraic groups $\pi:G\to Q$ satisfies the universal property: for all $f:G\to H$,
+	\begin{center}
+		\begin{tikzcd}
+			K=\ker\pi\ar[r]\ar[rd,"0"] & G\ar[r,"\pi"]\ar[d,"f"] & Q\ar[dl,dashed,"!"]\\
+			& H &
+		\end{tikzcd}
+	\end{center}
+\end{prop}
+
 \end{document}

### 2019-11-18 Algebraic Geometry/Algebraic-Geometry.tex
update

@@
 
+\section{November 15th, 2019}
+Thinking about quadradic forms (avoiding characteristic 2 for a moment): Any quadratif form $Q(x_0,\dots,x_n)$ is equivalent
+(via a linear change of coordinates) to $x_0^2+\cdots+x_m^2$ (when $k=\bar k$) for some $0\le m\le n$. Thus $Q$ is nondegenerate 
+if and only if $m=n$.
+
+Think about the rational map $\pi:\P n\dashrightarrow \P m$ sending $(x_0:\cdots:x_n)\mapsto (x_0:\cdots:x_m)$. Notice this is not 
+defined at $x_0=\cdots=x_m=0$, a linear subspace $L\subseteq \P n$ called the \textit{center of $\pi$}.
+For instance, the map $\P3\dashrightarrow\P2$ is not defined at $(0:0:0:1)$. In affine coordinates $\A3\subseteq \P3$, 
+where we set $x_3=1$. Here we think of setting $z=1$ (since we want to send $(x,y,z)\mapsto(x/z,y/z)=(x:y:z)$).
+
+Looking at where points go in the plane $\A2\subseteq \A3\subseteq \P3$, we see that it is somehow mapped via the line through the origin. 
+This is a ``pinhole camera'' projection of $\P3$ onto $\P2$.
+\begin{prob}
+	Given $\A3\to\A2$ via the map $(x,y,z)\mapsto (x,y)$, can this be recognized as the restiction of a projective morphism?
+\end{prob}
+
+Notice that given $\pi:\P3\dashrightarrow\P2$, the center is a conic given by $x_0^2+x_1^2+x_2^2=0$. This is a cone! Sweet!
+
+Check this out: If we have the affine map $\A3\to\A1$ with the quadratic form $(x/y)^2+1=0$, consider the pullback 
+under this map: the solutions to this in $\bbC$ are $x=\pm iy$, and for each of these points $\alpha$, the point $(x/y)-\alpha=0$ 
+pulls back to $x-\alpha y=0$, a plane! But we get two roots, so two planes in $\P3\supset \A3$. What is the intersection of these? The center!
+
+\subsection{Cremona}
+We have the Cremona map $\P2\dashrightarrow\P2$ sending $(x,y,z)\mapsto(1/x,1/y,1/z)$. This is ``obviously'' a birational involution. 
+At least involutivity is apparent.
+
+But we can scale by $xyz$ and get the map $(x,y,z)\mapsto(yz,xz,xy)$. Notice it is undefined at $(1,0,0),(0,1,0)$, and $(0,0,1)$.
+Now notice that the original map wasn't defined where \textit{any of the coordinates were zero} while the latter only requires that no more 
+than two are zero. Why is this kosher?!
+
+The idea here is that each of the lines $x=0$, etc are sent to POINTS $(1,0,0)$, which are then not in the domain of the cremona map.
+Thus the idea of involution is no longer meaningful!
+
+\begin{prob}
+	Consider the form $Q=xy-zw$ and its zero set in $\P3$. Then look at its image under the map $(x,y,z,w)\mapsto(y,z,w)$. SHow $\pi$ is birational and write down an 
+	inverse then compute $\pi\circ \pi^{-1}$ using cremona.
+\end{prob}
+
+\section{November 18th, 2019}
+Apparently Hitler loved cake.
+
+Today we are going to do some more examples. However, first we will consider 
+\subsection{Last week's problems}
+Consider the standard projection $\A3\to\A2$. We wanted to think of it as a linear map on projective spaces. Recall 
+\begin{defn}
+	A \textbf{rational map} $f:X\dashrightarrow Y$ is an equivalence of morphisms $f_U:U\to Y$ where $\varnothing\ne U\subseteq X$ is open. The equivalence 
+	relation is that $f_U=f_V$ if $f_U|_{U\cap V}=f_V|_{U\cap V}$.
+\end{defn}
+What does it mean for a map to be birational? One idea could be
+\begin{prob}
+	Make a category $\mathbf{Bir}$ of irreducbile varieties with mapes being rational maps (eq. classes). Try to think about inverses in this category.
+\end{prob}
+\begin{defn}
+	A rational map $f:X\dashrightarrow Y$ is \textbf{birational} if there exists a $g:Y\dashrightarrow X$ such that 
+	\[fg=\id_Y\qquad gf =\id_X\]
+
+	$f$ is \textbf{dominant} if there exists an open $V\subseteq Y$ that is contained in $\Im \tilde f$ for some representative $\tilde f$ of $f$.
+\end{defn}
+\begin{prob}
+	Show that if $f:X\dashrightarrow Y$ is birational then there exist nonempty opens $U$ and $V$ in $X$ and $Y$, respectively such that $f$ induces an isomorphism $U\xrightarrow{\sim} V$ of varieties.
+\end{prob}
+\begin{defn}
+	An irreducible variety is \textbf{rational} if it is birational to $\P n$ for some $n$.
+\end{defn}
+\begin{rmk}
+	Roughly speaking, we think that ``most of $X$ can be uniquely described by $n$ algebraic parameters.'' Try to look at some examples! The points where this doesn't make sense is jsut the 
+\end{rmk}
+\begin{prob}
+	I am repeating some of the old problems to refresh their importance: try figuring out how a quadric projects from $\P2$ to $\P1$ by the usual projection map. There is actually an isomorphism!
+\end{prob}
+We've been sniffing around the idea that any non-degenerate quadric $Q\subseteq \P n$ is rational. What about cublics? WHich are rational? The cubic curves (defined by cubic equations in $\P2$)
+are genus one and NOT rational. The cubic surfaces (those in $\P3$) are apparently.
+
 \end{document}

### 2019-11-18 Algebraic Groups/Algebraic-Groups.tex
update

@@
 Notes by Nico Courts}
-\date{Autumn 2019/ Winter and Spring 2020}
+\date{Autumn 2019/ Winter 2020}
 \maketitle
@@
 \begin{prop}
-	And finite dimentional representation $V$ of $G$ embeds $V\subseteq \Gamma(G)^{\oplus n}$.
+	And finite dimensional representation $V$ of $G$ embeds $V\subseteq \Gamma(G)^{\oplus n}$.
 \end{prop}
@@
 
+\section{November 13, 2019}
+The big theorem here:
+\begin{thm}
+	Let $G$ be an algebraic group over $k$. Let $H\subseteq G$ be an algebraic subgroup (over $k$). Then the
+	quotient $G/H$ exists as a quasi-projective scheme over $k$.
+
+	Moreover if $H\subseteq G$ is normal, then $G/H$ is also an algebraic group.
+\end{thm}
+
+Today we are going to prove/see the case when $G$ is a linear algebraic group and smooth (reduced) and that $H\subseteq G$ is normal and $k=\bar k$.
+The closedness of $k$ and the smoothness can be dropped at the expense of a couple extra lectures, probably.
+
+\begin{defn}
+	A map $G\xrightarrow{\pi} Q$ of algebraic groups over $k$ is a \textbf{quotient} if $\pi$ is faithfully flat.
+\end{defn}
+
+Using descent, we showed 
+\begin{prop}
+	If $G\xrightarrow{\pi} Q$ is a quotient, then  if we have another map $G\xrightarrow{\phi} P$ and furthermore that $\ker\pi\subseteq \ker\phi$,
+	then we get a (unique) factorization $Q\to P$.
+\end{prop}
+
+We also have
+\begin{prop}
+	Any map of linear algebrai groups $G\xrightarrow{\phi} H$ factors as $G\to Q\hookrightarrow H$ where $G\to Q$ is the quotient map.
+\end{prop}
+\begin{prf}
+	In the case that $G$ is reduced, we get a surjection onto $\Im\phi^\ast$ of $\Gamma(H)$, and $Q=\Spec\Im(\phi^\ast)$.
+	Then it just takes showing that the surjection $\Gamma(H)\twoheadrightarrow \Im(\phi^\ast)$ induces a closed immersion of $Q$ in $H$.
+
+	We also need a lemma that I missed but it can be found both in Milne and Waterhouse. This is where the heavy lifting is done.
+\end{prf}
+\section{November 15th, 2019}
+Today we are going to be talking about properties of algebraic groups as well as properties of lements of $g\in G$. Today, we will be focusing on 
+$g\in \GL(V)$, so essentially talking linear algebra.
+
+\begin{defn}
+	Let $V$ be a finite dimensional vector space over any field $k$. Let $g\in \GL(V)(k)$. Then $g$ is
+	\begin{itemize}
+		\item \textbf{diagonalizable} if there exists a basis of eigenvectors.
+		\item \textbf{semisimple} if there exsits an extension $k'/k$ such that $g\otimes\id\in\GL(V\otimes_kk')$ is diagonalizable.
+		\item \textbf{unipotent} if $g-\id$ is nilpotent.
+		\item \textbf{triagonalizable} if there exists a basis so that $g$ is upper triangular.
+	\end{itemize}
+\end{defn}
+
+Let $P_g\in k[T]$ be the characteristic polynomial for $g$. We say the \textit{eigenvalues of $g$ are in $k$} if $P_g$ splits over $k$. Thus if $g$ is diagonalizable, 
+the eigenvalues are in $k$ and this is exactly equivalent to $g$ being triagonalizable. Furthermore $g$ is unipotent if and only if 
+all of its eigenvalues are 1 or 0 (in $k$).
+\begin{prop}
+	If the eigenvalues $\lambda_i$ are in $k$, then 
+	\[V\cong \oplus\ker(g-\lambda_i\id)^{a_i})\]]
+	which gives us our Jordan decomposition in terms of generalized eigenspaces.
+\end{prop}
+
+\subsection{Jordan Decomposition}
+\begin{thm}
+	Let $k$ be perfect and let $g\in\GL(V)$ with $V$ finite dimensional over $k$. Then there exist unique $g_s$ and $g_u$ in $\GL(V)$ such that 
+	\begin{itemize}
+		\item $g=g_sg_u=g_ug_s$
+		\item $g_s$ is semisimple and $g_u$ is unipotent.
+	\end{itemize}
+	Moreover, $g_s$ and $g_u$ are polynomials in $g$.
+\end{thm}
+\begin{prf}
+	Assume first that $P_g(T)$ splits over $k$. Thus we can decompose 
+	\[V\cong V_{\lambda_i}\]
+	where $V_{\lambda_i}=\ker((g-\lambda_i\id)^{a_i})$.
+	This gives us a Jordan decomposition of $g$ as a matrix with eigenvalues on the diagonal and is upper triangular. Take $g_s$ to be $\diag(\lambda_1,\cdots,\lambda_1,\dots,\lambda_s)$, 
+	the diagonal of the matrix in this form. This is invertible (since $g\in\GL(V)$) so we set $g_u=g_s^{-1}g$. It is clear from this that these matrices commute and that they are 
+	semisimple and unipotent, respectively.
+
+	To get that these are polynomials in $g$, you go through an argument I missed. The uniqueness comes from looking at 
+	\[g=g_sg_u=h_sh_u\]
+	and taking 
+	\[h_s^{-1}g_s=h_ug_u^{-1}\]
+	and since everything commutes (check this), you get that these two elements are both semisimple and unipotent. But then they are diagonalizable with all 
+	eigenvalues 1, so they are the identity and we are done.
+
+	For the more general case, choose a finite extension $k'/k$ such that $P_g$ splits. Since $k$ is perfect, this is seperable. So we can assume it is Galois and 
+	set $G=\Gal(k'/k)$. So $g\otimes \id\in\GL(V\otimes_k k')$ factors uniquely as $g\otimes \id=g_ug_s$ with $g_u,g_s\in\GL(V)(k')$.
+
+	But then using the action of $G$ on $\GL(V)(k')$, by uniqueness of this decomposition $\sigma g_s=g_s$ and $\sigma g_u=g_u$ for all $\sigma\in G$ so 
+	the elements are in fact in the field fixed by $G$ (i.e. $k$). Similarly $G$ acts on $Q(T)\in k'[T]$ but must fix the polynomials describing the factors as polynomials in $g$.
+\end{prf}
+
+As an example, consider a field $k$ of characteristic 2 and pick $\alpha\in k/k^2$ (obviously $k$ is not perfect). Then the matrix $g=(\begin{smallmatrix}
+	0& 1\\\alpha,0
+\end{smallmatrix})$ has polynomial $T^2-\alpha$, so with a single repeated eigenvalue $\sqrt a\notin k$.
+
+\subsection{A broader context}
+So then if $\iota:G\hookrightarrow\GL(V)$ is inclusion (faithful representation) we can decompose $\iota(g)=g_ug_s$ using the above results. Some properties of this phenomenon:
+\begin{prop}
+	If $L:V\to W$ is a linear transformation and $g\in\GL(V)$ and $h\in\GL(W)$ such that $L\circ g=h\circ L$,
+	then $L\circ g_\ast=h_\ast\circ L$ for $\ast=u,s$.
+\end{prop}
+\begin{prop}
+	\begin{align*}
+		(g\oplus h)_\ast&=g_\ast\oplus h_\ast\\
+		(g\otimes h)_\ast&=g_\ast\otimes h_\ast
+	\end{align*}
+	for $\ast=s,u$.
+\end{prop}
+\begin{prop}
+	If $\rho:\GL(V)\to \GL(W)$ is a representation, then $\rho(g)_\ast=\rho(g_\ast)$.
+\end{prop}
+
+\begin{defn}
+	For $g\in G(k)$, $g$ is \textbf{semisimple (resp. unipotent)} if for all representations $\rho:G\to \GL(V)$,
+	$\rho(g)$ is semisimple (resp. unipotent).
+\end{defn}
+
+\section{November 18th, 2019}
+\subsection{Generalizing the Jordan decomposition}
+The further generalization of the Jordan decomposition we saw last time is
+\begin{thm}
+	If $k$ is perfect and $G$ is a linear algebraic group over $k$, then for any $g\in G(k)$, there exist unique $g_s,g_u\in G(k)$ such that 
+	\begin{itemize}
+		\item $g=g_sg_u=g_ug_s$
+		\item $g_s$ is semisimple and $g_u$ is unipotent.
+		\item $g_s$ and $g_u$ are polynomials in $g$.
+	\end{itemize}
+\end{thm}
+Last time we proved this for $G=\GL(V)$. Then we saw a collection of propositions that gave us that semisimplicity and unipotence act nicely 
+with respect to direct sum, tensor, and representations. This allowed us to define what semisimple and unipotent objects are in $G(k)$. This gave us 
+a result which I think I botched last time:
+\begin{prop}
+	If $G$ is alinear algebraic group and $V$ is a faithful finite dimentional representation ($G\hookrightarrow \GL(V)$ is a closed embedding), then eery representation 
+	of $G$ is obtained from $V$ using sums, tensors, subrepresentations, and quotients.
+\end{prop}
+\begin{rmk}
+	An aside: what does it mean for a subspace $W\subset V$ to be fixed by $g\in G$ ($V\in \lmod G$)? If $V$ is finite-dimensional, then one can just take the image of $G$ in $\GL(V)$
+	and ask whether $gW\subseteq W$.
+
+	More generally since $g$ is a $k$-point, there is a maximal ideal $\frakm_g\lhd \Gamma(G)$ such that 
+	\[\Gamma(G)/\frakm_g\cong k\] 
+	with quotient map $q$. Then this gives us a map $\Gamma(G)\otimes V\xrightarrow{q\otimes\id} k\otimes V\cong V$ which, when composed with the coaction $V\to\Gamma(G)\otimes V$,
+	gives us a map $m_g:V\to V$. Then we can say that $W$ is fixed by $g$ if $m_g(W)\subseteq W$.
+\end{rmk}
+\begin{prf}[in general]
+	Since $G$ is linear algebraic, we can embed $G\hookrightarrow\GL_n$ for some $n$. This gives us a way to decompose $g=g_sg_u$, but the question is 
+	whether $g_s$ and $g_u$ actually lie in (the image of) $G$.
+Thus for every representation $\rho:G\to \GL(W)$, we get a pair of elements $\rho(g)_s$ and $\rho(g)_u$.
+Due to the properties we saw before,
+	these are nice.
+
+	The idea here is that one representation we have at our disposal is the regular representation.
+In general this is large, but it recovers our group $G$ in some way and this will be key to our discussion.
+	Let $V=\Gamma(\GL_n)$, the regular representation of $\GL_n$. Then we get maps 
+	\[G\hookrightarrow\GL_n\to\GL(V)\]
+	and we have an ideal $I\subseteq \Gamma(GL_n)$ where $I$ is the ideal defining $G$. Then the claim is that $g\in G(k)$ stabilizes $I$. This is in Waterhouse chapter 9.
+\end{prf}
+
+\subsection{Group properties}
+This enables us to make more classifications of groups:
+\begin{prop}
+	We say that $G$ is \textbf{diagonalizable} if there is a faithful representation $G\hookrightarrow\GL_n$ that factors through the 
+	diagonal matrices in $\GL_n$.
+\end{prop}
+\begin{rmk}
+	This is equivalent to the earlier definition of there being a closed embedding of $G\hookrightarrow\Gm^n$.
+\end{rmk}
+\begin{defn}
+	We say that $G$ is of \textbf{multiplicative type} if $G\times_k\bar k$ is diagonalizable.
+\end{defn}
+
+\begin{prop}
+	If $G$ is a commutative linear algebraic group over $K$ and all elements $g\in G(k')$ for $k\to k'$ a field extension are semisimple, then $G$ is of multiplicative type.
+\end{prop}
+\begin{prf}
+	We can assume that $k=\bar k$ and we know that all $g\in G(k)$ are diagonalizable. Then we use the fact that commuting diagonalizable 
+	matrices are simultaneously diagonalizable. The result follows.
+\end{prf}
+\begin{rmk}
+	The proof of the linear algebra fact above comes from taking your commuting diagonalizable elements $g$ and $h$ and showing that each 
+	$\lambda$ eigenspace of $g$ is fixed by $h$ and vice versa, giving us the same number and dimension of eigenspaces.
+\end{rmk}
+
 \end{document}

### 2019-11-25 Algebraic Geometry/Algebraic-Geometry.tex
update

@@
 \end{prob}
-We've been sniffing around the idea that any non-degenerate quadric $Q\subseteq \P n$ is rational. What about cublics? WHich are rational? The cubic curves (defined by cubic equations in $\P2$)
+We've been sniffing around the idea that any non-degenerate quadric $Q\subseteq \P n$ is rational. What about cublics? Which are rational? The cubic curves (defined by cubic equations in $\P2$)
 are genus one and NOT rational. The cubic surfaces (those in $\P3$) are apparently.
 
+\section{November 25th, 2019}
+Last time we talked about quadrics and cubics and stuff! We asked whether smooth quadrics are all rational (birational to $\P n$). We also asked when cubics are rational.
+
+We saw that when $Q\subseteq \P2$ that $Q\cong \P1.$ In general we can see that these are rational! This involves taking any point on the quadric 
+and then project onto the other one. But the important thing to notice here is that this required us to pick a point and we break the symmetry of having 
+two points in each fiber.
+
+What about cubics? Here the center of a projection $\P3\dashrightarrow\P1$ is a \textit{line}. Let $S=Z(\text{cubic form})$. Then the fiber over $p\in\P1$ is 
+a plane: $\P2\hookrightarrow\P3$ via $(x_0:x_1:x_2)\mapsto(L_0:L_1:L_2:L_3)$ where the $L_i$ are linear forms in the $x_i$. Then if $F$ is our cubic form,
+we can consider the slice by this plane: $F(L_0,\dots,L_3)=G(x_0,\dots,x_3)$. For next time: what if the center $l\subseteq S$ (the center of projection is in $S$). One question:
+does the intersection $G$ of a plane containing such a line factor as a product of a quadric and a linear factor?
+
 \end{document}

### 2019-11-25 Algebraic Groups/Algebraic-Groups.tex
update

@@
 
+\section{November 22, 2019}
+From last time (which I missed): we had
+\begin{prop}
+	Let $G$ be a linear algebraic group over $k$. Then the following are equivalent:
+	\begin{itemize}
+		\item $G$ is diagonalizable (i.e. $G=D(A)$ for some finitely generated abelian group $A$)
+		\item There exists a faithful representation $G\to \GL(V)\cong \GL_n$ which maps into the diagonal matrices.
+		\item $\Gamma(G)$ is spanned by grouplike elements.
+		\item (Discussed below) All $G$ representations are direct sums of one-dimensional ones.
+	\end{itemize}
+\end{prop}
+Earlier this quarter we saw:
+\begin{prop}
+	If $G=D(A)$, then any $G$ representation $V$ can be decomposed as 
+	\[V=\oplus_{a\in A V_a}\]
+	where 
+	\[V_a=\{v\in V|\sigma^\ast(v)=t^a\otimes v\}\]
+\end{prop}
+\begin{rmk}
+	Here we used the fact (shown yesterday?) that the characters of an abelian group are in bijection with elements.
+\end{rmk}
+Note that if $k_a$ is the one-dimensional representation $k_a\to\Gamma(G)\otimes k_a$ via $1\mapsto t^1\otimes 1$. One can 
+quickly show that $V_a\cong k_a$, so we get that every representation decomposes as a direct sum of one-dimensional subgroups. In particular, 
+$G=D(A)$ are \textit{linearly reductive}.
+
+\subsection{Other properties of diagonalizable groups}
+First of all, these groups are commutative. That is, $Hom_\AlgGrp(G,\Gm)\cong A$.
+
+Next, we claim that $\Hom_\AlgGrp(G,\Ga)=0$. To see this, assume that $f:G\to \Ga$ is such a map and consider the 
+induced map $f^\ast:k[x]\to k[A]$. Then following through the commutative diagram for comultiplication, we get
+\[f^\ast(x)\otimes f^\ast(x)=f^\ast(x)\otimes 1+1\otimes f^\ast(x)\]
+by the diagonalizability of $G$ and the fact that $\Ga$ is ``infinitesimal'' (I forget the actual term from Hopf algebras.)
+
+\begin{defn}
+	A linear algebraic group $G$ over $k$ is \textbf{of multiplicative type} if there exists a finite extension $k'/k$ such that $G_{k}=G\times_{\Spec k}\Spec k'$ is diagonaliable.
+\end{defn}
+\begin{rmk}
+	This is equivalent to $G_{\bar k}$ being diagonalizable. One direction is obvious. For the other direction, Supposed we ahve a faithful 
+	representation $\bar V$ of $\bar G=G_{\bar k}$ by using some descent ideas.
+\end{rmk}
+
+\begin{thm}
+	Let $G$ be an algebraic group over $k$. Then the following are equivalent:
+	\begin{itemize}
+		\item $G$ is of multiplicative type.
+		\item $G$ is commutative and $\Hom(G,\Ga)=0$.
+		\item $G$ is commutative and $\Gamma(G)$ is co\'etale.
+		\item $G_{k^s}$ is diagonalizable. That is, there exists $k'/k$ that is finite and seperable such that $G_{k'}$ is seperable.
+	\end{itemize}
+\end{thm}
+What does that mean?
+\begin{defn}
+	Let $(C,\Delta,\varepsilon)$ be a coalgebra. Let $A=C^\vee$ be the $k$-linear dual. This gives us (using that $C^\vee\otimes C^\vee\hookrightarrow (C\otimes C)^\vee$)
+	an algebra. Then 
+	\begin{itemize}
+		\item If $C$ is finite dimensional, we say that $C$ is \textbf{co\'etale} if $A$ is commutative and if $A=C^\vee$ is \'etale. That is 
+		$A=\prod_{i=1}^n k_i'$ where $k_i'/k$ are all finite and seperable extensions.
+		\item If $C$ is infinite dimensional, then $C$ is \textbf{co\'etale} if it is a union of finite dimensional co\'etales.
+	\end{itemize}
+\end{defn}
+\begin{ex}
+	$G=\Gm$, $\Gamma(G)=k[t]_t$. Then let $C=k\langle t^d\rangle\subseteq \Gamma(G)$. Then $A=C^\vee$ is just $k$, so it is co\'etale.
+	Furthermore $k\langle t^{-d},\dots,t,\dots,t^d\rangle$ has that $C^\vee\cong \prod k$, so is as well.
+
+	Finally this gives us that $\Gamma(\Gm)$ is co\'etale since it is a union of these things.
+\end{ex}
+As an aside, if $V\to \Gamma(G)\otimes V$ is a representation, then $\Im(V\otimes V^\vee\to \Gamma(G))$ is a coalgebra.
+
+\section{November 25th, 2019}
+Today we finish our discussion of group schemes of multiplicative type. There was one implication we skipped in Friday's discussion:
+that if $G$ is commutative then $\Hom(G,\Ga)=0$ implies that $\Gamma(G)$ is co\'etale.
+\subsection{Continuing from last time}
+\begin{prf}[cont.]
+	Assume that $k=\bar k$ and let $C\subseteq\Gamma(G)$ be a subcoalgebra of finite dimension. Let $A=C^\vee$ and $\langle\cdot,\cdot\rangle$ be 
+	the natural pairing $C\times A\to k$. If $A$ is not \'etale, then since $A$ is finite dimensional,
+	\[A=A_1\times A_n\]
+	where the $A_i$ are local Artinian $k$-algebras.
+
+	So there exists a projection map $\pi:A\to k[\varepsilon]/\varepsilon^2\cong k\oplus k\varepsilon$ that sends 
+	\[x\mapsto \langle x,c\rangle+\varepsilon\langle x,d\rangle\]
+	for some $c,d\in C$.
+	Then a subclaim is to prove that
+	\begin{itemize}
+		\item $\Delta(c)=c\otimes c$
+		\item $\Delta(d)=c\otimes d+d\otimes c$
+		\item $\varepsilon(c)=1$
+	\end{itemize}
+	Then by also showing that $c$ is invertible, one can compute 
+	\[\Delta(dc^{-1})=1\otimes c^{-1}d+c^{-1}d\otimes 1\]
+	which then gives us a (nontrivial!) map $G\to \Ga$, which contradicts the assumption.
+\end{prf}
+\begin{ex}[Non co\'etale example]
+	Consider $\Ga=\Spec k[x]$ with $\delta(x)=x\otimes 1+1\otimes x$. Then consider th subcoalgebra spanned by $\langle 1,\dots,x^{n-1}\rangle$
+	and the dual algebra $A=C^\vee$ spanned by $e_i$. One can compute that 
+	\[e_i\cdot e_j=\binom{i+j}{i}e_{i+j}\]
+	if $i+j\le n-1$ and it is zero otherwise. But then in particular $e_{n-1}^2=0$, so $A$ isn't even reduced!
+
+	One could ask whether this still holds for any choice of subcoalgebra, but we aren't doing that.
+\end{ex}
+\begin{cor}
+	Let $G$ be a linear algebraic group over $k$. Assume that $G$ is commutative (and maybe smooth). Then $G$ is of multiplicative type 
+	if and only if $G(\bar k)$ consists purely of semisimple elements.
+\end{cor}
+\begin{cor}
+	Let 
+	\[1\to G'\to G\to G''\to 1\]
+	be a short exact sequence. If $G$ is commutative, then $G$ is of multiplciative type if and only if both $G'$ and $G''$ are.
+\end{cor}
+\begin{rmk}
+	That subgroups and quotients of multiplicative type group schemes are multiplicative type is not hard to see. To see this, apply $\Hom(-,\Ga)$ to the sequence and notice that quotients and subgroups 
+	of commutative groups are commutative.
+
+	The nontrivial direction is that the collection of multiplicative type groups is closed under extension. 
+\end{rmk}
+
 \end{document}

### 2019-12-02 Algebraic Geometry/Algebraic-Geometry.tex
update

@@
 
+\section{December 2nd, 2019}
+Today (and through the rest of the week) we are going to talk about sheaves! Woot.
+
+We're going to have a long time in between our last class this week and our first class in January. Here's a tidbit to tide you over:
+\begin{prob}
+	Homework for the break: Do all the problems in section 1 of chapter II of Hartshorne.
+\end{prob}
+\subsection{Sheaves}
+There is a mantra! \textit{Sheaves are a geometric manifestation of the flow of information from local to global.}
+\begin{ex}
+	Let $X\in\SpcFun$. Then $\O$ is the sheaf of regular functions on $X$.
+\end{ex}
+\begin{rmk}
+	Recall that we had that $\O(U)$ was defined for every open $U$ and that regulartity was preserved by 
+	restrictions and finally that regularity is local (it glues). These give us the ideas we will use to define a sheaf.
+\end{rmk}
+I love this little gluing diagram: given an open covering $U_i$ of $U$,
+\begin{center}
+	\begin{tikzcd}
+		\O(U)\ar[r] & \prod_i \O(U_i)\ar[r,shift left]\ar[r,shift right] & \prod_{i,j}\O(U_i\cap U_j)
+	\end{tikzcd}
+\end{center}
+where the top map sends
+\[(\varphi_i)_i\mapsto (\varphi_i|_{U_i\cap U_j})_{(i,j)}\]
+and the bottom does 
+\[(\varphi_i)_i\mapsto (\varphi_j|_{U_i\cap U_j})_{(i,j)}.\]
+That is, $\O(U)$ is the equalizer of these maps.
+
+We took some time to talk about limits as (right) adjoints to the constant diagram functor. Hadn't seen that before.
+We concluded with a bunch of cateory theory, which was really fun. Ind and Pro categories.
+\subsection{More examples}
+Let either
+\begin{itemize}
+	\item $X$ and $Y$ be spaces
+	\item $h_Y:U\mapsto h_Y(U)$ is a sheaf on $X$ for all open $U$ in $X$.
+\end{itemize}
+
 \end{document}

### 2019-12-02 Algebraic Groups/Algebraic-Groups.tex
update

@@
 \begin{ex}[Non co\'etale example]
-	Consider $\Ga=\Spec k[x]$ with $\delta(x)=x\otimes 1+1\otimes x$. Then consider th subcoalgebra spanned by $\langle 1,\dots,x^{n-1}\rangle$
+	Consider $\Ga=\Spec k[x]$ with $\Delta(x)=x\otimes 1+1\otimes x$. Then consider th subcoalgebra spanned by $\langle 1,\dots,x^{n-1}\rangle$
 	and the dual algebra $A=C^\vee$ spanned by $e_i$. One can compute that 
@@
 
+\section{December 2nd, 2019}
+We are going to continue our discussion on Jordan decomposition by visiting the idea of 
+\subsection{Unipotent group schemes}
+Recall that if $G$ is a linear algebraic group over $k$ and $g\in G(k)$, then 
+\begin{itemize}
+	\item $g$ is semisimple if $\rho(g)$ is for all representations $\rho:G\to \GL(V)$ ($\rho(g)\otimes \bar k$ is diagonalizable)
+	\item $g$ is unipotent if $\rho(g)$ is for all representations $\rho$ ($\rho(g)-1$ is nilpotent).
+\end{itemize}
+
+We also said that, if $k$ is perfect, $g=g_sg_u=g_ug_s$ and this is a unique decomposition (the Jordan decomposition).
+
+When we are talking about an arbitrary group scheme, we no longer use ``semisimple'' but rather ``multiplicative type'': 
+If $G\subseteq \Gm^n\subseteq\GL_n$, then we say that $G$ is \textbf{diagonaliable}. If $G_{\bar k}$ is diagonalizable, then 
+we say $G$ is \textbf{of multiplicative type.}
+
+Recall that when $G$ is smooth, we saw that $G$ was of multiplicative type if and only if $G$ is commutative and $G(\bar k)$ consists entirely of semisimple objects.
+
+Some questions we want to answer: Is there a characterization of groups containing unipotent elements? Perhaps more interestingly:
+Do we have structure results for factoring algebraic groups analogous to the Jordan decomposition?
+
+\begin{thm}[Kolchin fixed point theorem]
+	Let $V$ be a finite dimensional vector space over $k$ and let $G\subseteq \GL(V)$ be an abstract subgroup containing unipotent elements.
+	Then there exists a nonzero $v\in V$ fixed by all $g\in G$.
+\end{thm}
+\begin{prf}
+	Begin by setting 
+	\[V^G=\{v\in V|\forall g\in G, gv=v\}.\]
+	It is easy to see that if $k'/k$ is a field extension, then 
+	\[(V\otimes_k k')^G=V^G\otimes_k k'\]
+	thus we can assume that $k=\bar k$ (since we want to show that $V^G$ is nonempty).
+
+	We can also assume that $V$ is a simple $G$-module, since it suffices to find a fixed point in a submodule.
+	This reduces the proof to showing that $V^G=V$.
+
+	There are two different proofs of this: first, we want to show that for all $g\in G$, that $g-I_V=0$. Suppose there were 
+	an element $g'$ not equal to the identity. We know that for all $g\in G$, (since $g$ is unipotent):
+	\[\tr (g)=\dim V\quad\Rightarrow\quad \tr(gg')=\dim V\quad\Rightarrow\quad \tr(g(g'-I_V))=0\]
+
+	Let $E=\{f:V\to V|\tr(gf)=0,\forall g\in G\}$, which contains $g'-I_V\ne 0$. $E$ is $G$-invariant under composition.
+	Then let $g'-I_V\in X\subseteq E$ where $X$ is a simple $G$-module containing this element. THen for any $v\in V$, define 
+	\[\varphi_v:X\to V\quad\text{via}\quad f\mapsto f(v)\]
+	because $g'\ne I_V$, there exists a $v\in V$ such that $g'v\ne v$. Thus for this specific $v$, $\varphi_v$ is an isomorphism!
+
+	In particular, this means there exists an $f\in X$ such that $f(v)=v$. THen take 
+	\[A=\{L:V\to V| L\text{ linear, commuting with $G$}\}\]
+	which is a division ring over $k=\bar k$, which forces $A=k$.
+
+	But then for all $w\in V$, we get the composition 
+	\begin{center}
+		\begin{tikzcd}
+			V\ar[r,"\phi_v^{-1}"]\ar[rr,bend right] & X\ar[r,"\phi_w"] & V
+		\end{tikzcd}
+	\end{center}
+	and then we are supposed to show that our choice of $f$ and $v$ above send all other $w\in V$ to the span of $v$,
+	so $f$ has trace one. Something is fishy in this proof, but it's in Waterhouse.
+\end{prf}
+\begin{prf}[Less tricky, more fancy]
+	By Wedderburn, if $A$ is a ring and $M$ is a faithful left $A$-module, and if $B=\End_A(M)$, then if $M$ is simple and finitely generated over $B$, 
+	then 
+	\[\End_B(M)=A.\]
+	The more traditional setting here is when $A=k$, $M=k^n$, and $B=M_n(k)$. Then $\End_{M_n(k)}(k^n)=k.$
+
+	Now let $A\subseteq \End_k(V)$ generated by $G$. Let $M=V$ (which is assumed to be simple as above). Then by letting $B=\End_A(V)=k$, so by the above 
+	result, $\End_k(V)=A$ so $A$ is simple! Then let $I=\langle g-I_V:g\in G\rangle$, a two-sided ideal of $A$, so is either zero (in which case we are done) or is all of $A$.
+
+\end{prf}
+
 \end{document}

### 2019-12-04 Algebraic Geometry/Algebraic-Geometry.tex
update

@@
 \subsection{More examples}
-Let either
 \begin{itemize}
-	\item $X$ and $Y$ be spaces
+	\item Let $X$ and $Y$ be spaces
 	\item $h_Y:U\mapsto h_Y(U)$ is a sheaf on $X$ for all open $U$ in $X$.
+	\item If $X$ is a manifold and $V\to X$ is a vector bundle over $X$, then you get a variety of sheaves (discrete, cts, flat, holomorphic sections).
 \end{itemize}
 
+\section{December 4th, 2019}
+Let $\calC$ be a category (think of $\Set$).
+\subsection{More Sheafiness}
+\begin{defn}
+	A \textbf{presheaf} with values in $\calC$ on $X$ is an $\calF$ such that 
+	\begin{itemize}
+		\item $\forall U\subseteq X$ (open) we have $\calF(U)\in\calC$
+		\item $\forall V\subseteq U\subseteq X$, we get a restriction map $\rho_{VU}:F(U)\to F(V)$ which composes with other restrictions.
+		\item $\rho_{UU}=\id_{F(U)}$
+	\end{itemize}
+\end{defn}
+\begin{rmk}
+	Of course this is all equivalent to saying a presheaf is a functor 
+	\[\calF:\mathbf{Open}(X)^{op}\to \calC.\]
+\end{rmk}
+
+Now let's say that $\calC$ has products (including the empty product/final object). Then definitely the sheaf condition makes sense. 
+We get the coordinate kind of definition if we imaging $\calC$ is $\Set$ (or at least concrete).
+\begin{defn}
+	A \textbf{sheaf} is a presheaf that satisfies the sheaf condition for all open coverings of all open sets.
+\end{defn}
+Some examples:
+\begin{itemize}
+	\item $\O=h_{\A1}$
+	\item $h_Y$ is a sheaf on $X$ as we saw above.
+\end{itemize}
+\begin{rmk}
+	There are two categories:
+	\[\PreSh=\Func(\Open(X)\op,\calC)\]
+	and a full subcaltegory $\Sh$ of sheaves.
+\end{rmk}
+\subsection{Stalks}
+\begin{defn}
+	Given a $p\in X$W and $\calC\in \PreSh(X)$, the \textbf{stalk of $\calF$ at $p$} is
+	\[\calF_p=\colim_{p\in U}\calF(U)\]
+\end{defn}
+\begin{rmk}
+	If $\calC$ is $\Set$, then we can realize this as $\sqcup_{p\in U}\calF(U)$ modulo the relation that two maps are the same
+	if they agree on a small neighborhood of $p$.
+\end{rmk}
+As a matter of notation, we say that if $p\in V$ and $f\in\calF(V)$, then there is a value $f_p\in \calF_p$,
+called the \textbf{germ of $f$ at $p$.}
+
+\begin{thm}[Basic]
+	A map of sheaves $\calF\to \calG$ is an isomorphism (resp. mono/epi) if and only if for all $p\in X$ the induced map $\calF_p\to\calG_p$ is an iso (resp. mono, epi).
+\end{thm}
+\begin{prob}
+	For next time: Classify all (pre)sheaves on the space with two points $x$ and $y$ with topology $\{\varnothing,\{x\},\{x,y\}\}$. Also think about monos, epis, and isos of (pre)sheaves).
+\end{prob}
+
 \end{document}

### 2019-12-04 Algebraic Groups/Algebraic-Groups.tex
update

@@
 	result, $\End_k(V)=A$ so $A$ is simple! Then let $I=\langle g-I_V:g\in G\rangle$, a two-sided ideal of $A$, so is either zero (in which case we are done) or is all of $A$.
-
-\end{prf}
+\end{prf}
+
+\section{December 4th, 2019}
+Recall the Kolchin FPT from last time. 
+\subsection{More unipotent stuff}
+\begin{cor}
+	If $V$ is a finite dimensional vector space over $k$ and $G\to\GL(V)$ is an abstract group homomorphism whose image consists of unipotent elements, then 
+	there exists a basis of $V$ such that $G\subseteq U_n\subseteq\GL_n$ where $U_n$ is the group of upper-triangular matrices with ones on the diagonal.
+\end{cor}
+\begin{cor}
+	If $G$ is asmooth linear algebraic group over $k=\bar k$ such that $G(k)$ consists of unipotents, then $G\subseteq U_n$ as a closed subgroup.
+\end{cor}
+Examples include $\Ga$ (included as $(\begin{smallmatrix}
+	1&a\\0&1
+\end{smallmatrix})$) which is unipotent. In characteristic $p$, $\bbZ/p$ is unipotent. There is also some weird $\alpha_p$ stuff.
+\begin{defn}
+	A linear algebraic group $G$ over $k$ is unipotent if for all nonzero $G$-representations, $V^G\ne 0$.
+\end{defn}
+\begin{rmk}
+	This is equivalent if we require $V$ to be finite dimensional over $k$.
+\end{rmk}
+\begin{prop}
+	$G$ is unipotent if and only if for all finite dimensional representations $\rho:G\to \GL(V)$, there exists a basis 
+	such that $\rho(G)\subseteq U_n$.
+\end{prop}
+\begin{prf}
+	Choose a filtration of $V$:
+	\[0=V_0\subset V_1\subset\cdots\subset V_n=V\]
+	with all quotients $W_i\eqdef V_i/V_{i-1}$ simple. But then since $W_i^G\ne 0$ by the definition of unipotence, $W_i^G=W_i$.
+	So then $W_i$ is the trivial representation and apparently that finishes the forward direction.
+
+	The reverse direction is easy (pick a basis and $v_1$ is fixed by $\rho(G)$).
+\end{prf}
+\begin{thm}
+	Let $G$ be a linear algebraic group over $k$. Then $G$ is unipotent if and only if $G$ is closed in $U_n$ for some $n$.
+\end{thm}
+\begin{lem}
+	If $G$ is unipotent, so is any subgroup, quotient, or extension.
+\end{lem}
+\subsection{Induced representation}
+Let $H\to G$ be a linear algebraic group homomorphism. Then we have a functor 
+\[\operatorname{Ind}_H^G:\Rep(H)\to\Rep(G)\]
+that is defined via 
+\[\operatorname{Ind}_H^G(V)=\Mor_{\Sch/k}^H(G,\bbA(V))=\{G\to \bbA(V)|\text{$H$ invariant}\}=(\Gamma(G)\otimes_k V)^H\]
+Notice that the inclusion functor $f:\mathbf{B}H\to\mathbf{B}G$ gives us two maps, $f_\ast$ (induction) and $f^\ast$ (forgetful).
 

### 2020-01-17 Algebraic Geometry/Algebraic-Geometry.tex
Update

@@
 
+\part{Quarter 1: Spaces with functions}
 \section{September 25, 2019}
@@
 
+\part{Quarter 2: Sheaves and Schemes}
+I missed the first five lectures of this quarter because I was travelling for a conference. Sorry about that!
+
+\section{January 17th, 2020}
+\subsection{SCHEMES}
+Let's just dive right in. Last time they were talking about things:
+\begin{itemize}
+	\item Is the category of spaces with functions a full subcategory of locally ringed spaces?
+	\item Thinking about $\Hom(\Spec A,\O_A),(\Spec B,\O_B))$
+\end{itemize}
+In the later case, we noticed that if we had a pair of maps $(f,f^\sharp)$ between $\Spec$s and the structure sheafs. Then $\Gamma(f^\sharp):B\to A$
+gives us a local condition: if $\varphi=\Gamma(f^\sharp)$, $f$ must be the map induced by $\varphi$.
+
+\subsection{Talking about affine schemes}
+Let $A$ be a (commutative unital) ring. Define a locally ringed space $\Spec A$ where 
+\begin{itemize}
+	\item The underlying space $|\Spec A|$ is the set of prime ideals of the ring.
+	\begin{itemize}	
+		\item Notice before we used the maximals but getting all the primes gives us something extra...
+		\item The topology is the Zariski topology.
+		\item This looks like a closed point for every prime in $\bbZ$ as well as the dense generic point $(0)$. This is the French model for the Italian idea of a ``general point.''
+		\item A basis for this topology is $D(f)\subseteq |\Spec A|$ defined as the nonvanishing locus:
+		\[D(f)=\{\frakp|f\notin\frakp\}\]
+		or in other words, $f(\frakp)\ne 0$.
+	\end{itemize}
+	\item We want 
+	\[\Gamma(|\Spec A|,\O)=\O(|\Spec A|)=A.\]
+	as well as $\O(D(f))=A_f$, the localization. We know that $D(f)$ defines a basis for the topology for $\Spec A$, so this implies
+	this gives us a way to define the sheaf locally.
+
+	So it suffices to show that $\O:\calB\op\to\Ring$ forms a sheaf (satisfies the sheaf condition). Here's the (``bestest'') proof:
+	\begin{prf}[Bestest]
+		Algebraic formulation: Begin by reducing to $U=|\Spec A|$. Then take an open cover $U_i$ of $U$. Look at descent theory! It's also in SGA.
+	\end{prf}
+\end{itemize}
+\begin{rmk}
+	We can think of the value of $f$ at a prime $\frakp$ as follows: $f\in A$ is an element of a ring and 
+	\[f(\frakp)\eqdef \bar f\in \kappa(A/\frakp)\]
+	the residue field.
+\end{rmk}
+\begin{rmk}
+	To see why it suffices to define the sheaf on a basis: let $\calB$ be a basis for a topological space $X$. Then let $F:\calB\op\to \Ab$ be a sheaf 
+	on $\calB$. Then there is some work to do, but the statement is that the category of sheaves on $X$ is equivalent to the category of sheaves on $\calB$.
+
+	The idea is as follows: to get essential surjectivity, you take a sheaf on $\calB$ and take the sheafification to get a sheaf on $X$. 
+\end{rmk}
+
 \end{document}

### 2020-01-17 Algebraic Groups/Algebraic-Groups.tex
Update

@@
 
+\part{Quarter 1: Structure Theory}
 \section{September 25, 2019}
@@
 
+\part{Quarter 2: Representation Theory}
+\section{January 17th, 2020}
+\subsection{Recap}
+Let $T\subseteq B\subseteq G$ be a maximal torus, Borel, and semisimple algebraic group. Then we talked about:
+\begin{itemize}
+	\item The infinitesimal theory by studying $\Lie G$
+	\item Borel subgroups, parabolic subgroups, and the flag variety $G/B$.
+	\item We also saw the categorization of split connected semisimple algebraic groups according to their root data:
+	This is either given by $(X,R,X^\vee,R^\vee)$ of a group $X$ and a root system $R$ and their duals or equivalently a 
+	weight lattice $\Lambda$ and root latice $\Lambda_r=\bbZ R$. Then we can define 
+	\[Z/\bbZ R\cong \Lambda/\Lambda_r\]
+	to be the \textbf{fundamental group.}
+	\item Representation theory: $\Rep G$ is what is called a \textbf{highest weight category.} Let $\Lambda=X$. Then 
+	we have the \textbf{dominant roots}
+	\[\Lambda_+=\{\lambda\in\Lambda|\langle\lambda,\alpha^\vee\rangle\ge 0,\alpha\in R^+\}\]
+	and for every dominant root $\lambda$, we have the \textbf{(co)standard} modules $\Delta(\lambda)$ and $\nabla(\lambda)$.
+\end{itemize}
+\begin{defn}
+	Let $\lambda\in\Lambda=\Hom(T,\Gm)$. Define the $T$-module $\calR_\lambda$ by 
+	\[t\cdot 1\eqdef \lambda(t)1.\]
+	Then $T(k)=(k^\times)^\Gamma$ gives us $\lambda(t)\in k^\times$.
+\end{defn}
+
+\begin{ex}
+	Let $G=\GL_n$ and $B$ the upper triangular matrices and $T$ the diagonal ones. Then the unipotent matrices $\calU$ of upper triangulars with ones on the diagonal 
+	include into $B$ and we have a section $T\to B$ giving us $B=T\rtimes U$.
+
+	Then we consider $k_\lambda$ as a representation of $B$ by having $\calU$ act trivially. Then we define 
+	\[\nabla(\lambda)=\Ind_B^G k_\lambda\]
+	where $\Ind_B^G$ is the right adjoint to the restriction functor $\Res^G_B$.
+\end{ex}
+\begin{rmk}
+	The induction functor always exists, but the \textit{left} adjoint to $\Res$ doesn't. When it does, it is called \textbf{coinduction.}
+\end{rmk}
+
+\subsection{The Associated Sheaf}
+Given $H\subseteq G$, there is a functor
+\[\calF:\Rep H\to \Sch G/H.\]
+this leads to a fact:
+\begin{prop}
+	\[\Ind_B^G k_\lambda=\nabla(\lambda)\cong H^0(G/B,\calF_{G/B}(k_\lambda))\]
+\end{prop}
+
+The standard module $\Delta(\lambda)$ is essentially the dual (after applying an element of the Weyl group of longest length).
+\begin{thm}
+	Let $\lambda\in\Lambda_+$. Then $\nabla(\lambda)$ has a simple socle, denoted $L(\lambda)$. Furthermore, $\{L(\lambda)|\lambda\in\Lambda_+\}$ is a complete set of $G$-modules.
+\end{thm}
+
+\subsection{Other things to focus on:}
+Look at Humphreys' \textit{Linear Algebraic Groups}. There is an appendix on root systems that is
+
+\subsection{Examples}
+Let $k=\bar k$ and $G=G(\bar k)$. For type $A_{n+1}$, we have $\SL_n$ as well as $\operatorname{PGL}_n$. In the former case, $\Lambda/\Lambda_r=\bbZ/n\bbZ$ and in the latter this quotient is trivial.
+
+For type $C_n$, we get $\operatorname{Sp}_{2n}$. Let $\Sigma$ be the matrix given by $-1$s and 1s on the antidiagonal and zeros elsewhere (negative in the ``third quadrant.'')
+Then the matrices in this are those preserving the associated form: that is 
+\[A^T\Sigma A=\Sigma\]
+
+For type $B_n$, we look at the $2n+1$ square matrix $\Sigma$ that is block diagonal with blocks $(1)$ and the previous matrix. Then $\operatorname{SO}(2n+1,k)$ are the matrices preserving $\Sigma$.
+The type $D_n$ matrices are the ones preserving the matrix given by (positive) ones on the antidiagonal.
+
+\subsection{More Generally...}
+Let $A$ be a separable associative unital algebra over $k$. (That is, $A_{\bar k}$ is a product of simple algebras). Fix an involution $\sigma:A\to A$ 
+($\sigma^2=\id$). There is always an associated bilinear form for these involutions, which can be symmetric or skew-symmetric. It either preserves the base field or else 
+it descends to an involution of this field.
+
+\begin{defn}
+	$\GL_{1,A}$ is a group scheme whose $R$-points ($R$ a ring) are given by 
+	\[\GL_{1,A}(R)=\left((A\otimes_k R)^\times\right)^k\]
+	which contains a subgroup scheme 
+	\[\operatorname{Iso}(A,\sigma)(R)\eqdef \{a\in(A\otimes_k R)^\times|a\cdot\sigma_R(a)=1\}\]
+
+	We also have $\Aut(A,\sigma)\subset \Aut(A)$ defined as 
+	\[\Aut(A,\sigma)(R)\eqdef\{\alpha\in\Aut_R(A_R)|\alpha\circ\sigma_R=\sigma_R\circ\alpha\}\]
+
+	And finally,
+	\[\operatorname{Sim}(A,\sigma)\eqdef \{a\in A_R^\times|a\cdot\sigma_R(a)\in k_R^\times\}\]
+\end{defn}
+
+Now if $A$ is a central simple algebra and $\sigma$ is a symplectic form, then 
+\begin{itemize}
+	\item $\operatorname{Sp}(A,\sigma)=\operatorname{Iso}(A,\sigma)$
+	\item $\operatorname{GSp}(A,\sigma)=\operatorname{Sim}(A,\sigma)$
+	\item $\operatorname{PGSp}(A,\sigma)=\Aut(A,\sigma)$
+\end{itemize}
+
+Want to know more? Look at \textit{The Book of Involutions.}
+
 \end{document}

### 2020-01-24 Algebraic Geometry/Algebraic-Geometry.tex
Update

@@
 
+\section{January 22, 2020}
+Recall that we have the set $\Spec A$ of primes of $A$ and we know the value of the structure sheaf $\O$ on special sets:
+\[\O(D(f))=A_f\]
+Then we interpolate between these sets to get the value of the sheaf at specific sets.
+
+\subsection{Connection with Locally Ringed Spaces}
+
+\begin{prop}
+	Say $(X,\O_X)$ is a LRS. Then the map
+	\[\Hom_{LRS}((X,\O_X),\Spec A)\to \Hom_{\Ring}(A,\O_X(X))\]
+	where we send $(f,f^\sharp)\mapsto f^\sharp(\Spec A)$, is a bijection.
+\end{prop}
+\begin{prf}
+	Why is this surjective? Given a map $\varphi:A\to \O_X(X)$ we want to concoct a map of schemes $A\to\Spec A$.
+	Let $p\in X$ and $f(p)=(\rho_p\circ\varphi)^{-1}(m_p)$ where $\rho_p$ is the localization at $p$ map.
+
+	Why is $f$ continuous under this definition? It is enough to chekc that $f^{-1}(D(a))\subseteq X$ is open,
+	which is the set of all $x\in X$ such that $[\varphi(x)]\notin m_x\in \O_{X,x}$ at x. In other words, we are finding the $x$ such that $\overline{\varphi(a)}\ne 0\in k(x)$ (the residue field.)
+
+	Then to get the map $f^\sharp:\O_{\Spec A}\to f_\ast\O_X$.
+\end{prf}
+
+\begin{lem}
+	Let $(X,\O_X)$ be a locally ringed space. Let $f\in\O_X(X)$. Then 
+	\[D(f)=\{x\in X|f\notin m_x\in\O_{X,x}\}\subset X\]
+	is open. (Note we are writing the germ $g_x$ as $g$).
+\end{lem}
+\begin{prf}
+	Since $g\notin m_x$, so $g$ is a unit in $\O_{X,x}$. Thus we have an $h\in\O_{X,x}$
+	such that $gh=1$. Thus there exists an $x\in U$ and $\tilde h\in \O_X(U)$ such that $g|_U\tilde h=1\in\O_X(U)$
+	so $U\subseteq D(g)$.
+\end{prf}
+
+\begin{rmk}
+	Look at ``sheaf on a base.'' This is the idea of interpolation after defining the sheaf on the distinguished 
+	opens $D(f)$. It is in Vakil and EGA chapter zero and many others.
+\end{rmk}
+
+\subsection{Schemes}
+So what is a scheme?
+\begin{defn}
+	A \textbf{scheme} is a locally ringed spaces such that there exists an open covering $U_i\subseteq X$ with $(U_i, \O_X(U_i))$ is an affine scheme for all $i$.
+\end{defn}
+Some examples:
+\begin{itemize}
+	\item $\Spec \bbZ$. $\bbZ$ is special because it is an initial object in $\Ring$. So for any $X$ we have a 
+	canonically-defined map $f:X\to \Spec \bbZ$, where we assign to every point in $X$ a prime number: this is the characteristic of the residue field $k(x)$!
+	\item $\Spec k$ where $k$ is a field. Here we have a single point and $\O(\ast)=k$. Then the endomorphisms of this scheme are 
+	the endomorphisms $\Hom(k,k)$, which is often interesting! There is a unique one when $k=\bbQ$, but not over $\bbC$!
+\end{itemize}
+
+\begin{defn}
+	For a fixed scheme $S$, the category of $S$ schemes $\Sch/S$ has objects $X\to S$ and 
+	$\Hom(X\to S,Y\to X)$ are given by diagrams.
+\end{defn}
+
+\section{January 24th, 2019}
+Today we are going to continue talking about $S$-schemes (or schemes over $S$). This is just the slice category $\Sch/S$!
+Notice to really make things work we have to make the following definition:
+\[(g,g^\sharp)\circ(f,f^\sharp)=(g\circ f,g_\ast(f^\sharp)\circ g^\sharp)\]
+\begin{rmk}
+	There is a section in Hartshorne where he talks about the difference between $k$-varieties and $k$-schemes. You should read it.
+\end{rmk}
+
+\subsection{Examples of Schemes}
+Of course we have affine schemes by construction. Other examples we have seen are $\Spec R$ for $R$ a DVR or product of fields. These both 
+have two points.
+
+\begin{rmk}
+	Look up the ``Punctual Hilbert Scheme'' that talks about schemes with one point with different-sized tangent spaces. For instance, $k[x,y]/(x,y)^2$.
+\end{rmk}
+
+Remember the idea of the line with two origins? Well you can take two copies of $\Spec k[[t]]$ (which contains both an open and a closed point) 
+and glue along the open point and get something analogous. This is no loner affine, but that is not obvious!
+
+\subsection{Proj}
+Let $A=\oplus A_i$ be a graded ring. Then an element of $A$ is homogeneous if it lives in some $A_i$. An ideal is homogeneous if $I$ 
+is generated by homogeneous elements.
+
+Some examples: $A=k[x_0,\dots,x_n]$ with $A_i$ the monomials of degree $i$. The irrelevant ideal $A^+$ is the 
+ideal of all stuff in positive degree.
+
+Then $\Proj$ makes a schemes from a graded ring. As a set,
+\[|\Proj A|=\{\frakp\subseteq A| \frakp\text{ is a homogeneous prim such that } A^+\not\subset\frakp\}\]
+and we assign to this set the Zariski topology where if $I\subseteq A$ is homogeneous, then
+\[V(I)=\{\frakp\in|\Proj A||\frakp\supset I\}\]
+and, in particular, notice that $V(A^+)=\varnothing.$
+
+Then we need to assign to this space a sheaf. Let $f\in A$ be a homogeneous element. Then 
+\[\O(D(f))=A_{(f)}=(A_f)_0\]
+which is the degree zero part of the usual ring $A_f$.
+
+For example, if $A=k[x_0,x_1]$, and if $f=x_0$, then $A_{(f)}=k\left[\frac{x_1}{x_0}\right]$.
+
+Our main result: 
+\begin{thm}
+	$(|\Proj A|,\O)$ is a locally ringed space. Furthermore given a homogeneous element $f$ with $\deg f>0$, the map 
+	\[A_{(f)}\to\O(D(f))\]
+	induces an isomorphism 
+	\[(D(f),\O_{D(f)})\xrightarrow{\sim}\Spec A_{(f)}.\]
+\end{thm}
+Another example: consider $\bbP_\bbZ^n=\Proj\bbZ[x_0,\dots,x_n]$.
+
 \end{document}

### 2020-01-24 Algebraic Groups/Algebraic-Groups.tex
Update

@@
 \begin{defn}
-	Let $\lambda\in\Lambda=\Hom(T,\Gm)$. Define the $T$-module $\calR_\lambda$ by 
+	Let $\lambda\in\Lambda=\Hom(T,\Gm)$. Define the $T$-module $k	_\lambda$ by 
 	\[t\cdot 1\eqdef \lambda(t)1.\]
@@
 
+\section{January 22nd, 2020}
+Now we begin the first part of this class: Infinitesimal theory, derivationts, differentails, and Lie algebras.
+
+\subsection{Tangent Spaces}
+If we have a variety $V$, we can define the dual space of $V$ at $x$ by considering 
+\[(m_{\O_x}/m^2_{\O_x})^\ast\]
+then we can talk about the cotangent space $m/m^2$.
+
+More formally, set $k$ to be a field (note in Jantzen he uses a commutative ring). Let $A$ be a finitely generated \textbf{commutative} $k$-algebra,
+so there exists a surjection $k[x_1,\dots,x_n]\to A$. 
+\begin{defn}
+	Let $M\in\lmod A$. Then a map $D:A\to M$ is a \textbf{derivation} if 
+	\[D(ab)=aD(b)=bD(a).\]
+	We say that it is a \textbf{$k$-derivation} or $k$-linear if $D(k)=0$. 
+\end{defn}
+
+As an example, let $B=k[x_1,\dots,x_n]$ and let $\Omega_B$ be the free $B$ module of rank $n$, generated by $dx_1,\dots,dx_n$.
+Then the map $d_B:B\to\Omega_B$ sending $x_i\mapsto dx_i$ and $1\mapsto 0$ is a derivation.
+\begin{lem}
+	Let $M$ be a $B$ module. Then 
+	\[\operatorname{Der}_k(B,M)\stackrel{\sim}{\to}\Hom_B(\Omega_B,M).\]
+\end{lem}
+\begin{prf}
+	Given a derivation $D:B\to M$, we can define a map $\varphi:\Omega_B\to M$ via 
+	\[\varphi(dx_i)=D(x_i)\]
+	and extending $B$-linearly. One has to prove that this defines an algebra morphism. 
+\end{prf}
+\begin{rmk}
+	Another way to phrase this is $\Omega_B$ is a universal object in $\lmod B$ for derivations.
+\end{rmk}
+\begin{thm}
+	Let $A$ be a finitely generated commutative $k$-algebra. Then there exists a (universal) $A$-module $\Omega_A$
+	and a (universal) derivation $d_A:A\to \Omega_A$ such that for any $A$-module $M$,
+	\[\operatorname{Der}_k(A,M)\stackrel{\sim}{\to}\Hom_A(\Omega_A,M).\]
+\end{thm}
+\begin{rmk}
+	The universality of the pair $(\Omega_A,d_A)$ is saying that it is unique up to (unique) isomorphism.
+\end{rmk}
+\begin{prf}
+	To see this, we use the last result as well as the fact that 
+	\[\Omega_A=\frac{\Omega_B}{I\Omega_B+B(d_B(I))}.\]
+
+	If $I=(f_1,\dots,f_m)$ then $\Omega_A$ is an $A$-module on generators, then $\Omega_A$ is an $A$-module on generators 
+	$dx_1,\dots,dx_n$ subject to 
+	\[\sum_{i=1}^n\frac{\partial f_j}{\partial x_i}dx_i=0\]
+	for all $1\le i\le m$.
+
+	The next part of the proof consists of showing that given a derivation $\tilde D:B\twoheadrightarrow A\xrightarrow{D}M$ and the associated map $\tilde\varphi:\Omega_B\to M$, 
+	the latter factors through $\Omega_A$. SHow it vanishes on both part of the sum above.
+\end{prf}
+\begin{rmk}
+	Note: we just proved that the functor $\Der_k:\lmod A\to\Vect_k$ is representable! The uniqueness of the representing object is then a consequence of Yoneda lemma!
+\end{rmk}
+
+\begin{ex}
+	Let $A=k[x,y]/(x^2+y^2-1)$. Then $\Omega_A$ is generated by $d_x,d_y$ with $2xd_x+2yd_y=0$. When $\ch k=2$,
+	we get that $\Omega_A$ is free of rank 2.
+
+	This defines a group scheme (think matrices in $\SL_2\subseteq\GL_2$) and this gives us a surjection of $k[\SL_2]\to A$.
+
+	When $\ch k\ne 2$, we get that $\Omega_A$ is a free $A$-module of rank 1, generated by $dt=xdy-ydx$. I have no idea what this was,
+	but she wrote $dy=-xdt.$
+\end{ex}
+\begin{rmk}
+	Notice that the reason we lost smoothness in characteristic 2 was that we lost reducedness: now $(x+y-1)^2=0$. There 
+	is a general theorem (see deep in Waterhouse) that an affine group scheme is smooth only if its Hopf algebra is reduced.
+\end{rmk}
+\subsection{Derivations on Hopf algebras.}
+Recall that if $A$ is a Hopf algebra, then $I=\ker\varepsilon$ is the \textbf{augmentation ideal.}
+\begin{thm}
+	Let $A$ be a finitely generated commutative Hopf algebra over $k$. Then $\Omega_A$ is a free $A$-module.
+
+	In fact, \[\Omega_A\cong A\otimes_k I/I^2\]. The map $d_A:A\to A\otimes I/I^2$ is given by 
+	\[a\mapsto a'\otimes \pi(a'')\]
+	where the primes indicate Sweedler notation and $\pi:A\to I/I^2$ is a linear map (not a homomorphism!) that takes a (non-unique) 
+	splitting $A=I\oplus k\cdot 1$ and then project on the first coordinate and take a quotient by $I^2$. It remains to prove that, once we pass to the 
+	quotient, the choice of direct sum complement is irrelevant.
+\end{thm}
+
+\begin{lem}
+	We say that $D:A\to M$ is an $\varepsilon$-derivation (for $\varepsilon:A\to k$) if 
+	\[D(ab)=\varepsilon(a)D(b)+\varepsilon(b)D(a).\]
+	The universal module of $\varepsilon$-derivations is 
+	\[\Omega_A\otimes_\varepsilon k=\Omega_A/I\Omega_A=I/I^2\]
+\end{lem}
+
+\section{January 24th, 2020}
+Let $A$ be a finitely generated commutative Hopf algebra over $k$. Often we can drop finite generation 
+because any Hopf algebra is a limit of finitely generated ones (because of finite global dimension), so 
+these arguments still hold if they can pass to the limit.
+
+Let $\varepsilon:A\to k$ be the counit and $I=\ker\varepsilon$ be the augmentation ideal.
+\begin{rmk}
+	We have this duality between algebra and geometry, so the counit can be seen as an element of $G_A(k)=\Hom(A,k)$,
+	specifically the unit element of the group. :)
+\end{rmk}
+
+\begin{lem}
+	There is a series of isomorphisms
+	\[\Der_\varepsilon(A,M)\simeq \Hom_k(I/I^2,M)\simeq\Hom_A(A\otimes_kI/I^2,M)\]
+\end{lem}
+This can be proved relatively simply. It is a good exercise.
+\begin{thm}
+	\begin{itemize}
+		\item $\Omega_A\simeq A\otimes_k I/I^2$
+		\item $d_A:A\to \Omega_A=A\otimes I/I^2$
+	\end{itemize}
+	where the map $d_A$ is the one sending $a\mapsto\sum a'\otimes \pi(a'')$
+\end{thm}
+\begin{prf}
+	Recall that $\pi:A\to I/I^2$ is the map that sends $A\cong k\cdot 1\oplus I$ to its projection to $I$ modulo $I^2$.
+
+	Let $M$ be an $A$ module. Give $A\oplus M$ an $A$-algebra structure via 
+	\[(a,m)\cdot(a',m')=(aa',am'+a'm).\]
+	Then let $G_A(-)=\Hom(A,0)$ and notice 
+	\[G_A(A\oplus M)=\Hom_\Alg(A,A\oplus M)\]
+
+	Then we claim that 
+	\[\Hom(A,A\oplus M)=\{(\varphi,D)|\varphi\in\Hom(A,A),D\in \Der_\varphi(A,M)\}\]
+	then notice that 
+	\begin{align*}
+		(\varphi,D)(ab) &= (\varphi(a)\varphi(b),\varphi(a)D(b)+D(a)\varphi(b))\\
+		&=(\varphi(a),D(a))\cdot(\varphi(b),D(b)).
+	\end{align*}
+	so this is indeed an algebra morphism. The other direction is clear, so the bijetion holds.
+
+	Now consider the map $p:A\oplus M\to A$ and the induced map
+	\[\ker p_\ast\hookrightarrow G_A(A\oplus M)\xrightarrow{G_A(A)}\]
+	but the projection map admits a section, so $G_A(A\oplus M)$ is a semidirect product. What is $\ker p_\ast$, though?
+
+	\[\ker p_\ast=\{(\varphi,D)|p_\ast(\varphi,D)=\varphi=\id\in G_A(A)\}\]
+	so in particular $\varphi=\varepsilon$. Thus $\ker p_\ast\cong \Der_\varepsilon(A,M)$,
+	hence $G_A(A\oplus M)\simeq \Der_\varepsilon(A,M)\rtimes G_A(A)$.
+
+	For all $D'\in \Der_\varepsilon(A,M)$, the element $(D',\id_A)\in\Der_\varepsilon(A,M)\rtimes G_A(A)=G_A(A\oplus M)$
+	and this gives us (why?) a ``hidden'' correspondence between the $\varepsilon$ derivations and $k$-linear derivations.
+
+	To find this, take $D'\in\Der_\varepsilon(A,M)$ and consider the element $(\varepsilon, D')\in\ker p_\ast$.
+	Multiplying, we get (recalling that multiplication of functions is given by 
+	\[(\varphi\cdot\varphi')(x)=m\circ (\varphi\otimes \varphi')\circ \Delta(x)\]
+	in a Hopf algebra)
+	\[(\varepsilon, D')\cdot(\id_A,0)=(\varepsilon\cdot\id_A,a\mapsto \sum a'D'(a''))\]
+
+	Now use the map $d_A^\varepsilon:A\to A\otimes I/I^2$ via $a\mapsto 1\otimes \pi(a)$ and compute $\id_A\cdot d_A^\varepsilon$, which gives us our map 
+	$a\mapsto \sum a'\otimes \pi(a'')$ which proves things apparently but some logic is missing.
+\end{prf}
+
+\subsection{Our next goal}
+We want to show that, in characteristic zero, any algebraic group scheme is reduced (and thus smooth).
+\begin{lem}
+	Let $X-1,\dots,x_r$ be a basis for $I/I^2$. Then $\{x_i^{m_i}\cdots x_r^{m_r}|\sum m_i=n\}$
+	is a basis for $I^n/I^{n+1}$.
+\end{lem}
+\begin{rmk}
+	Notice that this definitely doesn't work if our algebra is not a Hopf algebra! 
+\end{rmk}
+\begin{prf}
+	Essentially this boils down to the existsnce of derivations with special properties: there is a $D_i:A\to A$ for all $1\le i\le r$
+	such that 
+	\[D_i(x_j)=\delta_{ij}\pmod{I},\]
+	essentially giving us differential operators that distinguish variables.
+	To see why the result follows, notice that if this claim is true then the Leibniz rule and induction gets us 
+	\[D_r^{m_r}D_{r-1}^{m_{r-1}}\cdots D_1^1(x_1^{m_1}\cdots x_r^{m_r})=m_1!\cdots m_r!\pmod{I}\]
+	which implies the statement.
+
+	To generate these operators, we will use the last result. Recall that $\Der_\varepsilon(A,A)=\Hom_k(I/I^2,A)$. Then consider
+	\begin{figure}[h]
+		\centering
+		\begin{tikzcd}
+			A\ar[rrd]\ar[d,"\pi"] &  &\\
+			I/I^2\ar[r,"d_i"] & k \ar[r,hook] & A
+		\end{tikzcd}
+	\end{figure}
+
+	\noindent where $d_i(x_j)=\delta_{ij}$. But notice that with such a $d_i$ we get a $k$-derivation $D_i:A\to A$
+	defined by 
+	\[D_i(a)=\sum a' d_i(\pi(a''))\]
+	and 
+	\[\varepsilon D_i(a)=\sum \varepsilon(a') d_i(\pi(a''))=d_i(\pi(\sum\varepsilon(a')a''))=d_i(\pi(a))=d_i(\pi(x_j))=\delta_{ij}\]
+\end{prf}
+
 \end{document}

### 2020-01-27 Algebraic Geometry/Algebraic-Geometry.tex
Update

@@
 
+\section{January 27th, 2020}
+
+\subsection{Absolute properties of schemes}
+\begin{defn}
+	A schemes is:
+	\begin{enumerate}
+		\item \textbf{connected} (resp. \textbf{irreducible}) if the topological space $|X|$ is.
+		\item \textbf{reduced} if for all $U\subseteq X$ $\O_X(U)$ is reduced.
+		\item \textbf{integral} if for all $U\subseteq X$ $\O_X(U)$ is an integral domain (iff $X$ is reduced and irreducible).
+		\item \textbf{quasicompact} if $|X|$ is quasicompact.
+		\item \textbf{locally Noetherian} if there is an open cover $\Spec A_i=U_i\subseteq X$ with $A_i$ Noetherian.
+		\item \textbf{Noetherian} if $X$ is locally Noetherian and quasicompact (iff there is a finite Noetherian open cover).
+		\item \textbf{normal} if there is an open cover $\Spec A_i=U_i\subseteq X$ with $A_i$ normal (integrally closed domain).
+	\end{enumerate}
+\end{defn}
+
+A question: if $\Spec A$ is locally Noetherian, is $A$ Noetherian? The answer is yes! We need a sublemma:
+\begin{lem}
+	Suppose that $\Spec B\subseteq \Spec A$ is open with $f\in A$ such that $D(f)\subseteq\Spec B$. Then $D(f)=D(g)$
+	where $g=f|_{\Spec B}\in\O(\Spec B)=B.$
+\end{lem}
+Then we may replace the $\Spec A_i$ by $\Spec A_{f_i}$ such that $A_{f_i}$ is Noetherian. But then $A\to \prod_1^n A_{f_i}$ is a faithfully flat 
+extension of a Noetherian ring, so it follows from the following:
+\begin{lem}
+	Let $A\to B$ be faithfully flat. Then $B$ being Noetherian implies $A$ is.
+\end{lem}
+You can see this by tensoring up a chain $I_\bullet$ in $A$ to a sequence $I_\bullet\otimes B\cong (I_iB)_{i\in\bbN}$. This is a sequence in $B$.
+
+\subsection{Examples}
+Consider $f\in k[x,y]$. THen $\Spec k[x,y]/(f)$ is integral.
+Consider $n$ copies of $\A1$ glued along $\A1\setminus 0$. This is also integral.
+
+Think about $\Spec k[x,y]/(xy)$. Draw a picture. Is it reduced/ireducible/connected?
+
 \end{document}

### 2020-01-27 Algebraic Groups/Algebraic-Groups.tex
Update

@@
 
+\section{January 27. 2020}
+We'll pick back up with proving that a finitely generated Hopf algebra in characterisitc zero is reduced.
+
+\subsection{The result}
+\begin{thm}
+	Let $A$ be a finitely generated Hopf algebra over $k$ with $\ch k=0$. THen $A$ is reduced.
+\end{thm}
+\begin{prf}
+	Let $Y\in I$ be a nilpotent with $y^m=0$. We want to show that $y=0$. It sufficies to show this for $m=2$.
+	We claim that $y\in \cap_{n\ge 1}I^n$. To see this, suppose that $y\in I^n\setminus I^{n+1}$.
+	then we can write 
+	\[y=\sum_{\sum m_i=n}a_{\underline m}x^{\underline m}+I^{n+1}\]
+	by the previous lemma. Then consider 
+	\[y^2=p(x_1,\dots,x_\Gamma)+I^{2n+1}\]
+	where $p$ is some monomial of degree $2n$ with some nonzero coefficient. Hence $y^2$ is nonzero mod $I^{2n+1}$, a contradiction!
+	Hence $y\in\cap I^n$.
+
+	We actually need that $k=\bar k$. Let $g:A\to k$ be a point in $G_A$. Then define the translation map
+	\[T_g:A\xrightarrow{\Delta}A\otimes A\xrightarrow{g\otimes \id}k\otimes A\simeq A\]
+	which is a map of algebras and is, in fact, invertible.
+
+	Define $\frakm_g=\ker g$ (think this is the maximal corresponding to the point $g$ in the spectrum). Then we claim that $T_g(I)=\frakm_g$.
+	If $y\in I$ then $y\in \frakm_G$ (since it's nilpotent it lies in every maximal). Thus 
+	by the previous claim $y\in\cap_{n\ge 1} \frakm_g^n$ and by the Krull intersection theorem, $y=0$ in $A_{\frakm_g}$ for all $\frakm_g$,
+	whence $y=0.$
+\end{prf}
+
+\subsection{Properties of \texorpdfstring{$\Omega_A$}{Omega A}}
+\begin{itemize}
+	\item $\Omega_{A\times B}\cong\Omega_A\times\Omega_B$
+	\item Let $S$ be a multiplicative system in $A$. Then  
+	\[\Omega_{S^{-1}A}\cong \Omega_A\otimes_A S^{-1}A\]
+	and in particular if $A$ is an integral domain, let $K=\operatorname{Frac}(A).$ Then 
+	\[\Omega_{K/k}\cong \Omega_A\otimes_A K.\]
+	\item Let $L/k$ be a finitely generated seperable field extension. That is we have 
+	\[L\supset E=k(x_1,\dots,x_\Gamma)\supset k\]
+	where $L/E$ is finite. Then $\dim_L\Omega_{L/k}=\operatorname{trdeg}_k L$ and $dx_1,\dots,dx_\Gamma$ forma a basis for $\Omega_{L/k}$.
+	\item Let $G$ be an algebraic group scheme. Maybe assume $k=\bar k$. Assume that $k[G]$ is an integral domain. Let $K=\operatorname{Frac}(k[G])$. Then 
+	\[\rank\Omega_{k[G]}=\operatorname{trdeg}_k K\]
+\end{itemize}
+The last item merits some discussion:
+\begin{prf}
+	By the third item above, we get 
+	\[\Omega_{K/k}=\Omega_{k[G]}\otimes_{k[G]}K\]
+	so 
+	\[\operatorname{trdeg}_k K=\dim_K\Omega_{K/k}=\rank \Omega_{k[G]}\]
+\end{prf}
+\begin{defn}
+	Let $G$ be an algebraic group scheme over $k$. Then $G$ is smooth if 
+	\[\dim G=\Omega_{k[G]}\]
+	where $\dim G=\dim_{\text{Krull}} k[G]$.
+\end{defn}
+\begin{thm}\label{thm:smoothgrps}
+	Let $G$ be an algebraic group scheme over $k$. Then if
+	\[\bar k[G]=k[G]\otimes_k \bar k\]
+	is reduced, $G$ is smooth.
+\end{thm}
+\begin{cor}
+	In characteristic zero, all algebraic group schemes are smooth.
+\end{cor}
+\begin{rmk}
+	Smoothness is independent of extending scalars (as the rank and dimension in the definition are), but 
+	you do need to extend scalars enough so that the nilpotents aren't ``hiding''. Thus it suffices to only extend to 
+	the seperable closure of $k$ or any perfect field $L/k$.
+\end{rmk}
+The following discussion and proof can be found in section 14 (and partially 6.6) of Waterhouse.
+\begin{prf}[of thm.~\ref{thm:smoothgrps}]
+	First extend scalars up to $\bar k$. Second recall (from section 6.6) that for $A=k[G]$, the following are equivalent:
+	\begin{itemize}
+		\item $\pi_0(A)$ is trivial
+		\item $\Spec A$ is connected
+		\item $\Spec A$ is irreducible
+		\item $A/\operatorname{Nil}(A)$ is an integral domain.
+	\end{itemize}
+	The second two are actually a statement about algebraic geometry (recall the geometry doesn't see the nilradical).
+
+	Assume that $k[G]$ is reduced. THen we can assume that $G=G^0$ is connected since 
+	\[k[G]\simeq k[G^0]\times\cdots\times k[G^0]\]
+	where we have $|\pi_0(k[G])|$ copies in the product. Then we use the first property in the last section.
+
+	For $G^0$ connected, the general properties in the last section imply that $k[G^0]$ is irreducible implies that $k[G^0]/\operatorname{Nil}(k[G^0])$ is 
+	an integral gdomain. By assumption, the nilradical is trivial, so $k[G^0]$ itself is an integral domain.
+
+	Let $K=\operatorname{Frac}(k[G^0])$. Then
+	\[\dim G^0=\dim k[G^0]\stackrel{\text{AG}}{=}\operatorname{trdeg}_kK=\rank \Omega_{k[G^0]}\]
+	where the $AG$ referenced above is essentially the Noether normalization lemma.
+\end{prf}
+
 \end{document}

### 2020-01-29 Algebraic Geometry/Algebraic-Geometry.tex
update

@@
 \subsection{Examples}
-Consider $f\in k[x,y]$. THen $\Spec k[x,y]/(f)$ is integral.
+Consider $f\in k[x,y]$. Then $\Spec k[x,y]/(f)$ is integral.
 Consider $n$ copies of $\A1$ glued along $\A1\setminus 0$. This is also integral.
@@
 
+\section{January 29th, 2020}
+\subsection{Example}
+Consdier the tangent space (the $\Hom$ from $\Spec k[\varepsilon]/(\varepsilon^2)$) to the space $\Spec k[x,y]/(xy)$ (the axes in $\bbR^2$).
+
+Looking at the algebra maps, you can map $x$ and $y$ freely to multiples of $\varepsilon$, giving us a two dimensional thing! Otherwise you can compute you only have one 
+degree of freedom. It is (locally) Noetherian, reducible, integral, and other stuff.
+
+You can do this with higher-order vectors as well! Think about what it means to map $\Spec k[\varepsilon]/(\varepsilon^n)$ into a scheme.
+
+\brk
+
+Think more about the countable number of lines glued together in a criss-cross. This is a scheme! It is locally Noetherian but not Noetherian. Also think about the line with infinitely many origins.
+
+\subsection{AG is relative}
+\begin{defn}
+	A \textbf{morphism $f:X\to Y$ of schemes} is 
+	\begin{itemize}
+		\item \textbf{quasicompact} if there is an open cover $U_i\subseteq Y$ with $U_i$ and $f^{-1}(U_i)$ quasicompact.
+		\item \textbf{locally of finite type} if there exists an open cover $\Spec A_i=U_i\subseteq Y$ such that each $f^{-1}(U_i)$ has an open cover $\Spec B_{ij}$ such that each map $A_i\to B_{ij}$ 
+		makes $B_{ij}$ into a finitely generated $A_i$ algebra.
+		\item \textbf{of finite type} if $f$ is locally of finite type and quasicompact.
+		\item \textbf{affine} if there exists an open covering $\Spec A\subseteq Y$ such that $f^{-1}(U_i)$ is affine.
+		\item \textbf{finite} if it is affine and there is an open cover $\Spec A_i\subseteq Y$ such that $f^{-1}(U_i)=\Spec B_i\to \Spec A_i$ is (module-)finite.
+		\item \textbf{quasifinite} if the map of topological spaces has finite fibers.
+		\item \textbf{locally quasifinite} if there exists a cover $U_i\subseteq Y$ and $V_{ij}\subseteq X$ such that $V_{ij}\to U_i$ is quasifinite.
+	\end{itemize}
+\end{defn}
+
+This is a theorem with a thousand different formulations so you may not find this one online:
+\begin{thm}[Zariski's Main Theorem]
+	Finite morphisms are precisely the proper and quasifinite ones.
+\end{thm}
+
+\subsection{An observation}
+Let $X$ a variety. Then how do we understand this in the language of schemes? It is an \textit{integral} $k$-scheme such that the map 
+$X\to \Spec k$ is
+\begin{itemize}
+	\item of finite type,
+	\item and separated (will see this later).
+\end{itemize}
+So finite type is aprt of ``relative variety.'' We can imagine a family of varieties given by the zero sets 
+of $y^^2=x^3+tx+4$ and consider this as a closed subvariety in $\A 2$ over $\A1$. Then we could specialize at any prime and recover a variety in $\A2_\bbZ$.
+Furthermore varying $4$ can get us more.
+
+\subsection{Another observation}
+If $f:X\to Y$ is finite, then $f$ is quasifinite. To see this, let $y\in \Spec A\subset Y$. Then since $f$ is finite,
+$f^{-1}(\Spec A)=\Spec B$ and $A\to B$ is a finite ring map (some ordering needs to be fixed there on choosing $A$ and $B$).
+Thus without loss of generality, we can assume $X=\Spec A$ and $Y=\Spec B$. So then $y\in Y$ if and only if $p\lhd A$ is prime.
+
 \end{document}

### 2020-01-29 Algebraic Groups/Algebraic-Groups.tex
update

@@
 
+\section{January 29th, 2020}
+As a reminder: we are not having classes February 10-14. The following Monday is president's day. So keep that in mind. :)
+
+\subsection{Lie algebras for \texorpdfstring{$G$}{G}}
+\begin{defn}
+	If $G$ is an algebraic group scheme over $k$, then $\Lie G$ is the space of left-invariant derivations $D:k[G]\to k[G]$.
+\end{defn}
+What is a left-invariant derivation? Let $A=k[G]=k^G$. The Yoneda lemma allows us to identify this with $\Hom(G,\A1)$.
+But given $f:G\to k$, we get an action 
+\[g\cdot f(-)=(T_g f)(-)=f(g-)\]
+Note that this actually defines a \textit{right action}, but the left and right invariant derivations are canonically isomorphic and this 
+will help us with the Lie algebra case.
+
+Now let $x\in G$ and by abuse of notation let $x=\operatorname{ev}_x:A\to k$ be the evaluation map. This gives us a diagram:
+\begin{figure}[h]
+	\begin{tikzcd}
+		A\ar[r,"\Delta"]\ar[d,"T_g"] & A\otimes A\ar[r,"(g;\operatorname{id})"] & A\ar[d,"x"]\\
+		A\ar[rr,"x"] & & k
+	\end{tikzcd}
+\end{figure}
+
+Let $T:A\to A$ be a $k$ map. THen $T$ is left invariant if $T_g\circ T=T\circ T_g$ for all $G$. 
+I wrote some notes in my digital pad. We did some computation. The result is that it is equivalent to having $\Delta\circ T=(\id\otimes T\circ \Delta)$.
+
+Thus we can restate the definition of $\Lie G$ as 
+\[\Lie G=\{D\in\Der_k(A,A)|\Delta\circ D=(\id\otimes D)\circ \Delta\}.\]
+\begin{lem}
+	$\Lie G$ has the structure of a Lie algebra with bracket 
+	\[[D_1,D_2]=D_1\circ D_2-D_2\circ D_1\]
+\end{lem}
+To prove that this we need to show the three things for the Lie bracket hold! The $k$ bilinearity and $[D,D]=0$ (careful! We want to consider characteristic 2!)
+are easy to show but the Jacobi identity requires some work.
+
+Notice that it is equivalent to saying that the operator $\ad D$ is a derivation:
+\[\ad D([D_1,D_2])=[\ad D(D_1),D_2]+[D_1,\ad D(D_2)]\]
+and one can check this. It has nothing to do with left invariance!
+
+\subsection{Another way to think of this algebra}
+\begin{lem}
+	There are natural bijections between 
+	\begin{enumerate}
+		\item $\Lie G$ (the left invariant derivations on $k[G]$)
+		\item $\Der_\varepsilon(k[G],k)$
+		\item $\ker (G(k[\tau]/(\tau^2))\to G(k))$ where the map is induced from the projection onto the constants in $k[\tau]/(\tau^2)$.
+	\end{enumerate}
+\end{lem}
+\begin{prf}
+	We can use that $G(k[\tau]/(\tau^2))=\Hom(A,k\oplus k\tau$ and then study the kernel.
+\end{prf}
+
 \end{document}

### 2020-02-03 Algebraic Geometry/Algebraic-Geometry.tex
update

@@
 
+\section{January 31, 2020}
+Today we are going to talk about fiber products. Max thinks they should be ``fibered products'' but he's wrong.
+
+\subsection{Fiber products are real good}
+\begin{defn}
+	A fiber product $X\times_Z Y$ is the limit of $X\to Z\leftarrow Y$.
+	You know the drill.
+\end{defn}
+Equivalently the fiber product of schemes is the fiber product of the functor $\Sch\op\to \Set$ that it represents.
+The reason this is worth mentioning is that one can use the fuctor of points perspective to define this as a fiber product 
+on sets!
+
+\begin{thm}
+	Fiber products exist in $\Sch$.
+\end{thm}
+\begin{prf}
+	We gonna do this.
+	\begin{defn}
+		A functor $\calF:\Sch\op\to\Set$ is a \textbf{sheaf} if for all $W\in\Sch$, the induced functor 
+		$\calF_W:\mathbf{Open}(W)\op\to\Set$ is a sheaf.
+	\end{defn}
+	Kinda obvious. Then the idea here is that the Yoneda embedding actually maps into sheaves on schemes (not just presheaves). Furthermore 
+	since the category of sheaves has all limits (CHECK THIS! It's easy), this means we are just trying to show that the sheaf $h_X\times_{h_Y}h_Z$ is representable 
+	by a scheme.
+
+	\begin{defn}
+		A natural transformation $\calG\to \calF$ in $\Func(\Sch\op,\Set)$ is an \textbf{open subfunctor} if, 
+		for all schemes $W$ and all $h_W\to F$, the pullback $G\times_F h_W\to h_W$ is representable by 
+		an open immersion $h_U\to h_W$ ($U\subseteq W$ open).
+	\end{defn}
+	We're trying to define a thing like a topology on the category of schemes.
+	\begin{defn}
+		Given a fucntor $\calF\in\PreSh(\Sch)$, an \textbf{open covering} is a collection of open 
+		subfunctors $\calG_i\hookrightarrow\calF$ such that for all $W\in\Sch$ $h_{U_i}=\calG_i\times\calF h_W\to h_W$.
+	\end{defn}
+	\begin{prop}
+		A sheaf $\calF:\PreSh(\Sch)$ is representable by a scheme if and only if $\calF$ admits an open covering by functors 
+		representable by affine schemes.
+	\end{prop}
+	You can try to prove it! It also is done in Demazure and Gabriel although I am pretty sure that is in French and/or inscrutable. 
+	Max says it's not too hard, though. You can kinda follow the argument in Hartshorne for the analogous thought coming from the category of locally ringed spaces.
+
+	\begin{lem}
+		Let $U,V,W$ be opens in $X,Y,Z$, resp. such that $f(U)\subset W\supset g(V)$. Then $h_U\times_{h_W}h_V\to h_X\times_{h_Z}h_Y$ is an open subfunctor.
+	\end{lem}
+	Again easy to check. This gives us (from affine coverings of $X,Y,Z$) an open covering by $h_{\Spec A}\times_{h_{\Spec C}}h_{\Spec B}$.
+
+	From here we can use (among other things) that the restricted Yoneda $\Sch\to\PreSh(\Aff)$ is fully faithful.
+\end{prf}
+You can read all of today's lecture in Vakil chapter 9.
+\section{February 3, 2020}
+Today Adam is giving the lecture! The title is
+
+\subsection{Category theory for algebraic geometry}
+This lecture is assuming that we know about categories, functors, and natural transformations.
+We will be focusing on limits and colimits.
+
+A motivating question: say we have an open cover $\cup U_i=X$. How can we relate the sheaves on $X$ to those on the $U_i$? Saw we have sheaves $\calF_i$ on $U_i$ along with 
+isomorphisms $\phi_{ij}:\calF_i|_{U_i\cap U_j}\to \calF_j|_{U_i\cap U_j}$ (along with a cocycle condition we'll discuss later). How can a sheave $\calF$ on $X$ 
+be related to these?
+
+Take an open $V\subseteq X$. Then for all $i,j,$ we have maps
+\begin{figure}[h]
+	\centering
+	\begin{tikzcd}
+		& \calF(V)\ar[dl]\ar[dr] &\\
+		\calF(V\cap U_i)\ar[d] & & \calF(V\cap U_j)\ar[d]\\
+		\calF_i(V\cap U_i)\ar[d] & & \calF_j(V\cap U_j)\ar[d]\\
+		\calF_i(V\cap U_i\cap U_j)\ar[rr,"\varphi_{ij}"] & & \calF_j(V\cap U_i\cap U_j)
+	\end{tikzcd}
+\end{figure}
+
+The gluing idea is to define $\calF(V)$ to be the limit of the bottom portion of the diagram. To do this in general,
+let $\calC$ be a category and $K:\calJ\to\calC$ for some other category $\calJ$. Define a cone over $K$ with summit $c\in \calC$ 
+to be the usual thing. Same for a cone under a functor. Then limits and colimits. Most of this is 
+basically the same as Emily Riehl.
+
+He uses the fact that $\operatorname{cone}(-,K):\calC\op\to\Set$ (this assumes $\calJ$ is small) is a functor,
+and furthermore that 
+\[\Hom_\calC(-,\lim K)\simeq \operatorname{cone}(-,K)\]
+which is essentially a restatement of the universal property of a limit.
+
+Some examples:
+\begin{itemize}
+	\item In $\Set$, the limit of $A\hookrightarrow X\hookleftarrow B$ is $A\cap B$.
+	\item Equalizers are limits!
+\end{itemize}
+
 \end{document}

### 2020-02-03 Algebraic Groups/Algebraic-Groups.tex
update

@@
 
+\section{January 31, 2020}
+Recall that we had $\Lie G$ as well as bijections between a couple different things.
+Today we are going to talk more about that. We scribbled out some easy computations for why (b) and (c) are equivalent.
+The idea here is that any map $A\to k\oplus k\tau$ can be written as $(\varphi,d)$ where $d$ is a $\varphi$ derivation.
+
+To show the equivalence of $(a)$ and $(b)$, you use the maps $D\mapsto \varepsilon D$ and the inverse $d\mapsto (\id\otimes d)\circ\Delta$.
+One needs to check that you get derivations and that the other thing
+is left-invariant. To show that they are inverses, consider 
+\[\varepsilon(\id\bar\otimes d)\circ \Delta(x)=\sum\varepsilon(a')d(a'')=d(\sum \varepsilon(a')a'')=d(a).\]
+Leave the other direction as an exercise.
+
+Let's check left invariance: we need to check (by the discussion the other day) that $\Delta\circ D=(d\otimes D)\circ\Delta.$ Checking this si doing a little diagram chase.
+Part of the diagram is associativity.
+
+\subsection{Lie structures}
+A natural question to ask: if we have this isomorphism in this last lemma, what are the brackets on the other two objects?
+
+For the first, let $d_1,d_2\in\Der_\varepsilon(A,k)$. Let $D_i=(\id\bar\otimes d_i)\circ\Delta$. Do some work and chase the diagrams for $[D_1,D_2]$!
+Then you find that 
+\[[d_1,d_2]=(d_1\bar\otimes d_2-d_2\bar\otimes d_1)\circ\Delta\]
+which really underlines that we need the group/Hopf structure on these things to really be able to do this! This is what makes these equivalences work!
+
+For the kernel $\ker p_\ast$, take two elements $g_1=(\varepsilon, d_1)\in k[\tau_1]$ and $g_2=(\varepsilon, d_2)\in k[\tau_2]$ where we 
+are using different nilpotents so that we have enough degrees of freedom. Consider $R=k[\tau_1,\tau_2]$ and consider the inclusion of $g_1$ and $g_2$ in $G(R)=\Hom(A,R)$.
+
+So we think of $g_1=\varepsilon+d_1\tau_1$ and $g_2=\varepsilon+d_2\tau_2$ and we compute the multiplication as:
+\[g_1g_2=\varepsilon+d_1\tau_1+d_2\tau_2+((d_1\bar\otimes d_2)\circ\Delta)\tau_1\tau_2\]
+and coompute $g_2g_1$ as well. Then we really don't have additive inverses in $G$, but we can compute 
+\[g_1g_2=(\varepsilon+[d_1,d_2]\tau_1\tau_2)g_2g_1.\]
+
+How do we use this in practice? Consider $\GL_n(S)$. The Lie algebra consists of the matrices $M\in\GL_n(k[\tau])$.
+Each such matrix can be written as $M_1+M_2\tau$. But we want the condition that the projection onto $\GL_n(k)$ is the identity,
+so really we have 
+\[\Lie\GL_n = \{I+M\tau\}\subset\GL_n(k[\tau])\]
+but any such matrix is invertible if $M$ is, so it is no further condition. This gets you your isomorphism with all of $\gl_n$.
+
+\section{February 3rd, 2020}
+Working from last time, we have some corollaries:
+\begin{cor}
+	Let $f:H\to G$ be a map. Then 
+	\begin{itemize}
+		\item $f$ induces a map $\Lie H\to \Lie G$ ($\Lie$ is functorial).
+		\item If $f$ is inclusion, then $\Lie H\hookrightarrow\Lie G$
+	\end{itemize}
+\end{cor}
+\begin{rmk}
+	This gives us the full equivalences 
+	\[\Lie G\simeq\Der_\varepsilon(k[G],k)\simeq (I/I^2)^\ast\simeq T_e G.\]
+\end{rmk}
+and from this we get 
+\begin{cor}
+	$G$ is smooth (or reduced) if and only if $\dim G=\dim_k \Lie G$.
+\end{cor}
+where the above is true since $\dim_k\Lie G=\rank \Omega_{k[G]}$.
+
+\subsection{Examples}
+We talked about $\GL_n$ last time. Also if we have $I+\tau_1M_1,I+\tau_2 M_2\in\Lie \GL_n=\gl_n$, the bracked can quickly be seen to be 
+\[I+\tau_1\tau_2[M_1,M_2]\]
+as one would hope.
+
+Now consider $\SL_n$. Then $\Lie\SL_n\subset\gl_n$. To analyze this, consider $\SL_n(k[\tau])$ 
+and we are looking for the ones with determinant one. But 
+\[\det(I+\tau M)=p(\tau)\]
+where $p$ is the characteristic polynomial for $I+\tau M$ over $k$. But the terms above the linear term die and the linear term is the trace (of $M$):
+\[\det(I+\tau M)=1+(\tr M)\tau=1\]
+so the condition is $\tr M=0$.
+
+\subsection{Positive characteristic}
+Recall that $\Ga$ is the group scheme represented by $k[t]$ where $t$ is primitive. $\Gm$ is 
+represented by $k[t,t^{-1}]$ where $t$ is grouplike.
+
+These are both one-dimensional smooth group schmemes, so the Lie algebras $\frakg_a=\Lie \Ga$ and $\frakg_m=\Lie\Gm$
+are both one-dimensional. If $\ch k=0$, then, there is only one Lie structure! But what if not?
+Assume for now $\ch k\ne 2$.
+
+Do the following problem! It's not hard.
+\begin{prob}
+	Let $\ch k=p$. Take a derivation $D\in \Der_k(A,A)$ and show that the $p^{th}$ power of $D$ is again a derivation.
+\end{prob}
+
+This give rise to a ``$p^{th}$ power map'' which gives rise to a \textbf{restricted Lie algebra}. See Weibel for a good 
+introduction (he had to really learn it well to write about it so it is very clear).
+
+Go back to the case of $\Ga$ and $\Gm$. $\frakg_a=\Der_\varepsilon(k[t],k)=\langle d_t=\frac{d}{dt}\rangle$
+such that $d_tt=1$.
+
+Then $D_t=(\id\bar\otimes d_t)\circ\Delta$ as the corresponding derivation on $k[G]$, and we can compute 
+\begin{align*}
+	D_t(t)&= (\id\bar\otimes d_t)\circ \Delta(t)\\
+	&= (\id\bar\otimes d_t)(1\otimes t+t\otimes 1)=1
+\end{align*}
+and so $D_t^2(t)=0$. Thus $D_t^p(t)=0$, so $\frakg_a$ is a trivial $p$ Lie algebra: $\frakg_a=\langle d_t\rangle$
+where $d_t^p=0$.
+
+In the case of $\Gm$, we can compute 
+\[D_t(t)=t\]
+so then $D_t^p=D_t$, so we get $\frakg_m=\langle d_t\rangle$ with $d_t^p=d_t$. The punchline here is 
+that the two infinitessimal theories, at first glance, appear to coincide although they actually vary 
+in the right light. $\frakg_a\cong\frakg_m$ as Lie algebras, but not as \textit{restricted} Lie algebras.
+
+\subsection{Frobenius Kernels}
+For reference, this can be found in Jantzen part I section 7. Some of the proof will be omitted, but can be found there.
+
+\brk
+
+Let $\ch k=p>0$ and let $G$ be an algebraic group scheme over $k$. We assume here that $k=\bar k$, but it suffices that 
+$k$ is perfect. This fixes some problems that arise. Recall the Frobenius map $F:\A n\to \A n$
+that sends $(a_1,\dots,a_n)\mapsto(a_1^p,\dots,a_n^p)$. The \textbf{Frobenius twist} of $G$ is the image 
+$G^{(1)}$ of $G$ under the Frobenius twist.
+
+Then we can define the \textbf{Frobenius kernel}
+\[G_{(1)}\ker F\hookrightarrow G\xrightarrow{F} G^{(1)}\]
+which one can show is a connected ($k[G_{(1)}]$ is local) and finite ($k[G_{(1)}]$ is finite dimensional) and a group scheme.
+
+Is the Frobenius kernel smooth? Usually no! The finite dimensionality of $k[G_{(1)}]$ means the dimension of $G$ is zero,
+so $\Lie G$ is trivial. But this happens only if $G$ is trivial.
+
+\brk
+
+An example: $\Ga$. Consider the Frobenius map $F:\Ga\to \Ga$ via $r\mapsto r^p$. Then the kernel 
+is the collection of $p$-nilpotents, 
+\[{\Ga}_{(1)}(R)=\{r\in R|r^p=0\}\]
+and one can compute that 
+\[k[{\Ga}_{(1)}]\cong k[t]/t^p\]
+with $t$ being a primitive element. We call ${\Ga}_{(1)}=\alpha_p$.
+
+When we switch the picture to $\Gm$, we are now looking at $p^{th}$ roots of unity! The algebra is 
+$k[t,t^{-1}]/(t^p-1)=k[t]/(t^p-1)\simeq k[t-1]/(t-1)^p\simeq k[u]/u^p$. So the coordinate algebra is isomorphic 
+to that in the last case, but with a different coproduct.
+
 \end{document}

### 2020-02-12 Algebraic Geometry/Algebraic-Geometry.tex
Update and fix some old stuff

@@
 
+\section{February 5th, 2020}
+Today Adam is talking more about limits!
+
+Recall that Yoneda says that the natural transformations 
+\[\Hom(\Hom_\calC(-,\lim k),\operatorname{cone}(-,K))\simeq\operatorname{cone}(\lim K,K)\]
+
+\subsection{Equilizers}
+Some clarity here: an equilizer in $\Ab$ can be a couple things:
+\begin{itemize}
+	\item The limit of a pair of parallel functors.
+	\item It is the same as saying that the following is exact:
+	\[0\to A\to B\xrightarrow{f-g} C\]
+\end{itemize}
+
+\subsection{Base change}
+This is one of the most important (co)limits in this subject (and in number theory and representation theory).
+If $A$ is a $k$-algebra and there is an inclusion $k\hookrightarrow L$. How can we extend $A$ to an $L$-algebra?
+
+Say that $B$ is an extension of $A$ that is also an $L$-algebra. Then if the universe loves us, we should get 
+\begin{figure}[h]
+	\centering
+	\begin{tikzcd}
+		k\ar[r]\ar[d] & A\ar[d]\\
+		L\ar[r] & B
+	\end{tikzcd}
+\end{figure}
+so if we want the ``best'' one, we need to find the colimit of $L\leftarrow k\to A$. To define this thing we just 
+trace through and show that the object with the usual axioms of the tensor product falls out. :)
+
+\begin{prop}
+	The following is a pullback (in schemes, locally ringed spaces, etc)
+	\begin{figure}[h]
+		\centering
+		\begin{tikzcd}
+			\Spec A\otimes_k L\ar[r]\ar[d] & \Spec L\ar[d]\\
+			\Spec A\ar[r] & \Spec k
+		\end{tikzcd}
+	\end{figure}
+\end{prop}
+\begin{rmk}
+	Why is it called a fiber product? Look at what happens in $\Set$. Look at the limit of the diagram 
+	$\ast\stackrel{y}{\hookrightarrow} Y \xleftarrow{f}X$. The pullback is literally the fiber over $y\in Y$.
+\end{rmk}
+Proving the following is a good excercise. Go try it! It's a fun categorical proof. You can do it by looking at cones. 
+\newpage
+\begin{lem}
+	If $\pi:X\to Y$ is a monomorphism, then in the pullback square 
+	\begin{figure}[h]
+		\centering
+		\begin{tikzcd}
+			X'\ar[r]\ar[d,"\pi'"] & X\ar[d,"\pi"]\\
+			Y'\ar[r] & Y
+		\end{tikzcd}
+	\end{figure}
+
+	\noindent $\pi'$ is also mono.
+\end{lem}
+As a corollary of this, if we have an open $U\subseteq X$ in a scheme and $\pi:Y\to X$ a morphism, then 
+$\pi^{-1}(U)$ is open in $Y$.
+
+\subsection{Gluing, baby}
+For the setup, remember that $X$ is a topological space and $X=\cup_i U_i$ open cover. We are given sheaves 
+$\calF_i$ on $U_i$ together with isomorphisms 
+\[b_{ij}:\calF_i|_{U_i\cap U_j}\to\calF_j|_{U_i\cap U_j}.\]
+
+We saw that the right way to glue to an $\calF$ on $X$ is to take 
+$\calF(V)$ to be a limit of the diagram we drew yesterday (just the bottom two rows).
+Then there are two questions we ask:
+\begin{itemize}
+	\item Is $\calF|_{U_i}=\calF_i$? This is answered by the \textit{cocycle condition}.
+	\item Is $\calF$ a sheaf? This is due to the fact that limits commute with limits.
+\end{itemize}
+
+\section{February 7th, 2020}
+Max's back. Time to talk about 
+\subsection{Separatedness and Properness}
+These are replacements for the Hausdorff and Compact conditions, neither of which we get in algebraic geometry. Let's get some definitions 
+\begin{defn}
+	A morphism $f:X\to Y$ is \textbf{separated} if the diagonal $\Delta:X\to X\times_Y X$ is a closed immersion. This 
+	replaces Hausdorff.
+\end{defn}
+Recall that a closed immersion is a homeomorphism onto a closed subspace such that the map of rings is surjective. 
+The prototype you should be thinking of is $\Spec k[x_1,\dots,x_m]/(f_1,\dots,f_n)$ as a closed subspace of $\A m$.
+\begin{defn}
+	A morphism $f:X\to Y$ is \textbf{closed} if the closed subsets of $X$ are sent to closed subsets of $Y$ (as topological spaces).
+
+	It is \textbf{universally closed} if for all $T\to Y$, the base change map $f_T:X\times_Y T\to T$ is closed.
+\end{defn}
+\begin{ex}
+	Consider $\A1_k\to\Spec k$. This is closed, but not universally closed! Consider the base change via $\A1\to\Spec k$:
+	$\A1\times\A1=\A2\to \A1$. In coordinates, we get the map of algebras $k[s]\hookrightarrow k[t,s]$. But the hyperbola 
+	cut out by $ts-1$ in $\A2$ projects onto $\A1\setminus 0$ under this map, which is super not closed.
+
+	Is $\A1\to \Spec k$ separated? Hell yes! Look at the ring map $A\times A\to A$ by multiplication. This is surjective.
+	And unravelling the definitions gets you the result: every morphism of affines is separated! So we don't really need to check 
+	that the inclusion is a topological embedding onto a closed set.
+\end{ex}
+
+I (well, Max) claim(s): If $f:X\to Y$ is separated and $\sigma:Y\to X$ is a section, then $\sigma$ is closed.
+In fact, check this out: \textbf{any section is a pullback of the diagonal} or even \textbf{``The diagonal is the universal section.''}
+
+\begin{figure}[h]
+	\centering
+	\begin{tikzcd}
+		X\ar[r]\ar[d,bend right,"f",swap]\ar[rr,bend left,"\id"] & X\times_Y \tilde X\ar[r]\ar[d,swap,"f_X"] & \tilde X\ar[d,"f"]\\
+		Y\ar[u,"\sigma",swap] \ar[r,"\sigma"] & X\ar[u,bend right,swap,"\Delta"] \ar[r,"f"] & Y
+	\end{tikzcd}
+\end{figure}
+Check it out. Assuming everything is a pullback square (remember the neat two of three thing for pullbacks) we get that $\Delta$ is 
+a pullback through itself of the diagonal! This gives us another result: $\Delta$ is closed if and only if sections are universally closed.
+
+So sections are closed, but ot universally closed! For instance the line with two origins is not separated!
+
+\begin{defn}
+	A morphism $f:X\to Y$ is \textbf{proper} if it is of finite type, separated, and universally closed.
+\end{defn}
+Note: $f$ is separated if and only if ``sections of $f$'' are universally closed. An example, $\P n$ is proper over $\Spec\bbZ$.
+So of course it is proper over anything.
+
+\section{February 10th, 2020}
+Today Taffy is talking about base change!
+
+\subsection{Base change}
+Notice this cool thing that we see in topological spaces: let $f:U\hookrightarrow Y$ be an open embedding. Then consider the pullback
+\begin{figure}[h]
+	\centering
+	\begin{tikzcd}
+		X\times_Y U\ar[r]\ar[d,"g"] & U\ar[d,"f"]\\
+		X\ar[r] & Y
+	\end{tikzcd}
+\end{figure}
+where we see $X\times_Y U\cong f^{-1}(U)$ and $g:f^{-1}(U)\hookrightarrow X$ is the open embedding. Here we just did base change and we saw a property being preserved!
+
+In $\Sch$, we can do the same kind of thing: Let $\calX$ and $\calY$ be schemes over $k$.
+
+I just listened to the rest. Good talk!
+
+\section{February 12th, 2020}
+Today we are going to talk more about seperatedness and properness and in particular the so-called 
+\subsection{Valuative Criteria}
+The idea here is that checking for separatedness or properness can be quite daunting sometimes. We will give a way to do this!
+\begin{defn}
+	A domain $R$ is a \textbf{valuation ring} if for every $x\in \kappa(R)^\times$, we have $x\in R$ or $x^{-1}\in R$.
+\end{defn}
+\begin{lem}
+	Consider loal rings contained in a field $k$ (with fraction field $k$). Say that $(S,\frakm_S)$ \textit{dominates} $(R,\frakm_R)$ if $R\subset S$ 
+	and furthermore $\frakm_S\cap R=\frakm_R$. This gives a partial ordering.
+
+	Then the valuation rings in $k$ are precisely the maximal elements for this ordering.
+\end{lem}
+\begin{defn}
+	A morphism $f:X\to Y$ is \textbf{quasi-separated} if the diagonal $\Delta:X\to X\times_Y X$ is quasicompact.
+\end{defn}
+THis is always true if $X$ is locally Noetherian. This will be true in almost every case you see in real life, because it 
+is really necessary to make the entire theory work.
+
+\begin{defn}
+	A \textbf{test diagram} for a morphism $f:X\to Y$ is a diagram as in the figure below	
+	where $i:R\hookrightarrow k=\kappa(R)$ is a vaulation ring.
+	A \textbf{filling} for a test diagram is an arrow $\gamma:\Spec R\to X$ making this diagram commute.
+\end{defn}
+\begin{figure}[h]\label{fig:testdiagram}
+	\centering
+	\begin{tikzcd}
+		\Spec k\ar[r]\ar[d,"i^\ast"] & X\ar[d,"f"]\\
+		\Spec R\ar[r] & Y
+	\end{tikzcd}
+	\caption{A test diagram for $f$}
+\end{figure}
+\begin{thm}
+	\begin{enumerate}
+		\item A quasiseparated morphism $f:X\to Y$ is separated if and only if every test diagram has at most one filling.
+		\item A quasicompact morphism $f:X\to Y$ is universally closed if and only if every test diagram admits at least one filling.
+		\item A quasisemparated morphism of finite type $f:X\to Y$ is proper sif and only if every test digram admints precisely one filling.
+	\end{enumerate}
+\end{thm}
+\begin{lem}
+	The entire theorem above follows from (b).
+\end{lem}
+\begin{prf}
+	To get (a) from (b), $f$ is quasiseparated iff $\Delta:X\to X\times_YX$ is quasicompact. Note that $f$ is separated iff $\Delta$ is 
+	universally closed. Further, note that $\Delta$ is always an immersion. One can show that $\Delta$ is a closed immersion into an open subscheme (look in Hartshorne and use a cover by open affines).
+
+	But then the test diagram for $f$ with two fillings $\alpha,\beta$ gives us a test diagram for $\Delta$ with $(\alpha,\beta)$ along the bottom.
+	What does it mean for this new diagram to have a filling? If we ahve $t:\Spec R\to X$, we get that 
+	\[(\alpha,\beta)=\Delta\circ t=(t,t)\]
+	so we have a filling if and only if $\alpha=\beta$!
+
+	That (b) implies (c) follows from the definitions!
+\end{prf}
+\begin{defn}
+	Given a scheme $X$, a point $s\in X$ is a \textbf{specialization} of a point $t\in X$ (equivalently, $t$ 
+	is a generization of $s$) if $s\in\overline{\{t\}}$.
+\end{defn}
+This is a topological property, but we want to capture it algebraically. We need to pass to limiting algebraic object 
+to do it most efficiently. As a mater of notation, we write $t\rightsquigarrow s$.
+\begin{lem}
+	Suppose $t\rightsquigarrow s$ in $X$ and we are given $K\supset \kappa(t)$ (residue field). Then there exists a vaulation ring $R\subset K=\operatorname{Frac} R$ and 
+	morphism $\Spec R\to X$ such that $(0)\mapsto t$ and $\frakm_R\mapsto s$.
+\end{lem}
+A ``proof'' of this is to look at $\O_{\overline{\{t\}},s}\subset k$ and find $R$ using Zorn's lemma with respect to domination.
+
 \end{document}

### 2020-02-12 Algebraic Groups/Algebraic-Groups.tex
Update and fix some old stuff

@@
 
+\section{February 5th, 2020}
+Today's plan: we're going to go back and talk a bit more about the Frobenius twist to understand it better.
+Then we'll talk about restricted enveloping algebras and then do some synthesis.
+
+\subsection{Frobenius twist}
+Let's saw we have a commutative ring with $R$ modules $A$ and $B$. Then we can take the pushout to get $A\otimes_R B$. Instead, let's
+take a $k$-algebra and an automorphism $\varphi:k\to k$ we take the pushout 
+\begin{figure}[h]
+	\centering
+	\begin{tikzcd}
+		k\ar[r]\ar[d,"\varphi"] & A\ar[d]\\
+		k\ar[r] & A\otimes_\varphi k
+	\end{tikzcd}
+\end{figure}
+which is the twist of $A$ by the automorphism $\varphi$. Here the scalars are controlled by $\varphi$ since 
+\begin{align*}
+	\lambda(a\otimes \mu)&= a\otimes \lambda\mu\\
+	&=\varphi^{-1}(\lambda)a\otimes \mu
+\end{align*}
+whenever this all makes sense.
+
+Now letting $\varphi(x)=x^p$ we get the Frobenius twist $A\otimes_\varphi k\eqdef A^{(1)}$. Whenever $k$ is perfect, this means that if $A$ is an 
+abelian group, so is $A^{(1)}$.
+
+We always have the map $a^{(1)}\eqdef a\otimes 1\to a^p$ end extending linearly to the map 
+\[a\otimes\lambda=\lambda^{1/p}a\otimes 1\mapsto \lambda a^p.\]
+\begin{rmk}
+	While this gives us a map $A^{(1)}\to A$, we don't ge a map in the other direction or $M^{(1)}\to M$ for a vector space $M$.
+	But we do get a map $M^{(1)}\to S^p(M)$.
+\end{rmk}
+\begin{rmk}
+	Let $X/k$ be a scheme. Then we can define $X^{(1)}/k$. If $X=G$ is a group scheme, then if $A=k[G]$,
+	we can define $G^{(1)}$ to be the group scheme such that 
+	\[k[G^{(1)}]=k[G]^{(1)}\]
+\end{rmk}
+
+You can read more about these in Jantzen, but the map $A^{(1)}\to A$ gives us a map $G\to G^{(1)}$
+and the kernel of this map is the Frobenius kernel.
+
+\subsection{Representation theory}
+\begin{rmk}
+	If $M\in\Rep G$, then this induces a representation of $\Lie G$. Why is this true? Well we can use the functoriality 
+	of $\Lie$! In particular, a representation $M$ of $G$ is equipped with a morphism $G\to \GL(M)$, yielding a lie algebra morphism 
+	\[\Lie G\to \Lie GL(M)=\gl(M)\]
+\end{rmk}
+
+The Frobenius twist also gives us maps on the level of representations. That is, we get a functor 
+\[(-)^{(1)}:\Rep G\to \Rep G\]
+More clearly, if $M\in\Rep G$ we have this vector space as well as a coaction 
+\[k[G]\to M\otimes k[G].\]
+This gives us a map 
+\[M^{(1)}\to M^{(1)}\otimes k[G]^{(1)}\to M^{(1)}\otimes k[G]\]
+so $M\in \operatorname{comod}k[G]\simeq \Rep G$.
+
+If we have two representations $M,N\in\Rep G$, then the map 
+\[\Hom(M,N)\to \Hom(M^{(1)},N^{(1)})\]
+Considering $\End(M)\to \End(M^{(1)})$, this gives us a map $\GL(M)\to\GL(M^{(1)})$.
+
+\begin{rmk}
+	The functor $(-)^{(1)}$ is what is called a ``strict polynomial functor''. :)
+\end{rmk}
+
+\begin{ex}
+	${\Ga}_{(1)}\simeq \alpha_p$ is the group represented by $k[t]/t^p$ with $t$ primitive. ${\Gm}_{(1)}$
+	is the same algebra with $t$ grouplike.
+\end{ex}
+\begin{ex}
+	We have already seen the coordinate algebra of $\GL_n$. What is the Frobenius twist? It is taking a matrix 
+	where we take all entries to the $p^{th}$ power (not the whole matrix!). So then the kernel is 
+	\[{\GL_n}_{(1)}=\{(a_{ij})|a_{ij}^p=\delta_{ij}\}\]
+
+	If you try to evaluate at a field, you won't get much: only one point. Thus this is a highly-singluar connected, finite 
+	group scheme with coordinate algebra 
+	\[k[x_{ij}]/(x_{ij}^p-\delta_{ij})\]
+
+	This gives you the ``infinitesimal neighborhood of 1''.
+\end{ex}
+
+\subsection{Restricted Lie algebra}
+\begin{rmk}
+	We never mentioned what happens for $\gl_n$! Here we can honestly multiply matrices, so we can define $A^{[p]}=A^p$. This tells you that for 
+	matrix groups we can realize their power operation in this way, but of course this is undesirable as it is dependent on embedding.
+\end{rmk}
+
+Recall that the universal enveloping algebra of a Lie algebra $\frakg$ is 
+\[\calU(\frakg)=T(\frakg)/\langle x\otimes y-y\otimes x-[x,y]| x,y\in\frakg)\]
+\begin{defn}
+	If $\frakg$ is a restricted Lie algebra (say $\Lie G$), then the \textbf{restricted Lie algebra} is 
+	\[u(\frakg)=\calU(\frakg)/\langle x^p-x^{[p]}|x\in\frakg)\]
+\end{defn}
+We have PBW bases for both $\calU(\frakg)$ and $u(\frakg)$. The latter is better because if $x_1,\dots,x_n$ are a basis for $\frakg$,
+the basis is 
+\[\{x_1^{a_1}\cdots x_n^{a_n}|a_i\le p-1\}.\]
+
+One can show that the ideal generated by $x^p-x^{[p]}$ is a Hopf ideal, so one gets that $u(\frakg)$ is again a cocommutative Hopf algebra, 
+but this time it is finite dimensional! Then one can consider just the $p$-restricted representations of $\frakg$ and one 
+finds that this category is equivalent to $\lmod {u(\frakg)}$.
+\begin{ex}
+	If $\frakg_m=\Lie\Gm$, we can compute 
+	\[u(\frakg_m)=k[t]/(t^p-1)\]
+	but now $t$ is primitive! In fact, $u(\frakg_m)=k[\mu_p]^\ast=k[\bbZ/p]$.
+\end{ex}
+The above example highlights an important result: if $\frakg=\Lie G$,
+\[u(\frakg)\xrightarrow{\sim}k[G_{(1)}]^\ast\]
+and so we get $\Rep\frakg\simeq \Rep G_{(1)}$. Recall in manifolds that the theories are almost of Lie groups and algebras 
+are very tightly connected (for connected compact manifolds). Here we don't get that! However we get a shadow of it:
+
+Note that we actually get a nested sequence of Frobenius kernels:
+\[G\supset G_{(n)}\supset\cdots\supset G_{(1)}\]
+meaning that in this way we get better and better approximations of $G$.
+
+\section{February 7th, 2020}
+Today we are going to talk about induction and restriction.
+\subsection{The fixed points functor}
+Recall the \textit{fixed point functor}: let $G$ be an algebraic group scheme and $M\in\Rep G$ (which we identify with $\lmod G$).
+Given an action of $G$ on $M$, we get 
+\[M^G=\{m\in M|\forall R, g\in G(R), g\cdot(m\otimes 1_R)=m\otimes 1_R\}\]
+This definition becomes difficult because it requires an extension of scalars and when you have a lot of nilpotents it messes with things.
+
+An alternative method: define 
+\[M^G=\{m\in M|\Delta_M(m)=m\otimes 1\}\]
+where $\Delta_M:M\to M\otimes k[G]$ is the coaction map. We can also write 
+\[M^G=\ker(\Delta_M-\id_M\otimes 1).\]
+
+This does indeed create a functor $(-)^G:\Rep G\to \Vectk$.
+\begin{prop}
+	$(-)^G$ is left exact.
+\end{prop}
+\begin{rmk}
+	$\Rep G$ has an internal Hom. That is, given $M,M'\in\Rep G$, we have an action of $G$ on $\Hom_k(M,M')$ (note: note $G$-equivariant maps!) via the diagonal (in the usual sense of 
+	Hopf algebras). All of this, together with tensor-(internal) hom adjunction
+	\[\Hom_A(M\otimes_k N, M')\cong\Hom_G(M,,\Hom_k(N,M')),\]
+	gives the category of Hopf modules the structure of a rigid monoidal category.
+
+	But now (as vector spaces), we have the isomorphism 
+	\[\Hom_k(M,M')^A\cong \Hom_A(M,M').\]
+\end{rmk}
+
+Let's take a moment to consider $k[G]$ as a $G$-module. The \textbf{right regular representation} on $k[G]$
+is as follows: given $f\in k[G]$, we have 
+\[\rho_r(g)(f)(x)=(g\cdot f)(x)=f(x\cdot g)\]
+and 
+\[\rho_l(g)(f)(x)=(g\ast f)(x)=f(g^{-1}\cdot x)\]
+and by composing the two, we get $\rho_r\circ\rho_l=\rho_l\circ\rho_g$ to get the adjoint representation on $k[G]$.
+
+Now as comodules, we have the action 
+\[\Delta:k[G]\to k[G]\otimes k[G]\]
+which is right regular representation. \textbf{Why?} We can define the left regular representation by 
+\[\tau\circ(S\otimes \id)\circ\Delta.\]
+As an exercise, determine whether this is the same element as $(\id\otimes S)\circ \Delta$.
+
+\subsection{Induction and restriction}
+Given $H\hookrightarrow G$, this gives us an obvious functor 
+\[\Res_H^G:\Rep G\to \Rep H\]
+Then we claim (but don't prove here) that $\Res$ has a right adjoint. In some cases, there is a left adjoint as well although 
+we don't have that in general.
+
+In this case, with groups, we can define 
+\[\Ind_H^G M\eqdef (M\otimes_kk[G])^H\]
+where each are $H$-modules ($k[g]$ with the right regular action) and the action on the tensor is via the left regular action.
+
+Another point of view: remember we can define the affine group scheme $M_a$ where $M_a(X)=M\otimes X$. Then 
+\[M\otimes k[G]=M_a(k[G])\]
+and then we can look at the morphisms $G\to M_a$ and consider 
+$(M\otimes k[G])^H$ to be the \textit{$H$-invariant functions $f:G\to M$} where $H$ actions on $\Mor(G,M_a)$ via 
+\[(h\cdot f)(g)=hf(gh)\]
+and so 
+\[\Mor(G,M_a)^H=\{f:G\to M|f(gh)=h^{-1}f(g),\forall R, g\in G(R)\}.\]
+
+\begin{rmk}
+	Notice that 
+	\[\Ind_H^G=(-)^H\circ(-\otimes k[G])\]
+	and so (since tensor is exact and restriction is left exact), $\Ind_H^G$ is left exact.
+\end{rmk}
+
+\subsection{Frobenius reciprocity}
+Define the evaluation map $\varepsilon_M:\Ind_H^G M\to M$ via 
+\[m\otimes f\mapsto \varepsilon(f)m\]
+Using the other form, if we have $f\in\Mor(G,M)^H$, we get $f\mapsto f(1)\in M$.
+
+\begin{prop}\label{prop:frobenius}
+	\begin{enumerate}
+		\item $\varepsilon_M:\Ind_H^G M\to M$ is an $H$-module map.
+		\item (\textit{Frobenius reciprocity}) $\varepsilon_M$ induces an isomorphism 
+		\[\Hom_G(N,\Ind_H^G M)\xrightarrow{\sim}\Hom_H(\Res_H^G N, M).\]
+	\end{enumerate}
+\end{prop}
+\begin{cor}
+	\begin{enumerate}
+		\item $\Ind$ is transitive: $\Ind_H^G=\Ind_{H'}^G\circ\Ind_H^{H'}$.
+		\item $\Ind$ commutes with field extensions.
+		\item $\Ind$ preserves injectives.
+	\end{enumerate}
+\end{cor}
+\begin{rmk}
+	Recall the proof that $\lmod R$ has enough projectives: the idea was to construct projective in $\Ab$ and then 
+	use a kind of induction to bring them up to the module category. In general, property (c) above empowers us to show 
+	that categories have enough injectives!
+\end{rmk}
+We have a week off! So here's some homework:
+\begin{prob}
+	Prove proposition \ref{prop:frobenius}. Notice if we have 
+	\[N\xrightarrow{f}\Ind_H^G\xrightarrow{\varepsilon_M} M\]
+	and define $\tilde\Psi^H:N\to \Ind_H^G M$ via the $H$-map $\psi:N\to M$ where 
+	\[\tilde\Psi:N\to \Mor(G,M)\]
+	via $n\mapsto \tilde\Psi_n:G\to M$ where $\tilde\Psi_n(g)=\Psi(g^{-1}n)$. Then show $\tilde\Psi_n$ is $H$-invariant, a $G$-map, and prove there are mutual inverses.
+\end{prob}
+
 \end{document}

### 2020-02-14 Algebraic Geometry/Algebraic-Geometry.tex
Update

@@
 \end{figure}
-\begin{thm}
+\begin{thm}\label{thm:fillings}
 	\begin{enumerate}
@@
 
+\section{February 14th, 2020}
+Max recommends probelm 4.10 from Hartshorne. It is the proof of Chow's lemma. It is not an easy proof, 
+but it is a very important one.
+
+\subsection{More generization and specialization}
+
+\begin{lem}
+	Given a quasicompact morphism $f:X\to Y$, $f(X)$ is closed in $Y$ if and only if it is \textit{closed under specialization}.
+\end{lem}
+\begin{prf}
+	The conclistion is stable udner localiation on $U_i$, so we may assume $Y=\Spec A$. Then since $f$ is quasicompact, 
+	$X=\cup_1^n\Spec B_i$ where $\Spec B_i=V_i$. But then 
+	\[\Spec\prod B_i=\sqcup V_i\to X\to Y\]
+	has the same image. Therefore we can assume that $X=\Spec B$ is affine.
+
+	So we have a morphism $\Spec B\to \Spec A$, giving us a ring map $\varphi:A\to B$. So now if $I=\ker\varphi$,
+	we get that the map $X\to Y$ factors through $\Spec A/I\hookrightarrow Y$, which is closed. So we can assume that $\varphi$ is injective.
+	We want to show that $f(X)=Y$. \textit{Notice that we haven't checked that the assumption is closed under these reductions.}
+
+	Now take any $p\in \Spec A$ and take a minimal prime $p_0\subseteq p$ (which exists by the standard Zorn argument).
+	So now $p_0\rightsquigarrow p$. If $p_0\in f(X)$, then $p\in f(X),$ so $f(X)=Y$. Consider $A\hookrightarrow B$ and localize at $p_0$ (which preserves injectivity).
+	and since $A_{p_0}$ has a single prime, you get a map with $p_0$ in the image. Thus you get $p$ and surjectivity.
+\end{prf}
+
+Now we are going to prove:
+\begin{prf}[of thm.~\ref{thm:fillings}(b)]
+	Assume that $f:X\to Y$ is universally closed. We want to show that fillings exist for any test diagram. Notice that fillings for test diagrams are equivalent 
+	to finding sections (since we can assume $Y=\Spec R$):
+	\begin{figure}[h]
+		\centering
+		\begin{tikzcd}
+			\Spec k\ar[r,"\iota"]\ar[d] & X=X\times_{\Spec R}\Spec R\ar[d]\\
+			\Spec R\ar[r,"\id"] & \Spec R\ar[u,bend right,dashed]
+		\end{tikzcd}
+	\end{figure}
+	Let $x\in X$ be in the image of $\iota(\Spec k)$ and let $k$ be the resitude field of $x$. Let $z=\overline{\{x\}}\subseteq X$
+	Now $(0)\subseteq f(z)\subset\Spec R$ is closed, so there exists $z\in Z$ such that $f(z)=\frakm_R$. I got lost at the end, but there were some diagrams that 
+	boiled down to our ring being a valuation ring.
+
+	Now assume that we have our fillings. Using the lemma, we can show that the image is cloed under specialization (warning: implicit base change being hidden here. If 
+	we have $f:X\to Y$, $X_T\to T$).
+\end{prf}
+
 \end{document}

### 2020-02-19 Algebraic Geometry/Algebraic-Geometry.tex
bump

@@
 
+\section{February 19th}
+Today Christine is talking!
+\begin{lem}[Affine Communication Lemma]
+	Let $\scrP$ be a property of some affines of a scheme $X$ such that
+	\begin{enumerate}
+		\item If $\Spec A\hookrightarrow X$ has $\scrP$ then for all $f\in A$, $\Spec A_f\hookrightarrow X$ has $\scrP$.
+		\item If $(f_1,\dots,f_n)=A$ and $\Spec A_f\hookrightarrow X$ have $\scrP$ then $\Spec A\hookrightarrow X$ has $\scrP$.
+	\end{enumerate}
+	Further if $X=\cup_{i\in\scrI}\Spec A_i$ and the $\Spec A\hookrightarrow X$ have $\scrP$, then so does $X$.
+\end{lem}
+
 \end{document}

### 2020-02-19 Algebraic Groups/Algebraic-Groups.tex
bump

@@
 
+\section{February 19th, 2020}
+Welcome back! Today we're taking about the proof for that last problem (Frob. Reciprocity)
+\begin{prf}
+	Recall we have the map $\varepsilon_M:\Ind_H^G M\to M$, where we think of 
+	\[\Ind_H^G M=(M\otimes k[G])^H=\{f:G\to M|h^{-1}f(-)=f(-\cdot h), gf(-)=f(g^{-1}\cdot -)\}\]
+	and $\varepsilon_M(f)=f(1)\in M$. Then the map giving us our Frobenius reciprocity is just composing with $\varepsilon_M$.
+
+	We will prove some of this. Let $\tilde\psi:N\to \Ind_H^G M$. Thus it corresponds, for each $n\in N$, to a map 
+	\[\tilde\psi_n:G\to M\]
+	and we need to define this map, check it is $H$-invariant, check that the resulting $\tilde\psi$ is $G$-invariant, and then check that this and $(\varepsilon_M\circ -)$ and this map are mutual inverses.
+
+	Then we busted out paper and wrote stuff down. I checked $G$ invariance.
+\end{prf}
+\begin{cor}
+	\begin{enumerate}
+		\item $\Ind$ is transitive.
+		\item $\Ind$ takes injectives to injectives.
+	\end{enumerate}
+\end{cor}
+The way to prove this is to notice that it is a right adjoint to an exact functor.
+
+\subsection{Projection formula}
+\begin{thm}[Tensor identity/Projection formula]
+	If $M\in\lmod H$ and $N\in\lmod G$,
+	\[\Ind_H^G(M\otimes N)\simeq (\Ind_H^GM)\otimes N\]
+\end{thm}
+\begin{rmk}
+	If $k$ is a ring instead of a field, we need to make an additional assumption: that $N$ is flat over $k$.
+\end{rmk}
+\begin{prf}
+	We'll write down the maps at least. Since 
+	\[L\eqdef \Ind_H^G(M\otimes N)=(M\otimes N\otimes k[G])^H\simeq\{f:G\to M\otimes N|f(gh)=(h^{-1}\otimes h^{-1})f(g)\}\]
+	on the right, we have 
+	\[R\eqdef\Ind_H^GM\otimes N=(M\otimes k[G])^H\otimes N\simeq\{f:G\to M\otimes N|f(xh)=(h^{-1}\otimes 1)f(x)\}\]
+
+	Now we can write down two maps 
+	\[\alpha,\beta:M\otimes N\otimes k[G]\to M\otimes N\otimes k[G]\]
+	(thinking of these as maps $G\to M\otimes N$) in the following way:
+	\[\alpha(f)(g)=(1\otimes g)f(g)\quad\text{and}\quad \beta(f)(g)(1\otimes g^{-1}f(g))\]
+	and we claim:
+	\begin{enumerate}
+		\item $\alpha$ and $\beta$ are mutual inverses;
+		\item $\alpha(L)=R$ and $\beta(R)=L$;
+		\item $\alpha,\beta$ are $G$-invariant.
+	\end{enumerate}
+\end{prf}
+\begin{rmk}
+	This is a bit tedious, but this whole induction business is spelled out in Jantzen.
+\end{rmk}
+
+\subsection{What do we gain from all this?}
+Well first, notice that $\Ind_e^G k=k[G]$, which implies taht $k[G]$ is injective. Then we have 
+\begin{cor}
+	Let $M\in \lmod k$ and notice 
+	\[\Ind_e^G M\simeq\Ind_e^G(k\otimes k M)\simeq k[G]\otimes M_\text{triv}\]
+	which is just a sum of copies of $k[G]$, which is again projective.
+\end{cor}
+\begin{cor}
+	Let $M\in \Rep G$. Then 
+	\[M\otimes k[G]\simeq M_\text{triv}\otimes k[G]\]
+\end{cor}
+\begin{prf}
+	$M_\text{triv}\times k[G]\simeq\Ind_e^G M=\Ind_e^G(k\otimes M)\simeq \Ind_e^Gk\otimes M\simeq k[G]\otimes M.$
+\end{prf}
+\begin{rmk}
+	The map between these things is nontrivial (we used some twisting in the tensor identity, for example). The map is 
+	\[m\otimes f\mapsto (1\otimes f)(1\otimes S)\Delta_M(m)\]
+	where $S$ is the coinverse.
+\end{rmk}
+
+\begin{thm}[``$\Rep G$ has enough injectives'']
+	\begin{enumerate}
+		\item $k[G]$ is injective.
+		\item $\forall M\in\Rep G$, there is an injective $I$ with $M\hookrightarrow I$.
+		\item Any injetive $I$ is a direct summand of $\bigoplus k[G]$.
+	\end{enumerate}
+\end{thm}
+Corollary 2 gets us our injectives containing $M$. This sets us up to do homological algebra, which we will do next time.
+
 \end{document}

### 2020-02-28 Algebraic Geometry/Algebraic-Geometry.tex
First update in a while

@@
 
+\section{February 21, 2020}
+Today we are going to talk about sheaves of modules. Let $(X,\O_X)$ be a ringed space. Then 
+\begin{defn}
+	An $\O_X$-module (``sheaf of $\O_X$ modules'') is a sheaf of abelian groups $\calF$ along with a morphism 
+	\[\O_X\to \iHom(\calF,\calF)\]
+	of sheaves of rings. Equivalently it is an action 
+	\[\O_X\times\calF\to \calF\]
+	or (locally) $\calF(U)$ is an $\O_X(U)$-module and the ``experience is functorial.''
+
+	A homomorphism of $\O_X$-modules is a morphism $\varphi:\calF\to\calG$ of abelian sheaves that is compatible with the $\O_X$-action.
+\end{defn}
+\begin{lem}
+	The category $\lmod{\O_X}$ of $\O_X$ modules is an abelian category.
+\end{lem}
+
+\section{February 24, 2020}
+We talked about the sheaf of sections of a vector bundle on a manifold. I didn't write it down but it is a good thought.
+Here are some
+
+\subsection{Moar sheavz}
+Let $X=\Spec A$ and let $M\in \lmod A$. We create a sheaf of modules $\tilde M$ by saying $\tilde M(D(f))=M_f$. 
+The right thing to do is $B=\prod A_{f_i}$. Then max did a bunch of shit. The idea is that this is almost exactly the 
+same construction that you use to create the structure sheaf. There is some cochain stuff going on here that is quite cool.
+
+So the functor $\tilde -:\lmod A\to \lmod{\O_X}$ is a functor and is exact, commutes with $\otimes,$ and satisfies $\Gamma(\Spec A,\tilde M)=M$ and 
+for all $p\in X$, $\tilde M_p=M_p$. It is also fully faithful but \textbf{not essentially surjective}.
+
+To show this last idea, consider the space with two points: $R= k[[t]]$ and let $U\subset\Spec R$ be the open point. THen define 
+a $\O_{K(R)}$ module on the localization $K(R)$ at $U$. Just let this sheaf be $\tilde K(R)$. But then consider the 
+\[\Gamma(\Spec R,\tilde K(R))=0\quad\text{and}\quad\Gamma(K(R),\tilde K(R))=K(R)\]
+which doesn't arise as the ``sheafification'' (DON'T read too much into this word although it may kinda be true) of an $R$-module.
+
+\subsection{(Quasi-)Coherent sheaves}
+Hey, don't read Hartshorne for coherent sheaves, apparently.
+
+\begin{defn}
+	Let $X$ be a scheme. Then an $\O_X$ modules $\scrF$ is \textbf{quasicoherent} if there exists an open affine cover $U_i=\Spec A_i$ and modules $M_i\in\lmod{A_i}$
+	such that for all $i$, $\scrF|_{U_i}=\tilde M_i$.
+\end{defn}
+Better:
+\begin{defn}
+	$\scrF$ is quasicoherent if and only if there exists an open (not necessarily affine!) cover $U_i\subseteq X$ such that there exists a presentation 
+	\[\O_{U_i}^{\oplus I}\to \O_{U_i}^{\oplus J}\to \scrF|_{U_i}\to 0\]
+	for all $i$ and for some index sets $I$ and $J$.
+\end{defn}
+
+The right thing to read is FAC by Serre where he does this whole thing (in French).
+
+\subsection{What does coherence mean?}
+Quasicoherence gives us $\calO^I\to\calO^J\to \scrF\to 0$ is exact. We can say that $\scrF$ is quasicoherent finitely genreated if $J$ can be taken 
+to be finite. It is quasicoherent finitely presented if both $I$ and $J$ can be taken to be finite.
+
+A sheaf is \textbf{coherent} if for ANY finite $J$ and surjection $\O^J\to \scrF$, $\ker(\pi)$ is finitely generated.
+In the Noetherian case, all of these coincide.
+
+WARNING!
+Coherence is not stable under pullback.
+So what kind of perations can we do?
+IF we have $f:X\to Y$ is a map of ringed spaces, we get the pullback $f_\ast:\O_X$-mod$\to\O_Y$-mod and we want to have a left
+adjoint $f^\ast$. Recal that we have $f^\sharp:\O_Y\to f_\ast\O_X$, where the codomain acts on $f_\ast\scrF$.
+
+How do we define the adjoint? $f^\sharp$. Wow I just stopped in the middle of a thought.
+
+\section{February 26th, 2020}
+We had a functorial experience. We showed that these adjoint maps $f_\ast,f^\ast$ exist on the level of (abelian)
+sheaves on $X$ and $Y$ and also they exist (THOUGH ARE DIFFERENT) when we're talking about $\O_X$ and $\O_Y$ modules. 
+These also restrict to maps on the quasicoherent sheaves (and finitely presented ones) but not, in general, to the coherent sheaves.
+
+\begin{thm}
+	The map $\tilde -:\lmod{A}\to \mathbf{QCoh}_{\Spec A}$ is an equivalence of categories.
+\end{thm}
+\begin{cor}
+	If $X$ is quasi-compact and quasi-separated, then for all $f\in \Gamma(X,\O)$ and $s\in \scrF(D(f))$, and $\scrF\in \mathbf{QCoh}_X$,
+	there exists a positive integer $n$ such that $f^ns$ extends to a global section $\sigma\in\Gamma(X,\scrF)$. In other words, 
+	all poles of $s$ have finite order.
+\end{cor}
+
+\subsection{Next time...}
+\begin{prop}
+	Let $f:X\to Y$ be a morphism of schemes. Then 
+	\begin{enumerate}
+		\item $f^\ast\mathbf{QCoh}_Y\subseteq\mathbf{QCoh}_X$ (we already showed this).
+		\item If $f$ is quasi-compact, then $f_\ast\mathbf{QCoh}_X\subseteq\mathbf{QCoh}_Y$.
+	\end{enumerate}
+\end{prop}
+
 \end{document}

### 2020-02-28 Algebraic Groups/Algebraic-Groups.tex
First update in a while

@@
 
+\section{February 21st, 2020}
+We did a lot of work last time to show that $k[G]$ was injective and furthermore that every module injects into 
+an injective module. In general, finding projectives is much harder.
+
+\subsection{Projectives}
+\begin{lem}
+	Let $E$ be a simple $G$-module. Then there exists an injective hull $Q_E$ such that 
+	\[E\hookrightarrow Q_E\]
+	with $\operatorname{soc}Q_E=E$ and $Q_E$ is indecomposable.
+\end{lem}
+\begin{rmk}
+	Any injective in $\Rep G$ is a direct sum of indecomposables.
+\end{rmk}
+\begin{lem}
+	Let $P$ be a projective $G$-module. Then $P$ is injective.
+\end{lem}
+\begin{prf}
+	Let $P$ be a projective module. We claim that for any $G$-module $M$, $P\otimes_k M$ is again projective. This follows from tensor-hom adjointness.
+	So $\Hom(-,P)$ is exact. Let 
+	\[0\to M_1\to M\to M_2\to 0\]
+	be a short exact sequence of finite dimensional modules. Dualizing is exact, as is $-\otimes P$, so 
+	\[0\leftarrow M_1^\ast\otimes P\leftarrow M^\ast\otimes P\leftarrow M_2\otimes P\leftarrow 0\]
+	is exact. Thus $M\otimes P$ splits as $(M_1\otimes P)\oplus M'$.
+
+	But then notice that 
+	\[\Hom_G(M_1,P)\simeq(\Hom_k(M,P))^G\simeq(M_1^\ast\otimes P)^G\]
+	but then $(M^\ast\otimes P)^G$ splits as 
+	\[(M_1^\ast\otimes P)^G\oplus M'^G\]
+	and this proves surjectivity.
+\end{prf}
+\begin{rmk}
+	Notice that this alst proof could possibly be better. The finite dimensionality part could be able to be weakened/dropped by
+	avoiding the tensor product representation.
+\end{rmk}
+
+\subsection{Frobenius categories}
+The upshot to all of this is that 
+\begin{prop}
+	If there exists a projective module $P$ in $\Rep G$, then the collections of injectives and projectives are the same.
+\end{prop}
+\begin{prf}
+	Let $P$ be projective. Let $P_1\subseteq P$ be a finite-dimensional submodule. Then 
+	\[M\hookrightarrow M\otimes \End_k(P_1)\simeq M\otimes P_1^\ast\otimes P_1\hookrightarrow M\otimes P_1^\ast\otimes P\]
+	which we have seen is both injective and projective.
+
+	So take a simple $M$ with injective hull $Q_M$. Then we get maps $Q_M\to M\otimes P_1^\ast\otimes P$ and $M\otimes P_1^\ast\otimes P\to Q_M$ 
+	by leveraging injectivity and projectivity. Thus $Q_M$ is a direct summand of $M\otimes P_1^\ast\otimes P$, which is projective. 
+	A little clean up gets us the result.
+\end{prf}
+
+\begin{rmk}
+	When $P$ is injective iff projective, the category $\Rep G$ is a Frobenius category.
+\end{rmk}
+
+\begin{rmk}
+	FOr smooth algebraic groups $\Gm$, $\Ga$, and $\GL_n$, there are no projective modules.
+	If $G$ is finite (finite groups, Frob kernels ${\Ga}_{(r)}$, $\mu_p$, ${\GL_n}_{(r)}$) we have that the collections of projective and injectives are the same.
+\end{rmk}
+\begin{rmk}
+Domkin put together a complete list of affine group schemes with projective modules.
+\end{rmk}
+
+\subsection{Cohomology}
+We know now that $\Rep G$ has enough injectives, so if $\calF:\Rep G\to \Vectk$ is left exact we can take an 
+injective resolution $M\to I^\bullet$ and then 
+\[(\calR^n\calF)(M)\eqdef H^n(\calF(I^\bullet)).\]
+
+Recall the Grothendieck spectral sequence:  Let $\calF:\calC\to \calD$ and $\calF:\calD\to \calE$. Then say 
+we want to compute the derived functors of $\calF'\circ\calF$. Assume that $\calF'$ is left exact and that 
+$\calF$ takes injectives to $\calF'$ acyclic objects. Then we have a spectral sequence whose second page is 
+\[E_2^{m,n}=\calR^m\calF'\circ\calR^n\calF\Rightarrow \calF^{m+n}(\calF'\circ\calF)\]
+
+\begin{rmk}
+	Notice that in the Grothendieck spectral sequence if $\calF'$ is exact (so $\calR^n\calF=0$ when $n>0$),
+	you get 
+	\[\calF'\circ\calR^n\calF\xrightarrow{\sim}\calR^n(\calF'\circ\calF)\]
+
+	You can do the same thing with exactness of $\calF$.
+\end{rmk}
+
+\subsubsection{Notation}
+We will use the following notation:
+\begin{align*}
+	\Ext_G^n(M,-)=\calR^n(\Hom_G(M,-))\\
+	H^n(G,M)=\Ext_G^n(k,M)\\
+	H^n(G,k)=\Ext^n_G(k,k)\\
+	H^\ast(G,k)=\oplus_0^\infty\Ext_G^n(k,k)
+\end{align*}
+and this last object is called the cohomology ring of $G$.
+
+\subsubsection{Applications}
+\begin{prop}
+	We have the derived adjointness 
+	\[\Ext_G^n(M\otimes V,N)\simeq \Ext_G^n(M,\Hom(V,N))\]
+\end{prop}
+
+\section{February 24th, 2020}
+Today we are going to talk more about the Grothendieck spectral sequence.
+\subsection{Examples}
+Let's begin by proving 35.5.2
+\begin{prf}
+	Let $\calF(-)=M\otimes -$ and $\calF'(-)=\Hom_G(-,N)$. Then 
+	\[\calF'\circ\calF(V)=\Hom(M\otimes V,N)\]
+	and so $\scrR^n(\calF')\circ \calF(V)=\Ext^n(M\otimes V,N)$
+	but we know that this is isomorphic to $\scrR^n(\calF'\circ\calF)(V)$,
+
+	I need to work this out once to understand the details. It involved using the regular tensor-hom identity and 
+	the convergence of the Grothendieck spectral sequence again.
+\end{prf}
+
+\begin{prop}
+	\[\Ext_G^n(M,\scrR^m\Ind_H^G N)\Rightarrow \Ext_H^{n+m}(m,N)\]
+\end{prop}
+\begin{prf}
+	Leverage Frobenius reciprocity and use the fact $\scrF=\Ind_H^G$ is left exact and takes injectives to injectives. Then if we let $\scrF'=\Hom_G(M,-)$,
+	we get $\scrF'\circ\scrF(N)=\Hom_H(M,N)$ and it pops out.
+\end{prf}
+
+\begin{defn}
+	We say a subgroup scheme $H\le G$ is \textbf{exact in $G$} if $\Ind_H^G$ is exact. Equivalently if 
+	$\scrR^{n>0}\Ind_H^G=0$.
+\end{defn}
+\begin{cor}
+	If $H$ is exact, we have an isomorphism 
+	\[\Ext_G^n(M,\Ind_H^G N)\xrightarrow{\sim}\Ext_H^n(M,N)\]
+\end{cor}
+This is, of course, a much easier thing to work with. But an interesting question is when $H$ is exact.
+\begin{prop}
+	Let $k$ be a field. Then $H$ is exact if and only if $k[G]$ is an injective $H$ module.
+\end{prop}
+\begin{rmk}
+	The direction ``$\Leftarrow$'' works when $k$ is a commutative ring.
+\end{rmk}
+\begin{rmk}
+	$k[G]=\Ind_1^G k=\Ind_H^G(\Ind_1^H k)=\Ind_H^G(k[H])$
+\end{rmk}
+\begin{prf}
+	Take a short exact sequence $0\to M_1\to M\to M_2\to 0$ of $H$-modules. Then we do the old argument:
+	take a splitting of $M\otimes k[G]$ and we get that $M_1\otimes k[G]$ is a summand of $M\otimes k[G]$.
+	Then applying invariants preserves this decomposition.
+
+	The rest of this (and all of these properties) live in Jantzen chapter 1 section 4.
+\end{prf}
+
+\begin{rmk}
+	If $H$ is a finite groups scheme, then $H$ is exact (of everything)! The motivation for this: say we have $H\le G$ 
+	any two affine group schemes.
+
+	In analogy, there is a statement about Hopf algebras (see my notes for that). If $A\hookrightarrow B$ is a map of two Hopf algebras (perhaps if $B$ is cosemisimple?)
+	This idea ports over here somehow. It can be proved via sheaf cohomology.
+\end{rmk}
+
+Next we're going to try to write down the ``generalized version'' of the tensor identity using the Grothendieck spectral sequence.
+\[(\scrR^n\Ind_H^G)\circ(-\otimes \Res^G_H N)\cong (\scrR^n\Ind_H^G(-))\otimes N\]
+
+\section{February 26th, 2020}
+This is all from Jantzen chapter I section 5. Apparently it is a bit hand-wavy.
+\subsection{Associated sheaves}
+Our goal is to describe $\scrR^n\Ind_H^G(M)$ as $H^n(G/H,\calL(M))$ where $\calL(M)$ is a sheaf on $G/H$.
+We are going to think of $\scrL:\Rep H\to \Sh G/H$ as a functor.
+
+\subsubsection{Quotients}
+The history of these quotients go back to Demazure and Gabriel. The idea is that we don't genreally have 
+quotients in $\Sch$, so we extend to the \textbf{sheaves in the fppf topology}. This is a Grothendieck topology that stands for ``
+finitely presented faithfully flat''. It's French, don't question it.
+
+When we have an algebraic group $G$ acting on a scheme $X$, we want to know when the quotient $X/G$ in this broader context 
+is actually a scheme.
+
+Let us assume that $G$ acts freely (pointwise). We assume that $k$ is a field. We can drop this, but then we have 
+to carry around the assumption of flatness everywhere. Let $X$ be a scheme of finitel type (algebraic) over $k$ and $G$ an 
+affine group scheme (also of finite type). Then we have 
+\begin{prop}
+	If $G$ acts freely and $X/G$ is a scheme, then 
+	\[X(R)/G(R)\hookrightarrow (X/G)(R)\]
+\end{prop}
+\begin{prop}
+	($G$ always acting freely) If $X/G$ is a scheme, there exists an isomorphism 
+	\[X\times G\xrightarrow{\sim}X\times_{X/G}X\quad\text{via}\quad (x,g)\mapsto (x,xg)\]
+	where on the level of points, this is $\{(x,x')|x=x'g\}$ (we are using right actions obviously).
+\end{prop}
+From here on out, we are going to assume that $X/G$ is a scheme unless we say something to the contrary.
+\begin{ex}
+	$\P n(R)$ is, as a set, the set of free rank one direct summands of $R^{n+1}$. Work out how (and if!) this is true.
+	Then we can define 
+	\[\A{n+1}_\ast(R)=\{(a_0,\dots,a_n)\in R^{n+1}|\sum Ra_i=R\}\]
+	which is a subscheme of $\A{n+1}$. Now we claim that 
+	\[\A{n+1}_\ast/\Gm\xrightarrow{\sim}\P{n}\]
+	And we had some trouble making this precise. This is in Jantzen, however.
+\end{ex}
+You can do similar things with the Grassmann schemes $\operatorname{Gr}(k,n)$.
+\begin{thm}
+	Let $G$ be a finite group scheme and $X$ an affine scheme with a free $G$ action. Then $X/G$ is an affine scheme with 
+	coordinate ring $k[X]^G$.
+\end{thm}
+\begin{thm}
+	Let $H\le G$ be algebraic group schemes (sometimes requires that $H$ is closed, but this is automatic when we look at it from the functorial 
+	perspective and require that $H$ is an algebraic subgroup scheme). Then (with $H$ acting on the right of $G$) $G/H$ (or in the other case $H\setminus G$) is a scheme.
+	In particular, $G\times H\simeq G\times_{G/H}G$.
+\end{thm}
+
+\section{February 28th, 2020}
+\subsection{Associated Sheaves}
+The set up is the following: we have a scheme $X/k$, algebraic group scheme $G/k$ acting freely on $X$ such that $X/G$ a scheme.
+We are constructing a functor 
+\[\calL:\Rep G\to \Sh X/G\]
+and actually to $\O_{X/G}$ modules. Let $\pi:X\to X/G$ be affine, flat and algebraic.
+\begin{thm}
+	\begin{enumerate}
+		\item If $Y\subset X/G$ is affine (open), then $\pi^{-1}Y\subset X$ is affine (open). T
+		\item If $Y\subset Y/G$ is affine open, then we get the map $\pi:\pi^{-1}Y\to Y$ induces a $k[Y]$ structure on $k[\pi^{-1}Y]$ such that it is a faithfully flat, finitely presented $k[Y]$ module.
+	\end{enumerate}
+\end{thm}
+
+Now we construct things in the following way. Let $M$ be a $G$-modules. Then for any open $U\subset Y$, define 
+\[\calL(M)(U)=\left(M\otimes_k k[\pi^{-1}U]\right)^G\]
+where we put the diagonal action on the tensor product. Notice $\pi^{-1}(U)\subset X$.
+
+Why is this the right thing to do? Notice that 
+\[M\otimes k[\pi^{-1}U]=\operatorname{Mor}(\pi^{-1}U,M_a)=\{f:\pi^{-1}U\to M_a|f(xg)=g^{-1}f(x)\}.\]
+We claim that $\calL(M)$ is a sheaf, and specifically a sheaf of $\O_{X/G}$-modules. That is, the map $\pi:X\to X/G$ gives us 
+a $\O_{X/G}$ module structure on $\O_X$ via $\pi^{ast}:\pi^{-1}\O_{X/G}\to \O_X$. On any $U$ we have, for 
+$f_1\in \O_{X/G}(U)$ and $f\in\calL(M)(U)$, an element 
+\[f_1\cdot f\in\calL(M)(U).\]
+
+\begin{prop}
+	\begin{enumerate}
+		\item $\calL$ is a sheaf of $\O_{X/G}$ modules.
+		\item $\calL$ is exact.
+		\item $\calL(M)$ is quasi-coherent.
+		\item If $M$ is a finite dimensional $G$-module, $\calL(M)$ is coherent.
+	\end{enumerate}
+\end{prop}
+We may talk about a proof of this next time.
+\begin{ex}
+	Let $M=k$. Then 
+	\[\calL(k)(U)=k[\pi^{-1}U]^G=\{f:\pi^{-1}U\to k_a=\A1|f(xg)=f(x)\}\]
+	so we are just looking for invariant functions. The claim is that this is actually 
+	\[\calL(k)(U)=\{f:\pi^{-1}U/G\to \A1\}=\{f:U\to A1\}=k[U]=\O_{X/G}(U)\]
+\end{ex}
+We want to know what happens to injective modules. All of our injectives are going to look like $M\otimes k[G]$ or direct summands of this or will be 
+otherwise comparable to this. So if we have $\pi:X\to X/G$,
+\[\calL_{X/G}(M\otimes k[G])\simeq \pi_\ast\calL_{X/e}(M)\]
+This is requires some thought and computation.
+
+\begin{prop}
+	\begin{enumerate}
+		\item The functor sending 
+		\[M\mapsto H^0(X/G,\calL(M))=\Gamma(\calL(M))\]
+		is left exact (but not right exact!)
+		\item If $X$ is affine, then the derived functors are 
+		\[M\mapsto H^n(X/G,\calL(M))\]
+	\end{enumerate}
+\end{prop}
+\begin{prf}
+	The second part looks almost obvious from the Grothendieck spectral sequence. We need, however, 
+	that $\calL$ takes injective $G$-modules to acyclic sheaves. Why is that true? Take $M\otimes k[G]$ an 
+	injective $G$-module, then we saw that 
+	\[\calL(M\otimes k[G])\simeq \pi_\ast\calJ_{X/e}(M).\]
+	We need to check that 
+	\[H^n(X/G,\pi_\ast\calL_{X/e}(M))\]
+	vanishes for all $n\ge 1$. THere is a problem in Hartshorne that gives this to us for an affine morphism between separated (something else) schemes.
+\end{prf}
+\begin{prop}
+	Let $H\le G$. Then 
+	\begin{enumerate}
+		\item $\calR^n\Ind_H^G(M)\simeq H^n(G/H,\calL_{G/H}(M))$.
+		\item If $G/H$ is Noetherian, then $\scrR^n\Ind_H^G(M)=0$ for $n>\dim G/H$.
+		\item If $G/H$ is projective, then $\scrR^n\Ind_H^G(M)$ is finitely generated whenever $M$ is finite dimensional.
+		\item If $G/H$ is affine, $\Ind_H^G$ is exact.
+	\end{enumerate}
+\end{prop}
+\begin{prf}
+	For the first statement, We are just taking higher derived functors of 
+	\[H^0(GH,\calL(M))=\calL(M)(G/H)=(M\otimes k[G])^H=\Ind_H^G(M)\]
+	where the isomorphisms here are natural. So the derived functors are the same!
+\end{prf}
+\begin{cor}
+	If $H$ is a finite group scheme, then $\Ind_H^G$ is exact.
+\end{cor}
+\begin{prf}
+	When know that $G/H$ is affine and in fact is $\Spec k[G]^H$.
+\end{prf}
+
 \end{document}

### 2020-03-07 Algebraic Geometry/Algebraic-Geometry.tex
Update because all the classes are ending

@@
 
+\section{March 2nd, 2020}
+Sorry, I missed Friday. Today we are continuing to talk about 
+\subsection{Closed Subschemes}
+We saw last time that there is a bijection between closed subschemes $\iota:Y\hookrightarrow X$ (modulo some relation?) and the collection of quasicoherent sheaves of ideals $\calI\subseteq\O_X$.
+Last time Max showed that it is surjective.
+
+To show injectivity, we go through a few steps: we begin by showing that the topological spaces underlying 
+two closed immersions $Y$ and $Y'$ with an iso $g:\iota_\ast\O_Y\to \iota'_\ast\O_{Y'}$. Then we continued by 
+constructing stuff.
+
+\section{March 4th, 2020}
+Today we are talking about sheaves on proj. The
+
 \end{document}

### 2020-03-07 Algebraic Groups/Algebraic-Groups.tex
Update because all the classes are ending

@@
 \DeclareMathOperator{\SL}{SL}
+\DeclareMathOperator{\Dist}{Dist}
 
@@
 \begin{prf}
-	When know that $G/H$ is affine and in fact is $\Spec k[G]^H$.
-\end{prf}
+	We know that $G/H$ is affine and in fact is $\Spec k[G]^H$.
+\end{prf}
+\section{March 2nd, 2020}
+Some reminders: we reinterpreted our induction functor in terms of sheaf cohomology. In our particular case, 
+we want to restrict to 
+\[\calL:\Rep H\to \O_{G/H}\text{-}\mathbf{mod}\]
+where $H\le G$ and we are sending $M\mapsto \calL(M)$ where now 
+\[\calR^n\Ind_H^G M\simeq H^n(G/H,\calL(M)).\]
+
+\subsection{A Preview}
+When we have a Borel subgroup $B\subseteq G$ (e.g. the upper triangulars in $\GL_n$), 
+then we get the functor 
+\[\Rep B\to \O_{G/B}\text{-}\mathbf{mod}\]
+where we send particular representations $\lambda$ to $H^0(G/B,\calL(k_\lambda))$.
+
+\subsection{Leftovers}
+\begin{prop}
+	$\calL$ is exact.
+\end{prop}
+\begin{prf}
+	It suffices to show that the association of $M$ to $\calL(M)(U)$, for $U$ an affine open, is exact. To 
+	prove this, consider $\tilde U=\pi^{-1}\subset X\to U\subset X/G$, which is faithfully flat---that is, the ring map 
+	$\pi_\ast:k[U]\to k[\tilde U]$ is. So it suffices to show that the map 
+	\[M\mapsto k[\tilde U]\otimes_{k[U]}\calL(M)(U)=k[\tilde U]\otimes_{k[U]}(k[\tilde U]\otimes M)^G\]
+	is exact. Recall that we have that $X\times G\simeq X\times_{X/G} X$ via the map $(g,x)\mapsto (g,gx).$
+	So via the same map, we get $\tilde U\times_U\tilde U\simeq \tilde U\times G$ where $G$ acts on the right factor.
+
+	This gives us an iso 
+	\[k[\tilde U]\otimes_{k[U]}k[\tilde U]\simeq k[\tilde U]\otimes k[G]\]
+	and so 
+	\[k[\tilde U]\otimes_{k[U]}(k[\tilde U]\otimes M)^G=(k[\tilde U]\otimes_{k[U]}k[\tilde U]\otimes M)^G\simeq (k[\tilde U]\otimes k[G]\otimes M)^G\]
+	where in both of the cases below $G$ acts diagonally on the last two tensor factors.
+
+	Then applying the tensor identity, we get that the above is
+	\[(k[U]\otimes k[G]\otimes M_{tr})^G\cong k[\tilde U]\otimes M\otimes (k[G])^G\simeq k[\tilde U]\otimes M\]
+	which is exact. 
+\end{prf}
+The next proposition helps us to show that $\calL$ takes injectives to acyclics so we can apply the Grothendieck spectral sequence.
+\begin{prop}
+	$\calL_{X/G}(M\otimes k[G])\simeq \pi_\ast \calL_{X/e}(M)$.
+\end{prop}
+\begin{prf}
+	It is enough to show that for all $U\subseteq X/G$ open and affine, the two sheaves have the same values. Recall that 
+	\[\calL_{X/G}(M\otimes k[G])(U)\cong\Mor(\pi^{-1}(U),M\otimes k[G])^G\cong\Mor(\pi^{-1}U\times G,M)^G\]
+	where in the last iso is induced from a kind of $\times$-hom adjunction.
+
+	Then one can define a map on the level of varieties $(x.g)\mapsto (xg,g)$ from $\pi^{-1}U\times G$ to itself where 
+	now $\pi^{-1}U$ has the trivial action.  Thus we get 
+	\[Mor(\pi^{-1}U_{tr},M)^G\simeq\Mor(\pi^{-1}U,M)=\pi_\ast\calL_{X/e}(M)\]
+\end{prf}
+\subsection{Special Case}
+Say we have $\pi:X\to X/G$ which is locally trivial: for every $U_i\subset X/G$, there is a local section $\sigma_i:U_i\to X$ with 
+the properties:
+\begin{itemize}
+	\item $\cup U_i=X/G$
+	\item $\pi\circ\sigma_i=\id_{U_i}$
+	\item $\pi^{-1}(U_i)=U_i\times G$
+\end{itemize}
+A thing we should discuss is the grassmannian. Let's work over a field $k$. Recall that $\operatorname{Gr}(r,n)$ (or similar notation) means the 
+$r$ dimensional subspaces in $\A n$. We have a natural action of $\GL_n$ on $\A n$ that induces an action on $\operatorname{Gr}(r,n)$. Then one can show that 
+$\GL_n/\operatorname{stab}_{\GL_n}(v)=\Im\pi\simeq\operatorname{Gr}(r,n).$ The stabilizer is a parabolic subgroup. We hope to see more of this 
+on Wednesday.
+
+\subsection{More stuff}
+\begin{thm}
+	If $G$ acts on $X$ and $M\in \Rep G$ is locally trivial (same assumptions everywhere), then if $\pi:X\to X/G$,
+	we get that $X\times^G M= X\times M/G\to X/G$ is locally trivial.
+\end{thm}
+\begin{rmk}
+	In particular, if $H\le G$ and if $M\in \Rep H$, the map $G\times^H M\to G/H$ is a vector bundle.
+\end{rmk}
+\begin{thm}
+	Let $H\le G$ and $M\in \Rep H$. Then $G\times ^H M\to G/H$ has the property that $\calL(M)$ is the sheaf of local sections of this vector bundle. That 
+	is, if $U\subset G/H$,
+	\[\Gamma(U,G\times^H M)\simeq \calL(M)(U)\]
+\end{thm}
+
+\section{March 4, 2020}
+Today we are talking about 
+\subsection{Distribution algebras for an algebraic group}
+The idea here is that we know something about representations of algebras and we want to relate
+the representations of an algebraic group to the reps of algebras.
+
+One idea is to notice that $\Rep G\simeq \lcomod{k[G]}$ and then dualize to get the modules of the dual of $k[G]$.
+This can blow up, of course, so we have to refine this idea.
+
+Let $X$ be an affine scheme and let $k[X]$ be its ring of functions. Let $x\in X(k)$ and let 
+\[I_x=\{f\in k[X]|f(x)=0\}\]
+Then we define 
+\begin{defn}
+	$\mu:k[X]\to k$ is a distribution of order $\le n$ supported at $x$ if $\mu(I_x^{n+1})=0$.
+	Then $\operatorname{Dist}_n(X,x)\cong \left(k[X]/I_x^{n+1}\right)^\times$ is the collection of all such $\mu$.
+
+	The subset of $\mu$ satisfying $\mu(1)=0$ is $\operatorname{Dist}_n^+(X,x)$ and we write $\operatorname{Dist}(X,x)$ and $\operatorname{Dist}^+(X,x)$
+	for the unions over all $n$.
+\end{defn}
+Notice that $\Dist^+_n(X,x)\cong \left(I_x/I_x^{n+1})\right)^\times$, so 
+\[\Dist_1^+(X,x)\simeq (I_x/I_x^2)^\times\simeq T_xX.\]
+
+Now $\Dist(X,x)$ is a $k[X]$-module via ($f,f_1\in k[X],\mu:k[X]\to k$)
+\[(f\mu)(f_1)=\mu(ff_1)\]
+\begin{rmk}
+	If $X$ is a (non-affine) scheme over $k$,
+	\[\Dist_n(X,x)\simeq \left(\O_{X,x}/\frakm_{X,x}^{n+1}\right)^\times\]
+\end{rmk}
+Let's go back to the affine case. This is functorial: if we have a map $\varphi:X\to Y$ and algebra map 
+$\varphi^\ast:k[Y]\to k[X]$ that satisfies $(\varphi^\ast)^{-1}(I_x)=I_{\varphi(x)}$. Thus we get a map 
+\[d\varphi_x:\Dist_n(X,x)\to \Dist_n(Y,\varphi(x))\]
+
+Also $\Dist(X,x)$ is a cocommutative coalgebra where the diagonal map $\Delta:X\to X\times X$ gives us a map 
+\[\Dist(X,x)\to \Dist(X,x)\otimes \Dist(X,x).\]
+
+\begin{ex}
+	Let $X=\A1$ and let $x=0$. Recall $k[\A1]\simeq k[T]$. Then $I=I_0=(T)$.
+	Therefore 
+	\[\Dist_n(\A1,0)\simeq (k[T]/T^{n+1})^\times\]
+	and 
+	\[\Dist(\A1,0)\simeq \oplus_{r\ge 0}k\gamma_r\]
+	where we let $\gamma_r:k[T]\to k$ via $\gamma_r(T^n)=\delta_{n,r}$.
+
+	Therefore $\gamma_1=\frac{\partial}{\partial T}$ and if the characteristic is zero, $\gamma_r=\frac{1}{r!}\frac{\partial}{\partial T^r}$.
+	Remember that we have a natural algebra structure from $\A1$ as a group scheme. The coalgebra structure comes from $\gamma_1$ 
+	being primitive and extending up to the rest.
+\end{ex}
+
+Now let $X=G$ be a (connected!) group scheme and $x=1$. Then $\Dist(G)\eqdef\Dist(G,1)$. 
+Then we can get an algebra structure in the following way: if $\mu,\nu:k[G]\to k$,
+\[\mu\nu=\mu\bar\otimes\nu\circ\Delta\]
+as we usually do with these things.
+
+Now $\Dist G$ is a filtered algebra so that 
+\[\Dist_n(G)\Dist_m(G)\subseteq\Dist_{n+m}(G).\]
+\begin{rmk}
+	In $k[G]$ we have a comultiplication with a special form. This can be read in Jantzen section I.2.4. If $I$ is the 
+	augmentation ideal of $k[G]$, and $f\in I$,
+	\[\Delta(f)=f\otimes 1+1\otimes f+I\otimes I\]
+
+	Then multiplicativity of $\Delta$ gives us 
+	\[\Delta(f_1\cdots f_n)=\prod_1^n(f_i\otimes 1+1\otimes f_n)+\sum_1^{n-1}I^r\otimes I^{n-r}\]
+\end{rmk}
+
+If we have $\mu,\nu\in\Dist_n$, then $[\mu,\nu]=\mu\nu-\nu\mu\in\Dist_{n+m-1}(G)$.
+Thus we have a filtration $\Dist_n\subseteq \Dist_{n+1}$ and the associated graded ring 
+$\oplus \Dist_i/\Dist_{i-1}$ is commutative. This is reminiscent of the universal enveloping algebra of a Lie algebra!
+\begin{ex}
+	Let $G=\Ga$ again. THen $k[T]=k[\Ga]$ and think of $\gamma_r=\frac{1}{r!}\frac{\partial^r}{\partial T^r}$.
+	What is the multiplication? Ends up it is the divided power algebra structure:
+	\[\gamma_n\gamma_n=\binom{n+m}{n}\gamma_{n+m}.\]
+	How do we see this? Consider 
+	\[\Delta(T)=1\otimes T+T\otimes 1\quad\Rightarrow\quad \Delta(T^l)=\sum_0^l\binom{l}{i}T^i\otimes T^{l-i}\]
+	so then 
+	\[\gamma_n\gamma_m(T^l)=\sum\binom{l}{i}\gamma_n(T^i)\gamma_m(T^{l-i})\]
+	where the summands are all zero except when $l=n+m$ and $i=n$.
+
+	In characteristic zero, $\gamma_r$ is honestly $\frac{1}{r!}\gamma_1^r$ so it ends up that this is just isomorphic
+	to a polynomial algebra $k[\gamma_1]$. This is not true in positive characteristic, because $\gamma_1^p$ dies, so we get a new 
+	generator $\gamma_p$ and so on.
+\end{ex}
+Now let $\frakg=\Lie G$. Then since $\frakg\simeq\Dist^+_1(G)$ which maps into $\Dist(G)$, the universal property 
+of the universal enveloping algebra gets us us the map 
+\[\calU(\frakg)\to \Dist(G)\]
+\begin{prop}
+	In characteristic zero, this map is an isomorphism.
+\end{prop}
+In positive characteristic, this fails. However we still ahve an injective map 
+\[u(\frakg)\hookrightarrow\Dist(G)\]
+where $u(\frakg)$ is the usual restricted enveloping algebra.
+
+\section{March 6th, 2020}
+Let $M\in\Rep G$ and let $\mu\in \Dist G$. THen define an action of $\Dist G$ on $M$ by 
+\[\mu m=(\id\bar\otimes \mu)\circ\Delta_M\]
+and we get a map $\Rep G\to\Dist G$. WHen $G$ is an algebraic gorup scheme over $k$,
+then the map on homs is an isomorphism. Notice: $\Rep G$ lands in locally finite $\Dist G$ modules, 
+but not all locally finite $\Dist G$ algebras come from $G$-mod.
+
+We're going fast today because I asked about $\operatorname{PGL}_n$. Whoops. Gonna just listen.
+
+Check out Jantzen for the development of roots and weights and all. But here are some theorems!
+\begin{thm}
+	Let $\lambda$ be a dominant weight. Then $H^0(\lambda)\eqdef\Ind_B^G k_\lambda$ is finite dimensional and nonzero, but $H^i(\lambda)=0$ for all $i>0$.
+\end{thm}
+\begin{thm}
+	$H^0(\lambda)$ has a simple socle called $L(\lambda)$. It's obviously also finite dimensional.
+\end{thm}
+\begin{thm}
+	$\{L(\lambda)|\lambda\in X^+\}$ is a complete and irredundant set of representations for $G$.
+\end{thm}
 
